### 2021-10-25 _paper.rmd
init structure.

---
title: "Methods for understanding the variable importance of local explanations of black-box models"
# author:
#   - name: Nicholas Spyrison
#     address: |
#       | Monash University
#       | Faculty of Information Technology
#       | ORCiD: 0000-0002-8417-0212
#     email: \email{nicholas.spyrison@monash.edu}
#     orcid: 0000-0002-8417-0212
#   - name: Dianne Cook
#     affiliation: Monash University
#     address: |
#       | ORCiD: 000-0002-3813-7155
#     orcid: 000-0002-3813-7155
#   - name: Kimbal Marriott
#     affiliation: Monash University
#     address: |
#       | ORCiD: 0000-0002-9813-0377
#     orcid: 0000-0002-9813-0377
# keywords:
#   formatted: [XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, DR, dimension reduction, R]
#   plain:     [XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, DR, dimension reduction, R]
bibliography: cheem_paper.bib
abstract: |
  ABSTRACT NEEDED
# documentclass: jdssv
# classoption: article
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    #template: template/jdssv_template.tex
    number_sections: yes
    fig_caption: yes
appendix:
  # - "appendix-1.Rmd"
# title:
#   formatted: "Methods for understanding the variable importance of local explanations of black-box models"
#   plain:     "Methods for understanding the variable importance of local explanations of black-box models"
#   short:     "Use of the manual tour to visually interrogate local explanations"
editor_options:
  chunk_output_type: console
urlcolor: blue
linkcolor: red
header-includes:
 - \usepackage{setspace} ### for title page spacing
 - \usepackage{hyperref} ### for all sorts of linking
 - \usepackage{graphicx} ### to insert Monash crest
---
```{r include=FALSE, cache=FALSE}
require("knitr")
require("kableExtra")
require("magrittr")
## Work packages
require("cheem")
require("spinifex")
## chunk options
knitr::opts_chunk$set(
  fig.align = "center",
  echo = FALSE, ## Code
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  cache = TRUE,
  cache.lazy = FALSE)
```

<!-- # cheat sheet {#sec:cheatsheet} -->
<!-- A bib reference [@wickham_visualizing_2015]. -->
<!-- A [Section intro](#sec:cheatsheet) reference, alternatively, section ef{sec:intro} (with no @; \\ref{sec:intro}). -->
<!-- ```{r crest, echo=FALSE, out.height = "10%", out.width = "10%", fig.cap = "A caption for crest figure"} -->
<!-- knitr::include_graphics("./figures/crest.jpg") -->
<!-- ``` -->
<!-- A figure \@ref(fig:crest) reference (with @; \\@ref(fig:crest). -->
<!-- ref:myFig-cap) Separate caption created above the R chunk -->
<!-- ```{r step2, echo=F, fig.cap = "(ref:myFig-cap)"} -->
<!-- knitr::include_graphics("./figures/crest.jpg") -->
<!-- ``` -->

<!-- GENERAL STRUCTURE:: -->
<!-- -- couple pages of general intro -->
<!-- -- tie into XAI -->
<!-- -- explain tour & manual tour -->
<!-- -- example: details on SHAP values -->
<!-- -- what you could learn about from classification and what from regression -->
<!-- -- section on software/interface: -->

# Abstract

Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models. The lack of transparency in such models has actualized as research toward explainable AI (XAI). While there are

# Introduction {#sec:intro}
<!-- WHAT TOPICS, and ISSUES addressed (motivation) -->
## MODELING
## XAI & interpretability crisis
## Local explanations
## Data visualization tours

# SHAP local explanation
## Variable importance, permuting over the X's included
## Visualizing and break down plots

# Interrogate variable imporances of the local explanations: Cheem
## RF model
## shap matrix
## Linked global approximations of data and local explanations
## Applcation of the manual tour to

# Design of cheem
## 1) Penguins speicies classification
## 2) FIFA wage regression 
## 3)?
## 4)?

# Software Infrastructure

# Discsussion

# Acknoledgements

# References

### 2021-10-25 _paper.tex
init structure.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Methods for understanding the variable importance of local explanations of black-box models},
  pdfkeywords={blah, blah},
  colorlinks=true,
  linkcolor=red,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Methods for understanding the variable importance of local explanations of black-box models}
\author{true \and true}
\date{}

\begin{document}
\maketitle
\begin{abstract}
ABSTRACT NEEDED
\end{abstract}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\hypertarget{sec:intro}{%
\section{Introduction}\label{sec:intro}}

\hypertarget{modeling}{%
\subsection{MODELING}\label{modeling}}

\hypertarget{xai-interpretability-crisis}{%
\subsection{XAI \& interpretability crisis}\label{xai-interpretability-crisis}}

\hypertarget{local-explanations}{%
\subsection{Local explanations}\label{local-explanations}}

\hypertarget{data-visualization-tours}{%
\subsection{Data visualization tours}\label{data-visualization-tours}}

\hypertarget{shap-local-explanation}{%
\section{SHAP local explanation}\label{shap-local-explanation}}

\hypertarget{variable-importance-permuting-over-the-xs-included}{%
\subsection{Variable importance, permuting over the X's included}\label{variable-importance-permuting-over-the-xs-included}}

\hypertarget{visualizing-and-break-down-plots}{%
\subsection{Visualizing and break down plots}\label{visualizing-and-break-down-plots}}

\hypertarget{interrogate-variable-imporances-of-the-local-explanations-cheem}{%
\section{Interrogate variable imporances of the local explanations: Cheem}\label{interrogate-variable-imporances-of-the-local-explanations-cheem}}

\hypertarget{rf-model}{%
\subsection{RF model}\label{rf-model}}

\hypertarget{shap-matrix}{%
\subsection{shap matrix}\label{shap-matrix}}

\hypertarget{linked-global-approximations-of-data-and-local-explanations}{%
\subsection{Linked global approximations of data and local explanations}\label{linked-global-approximations-of-data-and-local-explanations}}

\hypertarget{applcation-of-the-manual-tour-to}{%
\subsection{Applcation of the manual tour to}\label{applcation-of-the-manual-tour-to}}

\hypertarget{design-of-cheem}{%
\section{Design of cheem}\label{design-of-cheem}}

\hypertarget{penguins-speicies-classification}{%
\subsection{1) Penguins speicies classification}\label{penguins-speicies-classification}}

\hypertarget{fifa-wage-regression}{%
\subsection{2) FIFA wage regression}\label{fifa-wage-regression}}

\hypertarget{section}{%
\subsection{3)?}\label{section}}

\hypertarget{section-1}{%
\subsection{4)?}\label{section-1}}

\hypertarget{software-infrastructure}{%
\section{Software Infrastructure}\label{software-infrastructure}}

\hypertarget{discsussion}{%
\section{Discsussion}\label{discsussion}}

\hypertarget{acknoledgements}{%
\section{Acknoledgements}\label{acknoledgements}}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}

### 2021-10-25 figures/_make_figures.r

@@ -1,0 +1,519 @@
+# ch5_fig1_shap_distr_bd ----
+require("DALEX")
+require("dplyr")
+require("ggplot2")
+
+## Local func
+my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
+  ## Remade from: iBreakDown:::print.break_down_uncertainty
+  data.frame(
+    player = player_tag,
+    label  = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+    variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+    value  = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
+    ## Of the distribution of local attributions:
+    min    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
+    q1     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
+    median = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), median, na.rm = TRUE),
+    q3     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.75, na.rm = TRUE),
+    max    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), max, na.rm = TRUE))
+}
+my_parts_distribution <- function(pred_parts, player_tag = "<tag unused>"){
+  df <- data.frame(
+    player = player_tag,
+    label = pred_parts$label,
+    variable = pred_parts$variable_name,
+    value = pred_parts$variable_value, ## Obs value of X
+    contribution = pred_parts$contribution, ## SHAP contribution
+    B_perm_num = pred_parts$B
+  )
+  rownames(df) <- paste(pred_parts$label, pred_parts$variable, pred_parts$B, sep = ": ")
+  return(df)
+}
+.lvl_ord <- c("react", "off", "mvm", "def", "pwr", "acc", "bmi", "age", "gk")
+my_bd_df <- function(break_down, player_tag = "<tag unused>"){
+  df <- data.frame(
+    player = player_tag,
[…]

### 2021-10-25 _paper.rmd
abstract

@@
 abstract: |
-  ABSTRACT NEEDED
+  Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
+In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of explainable AI (XAI).
+XAI attempts to shed light on these models, one such approach is the use of local explanations.
+A local explanation of a model give a point-estimate of linear variable importance in the vicinity of one observation.
+We extract explanations for each observation, and approximate data and this attribution space side-by-side with linked brushing.
+After identifying an observation of interest its local explanation is used as a 1D projection basis.
+We then manipulate the magnitude of the variable contributions with a technique called the tour.
+This tour animates many projections over small changes in the projection basis.
+Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
+The implementation of our framework is available as an R package called cheem available at [github.com/nspyrison/cheem](https://github.com/nspyrison/cheem).
 # documentclass: jdssv
@@
 
-# Abstract
-
-Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-The lack of transparency in such models has actualized as research toward explainable AI (XAI).
-While there are
-
 # Introduction {#sec:intro}
@@
 ## shap matrix
-## Linked global approximations of data and local explanations
-## Applcation of the manual tour to
+## Linked global approximations of data and local attribution spaces
+## Applcation of the manual tour to intterogate local explanations
+## Classification task
+## Regression task
 
-# Design of cheem
+# Application Design
 ## 1) Penguins speicies classification
@@
 # Software Infrastructure
+## Extend spinifex, consume DALEX & treeshap
+## Preprocess
+## Runtime rendering
 

### 2021-10-25 _paper.tex
abstract

@@
   pdftitle={Methods for understanding the variable importance of local explanations of black-box models},
-  pdfkeywords={blah, blah},
   colorlinks=true,
@@
 \title{Methods for understanding the variable importance of local explanations of black-box models}
-\author{true \and true}
-\date{}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 \begin{abstract}
-ABSTRACT NEEDED
+Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
+In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of explainable AI (XAI).
+XAI attempts to shed light on these models, one such approach is the use of local explanations.
+A local explanation of a model give a point-estimate of linear variable importance in the vicinity of one observation.
+We extract explanations for each observation, and approximate data and this attribution space side-by-side with linked brushing.
+After identifying an observation of interest its local explanation is used as a 1D projection basis.
+We then manipulate the magnitude of the variable contributions with a technique called the tour.
+This tour animates many projections over small changes in the projection basis.
+Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
+The implementation of our framework is available as an R package called cheem available at \href{https://github.com/nspyrison/cheem}{github.com/nspyrison/cheem}.
 \end{abstract}
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
 
@@
 
-\hypertarget{linked-global-approximations-of-data-and-local-explanations}{%
-\subsection{Linked global approximations of data and local explanations}\label{linked-global-approximations-of-data-and-local-explanations}}
+\hypertarget{linked-global-approximations-of-data-and-local-attribution-spaces}{%
+\subsection{Linked global approximations of data and local attribution spaces}\label{linked-global-approximations-of-data-and-local-attribution-spaces}}
 
-\hypertarget{applcation-of-the-manual-tour-to}{%
-\subsection{Applcation of the manual tour to}\label{applcation-of-the-manual-tour-to}}
+\hypertarget{applcation-of-the-manual-tour-to-intterogate-local-explanations}{%
+\subsection{Applcation of the manual tour to intterogate local explanations}\label{applcation-of-the-manual-tour-to-intterogate-local-explanations}}
 
-\hypertarget{design-of-cheem}{%
-\section{Design of cheem}\label{design-of-cheem}}
+\hypertarget{classification-task}{%
+\subsection{Classification task}\label{classification-task}}
+
+\hypertarget{regression-task}{%
+\subsection{Regression task}\label{regression-task}}
+
+\hypertarget{application-design}{%
+\section{Application Design}\label{application-design}}
 
@@
 
+\hypertarget{extend-spinifex-consume-dalex-treeshap}{%
+\subsection{Extend spinifex, consume DALEX \& treeshap}\label{extend-spinifex-consume-dalex-treeshap}}
+
+\hypertarget{preprocess}{%
+\subsection{Preprocess}\label{preprocess}}
+
+\hypertarget{runtime-rendering}{%
+\subsection{Runtime rendering}\label{runtime-rendering}}
+
 \hypertarget{discsussion}{%

### 2021-10-26 _paper.rmd
intro section

@@
 <!-- A bib reference [@wickham_visualizing_2015]. -->
-<!-- A [Section intro](#sec:cheatsheet) reference, alternatively, section ef{sec:intro} (with no @; \\ref{sec:intro}). -->
+<!-- A [Section intro](#sec:cheatsheet) reference, alternatively, section {sec:intro} (with no @; \\ref{sec:intro}). -->
 <!-- ```{r crest, echo=FALSE, out.height = "10%", out.width = "10%", fig.cap = "A caption for crest figure"} -->
@@
 
-<!-- GENERAL STRUCTURE:: -->
-<!-- -- couple pages of general intro -->
-<!-- -- tie into XAI -->
-<!-- -- explain tour & manual tour -->
-<!-- -- example: details on SHAP values -->
-<!-- -- what you could learn about from classification and what from regression -->
-<!-- -- section on software/interface: -->
-
 # Introduction {#sec:intro}
 <!-- WHAT TOPICS, and ISSUES addressed (motivation) -->
-## MODELING
-## XAI & interpretability crisis
-## Local explanations
-## Data visualization tours
 
-# SHAP local explanation
+<!-- ## Predictive models -->
+<!-- history of regression and classification -->
+Mathematically rigorous approaches to predictive modeling are attributed to the method of least squares, over two centuries ago by Legendre and Gauss in 1805 and 1809 respectively.
+In 1886 Francis Galton coined the term _regression_ to refer to continuous, quantitative predictions.
+While _classification_ refers discretion predictions as introduced by Fisher in 1936.
+
+<!-- introduce explanatory vs predictive modeling -->
+Breiman and Shmueli [@breiman_statistical_2001, @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling is performed for to predict new or future out-of-sample observations.
+This distinction draws attention to the divide between interpretable models and black-box models.
+In explanatory modeling the interpretable is a key feature for drawing inferential conclusions.
+While predictive modeling may opt for potentially more accurate black-box models.
+The intended use of a model has important implications for which methods are used and the development of those models.
+
+<!-- ## XAI & interpretability crisis -->
+Predictive model and black box modeling is becoming increasingly common, but not without controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
+Applications have been known to reflect common biases against sex [@dastin_amazon_2018, @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
+This is a common issue stemming from biases the in sample data are violate ethical principals.
+Another issue is that of data-drift, when new data is outside the support of latent or exogenous explanatory variables.
+Data-drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
+Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@Adadi_peeking_2018, @arrieta_explainable_2020].
+
+<!-- ## Local explanations -->
+One branch of XAI is local explanations, which take a variable attribution approach to bring transparency to a model.
+Local explanations attempt to approximate a linear variable importance at the location of one observation.
+There are many such local explanations, any of which is works with our approach (assuming model-explanation compatibility).
+
+<!-- SHAP -->
+However, to illustrate our work we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010, @strumbelj_explaining_2014].
+The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
+To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP [@lundberg_consistent_2018].
+<!-- 2nd half is more application details; move lower lower? -->
+
+<!-- ## Data visualization tours -->
+In multivariate data visualization a _tour_ [@asimov_grand_1994, @buja_grand_1996, @lee_review_2021] is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
+Tours are viewed as an animation over small changes to a projection basis.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
+The intuition is similar to watching the shadow of hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
+<!-- ## manual tours --> There are various types of tours, which are distinguished be the generation of sequence of projection bases.
+In a _manual_ tour [@cook_manual_1997, @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
+<!--tours and models --> Applying tours in conjunction with models has been previously done,  _ie_ for exploring various statistical model fits [@wickham_removing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013, @da_silva_projection_2021].
+
+The approach purposed below is to use the manual tour as means to interrogate a local explanation; it see if its variable importance are good explanation for the model predictions.
+We make R package `cheem` with an interactive application to facilitate analysis.
+By viewing approximations of data- and attribution- space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
+We give case studies of toy and modern datasets for both classification and regression tasks.
+
+The rest of paper is organized as follows.
+The next section [SHAP](#sec:SHAP) covers the background of the local explanation SHAP and the traditional visuals produced from it.
+The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis.
+Following that, [Software Instructure](#sec:softwareinfrastructure) discusses the backend details of the package and preprocessing.
+The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
+We conclude with [Discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
+
+# SHAP local explanation {#sec:SHAP}
 ## Variable importance, permuting over the X's included
@@
 
-# Interrogate variable imporances of the local explanations: Cheem
-## RF model
-## shap matrix
-## Linked global approximations of data and local attribution spaces
-## Applcation of the manual tour to intterogate local explanations
-## Classification task
+<!-- # Interrogate variable importance of the local explanations: Cheem {#sec:cheem} -->
+<!-- ## RF model -->
+<!-- ## shap matrix -->
+<!-- ## Linked global approximations of data and local attribution spaces -->
+<!-- ## Application of the manual tour to interrogate local explanations -->
+<!-- ## Classification task -->
+<!-- ## Regression task -->
+
+# Application Design {#sec:applicationdesign}
+## Preprocessing
+## Global view
+## Cheem tour
+## Classifcation task
 ## Regression task
 
-# Application Design
+# Software Infrastructure {#sec:softwareinfrastructure}
+## Extend spinifex, consume DALEX & treeshap
+## Preprocess
+## Runtime rendering
+
+# Case Studies {#sec:casestudies}
 ## 1) Penguins speicies classification
@@
 ## 4)?
-
-# Software Infrastructure
-## Extend spinifex, consume DALEX & treeshap
-## Preprocess
-## Runtime rendering
 

### 2021-10-26 _paper.tex
intro section

@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\hypertarget{modeling}{%
-\subsection{MODELING}\label{modeling}}
+Mathematically rigorous approaches to predictive modeling are attributed to the method of least squares, over two centuries ago by Legendre and Gauss in 1805 and 1809 respectively.
+In 1886 Francis Galton coined the term \emph{regression} to refer to continuous, quantitative predictions.
+While \emph{classification} refers discretion predictions as introduced by Fisher in 1936.
 
-\hypertarget{xai-interpretability-crisis}{%
-\subsection{XAI \& interpretability crisis}\label{xai-interpretability-crisis}}
+Breiman and Shmueli (\textbf{shmueli\_explain\_2010?}) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed for to predict new or future out-of-sample observations.
+This distinction draws attention to the divide between interpretable models and black-box models.
+In explanatory modeling the interpretable is a key feature for drawing inferential conclusions.
+While predictive modeling may opt for potentially more accurate black-box models.
+The intended use of a model has important implications for which methods are used and the development of those models.
 
-\hypertarget{local-explanations}{%
-\subsection{Local explanations}\label{local-explanations}}
+Predictive model and black box modeling is becoming increasingly common, but not without controversy and issues (\textbf{kodiyan\_overview\_2019?}).
+Applications have been known to reflect common biases against sex (\textbf{duffy\_apple\_2019?}), race (\textbf{larson\_how\_2016?}), and age (\textbf{diaz\_addressing\_2018?}).
+This is a common issue stemming from biases the in sample data are violate ethical principals.
+Another issue is that of data-drift, when new data is outside the support of latent or exogenous explanatory variables.
+Data-drift can lead to worse predictions (\textbf{salzberg\_why\_2014?}).
+Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI Arrieta et al.
+(2020).
 
-\hypertarget{data-visualization-tours}{%
-\subsection{Data visualization tours}\label{data-visualization-tours}}
+One branch of XAI is local explanations, which take a variable attribution approach to bring transparency to a model.
+Local explanations attempt to approximate a linear variable importance at the location of one observation.
+There are many such local explanations, any of which is works with our approach (assuming model-explanation compatibility).
 
-\hypertarget{shap-local-explanation}{%
-\section{SHAP local explanation}\label{shap-local-explanation}}
+However, to illustrate our work we apply the model-agnostic explanation SHAP (\textbf{strumbelj\_explaining\_2014?}).
+The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
+To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP (\textbf{lundberg\_consistent\_2018?}).
+
+In multivariate data visualization a \emph{tour} Lee et al.
+(2021) is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
+Tours are viewed as an animation over small changes to a projection basis.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
+The intuition is similar to watching the shadow of hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
+There are various types of tours, which are distinguished be the generation of sequence of projection bases.
+In a \emph{manual} tour Spyrison and Dianne Cook (2020) this path is defined by changing the contribution of a selected variable.
+Applying tours in conjunction with models has been previously done, \emph{ie} for exploring various statistical model fits (\textbf{wickham\_removing\_2015?}), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path (\textbf{da\_silva\_projection\_2021?}).
+
+The approach purposed below is to use the manual tour as means to interrogate a local explanation; it see if its variable importance are good explanation for the model predictions.
+We make R package \texttt{cheem} with an interactive application to facilitate analysis.
+By viewing approximations of data- and attribution- space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
+We give case studies of toy and modern datasets for both classification and regression tasks.
+
+The rest of paper is organized as follows.
+The next section \protect\hyperlink{sec:SHAP}{SHAP} covers the background of the local explanation SHAP and the traditional visuals produced from it.
+The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis.
+Following that, \protect\hyperlink{sec:softwareinfrastructure}{Software Instructure} discusses the backend details of the package and preprocessing.
+The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
+We conlude with \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
+
+\hypertarget{sec:SHAP}{%
+\section{SHAP local explanation}\label{sec:SHAP}}
 
@@
 
-\hypertarget{interrogate-variable-imporances-of-the-local-explanations-cheem}{%
-\section{Interrogate variable imporances of the local explanations: Cheem}\label{interrogate-variable-imporances-of-the-local-explanations-cheem}}
+\hypertarget{sec:applicationdesign}{%
+\section{Application Design}\label{sec:applicationdesign}}
 
-\hypertarget{rf-model}{%
-\subsection{RF model}\label{rf-model}}
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
 
-\hypertarget{shap-matrix}{%
-\subsection{shap matrix}\label{shap-matrix}}
+\hypertarget{global-view}{%
+\subsection{Global view}\label{global-view}}
 
-\hypertarget{linked-global-approximations-of-data-and-local-attribution-spaces}{%
-\subsection{Linked global approximations of data and local attribution spaces}\label{linked-global-approximations-of-data-and-local-attribution-spaces}}
+\hypertarget{cheem-tour}{%
+\subsection{Cheem tour}\label{cheem-tour}}
 
-\hypertarget{applcation-of-the-manual-tour-to-intterogate-local-explanations}{%
-\subsection{Applcation of the manual tour to intterogate local explanations}\label{applcation-of-the-manual-tour-to-intterogate-local-explanations}}
-
-\hypertarget{classification-task}{%
-\subsection{Classification task}\label{classification-task}}
+\hypertarget{classifcation-task}{%
+\subsection{Classifcation task}\label{classifcation-task}}
 
@@
 
-\hypertarget{application-design}{%
-\section{Application Design}\label{application-design}}
+\hypertarget{sec:softwareinfrastructure}{%
+\section{Software Infrastructure}\label{sec:softwareinfrastructure}}
+
+\hypertarget{extend-spinifex-consume-dalex-treeshap}{%
+\subsection{Extend spinifex, consume DALEX \& treeshap}\label{extend-spinifex-consume-dalex-treeshap}}
+
+\hypertarget{preprocess}{%
+\subsection{Preprocess}\label{preprocess}}
+
+\hypertarget{runtime-rendering}{%
+\subsection{Runtime rendering}\label{runtime-rendering}}
+
+\hypertarget{sec:casestudies}{%
+\section{Case Studies}\label{sec:casestudies}}
 
@@
 
-\hypertarget{software-infrastructure}{%
-\section{Software Infrastructure}\label{software-infrastructure}}
-
-\hypertarget{extend-spinifex-consume-dalex-treeshap}{%
-\subsection{Extend spinifex, consume DALEX \& treeshap}\label{extend-spinifex-consume-dalex-treeshap}}
-
-\hypertarget{preprocess}{%
-\subsection{Preprocess}\label{preprocess}}
-
-\hypertarget{runtime-rendering}{%
-\subsection{Runtime rendering}\label{runtime-rendering}}
-
 \hypertarget{discsussion}{%
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
+Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
+2020.
+{``Explainable {Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies, Opportunities and Challenges Toward Responsible {AI}.''} \emph{Information Fusion} 58: 82--115.
+
+\leavevmode\hypertarget{ref-asimov_grand_1994}{}%
+Asimov, Daniel A., and Andreas Buja.
+1994.
+{``Grand Tour via Geodesic Interpolation of 2-Frames.''} In \emph{Visual Data Exploration and Analysis}, 2178:145--53.
+International Society for Optics; Photonics.
+\url{https://doi.org/10.1117/12.172065}.
+
+\leavevmode\hypertarget{ref-cook_manual_1997}{}%
+Cook, Dianne, and Andreas Buja.
+1997.
+{``Manual Controls for High-Dimensional Data Projections.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
+\url{https://doi.org/10.2307/1390747}.
+
+\leavevmode\hypertarget{ref-lee_review_2021}{}%
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Earo Wang, Nick Spyrison, and H.
+Sherry Zhang.
+2021.
+{``A {Review} of the {State}-of-the-{Art} on {Tours} for {Dynamic} {Visualization} of {High}-{Dimensional} {Data}.''} \emph{arXiv:2104.08016 {[}Cs, Stat{]}}, April.
+\url{http://arxiv.org/abs/2104.08016}.
+
+\leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
+Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An r Package for Creating a Manual Tour of Low-Dimensional Projections of Multivariate Data.''} \emph{The R Journal} 12 (1): (accepted).
+
+\end{CSLReferences}
 

### 2021-10-26 _paper.rmd
started SHAP, and touchup figure 1; shap distributions and breakdown plot

@@
 To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP [@lundberg_consistent_2018].
-<!-- 2nd half is more application details; move lower lower? -->
+<!-- 2nd half is more application details; move lower? -->
 
@@
 # SHAP local explanation {#sec:SHAP}
+
+SHaply Additive exPlanations, or SHAP [@lundberg_unified_2017] approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score [@shapley_value_1953].
+
+An observation's SHAP values were originally used to additvely explain the difference from the intercept to the prediction.
+This sort of explanation is predicated on the contribution of the previous variables, making it asymmetric across variable ordering.
+However, viewing several of these can highlight the non-linear weightings within a single model.
+We will use soccer data from FIFA 2020 season [@leone_fifa_2020] to illustrate this.
+We have 5000 observations of 9 aggregated skill measures and use a use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
+We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
+We expect to see a difference in attribution of the variable importance across the two positions of the players.
+Figure \@ref(fig:shapdistrbd) highlight that while the players have quite different wages, when we normalize the attribution we see that skill aggregate importance is different and in a way that we would expect; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+
+(ref:shapdistrbd-cap) Test caption
+```{r shapdistrbd, echo=F, fig.cap = "TODO:XXX Figure reference"}
+knitr::include_graphics("./figures/shap_distr_bd.pdf")
+```
+
 ## Variable importance, permuting over the X's included

### 2021-10-26 _paper.tex
started SHAP, and touchup figure 1; shap distributions and breakdown plot

@@
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
-We conlude with \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
+We conclude with \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
 
@@
 \section{SHAP local explanation}\label{sec:SHAP}}
+
+SHaply Additive exPlanations, or SHAP (Lundberg and Lee 2017) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score (\textbf{shapley\_value\_1953?}).
+
+An observation's SHAP values were originally used to additvely explain the difference from the intercept to the prediction.
+This sort of explanation is predicated on the contribution of the previous variables, making it asymmetric across variable ordering.
+However, viewing several of these can highlight the non-linear weightings within a single model.
+We will use soccer data from FIFA 2020 season (Leone 2020) to illustrate this.
+We have 5000 observations of 9 aggregated skill measures and use a use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
+We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
+We expect to see a difference in attribution of the variable importance across the two positions of the players.
+Figure \ref{fig:shapdistrbd} highlight that while the players have quite different wages, when we normalize the attribution we see that skill aggregate importance is different and in a way that we would expect; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+
+\begin{figure}
+
+{\centering \includegraphics{./figures/shap_distr_bd}
+
+}
+
+\caption{TODO:XXX Figure reference}\label{fig:shapdistrbd}
+\end{figure}
 
@@
 
+\leavevmode\hypertarget{ref-leone_fifa_2020}{}%
+Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
+
+\leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
+Lundberg, Scott, and Su-In Lee. 2017. {``A Unified Approach to Interpreting Model Predictions.''} \emph{arXiv Preprint arXiv:1705.07874}.
+
 \leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%

### 2021-10-26 figures/_ch5_cheem.r

@@ -1,519 +1,0 @@
-# ch5_fig1_shap_distr_bd ----
-require("DALEX")
-require("dplyr")
-require("ggplot2")
-
-## Local func
-my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
-  ## Remade from: iBreakDown:::print.break_down_uncertainty
-  data.frame(
-    player = player_tag,
-    label  = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
-    variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
-    value  = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
-    ## Of the distribution of local attributions:
-    min    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
-    q1     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
-    median = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), median, na.rm = TRUE),
-    q3     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.75, na.rm = TRUE),
-    max    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), max, na.rm = TRUE))
-}
-my_parts_distribution <- function(pred_parts, player_tag = "<tag unused>"){
-  df <- data.frame(
-    player = player_tag,
-    label = pred_parts$label,
-    variable = pred_parts$variable_name,
-    value = pred_parts$variable_value, ## Obs value of X
-    contribution = pred_parts$contribution, ## SHAP contribution
-    B_perm_num = pred_parts$B
-  )
-  rownames(df) <- paste(pred_parts$label, pred_parts$variable, pred_parts$B, sep = ": ")
-  return(df)
-}
-.lvl_ord <- c("react", "off", "mvm", "def", "pwr", "acc", "bmi", "age", "gk")
-my_bd_df <- function(break_down, player_tag = "<tag unused>"){
-  df <- data.frame(
-    player = player_tag,
[…]

### 2021-10-26 figures/_make_figures.r

@@ -30,7 +30,7 @@ my_parts_distribution <- function(pred_parts, player_tag = "<tag unused>"){
   rownames(df) <- paste(pred_parts$label, pred_parts$variable, pred_parts$B, sep = ": ")
   return(df)
 }
-.lvl_ord <- c("react", "off", "mvm", "def", "pwr", "acc", "bmi", "age", "gk")
+.lvl_ord <- c("reaction", "offense", "movement", "defense", "power", "accuracy", "BMI", "age", "goalkeeping")
 my_bd_df <- function(break_down, player_tag = "<tag unused>"){
   df <- data.frame(
     player = player_tag,
@@ -73,20 +73,20 @@ if(F) ## View corrplot?
 dat <- .dat_less_ys %>%
   dplyr::mutate(
     .keep = "none",
-    bmi = (weight_kg+(height_cm/100L)^2L),
+    BMI = (weight_kg+(height_cm/100L)^2L),
     age = age,
-    react = movement_reactions,
-    off = (attacking_finishing+skill_long_passing+attacking_volleys+
+    reaction = movement_reactions,
+    offense = (attacking_finishing+skill_long_passing+attacking_volleys+
              power_long_shots+skill_curve+mentality_positioning+attacking_crossing+
              attacking_short_passing+skill_dribbling+skill_ball_control)/10L,
-    def = (defending_sliding_tackle+mentality_interceptions+
+    defense = (defending_sliding_tackle+mentality_interceptions+
              defending_standing_tackle+defending_marking+mentality_aggression)/5L,
-    acc = (attacking_heading_accuracy+power_shot_power)/2L,
-    mvm = (movement_sprint_speed+movement_balance+movement_acceleration+
+    accuracy = (attacking_heading_accuracy+power_shot_power)/2L,
+    movement = (movement_sprint_speed+movement_balance+movement_acceleration+
              mentality_vision+mentality_composure+movement_agility+
              mentality_penalties+skill_fk_accuracy+power_stamina+movement_reactions)/10L,
-    pwr = (power_strength+power_jumping)/2L,
-    gk = (goalkeeping_diving+goalkeeping_positioning+goalkeeping_reflexes+
+    power = (power_strength+power_jumping)/2L,
+    goalkeeping = (goalkeeping_diving+goalkeeping_positioning+goalkeeping_reflexes+
[…]

### 2021-10-27 _paper.rmd
-finish section 2 -touch up shap distribution breakdown.

--- 2021-10-26 _paper.rmd
@@
 <!-- A figure \@ref(fig:crest) reference (with @; \\@ref(fig:crest). -->
-<!-- ref:myFig-cap) Separate caption created above the R chunk -->
-<!-- ```{r step2, echo=F, fig.cap = "(ref:myFig-cap)"} -->
-<!-- knitr::include_graphics("./figures/crest.jpg") -->
+<!-- (ref:baseplotcap) Separate caption created above the R chunk -->
+<!-- ```{r baseplot, echo=F, fig.cap = "(ref:baseplotcap)"} -->
+<!-- plot(1,2) -->
 <!-- ``` -->
@@
 # SHAP local explanation {#sec:SHAP}
-## Variable importance, permuting over the X's included
-## Visualizing and break down plots
 
-<!-- # Interrogate variable importance of the local explanations: Cheem {#sec:cheem} -->
-<!-- ## RF model -->
-<!-- ## shap matrix -->
-<!-- ## Linked global approximations of data and local attribution spaces -->
-<!-- ## Application of the manual tour to interrogate local explanations -->
-<!-- ## Classification task -->
-<!-- ## Regression task -->
+<!-- SHAP and history -->
+SHaply Additive exPlanations, or SHAP [@lundberg_unified_2017] approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score [@shapley_value_1953].
+
+<!-- An observation's SHAP values were originally used to additvely explain the difference from the intercept to the prediction.
+This sort of explanation is predicated on the contribution of the previous variables, making it asymmetric across variable ordering.
+-->
+
+TO illustrate SHAP and its original use we use soccer data from FIFA 2020 season [@leone_fifa_2020].
+We have 5000 observations of 9 aggregated skill measures and use a use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
+We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
+We expect to see a difference in attribution of the variable importance across the two positions of the players.
+
+<!-- Illustration -->
+Figure \@ref(fig:shapdistrbd) illustrates the SHAP values of these players.
+Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
+In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
+Panel c) shows a simplified breakdown plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
+Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing sum of the SHAP values.
+
+```{r shapdistrbd, echo=F, fig.cap = "Illustration of the distribution of SHAP attributions, the SHAP values, and a breakdown plot, the typical visual of SHAP local explanations.
+For FIFA 2020 data, of a random forest model regressing wages from 9 skill attributes for a star offensive and defensive player.
+a) The players have very different wages.
+b) Shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
+The median of these distributions are the final SHAP values, notice that that the variable importance differs across the exogenous information of player position.
+These explanations make sense; the variable importances make sense in light of the position of the player.
+c) Breakdown plots of the observations the explanation used to additively explain the difference beween the intercept and prediction"}
+knitr::include_graphics("./figures/shap_distr_bd.pdf")
+```
+
+In summary, this highlights how local explanations bring transparency to a model at least in the vicinity of their observations.
+In this instance we showed how two very different soccer players receive different profiles of variable importance to explain the the prediction of their wages.
+In the following section we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
 

### 2021-10-27 _paper.tex
-finish section 2 -touch up shap distribution breakdown.

@@
 We expect to see a difference in attribution of the variable importance across the two positions of the players.
-Figure \ref{fig:shapdistrbd} highlight that while the players have quite different wages, when we normalize the attribution we see that skill aggregate importance is different and in a way that we would expect; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+
+Figure \ref{fig:shapdistrbd} illustrates the SHAP values of these players.
+Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
+In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
+Panel c) shows a simplified breakdown plot (\textbf{gosiewska\_ibreakdown\_2019?}), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
+Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to
 
@@
 
-\caption{TODO:XXX Figure reference}\label{fig:shapdistrbd}
+\caption{Illustration of the distribution of SHAP attributions, the SHAP values, and a breakdown plot, the typical visual of SHAP local explanations.
+For FIFA 2020 data, of a random forest model regressing wages from 9 skill attributes for a star offensive and defensive player.
+a) The players have very different wages.
+b) Shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
+The median of these distributions are the final SHAP values, notice that that the variable importance differs across the exogenous information of player position.
+These explanations make sense; the variable importances make sense in light of the position of the player.
+c) Breakdown }\label{fig:shapdistrbd}
 \end{figure}
-
-\hypertarget{variable-importance-permuting-over-the-xs-included}{%
-\subsection{Variable importance, permuting over the X's included}\label{variable-importance-permuting-over-the-xs-included}}
-
-\hypertarget{visualizing-and-break-down-plots}{%
-\subsection{Visualizing and break down plots}\label{visualizing-and-break-down-plots}}
 

### 2021-10-27 figures/_make_figures.r

@@ -47,6 +47,8 @@ my_bd_df <- function(break_down, player_tag = "<tag unused>"){
   df$cumulative <- (df$cumulative - min(df$cumulative)) /
     (max(df$cumulative) - min(df$cumulative))
   df$last_cumulative <- c(NA, df$cumulative[-.n])
+  df$variable_num <- 1:.n
+  df$next_variable_num <- c(2:.n, NA)
   rownames(df) <- paste(break_down$label, break_down$variable, break_down$B, sep = ": ")
   return(df)
 }
@@ -166,7 +168,7 @@ dist_df$variable <- factor(dist_df$variable, levels = rev(.lvl_ord))
     scale_color_brewer(palette = "Dark2") +
     scale_fill_brewer(palette = "Dark2") +
     labs(title="SHAP distribution",
-         y = "variable", x = "normalize SHAP\n the median of the contributions while permuting X's") +
+         y = "variable", x = "normalized SHAP values\n the median of the contributions while permuting X's") +
     theme(legend.position = "off")
 )
 
@@ -186,17 +188,23 @@ bd_df_dijk <- my_bd_df(bd_dijk, "van Dijk (defense)")
 ## Bind, by row
 bd_df <- rbind(bd_df_messi, bd_df_dijk)
 bd_df <- bd_df[is.na(bd_df$variable) == FALSE, ]
-(g_bd <- ggplot(bd_df) +
+(g_bd <- ggplot() + #scale_y_continuous(limits = c(0, 1)) +
+    # ## vertical "lines"
+    # geom_segment(aes(x=variable_num, xend=variable_num, y=variable, yend=last_cumulative),
+    #              data = bd_df, color = "black", size = .5) +
+    ## horizontal "bars"
     geom_segment(aes(x=cumulative, xend=last_cumulative, y=variable, yend=variable, color=player),
-                 size=1.5, alpha=.8) + facet_grid(col=vars(player))+
+                 data = bd_df, size=1.5, alpha=.8) + 
+    facet_grid(col=vars(player)) +
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
-    labs(title="Breakdown profile of predictions",
-         y = "variable", x = "contribution to prediction | variable order") +
+    labs(title="Breakdown plot",
+         y = "variable", x = "normalized contribution to prediction | variable order") +
     theme(legend.margin = margin(0,0,0,0),
           legend.position = "bottom",
           axis.text.x = element_blank(),
[…]

### 2021-10-29 _paper.rmd
-## preprocessing - bib dump and fixes.

@@
 Data-drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
-Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@Adadi_peeking_2018, @arrieta_explainable_2020].
+Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@adadi_peeking_2018, @arrieta_explainable_2020].
 
@@
 <!-- ## Data visualization tours -->
-In multivariate data visualization a _tour_ [@asimov_grand_1994, @buja_grand_1996, @lee_review_2021] is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
+In multivariate data visualization a _tour_ [@asimov_grand_1985, @buja_grand_1986, @lee_review_2021] is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
 Tours are viewed as an animation over small changes to a projection basis.
@@
 In a _manual_ tour [@cook_manual_1997, @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
-<!--tours and models --> Applying tours in conjunction with models has been previously done,  _ie_ for exploring various statistical model fits [@wickham_removing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013, @da_silva_projection_2021].
+<!--tours and models --> Applying tours in conjunction with models has been previously done,  _ie_ for exploring various statistical model fits [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013, @da_silva_projection_2021].
 
@@
 We make R package `cheem` with an interactive application to facilitate analysis.
-By viewing approximations of data- and attribution- space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
+By viewing approximations of data- and attribution-space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
 We give case studies of toy and modern datasets for both classification and regression tasks.
@@
 # Application Design {#sec:applicationdesign}
-## Preprocessing
+
+Below we illustrate the two primary displays of the application: the global view and the tour view.
+Then we'll cover what we takeaway from the classification and regression tasks.
+Lastly, we discuss the preprocessing that needs to be done before display.
+
+<!-- TODO:XXX: -->
 ## Global view
@@
 ## Regression task
+
+## Preprocessing
+
+The benefit of having dynamic interaction with data is predicated on a reasonably small render time.
+The it is important to preprocess as much work as possible so that application resources can be used efficiently.
+Below we discuss the steps and details of the reprocessing.
+<!-- Fitting models regularly has non-negligible runtime and the exaction of a single observations local explanations is a similarly computationally expensive operation. -->
+
+(ref:citeRf) [@liaw_classification_2002]
+(ref:citeTs) [@kominsarczyk_treeshap_2021]
+\begin{itemize}
+	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic (color/shape) variable can be mapped typically a categorical variable exogenous to the model.
+	\item \textbf{Model:} any model can be used with this method.
+Currently we apply random forest models via the package \textbf{randomForest} (ref:citeRf) to mitigate the runtime of our local explanation which requires tree-based models.
+	\item \textbf{Local explanation:} any model-compatible linear explanation could be used.
+We apply tree SHAP, a more computationally efficient variant of SHAP applicable to tree-based models.
+This is done with the package \textbf{treeshap} (ref:citeTs), hosted on GitHub only].
+The global view shows all observations in attribution space requiring that we must extract the variable weightings from \emph{all} observations rather than just one.
+	\item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components.
+The within class separability of points can also optionally be displayed as QQ plots of the Mahalanobis distances within those spaces.
+\end{itemize}
+
+<!-- note on time of execution -->
+The time to preprocess the data will vary significantly with the choice of model and local explanation.
+However, for reference the FIFA data, 5000 observation of 9 explanatory variable, took 1.5 seconds to create PCA and Mahalanobis distances for both the data and attribution spaces.
+On the same data, a modestly hyper-parametered random forest model fit in 2.6 seconds, while extracting the tree SHAP values of each observation took 278 seconds combined.
+These time were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the run time will be spent on the local attribution.
+This makes tree SHAP a good candidate to start with.
+The package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes that are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+
+<!-- render time -->
+The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters.
+The application is made with __shiny__ [@chang_shiny_2021].
+The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievet_interactive_2020].
+The tour view is created in __spinifex__ [@spyrison_spinifex_2020], which creates the manual tour basis array, and facilitates similar rendering to html widget.
 

### 2021-10-29 _paper.tex
-## preprocessing - bib dump and fixes.

@@
 
-Breiman and Shmueli (\textbf{shmueli\_explain\_2010?}) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed for to predict new or future out-of-sample observations.
+Breiman and Shmueli Shmueli (2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed for to predict new or future out-of-sample observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
@@
 
-Predictive model and black box modeling is becoming increasingly common, but not without controversy and issues (\textbf{kodiyan\_overview\_2019?}).
-Applications have been known to reflect common biases against sex (\textbf{duffy\_apple\_2019?}), race (\textbf{larson\_how\_2016?}), and age (\textbf{diaz\_addressing\_2018?}).
+Predictive model and black box modeling is becoming increasingly common, but not without controversy and issues Kodiyan (2019).
+Applications have been known to reflect common biases against sex Duffy (2019), race (Larson et al.
+2016), and age (Díaz et al.
+2018).
 This is a common issue stemming from biases the in sample data are violate ethical principals.
 Another issue is that of data-drift, when new data is outside the support of latent or exogenous explanatory variables.
-Data-drift can lead to worse predictions (\textbf{salzberg\_why\_2014?}).
+Data-drift can lead to worse predictions Salzberg (2014).
 Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI Arrieta et al.
@@
 
-However, to illustrate our work we apply the model-agnostic explanation SHAP (\textbf{strumbelj\_explaining\_2014?}).
+However, to illustrate our work we apply the model-agnostic explanation SHAP Štrumbelj and Kononenko (2014).
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
-To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP (\textbf{lundberg\_consistent\_2018?}).
+To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP (S.
+M.
+Lundberg, Erion, and Lee 2018).
 
-In multivariate data visualization a \emph{tour} Lee et al.
+In multivariate data visualization a \emph{tour} S.
+Lee et al.
 (2021) is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
@@
 There are various types of tours, which are distinguished be the generation of sequence of projection bases.
-In a \emph{manual} tour Spyrison and Dianne Cook (2020) this path is defined by changing the contribution of a selected variable.
-Applying tours in conjunction with models has been previously done, \emph{ie} for exploring various statistical model fits (\textbf{wickham\_removing\_2015?}), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path (\textbf{da\_silva\_projection\_2021?}).
+In a \emph{manual} tour Spyrison and Cook (2020) this path is defined by changing the contribution of a selected variable.
+Applying tours in conjunction with models has been previously done, \emph{ie} for exploring various statistical model fits (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path Silva, Cook, and Lee (2021).
 
@@
 We make R package \texttt{cheem} with an interactive application to facilitate analysis.
-By viewing approximations of data- and attribution- space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
+By viewing approximations of data- and attribution-space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
 We give case studies of toy and modern datasets for both classification and regression tasks.
@@
 
-SHaply Additive exPlanations, or SHAP (Lundberg and Lee 2017) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score (\textbf{shapley\_value\_1953?}).
+SHaply Additive exPlanations, or SHAP (S.
+Lundberg and Lee 2017) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score (Shapley 1953).
 
-An observation's SHAP values were originally used to additvely explain the difference from the intercept to the prediction.
-This sort of explanation is predicated on the contribution of the previous variables, making it asymmetric across variable ordering.
-However, viewing several of these can highlight the non-linear weightings within a single model.
-We will use soccer data from FIFA 2020 season (Leone 2020) to illustrate this.
+TO illustrate SHAP and its original use we use soccer data from FIFA 2020 season (Leone 2020).
 We have 5000 observations of 9 aggregated skill measures and use a use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
@@
 We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
-Panel c) shows a simplified breakdown plot (\textbf{gosiewska\_ibreakdown\_2019?}), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to
+Panel c) shows a simplified breakdown plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
+Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing sum of the SHAP values.
 
@@
 These explanations make sense; the variable importances make sense in light of the position of the player.
-c) Breakdown }\label{fig:shapdistrbd}
+c) Breakdown plots of the observations the explanation used to additively explain the difference beween the intercept and prediction}\label{fig:shapdistrbd}
 \end{figure}
+
+In summary, this highlights how local explanations bring transparency to a model at least in the vicinity of their observations.
+In this instance we showed how two very different soccer players receive different profiles of variable importance to explain the the prediction of their wages.
+In the following section we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
 
@@
 
-\hypertarget{preprocessing}{%
-\subsection{Preprocessing}\label{preprocessing}}
+Below we illustrate the two primary displays of the application: the global view and the tour view.
+Then we'll cover what we takeaway from the classification and regression tasks.
+Lastly, we discuss the preprocessing that needs to be done before display.
 
@@
 \subsection{Regression task}\label{regression-task}}
+
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
+
+The benefit of having dynamic interaction with data is predicated on a reasonably small render time.
+The it is important to preprocess as much work as possible so that application resources can be used efficiently.
+Below we discuss the steps and details of the reprocessing.
+
+\begin{itemize}
+    \item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic (color/shape) variable can be mapped typically a categorical variable exogenous to the model.
+    \item \textbf{Model:} any model can be used with this method.
+Currently we apply random forest models via the package \textbf{randomForest} (Liaw and Wiener 2002) to mitigate the runtime of our local explanation which requires tree-based models.
+    \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
+We apply tree SHAP, a more computationally efficient variant of SHAP applicable to tree-based models.
+This is done with the package \textbf{treeshap} (Kominsarczyk et al.
+2021), hosted on GitHub only].
+The global view shows all observations in attribution space requiring that we must extract the variable weightings from \emph{all} observations rather than just one.
+    \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components.
+The within class separability of points can also optionally be displayed as QQ plots of the Mahalanobis distances within those spaces.
+\end{itemize}
+
+The time to preprocess the data will vary significantly with the choice of model and local explanation.
+However, for reference the FIFA data, 5000 observation of 9 explanatory variable, took 1.5 seconds to create PCA and Mahalanobis distances for both the data and attribution spaces.
+On the same data, a modestly hyper-parametered random forest model fit in 2.6 seconds, while extracting the tree SHAP values of each observation took 278 seconds combined.
+These time were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the run time will be spent on the local attribution.
+This makes tree SHAP a good candidate to start with.
+The package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes that are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+
+The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters.
+The application is made with \textbf{shiny} (Chang et al.
+2021).
+The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (\textbf{sievet\_interactive\_2020?}).
+The tour view is created in \textbf{spinifex} (Spyrison and Cook 2020), which creates the manual tour basis array, and facilitates similar rendering to html widget.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
+Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
+
 \leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
@@
 
-\leavevmode\hypertarget{ref-asimov_grand_1994}{}%
-Asimov, Daniel A., and Andreas Buja.
-1994.
-{``Grand Tour via Geodesic Interpolation of 2-Frames.''} In \emph{Visual Data Exploration and Analysis}, 2178:145--53.
-International Society for Optics; Photonics.
-\url{https://doi.org/10.1117/12.172065}.
+\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
+Asimov, Daniel.
+1985.
+{``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
+https://doi.org/\url{https://doi.org/10.1137/0906011}.
+
+\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
+Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
+
+\leavevmode\hypertarget{ref-buja_grand_1986}{}%
+Buja, Andreas, and Daniel Asimov.
+1986.
+{``Grand {Tour} {Methods}: {An} {Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer} {Science} and {Statistics}}, 63--67.
+New York, NY, USA: Elsevier North-Holland, Inc.
+\url{http://dl.acm.org/citation.cfm?id=26036.26046}.
+
+\leavevmode\hypertarget{ref-chang_shiny_2021}{}%
+Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
+2021.
+\emph{Shiny: Web Application Framework for r}.
+\url{https://CRAN.R-project.org/package=shiny}.
 
@@
 1997.
-{``Manual Controls for High-Dimensional Data Projections.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
+{``Manual {Controls} for {High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
 \url{https://doi.org/10.2307/1390747}.
 
+\leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
+Dastin, Jeffrey.
+2018.
+{``Amazon Scraps Secret {AI} Recruiting Tool That Showed Bias Against Women.''} \emph{Reuters}, October.
+\url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.
+
+\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
+Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
+2018.
+{``Addressing Age-Related Bias in Sentiment Analysis.''} In \emph{Proceedings of the 2018 Chi Conference on Human Factors in Computing Systems}, 1--14.
+
+\leavevmode\hypertarget{ref-duffy_apple_2019}{}%
+Duffy, Claire.
+2019.
+{``Apple Co-Founder {Steve} {Wozniak} Says {Apple} {Card} Discriminated Against His Wife.''} \emph{CNN}, November.
+\url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
+
+\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
+Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
+
+\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
+Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+
+\leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
+Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
+
+\leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
+Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
+
+\leavevmode\hypertarget{ref-larson_how_2016}{}%
+Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
+2016.
+{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''} \emph{ProPublica}, May.
+\url{https://www.propublica.org/article/how-we-analyzed-the-compas-recidivism-algorithm?token=RPR1E2qtzJltfJ0tS-gB_41kmfoWZAu4}.
+
+\leavevmode\hypertarget{ref-lazer_parable_2014}{}%
+Lazer, David, Ryan Kennedy, Gary King, and Alessandro Vespignani. 2014. {``The Parable of {Google} {Flu}: Traps in Big Data Analysis.''} \emph{Science} 343 (6176): 1203--5.
+
 \leavevmode\hypertarget{ref-lee_review_2021}{}%
-Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Earo Wang, Nick Spyrison, and H.
+Lee, Stuart, Dianne Cook, Natalia Da Silva, Ursula Laa, Earo Wang, Nick Spyrison, and H.
 Sherry Zhang.
 2021.
-{``A {Review} of the {State}-of-the-{Art} on {Tours} for {Dynamic} {Visualization} of {High}-{Dimensional} {Data}.''} \emph{arXiv:2104.08016 {[}Cs, Stat{]}}, April.
-\url{http://arxiv.org/abs/2104.08016}.
+{``A {Review} of the {State}-of-the-{Art} on {Tours} for {Dynamic} {Visualization} of {High}-Dimensional {Data}.''} \emph{arXiv Preprint arXiv:2104.08016}.
+
+\leavevmode\hypertarget{ref-lee_pptree_2013}{}%
+Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
 
@@
 
+\leavevmode\hypertarget{ref-liaw_classification_2002}{}%
+Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
+
+\leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
+Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
+
 \leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
@@
 
+\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
+O'neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
+
+\leavevmode\hypertarget{ref-salzberg_why_2014}{}%
+Salzberg, Steven. 2014. {``Why {Google} {Flu} {Is} {A} {Failure}.''} \emph{Forbes}, March. \url{https://www.forbes.com/sites/stevensalzberg/2014/03/23/why-google-flu-is-a-failure/}.
+
+\leavevmode\hypertarget{ref-shapley_value_1953}{}%
+Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
+
+\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
+Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
+
+\leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
+Silva, Natalia da, Dianne Cook, and Eun-Kyung Lee.
+2021.
+{``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
+
 \leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
-Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An r Package for Creating a Manual Tour of Low-Dimensional Projections of Multivariate Data.''} \emph{The R Journal} 12 (1): (accepted).
+Spyrison, Nicholas, and Dianne Cook.
+2020.
+{``Spinifex: An {R} {Package} for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of {Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
+\url{https://doi.org/10.32614/RJ-2020-027}.
+
+\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
+Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
+
+\leavevmode\hypertarget{ref-strumbelj_explaining_2014}{}%
+Štrumbelj, Erik, and Igor Kononenko. 2014. {``Explaining Prediction Models and Individual Predictions with Feature Contributions.''} \emph{Knowledge and Information Systems} 41 (3): 647--65.
+
+\leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
+Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
+Wickham, Hadley, Dianne Cook, and Heike Hofmann.
+2015.
+{``Visualizing Statistical Models: {Removing} the Blindfold.''} \emph{Statistical Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
+\url{https://doi.org/10.1002/sam.11271}.
 

### 2021-11-01 _paper.rmd
- remove maha - touch up times - add global_view capture and text.

@@
 ## Global view
+
+```{r globalview, echo=F, out.width = "100%", fig.cap = "Global view; approximations of the data and SHAP-spaces.
+Orietation of the basis contributions are illustrated on a unit circle.
+Linked brushing allows pbservations brushed in one plot to be selected in others.
+This selection is also used to in the proceeding view and their corresponding informaion is displayed in an interactive table.
+Hovering the cursor over an observation displays a tooltip with row number/name information.
+In the classification case, misclassified points are circled in red.
+This view prvides an orientenation to select a primary and comparison observation, key targets in the following tour."}
+knitr::include_graphics("./figures/global_view_penguins.PNG")
+```
+
 ## Cheem tour
@@
 	\item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components.
-The within class separability of points can also optionally be displayed as QQ plots of the Mahalanobis distances within those spaces.
 \end{itemize}
@@
 The time to preprocess the data will vary significantly with the choice of model and local explanation.
-However, for reference the FIFA data, 5000 observation of 9 explanatory variable, took 1.5 seconds to create PCA and Mahalanobis distances for both the data and attribution spaces.
-On the same data, a modestly hyper-parametered random forest model fit in 2.6 seconds, while extracting the tree SHAP values of each observation took 278 seconds combined.
+However, for reference the FIFA data, 5000 observation of 9 explanatory variable, took 0.6 seconds to create PCA for both the data and attribution spaces.
+On the same data, a modestly hyper-parametered random forest model fit in 2.9 seconds, while extracting the tree SHAP values of each observation took 254 seconds combined.
 These time were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the run time will be spent on the local attribution.

### 2021-11-01 _paper.tex
- remove maha - touch up times - add global_view capture and text.

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/global_view_penguins}
+
+}
+
+\caption{Global view; approximations of the data and SHAP-spaces.
+Orietation of the basis contributions are illustrated on a unit circle.
+Linked brushing allows pbservations brushed in one plot to be selected in others.
+This selection is also used to in the proceeding view and their corresponding informaion is displayed in an interactive table.
+Hovering the cursor over an observation displays a tooltip with row number/name information.
+In the classification case, misclassified points are circled in red.
+This view prvides an orientenation to select a primary and comparison observation, key targets in the following tour.}\label{fig:globalview}
+\end{figure}
+
 \hypertarget{cheem-tour}{%
@@
     \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components.
-The within class separability of points can also optionally be displayed as QQ plots of the Mahalanobis distances within those spaces.
 \end{itemize}

### 2021-11-02 _paper.rmd
grammerly and the tour section.

@@
 XAI attempts to shed light on these models, one such approach is the use of local explanations.
-A local explanation of a model give a point-estimate of linear variable importance in the vicinity of one observation.
+A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
 We extract explanations for each observation, and approximate data and this attribution space side-by-side with linked brushing.
@@
 Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
-The implementation of our framework is available as an R package called cheem available at [github.com/nspyrison/cheem](https://github.com/nspyrison/cheem).
+The implementation of our framework is available as an R package **cheem** available at [github.com/nspyrison/cheem](https://github.com/nspyrison/cheem).
 # documentclass: jdssv
@@
 In 1886 Francis Galton coined the term _regression_ to refer to continuous, quantitative predictions.
-While _classification_ refers discretion predictions as introduced by Fisher in 1936.
+While _classification_ refers to discrete predictions as introduced by Fisher in 1936.
 
 <!-- introduce explanatory vs predictive modeling -->
-Breiman and Shmueli [@breiman_statistical_2001, @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling is performed for to predict new or future out-of-sample observations.
+Breiman and Shmueli [@breiman_statistical_2001, @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling is performed to predict new or future out-of-sample observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
-In explanatory modeling the interpretable is a key feature for drawing inferential conclusions.
+In explanatory modeling, the interpretable is a key feature for drawing inferential conclusions.
 While predictive modeling may opt for potentially more accurate black-box models.
@@
 <!-- ## XAI & interpretability crisis -->
-Predictive model and black box modeling is becoming increasingly common, but not without controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
+Predictive model and black-box modeling is becoming increasingly common, but not without controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
 Applications have been known to reflect common biases against sex [@dastin_amazon_2018, @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-This is a common issue stemming from biases the in sample data are violate ethical principals.
-Another issue is that of data-drift, when new data is outside the support of latent or exogenous explanatory variables.
+This is a common issue stemming from biases in the in-sample, training data are violate ethical principles.
+Another issue is that of data-drift when new data is outside the support of latent or exogenous explanatory variables.
 Data-drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
@@
 One branch of XAI is local explanations, which take a variable attribution approach to bring transparency to a model.
-Local explanations attempt to approximate a linear variable importance at the location of one observation.
+Local explanations attempt to approximate linear variable importance at the location of one observation.
 There are many such local explanations, any of which is works with our approach (assuming model-explanation compatibility).
@@
 <!-- ## Data visualization tours -->
-In multivariate data visualization a _tour_ [@asimov_grand_1985, @buja_grand_1986, @lee_review_2021] is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
+In multivariate data visualization a _tour_ [@asimov_grand_1985, @buja_grand_1986, @lee_review_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
 Tours are viewed as an animation over small changes to a projection basis.
 Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
-The intuition is similar to watching the shadow of hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
-<!-- ## manual tours --> There are various types of tours, which are distinguished be the generation of sequence of projection bases.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
+<!-- ## manual tours --> There are various types of tours, which are distinguished by the generation of the sequence of projection bases.
 In a _manual_ tour [@cook_manual_1997, @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
@@
 
-The approach purposed below is to use the manual tour as means to interrogate a local explanation; it see if its variable importance are good explanation for the model predictions.
+The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is good explanation for the model predictions.
 We make R package `cheem` with an interactive application to facilitate analysis.
@@
 
-The rest of paper is organized as follows.
+The rest of this paper is organized as follows.
 The next section [SHAP](#sec:SHAP) covers the background of the local explanation SHAP and the traditional visuals produced from it.
@@
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
-We conclude with [Discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
+We conclude with a [Discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
 
@@
 SHaply Additive exPlanations, or SHAP [@lundberg_unified_2017] approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score [@shapley_value_1953].
+This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players that contribute to the score [@shapley_value_1953].
 
@@
 TO illustrate SHAP and its original use we use soccer data from FIFA 2020 season [@leone_fifa_2020].
-We have 5000 observations of 9 aggregated skill measures and use a use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
+We have 5000 observations of 9 aggregated skill measures and use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
 We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
-We expect to see a difference in attribution of the variable importance across the two positions of the players.
+We expect to see a difference in the attribution of the variable importance across the two positions of the players.
 
@@
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the prediction of the defensive player.
 We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
 Panel c) shows a simplified breakdown plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing sum of the SHAP values.
+Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
 
@@
 These explanations make sense; the variable importances make sense in light of the position of the player.
-c) Breakdown plots of the observations the explanation used to additively explain the difference beween the intercept and prediction"}
+c) Breakdown plots of the observations the explanation used to additively explain the difference between the intercept and prediction"}
 knitr::include_graphics("./figures/shap_distr_bd.pdf")
@@
 In summary, this highlights how local explanations bring transparency to a model at least in the vicinity of their observations.
-In this instance we showed how two very different soccer players receive different profiles of variable importance to explain the the prediction of their wages.
-In the following section we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
+In this instance, we showed how two very different soccer players receive different profiles of variable importance to explain the prediction of their wages.
+In the following section, we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
 
@@
 Below we illustrate the two primary displays of the application: the global view and the tour view.
-Then we'll cover what we takeaway from the classification and regression tasks.
+Then we'll cover what we take away from the classification and regression tasks.
 Lastly, we discuss the preprocessing that needs to be done before display.
 
-<!-- TODO:XXX: -->
 ## Global view
 
-```{r globalview, echo=F, out.width = "100%", fig.cap = "Global view; approximations of the data and SHAP-spaces.
-Orietation of the basis contributions are illustrated on a unit circle.
-Linked brushing allows pbservations brushed in one plot to be selected in others.
-This selection is also used to in the proceeding view and their corresponding informaion is displayed in an interactive table.
+The global view is an important context for exploring the separability of the data- and the local explanation's attribution-spaces, and is crucial in the selection of explanation to further interrogate and explore the structural sensitivity of.
+
+We show an approximation of these spaces with a projection through their first two principal components.
+The orientation of the variables are shown inscribed on a unit circle.
+While a single 2D projection will rarely encompass all of the structure of a higher-dimensional space, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
+
+```{r globalview, echo=F, out.width = "100%", fig.cap = "Global view screen capture; the approximations of the data and SHAP-spaces of the penguins data.
+Orientation of the basis contributions is illustrated on a unit circle.
+Linked brushing allows observations brushed in one plot to be selected in others.
+This selection is also used in the proceeding view and their corresponding information is displayed in an interactive table.
 Hovering the cursor over an observation displays a tooltip with row number/name information.
 In the classification case, misclassified points are circled in red.
-This view prvides an orientenation to select a primary and comparison observation, key targets in the following tour."}
+This view prpvides an orientation to select a primary and comparison observation, key targets in the following tour."}
 knitr::include_graphics("./figures/global_view_penguins.PNG")
@@
 
+This view offers dynamic interaction in several ways.
+A tooltip on hovering over a point that displays the row number/name and classification information if appropriate.
+Linked brushing allows for the selection of points (by click and drag) where those points will be highlighted in both plots.
+The information corresponding to the selected points is populated on a sortable table and the data powering the proceeding tour will also subset the data to the current selection.
+
 ## Cheem tour
+
+The primary observation identified via the global view is foundational to the production of the cheem tour.
+Namely, the linear attribution of that variable is used as a 1D projection basis.
+This is the approximate contributions of the variables that this model uses to justify its prediction for the observation.
+
+That normalized attribution of the primary observation is depicted as stacked bars where the horizontal width is the contribution.
+The bottom of this display is divided by the use case.
+In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
+In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of the observations is fixed to its prediction or residual.
+
+```{r cheemtour, echo=F, out.width = "100%", fig.cap = "Cheem tour screen capture; the primary observation's normalized SHAP values are the initial basis.
+This is the explanation of the linear variable importances for the model at this observation.
+The point in question was misclassified, it was predicted to be from the green cluster, while it was observed to be from the orange cluster.
+The top shows the contributions of the variables, with the dashed line being the primary observation's SHAP, its shape is closer to the orange cluster (observed group) while the position is pulled more toward the green (predicted group).
+The bottom shows the 1D projection with density and rug marks below.
+The dashed line is plausibly in the middle of the green density, the story that the explanation trying to sell us.
+Yet, when we play the tour animating on the contribution of bill length (b_l), the bottom dashed line is more regularly in the center of the observed orange cluster."}
+knitr::include_graphics("./figures/cheem_tour_penguins.PNG")
+```
+
+Data visualization tours animate many linear projections over small changes to the basis.
+The manual tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
+Doing allows an analyst to test an individual variable's sensitivity to the structure identified in the frame.
+The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution.
+In the next sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.
+
 ## Classifcation task
+
 ## Regression task
@@
 The benefit of having dynamic interaction with data is predicated on a reasonably small render time.
-The it is important to preprocess as much work as possible so that application resources can be used efficiently.
+It is important to preprocess as much work as possible so that application resources can be used efficiently.
 Below we discuss the steps and details of the reprocessing.
@@
 	\item \textbf{Model:} any model can be used with this method.
-Currently we apply random forest models via the package \textbf{randomForest} (ref:citeRf) to mitigate the runtime of our local explanation which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} (ref:citeRf) to mitigate the runtime of our local explanation which requires tree-based models.
 	\item \textbf{Local explanation:} any model-compatible linear explanation could be used.
@@
 The time to preprocess the data will vary significantly with the choice of model and local explanation.
-However, for reference the FIFA data, 5000 observation of 9 explanatory variable, took 0.6 seconds to create PCA for both the data and attribution spaces.
+However, for reference, the FIFA data, 5000 observations of 9 explanatory variables, took 0.6 seconds to create PCA for both the data and attribution spaces.
 On the same data, a modestly hyper-parametered random forest model fit in 2.9 seconds, while extracting the tree SHAP values of each observation took 254 seconds combined.
-These time were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the run time will be spent on the local attribution.
+These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the run time will be spent on the local attribution.
 This makes tree SHAP a good candidate to start with.
@@
 # Case Studies {#sec:casestudies}
-## 1) Penguins speicies classification
+## 1) Penguins species classification
 ## 2) FIFA wage regression
@@
 
-# Discsussion
+# Discussion {#sec:discussion}
 
-# Acknoledgements
+# Acknowledgments
 

### 2021-11-02 _paper.tex
grammerly and the tour section.

@@
 Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
-The implementation of our framework is available as an R package called cheem available at \href{https://github.com/nspyrison/cheem}{github.com/nspyrison/cheem}.
+The implementation of our framework is available as an R package \textbf{cheem} available at \href{https://github.com/nspyrison/cheem}{github.com/nspyrison/cheem}.
 \end{abstract}
@@
 In 1886 Francis Galton coined the term \emph{regression} to refer to continuous, quantitative predictions.
-While \emph{classification} refers discretion predictions as introduced by Fisher in 1936.
+While \emph{classification} refers to discrete predictions as introduced by Fisher in 1936.
 
-Breiman and Shmueli Shmueli (2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed for to predict new or future out-of-sample observations.
+Breiman and Shmueli Shmueli (2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed to predict new or future out-of-sample observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
-In explanatory modeling the interpretable is a key feature for drawing inferential conclusions.
+In explanatory modeling, the interpretable is a key feature for drawing inferential conclusions.
 While predictive modeling may opt for potentially more accurate black-box models.
@@
 
-Predictive model and black box modeling is becoming increasingly common, but not without controversy and issues Kodiyan (2019).
+Predictive model and black-box modeling is becoming increasingly common, but not without controversy and issues Kodiyan (2019).
 Applications have been known to reflect common biases against sex Duffy (2019), race (Larson et al.
@@
 2018).
-This is a common issue stemming from biases the in sample data are violate ethical principals.
-Another issue is that of data-drift, when new data is outside the support of latent or exogenous explanatory variables.
+This is a common issue stemming from biases in the in-sample, training data are violate ethical principles.
+Another issue is that of data-drift when new data is outside the support of latent or exogenous explanatory variables.
 Data-drift can lead to worse predictions Salzberg (2014).
@@
 Lee et al.
-(2021) is a sequence of linear projections of data onto a lower dimensional space, typically 1-3D.
+(2021) is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
 Tours are viewed as an animation over small changes to a projection basis.
 Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
-The intuition is similar to watching the shadow of hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
-There are various types of tours, which are distinguished be the generation of sequence of projection bases.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
+There are various types of tours, which are distinguished by the generation of the sequence of projection bases.
 In a \emph{manual} tour Spyrison and Cook (2020) this path is defined by changing the contribution of a selected variable.
@@
 
-The approach purposed below is to use the manual tour as means to interrogate a local explanation; it see if its variable importance are good explanation for the model predictions.
+The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is good explanation for the model predictions.
 We make R package \texttt{cheem} with an interactive application to facilitate analysis.
@@
 
-The rest of paper is organized as follows.
+The rest of this paper is organized as follows.
 The next section \protect\hyperlink{sec:SHAP}{SHAP} covers the background of the local explanation SHAP and the traditional visuals produced from it.
@@
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
-We conclude with \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
+We conclude with a \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
 
@@
 Lundberg and Lee 2017) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players contributing to the score (Shapley 1953).
+This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players that contribute to the score (Shapley 1953).
 
 TO illustrate SHAP and its original use we use soccer data from FIFA 2020 season (Leone 2020).
-We have 5000 observations of 9 aggregated skill measures and use a use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
+We have 5000 observations of 9 aggregated skill measures and use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
 We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
-We expect to see a difference in attribution of the variable importance across the two positions of the players.
+We expect to see a difference in the attribution of the variable importance across the two positions of the players.
 
@@
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the the prediction of the defensive player.
+In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the prediction of the defensive player.
 We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
 Panel c) shows a simplified breakdown plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing sum of the SHAP values.
+Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
 
@@
 These explanations make sense; the variable importances make sense in light of the position of the player.
-c) Breakdown plots of the observations the explanation used to additively explain the difference beween the intercept and prediction}\label{fig:shapdistrbd}
+c) Breakdown plots of the observations the explanation used to additively explain the difference between the intercept and prediction}\label{fig:shapdistrbd}
 \end{figure}
@@
 Below we illustrate the two primary displays of the application: the global view and the tour view.
-Then we'll cover what we takeaway from the classification and regression tasks.
+Then we'll cover what we take away from the classification and regression tasks.
 Lastly, we discuss the preprocessing that needs to be done before display.
@@
 \subsection{Global view}\label{global-view}}
+
+The global view is an important context for exploring the separability of the data- and the local explanation's attribution-spaces, and is crucial in the selection of explanation to further interrogate and explore the structural sensitivity of.
+
+We show an approximation of these spaces with a projection through their first two principal components.
+The orientation of the variables are shown inscribed on a unit circle.
+While a single 2D projection will rarely encompass all of the structure of a higher-dimensional space, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
 
@@
 
-\caption{Global view; approximations of the data and SHAP-spaces.
-Orietation of the basis contributions are illustrated on a unit circle.
-Linked brushing allows pbservations brushed in one plot to be selected in others.
-This selection is also used to in the proceeding view and their corresponding informaion is displayed in an interactive table.
+\caption{Global view screen capture; the approximations of the data and SHAP-spaces of the penguins data.
+Orientation of the basis contributions is illustrated on a unit circle.
+Linked brushing allows observations brushed in one plot to be selected in others.
+This selection is also used in the proceeding view and their corresponding information is displayed in an interactive table.
 Hovering the cursor over an observation displays a tooltip with row number/name information.
 In the classification case, misclassified points are circled in red.
-This view prvides an orientenation to select a primary and comparison observation, key targets in the following tour.}\label{fig:globalview}
+This view prpvides an orientation to select a primary and comparison observation, key targets in the following tour.}\label{fig:globalview}
 \end{figure}
+
+This view offers dynamic interaction in several ways.
+A tooltip on hovering over a point that displays the row number/name and classification information if appropriate.
+Linked brushing allows for the selection of points (by click and drag) where those points will be highlighted in both plots.
+The information corresponding to the selected points is populated on a sortable table and the data powering the proceeding tour will also subset the data to the current selection.
 
@@
 \subsection{Cheem tour}\label{cheem-tour}}
+
+The primary observation identified via the global view is foundational to the production of the cheem tour.
+Namely, the linear attribution of that variable is used as a 1D projection basis.
+This is the approximate contributions of the variables that this model uses to justify its prediction for the observation.
+
+That normalized attribution of the primary observation is depicted as stacked bars where the horizontal width is the contribution.
+The bottom of this display is divided by the use case.
+In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
+In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of the observations is fixed to its prediction or residual.
+
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_penguins}
+
+\}
+
+\textbackslash caption\{Cheem tour screen capture; the primary observation's normalized SHAP values are the initial basis.
+This is the explanation of the linear variable importances for the model at this observation.
+The point in question was misclassified, it was predicted to be from the green cluster, while it was observed to be from the orange cluster.
+The top shows the contributions of the variables, with the dashed line being the primary observation's SHAP, its shape is closer to the orange cluster (observed group) while the position is pulled more toward the green (predicted group).
+The bottom shows the 1D projection with density and rug marks below.
+The dashed line is plausibly in the middle of the green density, the story that the explanation trying to sell us.
+Yet, when we play the tour animating on the contribution of bill length (b\_l), the bottom dashed line is more regularly in the center of the observed orange cluster.\}\label{fig:cheemtour}
+\textbackslash end\{figure\}
+
+Data visualization tours animate many linear projections over small changes to the basis.
+The manual tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
+Doing allows an analyst to test an individual variable's sensitivity to the structure identified in the frame.
+The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution.
+In the next sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.
 
@@
     \item \textbf{Model:} any model can be used with this method.
-Currently we apply random forest models via the package \textbf{randomForest} (Liaw and Wiener 2002) to mitigate the runtime of our local explanation which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} (Liaw and Wiener 2002) to mitigate the runtime of our local explanation which requires tree-based models.
     \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
@@
 
-\hypertarget{penguins-speicies-classification}{%
-\subsection{1) Penguins speicies classification}\label{penguins-speicies-classification}}
+\hypertarget{penguins-species-classification}{%
+\subsection{1) Penguins species classification}\label{penguins-species-classification}}
 
@@
 
-\hypertarget{discsussion}{%
-\section{Discsussion}\label{discsussion}}
+\hypertarget{sec:discussion}{%
+\section{Discussion}\label{sec:discussion}}
 
-\hypertarget{acknoledgements}{%
-\section{Acknoledgements}\label{acknoledgements}}
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
 

### 2021-11-03 _paper.rmd
classifciation task.

@@
 The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution.
-In the next sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.
+In the following sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.
 
@@
 
+What information do we glean from using this method on a classifcation task?
+Typically we select a misclassified observation in comparison with correctly classified point that is nearby in data space.
+We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
+By default the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
+That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+
+Another way of thinking about this would be: what would the predictions be if the explanation were different.
+This is similar to the idea of _ceteris paribus_ profiles [@biecek_ceterisparibus_2020].
+_Ceteris paribus_ is Latin for “other things held constant” or “all else unchanged”.
+The profiles visualize what 'what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant.
+In contrast to _ceteris paribus_ profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.
+
 ## Regression task
+
+The regression case is not as discrete a feature. Instead the prediction or the residual becomes the comparison of accuracy.
 
@@
 The application is made with __shiny__ [@chang_shiny_2021].
-The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievet_interactive_2020].
+The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievert_interactive_2020].
 The tour view is created in __spinifex__ [@spyrison_spinifex_2020], which creates the manual tour basis array, and facilitates similar rendering to html widget.

### 2021-11-03 _paper.tex
classifciation task.

@@
 
+What information do we glean from using this method on a classifcation task?
+Typically we select a misclassified observation in comparison with correctly classified point that is nearby in data space.
+We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
+By default the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
+That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+
+Another way of thinking about this would be: what would the predictions be if the explanation were different.
+This is similar to the idea of \emph{ceteris paribus} profiles (\textbf{biecek\_ceterisparibus\_2020?}).
+\emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant.
+In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.
+
 \hypertarget{regression-task}{%
 \subsection{Regression task}\label{regression-task}}
+
+The regression case is not as discrete a feature. Instead the prediction or the residual becomes the comparison of accuracy.
 
@@
 2021).
-The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (\textbf{sievet\_interactive\_2020?}).
+The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (Sievert 2020).
 The tour view is created in \textbf{spinifex} (Spyrison and Cook 2020), which creates the manual tour basis array, and facilitates similar rendering to html widget.
@@
 
+\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
+Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
+
 \leavevmode\hypertarget{ref-da_silva_projection_2021}{}%

### 2021-11-04 _paper.rmd
starting to make use case figures.

@@
 It is important to preprocess as much work as possible so that application resources can be used efficiently.
+The work remaining at runtime should be solely responding to inputs and the rendering of figures and tables.
 Below we discuss the steps and details of the reprocessing.
-<!-- Fitting models regularly has non-negligible runtime and the exaction of a single observations local explanations is a similarly computationally expensive operation. -->
 
@@
 
-<!-- render time -->
-The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters.
+# Software Infrastructure {#sec:softwareinfrastructure}
+
+The above described method and application are implemented as an open source __R__ package, __cheem__ _TODO:XXX site github?
+cran?_.
+Preprocessing was facilited with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
 The application is made with __shiny__ [@chang_shiny_2021].
-The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievert_interactive_2020].
-The tour view is created in __spinifex__ [@spyrison_spinifex_2020], which creates the manual tour basis array, and facilitates similar rendering to html widget.
+The tour visual is an extension of __spinifex__ [@spyrison_spinifex_2020].
+Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive html widget with __plotly__ [@sievert_interactive_2020].
+The free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
 
-# Software Infrastructure {#sec:softwareinfrastructure}
-## Extend spinifex, consume DALEX & treeshap
-## Preprocess
-## Runtime rendering
+## Preprocessing
+
+Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
+It returns an a list containing data frames to be consumed by the application, higher level model performance, and runtime information.
+The return of this function can be then be saved to an .rds file for consumption in the application.
+
+## Linking brushing and highlighting
+
+The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
+Linked brushing also for a rectangular selection of observations.
+These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance highlighting in the tour.
+The distinction between the former and latter is subtle, but important for developers to grasp.
+In the former, self-contained html widgets can highlight directly within javascript without evaluation evaluation downstream in other reactive functions.
+In the latter, the use of `plotly::event_data()` returns the identity of the selected observations.
+This will require a reactive flush, but after that, the sky is the ceiling.
 
@@
 
+We would like to thank Professor Przemyslaw Biecek for his input in early in the project and to the broader MI\^2 lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages.
+This research was supported in part by Australian government Research Training Program (RTP) scholarships.
+
 # References

### 2021-11-04 figures/_make_figures_case_study.r

@@ -1,0 +1,164 @@
+# CASE STUDIES FOR CHEEM ----
+require("cheem")
+require("spinifex")
+require("dplyr")
+require("ggplot2")
+
+## Local function variants:
+
+static_global_view <- function(
+  cheem_ls,
+  primary_obs = NULL,
+  comparison_obs = NULL
+){
+  ## Prevent global variable warnings:
+  V1 <- V2 <- ggtext <- projection_nm <- layer_name <- tooltip <- NULL
+  .alpha <- logistic_tform(nrow(cheem_ls$decode_df), mid_pt = 500L)
+  global_view_df <- cheem_ls$global_view_df ## Init
+  is_classification <- cheem_ls$problem_type == "classification"
+  pred_clas <- as.factor(FALSE) ## If regression; dummy pred_clas
+  if(is_classification == TRUE) pred_clas <-
+    cheem_ls$decode_df$predicted_class %>%
+    rep_len(nrow(global_view_df)) %>%
+    as.factor()
+  
+  pts_highlight <- list()
+  ## Red misclassified points, if present
+  if(is_classification == TRUE){
+    .rn_misclass <- which(cheem_ls$decode_df$is_misclassified == TRUE)
+    .idx_misclas <- global_view_df$rownum %in% .rn_misclass
+    if(sum(.idx_misclas) > 0L){
+      .df <- global_view_df[.idx_misclas, ] # %>% highlight_key(~rownum)
+      pts_highlight <- c(
+        pts_highlight,
+        ggplot2::geom_point(ggplot2::aes(V1, V2), .df,
+                            color = "red", fill = NA,
+                            shape = 21L, size = 3L, alpha = .alpha)
+      )
+    }
+  }
+  ## Highlight comparison obs, if passed
+  if(is.null(comparison_obs) == FALSE){
+    .idx_comp <- global_view_df$rownum == comparison_obs
+    if(sum(.idx_comp) > 0L){
+      .df <- global_view_df[.idx_comp, ]
+      pts_highlight <- c(
+        pts_highlight,
+        ## Highlight comparison obs
+        ggplot2::geom_point(ggplot2::aes(V1, V2), #, color = pred_clas[.idx_comp]),
+                            .df, size = 3L, shape = 4L, color = "black")
+      )
+    }
+  }
+  ## Highlight shap obs, if passed
[…]

### 2021-11-05 _paper.rmd
init fifa and penguins structure for coffee and chocolates.

@@
 # Case Studies {#sec:casestudies}
-## 1) Penguins species classification
-## 2) FIFA wage regression
-## 3)?
-## 4)?
+## 1) penguins, species classification
+
+```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Penguins classificiation TODO: XXX"}
+knitr::include_graphics("./figures/case_penguins.pdf")
+```
+
+## 2) Chocolates, milk/dark chocolate classification
+
+TODO:XXX
+
+## 3) FIFA, wage regression
+```{r casefifa, echo=F, out.width = "100%", fig.cap = "FIFA 2020, regressing wages [log 2020 Euros] from skill measures. TODO: XXX"}
+knitr::include_graphics("./figures/case_fifa.pdf")
+```
+
+## 4) Coffee, rating regression
+
+TODO:XXX
+
+regress score color on species?
 

### 2021-11-05 _paper.tex
init fifa and penguins structure for coffee and chocolates.

@@
 Another way of thinking about this would be: what would the predictions be if the explanation were different.
-This is similar to the idea of \emph{ceteris paribus} profiles (\textbf{biecek\_ceterisparibus\_2020?}).
+This is similar to the idea of \emph{ceteris paribus} profiles (Biecek 2020).
 \emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant.
@@
 It is important to preprocess as much work as possible so that application resources can be used efficiently.
+The work remaining at runtime should be solely responding to inputs and the rendering of figures and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 
-The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters.
-The application is made with \textbf{shiny} (Chang et al.
-2021).
-The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (Sievert 2020).
-The tour view is created in \textbf{spinifex} (Spyrison and Cook 2020), which creates the manual tour basis array, and facilitates similar rendering to html widget.
-
 \hypertarget{sec:softwareinfrastructure}{%
@@
 
-\hypertarget{extend-spinifex-consume-dalex-treeshap}{%
-\subsection{Extend spinifex, consume DALEX \& treeshap}\label{extend-spinifex-consume-dalex-treeshap}}
+The above described method and application are implemented as an open source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX site github?
+cran?}.
+Preprocessing was facilited with models created via \textbf{randomForest} {[}liaw\_classification\_2002{]}, and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
+2021).
+The application is made with \textbf{shiny} (Chang et al.
+2021).
+The tour visual is an extension of \textbf{spinifex} (Spyrison and Cook 2020).
+Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive html widget with \textbf{plotly} (Sievert 2020).
+The free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
 
-\hypertarget{preprocess}{%
-\subsection{Preprocess}\label{preprocess}}
+\hypertarget{preprocessing-1}{%
+\subsection{Preprocessing}\label{preprocessing-1}}
 
-\hypertarget{runtime-rendering}{%
-\subsection{Runtime rendering}\label{runtime-rendering}}
+Provided with tidy and transformed data, the function \texttt{cheem\_ls} is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
+It returns an a list containing data frames to be consumed by the application, higher level model performance, and runtime information.
+The return of this function can be then be saved to an .rds file for consumption in the application.
+
+\hypertarget{linking-brushing-and-highlighting}{%
+\subsection{Linking brushing and highlighting}\label{linking-brushing-and-highlighting}}
+
+The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
+Linked brushing also for a rectangular selection of observations.
+These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance highlighting in the tour.
+The distinction between the former and latter is subtle, but important for developers to grasp.
+In the former, self-contained html widgets can highlight directly within javascript without evaluation evaluation downstream in other reactive functions.
+In the latter, the use of \texttt{plotly::event\_data()} returns the identity of the selected observations.
+This will require a reactive flush, but after that, the sky is the ceiling.
 
@@
 \hypertarget{penguins-species-classification}{%
-\subsection{1) Penguins species classification}\label{penguins-species-classification}}
+\subsection{1) penguins, species classification}\label{penguins-species-classification}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins}
+
+}
+
+\caption{Penguins classificiation TODO: XXX}\label{fig:casepenguins}
+\end{figure}
+
+\hypertarget{chocolates-milkdark-chocolate-classification}{%
+\subsection{2) Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
+
+TODO:XXX
 
 \hypertarget{fifa-wage-regression}{%
-\subsection{2) FIFA wage regression}\label{fifa-wage-regression}}
+\subsection{3) FIFA, wage regression}\label{fifa-wage-regression}}
 
-\hypertarget{section}{%
-\subsection{3)?}\label{section}}
+\begin{figure}
 
-\hypertarget{section-1}{%
-\subsection{4)?}\label{section-1}}
+{\centering \includegraphics[width=1\linewidth]{./figures/case_fifa}
+
+}
+
+\caption{FIFA 2020, regressing wages [log 2020 Euros] from skill measures. TODO: XXX}\label{fig:casefifa}
+\end{figure}
+
+\hypertarget{coffee-rating-regression}{%
+\subsection{4) Coffee, rating regression}\label{coffee-rating-regression}}
+
+TODO:XXX
+
+regress score color on species?
 
@@
 \section{Acknowledgments}\label{acknowledgments}}
+
+We would like to thank Professor Przemyslaw Biecek for his input in early in the project and to the broader MI\^{}2 lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages.
+This research was supported in part by Australian government Research Training Program (RTP) scholarships.
 
@@
 https://doi.org/\url{https://doi.org/10.1137/0906011}.
+
+\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
+Biecek, Przemyslaw. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
+
+\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
 

### 2021-11-05 figures/_make_figures_case_study.r

@@ -3,6 +3,7 @@ require("cheem")
 require("spinifex")
 require("dplyr")
 require("ggplot2")
+require("cowplot")
 
 ## Local function variants:
 
@@ -107,41 +108,46 @@ static_global_view <- function(
   Y <- as.integer(raw$species)
 }
 
-### cheem_ls -----
+### cheem_ls
 cheem_ls_peng <- cheem_ls( ## cheem_ls total: 3.51 sec elapsed
   x = X, y = Y, basis_type = "pca", class = clas)
 names(cheem_ls_peng)
 prim_obs <- 15L
 comp_obs <- 282L
 
-### Still shots of global and tours
+### Global view and cheem tour stills
+glob_view <- cheem::linked_global_view(cheem_ls_peng, prim_obs, comp_obs)
 glob_view <- static_global_view(
   cheem_ls_peng, prim_obs, comp_obs)
 
 bas <- basis_local_attribution(
   cheem_ls_peng$attr_df, rownum = prim_obs)
 ggt <- radial_cheem_ggtour(
-  cheem_ls_peng, basis=bas, mv_name=colnames(X)[1], 
-  primary_obs=prim_obs, comparison_obs=comp_obs)
+  cheem_ls_peng, basis=bas, mv_name=colnames(X)[1],
+  primary_obs=prim_obs, comparison_obs=comp_obs,
+  pcp_shape = 124, angle = 5)
+cheem_stills <- spinifex::filmstrip(ggt)
 
-### ch5_fig2_global_space -----
-
-
-### Save -----
+### Save
+cp <- cowplot::plot_grid(
+  glob_view + ggtitle("Global view"), 
+  cheem_stills + ggtitle("Cheem tour, select frames"),
+  labels = paste0(letters[1:2], ")"),
+  ncol = 1)#, rel_heights = c(1.5, 1))
 ggplot2::ggsave(
-  "./figures_from_script/ch5_fig2_global_space.pdf",
-  plot = p + theme(aspect.ratio=1), device = "pdf",
+  "./figures/case_penguins.pdf",
+  plot = cp, device = "pdf",
   width = 8, height = 4.8, units = "in")
 
-## Save interactive html widget
-ggp <- ggplotly(p, tooltip = "tooltip") %>%
-  config(displayModeBar = FALSE) %>% ## Remove html buttons
-  layout(dragmode = "select", showlegend = FALSE,
-         width = 640, height = 320) %>% ## Set drag left mouse
[…]

### 2021-11-09 _paper.rmd
- finish draft - grammerly edits - need to finish example cases.

--- 2021-11-04 _paper.rmd
@@
 <!-- ## XAI & interpretability crisis -->
-Predictive model and black-box modeling is becoming increasingly common, but not without controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
+Black-box models are becoming increasingly common, but not without teir share of controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
 Applications have been known to reflect common biases against sex [@dastin_amazon_2018, @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-This is a common issue stemming from biases in the in-sample, training data are violate ethical principles.
+These are all-too common issue, when endogenous or exogenous biases in the training data and identified and mimicked by models.
 Another issue is that of data-drift when new data is outside the support of latent or exogenous explanatory variables.
-Data-drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
+Data drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
 Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@adadi_peeking_2018, @arrieta_explainable_2020].
@@
 
-The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is good explanation for the model predictions.
+The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is a good explanation for the model predictions.
 We make R package `cheem` with an interactive application to facilitate analysis.
@@
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
-We conclude with a [Discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
+We conclude with a [discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
 
@@
 We show an approximation of these spaces with a projection through their first two principal components.
-The orientation of the variables are shown inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of a higher-dimensional space, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
+The orientation of the variables is shown inscribed on a unit circle.
+While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
 
-```{r globalview, echo=F, out.width = "100%", fig.cap = "Global view screen capture; the approximations of the data and SHAP-spaces of the penguins data.
+```{r globalview, echo=F, out.width = "100%", fig.cap = "Global view screen capture; the approximations of the data and SHAP-spaces of Palmer penguins data.
 Orientation of the basis contributions is illustrated on a unit circle.
@@
 In the classification case, misclassified points are circled in red.
-This view prpvides an orientation to select a primary and comparison observation, key targets in the following tour."}
+This view provides an orientation to select a primary and comparison observation, key targets in the following tour."}
 knitr::include_graphics("./figures/global_view_penguins.PNG")
@@
 This view offers dynamic interaction in several ways.
-A tooltip on hovering over a point that displays the row number/name and classification information if appropriate.
+A tooltip displays while the cursor hovers over a point displays the row number/name and classification information if appropriate.
 Linked brushing allows for the selection of points (by click and drag) where those points will be highlighted in both plots.
@@
 
-## Classifcation task
+## Classification task
 
-What information do we glean from using this method on a classifcation task?
-Typically we select a misclassified observation in comparison with correctly classified point that is nearby in data space.
+What information do we glean from using this method on a classification task?
+Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
 We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
-By default the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
 That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
@@
 _Ceteris paribus_ is Latin for “other things held constant” or “all else unchanged”.
-The profiles visualize what 'what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant.
-In contrast to _ceteris paribus_ profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.
+The profiles visualize what 'what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+In contrast to _ceteris paribus_ profiles, touring methods visualize all observations rather than one, While the variables are orthogonal dimensions, the 2D shadow they cast is not independent.
+That is, when one variable is rotated out of frame, other variables are effectively rotated into frame.
 
@@
 
-The regression case is not as discrete a feature. Instead the prediction or the residual becomes the comparison of accuracy.
+The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
 
@@
 \begin{itemize}
-	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic (color/shape) variable can be mapped typically a categorical variable exogenous to the model.
+	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic variable (color/shape)  can be mapped typically a categorical variable.
 	\item \textbf{Model:} any model can be used with this method.
@@
 However, for reference, the FIFA data, 5000 observations of 9 explanatory variables, took 0.6 seconds to create PCA for both the data and attribution spaces.
-On the same data, a modestly hyper-parametered random forest model fit in 2.9 seconds, while extracting the tree SHAP values of each observation took 254 seconds combined.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the run time will be spent on the local attribution.
+On the same data, with modest hyper parameters a random forest model fit in 2.9 seconds while extracting the tree SHAP values of each observation took 254 seconds combined.
+These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the time will be spent on the local attribution.
 This makes tree SHAP a good candidate to start with.
@@
 
-The above described method and application are implemented as an open source __R__ package, __cheem__ _TODO:XXX site github?
+The above-described method and application are implemented as an open-source __R__ package, __cheem__ _TODO:XXX site github?
 cran?_.
-Preprocessing was facilited with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
+Preprocessing was facilitated with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
 The application is made with __shiny__ [@chang_shiny_2021].
 The tour visual is an extension of __spinifex__ [@spyrison_spinifex_2020].
-Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive html widget with __plotly__ [@sievert_interactive_2020].
+Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive HTML widgets with __plotly__ [@sievert_interactive_2020].
 The free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
@@
 Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
-It returns an a list containing data frames to be consumed by the application, higher level model performance, and runtime information.
+It returns a list containing data frames to be consumed by the application, higher-level model performance, and runtime information.
 The return of this function can be then be saved to an .rds file for consumption in the application.
 
-## Linking brushing and highlighting
+## Interactive shiny application
 
@@
 Linked brushing also for a rectangular selection of observations.
-These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance highlighting in the tour.
+These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance, highlighting in the tour.
 The distinction between the former and latter is subtle, but important for developers to grasp.
-In the former, self-contained html widgets can highlight directly within javascript without evaluation evaluation downstream in other reactive functions.
+In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
 In the latter, the use of `plotly::event_data()` returns the identity of the selected observations.
@@
 # Case Studies {#sec:casestudies}
-## 1) Penguins species classification
-## 2) FIFA wage regression
-## 3)?
-## 4)?
+
+To illustrate the use of the cheem method we apply it to modern datasets, two classification examples and then two of regression.
+
+## 1) penguins, species classification
+
+Palm penguins data [@gorman_ecological_2014, @horst_palmerpenguins_2020] consist of four physical measurements of 3 species of penguins foraging near Plamer Station, Antarctica.
+The random forest model classifies the species of the penguin given the physical measurements.
+
+```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Species classification of Palmer penguin data.
+a) The primary observation is misclassified as an orange point and is being compared to a nearby correctly classified orange point."}
+knitr::include_graphics("./figures/case_penguins.pdf")
+```
+
+## 2) Chocolates, milk/dark chocolate classification
+
+TODO:XXX
+
+## 3) FIFA, wage regression
+
+The 2020 season FIFA[@leone_fifa_2020, @biecek_dalex_2018], contains many skill measurements of soccer/football players along with wage information.
+After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
+
+```{r casefifa, echo=F, out.width = "100%", fig.cap = "FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
+a) The primary observation is a star offensive player (L.
+Messi) as compared with a top defensive player (V.
+van Dijk).
+b) Starting from a basis where offensive skills ('*' toward the right, high predicted wages) are high we observe that how contingent this variable is for explaining the high wages.
+As the contribution decreases the primary observation moves left into the bulk of the other observations corresponding to lower predicted wages."}
+knitr::include_graphics("./figures/case_fifa.pdf")
+```
+
+The difference in their tree shap values highlights the fact that the random forest model found multiple ways to increase wages; that prediction accuracy was improved by varying linear variable importance that makes sense in light of the player positions that we exogenous to the model.
+
+## 4) Ames Housing prices 2018
+
+TODO:XXX
 
 # Discussion {#sec:discussion}
+
+The need to maintain the interpretability of black-box models is clear.
+One way to do this is by the use of local explanations of an observation.
+Local explanations approximate the linear variable importance to the model.
+Our idea is to visualize approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
+Then use the linear importance from the primary observation as a projection basis and explore a single variable's importance to the structure identified with use of the manual tour.
+
+We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generalized to any compatible model-explanation pair.
+We apply this to both the classification and regression task.
+In the former, it can be used to see which variables cause a misclassification as benchmarked against a nearby observation and explore changing the contribution of variables.
+In the regression case, we compare observations over exogenous .
+An observation with particularly extreme residual can be compared with nearby points more accurately predicted.
+
+We have created an open-source __R__ package __cheem__, available on CRAN, to facilitate this workflow including a dynamic application, with the ability to upload preprocessed data.
 
@@
 
+The namesake, cheem, refers to a fictional race of humanoid trees in the Doctor Who lore.
+Given that __DALEX__ pulls on from that universe and we originally apply tree shap explanations (specific to tree-based models) we found it fitting.
+
 # References

### 2021-11-09 _paper.tex
- finish draft - grammerly edits - need to finish example cases.

@@
 
-Predictive model and black-box modeling is becoming increasingly common, but not without controversy and issues Kodiyan (2019).
+Black-box models are becoming increasingly common, but not without teir share of controversy and issues Kodiyan (2019).
 Applications have been known to reflect common biases against sex Duffy (2019), race (Larson et al.
@@
 2018).
-This is a common issue stemming from biases in the in-sample, training data are violate ethical principles.
+These are all-too common issue, when endogenous or exogenous biases in the training data and identified and mimicked by models.
 Another issue is that of data-drift when new data is outside the support of latent or exogenous explanatory variables.
-Data-drift can lead to worse predictions Salzberg (2014).
+Data drift can lead to worse predictions Salzberg (2014).
 Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI Arrieta et al.
@@
 
-The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is good explanation for the model predictions.
+The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is a good explanation for the model predictions.
 We make R package \texttt{cheem} with an interactive application to facilitate analysis.
@@
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
-We conclude with a \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
+We conclude with a \protect\hyperlink{sec:discussion}{discussion} of the insights we draw from classification and regression tasks.
 
@@
 We show an approximation of these spaces with a projection through their first two principal components.
-The orientation of the variables are shown inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of a higher-dimensional space, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
+The orientation of the variables is shown inscribed on a unit circle.
+While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
 
@@
 
-\caption{Global view screen capture; the approximations of the data and SHAP-spaces of the penguins data.
+\caption{Global view screen capture; the approximations of the data and SHAP-spaces of Palmer penguins data.
 Orientation of the basis contributions is illustrated on a unit circle.
@@
 In the classification case, misclassified points are circled in red.
-This view prpvides an orientation to select a primary and comparison observation, key targets in the following tour.}\label{fig:globalview}
+This view provides an orientation to select a primary and comparison observation, key targets in the following tour.}\label{fig:globalview}
 \end{figure}
@@
 
-\hypertarget{classifcation-task}{%
-\subsection{Classifcation task}\label{classifcation-task}}
+\hypertarget{classification-task}{%
+\subsection{Classification task}\label{classification-task}}
 
-What information do we glean from using this method on a classifcation task?
-Typically we select a misclassified observation in comparison with correctly classified point that is nearby in data space.
+What information do we glean from using this method on a classification task?
+Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
 We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
-By default the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
 That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
@@
 This is similar to the idea of \emph{ceteris paribus} profiles (Biecek 2020).
-\emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant.
-In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.
+\emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, While the variables are orthogonal dimensions, the 2D shadow they cast is not independent.
+That is, when one variable is rotated out of frame, other variables are effectively rotated into frame.
 
@@
 \begin{itemize}
-    \item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic (color/shape) variable can be mapped typically a categorical variable exogenous to the model.
+    \item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic variable (color/shape)  can be mapped typically a categorical variable.
     \item \textbf{Model:} any model can be used with this method.
@@
 
-The above described method and application are implemented as an open source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX site github?
+The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX site github?
 cran?}.
-Preprocessing was facilited with models created via \textbf{randomForest} {[}liaw\_classification\_2002{]}, and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
+Preprocessing was facilitated with models created via \textbf{randomForest} {[}liaw\_classification\_2002{]}, and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
 2021).
@@
 The tour visual is an extension of \textbf{spinifex} (Spyrison and Cook 2020).
-Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive html widget with \textbf{plotly} (Sievert 2020).
+Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
 The free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
@@
 Provided with tidy and transformed data, the function \texttt{cheem\_ls} is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
-It returns an a list containing data frames to be consumed by the application, higher level model performance, and runtime information.
+It returns a list containing data frames to be consumed by the application, higher-level model performance, and runtime information.
 The return of this function can be then be saved to an .rds file for consumption in the application.
 
-\hypertarget{linking-brushing-and-highlighting}{%
-\subsection{Linking brushing and highlighting}\label{linking-brushing-and-highlighting}}
+\hypertarget{interactive-shiny-application}{%
+\subsection{Interactive shiny application}\label{interactive-shiny-application}}
 
@@
 Linked brushing also for a rectangular selection of observations.
-These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance highlighting in the tour.
+These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance, highlighting in the tour.
 The distinction between the former and latter is subtle, but important for developers to grasp.
-In the former, self-contained html widgets can highlight directly within javascript without evaluation evaluation downstream in other reactive functions.
+In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
 In the latter, the use of \texttt{plotly::event\_data()} returns the identity of the selected observations.
@@
 
+To illustrate the use of the cheem method we apply it to modern datasets, two classification examples and then two of regression.
+
 \hypertarget{penguins-species-classification}{%
 \subsection{1) penguins, species classification}\label{penguins-species-classification}}
+
+Palm penguins data Horst, Hill, and Gorman (2020) consist of four physical measurements of 3 species of penguins foraging near Plamer Station, Antarctica.
+The random forest model classifies the species of the penguin given the physical measurements.
 
@@
 
-\caption{Penguins classificiation TODO: XXX}\label{fig:casepenguins}
+\caption{Species classification of Palmer penguin data.
+a) The primary observation is misclassified as an orange point and is being compared to a nearby correctly classified orange point.}\label{fig:casepenguins}
 \end{figure}
@@
 
+The 2020 season FIFABiecek (2018), contains many skill measurements of soccer/football players along with wage information.
+After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
+
 \begin{figure}
@@
 
-\caption{FIFA 2020, regressing wages [log 2020 Euros] from skill measures. TODO: XXX}\label{fig:casefifa}
+\caption{FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
+a) The primary observation is a star offensive player (L.
+Messi) as compared with a top defensive player (V.
+van Dijk).
+b) Starting from a basis where offensive skills ('*' toward the right, high predicted wages) are high we observe that how contingent this variable is for explaining the high wages.
+As the contribution decreases the primary observation moves left into the bulk of the other observations corresponding to lower predicted wages.}\label{fig:casefifa}
 \end{figure}
 
-\hypertarget{coffee-rating-regression}{%
-\subsection{4) Coffee, rating regression}\label{coffee-rating-regression}}
+The difference in their tree shap values highlights the fact that the random forest model found multiple ways to increase wages; that prediction accuracy was improved by varying linear variable importance that makes sense in light of the player positions that we exogenous to the model.
+
+\hypertarget{ames-housing-prices-2018}{%
+\subsection{4) Ames Housing prices 2018}\label{ames-housing-prices-2018}}
 
@@
 
-regress score color on species?
-
 \hypertarget{sec:discussion}{%
 \section{Discussion}\label{sec:discussion}}
+
+The need to maintain the interpretability of black-box models is clear.
+One way to do this is by the use of local explanations of an observation.
+Local explanations approximate the linear variable importance to the model.
+Our idea is to visualize approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
+Then use the linear importance from the primary observation as a projection basis and explore a single variable's importance to the structure identified with use of the manual tour.
+
+We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generalized to any compatible model-explanation pair.
+We apply this to both the classification and regression task.
+In the former, it can be used to see which variables cause a misclassification as benchmarked against a nearby observation and explore changing the contribution of variables.
+In the regression case, we compare observations over exogenous .
+An observation with particularly extreme residual can be compared with nearby points more accurately predicted.
+
+We have created an open-source \textbf{R} package \textbf{cheem}, available on CRAN, to facilitate this workflow including a dynamic application, with the ability to upload preprocessed data.
 
@@
 This research was supported in part by Australian government Research Training Program (RTP) scholarships.
+
+The namesake, cheem, refers to a fictional race of humanoid trees in the Doctor Who lore.
+Given that \textbf{DALEX} pulls on from that universe and we originally apply tree shap explanations (specific to tree-based models) we found it fitting.
 
@@
 
+\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
+Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
+
 \leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
-Biecek, Przemyslaw. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
+---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
@@
 
+\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
+Gorman, Kristen B., Tony D.
+Williams, and William R.
+Fraser.
+2014.
+{``Ecological Sexual Dimorphism and Environmental Variability Within a Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS One} 9 (3): e90081.
+
 \leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
@@
 Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+
+\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
+Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
+Gorman.
+2020.
+{``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
 

### 2021-11-09 figures/_make_figures_case_study.r

@@ -116,38 +116,35 @@ prim_obs <- 15L
 comp_obs <- 282L
 
 ### Global view and cheem tour stills
-glob_view <- cheem::linked_global_view(cheem_ls_peng, prim_obs, comp_obs)
+# glob_view <- cheem::linked_global_view(cheem_ls_peng, prim_obs, comp_obs)
 glob_view <- static_global_view(
-  cheem_ls_peng, prim_obs, comp_obs)
+  cheem_ls_peng, prim_obs, comp_obs) +
+  ggplot2::coord_fixed() + 
+  theme(legend.position = "right",
+        legend.direction = "vertical", ## Levels within aesthetic
+        legend.box = "hrizontal",         ## Between aesthetic
+        legend.margin = ggplot2::margin(1L,1L,1L,1L, "mm")) +
+  labs(color='Predicted class', shape='Predicted class')
 
 bas <- basis_local_attribution(
   cheem_ls_peng$attr_df, rownum = prim_obs)
 ggt <- radial_cheem_ggtour(
   cheem_ls_peng, basis=bas, mv_name=colnames(X)[1],
   primary_obs=prim_obs, comparison_obs=comp_obs,
-  pcp_shape = 124, angle = 5)
+  do_add_pcp_segments = TRUE,
+  pcp_shape = 124, angle = 5) + theme(legend.position = "off")
 cheem_stills <- spinifex::filmstrip(ggt)
 
 ### Save
 cp <- cowplot::plot_grid(
-  glob_view + ggtitle("Global view"), 
+  glob_view + ggtitle("Global view"),
   cheem_stills + ggtitle("Cheem tour, select frames"),
   labels = paste0(letters[1:2], ")"),
   ncol = 1)#, rel_heights = c(1.5, 1))
 ggplot2::ggsave(
   "./figures/case_penguins.pdf",
   plot = cp, device = "pdf",
-  width = 8, height = 4.8, units = "in")
-
-# ## Save interactive html widget
-# ggp <- ggplotly(p, tooltip = "tooltip") %>%
-#   config(displayModeBar = FALSE) %>% ## Remove html buttons
-#   layout(dragmode = "select", showlegend = FALSE,
-#          width = 640, height = 320) %>% ## Set drag left mouse
-#   event_register("plotly_selected") %>% ## Reflect "selected", on release of the mouse button.
[…]

### 2021-11-15 figures/_make_figures_case_study.r
adj case studies, update stat_global view.

@@ -1,98 +1,142 @@
-# CASE STUDIES FOR CHEEM ----
-require("cheem")
-require("spinifex")
-require("dplyr")
-require("ggplot2")
-require("cowplot")
-
-## Local function variants:
-
-static_global_view <- function(
-  cheem_ls,
-  primary_obs = NULL,
-  comparison_obs = NULL
-){
-  ## Prevent global variable warnings:
-  V1 <- V2 <- ggtext <- projection_nm <- layer_name <- tooltip <- NULL
-  .alpha <- logistic_tform(nrow(cheem_ls$decode_df), mid_pt = 500L)
-  global_view_df <- cheem_ls$global_view_df ## Init
-  is_classification <- cheem_ls$problem_type == "classification"
-  pred_clas <- as.factor(FALSE) ## If regression; dummy pred_clas
-  if(is_classification == TRUE) pred_clas <-
-    cheem_ls$decode_df$predicted_class %>%
-    rep_len(nrow(global_view_df)) %>%
-    as.factor()
+#### CASE STUDIES FOR CHEEM ####
+# Setup ----
+{
+  require("cheem")
+  require("spinifex")
+  require("dplyr")
+  require("ggplot2")
+  require("cowplot")
   
-  pts_highlight <- list()
-  ## Red misclassified points, if present
-  if(is_classification == TRUE){
-    .rn_misclass <- which(cheem_ls$decode_df$is_misclassified == TRUE)
-    .idx_misclas <- global_view_df$rownum %in% .rn_misclass
-    if(sum(.idx_misclas) > 0L){
-      .df <- global_view_df[.idx_misclas, ] # %>% highlight_key(~rownum)
-      pts_highlight <- c(
-        pts_highlight,
-        ggplot2::geom_point(ggplot2::aes(V1, V2), .df,
-                            color = "red", fill = NA,
-                            shape = 21L, size = 3L, alpha = .alpha)
-      )
-    }
-  }
-  ## Highlight comparison obs, if passed
-  if(is.null(comparison_obs) == FALSE){
-    .idx_comp <- global_view_df$rownum == comparison_obs
-    if(sum(.idx_comp) > 0L){
-      .df <- global_view_df[.idx_comp, ]
-      pts_highlight <- c(
-        pts_highlight,
-        ## Highlight comparison obs
-        ggplot2::geom_point(ggplot2::aes(V1, V2), #, color = pred_clas[.idx_comp]),
[…]

### 2021-11-16 _paper.rmd
-rework section 2, local explanations -word smithing and header comments. -fix center alignment.

@@
 ---
-title: "Methods for understanding the variable importance of local explanations of black-box models"
+title: "Interogating the linear variable importance of local explanations of black-box models with animated linear projections"
 # author:
@@
 <!-- ## XAI & interpretability crisis -->
-Black-box models are becoming increasingly common, but not without teir share of controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
+Black-box models are becoming increasingly common, but not without their share of controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
 Applications have been known to reflect common biases against sex [@dastin_amazon_2018, @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
@@
 <!-- SHAP -->
-However, to illustrate our work we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010, @strumbelj_explaining_2014].
+To illustrate our work we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010, @strumbelj_explaining_2014].
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
@@
 The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
-<!-- ## manual tours --> There are various types of tours, which are distinguished by the generation of the sequence of projection bases.
+
+<!-- ## manual tours --> 
+There are various types of tours, which are distinguished by the sequence of projection bases.
 In a _manual_ tour [@cook_manual_1997, @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
@@
 
+<!-- Purposed approach -->
 The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is a good explanation for the model predictions.
-We make R package `cheem` with an interactive application to facilitate analysis.
+We provide a free and open-source R package `cheem` with an interactive application to facilitate analysis.
 By viewing approximations of data- and attribution-space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
@@
 The rest of this paper is organized as follows.
-The next section [SHAP](#sec:SHAP) covers the background of the local explanation SHAP and the traditional visuals produced from it.
-The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis.
-Following that, [Software Instructure](#sec:softwareinfrastructure) discusses the backend details of the package and preprocessing.
+The next section [Local explanation statistics](#sec:explanations) covers the background of the local explanation SHAP and the traditional visuals produced from it.
+[Tours and the radial tour](#sec:tour) digs deeper into the animation and what it shows.
+The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis and  discusses the backend details of the package and preprocessing.
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
@@
 
-# SHAP local explanation {#sec:SHAP}
+# Local explanation statistics {#sec:explanations}
+
+<!-- reminder of local explanation -->
+Consider a highly non-linear model.
+At face value its hard to say which variable are sensitive to the crossing of classification boundary or identify which variables caused an observation to have a relatively extreme residual.
+Local explanations shed light on these cases by approximating linear variable importance at in vicinity of one observation.
+
+<!-- Taxonomy of local explanations -->
+Figure 6 of @arrieta_explainable_2020 gives a broad summarization of the taxonomy and literature of explanation techniques.
+This includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016, @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
+There are a fewer number of model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017] and their variants are popular.
+
+<!-- Uses of local explanations -->
+These local explanation are used in a variety of ways depending on the data.
+For images saliency maps [@simonyan_deep_2014] a heatmap can be used to show which pixels were important for distinguishing pictures of wolfs or huskies for instance.
+In text analysis saliency can be used to highlight influential words [@vanni_textual_2018].
+In the case of numeric regression they can be used to explain variable additive contributions from intercept to prediction [@ribeiro_why_2016].
+
+## SHAP and tree SHAP
 
 <!-- SHAP and history -->
-SHaply Additive exPlanations, or SHAP [@lundberg_unified_2017] approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+SHaply Additive exPlanations, or SHAP approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
 This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players that contribute to the score [@shapley_value_1953].
 
-<!-- An observation's SHAP values were originally used to additvely explain the difference from the intercept to the prediction.
-This sort of explanation is predicated on the contribution of the previous variables, making it asymmetric across variable ordering.
--->
-
-TO illustrate SHAP and its original use we use soccer data from FIFA 2020 season [@leone_fifa_2020].
-We have 5000 observations of 9 aggregated skill measures and use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
-We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
+To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season [@leone_fifa_2020].
+We have 5000 observations of 9 skill measures (after aggregating variable with high correlation).
+A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
+We then extract the SHAP values of a star offensive player (L.
+Messi) and defensive player (V.
+van Dijk).
 We expect to see a difference in the attribution of the variable importance across the two positions of the players.
@@
 <!-- Illustration -->
-Figure \@ref(fig:shapdistrbd) illustrates the SHAP values of these players.
+Figure \@ref(fig:shapdistrbd) shows the SHAP values of these players.
+Panel a) shows these players recieve a sizable difference in wages.
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
@@
 
-In summary, this highlights how local explanations bring transparency to a model at least in the vicinity of their observations.
+In summary, this highlights how local explanations bring interpretability to a model at least in the vicinity of their observations.
 In this instance, we showed how two very different soccer players receive different profiles of variable importance to explain the prediction of their wages.
@@
 
-# Application Design {#sec:applicationdesign}
+# Tours and the radial tour
+
+# Application design {#sec:applicationdesign}
 
@@
 
+<!-- purpose -->
 The global view is an important context for exploring the separability of the data- and the local explanation's attribution-spaces, and is crucial in the selection of explanation to further interrogate and explore the structural sensitivity of.
 
+<!-- approximations of the spaces, PC1:2 -->
 We show an approximation of these spaces with a projection through their first two principal components.
@@
 
+<!-- Interaction -->
 This view offers dynamic interaction in several ways.
@@
 
+<!-- a function of the obs selected above -->
 The primary observation identified via the global view is foundational to the production of the cheem tour.
@@
 
-That normalized attribution of the primary observation is depicted as stacked bars where the horizontal width is the contribution.
-The bottom of this display is divided by the use case.
+<!-- Dipiction of a frame -->
+The SHAP values of the full dataset are shown as horizontal parrallel coordiate plot at the top with the selected observations highlighted.
 In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
-In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of the observations is fixed to its prediction or residual.
+In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of is fixed to the observations prediction and residual respectively.
+The current projection basis is depicted on the bottom, where the width of the bar is mapped to variables contribution to the horizontal axis.
+The starting frame is the normalized SHAP values of the primary observation.
 
-```{r cheemtour, echo=F, out.width = "100%", fig.cap = "Cheem tour screen capture; the primary observation's normalized SHAP values are the initial basis.
-This is the explanation of the linear variable importances for the model at this observation.
-The point in question was misclassified, it was predicted to be from the green cluster, while it was observed to be from the orange cluster.
-The top shows the contributions of the variables, with the dashed line being the primary observation's SHAP, its shape is closer to the orange cluster (observed group) while the position is pulled more toward the green (predicted group).
-The bottom shows the 1D projection with density and rug marks below.
-The dashed line is plausibly in the middle of the green density, the story that the explanation trying to sell us.
-Yet, when we play the tour animating on the contribution of bill length (b_l), the bottom dashed line is more regularly in the center of the observed orange cluster."}
+<!-- cheem tour caption -->
+(ref:cheemtour) Cheem tour screen capture; on classification random forest of simulated data, a misclassified observation is be compared with a nearby correctly classified observation.
+The top contains a horizontal display of the SHAP value from all observation with selected primary and comparison observations highlighted.
+The middle shows density and rug marks of the current frame with selections similarly highlighted.
+The bottom shows the projection basis, that is, each variables contribution to the horizontal position.
+The primary, misclassifed observation (dashed line), is plausibly in the middle of the green density, the story that the explanation trying to sell us.
+Yet, when we play the tour animating on the contribution of bill length (b_l), the bottom dashed line is more regularly in the center the it observed cluster, orange.
+By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.
+
+```{r cheemtour, echo=F, out.width = "100%", fig.cap = "(ref:cheemtour)"}
 knitr::include_graphics("./figures/cheem_tour_penguins.PNG")
@@
 
-# Software Infrastructure {#sec:softwareinfrastructure}
+## Backend infrastructure {#sec:infrastructure}
 
@@
 Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive HTML widgets with __plotly__ [@sievert_interactive_2020].
-The free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
+__DALEX__ [@biecek_dalex_2018] and
+the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
 
@@
 
-# Case Studies {#sec:casestudies}
+# Case studies {#sec:casestudies}
 

### 2021-11-16 _paper.tex
-rework section 2, local explanations -word smithing and header comments. -fix center alignment.

@@
 \hypersetup{
-  pdftitle={Methods for understanding the variable importance of local explanations of black-box models},
+  pdftitle={Interogating the linear variable importance of local explanations of black-box models with animated linear projections},
   colorlinks=true,
@@
 
-\title{Methods for understanding the variable importance of local explanations of black-box models}
+\title{Interogating the linear variable importance of local explanations of black-box models with animated linear projections}
 \author{}
@@
 
-Black-box models are becoming increasingly common, but not without teir share of controversy and issues Kodiyan (2019).
+Black-box models are becoming increasingly common, but not without their share of controversy and issues Kodiyan (2019).
 Applications have been known to reflect common biases against sex Duffy (2019), race (Larson et al.
@@
 
-However, to illustrate our work we apply the model-agnostic explanation SHAP Štrumbelj and Kononenko (2014).
+To illustrate our work we apply the model-agnostic explanation SHAP Štrumbelj and Kononenko (2014).
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
@@
 The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
-There are various types of tours, which are distinguished by the generation of the sequence of projection bases.
+
+There are various types of tours, which are distinguished by the sequence of projection bases.
 In a \emph{manual} tour Spyrison and Cook (2020) this path is defined by changing the contribution of a selected variable.
@@
 The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is a good explanation for the model predictions.
-We make R package \texttt{cheem} with an interactive application to facilitate analysis.
+We provide a free and open-source R package \texttt{cheem} with an interactive application to facilitate analysis.
 By viewing approximations of data- and attribution-space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
@@
 The rest of this paper is organized as follows.
-The next section \protect\hyperlink{sec:SHAP}{SHAP} covers the background of the local explanation SHAP and the traditional visuals produced from it.
-The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis.
-Following that, \protect\hyperlink{sec:softwareinfrastructure}{Software Instructure} discusses the backend details of the package and preprocessing.
+The next section \protect\hyperlink{sec:explanations}{Local explanation statistics} covers the background of the local explanation SHAP and the traditional visuals produced from it.
+\protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into the animation and what it shows.
+The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis and discusses the backend details of the package and preprocessing.
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
@@
 
-\hypertarget{sec:SHAP}{%
-\section{SHAP local explanation}\label{sec:SHAP}}
+\hypertarget{sec:explanations}{%
+\section{Local explanation statistics}\label{sec:explanations}}
 
-SHaply Additive exPlanations, or SHAP (S.
-Lundberg and Lee 2017) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+Consider a highly non-linear model.
+At face value its hard to say which variable are sensitive to the crossing of classification boundary or identify which variables caused an observation to have a relatively extreme residual.
+Local explanations shed light on these cases by approximating linear variable importance at in vicinity of one observation.
+
+Figure 6 of Arrieta et al.
+(2020) gives a broad summarization of the taxonomy and literature of explanation techniques.
+This includes a large number of model-specific explanations such as deepLIFT, Shrikumar, Greenside, and Kundaje (2017) a popular recursive method for estimating importance in neural networks.
+There are a fewer number of model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (S.
+Lundberg and Lee 2017) and their variants are popular.
+
+These local explanation are used in a variety of ways depending on the data.
+For images saliency maps (Simonyan, Vedaldi, and Zisserman 2014) a heatmap can be used to show which pixels were important for distinguishing pictures of wolfs or huskies for instance.
+In text analysis saliency can be used to highlight influential words (Vanni et al.
+2018).
+In the case of numeric regression they can be used to explain variable additive contributions from intercept to prediction (Ribeiro, Singh, and Guestrin 2016).
+
+\hypertarget{shap-and-tree-shap}{%
+\subsection{SHAP and tree SHAP}\label{shap-and-tree-shap}}
+
+SHaply Additive exPlanations, or SHAP approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
 This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players that contribute to the score (Shapley 1953).
 
-TO illustrate SHAP and its original use we use soccer data from FIFA 2020 season (Leone 2020).
-We have 5000 observations of 9 aggregated skill measures and use a random forest model to regress the wages, in 2020 Euros, from the skill measures.
-We then extract the SHAP values of a star offensive player (Messi) and defensive player (van Dijk).
+To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season (Leone 2020).
+We have 5000 observations of 9 skill measures (after aggregating variable with high correlation).
+A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
+We then extract the SHAP values of a star offensive player (L.
+Messi) and defensive player (V.
+van Dijk).
 We expect to see a difference in the attribution of the variable importance across the two positions of the players.
 
-Figure \ref{fig:shapdistrbd} illustrates the SHAP values of these players.
+Figure \ref{fig:shapdistrbd} shows the SHAP values of these players.
+Panel a) shows these players recieve a sizable difference in wages.
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
@@
 
+\hypertarget{tours-and-the-radial-tour}{%
+\section{Tours and the radial tour}\label{tours-and-the-radial-tour}}
+
 \hypertarget{sec:applicationdesign}{%
-\section{Application Design}\label{sec:applicationdesign}}
+\section{Application design}\label{sec:applicationdesign}}
 
@@
 
-That normalized attribution of the primary observation is depicted as stacked bars where the horizontal width is the contribution.
-The bottom of this display is divided by the use case.
+The SHAP values of the full dataset are shown as horizontal parrallel coordiate plot at the top with the selected observations highlighted.
 In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
-In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of the observations is fixed to its prediction or residual.
+In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of is fixed to the observations prediction and residual respectively.
+The current projection basis is depicted on the bottom, where the width of the bar is mapped to variables contribution to the horizontal axis.
+The starting frame is the normalized SHAP values of the primary observation.
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_penguins}
+{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_penguins}
 
-\}
+}
 
-\textbackslash caption\{Cheem tour screen capture; the primary observation's normalized SHAP values are the initial basis.
-This is the explanation of the linear variable importances for the model at this observation.
-The point in question was misclassified, it was predicted to be from the green cluster, while it was observed to be from the orange cluster.
-The top shows the contributions of the variables, with the dashed line being the primary observation's SHAP, its shape is closer to the orange cluster (observed group) while the position is pulled more toward the green (predicted group).
-The bottom shows the 1D projection with density and rug marks below.
-The dashed line is plausibly in the middle of the green density, the story that the explanation trying to sell us.
-Yet, when we play the tour animating on the contribution of bill length (b\_l), the bottom dashed line is more regularly in the center of the observed orange cluster.\}\label{fig:cheemtour}
-\textbackslash end\{figure\}
+\caption{Cheem tour screen capture; on classification random forest of simulated data, a misclassified observation is be compared with a nearby correctly classified observation.
+The top contains a horizontal display of the SHAP value from all observation with selected primary and comparison observations highlighted.
+The middle shows density and rug marks of the current frame with selections similarly highlighted.
+The bottom shows the projection basis, that is, each variables contribution to the horizontal position.
+The primary, misclassifed observation (dashed line), is plausibly in the middle of the green density, the story that the explanation trying to sell us.
+Yet, when we play the tour animating on the contribution of bill length (b\_l), the bottom dashed line is more regularly in the center the it observed cluster, orange.
+By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.}\label{fig:cheemtour}
+\end{figure}
 
@@
 
-\hypertarget{sec:softwareinfrastructure}{%
-\section{Software Infrastructure}\label{sec:softwareinfrastructure}}
+\hypertarget{sec:infrastructure}{%
+\subsection{Backend infrastructure}\label{sec:infrastructure}}
 
@@
 Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
-The free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
+\textbf{DALEX} (Biecek 2018) and
+the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
 
@@
 \hypertarget{sec:casestudies}{%
-\section{Case Studies}\label{sec:casestudies}}
+\section{Case studies}\label{sec:casestudies}}
 
@@
 
+\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
+Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
+2016.
+{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} \emph{arXiv:1602.04938 {[}Cs, Stat{]}}, February.
+\url{http://arxiv.org/abs/1602.04938}.
+
 \leavevmode\hypertarget{ref-salzberg_why_2014}{}%
@@
 
+\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
+Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
+2017.
+{``Learning Important Features Through Propagating Activation Differences.''} In \emph{International {Conference} on {Machine} {Learning}}, 3145--53.
+PMLR.
+
+\leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
+Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
+2016.
+{``Not Just a Black Box: {Learning} Important Features Through Propagating Activation Differences.''} \emph{arXiv Preprint arXiv:1605.01713}.
+
 \leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
@@
 {``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
+
+\leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
+Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
+2014.
+{``Deep Inside Convolutional Networks: {Visualising} Image Classification Models and Saliency Maps.''} In \emph{In {Workshop} at {International} {Conference} on {Learning} {Representations}}.
+Citeseer.
 
@@
 
+\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
+Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
+2018.
+{``Textual {Deconvolution} {Saliency} ({TDS}): A Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the 56th {Annual} {Meeting} of the {Association} for {Computational} {Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
+
 \leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%

### 2021-11-21 _paper.rmd
Application design and recent screen caps.

@@
 
+<!-- What-if & counterfactual analysis -->
+The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _Ceteris paribus_[@biecek_ceterisparibus_2020].
+Latin for “other things held constant” or “all else unchanged”, is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+It ignores correlations of the variables and imagines a case that was not observed.
+In contrast, our approach is a geometric explanation a of the factual observed case by varying the basis, the angle of the data object.
+Another, but related difference is that this geometric approach maintains orthogonal dimensions.
+That is to say when the contribution of some variables decrease the the contributions of others necessarily increases.
+
 The rest of this paper is organized as follows.
@@
 
+```{r}
+##TODO:XXX
+```
+
 # Application design {#sec:applicationdesign}
@@
 Lastly, we discuss the preprocessing that needs to be done before display.
+
+## Classification task
+
+What information do we glean from using this method on a classification task?
+Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
+We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
+That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+
+## Regression task
+
+The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
 
@@
 This view provides an orientation to select a primary and comparison observation, key targets in the following tour."}
-knitr::include_graphics("./figures/global_view_penguins.PNG")
+knitr::include_graphics("./figures/global_view_toy_class.PNG")
 ```
@@
 The starting frame is the normalized SHAP values of the primary observation.
+In figure \@ref(fig:cheemtourclass) we illustrate the classification case with simulated data; 3 clusters separated in the first 2 dimensions with 2 dimensions of noise.
 
-<!-- cheem tour caption -->
-(ref:cheemtour) Cheem tour screen capture; on classification random forest of simulated data, a misclassified observation is be compared with a nearby correctly classified observation.
-The top contains a horizontal display of the SHAP value from all observation with selected primary and comparison observations highlighted.
-The middle shows density and rug marks of the current frame with selections similarly highlighted.
-The bottom shows the projection basis, that is, each variables contribution to the horizontal position.
-The primary, misclassifed observation (dashed line), is plausibly in the middle of the green density, the story that the explanation trying to sell us.
-Yet, when we play the tour animating on the contribution of bill length (b_l), the bottom dashed line is more regularly in the center the it observed cluster, orange.
-By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.
+<!-- cheem tour classification caption -->
+(ref:cheemtourclass) Cheem tour, classification case.
+A random forest model predicting the class of the of simulated data.
+A misclassified observation is be compared with a nearby correctly classified observation.
+The top shows density and rug marks of the current frame with the positions of the primary and comparison observations shown as dashed and dotted lines.
+The bottom shows two key features; the distribution of attribution space shown as parallel coordinate lines across the variables and the bars show the current contribution of the variable to the projection.
+The primary, misclassifed observation (dashed line), is plausibly in the middle of the its prediction,  purple density, the story that the explanation trying to sell us.
+Yet, when we play the tour animating on the contribution of V1, the bottom dashed line is more regularly in the center the it observed cluster, green, By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.
 
-```{r cheemtour, echo=F, out.width = "100%", fig.cap = "(ref:cheemtour)"}
-knitr::include_graphics("./figures/cheem_tour_penguins.PNG")
+```{r cheemtourclass, echo=F, out.width = "100%", fig.cap = "(ref:cheemtourclass)"}
+knitr::include_graphics("./figures/cheem_tour_toy_class.PNG")
 ```
 
+<!-- tour animation -->
 Data visualization tours animate many linear projections over small changes to the basis.
@@
 
-## Classification task
+<!-- regression case -->
+In the regression case we compare an observation with a particularly extreme residual as compared with a neighboring point that is more  accurately predicted.
+The global tour is primarily the same without the misclassified circles, but the tooltip now shows the observation's residual.
+We simulate data, 5 variable from the uniform distribution (between 0 and 5) and $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
+We fit a random forest and extract the all tree SHAP values.
+The first thing we notice from the global view is the shape difference between the data- and attribution-spaces, which are primarily circular are triangular respectively.
+We also see that the random forest fit the data quite well viewing the predictions across the observations.
+Though linked brushing we find that the attribution space is highly correlated with the $y$ value.
+The highest values occupy 1 corner of the triangle, the lowest values cluster on the opposite side and the middle values form a ')' shape filling out the triangle in attribution space.
+In the regression case the tour view is a bit different, this is illustrated in figure \@ref(cheemtourreg).
+The the current frames basis and distribution are shown to the left.
+The horizontal positions of the two scatter plots are the resulting 1D projection of the basis while the heights are fixed to the observed y and the residuals.
 
-What information do we glean from using this method on a classification task?
-Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
-We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
-That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+<!-- tour regression caption -->
+(ref:cheemtourreg) Cheem tour, regression case.
+A random forest model predicting the continuous $y$ of the of simulated data where $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
+The current frames basis and the distribution of all attributions is shown on the left.
+The horizontal positions for the scatterplots are resulting projection through the 1D basis while the vertical heights are fixed to the observed y and residuals respectively.
 
-Another way of thinking about this would be: what would the predictions be if the explanation were different.
-This is similar to the idea of _ceteris paribus_ profiles [@biecek_ceterisparibus_2020].
-_Ceteris paribus_ is Latin for “other things held constant” or “all else unchanged”.
-The profiles visualize what 'what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
-In contrast to _ceteris paribus_ profiles, touring methods visualize all observations rather than one, While the variables are orthogonal dimensions, the 2D shadow they cast is not independent.
-That is, when one variable is rotated out of frame, other variables are effectively rotated into frame.
+```{r cheemtourreg, echo=F, out.width = "100%", fig.cap = "(ref:cheemtourreg)"}
+knitr::include_graphics("./figures/cheem_tour_toy_reg.PNG")
+```
 
-## Regression task
-
-The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
+Now that we have covered the classification and regression cases we will discuss the preprocessing, package infrastructure, and interactive features before proceeding to case studies with data from the wild.
 
@@
 
-## Backend infrastructure {#sec:infrastructure}
+## Package infrastructure {#sec:infrastructure}
 
@@
 
-## Preprocessing
-
-Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
+<!-- Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
 It returns a list containing data frames to be consumed by the application, higher-level model performance, and runtime information.
 The return of this function can be then be saved to an .rds file for consumption in the application.
+-->
 
-## Interactive shiny application
+## Interactive features
 

### 2021-11-21 _paper.tex
Application design and recent screen caps.

@@
 
+The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{Ceteris paribus}(Biecek 2020).
+Latin for ``other things held constant'' or ``all else unchanged,'' is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+It ignores correlations of the variables and imagines a case that was not observed.
+In contrast, our approach is a geometric explanation a of the factual observed case by varying the basis, the angle of the data object.
+Another, but related difference is that this geometric approach maintains orthogonal dimensions.
+That is to say when the contribution of some variables decrease the the contributions of others necessarily increases.
+
 The rest of this paper is organized as follows.
@@
 
+\hypertarget{classification-task}{%
+\subsection{Classification task}\label{classification-task}}
+
+What information do we glean from using this method on a classification task?
+Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
+We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
+That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+
+\hypertarget{regression-task}{%
+\subsection{Regression task}\label{regression-task}}
+
+The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
+
 \hypertarget{global-view}{%
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/global_view_penguins}
+{\centering \includegraphics[width=1\linewidth]{./figures/global_view_toy_class}
 
@@
 The starting frame is the normalized SHAP values of the primary observation.
+In figure \ref{fig:cheemtourclass} we illustrate the classification case with simulated data; 3 clusters separated in the first 2 dimensions with 2 dimensions of noise.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_penguins}
+{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_toy_class}
 
@@
 
-\caption{Cheem tour screen capture; on classification random forest of simulated data, a misclassified observation is be compared with a nearby correctly classified observation.
-The top contains a horizontal display of the SHAP value from all observation with selected primary and comparison observations highlighted.
-The middle shows density and rug marks of the current frame with selections similarly highlighted.
-The bottom shows the projection basis, that is, each variables contribution to the horizontal position.
-The primary, misclassifed observation (dashed line), is plausibly in the middle of the green density, the story that the explanation trying to sell us.
-Yet, when we play the tour animating on the contribution of bill length (b\_l), the bottom dashed line is more regularly in the center the it observed cluster, orange.
-By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.}\label{fig:cheemtour}
+\caption{Cheem tour, classification case.
+A random forest model predicting the class of the of simulated data.
+A misclassified observation is be compared with a nearby correctly classified observation.
+The top shows density and rug marks of the current frame with the positions of the primary and comparison observations shown as dashed and dotted lines.
+The bottom shows two key features; the distribution of attribution space shown as parallel coordinate lines across the variables and the bars show the current contribution of the variable to the projection.
+The primary, misclassifed observation (dashed line), is plausibly in the middle of the its prediction, purple density, the story that the explanation trying to sell us.
+Yet, when we play the tour animating on the contribution of V1, the bottom dashed line is more regularly in the center the it observed cluster, green, By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.}\label{fig:cheemtourclass}
 \end{figure}
@@
 
-\hypertarget{classification-task}{%
-\subsection{Classification task}\label{classification-task}}
+In the regression case we compare an observation with a particularly extreme residual as compared with a neighboring point that is more accurately predicted.
+The global tour is primarily the same without the misclassified circles, but the tooltip now shows the observation's residual.
+We simulate data, 5 variable from the uniform distribution (between 0 and 5) and \(y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon\).
+We fit a random forest and extract the all tree SHAP values.
+The first thing we notice from the global view is the shape difference between the data- and attribution-spaces, which are primarily circular are triangular respectively.
+We also see that the random forest fit the data quite well viewing the predictions across the observations.
+Though linked brushing we find that the attribution space is highly correlated with the \(y\) value.
+The highest values occupy 1 corner of the triangle, the lowest values cluster on the opposite side and the middle values form a `)' shape filling out the triangle in attribution space.
+In the regression case the tour view is a bit different, this is illustrated in figure \ref{cheemtourreg}.
+The the current frames basis and distribution are shown to the left.
+The horizontal positions of the two scatter plots are the resulting 1D projection of the basis while the heights are fixed to the observed y and the residuals.
 
-What information do we glean from using this method on a classification task?
-Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
-We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
-That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+\begin{figure}
 
-Another way of thinking about this would be: what would the predictions be if the explanation were different.
-This is similar to the idea of \emph{ceteris paribus} profiles (Biecek 2020).
-\emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
-In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, While the variables are orthogonal dimensions, the 2D shadow they cast is not independent.
-That is, when one variable is rotated out of frame, other variables are effectively rotated into frame.
+{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_toy_reg}
 
-\hypertarget{regression-task}{%
-\subsection{Regression task}\label{regression-task}}
+}
 
-The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
+\caption{Cheem tour, regression case.
+A random forest model predicting the continuous \(y\) of the of simulated data where \(y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon\).
+The current frames basis and the distribution of all attributions is shown on the left.
+The horizontal positions for the scatterplots are resulting projection through the 1D basis while the vertical heights are fixed to the observed y and residuals respectively.}\label{fig:cheemtourreg}
+\end{figure}
+
+Now that we have covered the classification and regression cases we will discuss the preprocessing, package infrastructure, and interactive features before proceeding to case studies with data from the wild.
 
@@
 \hypertarget{sec:infrastructure}{%
-\subsection{Backend infrastructure}\label{sec:infrastructure}}
+\subsection{Package infrastructure}\label{sec:infrastructure}}
 
@@
 
-\hypertarget{preprocessing-1}{%
-\subsection{Preprocessing}\label{preprocessing-1}}
-
-Provided with tidy and transformed data, the function \texttt{cheem\_ls} is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
-It returns a list containing data frames to be consumed by the application, higher-level model performance, and runtime information.
-The return of this function can be then be saved to an .rds file for consumption in the application.
-
-\hypertarget{interactive-shiny-application}{%
-\subsection{Interactive shiny application}\label{interactive-shiny-application}}
+\hypertarget{interactive-features}{%
+\subsection{Interactive features}\label{interactive-features}}
 

### 2021-11-22 _paper.rmd
Chocolates data figure

@@
 ---
-title: "Interogating the linear variable importance of local explanations of black-box models with animated linear projections"
+title: "Interrogating the linear variable importance of local explanations of black-box models with animated linear projections"
 # author:
@@
 Palm penguins data [@gorman_ecological_2014, @horst_palmerpenguins_2020] consist of four physical measurements of 3 species of penguins foraging near Plamer Station, Antarctica.
-The random forest model classifies the species of the penguin given the physical measurements.
+A random forest model was fit, classifying the species of the penguin given the physical measurements.
 

### 2021-11-22 figures/_make_figures_case_study.r

@@ -1,5 +1,5 @@
-#### CASE STUDIES FOR CHEEM ####
-# Setup ----
+# CASE STUDIES FOR CHEEM ----
+## Setup ----
 {
   require("cheem")
   require("spinifex")
@@ -7,237 +7,199 @@
   require("ggplot2")
   require("cowplot")
   
-  ## Local function variants:
+  setwd("~/R/cheem_paper")
+  fp <- "../cheem/inst/shiny_apps/cheem_initial/data/"
+  if(F)
+    dir("../cheem/inst/shiny_apps/cheem_initial/data/")
   
-  static_global_view <- function(
-    cheem_ls,
-    primary_obs = NULL,
-    comparison_obs = NULL,
-    color = NULL,
-    shape = NULL, 
-    height_px = 480L,
-    width_px = 960L
-  ){
-    ## Prevent global variable warnings:
-    V1 <- V2 <- ggtext <- projection_nm <- layer_name <- tooltip <- NULL
-    ## Initialize
-    global_view_df <- cheem_ls$global_view_df 
-    is_classification <- cheem_ls$type == "classification"
-    ## Aesthetics
-    .alpha <- logistic_tform(nrow(cheem_ls$decode_df))
-    ## setup shape and color
-    if(cheem_ls$type == "classification"){
-      if(is.null(color)) color <- cheem_ls$decode_df$predicted_class %>% as.factor()
-      if(is.null(shape)) shape <- cheem_ls$decode_df$predicted_class %>% as.factor()
-    }else{
-      ## Regression or unexpected type
-      if(length(unique(cheem_ls$decode_df$class)) > 1L){
-        ## Class defined
-        if(is.null(color)) color <- cheem_ls$decode_df$class %>% as.factor()
-        if(is.null(shape)) shape <- cheem_ls$decode_df$class %>% as.factor()
-      }else{
-        ## Class not defined
-        if(is.null(color)) color <- cheem_ls$decode_df$residual
-        if(is.null(shape)) shape <- as.factor(FALSE)
-      }
-    }
-    global_view_df$color <- color %>% rep_len(nrow(global_view_df))
-    global_view_df$shape <- shape %>% rep_len(nrow(global_view_df))
-    
-    ## Get the bases of the global view, map them
-    u_nms <- unique(global_view_df$layer_name)
[…]

### 2021-11-23 _paper.rmd
case study changes; figures will need updating tomorrow...

@@
 
-Palm penguins data [@gorman_ecological_2014, @horst_palmerpenguins_2020] consist of four physical measurements of 3 species of penguins foraging near Plamer Station, Antarctica.
+Palmer penguins data [@gorman_ecological_2014, @horst_palmerpenguins_2020] consist of 330 observations across four physical measurements of 3 species of penguins foraging near Palmer Station, Antarctica.
 A random forest model was fit, classifying the species of the penguin given the physical measurements.
 
-```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Species classification of Palmer penguin data.
-a) The primary observation is misclassified as an orange point and is being compared to a nearby correctly classified orange point."}
+```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Species classification of Palmer penguin data."}
 knitr::include_graphics("./figures/case_penguins.pdf")
@@
 
+In figure \@ref(fig:casepenguins) a misclassified point is contrasted with a correctly classified point of its observated class that is nearby in data-space.
+The attribution space from the treeshap local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution while primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
+The tour varies the contribution of of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
+Downplaying the contribution of bill length is crutial to linear explanation this observation be misclassified.
+
 ## 2) Chocolates, milk/dark chocolate classification
 
-TODO:XXX
+The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
+Each of which was labeled as being either milk or dark chocolates.
+With this data we can see if a manufacturer is giving an accurate portrayal of the chocolate.
+We are curious to see if there are chocolates that nutritionally look like milk chocolates that are labeled as dark chocolates, which may hold a higher market value.
+We should note that not all chocolates consist wholly of chocolate.
+The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable.
+A random forest model is fit classifying the type of chocolate.
+We'll select chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85% cocoa.
+
+```{r casechocolates, echo=F, out.width = "100%", fig.cap = "Chocolates data  type classification (milk or dark). "}
+knitr::include_graphics("./figures/case_chocolates.pdf")
+```
+
+From figure \@ref(fig:casechocolates) we similarly see that attribution space that is more separable relative to data-space.
+Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while there 8 of the inverse case.
+Calories from fat is the measure with the largest difference in treeshap attribution between these points.
 
@@
 After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
+The model was fit with 5000 observation of the 9 skill aggregates before being thinned to 500 player to gate occlusion and render time.
+We compare a leading offensive fielder (L.
+Messi) with that of a leading defensive fielder (V.
+van Dijk).
 
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
-a) The primary observation is a star offensive player (L.
+The primary observation is star offensive player (L.
 Messi) as compared with a top defensive player (V.
-van Dijk).
-b) Starting from a basis where offensive skills ('*' toward the right, high predicted wages) are high we observe that how contingent this variable is for explaining the high wages.
-As the contribution decreases the primary observation moves left into the bulk of the other observations corresponding to lower predicted wages."}
+van Dijk)."}
 knitr::include_graphics("./figures/case_fifa.pdf")
@@
 
-The difference in their tree shap values highlights the fact that the random forest model found multiple ways to increase wages; that prediction accuracy was improved by varying linear variable importance that makes sense in light of the player positions that we exogenous to the model.
+The difference in their tree SHAP values highlights the fact that the random forest model found multiple ways to predict wages in a way that makes sense; especially in the disparity of the importance of offensive and defensive skill measures while player position was exogenous to the model.
+Reaction skills were highly important in both cases.
+After removing reaction, we vary the contribution of offensive skills.
+This highlighting just how foundational offensive skills are to the explanation of L.
+Messi's wages.
 
-## 4) Ames Housing prices 2018
+## 4) Ames housing prices 2018
 
-TODO:XXX
+Ames 2018, housing data was subset to North Ames (neighborhood with the most house sales).
+Remaining are 338 house sales across 9 variables.
+Using interaction from the global view we select a house with an extreme negative residual and an accurate observation close to it in the data.
+
+```{r casefifa, echo=F, out.width = "100%", fig.cap = "Ames housing 2018 regressing log sales price [2018 USD]."}
+knitr::include_graphics("./figures/case_ames2018.pdf")
+```
+
+TODO:XXX after refresh
 

### 2021-11-24 _paper.rmd
upped dated case studies, minor smithing, and grammarly edits.

@@
 
+TODO:XXX
 ```{r}
@@
 
-The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
+The regression case is not as discrete a feature. Instead, we find it more meaningful to compare the observed value or  residual with the projection. In this case the horizontal positions
 
@@
 While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
+
+<!-- predicted by observed -->
+It is insightful to explore these two approximations against a visual of the model; prediction by observation is added.
+Linked brushing and preserved aesthetic features such as circling misclassifed observations helps link information from the different spaces together.
 
@@
 
+## Interactive features
+
+The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
+Linked brushing also for a rectangular selection of observations.
+These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance, highlighting in the tour.
+The distinction between the former and latter is subtle, but important for developers to grasp.
+In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
+In the latter, the use of `plotly::event_data()` returns the identity of the selected observations.
+This will require a reactive flush, but after that, the sky is the ceiling.
+
 ## Preprocessing
@@
 
+Installation and get started with the package can be acheived by running the following in __R__:
+
+```{r eval=FALSE, echo=TRUE}
+## Download the package:
+install.packages("cheem", dependencies = TRUE)
+## Try the app yourself:
+cheem::run_app()
+## Procss your own data for the app by following the examples in:
+?cheem::cheem_ls
+```
+
 <!-- Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
@@
 -->
-
-## Interactive features
-
-The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
-Linked brushing also for a rectangular selection of observations.
-These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance, highlighting in the tour.
-The distinction between the former and latter is subtle, but important for developers to grasp.
-In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
-In the latter, the use of `plotly::event_data()` returns the identity of the selected observations.
-This will require a reactive flush, but after that, the sky is the ceiling.
 
@@
 
-In figure \@ref(fig:casepenguins) a misclassified point is contrasted with a correctly classified point of its observated class that is nearby in data-space.
+In figure \@ref(fig:casepenguins) a misclassified point is contrasted with a correctly classified point of its observed class that is nearby in data-space.
 The attribution space from the treeshap local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution while primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
-The tour varies the contribution of of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crutial to linear explanation this observation be misclassified.
+The tour varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
+Downplaying the contribution of bill length is crucial to linear explanation this observation be misclassified.
 
@@
 Each of which was labeled as being either milk or dark chocolates.
-With this data we can see if a manufacturer is giving an accurate portrayal of the chocolate.
+With this data, we can see if a manufacturer is giving an accurate portrayal of the chocolate.
 We are curious to see if there are chocolates that nutritionally look like milk chocolates that are labeled as dark chocolates, which may hold a higher market value.
@@
 After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
-The model was fit with 5000 observation of the 9 skill aggregates before being thinned to 500 player to gate occlusion and render time.
+The model was fit with 5000 observations of the 9 skill aggregates before being thinned to 500 players to gate occlusion and render time.
 We compare a leading offensive fielder (L.
 Messi) with that of a leading defensive fielder (V.
-van Dijk).
+van Dijk), the same observations used in figure \@figure(fig:shapdistrbd).
 
@@
 
-The difference in their tree SHAP values highlights the fact that the random forest model found multiple ways to predict wages in a way that makes sense; especially in the disparity of the importance of offensive and defensive skill measures while player position was exogenous to the model.
-Reaction skills were highly important in both cases.
-After removing reaction, we vary the contribution of offensive skills.
-This highlighting just how foundational offensive skills are to the explanation of L.
-Messi's wages.
+With figure \@ref(fig:casefifa) we are going to test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, then offense has almost singular importance for the explanation of the offensive player.
+We vary the contribution of offensive skills.
+In the tour (3rd frame of b), offensive skills moved, and Messi is no longer separated from the group.
+We also notice that accuracy has rotated into the frame, which does maintain some separability.
 
@@
 Ames 2018, housing data was subset to North Ames (neighborhood with the most house sales).
-Remaining are 338 house sales across 9 variables.
+The remaining are 338 house sales across 9 variables.
 Using interaction from the global view we select a house with an extreme negative residual and an accurate observation close to it in the data.
 
-```{r casefifa, echo=F, out.width = "100%", fig.cap = "Ames housing 2018 regressing log sales price [2018 USD]."}
+```{r caseames, echo=F, out.width = "100%", fig.cap = "Ames housing 2018 regressing log sales price [2018 USD]."}
 knitr::include_graphics("./figures/case_ames2018.pdf")
@@
 
-TODO:XXX after refresh
+Figure \@ref(fig:caseames) shows the global view and extrema of the tour.
+The horizontal distance in the tour didn't show a large disparity between our selected points.
+This is not particularly surprising as most of the features have a sizable contribution.
+Rotating any one variable out of the frame will rotate other important variables into frame preserving most of the distance from intercept to prediction.
+However, the tour has revealed an interesting feature worth discussing.
+notice that the observations pivot about the origin, that is the basis roughly halfway between bases in frame 1 and 2 of panel b) the data is near a singular profile.
+This means that there is a basis orthogonal to this point that describes a lot of the variation.
+Knowing these singular frames can help direct to bases that have meaningful variation in the data.
 
@@
 Our idea is to visualize approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
-Then use the linear importance from the primary observation as a projection basis and explore a single variable's importance to the structure identified with use of the manual tour.
+Then use the linear importance from the primary observation as a projection basis and explore a single variable's importance to the structure identified with the use of the manual tour.
 
 We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generalized to any compatible model-explanation pair.
-We apply this to both the classification and regression task.
+We apply this to both the classification and regression tasks.
 In the former, it can be used to see which variables cause a misclassification as benchmarked against a nearby observation and explore changing the contribution of variables.
-In the regression case, we compare observations over exogenous .
-An observation with particularly extreme residual can be compared with nearby points more accurately predicted.
+In the regression case, we compare an observation with particularly extreme residual can be compared with nearby points more accurately predicted or another suitable benchmark.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input in early in the project and to the broader MI\^2 lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\^2 lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages.
 This research was supported in part by Australian government Research Training Program (RTP) scholarships.
 
-The namesake, cheem, refers to a fictional race of humanoid trees in the Doctor Who lore.
-Given that __DALEX__ pulls on from that universe and we originally apply tree shap explanations (specific to tree-based models) we found it fitting.
+The namesake, Cheem, refers to a fictional race of humanoid trees in  Doctor Who lore.
+Given that __DALEX__ pulls on from that universe and we originally apply tree SHAP explanations (specific to tree-based models) we found it fitting.
 

### 2021-11-24 _paper.tex
upped dated case studies, minor smithing, and grammarly edits.

@@
 \hypersetup{
-  pdftitle={Interogating the linear variable importance of local explanations of black-box models with animated linear projections},
+  pdftitle={Interrogating the linear variable importance of local explanations of black-box models with animated linear projections},
   colorlinks=true,
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs,array}
@@
 
-\title{Interogating the linear variable importance of local explanations of black-box models with animated linear projections}
+\title{Interrogating the linear variable importance of local explanations of black-box models with animated linear projections}
 \author{}
@@
 
+TODO:XXX
+
 \hypertarget{sec:applicationdesign}{%
@@
 
-The regression case is not as discrete a feature. Instead, the prediction or the residual becomes the comparison of accuracy.
+The regression case is not as discrete a feature. Instead, we find it more meaningful to compare the observed value or residual with the projection. In this case the horizontal positions
 
@@
 While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
+
+It is insightful to explore these two approximations against a visual of the model; prediction by observation is added.
+Linked brushing and preserved aesthetic features such as circling misclassifed observations helps link information from the different spaces together.
 
@@
 
+\hypertarget{interactive-features}{%
+\subsection{Interactive features}\label{interactive-features}}
+
+The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
+Linked brushing also for a rectangular selection of observations.
+These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance, highlighting in the tour.
+The distinction between the former and latter is subtle, but important for developers to grasp.
+In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
+In the latter, the use of \texttt{plotly::event\_data()} returns the identity of the selected observations.
+This will require a reactive flush, but after that, the sky is the ceiling.
+
 \hypertarget{preprocessing}{%
@@
 
-\hypertarget{interactive-features}{%
-\subsection{Interactive features}\label{interactive-features}}
+Installation and get started with the package can be acheived by running the following in \textbf{R}:
 
-The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
-Linked brushing also for a rectangular selection of observations.
-These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance, highlighting in the tour.
-The distinction between the former and latter is subtle, but important for developers to grasp.
-In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
-In the latter, the use of \texttt{plotly::event\_data()} returns the identity of the selected observations.
-This will require a reactive flush, but after that, the sky is the ceiling.
+\begin{Shaded}
+\begin{Highlighting}[]
+\DocumentationTok{\#\# Download the package:}
+\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
+\DocumentationTok{\#\# Try the app yourself:}
+\NormalTok{cheem}\SpecialCharTok{::}\FunctionTok{run\_app}\NormalTok{()}
+\DocumentationTok{\#\# Procss your own data for the app by following the examples in:}
+\NormalTok{?cheem}\SpecialCharTok{::}\NormalTok{cheem\_ls}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-Palm penguins data Horst, Hill, and Gorman (2020) consist of four physical measurements of 3 species of penguins foraging near Plamer Station, Antarctica.
-The random forest model classifies the species of the penguin given the physical measurements.
+Palmer penguins data Horst, Hill, and Gorman (2020) consist of 330 observations across four physical measurements of 3 species of penguins foraging near Palmer Station, Antarctica.
+A random forest model was fit, classifying the species of the penguin given the physical measurements.
 
@@
 
-\caption{Species classification of Palmer penguin data.
-a) The primary observation is misclassified as an orange point and is being compared to a nearby correctly classified orange point.}\label{fig:casepenguins}
+\caption{Species classification of Palmer penguin data.}\label{fig:casepenguins}
 \end{figure}
+
+In figure \ref{fig:casepenguins} a misclassified point is contrasted with a correctly classified point of its observed class that is nearby in data-space.
+The attribution space from the treeshap local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution while primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
+The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
+Downplaying the contribution of bill length is crucial to linear explanation this observation be misclassified.
 
@@
 
-TODO:XXX
+The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
+Each of which was labeled as being either milk or dark chocolates.
+With this data, we can see if a manufacturer is giving an accurate portrayal of the chocolate.
+We are curious to see if there are chocolates that nutritionally look like milk chocolates that are labeled as dark chocolates, which may hold a higher market value.
+We should note that not all chocolates consist wholly of chocolate.
+The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable.
+A random forest model is fit classifying the type of chocolate.
+We'll select chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85\% cocoa.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates}
+
+}
+
+\caption{Chocolates data  type classification (milk or dark). }\label{fig:casechocolates}
+\end{figure}
+
+From figure \ref{fig:casechocolates} we similarly see that attribution space that is more separable relative to data-space.
+Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while there 8 of the inverse case.
+Calories from fat is the measure with the largest difference in treeshap attribution between these points.
 
@@
 After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
+The model was fit with 5000 observations of the 9 skill aggregates before being thinned to 500 players to gate occlusion and render time.
+We compare a leading offensive fielder (L.
+Messi) with that of a leading defensive fielder (V.
+van Dijk), the same observations used in figure @figure(fig:shapdistrbd).
 
@@
 \caption{FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
-a) The primary observation is a star offensive player (L.
+The primary observation is star offensive player (L.
 Messi) as compared with a top defensive player (V.
-van Dijk).
-b) Starting from a basis where offensive skills ('*' toward the right, high predicted wages) are high we observe that how contingent this variable is for explaining the high wages.
-As the contribution decreases the primary observation moves left into the bulk of the other observations corresponding to lower predicted wages.}\label{fig:casefifa}
+van Dijk).}\label{fig:casefifa}
 \end{figure}
 
-The difference in their tree shap values highlights the fact that the random forest model found multiple ways to increase wages; that prediction accuracy was improved by varying linear variable importance that makes sense in light of the player positions that we exogenous to the model.
+With figure \ref{fig:casefifa} we are going to test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, then offense has almost singular importance for the explanation of the offensive player.
+We vary the contribution of offensive skills.
+In the tour (3rd frame of b), offensive skills moved, and Messi is no longer separated from the group.
+We also notice that accuracy has rotated into the frame, which does maintain some separability.
 
 \hypertarget{ames-housing-prices-2018}{%
-\subsection{4) Ames Housing prices 2018}\label{ames-housing-prices-2018}}
+\subsection{4) Ames housing prices 2018}\label{ames-housing-prices-2018}}
 
-TODO:XXX
+Ames 2018, housing data was subset to North Ames (neighborhood with the most house sales).
+The remaining are 338 house sales across 9 variables.
+Using interaction from the global view we select a house with an extreme negative residual and an accurate observation close to it in the data.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/case_ames2018}
+
+}
+
+\caption{Ames housing 2018 regressing log sales price [2018 USD].}\label{fig:caseames}
+\end{figure}
+
+Figure \ref{fig:caseames} shows the global view and extrema of the tour.
+The horizontal distance in the tour didn't show a large disparity between our selected points.
+This is not particularly surprising as most of the features have a sizable contribution.
+Rotating any one variable out of the frame will rotate other important variables into frame preserving most of the distance from intercept to prediction.
+However, the tour has revealed an interesting feature worth discussing.
+notice that the observations pivot about the origin, that is the basis roughly halfway between bases in frame 1 and 2 of panel b) the data is near a singular profile.
+This means that there is a basis orthogonal to this point that describes a lot of the variation.
+Knowing these singular frames can help direct to bases that have meaningful variation in the data.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input in early in the project and to the broader MI\^{}2 lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\^{}2 lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages.
 This research was supported in part by Australian government Research Training Program (RTP) scholarships.
 
-The namesake, cheem, refers to a fictional race of humanoid trees in the Doctor Who lore.
-Given that \textbf{DALEX} pulls on from that universe and we originally apply tree shap explanations (specific to tree-based models) we found it fitting.
+The namesake, Cheem, refers to a fictional race of humanoid trees in Doctor Who lore.
+Given that \textbf{DALEX} pulls on from that universe and we originally apply tree SHAP explanations (specific to tree-based models) we found it fitting.
 

### 2021-11-24 figures/_make_figures_case_study.r

@@ -13,32 +13,34 @@
     dir("../cheem/inst/shiny_apps/cheem_initial/data/")
   
   ## Load data:
-  penguins_ls   <- readRDS(paste0(fp, "1preprocess_penguins.rds"))
-  fifa_ls       <- readRDS(paste0(fp, "3preprocess_fifa.rds"))
-  ames2018_ls   <- readRDS(paste0(fp, "7preprocess_ames2018.rds"))
+  penguins_ls   <- readRDS(paste0(fp, "preprocess_penguins.rds"))
+  fifa_ls       <- readRDS(paste0(fp, "preprocess_fifa.rds"))
+  ames2018_ls   <- readRDS(paste0(fp, "preprocess_ames2018.rds"))
   chocolates_ls <- readRDS(paste0(fp, "preprocess_chocolates.rds"))
+  
+  .spaces <- paste(rep(" ", 3L), collapse = "")
+  .x_axis_title <- c("             x: PC1, y: PC2", "        x: PC1, y: PC2", "x: predicted, y: observed")
+  .x_axis_title <- paste(.x_axis_title, collapse = .spaces)
+  
+  .spaces <- paste(rep(" ", 5L), collapse = "")
+  .x_axis_title_reg <- c("            x: PC1, y: PC2", "        x: PC1, y: PC2", "x: predicted, y: observed")
+  .x_axis_title_reg <- paste(.x_axis_title_reg, collapse = .spaces)
 }
 
 ## Penguins classification ------
 {
   ## Data setup, spinifex::penguins
   names(penguins_ls)
-  prim_obs <- 15L
-  comp_obs <- 282L
+  prim_obs <- 124L
+  comp_obs <- 86L
   
   ### Global view and cheem tour stills
   .glob_view <- global_view(
     penguins_ls, prim_obs, comp_obs, as_ggplot = TRUE) +
-    theme_bw() +
-    labs(x = "PC1") +
-    theme(legend.position = "off",
-          axis.text  = element_blank(),
-          axis.ticks = element_blank())
-  # legend.direction = "vertical", ## Levels within aesthetic
-  # legend.box = "horizontal",     ## Between aesthetic
-  # legend.margin = ggplot2::margin(1L,1L,1L,1L, "mm")) +
-  # labs(color='Predicted class', shape='Predicted class')
-  
+    theme(legend.position = "right") +
[…]

### 2021-12-03 _paper.rmd
images to .png; to allow for printing .pdf (.pdf format with alpha causes `flattening` to hang).

@@
 Applications have been known to reflect common biases against sex [@dastin_amazon_2018, @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-These are all-too common issue, when endogenous or exogenous biases in the training data and identified and mimicked by models.
-Another issue is that of data-drift when new data is outside the support of latent or exogenous explanatory variables.
+These are all-too common of an issue, when endogenous or exogenous biases in the training data and identified and mimicked by models.
+Another issue is that of data drift when new data is outside the support of latent or exogenous explanatory variables.
 Data drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
@@
 In a _manual_ tour [@cook_manual_1997, @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
-<!--tours and models --> Applying tours in conjunction with models has been previously done,  _ie_ for exploring various statistical model fits [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013, @da_silva_projection_2021].
+<!--tours and models --> Applying tours in conjunction with models has been done in a couple contexts, _ie_ for exploring various statistical model fits [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013, @da_silva_projection_2021].
 
@@
 c) Breakdown plots of the observations the explanation used to additively explain the difference between the intercept and prediction"}
-knitr::include_graphics("./figures/shap_distr_bd.pdf")
+knitr::include_graphics("./figures/shap_distr_bd.png")
 ```
@@
 
-The regression case is not as discrete a feature. Instead, we find it more meaningful to compare the observed value or  residual with the projection. In this case the horizontal positions
+The regression case is not as discrete a feature.
+Instead, we find it more meaningful to compare the observed value or  residual with the projection.
+In this case the horizontal positions are the same, that is projection of the data through the attribution of the explanation and then varying as the radial tour changes the contributions of the selected variable.
+While the height of the points is fixed to the fixed either to the observed y and residuals for the middle and right panels respectively.
 
@@
 
-<!-- Dipiction of a frame -->
+<!-- Dupiction of a frame -->
 The SHAP values of the full dataset are shown as horizontal parrallel coordiate plot at the top with the selected observations highlighted.
@@
 ```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Species classification of Palmer penguin data."}
-knitr::include_graphics("./figures/case_penguins.pdf")
+knitr::include_graphics("./figures/case_penguins.png")
 ```
@@
 ```{r casechocolates, echo=F, out.width = "100%", fig.cap = "Chocolates data  type classification (milk or dark). "}
-knitr::include_graphics("./figures/case_chocolates.pdf")
+knitr::include_graphics("./figures/case_chocolates.png")
 ```
@@
 van Dijk)."}
-knitr::include_graphics("./figures/case_fifa.pdf")
+knitr::include_graphics("./figures/case_fifa.png")
 ```
@@
 ```{r caseames, echo=F, out.width = "100%", fig.cap = "Ames housing 2018 regressing log sales price [2018 USD]."}
-knitr::include_graphics("./figures/case_ames2018.pdf")
+knitr::include_graphics("./figures/case_ames2018.png")
 ```

### 2021-12-03 figures/_make_figures.r

@@ -109,6 +109,9 @@ Y <- .raw$wage_eur ## unscaled wages in Euros, assumed 2020 valuation.
 .hp_node <- if(.is_y_disc == TRUE) 1L else 5L
 .hp_node <- max(.hp_node, nrow(X) / 500L)
 
+
+## NOTE: this DALEX::predict_parts("SHAP")
+## NOT theeshap::treeshap()
 rf_mod <- randomForest::randomForest(Y~., data = data.frame(Y, X),
                                      mtry = .hp_mtry,
                                      nodesize = .hp_node,
@@ -168,7 +171,7 @@ dist_df$variable <- factor(dist_df$variable, levels = rev(.lvl_ord))
     scale_color_brewer(palette = "Dark2") +
     scale_fill_brewer(palette = "Dark2") +
     labs(title="SHAP distribution",
-         y = "variable", x = "normalized SHAP values\n the median of the contributions while permuting X's") +
+         y = "variable", x = "Normalized SHAP values\n the median of the contributions while permuting X's") +
     theme(legend.position = "off")
 )
 
@@ -199,7 +202,7 @@ bd_df <- bd_df[is.na(bd_df$variable) == FALSE, ]
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
     labs(title="Breakdown plot",
-         y = "variable", x = "normalized contribution to prediction | variable order") +
+         y = "variable", x = "Normalized contribution to prediction | variable order") +
     theme(legend.margin = margin(0,0,0,0),
           legend.position = "bottom",
           axis.text.x = element_blank(),
@@ -214,7 +217,7 @@ wages_df <- tibble::tibble(
     geom_segment(size=3L) +
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
-    labs(y = "player", x = "wages [2020 Euros]") +
+    labs(y = "player", x = "Wages [2020 Euros]") +
     theme(legend.position = "off"))
 ### Plot together
 require("patchwork")
@@ -227,8 +230,8 @@ require("cowplot")
 
 ## SAVE -----
 ggplot2::ggsave(
-  "./figures/shap_distr_bd.pdf",
-  cp, device = "pdf", width = 7, height = 8, units = "in")
+  "./figures/shap_distr_bd.png",
+  cp, device = "png", width = 7, height = 8, units = "in")
 
 
 # ch5_fig2_global_space -----
[…]

### 2021-12-03 figures/_make_figures_case_study.r

@@ -18,13 +18,9 @@
   ames2018_ls   <- readRDS(paste0(fp, "preprocess_ames2018.rds"))
   chocolates_ls <- readRDS(paste0(fp, "preprocess_chocolates.rds"))
   
-  .spaces <- paste(rep(" ", 3L), collapse = "")
-  .x_axis_title <- c("             x: PC1, y: PC2", "        x: PC1, y: PC2", "x: predicted, y: observed")
-  .x_axis_title <- paste(.x_axis_title, collapse = .spaces)
-  
-  .spaces <- paste(rep(" ", 5L), collapse = "")
-  .x_axis_title_reg <- c("            x: PC1, y: PC2", "        x: PC1, y: PC2", "x: predicted, y: observed")
-  .x_axis_title_reg <- paste(.x_axis_title_reg, collapse = .spaces)
+
+  .x_axis_title     <- "          x: PC1, y: PC2        x: PC1, y: PC2  x: predicted, y: observed"
+  .x_axis_title_reg <- "          x: PC1, y: PC2             x: PC1, y: PC2     x: predicted, y: observed"
 }
 
 ## Penguins classification ------
@@ -56,10 +52,11 @@
     labels = c("a)", "b)"),
     ncol = 1)#, rel_heights = c(1.5, 1))
 }
+
 ### Save
 ggplot2::ggsave(
-  "./figures/case_penguins.pdf",
-  plot = .cp, device = "pdf",
+  "./figures/case_penguins.png",
+  plot = .cp, device = "png",
   width = 6, height = 5, units = "in")
 .m <- gc()
 
@@ -98,8 +95,8 @@ ggplot2::ggsave(
 }
 ### Save
 ggplot2::ggsave(
-  "./figures/case_chocolates.pdf",
-  plot = .cp, device = "pdf",
+  "./figures/case_chocolates.png",
+  plot = .cp, device = "png",
   width = 6, height = 5, units = "in")
 .m <- gc()
 
@@ -135,8 +132,8 @@ ggplot2::ggsave(
 }
 ### Save
 ggplot2::ggsave(
-  "./figures/case_fifa.pdf",
-  plot = .cp, device = "pdf",
+  "./figures/case_fifa.png",
+  plot = .cp, device = "png",
   width = 5.5, height = 8, units = "in")
 .m <- gc()
 
@@ -173,14 +170,14 @@ ggplot2::ggsave(
 }
 ### Save
 ggplot2::ggsave(
-  "./figures/case_ames2018.pdf",
[…]

### 2021-12-04 _paper.rmd
Edits before restructuing to classification/regression case.

@@
   Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of explainable AI (XAI).
+In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
 XAI attempts to shed light on these models, one such approach is the use of local explanations.
@@
 Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
-The implementation of our framework is available as an R package **cheem** available at [github.com/nspyrison/cheem](https://github.com/nspyrison/cheem).
+The implementation of our framework is available as an __R__ package __cheem__ available at [github.com/nspyrison/cheem](https://github.com/nspyrison/cheem).
 # documentclass: jdssv
@@
 <!-- introduce explanatory vs predictive modeling -->
-Breiman and Shmueli [@breiman_statistical_2001, @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling is performed to predict new or future out-of-sample observations.
+Breiman and Shmueli [@breiman_statistical_2001; @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling is performed to predict new or future out-of-sample observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
@@
 <!-- ## XAI & interpretability crisis -->
-Black-box models are becoming increasingly common, but not without their share of controversy and issues [@oneil_weapons_2016, @kodiyan_overview_2019].
-Applications have been known to reflect common biases against sex [@dastin_amazon_2018, @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-These are all-too common of an issue, when endogenous or exogenous biases in the training data and identified and mimicked by models.
+Black-box models are becoming increasingly common, but not without their share of controversy and issues [@oneil_weapons_2016; @kodiyan_overview_2019].
+Applications have been known to reflect common biases against sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
+Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable which is then built into the model.
 Another issue is that of data drift when new data is outside the support of latent or exogenous explanatory variables.
-Data drift can lead to worse predictions [@lazer_parable_2014, @salzberg_why_2014].
-Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@adadi_peeking_2018, @arrieta_explainable_2020].
+Data drift can lead to worse predictions [@lazer_parable_2014; @salzberg_why_2014].
+Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@adadi_peeking_2018; @arrieta_explainable_2020].
 
@@
 <!-- SHAP -->
-To illustrate our work we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010, @strumbelj_explaining_2014].
+To illustrate our work we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010; @strumbelj_explaining_2014].
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
@@
 <!-- ## Data visualization tours -->
-In multivariate data visualization a _tour_ [@asimov_grand_1985, @buja_grand_1986, @lee_review_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
-Tours are viewed as an animation over small changes to a projection basis.
+In multivariate data visualization a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_review_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+Tours are viewed as an animation over small changes to the projection basis.
 Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
@@
 There are various types of tours, which are distinguished by the sequence of projection bases.
-In a _manual_ tour [@cook_manual_1997, @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
-<!--tours and models --> Applying tours in conjunction with models has been done in a couple contexts, _ie_ for exploring various statistical model fits [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013, @da_silva_projection_2021].
+In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
+<!--tours and models --> Applying tours to models has been done in a couple of contexts.
+Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013; @da_silva_projection_2021].
 
@@
 <!-- What-if & counterfactual analysis -->
-The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _Ceteris paribus_[@biecek_ceterisparibus_2020].
+The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _Ceteris paribus_ [@biecek_ceterisparibus_2020].
 Latin for “other things held constant” or “all else unchanged”, is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation a of the factual observed case by varying the basis, the angle of the data object.
-Another, but related difference is that this geometric approach maintains orthogonal dimensions.
-That is to say when the contribution of some variables decrease the the contributions of others necessarily increases.
+In contrast, our approach is a geometric explanation of the factual observed case by varying the basis, the angle of the data object.
+Another difference is that the basis is restricted on orthonormality.
+That is to say when the contribution of one variable decreases the contributions of others necessarily increases such that there is a full component in that direction.
 
+<!-- Paper structure -->
 The rest of this paper is organized as follows.
@@
 [Tours and the radial tour](#sec:tour) digs deeper into the animation and what it shows.
-The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis and discusses the backend details of the package and preprocessing.
+The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis, and discusses the backend details of the package and preprocessing.
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
-We conclude with a [discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
+We conclude with a [Discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
 
@@
 Consider a highly non-linear model.
-At face value its hard to say which variable are sensitive to the crossing of classification boundary or identify which variables caused an observation to have a relatively extreme residual.
-Local explanations shed light on these cases by approximating linear variable importance at in vicinity of one observation.
+At face value it is hard to say which variable(s) are sensitive to crossing of classification boundary or identify which variables caused an observation to have a relatively extreme residual.
+Local explanations shed light on these cases by approximating linear variable importance in the vicinity of one observation.
 
@@
 Figure 6 of @arrieta_explainable_2020 gives a broad summarization of the taxonomy and literature of explanation techniques.
-This includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016, @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
+This includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016; @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
 There are a fewer number of model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017] and their variants are popular.
@@
 <!-- Uses of local explanations -->
-These local explanation are used in a variety of ways depending on the data.
-For images saliency maps [@simonyan_deep_2014] a heatmap can be used to show which pixels were important for distinguishing pictures of wolfs or huskies for instance.
-In text analysis saliency can be used to highlight influential words [@vanni_textual_2018].
-In the case of numeric regression they can be used to explain variable additive contributions from intercept to prediction [@ribeiro_why_2016].
+These instance level explanations are used in a variety of ways depending on the data.
+In images saliency maps overlay or offset a heatmap indicating which pixels were important [@simonyan_deep_2014].
+For instance, or the presence of snow may be highlighted when distinguishing if a picture of contains a wolf or husky.
+In text analysis, Word-level contextual sentiment analysis can be used to highlight sentiment and magnitude of influential words [@vanni_textual_2018].
+In the case of numeric regression, they can be used to explain variable additive contributions from intercept to prediction [@ribeiro_why_2016].
 
@@
 <!-- SHAP and history -->
-SHaply Additive exPlanations, or SHAP approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players that contribute to the score [@shapley_value_1953].
+SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+This idea stems from the field of game theory where Shapley devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score [@shapley_value_1953].
 
+<!-- fifa example -->
 To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season [@leone_fifa_2020].
-We have 5000 observations of 9 skill measures (after aggregating variable with high correlation).
+We have 5000 observations of nine skill measures (after aggregating variable with high correlation).
 A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
@@
 Figure \@ref(fig:shapdistrbd) shows the SHAP values of these players.
-Panel a) shows these players recieve a sizable difference in wages.
+Panel a) shows these players receive a sizable difference in wages.
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
@@
 
-```{r shapdistrbd, echo=F, fig.cap = "Illustration of the distribution of SHAP attributions, the SHAP values, and a breakdown plot, the typical visual of SHAP local explanations.
-For FIFA 2020 data, of a random forest model regressing wages from 9 skill attributes for a star offensive and defensive player.
+```{r shapdistrbd, echo=F, fig.cap = "Illustration of the distribution of SHAP attributions and a breakdown plot.
+From FIFA 2020 data, a random forest model regresses wages from 9 skill attributes for a star offensive and defensive player.
 a) The players have very different wages.
 b) Shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
-The median of these distributions are the final SHAP values, notice that that the variable importance differs across the exogenous information of player position.
+The median of these distributions are the final SHAP values.
+The variable importance differs across the exogenous information of player position.
 These explanations make sense; the variable importances make sense in light of the position of the player.
@@
 It is insightful to explore these two approximations against a visual of the model; prediction by observation is added.
-Linked brushing and preserved aesthetic features such as circling misclassifed observations helps link information from the different spaces together.
+Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
 
@@
 Namely, the linear attribution of that variable is used as a 1D projection basis.
-This is the approximate contributions of the variables that this model uses to justify its prediction for the observation.
+This is the approximate contribution of the variables that this model uses to justify its prediction for the observation.
 
 <!-- Dupiction of a frame -->
-The SHAP values of the full dataset are shown as horizontal parrallel coordiate plot at the top with the selected observations highlighted.
+The SHAP values of the full dataset are shown as horizontal parallel coordinate plots at the top with the selected observations highlighted.
 In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
-In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of is fixed to the observations prediction and residual respectively.
-The current projection basis is depicted on the bottom, where the width of the bar is mapped to variables contribution to the horizontal axis.
+In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position is fixed to the observations prediction and residual respectively.
+The current projection basis is depicted on the bottom, where the width of a bar is mapped to the variable's contribution to the horizontal axis.
 The starting frame is the normalized SHAP values of the primary observation.
@@
 <!-- regression case -->
-In the regression case we compare an observation with a particularly extreme residual as compared with a neighboring point that is more  accurately predicted.
+In the regression case, we compare an observation with a particularly extreme residual as compared with a neighboring point that is more  accurately predicted.
 The global tour is primarily the same without the misclassified circles, but the tooltip now shows the observation's residual.
-We simulate data, 5 variable from the uniform distribution (between 0 and 5) and $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
-We fit a random forest and extract the all tree SHAP values.
+We simulate data, five variables from the uniform distribution in the support [0, 5] and $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
+We fit a random forest and extract the tree SHAP values from all observations.
 The first thing we notice from the global view is the shape difference between the data- and attribution-spaces, which are primarily circular are triangular respectively.
@@
 Though linked brushing we find that the attribution space is highly correlated with the $y$ value.
-The highest values occupy 1 corner of the triangle, the lowest values cluster on the opposite side and the middle values form a ')' shape filling out the triangle in attribution space.
-In the regression case the tour view is a bit different, this is illustrated in figure \@ref(cheemtourreg).
-The the current frames basis and distribution are shown to the left.
+The highest values occupy one corner of the triangle, the lowest values cluster on the opposite side, and the middle values form a ')' shape filling out the triangle in attribution space.
+In the regression case, the tour view is a bit different, this is illustrated in figure \@ref(fig:cheemtourreg).
+The current frames basis and distribution are shown to the left.
 The horizontal positions of the two scatter plots are the resulting 1D projection of the basis while the heights are fixed to the observed y and the residuals.
@@
 <!-- tour regression caption -->
-(ref:cheemtourreg) Cheem tour, regression case.
+(ref:cheemtourreg-cap) Cheem tour, regression case.
 A random forest model predicting the continuous $y$ of the of simulated data where $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
@@
 
-```{r cheemtourreg, echo=F, out.width = "100%", fig.cap = "(ref:cheemtourreg)"}
+```{r cheemtourreg, echo=F, out.width = "100%", fig.cap = "(ref:cheemtourreg-cap)"}
 knitr::include_graphics("./figures/cheem_tour_toy_reg.PNG")
@@
 The global view shows all observations in attribution space requiring that we must extract the variable weightings from \emph{all} observations rather than just one.
-	\item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components.
+	<!-- \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components. -->
 \end{itemize}
@@
 The time to preprocess the data will vary significantly with the choice of model and local explanation.
-However, for reference, the FIFA data, 5000 observations of 9 explanatory variables, took 0.6 seconds to create PCA for both the data and attribution spaces.
-On the same data, with modest hyper parameters a random forest model fit in 2.9 seconds while extracting the tree SHAP values of each observation took 254 seconds combined.
+For reference, the FIFA data, 5000 observations of 9 explanatory variables, took  2.9 seconds to fit a random forest models of modest hyperparmeters.
+Extracting the tree SHAP values of each observation took 254 seconds combined.
+PCA and statistics of the variables and attribution took 0.6 seconds.
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the time will be spent on the local attribution.
-This makes tree SHAP a good candidate to start with.
-The package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes that are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+This makes tree SHAP, with its reduced comupational complexity a good candidate to start with.
+Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 
-The above-described method and application are implemented as an open-source __R__ package, __cheem__ _TODO:XXX site github?
-cran?_.
+The above-described method and application are implemented as an open-source __R__ package, __cheem__ _TODO:XXX include  cran URL after published_.
 Preprocessing was facilitated with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
@@
 Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive HTML widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and
-the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
+__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
 
@@
 install.packages("cheem", dependencies = TRUE)
-## Try the app yourself:
-cheem::run_app()
-## Procss your own data for the app by following the examples in:
-?cheem::cheem_ls
+library("cheem")
+## Try the app:
+run_app()
+## Bring your own data: follow the examples in cheem_ls()
+?cheem_ls
 ```
@@
 
-## 1) penguins, species classification
+## 1) Penguin, species classification
 
-Palmer penguins data [@gorman_ecological_2014, @horst_palmerpenguins_2020] consist of 330 observations across four physical measurements of 3 species of penguins foraging near Palmer Station, Antarctica.
+Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] consist of 330 observations across four physical measurements of three species of penguins foraging near Palmer Station, Antarctica.
 A random forest model was fit, classifying the species of the penguin given the physical measurements.
@@
 A random forest model is fit classifying the type of chocolate.
-We'll select chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85% cocoa.
+We select a chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85% cocoa.
 
@@
 
-The 2020 season FIFA[@leone_fifa_2020, @biecek_dalex_2018], contains many skill measurements of soccer/football players along with wage information.
+The 2020 season FIFA [@leone_fifa_2020; @biecek_dalex_2018], contains many skill measurements of soccer/football players along with wage information.
 After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
@@
 Messi) with that of a leading defensive fielder (V.
-van Dijk), the same observations used in figure \@figure(fig:shapdistrbd).
+van Dijk), the same observations used in figure \@ref(fig:shapdistrbd).
 
@@
 
-Ames 2018, housing data was subset to North Ames (neighborhood with the most house sales).
+Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
 The remaining are 338 house sales across 9 variables.
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\^2 lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages.
 This research was supported in part by Australian government Research Training Program (RTP) scholarships.
@@
 The namesake, Cheem, refers to a fictional race of humanoid trees in  Doctor Who lore.
-Given that __DALEX__ pulls on from that universe and we originally apply tree SHAP explanations (specific to tree-based models) we found it fitting.
+__DALEX__ pulls on from that universe and we originally apply tree SHAP explanations which are specific to tree-based models.
 

### 2021-12-04 _paper.tex
images to .png; to allow for printing .pdf (.pdf format with alpha causes `flattening` to hang).
Edits before restructuing to classification/regression case.

@@
 Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of explainable AI (XAI).
+In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
 XAI attempts to shed light on these models, one such approach is the use of local explanations.
@@
 Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
-The implementation of our framework is available as an R package \textbf{cheem} available at \href{https://github.com/nspyrison/cheem}{github.com/nspyrison/cheem}.
+The implementation of our framework is available as an \textbf{R} package \textbf{cheem} available at \href{https://github.com/nspyrison/cheem}{github.com/nspyrison/cheem}.
 \end{abstract}
@@
 
-Breiman and Shmueli Shmueli (2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed to predict new or future out-of-sample observations.
+Breiman and Shmueli (Breiman 2001; Shmueli 2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed to predict new or future out-of-sample observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
@@
 
-Black-box models are becoming increasingly common, but not without their share of controversy and issues Kodiyan (2019).
-Applications have been known to reflect common biases against sex Duffy (2019), race (Larson et al.
+Black-box models are becoming increasingly common, but not without their share of controversy and issues (O'neil 2016; Kodiyan 2019).
+Applications have been known to reflect common biases against sex (Dastin 2018; Duffy 2019), race (Larson et al.
 2016), and age (Díaz et al.
 2018).
-These are all-too common issue, when endogenous or exogenous biases in the training data and identified and mimicked by models.
-Another issue is that of data-drift when new data is outside the support of latent or exogenous explanatory variables.
-Data drift can lead to worse predictions Salzberg (2014).
-Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI Arrieta et al.
-(2020).
+Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable which is then built into the model.
+Another issue is that of data drift when new data is outside the support of latent or exogenous explanatory variables.
+Data drift can lead to worse predictions (Lazer et al.
+2014; Salzberg 2014).
+Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI (Adadi and Berrada 2018; Arrieta et al.
+2020).
 
@@
 
-To illustrate our work we apply the model-agnostic explanation SHAP Štrumbelj and Kononenko (2014).
+To illustrate our work we apply the model-agnostic explanation SHAP (Strumbelj and Kononenko 2010; Štrumbelj and Kononenko 2014).
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
@@
 
-In multivariate data visualization a \emph{tour} S.
+In multivariate data visualization a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
 Lee et al.
-(2021) is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
-Tours are viewed as an animation over small changes to a projection basis.
+2021) is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+Tours are viewed as an animation over small changes to the projection basis.
 Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
@@
 There are various types of tours, which are distinguished by the sequence of projection bases.
-In a \emph{manual} tour Spyrison and Cook (2020) this path is defined by changing the contribution of a selected variable.
-Applying tours in conjunction with models has been previously done, \emph{ie} for exploring various statistical model fits (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path Silva, Cook, and Lee (2021).
+In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020) this path is defined by changing the contribution of a selected variable.
+Applying tours to models has been done in a couple of contexts.
+Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path (Y.
+D.
+Lee et al.
+2013; Silva, Cook, and Lee 2021).
 
@@
 
-The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{Ceteris paribus}(Biecek 2020).
+The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{Ceteris paribus} (Biecek 2020).
 Latin for ``other things held constant'' or ``all else unchanged,'' is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation a of the factual observed case by varying the basis, the angle of the data object.
-Another, but related difference is that this geometric approach maintains orthogonal dimensions.
-That is to say when the contribution of some variables decrease the the contributions of others necessarily increases.
+In contrast, our approach is a geometric explanation of the factual observed case by varying the basis, the angle of the data object.
+Another difference is that the basis is restricted on orthonormality.
+That is to say when the contribution of one variable decreases the contributions of others necessarily increases such that there is a full component in that direction.
 
@@
 \protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into the animation and what it shows.
-The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis and discusses the backend details of the package and preprocessing.
+The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis, and discusses the backend details of the package and preprocessing.
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
-We conclude with a \protect\hyperlink{sec:discussion}{discussion} of the insights we draw from classification and regression tasks.
+We conclude with a \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
 
@@
 (2020) gives a broad summarization of the taxonomy and literature of explanation techniques.
-This includes a large number of model-specific explanations such as deepLIFT, Shrikumar, Greenside, and Kundaje (2017) a popular recursive method for estimating importance in neural networks.
+This includes a large number of model-specific explanations such as deepLIFT, (Shrikumar et al.
+2016; Shrikumar, Greenside, and Kundaje 2017) a popular recursive method for estimating importance in neural networks.
 There are a fewer number of model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (S.
@@
 
-These local explanation are used in a variety of ways depending on the data.
-For images saliency maps (Simonyan, Vedaldi, and Zisserman 2014) a heatmap can be used to show which pixels were important for distinguishing pictures of wolfs or huskies for instance.
-In text analysis saliency can be used to highlight influential words (Vanni et al.
+These instance level explanations are used in a variety of ways depending on the data.
+In images saliency maps overlay or offset a heatmap indicating which pixels were important (Simonyan, Vedaldi, and Zisserman 2014).
+For instance, or the presence of snow may be highlighted when distinguishing if a picture of contains a wolf or husky.
+In text analysis, Word-level contextual sentiment analysis can be used to highlight sentiment and magnitude of influential words (Vanni et al.
 2018).
-In the case of numeric regression they can be used to explain variable additive contributions from intercept to prediction (Ribeiro, Singh, and Guestrin 2016).
+In the case of numeric regression, they can be used to explain variable additive contributions from intercept to prediction (Ribeiro, Singh, and Guestrin 2016).
 
@@
 
-SHaply Additive exPlanations, or SHAP approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate individual's contribution to cooperative games by permuting the players that contribute to the score (Shapley 1953).
+SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
+This idea stems from the field of game theory where Shapley devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score (Shapley 1953).
 
 To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season (Leone 2020).
-We have 5000 observations of 9 skill measures (after aggregating variable with high correlation).
+We have 5000 observations of nine skill measures (after aggregating variable with high correlation).
 A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
@@
 Figure \ref{fig:shapdistrbd} shows the SHAP values of these players.
-Panel a) shows these players recieve a sizable difference in wages.
+Panel a) shows these players receive a sizable difference in wages.
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
@@
 
-{\centering \includegraphics{./figures/shap_distr_bd}
+{\centering \includegraphics[width=29.17in]{./figures/shap_distr_bd}
 
@@
 
-\caption{Illustration of the distribution of SHAP attributions, the SHAP values, and a breakdown plot, the typical visual of SHAP local explanations.
-For FIFA 2020 data, of a random forest model regressing wages from 9 skill attributes for a star offensive and defensive player.
+\caption{Illustration of the distribution of SHAP attributions and a breakdown plot.
+From FIFA 2020 data, a random forest model regresses wages from 9 skill attributes for a star offensive and defensive player.
 a) The players have very different wages.
 b) Shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
-The median of these distributions are the final SHAP values, notice that that the variable importance differs across the exogenous information of player position.
+The median of these distributions are the final SHAP values.
+The variable importance differs across the exogenous information of player position.
 These explanations make sense; the variable importances make sense in light of the position of the player.
@@
 
-The regression case is not as discrete a feature. Instead, we find it more meaningful to compare the observed value or residual with the projection. In this case the horizontal positions
+The regression case is not as discrete a feature.
+Instead, we find it more meaningful to compare the observed value or residual with the projection.
+In this case the horizontal positions are the same, that is projection of the data through the attribution of the explanation and then varying as the radial tour changes the contributions of the selected variable.
+While the height of the points is fixed to the fixed either to the observed y and residuals for the middle and right panels respectively.
 
@@
 Namely, the linear attribution of that variable is used as a 1D projection basis.
-This is the approximate contributions of the variables that this model uses to justify its prediction for the observation.
+This is the approximate contribution of the variables that this model uses to justify its prediction for the observation.
 
-The SHAP values of the full dataset are shown as horizontal parrallel coordiate plot at the top with the selected observations highlighted.
+The SHAP values of the full dataset are shown as horizontal parallel coordinate plots at the top with the selected observations highlighted.
 In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
-In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position of is fixed to the observations prediction and residual respectively.
-The current projection basis is depicted on the bottom, where the width of the bar is mapped to variables contribution to the horizontal axis.
+In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position is fixed to the observations prediction and residual respectively.
+The current projection basis is depicted on the bottom, where the width of a bar is mapped to the variable's contribution to the horizontal axis.
 The starting frame is the normalized SHAP values of the primary observation.
@@
 
-In the regression case we compare an observation with a particularly extreme residual as compared with a neighboring point that is more accurately predicted.
+In the regression case, we compare an observation with a particularly extreme residual as compared with a neighboring point that is more accurately predicted.
 The global tour is primarily the same without the misclassified circles, but the tooltip now shows the observation's residual.
-We simulate data, 5 variable from the uniform distribution (between 0 and 5) and \(y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon\).
-We fit a random forest and extract the all tree SHAP values.
+We simulate data, five variables from the uniform distribution in the support {[}0, 5{]} and \(y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon\).
+We fit a random forest and extract the tree SHAP values from all observations.
 The first thing we notice from the global view is the shape difference between the data- and attribution-spaces, which are primarily circular are triangular respectively.
@@
 Though linked brushing we find that the attribution space is highly correlated with the \(y\) value.
-The highest values occupy 1 corner of the triangle, the lowest values cluster on the opposite side and the middle values form a `)' shape filling out the triangle in attribution space.
-In the regression case the tour view is a bit different, this is illustrated in figure \ref{cheemtourreg}.
-The the current frames basis and distribution are shown to the left.
+The highest values occupy one corner of the triangle, the lowest values cluster on the opposite side, and the middle values form a `)' shape filling out the triangle in attribution space.
+In the regression case, the tour view is a bit different, this is illustrated in figure \ref{fig:cheemtourreg}.
+The current frames basis and distribution are shown to the left.
 The horizontal positions of the two scatter plots are the resulting 1D projection of the basis while the heights are fixed to the observed y and the residuals.
@@
 The global view shows all observations in attribution space requiring that we must extract the variable weightings from \emph{all} observations rather than just one.
-    \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components.
+    <!-- \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components. -->
 \end{itemize}
@@
 The time to preprocess the data will vary significantly with the choice of model and local explanation.
-However, for reference, the FIFA data, 5000 observations of 9 explanatory variables, took 0.6 seconds to create PCA for both the data and attribution spaces.
-On the same data, with modest hyper parameters a random forest model fit in 2.9 seconds while extracting the tree SHAP values of each observation took 254 seconds combined.
+For reference, the FIFA data, 5000 observations of 9 explanatory variables, took 2.9 seconds to fit a random forest models of modest hyperparmeters.
+Extracting the tree SHAP values of each observation took 254 seconds combined.
+PCA and statistics of the variables and attribution took 0.6 seconds.
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the time will be spent on the local attribution.
-This makes tree SHAP a good candidate to start with.
-The package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes that are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+This makes tree SHAP, with its reduced comupational complexity a good candidate to start with.
+Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 
-The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX site github?
-cran?}.
+The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX include cran URL after published}.
 Preprocessing was facilitated with models created via \textbf{randomForest} {[}liaw\_classification\_2002{]}, and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
@@
 Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
-\textbf{DALEX} (Biecek 2018) and
-the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
+\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
 
@@
 \FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
-\DocumentationTok{\#\# Try the app yourself:}
-\NormalTok{cheem}\SpecialCharTok{::}\FunctionTok{run\_app}\NormalTok{()}
-\DocumentationTok{\#\# Procss your own data for the app by following the examples in:}
-\NormalTok{?cheem}\SpecialCharTok{::}\NormalTok{cheem\_ls}
+\FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
+\DocumentationTok{\#\# Try the app:}
+\FunctionTok{run\_app}\NormalTok{()}
+\DocumentationTok{\#\# Bring your own data: follow the examples in cheem\_ls()}
+\NormalTok{?cheem\_ls}
 \end{Highlighting}
@@
 
-\hypertarget{penguins-species-classification}{%
-\subsection{1) penguins, species classification}\label{penguins-species-classification}}
+\hypertarget{penguin-species-classification}{%
+\subsection{1) Penguin, species classification}\label{penguin-species-classification}}
 
-Palmer penguins data Horst, Hill, and Gorman (2020) consist of 330 observations across four physical measurements of 3 species of penguins foraging near Palmer Station, Antarctica.
+Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) consist of 330 observations across four physical measurements of three species of penguins foraging near Palmer Station, Antarctica.
 A random forest model was fit, classifying the species of the penguin given the physical measurements.
@@
 A random forest model is fit classifying the type of chocolate.
-We'll select chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85\% cocoa.
+We select a chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85\% cocoa.
 
@@
 
-The 2020 season FIFABiecek (2018), contains many skill measurements of soccer/football players along with wage information.
+The 2020 season FIFA (Leone 2020; Biecek 2018), contains many skill measurements of soccer/football players along with wage information.
 After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
@@
 Messi) with that of a leading defensive fielder (V.
-van Dijk), the same observations used in figure @figure(fig:shapdistrbd).
+van Dijk), the same observations used in figure \ref{fig:shapdistrbd}.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\^{}2 lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages.
 This research was supported in part by Australian government Research Training Program (RTP) scholarships.
@@
 The namesake, Cheem, refers to a fictional race of humanoid trees in Doctor Who lore.
-Given that \textbf{DALEX} pulls on from that universe and we originally apply tree SHAP explanations (specific to tree-based models) we found it fitting.
+\textbf{DALEX} pulls on from that universe and we originally apply tree SHAP explanations which are specific to tree-based models.
 

### 2021-12-04 figures/_make_figures.r

@@ -171,9 +171,8 @@ dist_df$variable <- factor(dist_df$variable, levels = rev(.lvl_ord))
     scale_color_brewer(palette = "Dark2") +
     scale_fill_brewer(palette = "Dark2") +
     labs(title="SHAP distribution",
-         y = "variable", x = "Normalized SHAP values\n the median of the contributions while permuting X's") +
-    theme(legend.position = "off")
-)
+         y = "Variable", x = "Normalized SHAP values\n the median of 25 permutations of the explanatory variables") +
+    theme(legend.position = "off"))
 
 ## Breakdowns & plot ----
 ## Messi Breakdown
@@ -202,7 +201,7 @@ bd_df <- bd_df[is.na(bd_df$variable) == FALSE, ]
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
     labs(title="Breakdown plot",
-         y = "variable", x = "Normalized contribution to prediction | variable order") +
+         y = "Variable", x = "Normalized contribution to prediction | variable order") +
     theme(legend.margin = margin(0,0,0,0),
           legend.position = "bottom",
           axis.text.x = element_blank(),
@@ -217,7 +216,7 @@ wages_df <- tibble::tibble(
     geom_segment(size=3L) +
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
-    labs(y = "player", x = "Wages [2020 Euros]") +
+    labs(y = "Player", x = "Wages [2020 Euros]") +
     theme(legend.position = "off"))
 ### Plot together
 require("patchwork")
@@ -247,7 +246,7 @@ ggplot2::ggsave(
   gg <- global_view(layer_ls, shap_obs, comp_obs,
                     color = "attr_proj.y_cor", as_ggplot = TRUE)
   
-  p## Save -----
+  ## Save -----
   ggplot2::ggsave(
     "./figures_from_script/ch5_fig2_global_space.png",
     plot = gg + theme(aspect.ratio = 1), device = "png",
@@ -263,7 +262,7 @@ ggplot2::ggsave(
   htmlwidgets::saveWidget(ggp, "./figures_from_script/ch5_fig2_global_space.html",
                           selfcontained = TRUE)
   
-  # ch5_fig3_cheem_initial_bas -----
+# ch5_fig3_cheem_initial_bas -----
   .df <- layer_ls$shap_df
   bas <- .df[shap_obs, -ncol(.df)] %>%
     as.matrix(nrow = 1L) %>% t() %>%

### 2021-12-10 _paper.rmd
- restructure paper into classification/regression - new screencaptures and remove old ones - a lot of grammarly edits - abstract causing error

@@
 title: "Interrogating the linear variable importance of local explanations of black-box models with animated linear projections"
-# author:
+shorttitle: Cheem
+#author:
 #  - name: Nicholas Spyrison
@@
 # keywords:
-#   formatted: [XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, DR, dimension reduction, R]
-#   plain:     [XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, DR, dimension reduction, R]
-bibliography: cheem_paper.bib
-abstract: |
-  Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
+#   formatted: [XAI, explanable artificial intelligence, black-box models, local # explanations, grand tour, manual tour, DR, dimension reduction, R]
+#   plain:     [XAI, explanable artificial intelligence, black-box models, local # explanations, grand tour, manual tour, DR, dimension reduction, R]
+#abstract: |
+#  Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
 In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
@@
 Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
-The implementation of our framework is available as an __R__ package __cheem__ available at [github.com/nspyrison/cheem](https://github.com/nspyrison/cheem).
+The implementation of our framework is available as an __R__ package __cheem__ available on CRAN.
 # documentclass: jdssv
-# classoption: article
+classoption: article
 output:
@@
     fig_caption: yes
-appendix:
+bibliography: cheem_paper.bib
+#appendix:
 #   - "appendix-1.Rmd"
@@
 header-includes:
- - \usepackage{setspace} ### for title page spacing
- - \usepackage{hyperref} ### for all sorts of linking
- - \usepackage{graphicx} ### to insert Monash crest
+ - \usepackage{setspace} ## for title page spacing
+ - \usepackage{hyperref} ## for all sorts of linking
 ---
@@
   fig.align  = "center",
-  echo = FALSE, ## Code
+  echo       = FALSE,
   collapse   = TRUE,
@@
 # Introduction {#sec:intro}
-<!-- WHAT TOPICS, and ISSUES addressed (motivation) -->
 
-<!-- ## Predictive models -->
-<!-- history of regression and classification -->
-Mathematically rigorous approaches to predictive modeling are attributed to the method of least squares, over two centuries ago by Legendre and Gauss in 1805 and 1809 respectively.
+<!-- History of regression and classification -->
+Mathematically rigorous approaches to predictive modeling are attributed to the least-squares method, over two centuries ago by Legendre and Gauss in 1805 and 1809, respectively.
 In 1886 Francis Galton coined the term _regression_ to refer to continuous, quantitative predictions.
@@
 
-<!-- introduce explanatory vs predictive modeling -->
-Breiman and Shmueli [@breiman_statistical_2001; @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling is performed to predict new or future out-of-sample observations.
+<!-- Introduce explanatory vs predictive modeling -->
+Breiman and Shmueli [@breiman_statistical_2001; @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling predicts new, out-of-sample, observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
-In explanatory modeling, the interpretable is a key feature for drawing inferential conclusions.
+In explanatory modeling, interpretability is key for drawing inferential conclusions.
 While predictive modeling may opt for potentially more accurate black-box models.
-The intended use of a model has important implications for which methods are used and the development of those models.
+The intended use has important implications for model selection and development.
 
-<!-- ## XAI & interpretability crisis -->
+<!-- XAI & interpretability crisis -->
 Black-box models are becoming increasingly common, but not without their share of controversy and issues [@oneil_weapons_2016; @kodiyan_overview_2019].
-Applications have been known to reflect common biases against sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable which is then built into the model.
-Another issue is that of data drift when new data is outside the support of latent or exogenous explanatory variables.
+Black-box models have been known to reflect common biases, including sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
+Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable, which is then built into the model.
+Another issue is data drift when new data is outside the support of latent or exogenous explanatory variables.
 Data drift can lead to worse predictions [@lazer_parable_2014; @salzberg_why_2014].
-Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI [@adadi_peeking_2018; @arrieta_explainable_2020].
+Such cases highlight the need to make models fair, accountable, ethical, and transparent, which has led to the movement of XAI [@adadi_peeking_2018; @arrieta_explainable_2020].
 
-<!-- ## Local explanations -->
+<!-- Local explanations -->
 One branch of XAI is local explanations, which take a variable attribution approach to bring transparency to a model.
 Local explanations attempt to approximate linear variable importance at the location of one observation.
-There are many such local explanations, any of which is works with our approach (assuming model-explanation compatibility).
+There are many such local explanations.
 
 <!-- SHAP -->
-To illustrate our work we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010; @strumbelj_explaining_2014].
+To illustrate our work, we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010; @strumbelj_explaining_2014].
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
-To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP [@lundberg_consistent_2018].
+To be exact, we apply a variant that enjoys a lower computational complexity, known as tree SHAP [@lundberg_consistent_2018].
 <!-- 2nd half is more application details; move lower? -->
@@
 <!-- ## Data visualization tours -->
-In multivariate data visualization a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_review_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_review_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
 Tours are viewed as an animation over small changes to the projection basis.
-Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
 The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
 
-<!-- ## manual tours --> 
-There are various types of tours, which are distinguished by the sequence of projection bases.
+<!-- manual tours --> 
+There are various types of tours, which are distinguished by the generation of projection bases.
 In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
@@
 <!-- Purposed approach -->
-The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is a good explanation for the model predictions.
-We provide a free and open-source R package `cheem` with an interactive application to facilitate analysis.
-By viewing approximations of data- and attribution-space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
+The approach proposed below is to use the radial, manual tour to interrogate a local explanation.
+After the identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
+We provide a free and open-source R package __cheem__ with an interactive application to facilitate analysis.
 We give case studies of toy and modern datasets for both classification and regression tasks.
@@
 <!-- What-if & counterfactual analysis -->
-The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _Ceteris paribus_ [@biecek_ceterisparibus_2020].
-Latin for “other things held constant” or “all else unchanged”, is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020].
+Latin for “other things held constant” or “all else unchanged” is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual observed case by varying the basis, the angle of the data object.
-Another difference is that the basis is restricted on orthonormality.
-That is to say when the contribution of one variable decreases the contributions of others necessarily increases such that there is a full component in that direction.
+In contrast, our approach is a geometric explanation of the factual, observed case by that varies contributions of the basis, essentially the orientation of the data object.
+Another difference is that the basis must maintain orthonormality.
+That is to say, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
 
 <!-- Paper structure -->
-The rest of this paper is organized as follows.
-The next section [Local explanation statistics](#sec:explanations) covers the background of the local explanation SHAP and the traditional visuals produced from it.
-[Tours and the radial tour](#sec:tour) digs deeper into the animation and what it shows.
-The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis, and discusses the backend details of the package and preprocessing.
+The remainder of this paper is organized as follows.
+The next section, [Local explanation statistics](#sec:explanations), covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
+[Tours and the radial tour](#sec:tour) digs deeper into these animations of continuous linear projections.
+The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis, preprocessing, and package infrastructure.
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
@@
 Consider a highly non-linear model.
-At face value it is hard to say which variable(s) are sensitive to crossing of classification boundary or identify which variables caused an observation to have a relatively extreme residual.
+At face value, it is hard to say which variable(s) are sensitive to the crossing of a classification boundary or identifies which variables caused an observation to have a relatively extreme residual.
 Local explanations shed light on these cases by approximating linear variable importance in the vicinity of one observation.
@@
 <!-- Taxonomy of local explanations -->
-Figure 6 of @arrieta_explainable_2020 gives a broad summarization of the taxonomy and literature of explanation techniques.
+Figure 6 of @arrieta_explainable_2020 gives comprehensive summarization of the taxonomy and literature of explanation techniques.
 This includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016; @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
@@
 <!-- Uses of local explanations -->
-These instance level explanations are used in a variety of ways depending on the data.
-In images saliency maps overlay or offset a heatmap indicating which pixels were important [@simonyan_deep_2014].
-For instance, or the presence of snow may be highlighted when distinguishing if a picture of contains a wolf or husky.
-In text analysis, Word-level contextual sentiment analysis can be used to highlight sentiment and magnitude of influential words [@vanni_textual_2018].
-In the case of numeric regression, they can be used to explain variable additive contributions from intercept to prediction [@ribeiro_why_2016].
+These instance-level explanations are used in a variety of ways depending on the data.
+In images, saliency maps overlay or offset a heatmap indicating which pixels were important [@simonyan_deep_2014].
+For instance, the presence of snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
+In the case of numeric regression, they are used to explain variable additive contributions from model intercept to the observation's prediction [@ribeiro_why_2016].
 
@@
 SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score [@shapley_value_1953].
+This idea stems from the field of game theory, where @shapley_value_1953 devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
 
@@
 To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season [@leone_fifa_2020].
-We have 5000 observations of nine skill measures (after aggregating variable with high correlation).
+We have 5000 observations of nine skill measures (after aggregating highly correlated variables).
 A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
@@
 In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the prediction of the defensive player.
-We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
+We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
 Panel c) shows a simplified breakdown plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
+Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
 
 ```{r shapdistrbd, echo=F, fig.cap = "Illustration of the distribution of SHAP attributions and a breakdown plot.
-From FIFA 2020 data, a random forest model regresses wages from 9 skill attributes for a star offensive and defensive player.
-a) The players have very different wages.
-b) Shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
-The median of these distributions are the final SHAP values.
+From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
+The players have very different wages, but b) shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
+The medians of these distributions are the final SHAP values.
 The variable importance differs across the exogenous information of player position.
-These explanations make sense; the variable importances make sense in light of the position of the player.
-c) Breakdown plots of the observations the explanation used to additively explain the difference between the intercept and prediction"}
+These explanations make sense; the variable importances make sense in light of the player's position.
+c) Breakdown plots of the observations the explanation used to additively explain the difference between the model intercept and the observation prediction."}
 knitr::include_graphics("./figures/shap_distr_bd.png")
@@
 
-In summary, this highlights how local explanations bring interpretability to a model at least in the vicinity of their observations.
-In this instance, we showed how two very different soccer players receive different profiles of variable importance to explain the prediction of their wages.
+In summary, this highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this instance, we showed how two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
 In the following section, we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
@@
 Below we illustrate the two primary displays of the application: the global view and the tour view.
-Then we'll cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing that needs to be done before display.
+Then we cover what we take away from the classification and regression tasks.
+Lastly, we discuss the preprocessing that before display.
+
+## Global view
+
+<!-- Purpose -->
+The global view provides an essential context of all observations and allows  exploration of the separability of the data- and attribution-spaces.
+While ultimately, its purpose is to facilitate the selection of a primary point and comparison point.
+
+<!-- Approximations of the spaces, PC1:2 -->
+An approximation of these spaces is given as the first two principal components of their respective spaces.
+This is shown side-by-side next to model information, the prediction, and the observed response variable.
+The orientation and magnitude  of the variables are inscribed on a unit circle.
+While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization real task at hand, the selection of observation, and nearby comparison.
+
+<!-- Predicted by observed -->
+It is insightful to explore these two approximations against a visual of the model; prediction by observation is also displayed.
+Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
+
+## Radial cheem tour
+
+<!-- a function of the obs selected above -->
+The global view facilitated the selection of a primary and optional comparison observation.
+The variable-level attribution of primary observation is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of  the contributions of the linear variables that best explain the difference between the model intercept and an observations prediction, not the local shape of the model surface.
+
+<!-- Tour start frame -->
+The initial frame is the normalized SHAP values of the primary observation.
+The current projection basis is depicted as the width of a bar, the variable's contribution to the horizontal axis.
+The normalized values of all observations are shown as vertical parallel coordinate plots.
+
+<!-- tour animation -->
+The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
+The default variable selected has the largest discrepancy between the primary and comparison observations attribution.
+In the following sections, we elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
+
+<!-- Segue to classification -->
+Now that we have introduced the global view and corresponding cheem radial tour let's discuss the differences between the classification and regression cases.
 
@@
 What information do we glean from using this method on a classification task?
-Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
+Typically we select a misclassified observation compared to a correctly classified point that is nearby in data space.
 We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
 By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
-That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+That is, we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+
+```{r classificationcase, echo=F, out.width = "100%", fig.cap = "Display illustrating the classification case.
+Both views are colored on predicted class and while red circles identify misclassified observations.
+The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
+The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction.
+We probe the support of the variable contributions by selecting a variable to vary the contribution."}
+knitr::include_graphics("./figures/app_classification.PNG")
+```
 
@@
 
-The regression case is not as discrete a feature.
-Instead, we find it more meaningful to compare the observed value or  residual with the projection.
-In this case the horizontal positions are the same, that is projection of the data through the attribution of the explanation and then varying as the radial tour changes the contributions of the selected variable.
-While the height of the points is fixed to the fixed either to the observed y and residuals for the middle and right panels respectively.
+The regression case, we opt to color the global view on a statistic, the observation's residual, the log Mahanabis distance in data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the radial tour, the horizontal positions are the same, the  attribution projection, then varying as the radial tour changes basis.
+While the vertical position is fixed to the observed response variable and residuals for the middle and right panels, respectively.
+This changes the display from 1D density to a scatterplot.
+The basis is still one component, the horizontal position, independent with the vertical position.
 
-## Global view
-
-<!-- purpose -->
-The global view is an important context for exploring the separability of the data- and the local explanation's attribution-spaces, and is crucial in the selection of explanation to further interrogate and explore the structural sensitivity of.
-
-<!-- approximations of the spaces, PC1:2 -->
-We show an approximation of these spaces with a projection through their first two principal components.
-The orientation of the variables is shown inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
-
-<!-- predicted by observed -->
-It is insightful to explore these two approximations against a visual of the model; prediction by observation is added.
-Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
-
-```{r globalview, echo=F, out.width = "100%", fig.cap = "Global view screen capture; the approximations of the data and SHAP-spaces of Palmer penguins data.
-Orientation of the basis contributions is illustrated on a unit circle.
-Linked brushing allows observations brushed in one plot to be selected in others.
-This selection is also used in the proceeding view and their corresponding information is displayed in an interactive table.
-Hovering the cursor over an observation displays a tooltip with row number/name information.
-In the classification case, misclassified points are circled in red.
-This view provides an orientation to select a primary and comparison observation, key targets in the following tour."}
-knitr::include_graphics("./figures/global_view_toy_class.PNG")
+```{r regressioncase, echo=F, out.width = "100%", fig.cap = "Display of the regression case.
+The global view can be colored on the correlation of the attribution projection and observed response.
+In the tour, the horizontal values are the same as the classification case.
+The vertical position is fixed to the observed response and residuals in the middle and right facets, respectively."}
+knitr::include_graphics("./figures/app_regression.PNG")
 ```
-
-<!-- Interaction -->
-This view offers dynamic interaction in several ways.
-A tooltip displays while the cursor hovers over a point displays the row number/name and classification information if appropriate.
-Linked brushing allows for the selection of points (by click and drag) where those points will be highlighted in both plots.
-The information corresponding to the selected points is populated on a sortable table and the data powering the proceeding tour will also subset the data to the current selection.
-
-## Cheem tour
-
-<!-- a function of the obs selected above -->
-The primary observation identified via the global view is foundational to the production of the cheem tour.
-Namely, the linear attribution of that variable is used as a 1D projection basis.
-This is the approximate contribution of the variables that this model uses to justify its prediction for the observation.
-
-<!-- Dupiction of a frame -->
-The SHAP values of the full dataset are shown as horizontal parallel coordinate plots at the top with the selected observations highlighted.
-In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
-In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position is fixed to the observations prediction and residual respectively.
-The current projection basis is depicted on the bottom, where the width of a bar is mapped to the variable's contribution to the horizontal axis.
-The starting frame is the normalized SHAP values of the primary observation.
-In figure \@ref(fig:cheemtourclass) we illustrate the classification case with simulated data; 3 clusters separated in the first 2 dimensions with 2 dimensions of noise.
-
-<!-- cheem tour classification caption -->
-(ref:cheemtourclass) Cheem tour, classification case.
-A random forest model predicting the class of the of simulated data.
-A misclassified observation is be compared with a nearby correctly classified observation.
-The top shows density and rug marks of the current frame with the positions of the primary and comparison observations shown as dashed and dotted lines.
-The bottom shows two key features; the distribution of attribution space shown as parallel coordinate lines across the variables and the bars show the current contribution of the variable to the projection.
-The primary, misclassifed observation (dashed line), is plausibly in the middle of the its prediction,  purple density, the story that the explanation trying to sell us.
-Yet, when we play the tour animating on the contribution of V1, the bottom dashed line is more regularly in the center the it observed cluster, green, By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.
-
-```{r cheemtourclass, echo=F, out.width = "100%", fig.cap = "(ref:cheemtourclass)"}
-knitr::include_graphics("./figures/cheem_tour_toy_class.PNG")
-```
-
-<!-- tour animation -->
-Data visualization tours animate many linear projections over small changes to the basis.
-The manual tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing allows an analyst to test an individual variable's sensitivity to the structure identified in the frame.
-The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution.
-In the following sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.
-
-<!-- regression case -->
-In the regression case, we compare an observation with a particularly extreme residual as compared with a neighboring point that is more  accurately predicted.
-The global tour is primarily the same without the misclassified circles, but the tooltip now shows the observation's residual.
-We simulate data, five variables from the uniform distribution in the support [0, 5] and $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
-We fit a random forest and extract the tree SHAP values from all observations.
-The first thing we notice from the global view is the shape difference between the data- and attribution-spaces, which are primarily circular are triangular respectively.
-We also see that the random forest fit the data quite well viewing the predictions across the observations.
-Though linked brushing we find that the attribution space is highly correlated with the $y$ value.
-The highest values occupy one corner of the triangle, the lowest values cluster on the opposite side, and the middle values form a ')' shape filling out the triangle in attribution space.
-In the regression case, the tour view is a bit different, this is illustrated in figure \@ref(fig:cheemtourreg).
-The current frames basis and distribution are shown to the left.
-The horizontal positions of the two scatter plots are the resulting 1D projection of the basis while the heights are fixed to the observed y and the residuals.
-
-<!-- tour regression caption -->
-(ref:cheemtourreg-cap) Cheem tour, regression case.
-A random forest model predicting the continuous $y$ of the of simulated data where $y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon$.
-The current frames basis and the distribution of all attributions is shown on the left.
-The horizontal positions for the scatterplots are resulting projection through the 1D basis while the vertical heights are fixed to the observed y and residuals respectively.
-
-```{r cheemtourreg, echo=F, out.width = "100%", fig.cap = "(ref:cheemtourreg-cap)"}
-knitr::include_graphics("./figures/cheem_tour_toy_reg.PNG")
-```
-
-Now that we have covered the classification and regression cases we will discuss the preprocessing, package infrastructure, and interactive features before proceeding to case studies with data from the wild.
 
@@
 
-The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
-Linked brushing also for a rectangular selection of observations.
-These points are highlighted within the other space of the global view _and_ crucially available for downstream consumption, for instance, highlighting in the tour.
-The distinction between the former and latter is subtle, but important for developers to grasp.
-In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
-In the latter, the use of `plotly::event_data()` returns the identity of the selected observations.
-This will require a reactive flush, but after that, the sky is the ceiling.
+<!-- reactive vs exploration interactions -->
+The application has several reactive selections that affect the data, coloring, and inputs for the tour to improve flexibility by extending the use cases.
+It also has interactive features in the global view that aid the analysis.
+
+<!-- Interaction except -->
+A tooltip displays while the cursor hovers over a point displays the observation number/name and classification information if appropriate.
+Linked brushing allows for the selection of points (with left click and drag) where those points will be highlighted in both plots.
+The information corresponding to the selected points is populated on a dynamic table.
+These interactions aid exploration of the spaces and ultimately the selection of the primary and comparison observations.
+
+```{r interactions, echo=F, out.width = "100%", fig.cap = "Illustration of data explorations interactions in the global view.
+This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
+Tooltips display when hovering over a point"}
+knitr::include_graphics("./figures/app_interactions.PNG")
+```
 
@@
 
-The benefit of having dynamic interaction with data is predicated on a reasonably small render time.
-It is important to preprocess as much work as possible so that application resources can be used efficiently.
-The work remaining at runtime should be solely responding to inputs and the rendering of figures and tables.
+The benefit of having dynamic interaction with data is predicated on a reasonably short render time.
+It is important to preprocess expensive operations so the application resources can be used efficiently.
+The work remaining at runtime is solely responding to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
 
-(ref:citeRf) [@liaw_classification_2002]
-(ref:citeTs) [@kominsarczyk_treeshap_2021]
+(ref:citeRf) @liaw_classification_2002
+(ref:citeTs) @kominsarczyk_treeshap_2021
 \begin{itemize}
-	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic variable (color/shape)  can be mapped typically a categorical variable.
+	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable.
+An optional categorical variable can be mapped to color and shape of obervations.
+Explanatory variables are scaled in visualization after modeling or creating local explanations.
 	\item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} (ref:citeRf) to mitigate the runtime of our local explanation which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)] for compatability with the local explanation which requires tree-based models.
 	\item \textbf{Local explanation:} any model-compatible linear explanation could be used.
-We apply tree SHAP, a more computationally efficient variant of SHAP applicable to tree-based models.
-This is done with the package \textbf{treeshap} (ref:citeTs), hosted on GitHub only].
-The global view shows all observations in attribution space requiring that we must extract the variable weightings from \emph{all} observations rather than just one.
-	<!-- \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components. -->
+We apply tree SHAP, a more computationally efficient variant of SHAP, that is applicable to tree-based models.
+This is done with the package \textbf{treeshap} [(ref:citeTs), hosted on GitHub only].
+The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
 \end{itemize}
@@
 <!-- note on time of execution -->
-The time to preprocess the data will vary significantly with the choice of model and local explanation.
-For reference, the FIFA data, 5000 observations of 9 explanatory variables, took  2.9 seconds to fit a random forest models of modest hyperparmeters.
+The time to preprocess the data will vary significantly with the model and local explanation.
+For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.9 seconds to fit a random forest model of modest hyperparameters.
 Extracting the tree SHAP values of each observation took 254 seconds combined.
-PCA and statistics of the variables and attribution took 0.6 seconds.
+PCA and statistics of the variables and attributions took 0.6 seconds.
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the time will be spent on the local attribution.
-This makes tree SHAP, with its reduced comupational complexity a good candidate to start with.
-Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+Increased model complexity or data dimensionality will quickly become an obstacle.
+This makes tree SHAP, with its reduced computational complexity, a good candidate to start with.
+Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 Preprocessing was facilitated with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
-The application is made with __shiny__ [@chang_shiny_2021].
+The application was made with __shiny__ [@chang_shiny_2021].
 The tour visual is an extension of __spinifex__ [@spyrison_spinifex_2020].
@@
 ```{r eval=FALSE, echo=TRUE}
-## Download the package:
+## Download the package
 install.packages("cheem", dependencies = TRUE)
+## Restart the R session to pick up the file structure may be needed
+restartSession()
+## Load cheem into session
 library("cheem")
-## Try the app:
+## Try the app
 run_app()
@@
 
-<!-- Provided with tidy and transformed data, the function `cheem_ls` is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations.
-It returns a list containing data frames to be consumed by the application, higher-level model performance, and runtime information.
-The return of this function can be then be saved to an .rds file for consumption in the application.
--->
-
 # Case studies {#sec:casestudies}
 
-To illustrate the use of the cheem method we apply it to modern datasets, two classification examples and then two of regression.
+To illustrate the use of the cheem method, we apply it to modern datasets, two classification examples and then two of regression.
 
@@
 
-In figure \@ref(fig:casepenguins) a misclassified point is contrasted with a correctly classified point of its observed class that is nearby in data-space.
-The attribution space from the treeshap local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution while primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
+In figure \@ref(fig:casepenguins), a misclassified point is contrasted with a correctly classified point of its observed class nearby in data-space.
+The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
+The primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
 The tour varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to linear explanation this observation be misclassified.
+Downplaying the contribution of bill length is crucial to the linear explanation of this observation be misclassified.
 
@@
 Each of which was labeled as being either milk or dark chocolates.
-With this data, we can see if a manufacturer is giving an accurate portrayal of the chocolate.
+With this data, we can see if a manufacturer gives an accurate portrayal of the chocolate.
 We are curious to see if there are chocolates that nutritionally look like milk chocolates that are labeled as dark chocolates, which may hold a higher market value.
@@
 A random forest model is fit classifying the type of chocolate.
-We select a chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85% cocoa.
+We selected a chocolate labeled dark, through predicted to be milk chocolate with a comparison with chocolate labeled 85% cocoa.
 
-```{r casechocolates, echo=F, out.width = "100%", fig.cap = "Chocolates data  type classification (milk or dark). "}
+```{r casechocolates, echo=F, out.width = "100%", fig.cap = "Chocolates data type classification (milk or dark)."}
 knitr::include_graphics("./figures/case_chocolates.png")
@@
 
-From figure \@ref(fig:casechocolates) we similarly see that attribution space that is more separable relative to data-space.
-Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while there 8 of the inverse case.
-Calories from fat is the measure with the largest difference in treeshap attribution between these points.
+From figure \@ref(fig:casechocolates) we similarly see that attribution space is more separable relative to data-space.
+Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while 8 of the inverse case.
+Calories from fat is the variable with the largest difference in treeshap attribution between these points.
 
@@
 
-The 2020 season FIFA [@leone_fifa_2020; @biecek_dalex_2018], contains many skill measurements of soccer/football players along with wage information.
+The 2020 season FIFA [@leone_fifa_2020; @biecek_dalex_2018], contains many skill measurements of soccer/football players and wage information.
 After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
-The model was fit with 5000 observations of the 9 skill aggregates before being thinned to 500 players to gate occlusion and render time.
+The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
 We compare a leading offensive fielder (L.
-Messi) with that of a leading defensive fielder (V.
-van Dijk), the same observations used in figure \@ref(fig:shapdistrbd).
+Messi) with that of a top defensive fielder (V.
+van Dijk), the same observations were used in figure \@ref(fig:shapdistrbd).
 
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
-The primary observation is star offensive player (L.
-Messi) as compared with a top defensive player (V.
+The primary observation is a star offensive player (L.
+Messi) compared with a top defensive player (V.
 van Dijk)."}
@@
 
-With figure \@ref(fig:casefifa) we are going to test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense has almost singular importance for the explanation of the offensive player.
+With figure \@ref(fig:casefifa) we will test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, then offense skills has almost singular importance for the explanation of the offensive player.
 We vary the contribution of offensive skills.
 In the tour (3rd frame of b), offensive skills moved, and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, which does maintain some separability.
+We also notice that accuracy has rotated into the frame, maintaining some separability.
 
-## 4) Ames housing prices 2018
+## 4) Ames housing 2018, sales price regression
 
 Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across 9 variables.
-Using interaction from the global view we select a house with an extreme negative residual and an accurate observation close to it in the data.
+The remaining are 338 house sales across nine variables.
+Using interaction from the global view, we select a house with an extreme negative residual and an accurate observation close to it in the data.
 
@@
 Figure \@ref(fig:caseames) shows the global view and extrema of the tour.
-The horizontal distance in the tour didn't show a large disparity between our selected points.
-This is not particularly surprising as most of the features have a sizable contribution.
-Rotating any one variable out of the frame will rotate other important variables into frame preserving most of the distance from intercept to prediction.
+The horizontal distance in the tour didn't show a significant disparity between our selected points.
+This is not particularly surprising as most variables have a sizable contribution.
+Rotating any one variable out of the frame will rotate other vital variables into the frame, preserving most of the distance from intercept to prediction.
 However, the tour has revealed an interesting feature worth discussing.
-notice that the observations pivot about the origin, that is the basis roughly halfway between bases in frame 1 and 2 of panel b) the data is near a singular profile.
-This means that there is a basis orthogonal to this point that describes a lot of the variation.
-Knowing these singular frames can help direct to bases that have meaningful variation in the data.
+Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames 1 and 2 of panel b) the data is near a singular profile.
+This means that there is a basis orthogonal to this point that describes sizable variation.
+Knowing these singular bases can point toward others that have meaningful variation in the data.
 
@@
 
-The need to maintain the interpretability of black-box models is clear.
-One way to do this is by the use of local explanations of an observation.
+The need to maintain the interpretability of black-box models is evident.
+One aspect uses local explanations of the model in the vicinity of an observation.
 Local explanations approximate the linear variable importance to the model.
-Our idea is to visualize approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
-Then use the linear importance from the primary observation as a projection basis and explore a single variable's importance to the structure identified with the use of the manual tour.
+Our contribution is to assess the explanations by examining the support of explanation, varying the contribution with a radial tour.
+First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
+Then normalized the linear importance from the explanation of the primary observation to use as the initial basis of a manual tour.
+The variable sensitivity to the structure identified in the explanation is explored with the tours varying basis.
 
-We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generalized to any compatible model-explanation pair.
-We apply this to both the classification and regression tasks.
-In the former, it can be used to see which variables cause a misclassification as benchmarked against a nearby observation and explore changing the contribution of variables.
-In the regression case, we compare an observation with particularly extreme residual can be compared with nearby points more accurately predicted or another suitable benchmark.
-
-We have created an open-source __R__ package __cheem__, available on CRAN, to facilitate this workflow including a dynamic application, with the ability to upload preprocessed data.
+We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generally used with any compatible model-explanation pairing.
+We apply it to the classification and regression tasks.
+We have created an open-source __R__ package __cheem__, available on CRAN, to facilitate preprocessing and exploration with the described interactive application.
+Toy and real data are provided or upload your data after preprocessing.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Pyhton__ packages.
-This research was supported in part by Australian government Research Training Program (RTP) scholarships.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages.
+This research was supported by Australian government Research Training Program (RTP) scholarships.
 
-The namesake, Cheem, refers to a fictional race of humanoid trees in  Doctor Who lore.
-__DALEX__ pulls on from that universe and we originally apply tree SHAP explanations which are specific to tree-based models.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations that are specific to tree-based models.
 

### 2021-12-10 _paper.tex
- restructure paper into classification/regression - new screencaptures and remove old ones - a lot of grammarly edits - abstract causing error

@@
 \documentclass[
-]{article}
+  article]{article}
 \usepackage{amsmath,amssymb}
@@
 \usepackage{hyperref}
-\usepackage{graphicx}
 \usepackage{booktabs}
@@
 \maketitle
-\begin{abstract}
-Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
-XAI attempts to shed light on these models, one such approach is the use of local explanations.
-A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
-We extract explanations for each observation, and approximate data and this attribution space side-by-side with linked brushing.
-After identifying an observation of interest its local explanation is used as a 1D projection basis.
-We then manipulate the magnitude of the variable contributions with a technique called the tour.
-This tour animates many projections over small changes in the projection basis.
-Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
-The implementation of our framework is available as an \textbf{R} package \textbf{cheem} available at \href{https://github.com/nspyrison/cheem}{github.com/nspyrison/cheem}.
-\end{abstract}
 
@@
 
-Mathematically rigorous approaches to predictive modeling are attributed to the method of least squares, over two centuries ago by Legendre and Gauss in 1805 and 1809 respectively.
+Mathematically rigorous approaches to predictive modeling are attributed to the least-squares method, over two centuries ago by Legendre and Gauss in 1805 and 1809, respectively.
 In 1886 Francis Galton coined the term \emph{regression} to refer to continuous, quantitative predictions.
@@
 
-Breiman and Shmueli (Breiman 2001; Shmueli 2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling is performed to predict new or future out-of-sample observations.
+Breiman and Shmueli (Breiman 2001; Shmueli 2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling predicts new, out-of-sample, observations.
 This distinction draws attention to the divide between interpretable models and black-box models.
-In explanatory modeling, the interpretable is a key feature for drawing inferential conclusions.
+In explanatory modeling, interpretability is key for drawing inferential conclusions.
 While predictive modeling may opt for potentially more accurate black-box models.
-The intended use of a model has important implications for which methods are used and the development of those models.
+The intended use has important implications for model selection and development.
 
 Black-box models are becoming increasingly common, but not without their share of controversy and issues (O'neil 2016; Kodiyan 2019).
-Applications have been known to reflect common biases against sex (Dastin 2018; Duffy 2019), race (Larson et al.
+Black-box models have been known to reflect common biases, including sex (Dastin 2018; Duffy 2019), race (Larson et al.
 2016), and age (Díaz et al.
 2018).
-Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable which is then built into the model.
-Another issue is that of data drift when new data is outside the support of latent or exogenous explanatory variables.
+Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable, which is then built into the model.
+Another issue is data drift when new data is outside the support of latent or exogenous explanatory variables.
 Data drift can lead to worse predictions (Lazer et al.
 2014; Salzberg 2014).
-Such issues highlight the need to make models fair, accountable, ethical, and transparent which has led to the movement of XAI (Adadi and Berrada 2018; Arrieta et al.
+Such cases highlight the need to make models fair, accountable, ethical, and transparent, which has led to the movement of XAI (Adadi and Berrada 2018; Arrieta et al.
 2020).
@@
 Local explanations attempt to approximate linear variable importance at the location of one observation.
-There are many such local explanations, any of which is works with our approach (assuming model-explanation compatibility).
+There are many such local explanations.
 
-To illustrate our work we apply the model-agnostic explanation SHAP (Strumbelj and Kononenko 2010; Štrumbelj and Kononenko 2014).
+To illustrate our work, we apply the model-agnostic explanation SHAP (Strumbelj and Kononenko 2010; Štrumbelj and Kononenko 2014).
 The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
-To be exact we apply a variant that enjoys a lower computational complexity, known as tree SHAP (S.
+To be exact, we apply a variant that enjoys a lower computational complexity, known as tree SHAP (S.
 M.
@@
 
-In multivariate data visualization a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
+In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
 Lee et al.
@@
 Tours are viewed as an animation over small changes to the projection basis.
-Structure in a projection can then be explored visually to see which variables contribute to the formation of the structure.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
 The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
 
-There are various types of tours, which are distinguished by the sequence of projection bases.
+There are various types of tours, which are distinguished by the generation of projection bases.
 In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020) this path is defined by changing the contribution of a selected variable.
@@
 
-The approach purposed below is to use the manual tour as means to interrogate a local explanation; a means of evaluating if its variable importance is a good explanation for the model predictions.
-We provide a free and open-source R package \texttt{cheem} with an interactive application to facilitate analysis.
-By viewing approximations of data- and attribution-space side-by-side, with linked brushing an analyst can identify observations of interest whose explanations are then rendered at the initial projection basis and explored with a manual tour to further interpret the variable importance of the local explanation.
+The approach proposed below is to use the radial, manual tour to interrogate a local explanation.
+After the identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
+We provide a free and open-source R package \textbf{cheem} with an interactive application to facilitate analysis.
 We give case studies of toy and modern datasets for both classification and regression tasks.
 
-The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{Ceteris paribus} (Biecek 2020).
-Latin for ``other things held constant'' or ``all else unchanged,'' is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
+Latin for ``other things held constant'' or ``all else unchanged'' is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual observed case by varying the basis, the angle of the data object.
-Another difference is that the basis is restricted on orthonormality.
-That is to say when the contribution of one variable decreases the contributions of others necessarily increases such that there is a full component in that direction.
+In contrast, our approach is a geometric explanation of the factual, observed case by that varies contributions of the basis, essentially the orientation of the data object.
+Another difference is that the basis must maintain orthonormality.
+That is to say, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
 
-The rest of this paper is organized as follows.
-The next section \protect\hyperlink{sec:explanations}{Local explanation statistics} covers the background of the local explanation SHAP and the traditional visuals produced from it.
-\protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into the animation and what it shows.
-The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis, and discusses the backend details of the package and preprocessing.
+The remainder of this paper is organized as follows.
+The next section, \protect\hyperlink{sec:explanations}{Local explanation statistics}, covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
+\protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into these animations of continuous linear projections.
+The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis, preprocessing, and package infrastructure.
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
@@
 Figure 6 of Arrieta et al.
-(2020) gives a broad summarization of the taxonomy and literature of explanation techniques.
+(2020) gives comprehensive summarization of the taxonomy and literature of explanation techniques.
 This includes a large number of model-specific explanations such as deepLIFT, (Shrikumar et al.
@@
 
-These instance level explanations are used in a variety of ways depending on the data.
-In images saliency maps overlay or offset a heatmap indicating which pixels were important (Simonyan, Vedaldi, and Zisserman 2014).
-For instance, or the presence of snow may be highlighted when distinguishing if a picture of contains a wolf or husky.
-In text analysis, Word-level contextual sentiment analysis can be used to highlight sentiment and magnitude of influential words (Vanni et al.
+These instance-level explanations are used in a variety of ways depending on the data.
+In images, saliency maps overlay or offset a heatmap indicating which pixels were important (Simonyan, Vedaldi, and Zisserman 2014).
+For instance, the presence of snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
 2018).
-In the case of numeric regression, they can be used to explain variable additive contributions from intercept to prediction (Ribeiro, Singh, and Guestrin 2016).
+In the case of numeric regression, they are used to explain variable additive contributions from model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
 
@@
 SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory where Shapley devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score (Shapley 1953).
+This idea stems from the field of game theory, where Shapley (1953) devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
 
 To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season (Leone 2020).
-We have 5000 observations of nine skill measures (after aggregating variable with high correlation).
+We have 5000 observations of nine skill measures (after aggregating highly correlated variables).
 A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
@@
 In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the prediction of the defensive player.
-We would likewise expect the profile of variable importance to be unique for star players of other positions as well, such as goalkeepers or middle fielders.
+We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
 Panel c) shows a simplified breakdown plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry with respect to the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
+Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
 
@@
 \caption{Illustration of the distribution of SHAP attributions and a breakdown plot.
-From FIFA 2020 data, a random forest model regresses wages from 9 skill attributes for a star offensive and defensive player.
-a) The players have very different wages.
-b) Shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
-The median of these distributions are the final SHAP values.
+From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
+The players have very different wages, but b) shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
+The medians of these distributions are the final SHAP values.
 The variable importance differs across the exogenous information of player position.
-These explanations make sense; the variable importances make sense in light of the position of the player.
-c) Breakdown plots of the observations the explanation used to additively explain the difference between the intercept and prediction}\label{fig:shapdistrbd}
+These explanations make sense; the variable importances make sense in light of the player's position.
+c) Breakdown plots of the observations the explanation used to additively explain the difference between the model intercept and the observation prediction.}\label{fig:shapdistrbd}
 \end{figure}
 
-In summary, this highlights how local explanations bring interpretability to a model at least in the vicinity of their observations.
-In this instance, we showed how two very different soccer players receive different profiles of variable importance to explain the prediction of their wages.
+In summary, this highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this instance, we showed how two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
 In the following section, we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
@@
 Below we illustrate the two primary displays of the application: the global view and the tour view.
-Then we'll cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing that needs to be done before display.
+Then we cover what we take away from the classification and regression tasks.
+Lastly, we discuss the preprocessing that before display.
+
+\hypertarget{global-view}{%
+\subsection{Global view}\label{global-view}}
+
+The global view provides an essential context of all observations and allows exploration of the separability of the data- and attribution-spaces.
+While ultimately, its purpose is to facilitate the selection of a primary point and comparison point.
+
+An approximation of these spaces is given as the first two principal components of their respective spaces.
+This is shown side-by-side next to model information, the prediction, and the observed response variable.
+The orientation and magnitude of the variables are inscribed on a unit circle.
+While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization real task at hand, the selection of observation, and nearby comparison.
+
+It is insightful to explore these two approximations against a visual of the model; prediction by observation is also displayed.
+Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
+
+\hypertarget{radial-cheem-tour}{%
+\subsection{Radial cheem tour}\label{radial-cheem-tour}}
+
+The global view facilitated the selection of a primary and optional comparison observation.
+The variable-level attribution of primary observation is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observations prediction, not the local shape of the model surface.
+
+The initial frame is the normalized SHAP values of the primary observation.
+The current projection basis is depicted as the width of a bar, the variable's contribution to the horizontal axis.
+The normalized values of all observations are shown as vertical parallel coordinate plots.
+
+The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
+The default variable selected has the largest discrepancy between the primary and comparison observations attribution.
+In the following sections, we elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
+
+Now that we have introduced the global view and corresponding cheem radial tour let's discuss the differences between the classification and regression cases.
 
@@
 What information do we glean from using this method on a classification task?
-Typically we select a misclassified observation in comparison with a correctly classified point that is nearby in data space.
+Typically we select a misclassified observation compared to a correctly classified point that is nearby in data space.
 We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
 By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
-That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+That is, we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}
+
+}
+
+\caption{Display illustrating the classification case.
+Both views are colored on predicted class and while red circles identify misclassified observations.
+The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
+The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction.
+We probe the support of the variable contributions by selecting a variable to vary the contribution.}\label{fig:classificationcase}
+\end{figure}
 
@@
 
-The regression case is not as discrete a feature.
-Instead, we find it more meaningful to compare the observed value or residual with the projection.
-In this case the horizontal positions are the same, that is projection of the data through the attribution of the explanation and then varying as the radial tour changes the contributions of the selected variable.
-While the height of the points is fixed to the fixed either to the observed y and residuals for the middle and right panels respectively.
-
-\hypertarget{global-view}{%
-\subsection{Global view}\label{global-view}}
-
-The global view is an important context for exploring the separability of the data- and the local explanation's attribution-spaces, and is crucial in the selection of explanation to further interrogate and explore the structural sensitivity of.
-
-We show an approximation of these spaces with a projection through their first two principal components.
-The orientation of the variables is shown inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it provides a reasonable starting point for the real task at hand, the selection of observation and nearby comparison.
-
-It is insightful to explore these two approximations against a visual of the model; prediction by observation is added.
-Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
+The regression case, we opt to color the global view on a statistic, the observation's residual, the log Mahanabis distance in data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the radial tour, the horizontal positions are the same, the attribution projection, then varying as the radial tour changes basis.
+While the vertical position is fixed to the observed response variable and residuals for the middle and right panels, respectively.
+This changes the display from 1D density to a scatterplot.
+The basis is still one component, the horizontal position, independent with the vertical position.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/global_view_toy_class}
+{\centering \includegraphics[width=1\linewidth]{./figures/app_regression}
 
@@
 
-\caption{Global view screen capture; the approximations of the data and SHAP-spaces of Palmer penguins data.
-Orientation of the basis contributions is illustrated on a unit circle.
-Linked brushing allows observations brushed in one plot to be selected in others.
-This selection is also used in the proceeding view and their corresponding information is displayed in an interactive table.
-Hovering the cursor over an observation displays a tooltip with row number/name information.
-In the classification case, misclassified points are circled in red.
-This view provides an orientation to select a primary and comparison observation, key targets in the following tour.}\label{fig:globalview}
+\caption{Display of the regression case.
+The global view can be colored on the correlation of the attribution projection and observed response.
+In the tour, the horizontal values are the same as the classification case.
+The vertical position is fixed to the observed response and residuals in the middle and right facets, respectively.}\label{fig:regressioncase}
 \end{figure}
-
-This view offers dynamic interaction in several ways.
-A tooltip displays while the cursor hovers over a point displays the row number/name and classification information if appropriate.
-Linked brushing allows for the selection of points (by click and drag) where those points will be highlighted in both plots.
-The information corresponding to the selected points is populated on a sortable table and the data powering the proceeding tour will also subset the data to the current selection.
-
-\hypertarget{cheem-tour}{%
-\subsection{Cheem tour}\label{cheem-tour}}
-
-The primary observation identified via the global view is foundational to the production of the cheem tour.
-Namely, the linear attribution of that variable is used as a 1D projection basis.
-This is the approximate contribution of the variables that this model uses to justify its prediction for the observation.
-
-The SHAP values of the full dataset are shown as horizontal parallel coordinate plots at the top with the selected observations highlighted.
-In the classification case, 1D density curves with underlying rug marks are drawn and colored according to their predicted classes.
-In the regression case, the horizontal position of the points comes from projection through the 1D attribution basis while the vertical position is fixed to the observations prediction and residual respectively.
-The current projection basis is depicted on the bottom, where the width of a bar is mapped to the variable's contribution to the horizontal axis.
-The starting frame is the normalized SHAP values of the primary observation.
-In figure \ref{fig:cheemtourclass} we illustrate the classification case with simulated data; 3 clusters separated in the first 2 dimensions with 2 dimensions of noise.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_toy_class}
-
-}
-
-\caption{Cheem tour, classification case.
-A random forest model predicting the class of the of simulated data.
-A misclassified observation is be compared with a nearby correctly classified observation.
-The top shows density and rug marks of the current frame with the positions of the primary and comparison observations shown as dashed and dotted lines.
-The bottom shows two key features; the distribution of attribution space shown as parallel coordinate lines across the variables and the bars show the current contribution of the variable to the projection.
-The primary, misclassifed observation (dashed line), is plausibly in the middle of the its prediction, purple density, the story that the explanation trying to sell us.
-Yet, when we play the tour animating on the contribution of V1, the bottom dashed line is more regularly in the center the it observed cluster, green, By varying this contribution we gain information of how sensitive this variable is to the explanation; the variable importance which led to a misclassification.}\label{fig:cheemtourclass}
-\end{figure}
-
-Data visualization tours animate many linear projections over small changes to the basis.
-The manual tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing allows an analyst to test an individual variable's sensitivity to the structure identified in the frame.
-The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution.
-In the following sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.
-
-In the regression case, we compare an observation with a particularly extreme residual as compared with a neighboring point that is more accurately predicted.
-The global tour is primarily the same without the misclassified circles, but the tooltip now shows the observation's residual.
-We simulate data, five variables from the uniform distribution in the support {[}0, 5{]} and \(y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon\).
-We fit a random forest and extract the tree SHAP values from all observations.
-The first thing we notice from the global view is the shape difference between the data- and attribution-spaces, which are primarily circular are triangular respectively.
-We also see that the random forest fit the data quite well viewing the predictions across the observations.
-Though linked brushing we find that the attribution space is highly correlated with the \(y\) value.
-The highest values occupy one corner of the triangle, the lowest values cluster on the opposite side, and the middle values form a `)' shape filling out the triangle in attribution space.
-In the regression case, the tour view is a bit different, this is illustrated in figure \ref{fig:cheemtourreg}.
-The current frames basis and distribution are shown to the left.
-The horizontal positions of the two scatter plots are the resulting 1D projection of the basis while the heights are fixed to the observed y and the residuals.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/cheem_tour_toy_reg}
-
-}
-
-\caption{Cheem tour, regression case.
-A random forest model predicting the continuous \(y\) of the of simulated data where \(y = x1 + x2 + x1*x2 + (x3 + x4 + x5) / 10 + \epsilon\).
-The current frames basis and the distribution of all attributions is shown on the left.
-The horizontal positions for the scatterplots are resulting projection through the 1D basis while the vertical heights are fixed to the observed y and residuals respectively.}\label{fig:cheemtourreg}
-\end{figure}
-
-Now that we have covered the classification and regression cases we will discuss the preprocessing, package infrastructure, and interactive features before proceeding to case studies with data from the wild.
 
@@
 
-The dynamic interaction with the global view is critical to the selection of the primary and comparison observations.
-Linked brushing also for a rectangular selection of observations.
-These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance, highlighting in the tour.
-The distinction between the former and latter is subtle, but important for developers to grasp.
-In the former, self-contained HTML widgets can highlight directly within javascript without evaluation downstream in other reactive functions.
-In the latter, the use of \texttt{plotly::event\_data()} returns the identity of the selected observations.
-This will require a reactive flush, but after that, the sky is the ceiling.
+The application has several reactive selections that affect the data, coloring, and inputs for the tour to improve flexibility by extending the use cases.
+It also has interactive features in the global view that aid the analysis.
+
+A tooltip displays while the cursor hovers over a point displays the observation number/name and classification information if appropriate.
+Linked brushing allows for the selection of points (with left click and drag) where those points will be highlighted in both plots.
+The information corresponding to the selected points is populated on a dynamic table.
+These interactions aid exploration of the spaces and ultimately the selection of the primary and comparison observations.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/app_interactions}
+
+}
+
+\caption{Illustration of data explorations interactions in the global view.
+This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
+Tooltips display when hovering over a point}\label{fig:interactions}
+\end{figure}
 
@@
 
-The benefit of having dynamic interaction with data is predicated on a reasonably small render time.
-It is important to preprocess as much work as possible so that application resources can be used efficiently.
-The work remaining at runtime should be solely responding to inputs and the rendering of figures and tables.
+The benefit of having dynamic interaction with data is predicated on a reasonably short render time.
+It is important to preprocess expensive operations so the application resources can be used efficiently.
+The work remaining at runtime is solely responding to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 \begin{itemize}
-    \item \textbf{Data:} a complete numerical matrix; explanatory and response variable, an optional aesthetic variable (color/shape)  can be mapped typically a categorical variable.
+    \item \textbf{Data:} a complete numerical matrix; explanatory and response variable.
+An optional categorical variable can be mapped to color and shape of obervations.
+Explanatory variables are scaled in visualization after modeling or creating local explanations.
     \item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} (Liaw and Wiener 2002) to mitigate the runtime of our local explanation which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)] for compatability with the local explanation which requires tree-based models.
     \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
-We apply tree SHAP, a more computationally efficient variant of SHAP applicable to tree-based models.
-This is done with the package \textbf{treeshap} (Kominsarczyk et al.
-2021), hosted on GitHub only].
-The global view shows all observations in attribution space requiring that we must extract the variable weightings from \emph{all} observations rather than just one.
-    <!-- \item \textbf{Global view:} The data- and attribution-spaces are approximated as their the first two principal components. -->
+We apply tree SHAP, a more computationally efficient variant of SHAP, that is applicable to tree-based models.
+This is done with the package \textbf{treeshap} [Kominsarczyk et al.
+(2021), hosted on GitHub only].
+The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
 \end{itemize}
 
-The time to preprocess the data will vary significantly with the choice of model and local explanation.
-For reference, the FIFA data, 5000 observations of 9 explanatory variables, took 2.9 seconds to fit a random forest models of modest hyperparmeters.
+The time to preprocess the data will vary significantly with the model and local explanation.
+For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.9 seconds to fit a random forest model of modest hyperparameters.
 Extracting the tree SHAP values of each observation took 254 seconds combined.
-PCA and statistics of the variables and attribution took 0.6 seconds.
+PCA and statistics of the variables and attributions took 0.6 seconds.
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the time will be spent on the local attribution.
-This makes tree SHAP, with its reduced comupational complexity a good candidate to start with.
-Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+Increased model complexity or data dimensionality will quickly become an obstacle.
+This makes tree SHAP, with its reduced computational complexity, a good candidate to start with.
+Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 2021).
-The application is made with \textbf{shiny} (Chang et al.
+The application was made with \textbf{shiny} (Chang et al.
 2021).
@@
 \begin{Highlighting}[]
-\DocumentationTok{\#\# Download the package:}
+\DocumentationTok{\#\# Download the package}
 \FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
+\DocumentationTok{\#\# Restart the R session to pick up the file structure may be needed}
+\FunctionTok{restartSession}\NormalTok{()}
+\DocumentationTok{\#\# Load cheem into session}
 \FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
-\DocumentationTok{\#\# Try the app:}
+\DocumentationTok{\#\# Try the app}
 \FunctionTok{run\_app}\NormalTok{()}
@@
 
-To illustrate the use of the cheem method we apply it to modern datasets, two classification examples and then two of regression.
+To illustrate the use of the cheem method, we apply it to modern datasets, two classification examples and then two of regression.
 
@@
 
-In figure \ref{fig:casepenguins} a misclassified point is contrasted with a correctly classified point of its observed class that is nearby in data-space.
-The attribution space from the treeshap local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution while primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
+In figure \ref{fig:casepenguins}, a misclassified point is contrasted with a correctly classified point of its observed class nearby in data-space.
+The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
+The primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
 The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to linear explanation this observation be misclassified.
+Downplaying the contribution of bill length is crucial to the linear explanation of this observation be misclassified.
 
@@
 A random forest model is fit classifying the type of chocolate.
-We select a chocolate labeled dark, through predicted to be milk chocolate with a comparison with a chocolate labeled 85\% cocoa.
+We selected a chocolate labeled dark, through predicted to be milk chocolate with a comparison with chocolate labeled 85\% cocoa.
 
@@
 
-\caption{Chocolates data  type classification (milk or dark). }\label{fig:casechocolates}
+\caption{Chocolates data type classification (milk or dark).}\label{fig:casechocolates}
 \end{figure}
 
-From figure \ref{fig:casechocolates} we similarly see that attribution space that is more separable relative to data-space.
-Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while there 8 of the inverse case.
-Calories from fat is the measure with the largest difference in treeshap attribution between these points.
+From figure \ref{fig:casechocolates} we similarly see that attribution space is more separable relative to data-space.
+Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while 8 of the inverse case.
+Calories from fat is the variable with the largest difference in treeshap attribution between these points.
 
@@
 
-The 2020 season FIFA (Leone 2020; Biecek 2018), contains many skill measurements of soccer/football players along with wage information.
+The 2020 season FIFA (Leone 2020; Biecek 2018), contains many skill measurements of soccer/football players and wage information.
 After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
-The model was fit with 5000 observations of the 9 skill aggregates before being thinned to 500 players to gate occlusion and render time.
+The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
 We compare a leading offensive fielder (L.
-Messi) with that of a leading defensive fielder (V.
-van Dijk), the same observations used in figure \ref{fig:shapdistrbd}.
+Messi) with that of a top defensive fielder (V.
+van Dijk), the same observations were used in figure \ref{fig:shapdistrbd}.
 
@@
 \caption{FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
-The primary observation is star offensive player (L.
-Messi) as compared with a top defensive player (V.
+The primary observation is a star offensive player (L.
+Messi) compared with a top defensive player (V.
 van Dijk).}\label{fig:casefifa}
@@
 
-With figure \ref{fig:casefifa} we are going to test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense has almost singular importance for the explanation of the offensive player.
+With figure \ref{fig:casefifa} we will test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, then offense skills has almost singular importance for the explanation of the offensive player.
 We vary the contribution of offensive skills.
 In the tour (3rd frame of b), offensive skills moved, and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, which does maintain some separability.
+We also notice that accuracy has rotated into the frame, maintaining some separability.
 
-\hypertarget{ames-housing-prices-2018}{%
-\subsection{4) Ames housing prices 2018}\label{ames-housing-prices-2018}}
+\hypertarget{ames-housing-2018-sales-price-regression}{%
+\subsection{4) Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
 
 Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across 9 variables.
-Using interaction from the global view we select a house with an extreme negative residual and an accurate observation close to it in the data.
+The remaining are 338 house sales across nine variables.
+Using interaction from the global view, we select a house with an extreme negative residual and an accurate observation close to it in the data.
 
@@
 Figure \ref{fig:caseames} shows the global view and extrema of the tour.
-The horizontal distance in the tour didn't show a large disparity between our selected points.
-This is not particularly surprising as most of the features have a sizable contribution.
-Rotating any one variable out of the frame will rotate other important variables into frame preserving most of the distance from intercept to prediction.
+The horizontal distance in the tour didn't show a significant disparity between our selected points.
+This is not particularly surprising as most variables have a sizable contribution.
+Rotating any one variable out of the frame will rotate other vital variables into the frame, preserving most of the distance from intercept to prediction.
 However, the tour has revealed an interesting feature worth discussing.
-notice that the observations pivot about the origin, that is the basis roughly halfway between bases in frame 1 and 2 of panel b) the data is near a singular profile.
-This means that there is a basis orthogonal to this point that describes a lot of the variation.
-Knowing these singular frames can help direct to bases that have meaningful variation in the data.
+Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames 1 and 2 of panel b) the data is near a singular profile.
+This means that there is a basis orthogonal to this point that describes sizable variation.
+Knowing these singular bases can point toward others that have meaningful variation in the data.
 
@@
 
-The need to maintain the interpretability of black-box models is clear.
-One way to do this is by the use of local explanations of an observation.
+The need to maintain the interpretability of black-box models is evident.
+One aspect uses local explanations of the model in the vicinity of an observation.
 Local explanations approximate the linear variable importance to the model.
-Our idea is to visualize approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
-Then use the linear importance from the primary observation as a projection basis and explore a single variable's importance to the structure identified with the use of the manual tour.
+Our contribution is to assess the explanations by examining the support of explanation, varying the contribution with a radial tour.
+First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
+Then normalized the linear importance from the explanation of the primary observation to use as the initial basis of a manual tour.
+The variable sensitivity to the structure identified in the explanation is explored with the tours varying basis.
 
-We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generalized to any compatible model-explanation pair.
-We apply this to both the classification and regression tasks.
-In the former, it can be used to see which variables cause a misclassification as benchmarked against a nearby observation and explore changing the contribution of variables.
-In the regression case, we compare an observation with particularly extreme residual can be compared with nearby points more accurately predicted or another suitable benchmark.
-
-We have created an open-source \textbf{R} package \textbf{cheem}, available on CRAN, to facilitate this workflow including a dynamic application, with the ability to upload preprocessed data.
+We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generally used with any compatible model-explanation pairing.
+We apply it to the classification and regression tasks.
+We have created an open-source \textbf{R} package \textbf{cheem}, available on CRAN, to facilitate preprocessing and exploration with the described interactive application.
+Toy and real data are provided or upload your data after preprocessing.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages.
-This research was supported in part by Australian government Research Training Program (RTP) scholarships.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
+This research was supported by Australian government Research Training Program (RTP) scholarships.
 
-The namesake, Cheem, refers to a fictional race of humanoid trees in Doctor Who lore.
-\textbf{DALEX} pulls on from that universe and we originally apply tree SHAP explanations which are specific to tree-based models.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations that are specific to tree-based models.
 

### 2021-12-10 figures/_make_figures.r

@@ -200,7 +200,7 @@ bd_df <- bd_df[is.na(bd_df$variable) == FALSE, ]
     facet_grid(col=vars(player)) +
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
-    labs(title="Breakdown plot",
+    labs(title = "Breakdown plot", color = "Players",
          y = "Variable", x = "Normalized contribution to prediction | variable order") +
     theme(legend.margin = margin(0,0,0,0),
           legend.position = "bottom",
@@ -219,18 +219,19 @@ wages_df <- tibble::tibble(
     labs(y = "Player", x = "Wages [2020 Euros]") +
     theme(legend.position = "off"))
 ### Plot together
-require("patchwork")
-(pw <- g_wage / g_shap / g_bd +
-    plot_layout(heights = c(1, 3, 3)))
 require("cowplot")
+if(F) ## With differing player wages
+  (cp <- cowplot::plot_grid(
+    g_wage, g_shap, g_bd, ncol = 1, 
+    rel_heights = c(1, 2, 2), labels=c("a)", "b)", "c)")))
 (cp <- cowplot::plot_grid(
-  g_wage, g_shap, g_bd, ncol = 1, 
-  rel_heights = c(1, 2, 2), labels=c("a)", "b)", "c)")))
+  g_shap, g_bd, ncol = 1, 
+  rel_heights = c(2, 2), labels=c("a)", "b)")))
 
 ## SAVE -----
 ggplot2::ggsave(
   "./figures/shap_distr_bd.png",
-  cp, device = "png", width = 7, height = 8, units = "in")
+  cp, device = "png", width = 7, height = 7, units = "in")
 
 
 # ch5_fig2_global_space -----
@@ -244,7 +245,7 @@ ggplot2::ggsave(
   
   ## Create ggplot
   gg <- global_view(layer_ls, shap_obs, comp_obs,
-                    color = "attr_proj.y_cor", as_ggplot = TRUE)
+                    color = "cor_attr_proj.y", as_ggplot = TRUE)
   
   ## Save -----
   ggplot2::ggsave(

### 2021-12-12 _paper.rmd
Manual and radial tour sections and grammerly edits. Emailing to Yang.

@@
 This distinction draws attention to the divide between interpretable models and black-box models.
-In explanatory modeling, interpretability is key for drawing inferential conclusions.
+In explanatory modeling, interpretability is vital for drawing inferential conclusions.
 While predictive modeling may opt for potentially more accurate black-box models.
@@
 <!-- ## Data visualization tours -->
-In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_review_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
 Tours are viewed as an animation over small changes to the projection basis.
@@
 There are various types of tours, which are distinguished by the generation of projection bases.
-In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020] this path is defined by changing the contribution of a selected variable.
+In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected variable.
 <!--tours and models --> Applying tours to models has been done in a couple of contexts.
@@
 <!-- Purposed approach -->
-The approach proposed below is to use the radial, manual tour to interrogate a local explanation.
-After the identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
+The proposed approach is to use the radial, manual tour to interrogate a local explanation.
+After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
 We provide a free and open-source R package __cheem__ with an interactive application to facilitate analysis.
@@
 The remainder of this paper is organized as follows.
-The next section, [Local explanation statistics](#sec:explanations), covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
+The following section, [Local explanation statistics](#sec:explanations), covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
 [Tours and the radial tour](#sec:tour) digs deeper into these animations of continuous linear projections.
@@
 In images, saliency maps overlay or offset a heatmap indicating which pixels were important [@simonyan_deep_2014].
-For instance, the presence of snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
 In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
@@
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the prediction of the defensive player.
+In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
 We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
 Panel c) shows a simplified breakdown plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
+Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to panel b), namely, by decreasing the sum of the SHAP values.
 
@@
 
-TODO:XXX
-```{r}
-##TODO:XXX
-```
+<!-- tours intro -->
+A data visualization _tour_ animates many linear projections over small changes in the basis.
+One of the key features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
+There  are various types of tours, which are distinguished by the selection of their basis paths [@lee_state_2021; @cook_grand_2008].
+
+## Manual tours and the radial subset
+
+<!-- Manual tour -->
+The manual tour [@cook_manual_1997] defines its basis path by manipulating the basis contribution of a selected variable.
+A manipulation dimension is appended onto the projection plane, with a full contribution given to the selected variable.
+The bases are then selected based on rotating this newly created manipulation space.
+A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
+This means that such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis [@karwowski_international_2006].
+However, this navigation is relatively time-consuming due to the huge volume of $p$-space.
+It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
+In this case, the projection of the normalized explanations is the feature of interest.
+
+<!-- radial tour -->
+More generally, the manual tour can change the contribution of the variable in any of the display dimensions.
+We will with a specific more directed interaction, namely, a _radial_ tour.
+In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution radius.
 
@@
 The orientation and magnitude  of the variables are inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization real task at hand, the selection of observation, and nearby comparison.
+While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization, given the real task at hand, the selection of observations to explore further.
 
@@
 <!-- Segue to classification -->
-Now that we have introduced the global view and corresponding cheem radial tour let's discuss the differences between the classification and regression cases.
+Now that we have introduced the global view and corresponding cheem radial tour, let's discuss the differences between the classification and regression cases.
 
@@
 This changes the display from 1D density to a scatterplot.
-The basis is still one component, the horizontal position, independent with the vertical position.
+The basis is still one component, the horizontal position, independent of the vertical position.
 
@@
 	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable.
-An optional categorical variable can be mapped to color and shape of obervations.
+An optional categorical variable can be mapped to the color and shape of observations.
 Explanatory variables are scaled in visualization after modeling or creating local explanations.
 	\item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)] for compatability with the local explanation which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)] for compatibility with the local explanation, which requires tree-based models.
 	\item \textbf{Local explanation:} any model-compatible linear explanation could be used.
@@
 The tour varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to the linear explanation of this observation be misclassified.
+Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.
 
@@
 
-From figure \@ref(fig:casechocolates) we similarly see that attribution space is more separable relative to data-space.
-Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while 8 of the inverse case.
+In figure \@ref(fig:casechocolates), we similarly see that attribution space is more separable relative to data-space.
+Interestingly there is not the class imbalance that we suspected; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
@@
 
-The 2020 season FIFA [@leone_fifa_2020; @biecek_dalex_2018], contains many skill measurements of soccer/football players and wage information.
+The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
 After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
@@
 
-With figure \@ref(fig:casefifa) we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense skills has almost singular importance for the explanation of the offensive player.
+With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for the explanation of the offensive player.
 We vary the contribution of offensive skills.
@@
 We have created an open-source __R__ package __cheem__, available on CRAN, to facilitate preprocessing and exploration with the described interactive application.
-Toy and real data are provided or upload your data after preprocessing.
+Toy and real data are provided, or upload your data after preprocessing.
 
@@
 We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages.
-This research was supported by Australian government Research Training Program (RTP) scholarships.
+This research was supported by Australian Government Research Training Program (RTP) scholarships.
 

### 2021-12-12 _paper.tex
Manual and radial tour sections and grammerly edits. Emailing to Yang.

@@
 There are various types of tours, which are distinguished by the generation of projection bases.
-In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020) this path is defined by changing the contribution of a selected variable.
+In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected variable.
 Applying tours to models has been done in a couple of contexts.
@@
 
-The approach proposed below is to use the radial, manual tour to interrogate a local explanation.
-After the identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
+The proposed approach is to use the radial, manual tour to interrogate a local explanation.
+After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
 We provide a free and open-source R package \textbf{cheem} with an interactive application to facilitate analysis.
@@
 The remainder of this paper is organized as follows.
-The next section, \protect\hyperlink{sec:explanations}{Local explanation statistics}, covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
+The following section, \protect\hyperlink{sec:explanations}{Local explanation statistics}, covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
 \protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into these animations of continuous linear projections.
@@
 In images, saliency maps overlay or offset a heatmap indicating which pixels were important (Simonyan, Vedaldi, and Zisserman 2014).
-For instance, the presence of snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
 In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
@@
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more important to the model for explaining the prediction of the defensive player.
+In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
 We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
 Panel c) shows a simplified breakdown plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to that of panel b), namely, by decreasing the sum of the SHAP values.
+Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to panel b), namely, by decreasing the sum of the SHAP values.
 
@@
 
-TODO:XXX
+A data visualization \emph{tour} animates many linear projections over small changes in the basis.
+One of the key features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
+There are various types of tours, which are distinguished by the selection of their basis paths (S.
+Lee et al.
+2021; Cook et al.
+2008).
+
+\hypertarget{manual-tours-and-the-radial-subset}{%
+\subsection{Manual tours and the radial subset}\label{manual-tours-and-the-radial-subset}}
+
+The manual tour (Cook and Buja 1997) defines its basis path by manipulating the basis contribution of a selected variable.
+A manipulation dimension is appended onto the projection plane, with a full contribution given to the selected variable.
+The bases are then selected based on rotating this newly created manipulation space.
+A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
+This means that such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
+However, this navigation is relatively time-consuming due to the huge volume of \(p\)-space.
+It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
+In this case, the projection of the normalized explanations is the feature of interest.
+
+More generally, the manual tour can change the contribution of the variable in any of the display dimensions.
+We will with a specific more directed interaction, namely, a \emph{radial} tour.
+In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution radius.
 
@@
 The orientation and magnitude of the variables are inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization real task at hand, the selection of observation, and nearby comparison.
+While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization, given the real task at hand, the selection of observations to explore further.
 
@@
 
-Now that we have introduced the global view and corresponding cheem radial tour let's discuss the differences between the classification and regression cases.
+Now that we have introduced the global view and corresponding cheem radial tour, let's discuss the differences between the classification and regression cases.
 
@@
     \item \textbf{Data:} a complete numerical matrix; explanatory and response variable.
-An optional categorical variable can be mapped to color and shape of obervations.
+An optional categorical variable can be mapped to the color and shape of observations.
 Explanatory variables are scaled in visualization after modeling or creating local explanations.
     \item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)] for compatability with the local explanation which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)] for compatibility with the local explanation, which requires tree-based models.
     \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
@@
 The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to the linear explanation of this observation be misclassified.
+Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.
 
@@
 
-From figure \ref{fig:casechocolates} we similarly see that attribution space is more separable relative to data-space.
-Interestingly there is not the class imbalance that we suspected; there are only 6 chocolates labeled as dark and predicted as milk, while 8 of the inverse case.
+In figure \ref{fig:casechocolates}, we similarly see that attribution space is more separable relative to data-space.
+Interestingly there is not the class imbalance that we suspected; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
@@
 
-The 2020 season FIFA (Leone 2020; Biecek 2018), contains many skill measurements of soccer/football players and wage information.
+The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
 After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
@@
 
-With figure \ref{fig:casefifa} we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense skills has almost singular importance for the explanation of the offensive player.
+With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for the explanation of the offensive player.
 We vary the contribution of offensive skills.
@@
 We have created an open-source \textbf{R} package \textbf{cheem}, available on CRAN, to facilitate preprocessing and exploration with the described interactive application.
-Toy and real data are provided or upload your data after preprocessing.
+Toy and real data are provided, or upload your data after preprocessing.
 
@@
 We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
-This research was supported by Australian government Research Training Program (RTP) scholarships.
+This research was supported by Australian Government Research Training Program (RTP) scholarships.
 
@@
 
+\leavevmode\hypertarget{ref-cook_grand_2008}{}%
+Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
+2008.
+{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual} {Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
+Berlin, Heidelberg: Springer Berlin Heidelberg.
+\url{https://doi.org/10.1007/978-3-540-33037-0_13}.
+
 \leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
@@
 
+\leavevmode\hypertarget{ref-karwowski_international_2006}{}%
+Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
+
 \leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
@@
 
-\leavevmode\hypertarget{ref-lee_review_2021}{}%
-Lee, Stuart, Dianne Cook, Natalia Da Silva, Ursula Laa, Earo Wang, Nick Spyrison, and H.
+\leavevmode\hypertarget{ref-lee_state_2021}{}%
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
 Sherry Zhang.
 2021.
-{``A {Review} of the {State}-of-the-{Art} on {Tours} for {Dynamic} {Visualization} of {High}-Dimensional {Data}.''} \emph{arXiv Preprint arXiv:2104.08016}.
+{``The State-of-the-Art on Tours for Dynamic Visualization of High-Dimensional Data.''} \emph{WIREs Computational Statistics} n/a (n/a): e1573.
+\url{https://doi.org/10.1002/wics.1573}.
 

### 2021-12-15 _paper.rmd
Abstract recovered, but not authors.

@@
 shorttitle: Cheem
-#author:
-#  - name: Nicholas Spyrison
+author:
+#- name: "Nicholas Spyrison"
   # address: |
@@
   # orcid: 0000-0002-8417-0212
-#  - name: Dianne Cook
+#- name: "Dianne Cook"
   # affiliation: Monash University
@@
   # orcid: 000-0002-3813-7155
-#  - name: Kimbal Marriott
+#- name: "Kimbal Marriott"
   # affiliation: Monash University
@@
 #   plain:     [XAI, explanable artificial intelligence, black-box models, local # explanations, grand tour, manual tour, DR, dimension reduction, R]
-#abstract: |
-#  Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
+abstract: |
+ Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
 In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
@@
   bookdown::pdf_document2:
+  #pdf_document:
     keep_tex: true
@@
 <!-- A bib reference [@wickham_visualizing_2015]. -->
-<!-- A [Section intro](#sec:cheatsheet) reference, alternatively, section {sec:intro} (with no @; \\ref{sec:intro}). -->
+<!-- A [Section intro](#sec:cheatsheet) reference, alternatively, section \@ref(sec:intro) (with @; \\@ref(sec:intro)). -->
 <!-- ```{r crest, echo=FALSE, out.height = "10%", out.width = "10%", fig.cap = "A caption for crest figure"} -->

### 2021-12-15 _paper.tex
Abstract recovered, but not authors.

@@
 \maketitle
+\begin{abstract}
+Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
+In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
+XAI attempts to shed light on these models, one such approach is the use of local explanations.
+A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
+We extract explanations for each observation, and approximate data and this attribution space side-by-side with linked brushing.
+After identifying an observation of interest its local explanation is used as a 1D projection basis.
+We then manipulate the magnitude of the variable contributions with a technique called the tour.
+This tour animates many projections over small changes in the projection basis.
+Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
+The implementation of our framework is available as an \textbf{R} package \textbf{cheem} available on CRAN.
+\end{abstract}
 

### 2021-12-22 _paper.rmd
edits from Jieyang and grammerly.

@@
 ---
-title: "Interrogating the linear variable importance of local explanations of black-box models with animated linear projections"
+title: "Interrogating the linear variable importance of local explanations of non-linear models with animated linear projections"
 shorttitle: Cheem
+keywords:
+- R
+- XAI
+- explanable artificial intelligence
+- black-box models
+- local explanations
+- grand tour
+- manual tour
+- radial tour
 author:
 #- name: "Nicholas Spyrison"
-  # address: |
 #     | Monash University
 #     | Faculty of Information Technology
+#   address: |
 #     | ORCiD: 0000-0002-8417-0212
@@
 #   orcid: 0000-0002-9813-0377
-# keywords:
-#   formatted: [XAI, explanable artificial intelligence, black-box models, local # explanations, grand tour, manual tour, DR, dimension reduction, R]
-#   plain:     [XAI, explanable artificial intelligence, black-box models, local # explanations, grand tour, manual tour, DR, dimension reduction, R]
 abstract: |
@@
 In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
-XAI attempts to shed light on these models, one such approach is the use of local explanations.
+XAI attempts to shed light on these models.
+One such approach is _local explanations_.
 A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
-We extract explanations for each observation, and approximate data and this attribution space side-by-side with linked brushing.
-After identifying an observation of interest its local explanation is used as a 1D projection basis.
-We then manipulate the magnitude of the variable contributions with a technique called the tour.
-This tour animates many projections over small changes in the projection basis.
-Doing so allows a user to visually explore the data space through the lens of this local explanation and interrogate its variable importance.
+We extract explanations for each observation and approximate data, and this attribution space side-by-side with linked brushing.
+After identifying an observation of interest, its local explanation is used as a 1D projection basis.
+We then manipulate the magnitude of the variable contributions with the _tour_ technique.
+The tour animates many projections over small changes in the projection basis as it changes the contribution of one variable.
+Doing so allows a user to visually explore the data space through the lens of this local explanation and test its support.
 The implementation of our framework is available as an __R__ package __cheem__ available on CRAN.
@@
 #appendix:
-#   - "appendix-1.Rmd"
 # title:
@@
 This distinction draws attention to the divide between interpretable models and black-box models.
+The intended use has important implications for model selection and development.
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for potentially more accurate black-box models.
-The intended use has important implications for model selection and development.
+While predictive modeling may opt for more accurate non-linear models, the additional hindrance to interpretation should not be taken lightly.
 
 <!-- XAI & interpretability crisis -->
-Black-box models are becoming increasingly common, but not without their share of controversy and issues [@oneil_weapons_2016; @kodiyan_overview_2019].
+Interpretability aside, black-box models are becoming increasingly common, but not without their share of controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
 Black-box models have been known to reflect common biases, including sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
@@
 Data drift can lead to worse predictions [@lazer_parable_2014; @salzberg_why_2014].
+While typically more resistant to data drift, linear models will similarly find biases in the data.
+However, their transparency goes a long way into exploring and mitigating the biases.
 Such cases highlight the need to make models fair, accountable, ethical, and transparent, which has led to the movement of XAI [@adadi_peeking_2018; @arrieta_explainable_2020].
@@
 
-<!-- SHAP -->
-To illustrate our work, we apply the model-agnostic explanation SHAP [@strumbelj_efficient_2010; @strumbelj_explaining_2014].
-The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
-To be exact, we apply a variant that enjoys a lower computational complexity, known as tree SHAP [@lundberg_consistent_2018].
-<!-- 2nd half is more application details; move lower? -->
+<!-- Data visualization tours -->
+In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+Tours are viewed as an animation over minor changes to the projection basis.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information of the structure and features of the object.
 
-<!-- ## Data visualization tours -->
-In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
-Tours are viewed as an animation over small changes to the projection basis.
-Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
-
-<!-- manual tours --> 
-There are various types of tours, which are distinguished by the generation of projection bases.
+<!-- Manual tours --> 
+There are various types of tours distinguished by the generation of projection bases.
 In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected variable.
@@
 <!-- Purposed approach -->
-The proposed approach is to use the radial, manual tour to interrogate a local explanation.
+The proposed approach uses the radial, manual tour to interrogate a local explanation.
 After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
 We provide a free and open-source R package __cheem__ with an interactive application to facilitate analysis.
-We give case studies of toy and modern datasets for both classification and regression tasks.
+We provide toy and modern datasets case studies for both classification and regression tasks.
 
@@
 The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020].
-Latin for “other things held constant” or “all else unchanged” is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+This phrase, Latin for “other things held constant” or “all else unchanged”, shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual, observed case by that varies contributions of the basis, essentially the orientation of the data object.
-Another difference is that the basis must maintain orthonormality.
+In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
+Another difference is that the basis must remain orthonormal.
 That is to say, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
@@
 [Tours and the radial tour](#sec:tour) digs deeper into these animations of continuous linear projections.
-The section [Application Design](#sec:applicationdesign) discusses the layout of the application, how it facilitates analysis, preprocessing, and package infrastructure.
+The section [Application Design](#sec:applicationdesign) discusses the visual layouts, how they facilitates analysis, data preprocessing, and package infrastructure.
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
-We conclude with a [Discussion](#sec:discussion) of the insights we draw from classification and regression tasks.
+We conclude with a [Discussion](#sec:cheemdiscussion) of the insights we draw from classification and regression tasks.
 
@@
 
-<!-- reminder of local explanation -->
+<!-- Reminder of local explanation -->
 Consider a highly non-linear model.
-At face value, it is hard to say which variable(s) are sensitive to the crossing of a classification boundary or identifies which variables caused an observation to have a relatively extreme residual.
+At face value, it is hard to say which variable(s) are sensitive to  crossing a classification boundary or identify which variables caused an observation to have a relatively extreme residual.
 Local explanations shed light on these cases by approximating linear variable importance in the vicinity of one observation.
@@
 <!-- Taxonomy of local explanations -->
-Figure 6 of @arrieta_explainable_2020 gives comprehensive summarization of the taxonomy and literature of explanation techniques.
-This includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016; @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
-There are a fewer number of model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017] and their variants are popular.
+Figure six of @arrieta_explainable_2020 gives comprehensive summarization of the taxonomy and literature of explanation techniques.
+The figure includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016; @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017] and their variants are popular.
 
 <!-- Uses of local explanations -->
-These instance-level explanations are used in a variety of ways depending on the data.
-In images, saliency maps overlay or offset a heatmap indicating which pixels were important [@simonyan_deep_2014].
+These instance-level explanations are used in various ways depending on the data.
+In images, saliency maps overlay or offset a heatmap indicating which pixels were necessary [@simonyan_deep_2014].
 For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
 In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
-In the case of numeric regression, they are used to explain variable additive contributions from model intercept to the observation's prediction [@ribeiro_why_2016].
+In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction [@ribeiro_why_2016].
 
@@
 
-<!-- fifa example -->
+<!-- Fifa example -->
 To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season [@leone_fifa_2020].
@@
 
+```{r shapdistrbd, echo=F, fig.align='center', fig.cap = "Illustration of the distribution of SHAP attributions and a break-down plot.
+From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
+The players have very different salaries, but a) shows the distributions of 25 permutations in the explanatory variables.
+The medians of these distributions are the final SHAP values.
+The variable importance differs across the exogenous information of player position.
+These explanations make sense; the variable importances seem realistic given the player's positions.
+b) Break-down plots of the observations using their explanations to additively cover the difference between the model intercept and the observation predictions."}
+knitr::include_graphics("./figures/shap_distr_bd.png")
+```
+
 <!-- Illustration -->
@@
 Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
+In the light of the player position, the difference in the variable importance makes sense; offense and movement skills are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
 We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
-Panel c) shows a simplified breakdown plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to panel b), namely, by decreasing the sum of the SHAP values.
+Panel b) shows a simplified break-down plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
+Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a), by decreasing sum of the SHAP values.
 
-```{r shapdistrbd, echo=F, fig.cap = "Illustration of the distribution of SHAP attributions and a breakdown plot.
-From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
-The players have very different wages, but b) shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
-The medians of these distributions are the final SHAP values.
-The variable importance differs across the exogenous information of player position.
-These explanations make sense; the variable importances make sense in light of the player's position.
-c) Breakdown plots of the observations the explanation used to additively explain the difference between the model intercept and the observation prediction."}
-knitr::include_graphics("./figures/shap_distr_bd.png")
-```
-
-In summary, this highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this instance, we showed how two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
-In the following section, we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
+<!-- Segue -->
+In summary, this figure highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
+In application we apply _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
+Tree SHAP only compatible with tree-based models; we illustrate on random forests.
+In the following section, we will use normalized explanations as the starting projection basis to further interrogate the explanation.
 
@@
 
-<!-- tours intro -->
+<!-- Tours intro -->
 A data visualization _tour_ animates many linear projections over small changes in the basis.
-One of the key features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
-There  are various types of tours, which are distinguished by the selection of their basis paths [@lee_state_2021; @cook_grand_2008].
+One of the critical features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
+There are various types of tours that are distinguished by selecting their basis paths [@lee_state_2021; @cook_grand_2008].
 
-## Manual tours and the radial subset
+## Manual tours and its radial case
 
 <!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its basis path by manipulating the basis contribution of a selected variable.
-A manipulation dimension is appended onto the projection plane, with a full contribution given to the selected variable.
+The manual tour [@cook_manual_1997] defines its basis path by manipulating a selected variable's basis contribution.
+A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
 The bases are then selected based on rotating this newly created manipulation space.
 A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
-This means that such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis [@karwowski_international_2006].
-However, this navigation is relatively time-consuming due to the huge volume of $p$-space.
+Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis [@karwowski_international_2006].
+However, this navigation is relatively time-consuming due to the vast volume of the display-space.
 It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
@@
 
-<!-- radial tour -->
-More generally, the manual tour can change the contribution of the variable in any of the display dimensions.
-We will with a specific more directed interaction, namely, a _radial_ tour.
-In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution radius.
+<!-- Radial tour -->
+More generally, the manual tour can change the contribution of a variable to the display dimensions.
+We will apply a more directed interaction, namely, a _radial_ tour.
+In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
 
-# Application design {#sec:applicationdesign}
+# cheem viewer {#sec:applicationdesign}
 
-Below we illustrate the two primary displays of the application: the global view and the tour view.
+Below we illustrate the two primary displays of the cheem viewer application: the global view and the tour view.
 Then we cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing that before display.
+Lastly, we discuss the preprocessing the data before application runtime.
 
@@
 <!-- Purpose -->
-The global view provides an essential context of all observations and allows  exploration of the separability of the data- and attribution-spaces.
-While ultimately, its purpose is to facilitate the selection of a primary point and comparison point.
+The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+While comparison of these spaces is interesting, the global view's purpose is to facilitate the selection of observations.
+The local explanation of these points will be explored in more detail.
 
@@
 An approximation of these spaces is given as the first two principal components of their respective spaces.
-This is shown side-by-side next to model information, the prediction, and the observed response variable.
+Model information, observed response by its prediction, is also provided.
 The orientation and magnitude of the variables are inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization, given the real task at hand, the selection of observations to explore further.
-
-<!-- Predicted by observed -->
-It is insightful to explore these two approximations against a visual of the model; prediction by observation is also displayed.
-Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
+<!-- Interactions -->Misclassified observations are circled in red if applicable.
+Linked brushing between the plots and tabular display of select points facilitates exploration of the spaces and the model.
+A single 2D projection will not encompass all of the structure of higher-dimensional space.
+However, it is a reasonable summarization given the real task at hand; the selection of observations to explore further.
 
@@
 
-<!-- a function of the obs selected above -->
+<!-- A function of the obs selected -->
 The global view facilitated the selection of a primary and optional comparison observation.
-The variable-level attribution of primary observation is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of  the contributions of the linear variables that best explain the difference between the model intercept and an observations prediction, not the local shape of the model surface.
+The variable-level attribution of the primary observation is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observation's prediction, not the local shape of the model surface.
 
@@
 
-<!-- tour animation -->
+<!-- Tour animation -->
 The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
-The default variable selected has the largest discrepancy between the primary and comparison observations attribution.
-In the following sections, we elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
-
-<!-- Segue to classification -->
-Now that we have introduced the global view and corresponding cheem radial tour, let's discuss the differences between the classification and regression cases.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The following sections elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
+<!-- Segue to classification -->Now that we have introduced the global view and corresponding cheem radial tour let us discuss the differences between the classification and regression cases.
 
@@
 What information do we glean from using this method on a classification task?
-Typically we select a misclassified observation compared to a correctly classified point that is nearby in data space.
-We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction.
+Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
+The initial frame is the linear attribution of that observation's local explanation.
 By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
@@
 
-```{r classificationcase, echo=F, out.width = "100%", fig.cap = "Display illustrating the classification case.
-Both views are colored on predicted class and while red circles identify misclassified observations.
+```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Display illustrating the classification case.
+Plots are  colored on predicted class and red circles indicate misclassified observations.
 The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
@@
 
-The regression case, we opt to color the global view on a statistic, the observation's residual, the log Mahanabis distance in data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
-In the radial tour, the horizontal positions are the same, the  attribution projection, then varying as the radial tour changes basis.
-While the vertical position is fixed to the observed response variable and residuals for the middle and right panels, respectively.
-This changes the display from 1D density to a scatterplot.
-The basis is still one component, the horizontal position, independent of the vertical position.
+In the regression case, the global view can be colored on a statistic to highlight the structure in the explanation space, including residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
+The vertical position is fixed to the observed response variable and residuals in the middle and right panels, respectively.
+Correspondingly, the display changes from univariate density to 2D scatterplot.
+The basis is still one component (horizontal) independent of the vertical position.
 
-```{r regressioncase, echo=F, out.width = "100%", fig.cap = "Display of the regression case.
+```{r regressioncase, echo=F, out.width = "100%", fig.cap="Display of the regression task.
 The global view can be colored on the correlation of the attribution projection and observed response.
-In the tour, the horizontal values are the same as the classification case.
-The vertical position is fixed to the observed response and residuals in the middle and right facets, respectively."}
+In the tour, the horizontal values are the same as the classification case; the projection through the basis.
+The vertical position is now maped to the observed y and residuals."}
 knitr::include_graphics("./figures/app_regression.PNG")
@@
 
-<!-- reactive vs exploration interactions -->
-The application has several reactive selections that affect the data, coloring, and inputs for the tour to improve flexibility by extending the use cases.
-It also has interactive features in the global view that aid the analysis.
+<!-- Reactive vs exploration interactions -->
+The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
+These reactive inputs make the software flexible and extensible.
+The application also has It also has more exploratory interactions to help link points in the data and extract other information not on plot axes.
 
-<!-- Interaction except -->
-A tooltip displays while the cursor hovers over a point displays the observation number/name and classification information if appropriate.
-Linked brushing allows for the selection of points (with left click and drag) where those points will be highlighted in both plots.
+<!-- Exploratory interactions -->
+A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
 The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and ultimately the selection of the primary and comparison observations.
+These interactions aid exploration of the spaces and, finally, identification of a primary and comparison observation.
 
@@
 This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over a point"}
+Tooltips display when hovering over an obersvation."}
 knitr::include_graphics("./figures/app_interactions.PNG")
@@
 
-The benefit of having dynamic interaction with data is predicated on a reasonably short render time.
-It is important to preprocess expensive operations so the application resources can be used efficiently.
-The work remaining at runtime is solely responding to inputs and rendering of visuals and tables.
+It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
+In order to keep render time low all of the computational operations should be prepared before runtime.
+The work remaining at runtime is solely reacting to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 	\item \textbf{Local explanation:} any model-compatible linear explanation could be used.
-We apply tree SHAP, a more computationally efficient variant of SHAP, that is applicable to tree-based models.
-This is done with the package \textbf{treeshap} [(ref:citeTs), hosted on GitHub only].
+We apply tree SHAP, a more computationally efficient variant of SHAP, compatible with tree-based models.
+Tree SHAP was calculated with the package \textbf{treeshap} [(ref:citeTs), hosted on GitHub only].
 The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
@@
 
-<!-- note on time of execution -->
+<!-- Note on time of execution -->
 The time to preprocess the data will vary significantly with the model and local explanation.
@@
 PCA and statistics of the variables and attributions took 0.6 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice it to say that the bulk of the time will be spent on the local attribution.
-Increased model complexity or data dimensionality will quickly become an obstacle.
+These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
+An increased of model complexity or data dimensionality will quickly become an obstacle.
 This makes tree SHAP, with its reduced computational complexity, a good candidate to start with.
@@
 
-The above-described method and application are implemented as an open-source __R__ package, __cheem__ _TODO:XXX include  cran URL after published_.
-Preprocessing was facilitated with models created via __randomForest__ [liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
+The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on [CRAN](https://CRAN.R-project.org/package=cheem).
+Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
 The application was made with __shiny__ [@chang_shiny_2021].
-The tour visual is an extension of __spinifex__ [@spyrison_spinifex_2020].
+The tour visual is built with __spinifex__ [@spyrison_spinifex_2020].
 Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive HTML widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] was a huge boon to understanding local explanations and how to apply them.
+__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a huge boon to understanding local explanations and how to apply them.
 
-Installation and get started with the package can be acheived by running the following in __R__:
+## Installation and getting started
+
+The following __R__ code will help getting up and running:
 
@@
 install.packages("cheem", dependencies = TRUE)
-## Restart the R session to pick up the file structure may be needed
+## Restart the R session so the IDE has the correct directory structure
 restartSession()
@@
 run_app()
-## Bring your own data: follow the examples in cheem_ls()
+
+# Processing your data
+## Install treeshap from github, to use as a local explainer
+remotes::install_github('ModelOriented/treeshap') ## Local 
+## Follow the examples in cheem_ls()
 ?cheem_ls
@@
 The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
-The primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
+The primary observation is between the predicted and observed clusters, a sign of uncertainty in the prediction.
 The tour varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
@@
 Each of which was labeled as being either milk or dark chocolates.
-With this data, we can see if a manufacturer gives an accurate portrayal of the chocolate.
-We are curious to see if there are chocolates that nutritionally look like milk chocolates that are labeled as dark chocolates, which may hold a higher market value.
+With this data, we can see if a manufacturer accurately portrays the chocolate.
+We are curious to see if there are chocolates that nutritionally look like milk chocolates labeled as dark chocolates, which may hold a higher market value.
 We should note that not all chocolates consist wholly of chocolate.
@@
 A random forest model is fit classifying the type of chocolate.
-We selected a chocolate labeled dark, through predicted to be milk chocolate with a comparison with chocolate labeled 85% cocoa.
+We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85% cocoa.
 
@@
 
-In figure \@ref(fig:casechocolates), we similarly see that attribution space is more separable relative to data-space.
-Interestingly there is not the class imbalance that we suspected; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
+In figure \@ref(fig:casechocolates),  we similarly see that attribution space is more separable than data-space.
+Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for the explanation of the offensive player.
+If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for explaining the offensive player.
 We vary the contribution of offensive skills.
-In the tour (3rd frame of b), offensive skills moved, and Messi is no longer separated from the group.
+In the tour (panel b, frame 3), offensive skills are removed, and Messi is no longer separated from the group.
 We also notice that accuracy has rotated into the frame, maintaining some separability.
@@
 However, the tour has revealed an interesting feature worth discussing.
-Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames 1 and 2 of panel b) the data is near a singular profile.
+Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames one and two of panel b) the data is near a singular profile.
 This means that there is a basis orthogonal to this point that describes sizable variation.
-Knowing these singular bases can point toward others that have meaningful variation in the data.
+Knowing these singular bases can
+point toward others with meaningful data variation.
 
@@
 Local explanations approximate the linear variable importance to the model.
-Our contribution is to assess the explanations by examining the support of explanation, varying the contribution with a radial tour.
-First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast, and ultimately, identify primary and comparison observations of interest.
-Then normalized the linear importance from the explanation of the primary observation to use as the initial basis of a manual tour.
-The variable sensitivity to the structure identified in the explanation is explored with the tours varying basis.
+Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
+First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast and, identify primary and comparison observations of interest.
+The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with radial tour.
+The variable sensitivity to the structure identified in the explanation is explored by the tours.
 
@@
 We apply it to the classification and regression tasks.
-We have created an open-source __R__ package __cheem__, available on CRAN, to facilitate preprocessing and exploration with the described interactive application.
+We have created an open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem), to facilitate preprocessing and exploration with the described interactive application.
 Toy and real data are provided, or upload your data after preprocessing.
@@
 This research was supported by Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this piece.
 
 The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations that are specific to tree-based models.
+__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
 

### 2021-12-22 _paper.tex
edits from Jieyang and grammerly.

--- 2021-12-12 _paper.tex
@@
 \hypersetup{
-  pdftitle={Interrogating the linear variable importance of local explanations of black-box models with animated linear projections},
+  pdftitle={Interrogating the linear variable importance of local explanations of non-linear models with animated linear projections},
+  pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
   colorlinks=true,
@@
 
-\title{Interrogating the linear variable importance of local explanations of black-box models with animated linear projections}
+\title{Interrogating the linear variable importance of local explanations of non-linear models with animated linear projections}
 \author{}
@@
 \maketitle
+\begin{abstract}
+Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
+In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
+XAI attempts to shed light on these models.
+One such approach is \emph{local explanations}.
+A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
+We extract explanations for each observation and approximate data, and this attribution space side-by-side with linked brushing.
+After identifying an observation of interest, its local explanation is used as a 1D projection basis.
+We then manipulate the magnitude of the variable contributions with the \emph{tour} technique.
+The tour animates many projections over small changes in the projection basis as it changes the contribution of one variable.
+Doing so allows a user to visually explore the data space through the lens of this local explanation and test its support.
+The implementation of our framework is available as an \textbf{R} package \textbf{cheem} available on CRAN.
+\end{abstract}
 
@@
 This distinction draws attention to the divide between interpretable models and black-box models.
+The intended use has important implications for model selection and development.
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for potentially more accurate black-box models.
-The intended use has important implications for model selection and development.
+While predictive modeling may opt for more accurate non-linear models, the additional hindrance to interpretation should not be taken lightly.
 
-Black-box models are becoming increasingly common, but not without their share of controversy and issues (O'neil 2016; Kodiyan 2019).
+Interpretability aside, black-box models are becoming increasingly common, but not without their share of controversy (O'neil 2016; Kodiyan 2019).
 Black-box models have been known to reflect common biases, including sex (Dastin 2018; Duffy 2019), race (Larson et al.
@@
 2014; Salzberg 2014).
+While typically more resistant to data drift, linear models will similarly find biases in the data.
+However, their transparency goes a long way into exploring and mitigating the biases.
 Such cases highlight the need to make models fair, accountable, ethical, and transparent, which has led to the movement of XAI (Adadi and Berrada 2018; Arrieta et al.
@@
 
-To illustrate our work, we apply the model-agnostic explanation SHAP (Strumbelj and Kononenko 2010; Štrumbelj and Kononenko 2014).
-The exact details of SHAP are tangent to the ideas of this work, but suffice it to say that SHAP approximates variable importance by taking the median importance over permutations of the explanatory variables.
-To be exact, we apply a variant that enjoys a lower computational complexity, known as tree SHAP (S.
-M.
-Lundberg, Erion, and Lee 2018).
-
 In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
@@
 2021) is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
-Tours are viewed as an animation over small changes to the projection basis.
+Tours are viewed as an animation over minor changes to the projection basis.
 Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the structural shape of the shadow change gleans insight into the shape and features of the object.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information of the structure and features of the object.
 
-There are various types of tours, which are distinguished by the generation of projection bases.
+There are various types of tours distinguished by the generation of projection bases.
 In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected variable.
@@
 
-The proposed approach is to use the radial, manual tour to interrogate a local explanation.
+The proposed approach uses the radial, manual tour to interrogate a local explanation.
 After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
 We provide a free and open-source R package \textbf{cheem} with an interactive application to facilitate analysis.
-We give case studies of toy and modern datasets for both classification and regression tasks.
+We provide toy and modern datasets case studies for both classification and regression tasks.
 
 The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-Latin for ``other things held constant'' or ``all else unchanged'' is a counterfactual analysis showing how an observation's prediction would change from a change in one explanatory variable given that other variables are held constant.
+This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual, observed case by that varies contributions of the basis, essentially the orientation of the data object.
-Another difference is that the basis must maintain orthonormality.
+In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
+Another difference is that the basis must remain orthonormal.
 That is to say, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
@@
 \protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into these animations of continuous linear projections.
-The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the layout of the application, how it facilitates analysis, preprocessing, and package infrastructure.
+The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the visual layouts, how they facilitates analysis, data preprocessing, and package infrastructure.
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
-We conclude with a \protect\hyperlink{sec:discussion}{Discussion} of the insights we draw from classification and regression tasks.
+We conclude with a \protect\hyperlink{sec:cheemdiscussion}{Discussion} of the insights we draw from classification and regression tasks.
 
@@
 Consider a highly non-linear model.
-At face value, it is hard to say which variable(s) are sensitive to the crossing of a classification boundary or identifies which variables caused an observation to have a relatively extreme residual.
+At face value, it is hard to say which variable(s) are sensitive to crossing a classification boundary or identify which variables caused an observation to have a relatively extreme residual.
 Local explanations shed light on these cases by approximating linear variable importance in the vicinity of one observation.
 
-Figure 6 of Arrieta et al.
+Figure six of Arrieta et al.
 (2020) gives comprehensive summarization of the taxonomy and literature of explanation techniques.
-This includes a large number of model-specific explanations such as deepLIFT, (Shrikumar et al.
+The figure includes a large number of model-specific explanations such as deepLIFT, (Shrikumar et al.
 2016; Shrikumar, Greenside, and Kundaje 2017) a popular recursive method for estimating importance in neural networks.
-There are a fewer number of model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (S.
+There are fewer model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (S.
 Lundberg and Lee 2017) and their variants are popular.
 
-These instance-level explanations are used in a variety of ways depending on the data.
-In images, saliency maps overlay or offset a heatmap indicating which pixels were important (Simonyan, Vedaldi, and Zisserman 2014).
+These instance-level explanations are used in various ways depending on the data.
+In images, saliency maps overlay or offset a heatmap indicating which pixels were necessary (Simonyan, Vedaldi, and Zisserman 2014).
 For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
@@
 2018).
-In the case of numeric regression, they are used to explain variable additive contributions from model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
+In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
 
@@
 
-Figure \ref{fig:shapdistrbd} shows the SHAP values of these players.
-Panel a) shows these players receive a sizable difference in wages.
-Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offensive and movement are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
-We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
-Panel c) shows a simplified breakdown plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observations prediction.
-Such additive approaches will show an asymmetry in the variable ordering, so we opt to fix the order to panel b), namely, by decreasing the sum of the SHAP values.
-
 \begin{figure}
@@
 
-\caption{Illustration of the distribution of SHAP attributions and a breakdown plot.
+\caption{Illustration of the distribution of SHAP attributions and a break-down plot.
 From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
-The players have very different wages, but b) shows the distributions of the attributions permuting over 25 permutations in the explanatory variables.
+The players have very different salaries, but a) shows the distributions of 25 permutations in the explanatory variables.
 The medians of these distributions are the final SHAP values.
 The variable importance differs across the exogenous information of player position.
-These explanations make sense; the variable importances make sense in light of the player's position.
-c) Breakdown plots of the observations the explanation used to additively explain the difference between the model intercept and the observation prediction.}\label{fig:shapdistrbd}
+These explanations make sense; the variable importances seem realistic given the player's positions.
+b) Break-down plots of the observations using their explanations to additively cover the difference between the model intercept and the observation predictions.}\label{fig:shapdistrbd}
 \end{figure}
 
-In summary, this highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this instance, we showed how two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
-In the following section, we will be using normalized explanations as the starting projection basis to interrogate the explanation further.
+Figure \ref{fig:shapdistrbd} shows the SHAP values of these players.
+Panel a) shows these players receive a sizable difference in wages.
+Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
+In the light of the player position, the difference in the variable importance makes sense; offense and movement skills are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
+We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
+Panel b) shows a simplified break-down plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
+Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a), by decreasing sum of the SHAP values.
+
+In summary, this figure highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
+In application we apply \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (S.
+M.
+Lundberg, Erion, and Lee 2018).
+Tree SHAP only compatible with tree-based models; we illustrate on random forests.
+In the following section, we will use normalized explanations as the starting projection basis to further interrogate the explanation.
 
@@
 A data visualization \emph{tour} animates many linear projections over small changes in the basis.
-One of the key features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
-There are various types of tours, which are distinguished by the selection of their basis paths (S.
+One of the critical features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
+There are various types of tours that are distinguished by selecting their basis paths (S.
 Lee et al.
@@
 
-\hypertarget{manual-tours-and-the-radial-subset}{%
-\subsection{Manual tours and the radial subset}\label{manual-tours-and-the-radial-subset}}
+\hypertarget{manual-tours-and-its-radial-case}{%
+\subsection{Manual tours and its radial case}\label{manual-tours-and-its-radial-case}}
 
-The manual tour (Cook and Buja 1997) defines its basis path by manipulating the basis contribution of a selected variable.
-A manipulation dimension is appended onto the projection plane, with a full contribution given to the selected variable.
+The manual tour (Cook and Buja 1997) defines its basis path by manipulating a selected variable's basis contribution.
+A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
 The bases are then selected based on rotating this newly created manipulation space.
 A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
-This means that such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
-However, this navigation is relatively time-consuming due to the huge volume of \(p\)-space.
+Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
+However, this navigation is relatively time-consuming due to the vast volume of the display-space.
 It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
@@
 
-More generally, the manual tour can change the contribution of the variable in any of the display dimensions.
-We will with a specific more directed interaction, namely, a \emph{radial} tour.
-In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution radius.
+More generally, the manual tour can change the contribution of a variable to the display dimensions.
+We will apply a more directed interaction, namely, a \emph{radial} tour.
+In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
 
 \hypertarget{sec:applicationdesign}{%
-\section{Application design}\label{sec:applicationdesign}}
+\section{cheem viewer}\label{sec:applicationdesign}}
 
-Below we illustrate the two primary displays of the application: the global view and the tour view.
+Below we illustrate the two primary displays of the cheem viewer application: the global view and the tour view.
 Then we cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing that before display.
+Lastly, we discuss the preprocessing the data before application runtime.
 
@@
 
-The global view provides an essential context of all observations and allows exploration of the separability of the data- and attribution-spaces.
-While ultimately, its purpose is to facilitate the selection of a primary point and comparison point.
+The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+While comparison of these spaces is interesting, the global view's purpose is to facilitate the selection of observations.
+The local explanation of these points will be explored in more detail.
 
 An approximation of these spaces is given as the first two principal components of their respective spaces.
-This is shown side-by-side next to model information, the prediction, and the observed response variable.
+Model information, observed response by its prediction, is also provided.
 The orientation and magnitude of the variables are inscribed on a unit circle.
-While a single 2D projection will rarely encompass all of the structure of higher-dimensional spaces, it is a reasonable summarization, given the real task at hand, the selection of observations to explore further.
-
-It is insightful to explore these two approximations against a visual of the model; prediction by observation is also displayed.
-Linked brushing and preserved aesthetic features such as circling misclassified observations help link information from the different spaces together.
+Misclassified observations are circled in red if applicable.
+Linked brushing between the plots and tabular display of select points facilitates exploration of the spaces and the model.
+A single 2D projection will not encompass all of the structure of higher-dimensional space.
+However, it is a reasonable summarization given the real task at hand; the selection of observations to explore further.
 
@@
 The global view facilitated the selection of a primary and optional comparison observation.
-The variable-level attribution of primary observation is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observations prediction, not the local shape of the model surface.
+The variable-level attribution of the primary observation is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observation's prediction, not the local shape of the model surface.
 
@@
 The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
-The default variable selected has the largest discrepancy between the primary and comparison observations attribution.
-In the following sections, we elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
-
-Now that we have introduced the global view and corresponding cheem radial tour, let's discuss the differences between the classification and regression cases.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The following sections elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
+Now that we have introduced the global view and corresponding cheem radial tour let us discuss the differences between the classification and regression cases.
 
@@
 \caption{Display illustrating the classification case.
-Both views are colored on predicted class and while red circles identify misclassified observations.
+Plots are  colored on predicted class and red circles indicate misclassified observations.
 The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
@@
 
-The regression case, we opt to color the global view on a statistic, the observation's residual, the log Mahanabis distance in data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
-In the radial tour, the horizontal positions are the same, the attribution projection, then varying as the radial tour changes basis.
-While the vertical position is fixed to the observed response variable and residuals for the middle and right panels, respectively.
-This changes the display from 1D density to a scatterplot.
-The basis is still one component, the horizontal position, independent of the vertical position.
+In the regression case, the global view can be colored on a statistic to highlight the structure in the explanation space, including residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
+The vertical position is fixed to the observed response variable and residuals in the middle and right panels, respectively.
+Correspondingly, the display changes from univariate density to 2D scatterplot.
+The basis is still one component (horizontal) independent of the vertical position.
 
@@
 
-\caption{Display of the regression case.
+\caption{Display of the regression task.
 The global view can be colored on the correlation of the attribution projection and observed response.
-In the tour, the horizontal values are the same as the classification case.
-The vertical position is fixed to the observed response and residuals in the middle and right facets, respectively.}\label{fig:regressioncase}
+In the tour, the horizontal values are the same as the classification case; the projection through the basis.
+The vertical position is now maped to the observed y and residuals.}\label{fig:regressioncase}
 \end{figure}
@@
 
-The application has several reactive selections that affect the data, coloring, and inputs for the tour to improve flexibility by extending the use cases.
-It also has interactive features in the global view that aid the analysis.
+The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
+These reactive inputs make the software flexible and extensible.
+The application also has It also has more exploratory interactions to help link points in the data and extract other information not on plot axes.
 
-A tooltip displays while the cursor hovers over a point displays the observation number/name and classification information if appropriate.
-Linked brushing allows for the selection of points (with left click and drag) where those points will be highlighted in both plots.
+A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
 The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and ultimately the selection of the primary and comparison observations.
+These interactions aid exploration of the spaces and, finally, identification of a primary and comparison observation.
 
@@
 This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over a point}\label{fig:interactions}
+Tooltips display when hovering over an obersvation.}\label{fig:interactions}
 \end{figure}
@@
     \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
-We apply tree SHAP, a more computationally efficient variant of SHAP, that is applicable to tree-based models.
-This is done with the package \textbf{treeshap} [Kominsarczyk et al.
+We apply tree SHAP, a more computationally efficient variant of SHAP, compatible with tree-based models.
+Tree SHAP was calculated with the package \textbf{treeshap} [Kominsarczyk et al.
 (2021), hosted on GitHub only].
@@
 
-The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX include cran URL after published}.
-Preprocessing was facilitated with models created via \textbf{randomForest} {[}liaw\_classification\_2002{]}, and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
+The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
+Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002), and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
 2021).
@@
 2021).
-The tour visual is an extension of \textbf{spinifex} (Spyrison and Cook 2020).
+The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
 Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
-\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.
+\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a huge boon to understanding local explanations and how to apply them.
 
-Installation and get started with the package can be acheived by running the following in \textbf{R}:
+\hypertarget{installation-and-getting-started}{%
+\subsection{Installation and getting started}\label{installation-and-getting-started}}
+
+The following \textbf{R} code will help getting up and running:
 
@@
 \FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
-\DocumentationTok{\#\# Restart the R session to pick up the file structure may be needed}
+\DocumentationTok{\#\# Restart the R session so the IDE has the correct directory structure}
 \FunctionTok{restartSession}\NormalTok{()}
@@
 \FunctionTok{run\_app}\NormalTok{()}
-\DocumentationTok{\#\# Bring your own data: follow the examples in cheem\_ls()}
+
+\CommentTok{\# Processing your data}
+\DocumentationTok{\#\# Install treeshap from github, to use as a local explainer}
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{) }\DocumentationTok{\#\# Local }
+\DocumentationTok{\#\# Follow the examples in cheem\_ls()}
 \NormalTok{?cheem\_ls}
@@
 The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
-The primary observation is in-between the predicted and observed clusters, a sign of uncertainty in the prediction.
+The primary observation is between the predicted and observed clusters, a sign of uncertainty in the prediction.
 The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
@@
 A random forest model is fit classifying the type of chocolate.
-We selected a chocolate labeled dark, through predicted to be milk chocolate with a comparison with chocolate labeled 85\% cocoa.
+We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85\% cocoa.
 
@@
 
-In figure \ref{fig:casechocolates}, we similarly see that attribution space is more separable relative to data-space.
-Interestingly there is not the class imbalance that we suspected; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
+In figure \ref{fig:casechocolates}, we similarly see that attribution space is more separable than data-space.
+Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for the explanation of the offensive player.
+If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for explaining the offensive player.
 We vary the contribution of offensive skills.
-In the tour (3rd frame of b), offensive skills moved, and Messi is no longer separated from the group.
+In the tour (panel b, frame 3), offensive skills are removed, and Messi is no longer separated from the group.
 We also notice that accuracy has rotated into the frame, maintaining some separability.
@@
 We apply it to the classification and regression tasks.
-We have created an open-source \textbf{R} package \textbf{cheem}, available on CRAN, to facilitate preprocessing and exploration with the described interactive application.
+We have created an open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate preprocessing and exploration with the described interactive application.
 Toy and real data are provided, or upload your data after preprocessing.
@@
 This research was supported by Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this piece.
 
 The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations that are specific to tree-based models.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
 
@@
 
-\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
-Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
-
-\leavevmode\hypertarget{ref-strumbelj_explaining_2014}{}%
-Štrumbelj, Erik, and Igor Kononenko. 2014. {``Explaining Prediction Models and Individual Predictions with Feature Contributions.''} \emph{Knowledge and Information Systems} 41 (3): 647--65.
-
 \leavevmode\hypertarget{ref-vanni_textual_2018}{}%

### 2021-12-22 _paper.rmd
interrogate to scrutinize

@@
 ---
-title: "Interrogating the linear variable importance of local explanations of non-linear models with animated linear projections"
+title: "Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections"
 shorttitle: Cheem
@@
 #   plain:     "Methods for understanding the variable importance of local explanations of black-box models"
-#   short:     "Use of the manual tour to visually interrogate local explanations"
+#   short:     "cheem: radial tour scrutinizing local explanations"
 editor_options:
@@
 <!-- Purposed approach -->
-The proposed approach uses the radial, manual tour to interrogate a local explanation.
+The proposed approach uses the radial, manual tour to scrutinize a local explanation.
 After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
@@
 Tree SHAP only compatible with tree-based models; we illustrate on random forests.
-In the following section, we will use normalized explanations as the starting projection basis to further interrogate the explanation.
+In the following section, we will use normalized explanations as the starting projection basis to further scrutinize the explanation.
 

### 2021-12-22 _paper.tex
interrogate to scrutinize

@@
 \hypersetup{
-  pdftitle={Interrogating the linear variable importance of local explanations of non-linear models with animated linear projections},
+  pdftitle={Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections},
   pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
@@
 
-\title{Interrogating the linear variable importance of local explanations of non-linear models with animated linear projections}
+\title{Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections}
 \author{}
@@
 
-The proposed approach uses the radial, manual tour to interrogate a local explanation.
+The proposed approach uses the radial, manual tour to scrutinize a local explanation.
 After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.

### 2021-12-29 _paper.rmd
-.mp4 of the case studies, - added mp4 url and embelishment to case studies. - need to validate that linkes work.

@@
 [Tours and the radial tour](#sec:tour) digs deeper into these animations of continuous linear projections.
-The section [Application Design](#sec:applicationdesign) discusses the visual layouts, how they facilitates analysis, data preprocessing, and package infrastructure.
+The section [Application Design](#sec:applicationdesign) discusses the visual layouts, how they facilitate analysis, data preprocessing, and package infrastructure.
 The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
@@
 These instance-level explanations are used in various ways depending on the data.
-In images, saliency maps overlay or offset a heatmap indicating which pixels were necessary [@simonyan_deep_2014].
+In images, saliency maps overlay or offset a heatmap indicating necessary pixels [@simonyan_deep_2014].
 For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
@@
 Panel b) shows a simplified break-down plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
-Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a), by decreasing sum of the SHAP values.
+Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a) by decreasing the  sum of the SHAP values.
 
@@
 In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
-In application we apply _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
-Tree SHAP only compatible with tree-based models; we illustrate on random forests.
-In the following section, we will use normalized explanations as the starting projection basis to further scrutinize the explanation.
+In application, we apply _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
+Tree SHAP only compatible with tree-based models; we illustrate random forests.
+The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
 
@@
 <!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its basis path by manipulating a selected variable's basis contribution.
+The manual tour [@cook_manual_1997] defines its basis path by manipulating a selected variable's contribution to the basis.
 A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
@@
 Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis [@karwowski_international_2006].
-However, this navigation is relatively time-consuming due to the vast volume of the display-space.
+However, this navigation is relatively time-consuming due to the vast volume of the display space.
 It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
@@
 Then we cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing the data before application runtime.
+Lastly, we discuss the preprocessing of the data before application runtime.
 
@@
 The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
-While comparison of these spaces is interesting, the global view's purpose is to facilitate the selection of observations.
+While the comparison of these spaces is interesting, the global view's purpose is to enable the selection of observations.
 The local explanation of these points will be explored in more detail.
@@
 An approximation of these spaces is given as the first two principal components of their respective spaces.
-Model information, observed response by its prediction, is also provided.
+Model information, the observed response by its prediction, is also provided.
 The orientation and magnitude of the variables are inscribed on a unit circle.
@@
 The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
-The following sections elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
-<!-- Segue to classification -->Now that we have introduced the global view and corresponding cheem radial tour let us discuss the differences between the classification and regression cases.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
+The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The following sections elaborate on the takeaways from applying this approach in classification and regression tasks, respectively.
+<!-- Segue to classification -->Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
 
@@
 The initial frame is the linear attribution of that observation's local explanation.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
-That is, we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation; we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
 
 ```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Display illustrating the classification case.
-Plots are  colored on predicted class and red circles indicate misclassified observations.
+Plots are  colored on predicted class, and red circles indicate misclassified observations.
 The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
@@
 In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
-The vertical position is fixed to the observed response variable and residuals in the middle and right panels, respectively.
+The vertical position is fixed to the observed response variable and residuals in the middle and right panels.
 Correspondingly, the display changes from univariate density to 2D scatterplot.
@@
 In the tour, the horizontal values are the same as the classification case; the projection through the basis.
-The vertical position is now maped to the observed y and residuals."}
+The vertical position is now mapped to the observed y and residuals."}
 knitr::include_graphics("./figures/app_regression.PNG")
@@
 These reactive inputs make the software flexible and extensible.
-The application also has It also has more exploratory interactions to help link points in the data and extract other information not on plot axes.
+The application also has more exploratory interactions to help link points in the data and extract other information, not on plot axes.
 
@@
 This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an obersvation."}
+Tooltips display when hovering over an observation."}
 knitr::include_graphics("./figures/app_interactions.PNG")
@@
 It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
-In order to keep render time low all of the computational operations should be prepared before runtime.
+To keep render time low, all computational operations should be prepared before runtime.
 The work remaining at runtime is solely reacting to inputs and rendering of visuals and tables.
@@
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
-An increased of model complexity or data dimensionality will quickly become an obstacle.
+An increase in model complexity or data dimensionality will quickly become an obstacle.
 This makes tree SHAP, with its reduced computational complexity, a good candidate to start with.
@@
 
-```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Species classification of Palmer penguin data."}
+(ref:casepenguins) Species classification of Palmer penguin data.
+We select an a chinstrap penguin that is mislabeled as an adelie.
+By varying the contribution to bill length, we that the explanation does not hold when bill length has a large contribution.
+The .mp4 animation of this tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
+```{r casepenguins, echo=F, out.width = "100%", fig.cap = "(ref:casepenguins)"}
 knitr::include_graphics("./figures/case_penguins.png")
@@
 Each of which was labeled as being either milk or dark chocolates.
-With this data, we can see if a manufacturer accurately portrays the chocolate.
-We are curious to see if there are chocolates that nutritionally look like milk chocolates labeled as dark chocolates, which may hold a higher market value.
+We can see if a manufacturer accurately portrays the chocolate with this data.
+We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value.
 We should note that not all chocolates consist wholly of chocolate.
@@
 
-```{r casechocolates, echo=F, out.width = "100%", fig.cap = "Chocolates data type classification (milk or dark)."}
+(ref:casechocolates) Chocolates data type classification (milk or dark).
+We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
+We vary the contribution of calories from fat.
+Animated tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
+```{r casechocolates, echo=F, out.width = "100%", fig.cap = "(ref:casechocolates)"}
 knitr::include_graphics("./figures/case_chocolates.png")
@@
 
-In figure \@ref(fig:casechocolates),  we similarly see that attribution space is more separable than data-space.
+Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space.
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
+In this case, it feels a little bit strange to call the selected observation a misclassification of the model.
+There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+This feels more like a measurement theory-related problem.
+In this case, the candy is being sold as dark chocolate, while nutritional value more closely resembles milk chocolates.
 
@@
 
-```{r casefifa, echo=F, out.width = "100%", fig.cap = "FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
+(ref:casefifa) FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
 The primary observation is a star offensive player (L.
 Messi) compared with a top defensive player (V.
-van Dijk)."}
+van Dijk).
+The animate radial tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4)
+```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
 knitr::include_graphics("./figures/case_fifa.png")
@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for explaining the offensive player.
+If we remove reaction and movement skills from the basis, offense skills are almost singularly important for explaining the offensive player.
 We vary the contribution of offensive skills.
-In the tour (panel b, frame 3), offensive skills are removed, and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, maintaining some separability.
+Offensive skills are removed in the tour (panel b, frame 3), and Messi
+is no longer separated from the group. We also notice that accuracy has rotated into the frame, maintaining some separability.
 
@@
 
-```{r caseames, echo=F, out.width = "100%", fig.cap = "Ames housing 2018 regressing log sales price [2018 USD]."}
+(ref:caseames) Ames housing 2018 regressing log sales price [2018 USD].
+Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap of a prediction.
+This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
+The corresponding animation is at [github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
+```{r caseames, echo=F, out.width = "100%", fig.cap = "(ref:caseames)"}
 knitr::include_graphics("./figures/case_ames2018.png")
@@
 Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast and, identify primary and comparison observations of interest.
-The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with radial tour.
-The variable sensitivity to the structure identified in the explanation is explored by the tours.
+First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
+The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.
+The tours explore the variable sensitivity to the structure identified in
+the explanation.
 

### 2021-12-29 _paper.tex
-.mp4 of the case studies, - added mp4 url and embelishment to case studies. - need to validate that linkes work.

@@
 
-\caption{Species classification of Palmer penguin data.}\label{fig:casepenguins}
+\caption{Species classification of Palmer penguin data.
+We select an a chinstrp penguin that is mislabeled as an adlie.
+By varying the cntribution to bill length we that the explaintion does not hold when bill length has a large contribtuion.
+The .mp4 animaion of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}}\label{fig:casepenguins}
 \end{figure}
@@
 
-\caption{Chocolates data type classification (milk or dark).}\label{fig:casechocolates}
+\caption{Chocolates data type classification (milk or dark).
+We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
+We vary the contribution form Calories from Fat.
+Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.}\label{fig:casechocolates}
 \end{figure}
@@
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
+In this case it feels a little bit strange to call the selected observation a misclassification of the model.
+There plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+This feels more like measurement theory related problem.
+In this case the candy is being sold as dark chocolate, while nutritional value more closely resemble milk chocolate.
 
@@
 
-\caption{FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
+\caption{FIFA 2020, regressing log wages {[}2020 Euros{]} from aggregations of skill measurements.
 The primary observation is a star offensive player (L.
 Messi) compared with a top defensive player (V.
-van Dijk).}\label{fig:casefifa}
+van Dijk).
+The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_fifa.mp4}}\label{fig:casefifa}
 \end{figure}
@@
 
-\caption{Ames housing 2018 regressing log sales price [2018 USD].}\label{fig:caseames}
+\caption{Ames housing 2018 regressing log sales price {[}2018 USD{]}.
+Because the SHAP values are relatively well distributed accross the variables, there is a lot of redundant information to explain the gap of a prediction.
+This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looing for, but certainly an interesting feature to find.
+The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}}\label{fig:caseames}
 \end{figure}

### 2021-12-29 figures/_make_figures_case_study.r

@@ -38,7 +38,9 @@
          x = .x_axis_title) + 
     ggtitle("Global view")
   .bas <- basis_attr_df(penguins_ls$attr_df, prim_obs)
-  .mv  <- manip_var_of_attr_df(penguins_ls$attr_df, prim_obs, comp_obs)
+  .mv  <- 1 ## seems better than the 3 from
+  ##manip_var_of_attr_df(penguins_ls$attr_df, prim_obs, comp_obs)
+  ## Cheem tour for stills
   .ggt <- radial_cheem_tour(
     penguins_ls, basis = .bas, manip_var = .mv,
     primary_obs = prim_obs, comparison_obs = comp_obs,
@@ -53,26 +55,38 @@
     ncol = 1)#, rel_heights = c(1.5, 1))
 }
 
-### Save
+### Save still shots for paper
 ggplot2::ggsave(
   "./figures/case_penguins.png",
   plot = .cp, device = "png",
   width = 6, height = 5, units = "in")
 .m <- gc()
 
+### Save .mp4, add GitHub urls to paper
+.ggt <- radial_cheem_tour(
+  penguins_ls, basis = .bas, manip_var = .mv,
+  primary_obs = prim_obs, comparison_obs = comp_obs,
+  do_add_pcp_segments = FALSE,
+  pcp_shape = 124, angle = .15) +## realistic angle
+  theme(legend.position = "top", legend.direction = "horizontal")
+.anim <- animate_gganimate(
+  .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom
+  #height = 10, width = 12, units = "cm", ## "px", "in", "cm", or "mm."
+  render = gganimate::av_renderer("./figures/case_penguins.mp4")) ## Alternative render
+## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4
 
 
 ## Chocolates classification -----
 {
   names(chocolates_ls)
-  prim_obs <- 22L
-  comp_obs <- 34L
+  prim_obs <- 22L #"Dark Chocolate Bar, Lindt, Switzerland"
+  comp_obs <- 34L #"85% Cocoa Dark French Chocolate, Thorntos, UK (2nd) 
   if(F)
-    global_view(chocolates_ls, 
+    global_view(chocolates_ls,
                 22, #"Dark Chocolate Bar, Lindt, Switzerland"
                 49) #"Pure Dark Chocolate, Hershey's, US"
   
[…]

### 2022-01-02 _paper.rmd
Smoothing over Intro, and grammarly edits.

@@
  Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
-XAI attempts to shed light on these models.
-One such approach is _local explanations_.
-A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
-We extract explanations for each observation and approximate data, and this attribution space side-by-side with linked brushing.
+In such models, increased predictive power comes at the cost of the opacity of factor analysis and variable interpretability.
+The loss of interpretability has led to the field of eXplainable AI (XAI).
+XAI attempts to shed light on these models by providing means for interpreting them.
+A _local explanation_ of a non-linear model gives a point-estimate of linear variable importance in the vicinity of one observation.
+After creating a model and exacting explanations for each observation, we explore the original variables, the explanation's attribution, and model information side-by-side with interactively display.
 After identifying an observation of interest, its local explanation is used as a 1D projection basis.
-We then manipulate the magnitude of the variable contributions with the _tour_ technique.
+We then manipulate the magnitude of a variable's contribution with a _tour_ technique.
 The tour animates many projections over small changes in the projection basis as it changes the contribution of one variable.
-Doing so allows a user to visually explore the data space through the lens of this local explanation and test its support.
-The implementation of our framework is available as an __R__ package __cheem__ available on CRAN.
+Doing so allows an analyst to visually explore the data through the lens of this local explanation and test the variable support where the explanation seems valid.
+The interactive application and code facilitating this methodology is available are the __R__ package __cheem__, available on CRAN.
 # documentclass: jdssv
@@
 
-<!-- History of regression and classification -->
-Mathematically rigorous approaches to predictive modeling are attributed to the least-squares method, over two centuries ago by Legendre and Gauss in 1805 and 1809, respectively.
-In 1886 Francis Galton coined the term _regression_ to refer to continuous, quantitative predictions.
-While _classification_ refers to discrete predictions as introduced by Fisher in 1936.
-
 <!-- Introduce explanatory vs predictive modeling -->
-Breiman and Shmueli [@breiman_statistical_2001; @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose such as hypothesis testing, while _predictive_ modeling predicts new, out-of-sample, observations.
-This distinction draws attention to the divide between interpretable models and black-box models.
+There are different reasons and emphases to fit a model.
+Breiman and Shmueli [@breiman_statistical_2001; @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample observations.
 The intended use has important implications for model selection and development.
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate non-linear models, the additional hindrance to interpretation should not be taken lightly.
+While predictive modeling may opt for more accurate non-linear models.
+The use of black-box models is becoming increasingly common, but not without their share of controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
+However, the loss of interpretation should not be taken lightly.
 
-<!-- XAI & interpretability crisis -->
-Interpretability aside, black-box models are becoming increasingly common, but not without their share of controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
-Black-box models have been known to reflect common biases, including sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable, which is then built into the model.
-Another issue is data drift when new data is outside the support of latent or exogenous explanatory variables.
-Data drift can lead to worse predictions [@lazer_parable_2014; @salzberg_why_2014].
-While typically more resistant to data drift, linear models will similarly find biases in the data.
-However, their transparency goes a long way into exploring and mitigating the biases.
-Such cases highlight the need to make models fair, accountable, ethical, and transparent, which has led to the movement of XAI [@adadi_peeking_2018; @arrieta_explainable_2020].
+<!-- Interpretability important for fighting baises -->
+Interpretability is vital for exploring and protecting against potential biases in any model.
+For instance, models regularly pick up on biases in the training data that have observed influence on the response variable, which is then built into the model.
+Such biases include sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
+Variable-level interpretability of models is essential in the evaluation models for such biases.
+
+<!-- Interpretability for data drift -->
+Another concern is that of data drift.
+Data drift is an unexpected shift in support of the variables.
+Non-linear models are regularly more sensitive to these changes.
+The interpretability of the model means more transparency to see where models' predictions may be plausible or completely unreliable.
 
 <!-- Local explanations -->
-One branch of XAI is local explanations, which take a variable attribution approach to bring transparency to a model.
+Explanabile Artificial Intelligence (XAI) is a recent branch of research that tries to increase the interpretability of black-box models.
+One way to do so is to use _local explanations_, which take a variable attribution approach to bring transparency to a model.
 Local explanations attempt to approximate linear variable importance at the location of one observation.
-There are many such local explanations.
+Even with a linear space or local approximation, the number of terms is usually sizable, thus challenging to visualize comprehensively.
 
 <!-- Data visualization tours -->
-In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear projections of data onto a lower-dimensional space.
 Tours are viewed as an animation over minor changes to the projection basis.
@@
 <!--tours and models --> Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path [@lee_pptree_2013; @da_silva_projection_2021].
+Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
 
@@
 The proposed approach uses the radial, manual tour to scrutinize a local explanation.
-After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
+After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by changing variable contribution with the radial tour.
 We provide a free and open-source R package __cheem__ with an interactive application to facilitate analysis.
-We provide toy and modern datasets case studies for both classification and regression tasks.
+We supply toy and modern datasets case studies for classification and regression tasks.
 
@@
 In application, we apply _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
-Tree SHAP only compatible with tree-based models; we illustrate random forests.
+Tree SHAP is only compatible with tree-based models; we illustrate random forests.
 The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
@@
 The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
-The following sections elaborate on the takeaways from applying this approach in classification and regression tasks, respectively.
+The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
 <!-- Segue to classification -->Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
@@
 
-In the regression case, the global view can be colored on a statistic to highlight the structure in the explanation space, including residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the regression case, the global view can be colored on a statistic to highlight the explanation space's structure.
+For this purpose, we include residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
 In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
@@
 These reactive inputs make the software flexible and extensible.
-The application also has more exploratory interactions to help link points in the data and extract other information, not on plot axes.
+The application also has more exploratory interactions to help link points across displays and reveal structure found in differnt spaces.
 
@@
 ```{r interactions, echo=F, out.width = "100%", fig.cap = "Illustration of data explorations interactions in the global view.
-This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
+This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
 Tooltips display when hovering over an observation."}
@@
 It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
-To keep render time low, all computational operations should be prepared before runtime.
-The work remaining at runtime is solely reacting to inputs and rendering of visuals and tables.
+All computational operations should be prepared before runtime.
+The work remaining when an appication is ran is solely reacting to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-This makes tree SHAP, with its reduced computational complexity, a good candidate to start with.
+With its reduced computational complexity, this makes tree SHAP a good candidate to start with.
 Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
@@
 (ref:casepenguins) Species classification of Palmer penguin data.
-We select an a chinstrap penguin that is mislabeled as an adelie.
-By varying the contribution to bill length, we that the explanation does not hold when bill length has a large contribution.
+We select a chinstrap penguin that is mislabeled as an adelie.
+By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
 The .mp4 animation of this tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
@@
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
-In this case, it feels a little bit strange to call the selected observation a misclassification of the model.
+In this case, it feels strange to call the selected observation a misclassification of the model.
 There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
@@
 (ref:caseames) Ames housing 2018 regressing log sales price [2018 USD].
-Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap of a prediction.
+Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
 This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
@@
 Figure \@ref(fig:caseames) shows the global view and extrema of the tour.
-The horizontal distance in the tour didn't show a significant disparity between our selected points.
+The horizontal distance in the tour did not show a significant disparity between our selected points.
 This is not particularly surprising as most variables have a sizable contribution.
@@
 Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
+First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
 The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.

### 2022-01-02 _paper.tex
Smoothing over Intro, and grammarly edits.

@@
 Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of opaque factor analysis, which has led to the field of eXplainable AI (XAI).
-XAI attempts to shed light on these models.
-One such approach is \emph{local explanations}.
-A local explanation of a model gives a point-estimate of linear variable importance in the vicinity of one observation.
-We extract explanations for each observation and approximate data, and this attribution space side-by-side with linked brushing.
+In such models, increased predictive power comes at the cost of the opacity of factor analysis and variable interpretability.
+The loss of interpretability has led to the field of eXplainable AI (XAI).
+XAI attempts to shed light on these models by providing means for interpreting them.
+A \emph{local explanation} of a non-linear model gives a point-estimate of linear variable importance in the vicinity of one observation.
+After creating a model and exacting explanations for each observation, we explore the original variables, the explanation's attribution, and model information side-by-side with interactively display.
 After identifying an observation of interest, its local explanation is used as a 1D projection basis.
-We then manipulate the magnitude of the variable contributions with the \emph{tour} technique.
+We then manipulate the magnitude of a variable's contribution with a \emph{tour} technique.
 The tour animates many projections over small changes in the projection basis as it changes the contribution of one variable.
-Doing so allows a user to visually explore the data space through the lens of this local explanation and test its support.
-The implementation of our framework is available as an \textbf{R} package \textbf{cheem} available on CRAN.
+Doing so allows an analyst to visually explore the data through the lens of this local explanation and test the variable support where the explanation seems valid.
+The interactive application and code facilitating this methodology is available are the \textbf{R} package \textbf{cheem}, available on CRAN.
 \end{abstract}
@@
 
-Mathematically rigorous approaches to predictive modeling are attributed to the least-squares method, over two centuries ago by Legendre and Gauss in 1805 and 1809, respectively.
-In 1886 Francis Galton coined the term \emph{regression} to refer to continuous, quantitative predictions.
-While \emph{classification} refers to discrete predictions as introduced by Fisher in 1936.
-
-Breiman and Shmueli (Breiman 2001; Shmueli 2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose such as hypothesis testing, while \emph{predictive} modeling predicts new, out-of-sample, observations.
-This distinction draws attention to the divide between interpretable models and black-box models.
+There are different reasons and emphases to fit a model.
+Breiman and Shmueli (Breiman 2001; Shmueli 2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample observations.
 The intended use has important implications for model selection and development.
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate non-linear models, the additional hindrance to interpretation should not be taken lightly.
+While predictive modeling may opt for more accurate non-linear models.
+The use of black-box models is becoming increasingly common, but not without their share of controversy (O'neil 2016; Kodiyan 2019).
+However, the loss of interpretation should not be taken lightly.
 
-Interpretability aside, black-box models are becoming increasingly common, but not without their share of controversy (O'neil 2016; Kodiyan 2019).
-Black-box models have been known to reflect common biases, including sex (Dastin 2018; Duffy 2019), race (Larson et al.
+Interpretability is vital for exploring and protecting against potential biases in any model.
+For instance, models regularly pick up on biases in the training data that have observed influence on the response variable, which is then built into the model.
+Such biases include sex (Dastin 2018; Duffy 2019), race (Larson et al.
 2016), and age (Díaz et al.
 2018).
-Such issues occur when biases existent in the training data, the model picks up on this influence on the response variable, which is then built into the model.
-Another issue is data drift when new data is outside the support of latent or exogenous explanatory variables.
-Data drift can lead to worse predictions (Lazer et al.
-2014; Salzberg 2014).
-While typically more resistant to data drift, linear models will similarly find biases in the data.
-However, their transparency goes a long way into exploring and mitigating the biases.
-Such cases highlight the need to make models fair, accountable, ethical, and transparent, which has led to the movement of XAI (Adadi and Berrada 2018; Arrieta et al.
-2020).
+Variable-level interpretability of models is essential in the evaluation models for such biases.
 
-One branch of XAI is local explanations, which take a variable attribution approach to bring transparency to a model.
+Another concern is that of data drift.
+Data drift is an unexpected shift in support of the variables.
+Non-linear models are regularly more sensitive to these changes.
+The interpretability of the model means more transparency to see where models' predictions may be plausible or completely unreliable.
+
+Explanabile Artificial Intelligence (XAI) is a recent branch of research that tries to increase the interpretability of black-box models.
+One way to do so is to use \emph{local explanations}, which take a variable attribution approach to bring transparency to a model.
 Local explanations attempt to approximate linear variable importance at the location of one observation.
-There are many such local explanations.
+Even with a linear space or local approximation, the number of terms is usually sizable, thus challenging to visualize comprehensively.
 
@@
 Lee et al.
-2021) is a sequence of linear projections of data onto a lower-dimensional space, typically 1-3D.
+2021) is a sequence of linear projections of data onto a lower-dimensional space.
 Tours are viewed as an animation over minor changes to the projection basis.
@@
 Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis path (Y.
+Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths (Y.
 D.
@@
 The proposed approach uses the radial, manual tour to scrutinize a local explanation.
-After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by the explanation as the contributions of the variables are varied with the radial tour.
+After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by changing variable contribution with the radial tour.
 We provide a free and open-source R package \textbf{cheem} with an interactive application to facilitate analysis.
-We provide toy and modern datasets case studies for both classification and regression tasks.
+We supply toy and modern datasets case studies for classification and regression tasks.
 
@@
 \protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into these animations of continuous linear projections.
-The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the visual layouts, how they facilitates analysis, data preprocessing, and package infrastructure.
+The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the visual layouts, how they facilitate analysis, data preprocessing, and package infrastructure.
 The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
@@
 These instance-level explanations are used in various ways depending on the data.
-In images, saliency maps overlay or offset a heatmap indicating which pixels were necessary (Simonyan, Vedaldi, and Zisserman 2014).
+In images, saliency maps overlay or offset a heatmap indicating necessary pixels (Simonyan, Vedaldi, and Zisserman 2014).
 For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
@@
 Panel b) shows a simplified break-down plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
-Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a), by decreasing sum of the SHAP values.
+Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a) by decreasing the sum of the SHAP values.
 
@@
 In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
-In application we apply \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (S.
+In application, we apply \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (S.
 M.
 Lundberg, Erion, and Lee 2018).
-Tree SHAP only compatible with tree-based models; we illustrate on random forests.
-In the following section, we will use normalized explanations as the starting projection basis to further scrutinize the explanation.
+Tree SHAP is only compatible with tree-based models; we illustrate random forests.
+The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
 
@@
 
-The manual tour (Cook and Buja 1997) defines its basis path by manipulating a selected variable's basis contribution.
+The manual tour (Cook and Buja 1997) defines its basis path by manipulating a selected variable's contribution to the basis.
 A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
@@
 Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
-However, this navigation is relatively time-consuming due to the vast volume of the display-space.
+However, this navigation is relatively time-consuming due to the vast volume of the display space.
 It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
@@
 The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
-The following sections elaborate on the takeaways we draw from applying this approach in classification and regression tasks, respectively.
-Now that we have introduced the global view and corresponding cheem radial tour let us discuss the differences between the classification and regression cases.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
+The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
+Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
 
@@
 The initial frame is the linear attribution of that observation's local explanation.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation.
-That is, we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation; we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
 
@@
 \caption{Display illustrating the classification case.
-Plots are  colored on predicted class and red circles indicate misclassified observations.
+Plots are  colored on predicted class, and red circles indicate misclassified observations.
 The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
@@
 
-In the regression case, the global view can be colored on a statistic to highlight the structure in the explanation space, including residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the regression case, the global view can be colored on a statistic to highlight the explanation space's structure.
+For this purpose, we include residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
 In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
-The vertical position is fixed to the observed response variable and residuals in the middle and right panels, respectively.
+The vertical position is fixed to the observed response variable and residuals in the middle and right panels.
 Correspondingly, the display changes from univariate density to 2D scatterplot.
@@
 In the tour, the horizontal values are the same as the classification case; the projection through the basis.
-The vertical position is now maped to the observed y and residuals.}\label{fig:regressioncase}
+The vertical position is now mapped to the observed y and residuals.}\label{fig:regressioncase}
 \end{figure}
@@
 These reactive inputs make the software flexible and extensible.
-The application also has It also has more exploratory interactions to help link points in the data and extract other information not on plot axes.
+The application also has more exploratory interactions to help link points across displays and reveal structure found in differnt spaces.
 
@@
 \caption{Illustration of data explorations interactions in the global view.
-This view has linked brushing of the points where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an obersvation.}\label{fig:interactions}
+This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
+Tooltips display when hovering over an observation.}\label{fig:interactions}
 \end{figure}
@@
 It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
-In order to keep render time low all of the computational operations should be prepared before runtime.
-The work remaining at runtime is solely reacting to inputs and rendering of visuals and tables.
+All computational operations should be prepared before runtime.
+The work remaining when an appication is ran is solely reacting to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
-An increased of model complexity or data dimensionality will quickly become an obstacle.
-This makes tree SHAP, with its reduced computational complexity, a good candidate to start with.
+An increase in model complexity or data dimensionality will quickly become an obstacle.
+With its reduced computational complexity, this makes tree SHAP a good candidate to start with.
 Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
@@
 \caption{Species classification of Palmer penguin data.
-We select an a chinstrp penguin that is mislabeled as an adlie.
-By varying the cntribution to bill length we that the explaintion does not hold when bill length has a large contribtuion.
-The .mp4 animaion of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}}\label{fig:casepenguins}
+We select a chinstrap penguin that is mislabeled as an adelie.
+By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
+The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}}\label{fig:casepenguins}
 \end{figure}
@@
 We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
-We vary the contribution form Calories from Fat.
+We vary the contribution of calories from fat.
 Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.}\label{fig:casechocolates}
@@
 
-In figure \ref{fig:casechocolates}, we similarly see that attribution space is more separable than data-space.
+Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space.
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
 Calories from fat is the variable with the largest difference in treeshap attribution between these points.
-In this case it feels a little bit strange to call the selected observation a misclassification of the model.
-There plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-This feels more like measurement theory related problem.
-In this case the candy is being sold as dark chocolate, while nutritional value more closely resemble milk chocolate.
+In this case, it feels strange to call the selected observation a misclassification of the model.
+There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+This feels more like a measurement theory-related problem.
+In this case, the candy is being sold as dark chocolate, while nutritional value more closely resembles milk chocolates.
 
@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, then offense skills have almost singular importance for explaining the offensive player.
+If we remove reaction and movement skills from the basis, offense skills are almost singularly important for explaining the offensive player.
 We vary the contribution of offensive skills.
-In the tour (panel b, frame 3), offensive skills are removed, and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, maintaining some separability.
+Offensive skills are removed in the tour (panel b, frame 3), and Messi
+is no longer separated from the group. We also notice that accuracy has rotated into the frame, maintaining some separability.
 
@@
 \caption{Ames housing 2018 regressing log sales price {[}2018 USD{]}.
-Because the SHAP values are relatively well distributed accross the variables, there is a lot of redundant information to explain the gap of a prediction.
-This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looing for, but certainly an interesting feature to find.
+Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
+This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
 The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}}\label{fig:caseames}
@@
 Figure \ref{fig:caseames} shows the global view and extrema of the tour.
-The horizontal distance in the tour didn't show a significant disparity between our selected points.
+The horizontal distance in the tour did not show a significant disparity between our selected points.
 This is not particularly surprising as most variables have a sizable contribution.
@@
 Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data and explanation spaces side-by-side, using dynamic interaction to compare and contrast and, identify primary and comparison observations of interest.
-The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with radial tour.
-The variable sensitivity to the structure identified in the explanation is explored by the tours.
+First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
+The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.
+The tours explore the variable sensitivity to the structure identified in
+the explanation.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
-Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
-
 \leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
@@
 
-\leavevmode\hypertarget{ref-lazer_parable_2014}{}%
-Lazer, David, Ryan Kennedy, Gary King, and Alessandro Vespignani. 2014. {``The Parable of {Google} {Flu}: Traps in Big Data Analysis.''} \emph{Science} 343 (6176): 1203--5.
-
 \leavevmode\hypertarget{ref-lee_state_2021}{}%
@@
 \url{http://arxiv.org/abs/1602.04938}.
-
-\leavevmode\hypertarget{ref-salzberg_why_2014}{}%
-Salzberg, Steven. 2014. {``Why {Google} {Flu} {Is} {A} {Failure}.''} \emph{Forbes}, March. \url{https://www.forbes.com/sites/stevensalzberg/2014/03/23/why-google-flu-is-a-failure/}.
 

### 2022-01-02 _paper.rmd
Couple laste touch up, and linting while refreshing thesis_ns

@@
 
-<!-- Interpretability important for fighting baises -->
+<!-- Interpretability & baises -->
 Interpretability is vital for exploring and protecting against potential biases in any model.
@@
 
-<!-- Interpretability for data drift -->
+<!-- Interpretability & data drift -->
 Another concern is that of data drift.
@@
 These reactive inputs make the software flexible and extensible.
-The application also has more exploratory interactions to help link points across displays and reveal structure found in differnt spaces.
+The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.
 
@@
 
-```{r interactions, echo=F, out.width = "100%", fig.cap = "Illustration of data explorations interactions in the global view.
+```{r interactions, echo=F, out.width="100%", fig.cap = "Illustration of data explorations interactions in the global view.
 This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
@@
 All computational operations should be prepared before runtime.
-The work remaining when an appication is ran is solely reacting to inputs and rendering of visuals and tables.
+The work remaining when an application is ran is solely reacting to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 The .mp4 animation of this tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
-```{r casepenguins, echo=F, out.width = "100%", fig.cap = "(ref:casepenguins)"}
+```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
 knitr::include_graphics("./figures/case_penguins.png")
@@
 Animated tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
-```{r casechocolates, echo=F, out.width = "100%", fig.cap = "(ref:casechocolates)"}
+```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
 knitr::include_graphics("./figures/case_chocolates.png")
@@
 We vary the contribution of offensive skills.
-Offensive skills are removed in the tour (panel b, frame 3), and Messi
-is no longer separated from the group. We also notice that accuracy has rotated into the frame, maintaining some separability.
+Offensive skills are removed in the tour (panel b, frame 3), and Messi is no longer separated from the group.
+We also notice that accuracy has rotated into the frame, maintaining some separability.
 
@@
 The corresponding animation is at [github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
-```{r caseames, echo=F, out.width = "100%", fig.cap = "(ref:caseames)"}
+```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
 knitr::include_graphics("./figures/case_ames2018.png")
@@
 This means that there is a basis orthogonal to this point that describes sizable variation.
-Knowing these singular bases can
-point toward others with meaningful data variation.
+Knowing these singular bases can point toward others with meaningful data variation.
 
-# Discussion {#sec:discussion}
+# Discussion {#sec:cheemdiscussion}
 
@@
 The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.
-The tours explore the variable sensitivity to the structure identified in
-the explanation.
+The tours explore the variable sensitivity to the structure identified in the explanation.
 
@@
 This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this piece.
+Thanks to Jieyang Chong for helping proofread this article.
 

### 2022-01-02 _paper.tex
Couple laste touch up, and linting while refreshing thesis_ns

@@
 This means that there is a basis orthogonal to this point that describes sizable variation.
-Knowing these singular bases can
-point toward others with meaningful data variation.
+Knowing these singular bases can point toward others with meaningful data variation.
 
-\hypertarget{sec:discussion}{%
-\section{Discussion}\label{sec:discussion}}
+\hypertarget{sec:cheemdiscussion}{%
+\section{Discussion}\label{sec:cheemdiscussion}}
 

### 2022-01-02 _paper.rmd
typo

@@
 <!-- Local explanations -->
-Explanabile Artificial Intelligence (XAI) is a recent branch of research that tries to increase the interpretability of black-box models.
+Explainable Artificial Intelligence (XAI) is a recent branch of research that tries to increase the interpretability of black-box models.
 One way to do so is to use _local explanations_, which take a variable attribution approach to bring transparency to a model.

### 2022-01-05 _paper.rmd
updates to title, abstract and intro

@@
 ---
-title: "Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections"
+title: "Exploring Local Explanations of Non-linear Models Using Animated Linear Projections"
 shorttitle: Cheem
@@
 #- name: "Nicholas Spyrison"
-#     | Monash University
-#     | Faculty of Information Technology
 #   address: |
@@
 abstract: |
- Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of the opacity of factor analysis and variable interpretability.
-The loss of interpretability has led to the field of eXplainable AI (XAI).
-XAI attempts to shed light on these models by providing means for interpreting them.
-A _local explanation_ of a non-linear model gives a point-estimate of linear variable importance in the vicinity of one observation.
-After creating a model and exacting explanations for each observation, we explore the original variables, the explanation's attribution, and model information side-by-side with interactively display.
-After identifying an observation of interest, its local explanation is used as a 1D projection basis.
-We then manipulate the magnitude of a variable's contribution with a _tour_ technique.
-The tour animates many projections over small changes in the projection basis as it changes the contribution of one variable.
-Doing so allows an analyst to visually explore the data through the lens of this local explanation and test the variable support where the explanation seems valid.
-The interactive application and code facilitating this methodology is available are the __R__ package __cheem__, available on CRAN.
+ Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models.
+The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
+XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation.
+These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (footbal salaries, house prices) response variables.
+The methods are implemented in the __R__ package __cheem__, available on CRAN.
 # documentclass: jdssv
-classoption: article
+#classoption: article
 output:
-  bookdown::pdf_document2:
-  #pdf_document:
+  #bookdown::pdf_document2:
+  pdf_document:
     keep_tex: true
@@
 ---
-```{r include=FALSE, cache=FALSE}
-require("knitr")
-require("kableExtra")
-require("magrittr")
+
+```{r loadlibraries, include=FALSE, cache=FALSE}
+library("knitr")
+library("kableExtra")
+library("magrittr")
 ## Work packages
-require("cheem")
-require("spinifex")
+library("cheem")
+library("spinifex")
 ## chunk options
@@
 There are different reasons and emphases to fit a model.
-Breiman and Shmueli [@breiman_statistical_2001; @shmueli_explain_2010] introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample observations.
+@breiman_statistical_2001 and @shmueli_explain_2010 introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample observations.
 The intended use has important implications for model selection and development.
@@
 The use of black-box models is becoming increasingly common, but not without their share of controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
-However, the loss of interpretation should not be taken lightly.
+However, the loss of interpretation presents a challenge.
 
@@
 <!-- Interpretability & data drift -->
-Another concern is that of data drift.
-Data drift is an unexpected shift in support of the variables.
-Non-linear models are regularly more sensitive to these changes.
-The interpretability of the model means more transparency to see where models' predictions may be plausible or completely unreliable.
+Another concern is that of data drift, which is a shift in support or domain of the explanatory variables.
+Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
+Better interpretability of the model means that there is more transparency where models' predictions may be plausible or completely unreliable.
 
 <!-- Local explanations -->
-Explainable Artificial Intelligence (XAI) is a recent branch of research that tries to increase the interpretability of black-box models.
-One way to do so is to use _local explanations_, which take a variable attribution approach to bring transparency to a model.
-Local explanations attempt to approximate linear variable importance at the location of one observation.
-Even with a linear space or local approximation, the number of terms is usually sizable, thus challenging to visualize comprehensively.
+Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
+A common approach is to use _local explanations_, which attempt to approximate linear variable importance at the location of each observation, or the predictions at a specific point in the data domain.
+Because these are point specific, a challenge is to visualize to more comprehensively understand a model.
 
@@
 <!-- Purposed approach -->
-The proposed approach uses the radial, manual tour to scrutinize a local explanation.
-After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by changing variable contribution with the radial tour.
-We provide a free and open-source R package __cheem__ with an interactive application to facilitate analysis.
-We supply toy and modern datasets case studies for classification and regression tasks.
+In our proposed approach we use the radial manual tour to scrutinize a local explanation.
+Additional interactivity allows the user to identify an observation of interest, then explore its local explanation by changing variable contribution with the radial tour.
+The methods are implemented in R package __cheem__.
+Example datasets are provided to illustrate usage for classification and regression tasks.
 
 <!-- What-if & counterfactual analysis -->
-The change in the projection basis might feel similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020].
-This phrase, Latin for “other things held constant” or “all else unchanged”, shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
+Using a radial tour can be considered similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020].
+This phrase, Latin for "other things held constant" or "all else unchanged", shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
 In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
-Another difference is that the basis must remain orthonormal.
-That is to say, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+A constraint in our approach is that the basis must remain orthonormal.
+That means, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions, and is thus an interpretable visualisation.
 
@@
 The remainder of this paper is organized as follows.
-The following section, [Local explanation statistics](#sec:explanations), covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
-[Tours and the radial tour](#sec:tour) digs deeper into these animations of continuous linear projections.
-The section [Application Design](#sec:applicationdesign) discusses the visual layouts, how they facilitate analysis, data preprocessing, and package infrastructure.
-The section [Case Studies](#sec:casestudies) illustrates several applications of this method.
-We conclude with a [Discussion](#sec:cheemdiscussion) of the insights we draw from classification and regression tasks.
+The following section, [Local explanation statistics](#sec:explanations), covers the background of the local explanation, and the traditional visuals produced.
+[Tours and the radial tour](#sec:tour) explains the animations of continuous linear projections.
+The section [Application Design](#sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
+The section [Case Studies](#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response variables.
+We conclude with a [Discussion](#sec:cheemdiscussion) of the insights gained and directions that might be explore in the future.
+(XXX DC Can you change the references to number of the section not title?)
 
@@
 The figure includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016; @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017] and their variants are popular.
+There are fewer model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP [@lundberg_unified_2017] and their variants are popular.
 

### 2022-01-06 _paper.rmd
done with SHAP and local explanations

@@
 There are different reasons and emphases to fit a model.
-@breiman_statistical_2001 and @shmueli_explain_2010 introduce the idea of distinguishing modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample observations.
+@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomizes modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample observations.
 The intended use has important implications for model selection and development.
@@
 <!-- Interpretability & baises -->
-Interpretability is vital for exploring and protecting against potential biases in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response variable, which is then built into the model.
-Such biases include sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018].
-Variable-level interpretability of models is essential in the evaluation models for such biases.
+Interpretability is vital for exploring and protecting against potential biases (e.g.
+sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]) in any model.
+For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) variable, which is then built into the model.
+Variable-level (feature-level) interpretability of models is essential in the evaluation models for such biases.
+It is also generally important for many problems, where it is not enough to accurately predict accurately but one must be able to explain which predictors are most responsible in generating a response value.
 
 <!-- Interpretability & data drift -->
-Another concern is that of data drift, which is a shift in support or domain of the explanatory variables.
+Another concern is that of data drift, which is a shift in support or domain of the explanatory variables (feature or predictors).
 Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
@@
 Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
-A common approach is to use _local explanations_, which attempt to approximate linear variable importance at the location of each observation, or the predictions at a specific point in the data domain.
-Because these are point specific, a challenge is to visualize to more comprehensively understand a model.
+A common approach is to use _local explanations_, which attempt to approximate linear variable importance at the location of each observation (instance), or the predictions at a specific point in the data domain.
+Because these are point specific, a challenge is visualizing them to more comprehensively understand a model.
+
+XXX DC Note that we need to make sure that the paper is readable for a machine learner, so a translation of terms is important.
+I've done this by parenthesizing the ML term in this introduction, which should be enough.
 
@@
 The remainder of this paper is organized as follows.
-The following section, [Local explanation statistics](#sec:explanations), covers the background of the local explanation, and the traditional visuals produced.
+The following section, [Local explanations](#sec:explanations) covers the background of the local explanation, and the traditional visuals produced.
 [Tours and the radial tour](#sec:tour) explains the animations of continuous linear projections.
@@
 
-# Local explanation statistics {#sec:explanations}
+# Local explanations {#sec:explanations}
 
@@
 Consider a highly non-linear model.
-At face value, it is hard to say which variable(s) are sensitive to  crossing a classification boundary or identify which variables caused an observation to have a relatively extreme residual.
-Local explanations shed light on these cases by approximating linear variable importance in the vicinity of one observation.
+It can be hard to determine whether small changes in a variable's value will make a class prediction change group or identify which variables contribute to an extreme residual.
+Local explanations shed light on these situations by approximating linear variable importance in the vicinity of a single observation or point.
 
 <!-- Taxonomy of local explanations -->
-Figure six of @arrieta_explainable_2020 gives comprehensive summarization of the taxonomy and literature of explanation techniques.
-The figure includes a large number of model-specific explanations such as deepLIFT, [@shrikumar_not_2016; @shrikumar_learning_2017] a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP [@lundberg_unified_2017] and their variants are popular.
+A comprehensive summary of the taxonomy and literature of explanation techniques is provided in Figure 6 of @arrieta_explainable_2020.
+It includes a large number of model-specific explanations such as deepLIFT [@shrikumar_not_2016; @shrikumar_learning_2017], a popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic explanations, of which LIME [@ribeiro_why_2016] SHAP [@lundberg_unified_2017], and their variants are popular.
 
@@
 These instance-level explanations are used in various ways depending on the data.
-In images, saliency maps overlay or offset a heatmap indicating necessary pixels [@simonyan_deep_2014].
-For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+In image classification, where pixels would correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels [@simonyan_deep_2014].
+For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
 In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
@@
 
-## SHAP and tree SHAP
+<!-- ## SHAP SHAP and history -->
+SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation conceptually by examining the effect of other variables on the contribution of the variable of interest on predicting the response.
+This explanations almost all point to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
+@strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
+However, it is also related to partial dependence plots (XXX Needs reference), used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+It could also be considered to be similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each variable in the context of all other candidate variables.
 
-<!-- SHAP and history -->
-SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory, where @shapley_value_1953 devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
+For our application, we use _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
+Tree SHAP is only compatible with tree-based models; we illustrate random forests.
+The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
 
 <!-- Fifa example -->
-To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season [@leone_fifa_2020].
-We have 5000 observations of nine skill measures (after aggregating highly correlated variables).
-A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
+XXX DC does this example come from/motivated by https://ema.drwhy.ai/UseCaseFIFA.html ?? it needs to be cited here. 
+To understand the use SHAP we use soccer data from FIFA 2020 season [@leone_fifa_2020].
+There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
+A random forest model is fit with log wages, in 2020 Euros, as the response, using the skill measures as predictors.
 We then extract the SHAP values of a star offensive player (L.
@@
 van Dijk).
-We expect to see a difference in the attribution of the variable importance across the two positions of the players.
+The results are displayed in Figure \@ref(fig:shapdistrbd).
+We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
+Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of variables is fixed, so the two observations can be more easily compared.
 
-```{r shapdistrbd, echo=F, fig.align='center', fig.cap = "Illustration of the distribution of SHAP attributions and a break-down plot.
-From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
-The players have very different salaries, but a) shows the distributions of 25 permutations in the explanatory variables.
-The medians of these distributions are the final SHAP values.
-The variable importance differs across the exogenous information of player position.
-These explanations make sense; the variable importances seem realistic given the player's positions.
-b) Break-down plots of the observations using their explanations to additively cover the difference between the model intercept and the observation predictions."}
+```{r shapdistrbd, echo=F, fig.align='center', fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries for FIFA 2020 players based on nine predictors corresponding to different skills.
+Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
+Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
+Plot b) is the break-down plot showing the SHAP values.
+It can be learned that reaction is important for explaining both players salaries.
+Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi."}
 knitr::include_graphics("./figures/shap_distr_bd.png")
@@
 
+XXX DC Can you remove "Variable" from the vertical axes?
+And remove reference to "permutations" in the figure text.
+Also, why include interaction and prediction - suggest removing these because there is no information.
+Perhaps switch the order, and put the breakdown plot first?
+
 <!-- Illustration -->
-Figure \@ref(fig:shapdistrbd) shows the SHAP values of these players.
+<!-- Figure \@ref(fig:shapdistrbd) shows the SHAP values of these players.
 Panel a) shows these players receive a sizable difference in wages.
@@
 Panel b) shows a simplified break-down plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
-Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a) by decreasing the sum of the SHAP values.
+Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a) by decreasing the sum of the SHAP values.-->
 
 <!-- Segue -->
-In summary, this figure highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
 In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
-In application, we apply _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
-Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
 
@@
 
-## Manual tours and its radial case
+<!-- ## Manual tours and its radial case-->
 
@@
 
-# cheem viewer {#sec:applicationdesign}
+# The cheem viewer {#sec:applicationdesign}
 

### 2022-01-06 _paper.tex
typo
done with SHAP and local explanations

@@
 \PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
 \documentclass[
-  article]{article}
+]{article}
 \usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \hypersetup{
-  pdftitle={Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections},
-  pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
+  pdftitle={Exploring Local Explanations of Non-linear Models Using Animated Linear Projections},
+  pdfkeywords={R, XAI, explanable artificial intelligence, black-box
+models, local explanations, grand tour, manual tour, radial tour},
   colorlinks=true,
-  linkcolor=red,
-  filecolor=Maroon,
-  citecolor=Blue,
-  urlcolor=blue,
+  linkcolor={red},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={blue},
   pdfcreator={LaTeX via pandoc}}
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 \setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \usepackage{setspace}
@@
 \usepackage{xcolor}
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-\title{Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections}
+\title{Exploring Local Explanations of Non-linear Models Using Animated
+Linear Projections}
 \author{}
@@
 \begin{abstract}
-Artificial Intelligence (AI) has seen a revitalization in recent years from the use of increasingly hard-to-interpret black-box models.
-In such models, increased predictive power comes at the cost of the opacity of factor analysis and variable interpretability.
-The loss of interpretability has led to the field of eXplainable AI (XAI).
-XAI attempts to shed light on these models by providing means for interpreting them.
-A \emph{local explanation} of a non-linear model gives a point-estimate of linear variable importance in the vicinity of one observation.
-After creating a model and exacting explanations for each observation, we explore the original variables, the explanation's attribution, and model information side-by-side with interactively display.
-After identifying an observation of interest, its local explanation is used as a 1D projection basis.
-We then manipulate the magnitude of a variable's contribution with a \emph{tour} technique.
-The tour animates many projections over small changes in the projection basis as it changes the contribution of one variable.
-Doing so allows an analyst to visually explore the data through the lens of this local explanation and test the variable support where the explanation seems valid.
-The interactive application and code facilitating this methodology is available are the \textbf{R} package \textbf{cheem}, available on CRAN.
+Artificial Intelligence (AI) has seen a revitalization in recent years
+with the prevalence of increasingly hard-to-interpret black-box models.
+The increased predictive power comes at the cost of interpretability,
+which has led to the emergence of eXplainable AI (XAI). XAI attempts to
+shed light on how models are using predictors, to arrive at a
+prediction, with a point estimate of the linear variable importance in
+the vicinity of each observation. These are simply linear projections,
+and as such can be further explored interactively to better understand
+the interaction between variables used to make predictions, across the
+predictive model surface. Here we describe interactive linear
+interpolation used for exploration at any observation, and illustrate
+with examples with categorical (penguin species, chocolate types) and
+quantitative (footbal salaries, house prices) response variables. The
+methods are implemented in the \textbf{R} package \textbf{cheem},
+available on CRAN.
 \end{abstract}
@@
 
-There are different reasons and emphases to fit a model.
-Breiman and Shmueli (Breiman 2001; Shmueli 2010) introduce the idea of distinguishing modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample observations.
-The intended use has important implications for model selection and development.
-In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate non-linear models.
-The use of black-box models is becoming increasingly common, but not without their share of controversy (O'neil 2016; Kodiyan 2019).
-However, the loss of interpretation should not be taken lightly.
+There are different reasons and emphases to fit a model. Breiman (2001),
+reiterated by Shmueli (2010), taxonomizes modeling based on its purpose;
+\emph{explanatory} modeling is done for some inferential purpose, while
+\emph{predictive} modeling focuses more on the predictions of
+out-of-sample observations. The intended use has important implications
+for model selection and development. In explanatory modeling,
+interpretability is vital for drawing inferential conclusions. While
+predictive modeling may opt for more accurate non-linear models. The use
+of black-box models is becoming increasingly common, but not without
+their share of controversy (O'Neil 2016; Kodiyan 2019). However, the
+loss of interpretation presents a challenge.
 
-Interpretability is vital for exploring and protecting against potential biases in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response variable, which is then built into the model.
-Such biases include sex (Dastin 2018; Duffy 2019), race (Larson et al.
-2016), and age (Díaz et al.
-2018).
-Variable-level interpretability of models is essential in the evaluation models for such biases.
+Interpretability is vital for exploring and protecting against potential
+biases (e.g.~sex (Dastin 2018; Duffy 2019), race (Larson et al. 2016),
+and age (Díaz et al. 2018)) in any model. For instance, models regularly
+pick up on biases in the training data that have observed influence on
+the response (output) variable, which is then built into the model.
+Variable-level (feature-level) interpretability of models is essential
+in the evaluation models for such biases. It is also generally important
+for many problems, where it is not enough to accurately predict
+accurately but one must be able to explain which predictors are most
+responsible in generating a response value.
 
-Another concern is that of data drift.
-Data drift is an unexpected shift in support of the variables.
-Non-linear models are regularly more sensitive to these changes.
-The interpretability of the model means more transparency to see where models' predictions may be plausible or completely unreliable.
+Another concern is that of data drift, which is a shift in support or
+domain of the explanatory variables (feature or predictors). Non-linear
+models are typically more sensitive, and do not extrapolate well outside
+of the training data domain. Better interpretability of the model means
+that there is more transparency where models' predictions may be
+plausible or completely unreliable.
 
-Explanabile Artificial Intelligence (XAI) is a recent branch of research that tries to increase the interpretability of black-box models.
-One way to do so is to use \emph{local explanations}, which take a variable attribution approach to bring transparency to a model.
-Local explanations attempt to approximate linear variable importance at the location of one observation.
-Even with a linear space or local approximation, the number of terms is usually sizable, thus challenging to visualize comprehensively.
+Explainable Artificial Intelligence (XAI) is an emerging field of
+research that tries to increase the interpretability of black-box
+models. A common approach is to use \emph{local explanations}, which
+attempt to approximate linear variable importance at the location of
+each observation (instance), or the predictions at a specific point in
+the data domain. Because these are point specific, a challenge is
+visualizing them to more comprehensively understand a model.
 
-In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
-Lee et al.
-2021) is a sequence of linear projections of data onto a lower-dimensional space.
-Tours are viewed as an animation over minor changes to the projection basis.
-Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information of the structure and features of the object.
+XXX DC Note that we need to make sure that the paper is readable for a
+machine learner, so a translation of terms is important. I've done this
+by parenthesizing the ML term in this introduction, which should be
+enough.
 
-There are various types of tours distinguished by the generation of projection bases.
-In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected variable.
-Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths (Y.
-D.
-Lee et al.
-2013; Silva, Cook, and Lee 2021).
+In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and
+Asimov 1986; S. Lee et al. 2021) is a sequence of linear projections of
+data onto a lower-dimensional space. Tours are viewed as an animation
+over minor changes to the projection basis. Structure in a projection
+can then be explored visually to see which variables contribute to the
+formation of that structure. The intuition is similar to watching the
+shadow of a hidden 3D object change as the object is rotated; watching
+the shape of the shadow change conveys information of the structure and
+features of the object.
 
-The proposed approach uses the radial, manual tour to scrutinize a local explanation.
-After identifying an observation of interest, its explanation can be evaluated by testing the support of the structure identified by changing variable contribution with the radial tour.
-We provide a free and open-source R package \textbf{cheem} with an interactive application to facilitate analysis.
-We supply toy and modern datasets case studies for classification and regression tasks.
+There are various types of tours distinguished by the generation of
+projection bases. In a \emph{manual} tour (Cook and Buja 1997; Spyrison
+and Cook 2020), the path is defined by changing the contribution of a
+selected variable. Applying tours to models has been done in a couple of
+contexts. Specifically for exploring various statistical model fits and
+classification boundaries (Wickham, Cook, and Hofmann 2015), and using
+tree- and forest-based approaches as a projection pursuit index to
+generate a tour basis paths (Y. D. Lee et al. 2013; da Silva, Cook, and
+Lee 2021).
 
-The change in the projection basis might feel similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
-It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
-Another difference is that the basis must remain orthonormal.
-That is to say, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+In our proposed approach we use the radial manual tour to scrutinize a
+local explanation. Additional interactivity allows the user to identify
+an observation of interest, then explore its local explanation by
+changing variable contribution with the radial tour. The methods are
+implemented in R package \textbf{cheem}. Example datasets are provided
+to illustrate usage for classification and regression tasks.
 
-The remainder of this paper is organized as follows.
-The following section, \protect\hyperlink{sec:explanations}{Local explanation statistics}, covers the background of the local explanation, SHAP, and the traditional visuals produced from it.
-\protect\hyperlink{sec:tour}{Tours and the radial tour} digs deeper into these animations of continuous linear projections.
-The section \protect\hyperlink{sec:applicationdesign}{Application Design} discusses the visual layouts, how they facilitate analysis, data preprocessing, and package infrastructure.
-The section \protect\hyperlink{sec:casestudies}{Case Studies} illustrates several applications of this method.
-We conclude with a \protect\hyperlink{sec:cheemdiscussion}{Discussion} of the insights we draw from classification and regression tasks.
+Using a radial tour can be considered similar to counterfactual, what-if
+analysis, such as \emph{ceteris paribus} (Biecek 2020). This phrase,
+Latin for ``other things held constant'' or ``all else unchanged'',
+shows how an observation's prediction would change from a marginal
+change in one explanatory variable given that other variables are held
+constant. It ignores correlations of the variables and imagines a case
+that was not observed. In contrast, our approach is a geometric
+explanation of the factual; it varies contributions of the variables by
+rotating the basis, a reorientation of the data object. A constraint in
+our approach is that the basis must remain orthonormal. That means, when
+the contribution of one variable decreases, the contributions of others
+necessarily increase such that there is a complete component in that
+direction. This also ensures that what is seen is strictly a
+low-dimensional projection from high-dimensions, and is thus an
+interpretable visualisation.
+
+The remainder of this paper is organized as follows. The following
+section, \protect\hyperlink{sec:explanations}{Local explanations} covers
+the background of the local explanation, and the traditional visuals
+produced. \protect\hyperlink{sec:tour}{Tours and the radial tour}
+explains the animations of continuous linear projections. The section
+\protect\hyperlink{sec:applicationdesign}{Application Design} discusses
+the visual layout in the interactive interface, how they facilitate
+analysis, data preprocessing, and package infrastructure. The section
+\protect\hyperlink{sec:casestudies}{Case Studies} illustrates the
+application to supervised learning with categorical and quantitative
+response variables. We conclude with a
+\protect\hyperlink{sec:cheemdiscussion}{Discussion} of the insights
+gained and directions that might be explore in the future. (XXX DC Can
+you change the references to number of the section not title?)
 
 \hypertarget{sec:explanations}{%
-\section{Local explanation statistics}\label{sec:explanations}}
+\section{Local explanations}\label{sec:explanations}}
 
-Consider a highly non-linear model.
-At face value, it is hard to say which variable(s) are sensitive to crossing a classification boundary or identify which variables caused an observation to have a relatively extreme residual.
-Local explanations shed light on these cases by approximating linear variable importance in the vicinity of one observation.
+Consider a highly non-linear model. It can be hard to determine whether
+small changes in a variable's value will make a class prediction change
+group or identify which variables contribute to an extreme residual.
+Local explanations shed light on these situations by approximating
+linear variable importance in the vicinity of a single observation or
+point.
 
-Figure six of Arrieta et al.
-(2020) gives comprehensive summarization of the taxonomy and literature of explanation techniques.
-The figure includes a large number of model-specific explanations such as deepLIFT, (Shrikumar et al.
-2016; Shrikumar, Greenside, and Kundaje 2017) a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (S.
-Lundberg and Lee 2017) and their variants are popular.
+A comprehensive summary of the taxonomy and literature of explanation
+techniques is provided in Figure 6 of Arrieta et al. (2020). It includes
+a large number of model-specific explanations such as deepLIFT
+(Shrikumar et al. 2016; Shrikumar, Greenside, and Kundaje 2017), a
+popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic explanations, of which LIME (Ribeiro,
+Singh, and Guestrin 2016) SHAP (S. Lundberg and Lee 2017), and their
+variants are popular.
 
-These instance-level explanations are used in various ways depending on the data.
-In images, saliency maps overlay or offset a heatmap indicating necessary pixels (Simonyan, Vedaldi, and Zisserman 2014).
-For instance, snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
-2018).
-In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
+These instance-level explanations are used in various ways depending on
+the data. In image classification, where pixels would correspond to
+predictors, saliency maps overlay or offset a heatmap indicating
+important pixels (Simonyan, Vedaldi, and Zisserman 2014). For instance,
+pixels corresponding to snow may be highlighted when distinguishing if a
+picture contains a wolf or husky. In text analysis, word-level
+contextual sentiment analysis can be used to highlight the sentiment and
+magnitude of influential words (Vanni et al. 2018). In the case of
+numeric regression, they are used to explain variable additive
+contributions from the model intercept to the observation's prediction
+(Ribeiro, Singh, and Guestrin 2016).
 
-\hypertarget{shap-and-tree-shap}{%
-\subsection{SHAP and tree SHAP}\label{shap-and-tree-shap}}
+SHaply Additive exPlanations (SHAP) approximates the variable importance
+in the vicinity of one observation conceptually by examining the effect
+of other variables on the contribution of the variable of interest on
+predicting the response. This explanations almost all point to Shapley
+(1953)'s method to evaluate an individual's contribution to cooperative
+games by permuting the players that contribute to the score. Strumbelj
+and Kononenko (2010) introduced the use of SHAP for local explanations
+in ML models. However, it is also related to partial dependence plots
+(XXX Needs reference), used to explain the effect of a variable by
+predicting the response for a range of values on this variable, after
+fixing the value of all other variables to their mean. It could also be
+considered to be similar to examining the coefficients from all subsets
+regression, as described in Wickham, Cook, and Hofmann (2015), which
+helps to understand the relative importance of each variable in the
+context of all other candidate variables.
 
-SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation by taking the median importance of a subset of permutations in the explanatory variables.
-This idea stems from the field of game theory, where Shapley (1953) devised a method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
+For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a
+lower computational complexity (S. M. Lundberg, Erion, and Lee 2018).
+Tree SHAP is only compatible with tree-based models; we illustrate
+random forests. The following section will use normalized explanations
+as the starting projection basis to further scrutinize the explanation.
 
-To illustrate SHAP and its original use, explaining the difference between the intercept and an observation's prediction, we use soccer data from FIFA 2020 season (Leone 2020).
-We have 5000 observations of nine skill measures (after aggregating highly correlated variables).
-A random forest model is fit to regress the log wages, in 2020 Euros, from the skill measures.
-We then extract the SHAP values of a star offensive player (L.
-Messi) and defensive player (V.
-van Dijk).
-We expect to see a difference in the attribution of the variable importance across the two positions of the players.
+XXX DC does this example come from/motivated by
+\url{https://ema.drwhy.ai/UseCaseFIFA.html} ?? it needs to be cited
+here. To understand the use SHAP we use soccer data from FIFA 2020
+season (Leone 2020). There are 5000 observations of 9 skill measures
+(after aggregating highly correlated variables). A random forest model
+is fit with log wages, in 2020 Euros, as the response, using the skill
+measures as predictors. We then extract the SHAP values of a star
+offensive player (L. Messi) and defensive player (V. van Dijk). The
+results are displayed in Figure @ref(fig:shapdistrbd). We expect to see
+a difference in the attribution of the variable importance across the
+two positions of the players, which would be interpreted as how the
+player's salary depends on this combination of skill sets. Plot (b) is a
+modified breakdown plot (Gosiewska and Biecek 2019) where the order of
+variables is fixed, so the two observations can be more easily compared.
 
@@
 
-\caption{Illustration of the distribution of SHAP attributions and a break-down plot.
-From FIFA 2020 data, a random forest model regresses wages from nine skill attributes for a star offensive and defensive player.
-The players have very different salaries, but a) shows the distributions of 25 permutations in the explanatory variables.
-The medians of these distributions are the final SHAP values.
-The variable importance differs across the exogenous information of player position.
-These explanations make sense; the variable importances seem realistic given the player's positions.
-b) Break-down plots of the observations using their explanations to additively cover the difference between the model intercept and the observation predictions.}\label{fig:shapdistrbd}
+\caption{Illustration of the use of SHAP values for a random forest model for salaries for FIFA 2020 players based on nine predictors corresponding to different skills.
+Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
+Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
+Plot b) is the break-down plot showing the SHAP values.
+It can be learned that reaction is important for explaining both players salaries.
+Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
 \end{figure}
 
-Figure \ref{fig:shapdistrbd} shows the SHAP values of these players.
-Panel a) shows these players receive a sizable difference in wages.
-Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offense and movement skills are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
-We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
-Panel b) shows a simplified break-down plot (Gosiewska and Biecek 2019), where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
-Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a) by decreasing the sum of the SHAP values.
+XXX DC Can you remove ``Variable'' from the vertical axes? And remove
+reference to ``permutations'' in the figure text. Also, why include
+interaction and prediction - suggest removing these because there is no
+information. Perhaps switch the order, and put the breakdown plot first?
 
-In summary, this figure highlights how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
-In application, we apply \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (S.
-M.
-Lundberg, Erion, and Lee 2018).
-Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
+In summary, these plots highlight how local explanations bring
+interpretability to a model, at least in the vicinity of their
+observations. In this instance, two players with different positions
+receive different profiles of variable importance to explain the
+prediction of their wages.
 
@@
 
-A data visualization \emph{tour} animates many linear projections over small changes in the basis.
-One of the critical features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
-There are various types of tours that are distinguished by selecting their basis paths (S.
-Lee et al.
-2021; Cook et al.
-2008).
+A data visualization \emph{tour} animates many linear projections over
+small changes in the basis. One of the critical features of the tour is
+the object permanence of the data points; one can track the relative
+changes of observations as the basis moves, as opposed to discretely
+jumping to an orthogonal view with no intermediate information. There
+are various types of tours that are distinguished by selecting their
+basis paths (S. Lee et al. 2021; Cook et al. 2008).
 
-\hypertarget{manual-tours-and-its-radial-case}{%
-\subsection{Manual tours and its radial case}\label{manual-tours-and-its-radial-case}}
+The manual tour (Cook and Buja 1997) defines its basis path by
+manipulating a selected variable's contribution to the basis. A
+manipulation dimension is appended onto the projection plane, giving a
+full contribution to the chosen variable. The bases are then selected
+based on rotating this newly created manipulation space. A crucial
+feature of the manual tour is that it allows users to control the
+variable contributions of the basis. Such manipulations can be selected
+and queued in advance or selected on the spot for human-in-the-loop
+analysis (Karwowski 2006). However, this navigation is relatively
+time-consuming due to the vast volume of the display space. It is
+advisable to use this method to explore the sensitivity of the variable
+contribution to a previously identified feature of interest. In this
+case, the projection of the normalized explanations is the feature of
+interest.
 
-The manual tour (Cook and Buja 1997) defines its basis path by manipulating a selected variable's contribution to the basis.
-A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
-The bases are then selected based on rotating this newly created manipulation space.
-A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
-Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
-However, this navigation is relatively time-consuming due to the vast volume of the display space.
-It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
-In this case, the projection of the normalized explanations is the feature of interest.
-
-More generally, the manual tour can change the contribution of a variable to the display dimensions.
-We will apply a more directed interaction, namely, a \emph{radial} tour.
-In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+More generally, the manual tour can change the contribution of a
+variable to the display dimensions. We will apply a more directed
+interaction, namely, a \emph{radial} tour. In a radial tour, the
+selected variable is allowed to change its magnitude of contribution but
+not its angle; it must move along the direction of its original
+contribution.
 
 \hypertarget{sec:applicationdesign}{%
-\section{cheem viewer}\label{sec:applicationdesign}}
+\section{The cheem viewer}\label{sec:applicationdesign}}
 
-Below we illustrate the two primary displays of the cheem viewer application: the global view and the tour view.
-Then we cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing of the data before application runtime.
+Below we illustrate the two primary displays of the cheem viewer
+application: the global view and the tour view. Then we cover what we
+take away from the classification and regression tasks. Lastly, we
+discuss the preprocessing of the data before application runtime.
 
@@
 
-The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
-While the comparison of these spaces is interesting, the global view's purpose is to enable the selection of observations.
+The global view provides an essential context of all observations and
+facilitates the exploration of the separability of the data- and
+attribution-spaces. While the comparison of these spaces is interesting,
+the global view's purpose is to enable the selection of observations.
 The local explanation of these points will be explored in more detail.
 
-An approximation of these spaces is given as the first two principal components of their respective spaces.
-Model information, the observed response by its prediction, is also provided.
-The orientation and magnitude of the variables are inscribed on a unit circle.
-Misclassified observations are circled in red if applicable.
-Linked brushing between the plots and tabular display of select points facilitates exploration of the spaces and the model.
-A single 2D projection will not encompass all of the structure of higher-dimensional space.
-However, it is a reasonable summarization given the real task at hand; the selection of observations to explore further.
+An approximation of these spaces is given as the first two principal
+components of their respective spaces. Model information, the observed
+response by its prediction, is also provided. The orientation and
+magnitude of the variables are inscribed on a unit circle. Misclassified
+observations are circled in red if applicable. Linked brushing between
+the plots and tabular display of select points facilitates exploration
+of the spaces and the model. A single 2D projection will not encompass
+all of the structure of higher-dimensional space. However, it is a
+reasonable summarization given the real task at hand; the selection of
+observations to explore further.
 
@@
 
-The global view facilitated the selection of a primary and optional comparison observation.
-The variable-level attribution of the primary observation is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observation's prediction, not the local shape of the model surface.
+The global view facilitated the selection of a primary and optional
+comparison observation. The variable-level attribution of the primary
+observation is normalized and used as the initial 1D basis in a radial
+tour. This is an approximation of the contributions of the linear
+variables that best explain the difference between the model intercept
+and an observation's prediction, not the local shape of the model
+surface.
 
-The initial frame is the normalized SHAP values of the primary observation.
-The current projection basis is depicted as the width of a bar, the variable's contribution to the horizontal axis.
-The normalized values of all observations are shown as vertical parallel coordinate plots.
+The initial frame is the normalized SHAP values of the primary
+observation. The current projection basis is depicted as the width of a
+bar, the variable's contribution to the horizontal axis. The normalized
+values of all observations are shown as vertical parallel coordinate
+plots.
 
-The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
-The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
-The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
-Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
+The radial tour creates a basis path by varying the contribution of a
+selected variable, fully into and out of a projection frame. Doing so
+tests an individual variable's sensitivity to the structure identified
+by the local explanation. The default variable selected has the largest
+discrepancy between the attribution of primary and comparison
+observations. The following sections elaborate on the takeaways from
+applying this approach in classification and regression tasks. Now that
+we have introduced the global view and corresponding cheem radial tour,
+let us discuss the differences between the classification and regression
+cases.
 
@@
 
-What information do we glean from using this method on a classification task?
-Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
-The initial frame is the linear attribution of that observation's local explanation.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation; we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+What information do we glean from using this method on a classification
+task? Typically we select a misclassified observation compared to a
+correctly classified point nearby in data space. The initial frame is
+the linear attribution of that observation's local explanation. By
+default, the manual tour varies the contribution of the variable with
+the largest difference between the primary and comparison observation;
+we can test the sensitivity of each variable to structure identified by
+the local explanation; we are exploring the support of the explanation,
+evaluating the support or robustness of the prediction.
 
@@
 
-In the regression case, the global view can be colored on a statistic to highlight the explanation space's structure.
-For this purpose, we include residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
-In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
-The vertical position is fixed to the observed response variable and residuals in the middle and right panels.
-Correspondingly, the display changes from univariate density to 2D scatterplot.
-The basis is still one component (horizontal) independent of the vertical position.
+In the regression case, the global view can be colored on a statistic to
+highlight the explanation space's structure. For this purpose, we
+include residuals, log Mahalanobis distance of data space (a measure of
+outlyingness), and the correlation of the attribution projection with
+the observed response. In the radial tour, the horizontal positions are
+the same, the basis projection of the radial tour. The vertical position
+is fixed to the observed response variable and residuals in the middle
+and right panels. Correspondingly, the display changes from univariate
+density to 2D scatterplot. The basis is still one component (horizontal)
+independent of the vertical position.
 
@@
 
-The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible.
-The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.
+The application has several reactive inputs that affect the data used,
+aesthetic display, and tour manipulation. These reactive inputs make the
+software flexible and extensible. The application also has more
+exploratory interactions to help link points across displays and reveal
+structure found in different spaces.
 
-A tooltip displays observation number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
-The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and, finally, identification of a primary and comparison observation.
+A tooltip displays observation number/name and classification
+information while the cursor hovers over a point. Linked brushing allows
+the selection of points (left click and drag) where those points will be
+highlighted across plots. The information corresponding to the selected
+points is populated on a dynamic table. These interactions aid
+exploration of the spaces and, finally, identification of a primary and
+comparison observation.
 
@@
 
-It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
-All computational operations should be prepared before runtime.
-The work remaining when an application is ran is solely reacting to inputs and rendering of visuals and tables.
-Below we discuss the steps and details of the reprocessing.
+It is vital to mitigate the render time of visuals, especially when
+users may want to iterate many times. All computational operations
+should be prepared before runtime. The work remaining when an
+application is ran is solely reacting to inputs and rendering of visuals
+and tables. Below we discuss the steps and details of the reprocessing.
+
+(ref:citeRf) Liaw and Wiener (2002) (ref:citeTs) Kominsarczyk et al.
+(2021)
 
@@
     \item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)] for compatibility with the local explanation, which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)] for compatibility with the local explanation, which requires tree-based models.
     \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
 We apply tree SHAP, a more computationally efficient variant of SHAP, compatible with tree-based models.
-Tree SHAP was calculated with the package \textbf{treeshap} [Kominsarczyk et al.
-(2021), hosted on GitHub only].
+Tree SHAP was calculated with the package \textbf{treeshap} [(ref:citeTs), hosted on GitHub only].
 The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
@@
 
-The time to preprocess the data will vary significantly with the model and local explanation.
-For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.9 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 254 seconds combined.
-PCA and statistics of the variables and attributions took 0.6 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
-An increase in model complexity or data dimensionality will quickly become an obstacle.
-With its reduced computational complexity, this makes tree SHAP a good candidate to start with.
-Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+The time to preprocess the data will vary significantly with the model
+and local explanation. For reference, the FIFA data, 5000 observations
+of nine explanatory variables, took 2.9 seconds to fit a random forest
+model of modest hyperparameters. Extracting the tree SHAP values of each
+observation took 254 seconds combined. PCA and statistics of the
+variables and attributions took 0.6 seconds. These runtimes were from a
+non-parallelized R session on a modern laptop, but suffice to say that
+the bulk of the time will be spent on the local attribution. An increase
+in model complexity or data dimensionality will quickly become an
+obstacle. With its reduced computational complexity, this makes tree
+SHAP a good candidate to start with. Alternatively, the package
+\textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which
+are attributed to fewer calls to the prediction function, partial
+implementation in C++, and efficient use of logical subsetting.
 
@@
 
-The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002), and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
-2021).
-The application was made with \textbf{shiny} (Chang et al.
-2021).
-The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
-Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
-\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a huge boon to understanding local explanations and how to apply them.
+The above-described method and application are implemented as an
+open-source \textbf{R} package, \textbf{cheem} available on
+\href{https://CRAN.R-project.org/package=cheem}{CRAN}. Preprocessing was
+facilitated with models created via \textbf{randomForest} (Liaw and
+Wiener 2002), and explanations calculated with \textbf{treeshap}
+(Kominsarczyk et al. 2021). The application was made with \textbf{shiny}
+(Chang et al. 2021). The tour visual is built with \textbf{spinifex}
+(Spyrison and Cook 2020). Both views are created first with first with
+\textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML
+widgets with \textbf{plotly} (Sievert 2020). \textbf{DALEX} (Biecek
+2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and
+Burzykowski 2021) were a huge boon to understanding local explanations
+and how to apply them.
 
 \hypertarget{installation-and-getting-started}{%
-\subsection{Installation and getting started}\label{installation-and-getting-started}}
+\subsection{Installation and getting
+started}\label{installation-and-getting-started}}
 
@@
 
-To illustrate the use of the cheem method, we apply it to modern datasets, two classification examples and then two of regression.
+To illustrate the use of the cheem method, we apply it to modern
+datasets, two classification examples and then two of regression.
 
 \hypertarget{penguin-species-classification}{%
-\subsection{1) Penguin, species classification}\label{penguin-species-classification}}
+\subsection{1) Penguin, species
+classification}\label{penguin-species-classification}}
 
-Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) consist of 330 observations across four physical measurements of three species of penguins foraging near Palmer Station, Antarctica.
-A random forest model was fit, classifying the species of the penguin given the physical measurements.
+Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill,
+and Gorman 2020) consist of 330 observations across four physical
+measurements of three species of penguins foraging near Palmer Station,
+Antarctica. A random forest model was fit, classifying the species of
+the penguin given the physical measurements.
+
+(ref:casepenguins) Species classification of Palmer penguin data. We
+select a chinstrap penguin that is mislabeled as an adelie. By varying
+the contribution to bill length, we observe the explanation does not
+hold when bill length has a significant contribution. The .mp4 animation
+of this tour can be found at
+\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}
 
@@
 
-\caption{Species classification of Palmer penguin data.
-We select a chinstrap penguin that is mislabeled as an adelie.
-By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
-The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}}\label{fig:casepenguins}
+\caption{(ref:casepenguins)}\label{fig:casepenguins}
 \end{figure}
 
-In figure \ref{fig:casepenguins}, a misclassified point is contrasted with a correctly classified point of its observed class nearby in data-space.
-The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
-The primary observation is between the predicted and observed clusters, a sign of uncertainty in the prediction.
-The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.
+In figure @ref(fig:casepenguins), a misclassified point is contrasted
+with a correctly classified point of its observed class nearby in
+data-space. The attribution space from the tree SHAP local explanations
+is a more separable space, where the comparison is squarely in the
+middle of the orange distribution. The primary observation is between
+the predicted and observed clusters, a sign of uncertainty in the
+prediction. The tour varies the contribution of bill length (b\_l) as
+this variable differs most from the contribution of the comparison
+observation. Downplaying the contribution of bill length is crucial to
+the linear explanation of this observation being misclassified.
 
 \hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{2) Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
+\subsection{2) Chocolates, milk/dark chocolate
+classification}\label{chocolates-milkdark-chocolate-classification}}
 
-The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
-Each of which was labeled as being either milk or dark chocolates.
-We can see if a manufacturer accurately portrays the chocolate with this data.
-We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value.
-We should note that not all chocolates consist wholly of chocolate.
-The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable.
-A random forest model is fit classifying the type of chocolate.
-We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85\% cocoa.
+The chocolates dataset consists of 88 observations of 10 nutritional
+measurements from their labels. Each of which was labeled as being
+either milk or dark chocolates. We can see if a manufacturer accurately
+portrays the chocolate with this data. We are curious to see if
+chocolates that nutritionally look like milk chocolates are labeled as
+dark chocolates, which may hold a higher market value. We should note
+that not all chocolates consist wholly of chocolate. The addition of
+other ingredients will decrease the predictive power of the model
+nutritional explanatory variable. A random forest model is fit
+classifying the type of chocolate. We selected a chocolate labeled dark,
+through predicted to be milk chocolate compared with a chocolate labeled
+85\% cocoa.
+
+(ref:casechocolates) Chocolates data type classification (milk or dark).
+We select a chocolate labeled as dark though a random forest model
+predicts it to be milk chocolate in light of the values on the
+nutritional label. We vary the contribution of calories from fat.
+Animated tour can be found at
+\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.
 
@@
 
-\caption{Chocolates data type classification (milk or dark).
-We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
-We vary the contribution of calories from fat.
-Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.}\label{fig:casechocolates}
+\caption{(ref:casechocolates)}\label{fig:casechocolates}
 \end{figure}
 
-Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space.
-Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-Calories from fat is the variable with the largest difference in treeshap attribution between these points.
-In this case, it feels strange to call the selected observation a misclassification of the model.
-There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-This feels more like a measurement theory-related problem.
-In this case, the candy is being sold as dark chocolate, while nutritional value more closely resembles milk chocolates.
+Figure @ref(fig:casechocolates) similarly shows that attribution-space
+is more separable than data-space. Interestingly, the class imbalance
+that we suspected was not observed; there are only six chocolates
+labeled as dark and predicted as milk, while eight of the inverse case.
+Calories from fat is the variable with the largest difference in
+treeshap attribution between these points. In this case, it feels
+strange to call the selected observation a misclassification of the
+model. There are plausible reasons that a manufacturer has incentives to
+cut corners and label their products different than what they are. This
+feels more like a measurement theory-related problem. In this case, the
+candy is being sold as dark chocolate, while nutritional value more
+closely resembles milk chocolates.
 
@@
 
-The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
-The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
-We compare a leading offensive fielder (L.
-Messi) with that of a top defensive fielder (V.
-van Dijk), the same observations were used in figure \ref{fig:shapdistrbd}.
+The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill
+measurements of soccer/football players and wage information. After
+aggregation of the skill measurements, we regress the log wages {[}2020
+euros{]} given just the skill aggregates. The model was fit from 5000
+observations of the nine skill aggregates before being thinned to 500
+players to mitigate occlusion and render time. We compare a leading
+offensive fielder (L. Messi) with that of a top defensive fielder (V.
+van Dijk), the same observations were used in figure
+@ref(fig:shapdistrbd).
+
+(ref:casefifa) FIFA 2020, regressing log wages {[}2020 Euros{]} from
+aggregations of skill measurements. The primary observation is a star
+offensive player (L. Messi) compared with a top defensive player (V. van
+Dijk). The animate radial tour can be found at
+\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_fifa.mp4}
 
@@
 
-\caption{FIFA 2020, regressing log wages {[}2020 Euros{]} from aggregations of skill measurements.
-The primary observation is a star offensive player (L.
-Messi) compared with a top defensive player (V.
-van Dijk).
-The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_fifa.mp4}}\label{fig:casefifa}
+\caption{(ref:casefifa)}\label{fig:casefifa}
 \end{figure}
 
-With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, offense skills are almost singularly important for explaining the offensive player.
-We vary the contribution of offensive skills.
-Offensive skills are removed in the tour (panel b, frame 3), and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, maintaining some separability.
+With figure @ref(fig:casefifa), we will test the premise of the local
+explanation. If we remove reaction and movement skills from the basis,
+offense skills are almost singularly important for explaining the
+offensive player. We vary the contribution of offensive skills.
+Offensive skills are removed in the tour (panel b, frame 3), and Messi
+is no longer separated from the group. We also notice that accuracy has
+rotated into the frame, maintaining some separability.
 
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{4) Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
+\subsection{4) Ames housing 2018, sales price
+regression}\label{ames-housing-2018-sales-price-regression}}
 
-Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across nine variables.
-Using interaction from the global view, we select a house with an extreme negative residual and an accurate observation close to it in the data.
+Ames 2018, housing data was subset to North Ames (the neighborhood with
+the most house sales). The remaining are 338 house sales across nine
+variables. Using interaction from the global view, we select a house
+with an extreme negative residual and an accurate observation close to
+it in the data.
+
+(ref:caseames) Ames housing 2018 regressing log sales price {[}2018
+USD{]}. Because the SHAP values are relatively well distributed across
+the variables, there is a lot of redundant information to explain the
+gap between select observations. This case may not be ideal for the
+cheem methodology, though we did find a near singular frame in the
+manual tour not, something we came looking for, but certainly, an
+exciting feature to find. The corresponding animation is at
+\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}
 
@@
 
-\caption{Ames housing 2018 regressing log sales price {[}2018 USD{]}.
-Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
-This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
-The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}}\label{fig:caseames}
+\caption{(ref:caseames)}\label{fig:caseames}
 \end{figure}
 
-Figure \ref{fig:caseames} shows the global view and extrema of the tour.
-The horizontal distance in the tour did not show a significant disparity between our selected points.
-This is not particularly surprising as most variables have a sizable contribution.
-Rotating any one variable out of the frame will rotate other vital variables into the frame, preserving most of the distance from intercept to prediction.
-However, the tour has revealed an interesting feature worth discussing.
-Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames one and two of panel b) the data is near a singular profile.
-This means that there is a basis orthogonal to this point that describes sizable variation.
-Knowing these singular bases can point toward others with meaningful data variation.
+Figure @ref(fig:caseames) shows the global view and extrema of the tour.
+The horizontal distance in the tour did not show a significant disparity
+between our selected points. This is not particularly surprising as most
+variables have a sizable contribution. Rotating any one variable out of
+the frame will rotate other vital variables into the frame, preserving
+most of the distance from intercept to prediction. However, the tour has
+revealed an interesting feature worth discussing. Notice that the
+observations pivot about the origin, the basis roughly halfway between
+bases in frames one and two of panel b) the data is near a singular
+profile. This means that there is a basis orthogonal to this point that
+describes sizable variation. Knowing these singular bases can point
+toward others with meaningful data variation.
 
@@
 
-The need to maintain the interpretability of black-box models is evident.
-One aspect uses local explanations of the model in the vicinity of an observation.
-Local explanations approximate the linear variable importance to the model.
-Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
-The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.
-The tours explore the variable sensitivity to the structure identified in the explanation.
+The need to maintain the interpretability of black-box models is
+evident. One aspect uses local explanations of the model in the vicinity
+of an observation. Local explanations approximate the linear variable
+importance to the model. Our contribution is to assess explanations by
+examining the support by varying the contributions with a radial tour.
+First, a global view visualizes approximations of the data space,
+explanation space, model predictions side-by-side, using dynamic
+interaction to compare and contrast and identify primary and comparison
+observations of interest. The normalized linear importance from the
+explanation of the primary observation becomes the feature of interest
+to further explore with the radial tour. The tours explore the variable
+sensitivity to the structure identified in the explanation.
 
-We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generally used with any compatible model-explanation pairing.
-We apply it to the classification and regression tasks.
-We have created an open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate preprocessing and exploration with the described interactive application.
-Toy and real data are provided, or upload your data after preprocessing.
+We have illustrated this method on random forest models using the tree
+SHAP local explanation, while it could be generally used with any
+compatible model-explanation pairing. We apply it to the classification
+and regression tasks. We have created an open-source \textbf{R} package
+\textbf{cheem}, available on
+\href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate
+preprocessing and exploration with the described interactive
+application. Toy and real data are provided, or upload your data after
+preprocessing.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
+We would like to thank Professor Przemyslaw Biecek for his input early
+in the project and to the broader \(\text{MI}^\text2\) lab group for the
+\textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
+This research was supported by Australian Government Research Training
+Program (RTP) scholarships. Thanks to Jieyang Chong for helping
+proofread this article.
 
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+The namesake, Cheem, refers to a fictional race of humanoid trees from
+Doctor Who lore. \textbf{DALEX} pulls on from that universe, and we
+initially apply tree SHAP explanations specific to tree-based models.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
-Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
-2020.
-{``Explainable {Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies, Opportunities and Challenges Toward Responsible {AI}.''} \emph{Information Fusion} 58: 82--115.
+\leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
+Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser,
+Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio
+Gil-López, Daniel Molina, and Richard Benjamins. 2020. {``Explainable
+{Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies,
+Opportunities and Challenges Toward Responsible {AI}.''}
+\emph{Information Fusion} 58: 82--115.
 
-\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
-Asimov, Daniel.
-1985.
-{``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
+\leavevmode\vadjust pre{\hypertarget{ref-asimov_grand_1985}{}}%
+Asimov, Daniel. 1985. {``The {Grand} {Tour}: A {Tool} for {Viewing}
+{Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and
+Statistical Computing} 6 (1): 128--43.
 https://doi.org/\url{https://doi.org/10.1137/0906011}.
 
-\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
-Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
+Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive
+Models in {R}.''} \emph{The Journal of Machine Learning Research} 19
+(1): 3245--49.
 
-\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
----------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
+---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus}
+{Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
-\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
-Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
+{Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive}
+{Models}}. CRC Press.
 
-\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
-Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
+\leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
+Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with
+Comments and a Rejoinder by the Author).''} \emph{Statistical Science}
+16 (3): 199--231.
 
-\leavevmode\hypertarget{ref-buja_grand_1986}{}%
-Buja, Andreas, and Daniel Asimov.
-1986.
-{``Grand {Tour} {Methods}: {An} {Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer} {Science} and {Statistics}}, 63--67.
-New York, NY, USA: Elsevier North-Holland, Inc.
-\url{http://dl.acm.org/citation.cfm?id=26036.26046}.
+\leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
+Buja, Andreas, and Daniel Asimov. 1986. {``Grand {Tour} {Methods}: {An}
+{Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on
+the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer}
+{Science} and {Statistics}}, 63--67. New York, NY, USA: Elsevier
+North-Holland, Inc. \url{http://dl.acm.org/citation.cfm?id=26036.26046}.
 
-\leavevmode\hypertarget{ref-chang_shiny_2021}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
-2021.
-\emph{Shiny: Web Application Framework for r}.
+\leavevmode\vadjust pre{\hypertarget{ref-chang_shiny_2021}{}}%
+Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke,
+Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara
+Borges. 2021. \emph{Shiny: Web Application Framework for r}.
 \url{https://CRAN.R-project.org/package=shiny}.
 
-\leavevmode\hypertarget{ref-cook_manual_1997}{}%
-Cook, Dianne, and Andreas Buja.
-1997.
-{``Manual {Controls} for {High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
+\leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
+Cook, Dianne, and Andreas Buja. 1997. {``Manual {Controls} for
+{High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of
+Computational and Graphical Statistics} 6 (4): 464--80.
 \url{https://doi.org/10.2307/1390747}.
 
-\leavevmode\hypertarget{ref-cook_grand_2008}{}%
-Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
-2008.
-{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual} {Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
+\leavevmode\vadjust pre{\hypertarget{ref-cook_grand_2008}{}}%
+Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham. 2008.
+{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual}
+{Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
 Berlin, Heidelberg: Springer Berlin Heidelberg.
@@
 
-\leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
-Dastin, Jeffrey.
-2018.
-{``Amazon Scraps Secret {AI} Recruiting Tool That Showed Bias Against Women.''} \emph{Reuters}, October.
+\leavevmode\vadjust pre{\hypertarget{ref-da_silva_projection_2021}{}}%
+da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee. 2021. {``A
+{Projection} {Pursuit} {Forest} {Algorithm} for {Supervised}
+{Classification}.''} \emph{Journal of Computational and Graphical
+Statistics}, 1--21.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dastin_amazon_2018}{}}%
+Dastin, Jeffrey. 2018. {``Amazon Scraps Secret {AI} Recruiting Tool That
+Showed Bias Against Women.''} \emph{Reuters}, October.
 \url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.
 
-\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
-Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
-2018.
-{``Addressing Age-Related Bias in Sentiment Analysis.''} In \emph{Proceedings of the 2018 Chi Conference on Human Factors in Computing Systems}, 1--14.
+\leavevmode\vadjust pre{\hypertarget{ref-diaz_addressing_2018}{}}%
+Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren
+Gergle. 2018. {``Addressing Age-Related Bias in Sentiment Analysis.''}
+In \emph{Proceedings of the 2018 Chi Conference on Human Factors in
+Computing Systems}, 1--14.
 
-\leavevmode\hypertarget{ref-duffy_apple_2019}{}%
-Duffy, Claire.
-2019.
-{``Apple Co-Founder {Steve} {Wozniak} Says {Apple} {Card} Discriminated Against His Wife.''} \emph{CNN}, November.
+\leavevmode\vadjust pre{\hypertarget{ref-duffy_apple_2019}{}}%
+Duffy, Claire. 2019. {``Apple Co-Founder {Steve} {Wozniak} Says {Apple}
+{Card} Discriminated Against His Wife.''} \emph{CNN}, November.
 \url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
 
-\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
-Gorman, Kristen B., Tony D.
-Williams, and William R.
-Fraser.
-2014.
-{``Ecological Sexual Dimorphism and Environmental Variability Within a Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS One} 9 (3): e90081.
+\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
+Gorman, Kristen B., Tony D. Williams, and William R. Fraser. 2014.
+{``Ecological Sexual Dimorphism and Environmental Variability Within a
+Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS
+One} 9 (3): e90081.
 
-\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
-Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
+\leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
+Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}:
+{Uncertainty} of Model Explanations for Non-Additive Predictive
+Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
-\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
-Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+\leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
+Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley}
+{Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
-\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
-Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
-Gorman.
-2020.
-{``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
+\leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
+Horst, Allison Marie, Alison Presmanes Hill, and Kristen B. Gorman.
+2020. {``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin
+Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
 
-\leavevmode\hypertarget{ref-karwowski_international_2006}{}%
-Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
+\leavevmode\vadjust pre{\hypertarget{ref-karwowski_international_2006}{}}%
+Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of
+{Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
 
-\leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
-Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
+\leavevmode\vadjust pre{\hypertarget{ref-kodiyan_overview_2019}{}}%
+Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using
+{AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring
+Tool.''} \emph{Researchgate Preprint}.
 
-\leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
-Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
+\leavevmode\vadjust pre{\hypertarget{ref-kominsarczyk_treeshap_2021}{}}%
+Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw
+Biecek. 2021. {``Treeshap.''} Model Oriented.
+\url{https://github.com/ModelOriented/treeshap}.
 
-\leavevmode\hypertarget{ref-larson_how_2016}{}%
-Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
-2016.
-{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''} \emph{ProPublica}, May.
+\leavevmode\vadjust pre{\hypertarget{ref-larson_how_2016}{}}%
+Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin. 2016.
+{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''}
+\emph{ProPublica}, May.
 \url{https://www.propublica.org/article/how-we-analyzed-the-compas-recidivism-algorithm?token=RPR1E2qtzJltfJ0tS-gB_41kmfoWZAu4}.
 
-\leavevmode\hypertarget{ref-lee_state_2021}{}%
-Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
-Sherry Zhang.
-2021.
-{``The State-of-the-Art on Tours for Dynamic Visualization of High-Dimensional Data.''} \emph{WIREs Computational Statistics} n/a (n/a): e1573.
+\leavevmode\vadjust pre{\hypertarget{ref-lee_state_2021}{}}%
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas
+Spyrison, Earo Wang, and H. Sherry Zhang. 2021. {``The State-of-the-Art
+on Tours for Dynamic Visualization of High-Dimensional Data.''}
+\emph{WIREs Computational Statistics} n/a (n/a): e1573.
 \url{https://doi.org/10.1002/wics.1573}.
 
-\leavevmode\hypertarget{ref-lee_pptree_2013}{}%
-Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
+\leavevmode\vadjust pre{\hypertarget{ref-lee_pptree_2013}{}}%
+Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013.
+{``{PPtree}: {Projection} Pursuit Classification Tree.''}
+\emph{Electronic Journal of Statistics} 7: 1369--86.
 
-\leavevmode\hypertarget{ref-leone_fifa_2020}{}%
-Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
+\leavevmode\vadjust pre{\hypertarget{ref-leone_fifa_2020}{}}%
+Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''}
+\url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
 
-\leavevmode\hypertarget{ref-liaw_classification_2002}{}%
-Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
+\leavevmode\vadjust pre{\hypertarget{ref-liaw_classification_2002}{}}%
+Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression
+by {randomForest}.''} \emph{R News} 2 (3): 18--22.
 
-\leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
-Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
+\leavevmode\vadjust pre{\hypertarget{ref-lundberg_consistent_2018}{}}%
+Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent
+Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv
+Preprint arXiv:1802.03888}.
 
-\leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
-Lundberg, Scott, and Su-In Lee. 2017. {``A Unified Approach to Interpreting Model Predictions.''} \emph{arXiv Preprint arXiv:1705.07874}.
+\leavevmode\vadjust pre{\hypertarget{ref-lundberg_unified_2017}{}}%
+Lundberg, Scott, and Su-In Lee. 2017. {``A Unified Approach to
+Interpreting Model Predictions.''} \emph{arXiv Preprint
+arXiv:1705.07874}.
 
-\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
-O'neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
+\leavevmode\vadjust pre{\hypertarget{ref-oneil_weapons_2016}{}}%
+O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data
+Increases Inequality and Threatens Democracy}. Crown.
 
-\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
-Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
-2016.
-{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} \emph{arXiv:1602.04938 {[}Cs, Stat{]}}, February.
+\leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
+Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin. 2016. {``"{Why}
+{Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any}
+{Classifier}.''} \emph{arXiv:1602.04938 {[}Cs, Stat{]}}, February.
 \url{http://arxiv.org/abs/1602.04938}.
 
-\leavevmode\hypertarget{ref-shapley_value_1953}{}%
-Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
+\leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
+Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton
+University Press.
 
-\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
-Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
+\leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
+Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical
+Science} 25 (3): 289--310.
 
-\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
-Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
-2017.
-{``Learning Important Features Through Propagating Activation Differences.''} In \emph{International {Conference} on {Machine} {Learning}}, 3145--53.
-PMLR.
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
+Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje. 2017.
+{``Learning Important Features Through Propagating Activation
+Differences.''} In \emph{International {Conference} on {Machine}
+{Learning}}, 3145--53. PMLR.
 
-\leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
-Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
-2016.
-{``Not Just a Black Box: {Learning} Important Features Through Propagating Activation Differences.''} \emph{arXiv Preprint arXiv:1605.01713}.
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_not_2016}{}}%
+Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul
+Kundaje. 2016. {``Not Just a Black Box: {Learning} Important Features
+Through Propagating Activation Differences.''} \emph{arXiv Preprint
+arXiv:1605.01713}.
 
-\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
-Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
+\leavevmode\vadjust pre{\hypertarget{ref-sievert_interactive_2020}{}}%
+Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data}
+{Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC.
+\url{https://plotly-r.com}.
 
-\leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
-Silva, Natalia da, Dianne Cook, and Eun-Kyung Lee.
-2021.
-{``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
+\leavevmode\vadjust pre{\hypertarget{ref-simonyan_deep_2014}{}}%
+Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman. 2014. {``Deep
+Inside Convolutional Networks: {Visualising} Image Classification Models
+and Saliency Maps.''} In \emph{In {Workshop} at {International}
+{Conference} on {Learning} {Representations}}. Citeseer.
 
-\leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
-Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
-2014.
-{``Deep Inside Convolutional Networks: {Visualising} Image Classification Models and Saliency Maps.''} In \emph{In {Workshop} at {International} {Conference} on {Learning} {Representations}}.
-Citeseer.
-
-\leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
-Spyrison, Nicholas, and Dianne Cook.
-2020.
-{``Spinifex: An {R} {Package} for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of {Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
+\leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
+Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An {R} {Package}
+for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of
+{Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
 \url{https://doi.org/10.32614/RJ-2020-027}.
 
-\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
-Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
-2018.
-{``Textual {Deconvolution} {Saliency} ({TDS}): A Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the 56th {Annual} {Meeting} of the {Association} for {Computational} {Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
+\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
+Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation
+of Individual Classifications Using Game Theory.''} \emph{The Journal of
+Machine Learning Research} 11: 1--18.
 
-\leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
-Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
+Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and
+Damon Mayaffre. 2018. {``Textual {Deconvolution} {Saliency} ({TDS}): A
+Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the
+56th {Annual} {Meeting} of the {Association} for {Computational}
+{Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
 
-\leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
-Wickham, Hadley, Dianne Cook, and Heike Hofmann.
-2015.
-{``Visualizing Statistical Models: {Removing} the Blindfold.''} \emph{Statistical Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
+\leavevmode\vadjust pre{\hypertarget{ref-wickham_ggplot2_2016}{}}%
+Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data}
+{Analysis}}. Springer-Verlag New York.
+\url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham_visualizing_2015}{}}%
+Wickham, Hadley, Dianne Cook, and Heike Hofmann. 2015. {``Visualizing
+Statistical Models: {Removing} the Blindfold.''} \emph{Statistical
+Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
 \url{https://doi.org/10.1002/sam.11271}.

### 2022-01-06 _paper.rmd
Figure changes from Di's slack comments.

@@
 # Introduction {#sec:intro}
+
+NS "\\@ref() Not linking now, will have to go back to \\ref{} after changes ...
 
@@
 van Dijk).
-The results are displayed in Figure \@ref(fig:shapdistrbd).
+The results are displayed in Figure \ref{fig:shapdistrbd}.
 We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
@@
 Perhaps switch the order, and put the breakdown plot first?
+
+XXX NS double check for text discussing normalization of shap values or scaling distributions of permutations; Specifically in figure one within each player, we divide all contributions by the norm of the median values.
 

### 2022-01-06 _paper.tex
Figure changes from Di's slack comments.

@@
 ]{article}
+\title{Exploring Local Explanations of Non-linear Models Using Animated
+Linear Projections}
+\author{}
+\date{\vspace{-2.5em}}
+
 \usepackage{amsmath,amssymb}
@@
 
-\title{Exploring Local Explanations of Non-linear Models Using Animated
-Linear Projections}
-\author{}
-\date{\vspace{-2.5em}}
-
 \begin{document}
@@
 
+NS ``\textbackslash{}(\textbf{ref?})() Not linking now, will have to go
+back to \textbackslash ref\{\} after changes \ldots{}
+
 There are different reasons and emphases to fit a model. Breiman (2001),
@@
 analysis, such as \emph{ceteris paribus} (Biecek 2020). This phrase,
-Latin for ``other things held constant'' or ``all else unchanged'',
+Latin for ``other things held constant'' or ``all else unchanged,''
 shows how an observation's prediction would change from a marginal
@@
 offensive player (L. Messi) and defensive player (V. van Dijk). The
-results are displayed in Figure @ref(fig:shapdistrbd). We expect to see
+results are displayed in Figure \ref{fig:shapdistrbd}. We expect to see
 a difference in the attribution of the variable importance across the
@@
 
-{\centering \includegraphics[width=29.17in]{./figures/shap_distr_bd}
+{\centering \includegraphics[width=25in]{./figures/shap_distr_bd}
 
@@
 information. Perhaps switch the order, and put the breakdown plot first?
+
+XXX NS double check for text discussing normalization of shap values or
+scaling distributions of permutations; Specifically in figure one within
+each player, we divide all contributions by the norm of the median
+values.
 
@@
 The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction.
-We probe the support of the variable contributions by selecting a variable to vary the contribution.}\label{fig:classificationcase}
+We probe the support of the variable contributions by selecting a variable to vary the contribution.}(\#fig:classificationcase)
 \end{figure}
@@
 In the tour, the horizontal values are the same as the classification case; the projection through the basis.
-The vertical position is now mapped to the observed y and residuals.}\label{fig:regressioncase}
+The vertical position is now mapped to the observed y and residuals.}(\#fig:regressioncase)
 \end{figure}
@@
 This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an observation.}\label{fig:interactions}
+Tooltips display when hovering over an observation.}(\#fig:interactions)
 \end{figure}
@@
 
-\caption{(ref:casepenguins)}\label{fig:casepenguins}
+\caption{(ref:casepenguins)}(\#fig:casepenguins)
 \end{figure}
@@
 
-\caption{(ref:casechocolates)}\label{fig:casechocolates}
+\caption{(ref:casechocolates)}(\#fig:casechocolates)
 \end{figure}
@@
 
-\caption{(ref:casefifa)}\label{fig:casefifa}
+\caption{(ref:casefifa)}(\#fig:casefifa)
 \end{figure}
@@
 
-\caption{(ref:caseames)}\label{fig:caseames}
+\caption{(ref:caseames)}(\#fig:caseames)
 \end{figure}

### 2022-01-06 figures/_make_figures.r

@@ -1,58 +1,60 @@
 # ch5_fig1_shap_distr_bd ----
-require("DALEX")
-require("dplyr")
-require("ggplot2")
-
-## Local func
-my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
-  ## Remade from: iBreakDown:::print.break_down_uncertainty
-  data.frame(
-    player = player_tag,
-    label  = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
-    variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
-    value  = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
-    ## Of the distribution of local attributions:
-    min    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
-    q1     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
-    median = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), median, na.rm = TRUE),
-    q3     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.75, na.rm = TRUE),
-    max    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), max, na.rm = TRUE))
+{
+  require("DALEX")
+  require("dplyr")
+  require("ggplot2")
+  set.seed(2022)
+  
+  ## Local func
+  my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
+    ## Remade from: iBreakDown:::print.break_down_uncertainty
+    data.frame(
+      player = player_tag,
+      label  = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+      variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
[…]

### 2022-01-06 figures/_make_figures_case_study.r

@@ -52,7 +52,7 @@
   .cp <- cowplot::plot_grid(
     .glob_view, .cheem_stills,
     labels = c("a)", "b)"),
-    ncol = 1)#, rel_heights = c(1.5, 1))
+    ncol = 1, rel_heights = c(1, 1.25))
 }
 
 ### Save still shots for paper
@@ -106,7 +106,7 @@ ggplot2::ggsave(
   .cp <- cowplot::plot_grid(
     .glob_view, .cheem_stills,
     labels = c("a)", "b)"),
-    ncol = 1)#, rel_heights = c(1.5, 1))
+    ncol = 1, rel_heights = c(1, 1.25))
 }
 
 ### Save Stills

### 2022-01-06 _paper.rmd
Working through Di's comments.

--- 2022-01-06 _paper.rmd
@@
 output:
-  #bookdown::pdf_document2:
-  pdf_document:
+  ## bookdown::pdf_document2: allows \@ref()
+  bookdown::pdf_document2:
+  ## pdf_document: requires \ref{} 
+  #pdf_document:
     keep_tex: true
@@
 ```
-
+<!-- bookdown::pdf_document2: allows \@ref(), pdf_document: requires \ref{} -->
 <!-- # cheat sheet {#sec:cheatsheet} -->
@@
 
-XXX DC Note that we need to make sure that the paper is readable for a machine learner, so a translation of terms is important.
-I've done this by parenthesizing the ML term in this introduction, which should be enough.
-
 <!-- Data visualization tours -->
@@
 The remainder of this paper is organized as follows.
-The following section, [Local explanations](#sec:explanations) covers the background of the local explanation, and the traditional visuals produced.
-[Tours and the radial tour](#sec:tour) explains the animations of continuous linear projections.
-The section [Application Design](#sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-The section [Case Studies](#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response variables.
-We conclude with a [Discussion](#sec:cheemdiscussion) of the insights gained and directions that might be explore in the future.
-(XXX DC Can you change the references to number of the section not title?)
+The following Section, \@ref(sec:explanations) covers the background of the local explanation, and the traditional visuals produced.
+Section \@ref(sec:tour) explains the animations of continuous linear projections.
+Section \@ref(sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
+Then Section \@ref(#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response variables.
+We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explore in the future.
 
@@
 
-<!-- ## SHAP SHAP and history -->
+<!-- ## SHAP and history -->
 SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation conceptually by examining the effect of other variables on the contribution of the variable of interest on predicting the response.
@@
 @strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
-However, it is also related to partial dependence plots (XXX Needs reference), used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+However, it is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
 It could also be considered to be similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each variable in the context of all other candidate variables.
@@
 <!-- Fifa example -->
-XXX DC does this example come from/motivated by https://ema.drwhy.ai/UseCaseFIFA.html ?? it needs to be cited here. 
-To understand the use SHAP we use soccer data from FIFA 2020 season [@leone_fifa_2020].
+Following the use case _Explanatory Model Analysis_ [@biecek_explanatory_2021], we use FIFA data to illustrate the use of SHAP.
+Consider soccer data from the FIFA 2020 season [@leone_fifa_2020].
 There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
-A random forest model is fit with log wages, in 2020 Euros, as the response, using the skill measures as predictors.
+A random forest model is fit regressing log wages [2020 Euros], from the skill measures.
 We then extract the SHAP values of a star offensive player (L.
@@
 
-XXX DC Can you remove "Variable" from the vertical axes?
-And remove reference to "permutations" in the figure text.
-Also, why include interaction and prediction - suggest removing these because there is no information.
-Perhaps switch the order, and put the breakdown plot first?
-
-<!-- Illustration -->
-<!-- Figure \@ref(fig:shapdistrbd) shows the SHAP values of these players.
-Panel a) shows these players receive a sizable difference in wages.
-Panel b) shows the underlying distribution of the SHAP attributions while permuting the explanatory variables, with the medians being the SHAP values.
-In the light of the player position, the difference in the variable importance makes sense; offense and movement skills are more important for the offensive player, while defensive and power skills are more informative to the model for explaining the prediction of the defensive player.
-We would likewise expect the profile of variable importance to be unique for star players of other positions, such as goalkeepers or middle fielders.
-Panel b) shows a simplified break-down plot [@gosiewska_ibreakdown_2019], where a local explanation is used to additively explain the difference from the intercept to the observation's prediction.
-Such additive approaches will show asymmetry in the variable ordering, so we opt to fix the order to panel a) by decreasing the sum of the SHAP values.-->
-
 <!-- Segue -->
@@
 There are various types of tours that are distinguished by selecting their basis paths [@lee_state_2021; @cook_grand_2008].
-
-<!-- ## Manual tours and its radial case-->
 

### 2022-01-06 _paper.tex
Working through Di's comments.

--- 2022-01-06 _paper.tex
@@
 ]{article}
+\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
+\author{}
+\date{\vspace{-2.5em}}
+
 \usepackage{amsmath,amssymb}
@@
   pdftitle={Exploring Local Explanations of Non-linear Models Using Animated Linear Projections},
-  pdfkeywords={R, XAI, explanable artificial intelligence, black-box
-models, local explanations, grand tour, manual tour, radial tour},
+  pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
   colorlinks=true,
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 \begin{abstract}
-Artificial Intelligence (AI) has seen a revitalization in recent years
-with the prevalence of increasingly hard-to-interpret black-box models.
-The increased predictive power comes at the cost of interpretability,
-which has led to the emergence of eXplainable AI (XAI). XAI attempts to
-shed light on how models are using predictors, to arrive at a
-prediction, with a point estimate of the linear variable importance in
-the vicinity of each observation. These are simply linear projections,
-and as such can be further explored interactively to better understand
-the interaction between variables used to make predictions, across the
-predictive model surface. Here we describe interactive linear
-interpolation used for exploration at any observation, and illustrate
-with examples with categorical (penguin species, chocolate types) and
-quantitative (footbal salaries, house prices) response variables. The
-methods are implemented in the \textbf{R} package \textbf{cheem},
-available on CRAN.
+Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models.
+The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
+XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation.
+These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (footbal salaries, house prices) response variables.
+The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
 \end{abstract}
@@
 
-There are different reasons and emphases to fit a model. Breiman (2001),
-reiterated by Shmueli (2010), taxonomizes modeling based on its purpose;
-\emph{explanatory} modeling is done for some inferential purpose, while
-\emph{predictive} modeling focuses more on the predictions of
-out-of-sample observations. The intended use has important implications
-for model selection and development. In explanatory modeling,
-interpretability is vital for drawing inferential conclusions. While
-predictive modeling may opt for more accurate non-linear models. The use
-of black-box models is becoming increasingly common, but not without
-their share of controversy (O'Neil 2016; Kodiyan 2019). However, the
-loss of interpretation presents a challenge.
+There are different reasons and emphases to fit a model.
+Breiman (2001), reiterated by Shmueli (2010), taxonomizes modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample observations.
+The intended use has important implications for model selection and development.
+In explanatory modeling, interpretability is vital for drawing inferential conclusions.
+While predictive modeling may opt for more accurate non-linear models.
+The use of black-box models is becoming increasingly common, but not without their share of controversy (O'Neil 2016; Kodiyan 2019).
+However, the loss of interpretation presents a challenge.
 
-Interpretability is vital for exploring and protecting against potential
-biases (e.g.~sex (Dastin 2018; Duffy 2019), race (Larson et al. 2016),
-and age (Díaz et al. 2018)) in any model. For instance, models regularly
-pick up on biases in the training data that have observed influence on
-the response (output) variable, which is then built into the model.
-Variable-level (feature-level) interpretability of models is essential
-in the evaluation models for such biases. It is also generally important
-for many problems, where it is not enough to accurately predict
-accurately but one must be able to explain which predictors are most
-responsible in generating a response value.
+Interpretability is vital for exploring and protecting against potential biases (e.g.~sex (Dastin 2018; Duffy 2019), race (Larson et al.
+2016), and age (Díaz et al.
+2018)) in any model.
+For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) variable, which is then built into the model.
+Variable-level (feature-level) interpretability of models is essential in the evaluation models for such biases.
+It is also generally important for many problems, where it is not enough to accurately predict accurately but one must be able to explain which predictors are most responsible in generating a response value.
 
-Another concern is that of data drift, which is a shift in support or
-domain of the explanatory variables (feature or predictors). Non-linear
-models are typically more sensitive, and do not extrapolate well outside
-of the training data domain. Better interpretability of the model means
-that there is more transparency where models' predictions may be
-plausible or completely unreliable.
+Another concern is that of data drift, which is a shift in support or domain of the explanatory variables (feature or predictors).
+Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
+Better interpretability of the model means that there is more transparency where models' predictions may be plausible or completely unreliable.
 
-Explainable Artificial Intelligence (XAI) is an emerging field of
-research that tries to increase the interpretability of black-box
-models. A common approach is to use \emph{local explanations}, which
-attempt to approximate linear variable importance at the location of
-each observation (instance), or the predictions at a specific point in
-the data domain. Because these are point specific, a challenge is
-visualizing them to more comprehensively understand a model.
+Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
+A common approach is to use \emph{local explanations}, which attempt to approximate linear variable importance at the location of each observation (instance), or the predictions at a specific point in the data domain.
+Because these are point specific, a challenge is visualizing them to more comprehensively understand a model.
 
-XXX DC Note that we need to make sure that the paper is readable for a
-machine learner, so a translation of terms is important. I've done this
-by parenthesizing the ML term in this introduction, which should be
-enough.
+In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
+Lee et al.
+2021) is a sequence of linear projections of data onto a lower-dimensional space.
+Tours are viewed as an animation over minor changes to the projection basis.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information of the structure and features of the object.
 
-In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and
-Asimov 1986; S. Lee et al. 2021) is a sequence of linear projections of
-data onto a lower-dimensional space. Tours are viewed as an animation
-over minor changes to the projection basis. Structure in a projection
-can then be explored visually to see which variables contribute to the
-formation of that structure. The intuition is similar to watching the
-shadow of a hidden 3D object change as the object is rotated; watching
-the shape of the shadow change conveys information of the structure and
-features of the object.
+There are various types of tours distinguished by the generation of projection bases.
+In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected variable.
+Applying tours to models has been done in a couple of contexts.
+Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths (Y.
+D.
+Lee et al.
+2013; da Silva, Cook, and Lee 2021).
 
-There are various types of tours distinguished by the generation of
-projection bases. In a \emph{manual} tour (Cook and Buja 1997; Spyrison
-and Cook 2020), the path is defined by changing the contribution of a
-selected variable. Applying tours to models has been done in a couple of
-contexts. Specifically for exploring various statistical model fits and
-classification boundaries (Wickham, Cook, and Hofmann 2015), and using
-tree- and forest-based approaches as a projection pursuit index to
-generate a tour basis paths (Y. D. Lee et al. 2013; da Silva, Cook, and
-Lee 2021).
+In our proposed approach we use the radial manual tour to scrutinize a local explanation.
+Additional interactivity allows the user to identify an observation of interest, then explore its local explanation by changing variable contribution with the radial tour.
+The methods are implemented in R package \textbf{cheem}.
+Example datasets are provided to illustrate usage for classification and regression tasks.
 
-In our proposed approach we use the radial manual tour to scrutinize a
-local explanation. Additional interactivity allows the user to identify
-an observation of interest, then explore its local explanation by
-changing variable contribution with the radial tour. The methods are
-implemented in R package \textbf{cheem}. Example datasets are provided
-to illustrate usage for classification and regression tasks.
+Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
+This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
+It ignores correlations of the variables and imagines a case that was not observed.
+In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
+A constraint in our approach is that the basis must remain orthonormal.
+That means, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions, and is thus an interpretable visualisation.
 
-Using a radial tour can be considered similar to counterfactual, what-if
-analysis, such as \emph{ceteris paribus} (Biecek 2020). This phrase,
-Latin for ``other things held constant'' or ``all else unchanged'',
-shows how an observation's prediction would change from a marginal
-change in one explanatory variable given that other variables are held
-constant. It ignores correlations of the variables and imagines a case
-that was not observed. In contrast, our approach is a geometric
-explanation of the factual; it varies contributions of the variables by
-rotating the basis, a reorientation of the data object. A constraint in
-our approach is that the basis must remain orthonormal. That means, when
-the contribution of one variable decreases, the contributions of others
-necessarily increase such that there is a complete component in that
-direction. This also ensures that what is seen is strictly a
-low-dimensional projection from high-dimensions, and is thus an
-interpretable visualisation.
-
-The remainder of this paper is organized as follows. The following
-section, \protect\hyperlink{sec:explanations}{Local explanations} covers
-the background of the local explanation, and the traditional visuals
-produced. \protect\hyperlink{sec:tour}{Tours and the radial tour}
-explains the animations of continuous linear projections. The section
-\protect\hyperlink{sec:applicationdesign}{Application Design} discusses
-the visual layout in the interactive interface, how they facilitate
-analysis, data preprocessing, and package infrastructure. The section
-\protect\hyperlink{sec:casestudies}{Case Studies} illustrates the
-application to supervised learning with categorical and quantitative
-response variables. We conclude with a
-\protect\hyperlink{sec:cheemdiscussion}{Discussion} of the insights
-gained and directions that might be explore in the future. (XXX DC Can
-you change the references to number of the section not title?)
+The remainder of this paper is organized as follows.
+The following Section, \ref{sec:explanations} covers the background of the local explanation, and the traditional visuals produced.
+Section \ref{sec:tour} explains the animations of continuous linear projections.
+Section \ref{sec:applicationdesign} discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
+Then Section @ref(\#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response variables.
+We conclude with Section \ref{sec:cheemdiscussion} of the insights gained and directions that might be explore in the future.
 
@@
 
-Consider a highly non-linear model. It can be hard to determine whether
-small changes in a variable's value will make a class prediction change
-group or identify which variables contribute to an extreme residual.
-Local explanations shed light on these situations by approximating
-linear variable importance in the vicinity of a single observation or
-point.
+Consider a highly non-linear model.
+It can be hard to determine whether small changes in a variable's value will make a class prediction change group or identify which variables contribute to an extreme residual.
+Local explanations shed light on these situations by approximating linear variable importance in the vicinity of a single observation or point.
 
-A comprehensive summary of the taxonomy and literature of explanation
-techniques is provided in Figure 6 of Arrieta et al. (2020). It includes
-a large number of model-specific explanations such as deepLIFT
-(Shrikumar et al. 2016; Shrikumar, Greenside, and Kundaje 2017), a
-popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME (Ribeiro,
-Singh, and Guestrin 2016) SHAP (S. Lundberg and Lee 2017), and their
-variants are popular.
+A comprehensive summary of the taxonomy and literature of explanation techniques is provided in Figure 6 of Arrieta et al.
+(2020).
+It includes a large number of model-specific explanations such as deepLIFT (Shrikumar et al.
+2016; Shrikumar, Greenside, and Kundaje 2017), a popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic explanations, of which LIME (Ribeiro, Singh, and Guestrin 2016) SHAP (S.
+Lundberg and Lee 2017), and their variants are popular.
 
-These instance-level explanations are used in various ways depending on
-the data. In image classification, where pixels would correspond to
-predictors, saliency maps overlay or offset a heatmap indicating
-important pixels (Simonyan, Vedaldi, and Zisserman 2014). For instance,
-pixels corresponding to snow may be highlighted when distinguishing if a
-picture contains a wolf or husky. In text analysis, word-level
-contextual sentiment analysis can be used to highlight the sentiment and
-magnitude of influential words (Vanni et al. 2018). In the case of
-numeric regression, they are used to explain variable additive
-contributions from the model intercept to the observation's prediction
-(Ribeiro, Singh, and Guestrin 2016).
+These instance-level explanations are used in various ways depending on the data.
+In image classification, where pixels would correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels (Simonyan, Vedaldi, and Zisserman 2014).
+For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
+2018).
+In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
 
-SHaply Additive exPlanations (SHAP) approximates the variable importance
-in the vicinity of one observation conceptually by examining the effect
-of other variables on the contribution of the variable of interest on
-predicting the response. This explanations almost all point to Shapley
-(1953)'s method to evaluate an individual's contribution to cooperative
-games by permuting the players that contribute to the score. Strumbelj
-and Kononenko (2010) introduced the use of SHAP for local explanations
-in ML models. However, it is also related to partial dependence plots
-(XXX Needs reference), used to explain the effect of a variable by
-predicting the response for a range of values on this variable, after
-fixing the value of all other variables to their mean. It could also be
-considered to be similar to examining the coefficients from all subsets
-regression, as described in Wickham, Cook, and Hofmann (2015), which
-helps to understand the relative importance of each variable in the
-context of all other candidate variables.
+SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation conceptually by examining the effect of other variables on the contribution of the variable of interest on predicting the response.
+This explanations almost all point to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
+Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
+However, it is also related to partial dependence plots (Molnar 2020), used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+It could also be considered to be similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each variable in the context of all other candidate variables.
 
-For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a
-lower computational complexity (S. M. Lundberg, Erion, and Lee 2018).
-Tree SHAP is only compatible with tree-based models; we illustrate
-random forests. The following section will use normalized explanations
-as the starting projection basis to further scrutinize the explanation.
+For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (S.
+M.
+Lundberg, Erion, and Lee 2018).
+Tree SHAP is only compatible with tree-based models; we illustrate random forests.
+The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
 
-XXX DC does this example come from/motivated by
-\url{https://ema.drwhy.ai/UseCaseFIFA.html} ?? it needs to be cited
-here. To understand the use SHAP we use soccer data from FIFA 2020
-season (Leone 2020). There are 5000 observations of 9 skill measures
-(after aggregating highly correlated variables). A random forest model
-is fit with log wages, in 2020 Euros, as the response, using the skill
-measures as predictors. We then extract the SHAP values of a star
-offensive player (L. Messi) and defensive player (V. van Dijk). The
-results are displayed in Figure @ref(fig:shapdistrbd). We expect to see
-a difference in the attribution of the variable importance across the
-two positions of the players, which would be interpreted as how the
-player's salary depends on this combination of skill sets. Plot (b) is a
-modified breakdown plot (Gosiewska and Biecek 2019) where the order of
-variables is fixed, so the two observations can be more easily compared.
+Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), we use FIFA data to illustrate the use of SHAP.
+Consider soccer data from the FIFA 2020 season (Leone 2020).
+There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
+A random forest model is fit regressing log wages {[}2020 Euros{]}, from the skill measures.
+We then extract the SHAP values of a star offensive player (L.
+Messi) and defensive player (V.
+van Dijk).
+The results are displayed in Figure \ref{fig:shapdistrbd}.
+We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
+Plot (b) is a modified breakdown plot (Gosiewska and Biecek 2019) where the order of variables is fixed, so the two observations can be more easily compared.
 
@@
 
-XXX DC Can you remove ``Variable'' from the vertical axes? And remove
-reference to ``permutations'' in the figure text. Also, why include
-interaction and prediction - suggest removing these because there is no
-information. Perhaps switch the order, and put the breakdown plot first?
-
-In summary, these plots highlight how local explanations bring
-interpretability to a model, at least in the vicinity of their
-observations. In this instance, two players with different positions
-receive different profiles of variable importance to explain the
-prediction of their wages.
+In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
 
@@
 
-A data visualization \emph{tour} animates many linear projections over
-small changes in the basis. One of the critical features of the tour is
-the object permanence of the data points; one can track the relative
-changes of observations as the basis moves, as opposed to discretely
-jumping to an orthogonal view with no intermediate information. There
-are various types of tours that are distinguished by selecting their
-basis paths (S. Lee et al. 2021; Cook et al. 2008).
+A data visualization \emph{tour} animates many linear projections over small changes in the basis.
+One of the critical features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
+There are various types of tours that are distinguished by selecting their basis paths (S.
+Lee et al.
+2021; Cook et al.
+2008).
 
-The manual tour (Cook and Buja 1997) defines its basis path by
-manipulating a selected variable's contribution to the basis. A
-manipulation dimension is appended onto the projection plane, giving a
-full contribution to the chosen variable. The bases are then selected
-based on rotating this newly created manipulation space. A crucial
-feature of the manual tour is that it allows users to control the
-variable contributions of the basis. Such manipulations can be selected
-and queued in advance or selected on the spot for human-in-the-loop
-analysis (Karwowski 2006). However, this navigation is relatively
-time-consuming due to the vast volume of the display space. It is
-advisable to use this method to explore the sensitivity of the variable
-contribution to a previously identified feature of interest. In this
-case, the projection of the normalized explanations is the feature of
-interest.
+The manual tour (Cook and Buja 1997) defines its basis path by manipulating a selected variable's contribution to the basis.
+A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
+The bases are then selected based on rotating this newly created manipulation space.
+A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
+Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
+However, this navigation is relatively time-consuming due to the vast volume of the display space.
+It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
+In this case, the projection of the normalized explanations is the feature of interest.
 
-More generally, the manual tour can change the contribution of a
-variable to the display dimensions. We will apply a more directed
-interaction, namely, a \emph{radial} tour. In a radial tour, the
-selected variable is allowed to change its magnitude of contribution but
-not its angle; it must move along the direction of its original
-contribution.
+More generally, the manual tour can change the contribution of a variable to the display dimensions.
+We will apply a more directed interaction, namely, a \emph{radial} tour.
+In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
 
@@
 
-Below we illustrate the two primary displays of the cheem viewer
-application: the global view and the tour view. Then we cover what we
-take away from the classification and regression tasks. Lastly, we
-discuss the preprocessing of the data before application runtime.
+Below we illustrate the two primary displays of the cheem viewer application: the global view and the tour view.
+Then we cover what we take away from the classification and regression tasks.
+Lastly, we discuss the preprocessing of the data before application runtime.
 
@@
 
-The global view provides an essential context of all observations and
-facilitates the exploration of the separability of the data- and
-attribution-spaces. While the comparison of these spaces is interesting,
-the global view's purpose is to enable the selection of observations.
+The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+While the comparison of these spaces is interesting, the global view's purpose is to enable the selection of observations.
 The local explanation of these points will be explored in more detail.
 
-An approximation of these spaces is given as the first two principal
-components of their respective spaces. Model information, the observed
-response by its prediction, is also provided. The orientation and
-magnitude of the variables are inscribed on a unit circle. Misclassified
-observations are circled in red if applicable. Linked brushing between
-the plots and tabular display of select points facilitates exploration
-of the spaces and the model. A single 2D projection will not encompass
-all of the structure of higher-dimensional space. However, it is a
-reasonable summarization given the real task at hand; the selection of
-observations to explore further.
+An approximation of these spaces is given as the first two principal components of their respective spaces.
+Model information, the observed response by its prediction, is also provided.
+The orientation and magnitude of the variables are inscribed on a unit circle.
+Misclassified observations are circled in red if applicable.
+Linked brushing between the plots and tabular display of select points facilitates exploration of the spaces and the model.
+A single 2D projection will not encompass all of the structure of higher-dimensional space.
+However, it is a reasonable summarization given the real task at hand; the selection of observations to explore further.
 
@@
 
-The global view facilitated the selection of a primary and optional
-comparison observation. The variable-level attribution of the primary
-observation is normalized and used as the initial 1D basis in a radial
-tour. This is an approximation of the contributions of the linear
-variables that best explain the difference between the model intercept
-and an observation's prediction, not the local shape of the model
-surface.
+The global view facilitated the selection of a primary and optional comparison observation.
+The variable-level attribution of the primary observation is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observation's prediction, not the local shape of the model surface.
 
-The initial frame is the normalized SHAP values of the primary
-observation. The current projection basis is depicted as the width of a
-bar, the variable's contribution to the horizontal axis. The normalized
-values of all observations are shown as vertical parallel coordinate
-plots.
+The initial frame is the normalized SHAP values of the primary observation.
+The current projection basis is depicted as the width of a bar, the variable's contribution to the horizontal axis.
+The normalized values of all observations are shown as vertical parallel coordinate plots.
 
-The radial tour creates a basis path by varying the contribution of a
-selected variable, fully into and out of a projection frame. Doing so
-tests an individual variable's sensitivity to the structure identified
-by the local explanation. The default variable selected has the largest
-discrepancy between the attribution of primary and comparison
-observations. The following sections elaborate on the takeaways from
-applying this approach in classification and regression tasks. Now that
-we have introduced the global view and corresponding cheem radial tour,
-let us discuss the differences between the classification and regression
-cases.
+The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
+Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
+The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
+Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
 
@@
 
-What information do we glean from using this method on a classification
-task? Typically we select a misclassified observation compared to a
-correctly classified point nearby in data space. The initial frame is
-the linear attribution of that observation's local explanation. By
-default, the manual tour varies the contribution of the variable with
-the largest difference between the primary and comparison observation;
-we can test the sensitivity of each variable to structure identified by
-the local explanation; we are exploring the support of the explanation,
-evaluating the support or robustness of the prediction.
+What information do we glean from using this method on a classification task?
+Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
+The initial frame is the linear attribution of that observation's local explanation.
+By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation; we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
 
@@
 
-In the regression case, the global view can be colored on a statistic to
-highlight the explanation space's structure. For this purpose, we
-include residuals, log Mahalanobis distance of data space (a measure of
-outlyingness), and the correlation of the attribution projection with
-the observed response. In the radial tour, the horizontal positions are
-the same, the basis projection of the radial tour. The vertical position
-is fixed to the observed response variable and residuals in the middle
-and right panels. Correspondingly, the display changes from univariate
-density to 2D scatterplot. The basis is still one component (horizontal)
-independent of the vertical position.
+In the regression case, the global view can be colored on a statistic to highlight the explanation space's structure.
+For this purpose, we include residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
+In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
+The vertical position is fixed to the observed response variable and residuals in the middle and right panels.
+Correspondingly, the display changes from univariate density to 2D scatterplot.
+The basis is still one component (horizontal) independent of the vertical position.
 
@@
 
-The application has several reactive inputs that affect the data used,
-aesthetic display, and tour manipulation. These reactive inputs make the
-software flexible and extensible. The application also has more
-exploratory interactions to help link points across displays and reveal
-structure found in different spaces.
+The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
+These reactive inputs make the software flexible and extensible.
+The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.
 
-A tooltip displays observation number/name and classification
-information while the cursor hovers over a point. Linked brushing allows
-the selection of points (left click and drag) where those points will be
-highlighted across plots. The information corresponding to the selected
-points is populated on a dynamic table. These interactions aid
-exploration of the spaces and, finally, identification of a primary and
-comparison observation.
+A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
+The information corresponding to the selected points is populated on a dynamic table.
+These interactions aid exploration of the spaces and, finally, identification of a primary and comparison observation.
 
@@
 
-It is vital to mitigate the render time of visuals, especially when
-users may want to iterate many times. All computational operations
-should be prepared before runtime. The work remaining when an
-application is ran is solely reacting to inputs and rendering of visuals
-and tables. Below we discuss the steps and details of the reprocessing.
-
-(ref:citeRf) Liaw and Wiener (2002) (ref:citeTs) Kominsarczyk et al.
-(2021)
+It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
+All computational operations should be prepared before runtime.
+The work remaining when an application is ran is solely reacting to inputs and rendering of visuals and tables.
+Below we discuss the steps and details of the reprocessing.
 
@@
     \item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)] for compatibility with the local explanation, which requires tree-based models.
+Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)] for compatibility with the local explanation, which requires tree-based models.
     \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
 We apply tree SHAP, a more computationally efficient variant of SHAP, compatible with tree-based models.
-Tree SHAP was calculated with the package \textbf{treeshap} [(ref:citeTs), hosted on GitHub only].
+Tree SHAP was calculated with the package \textbf{treeshap} [Kominsarczyk et al.
+(2021), hosted on GitHub only].
 The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
@@
 
-The time to preprocess the data will vary significantly with the model
-and local explanation. For reference, the FIFA data, 5000 observations
-of nine explanatory variables, took 2.9 seconds to fit a random forest
-model of modest hyperparameters. Extracting the tree SHAP values of each
-observation took 254 seconds combined. PCA and statistics of the
-variables and attributions took 0.6 seconds. These runtimes were from a
-non-parallelized R session on a modern laptop, but suffice to say that
-the bulk of the time will be spent on the local attribution. An increase
-in model complexity or data dimensionality will quickly become an
-obstacle. With its reduced computational complexity, this makes tree
-SHAP a good candidate to start with. Alternatively, the package
-\textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which
-are attributed to fewer calls to the prediction function, partial
-implementation in C++, and efficient use of logical subsetting.
+The time to preprocess the data will vary significantly with the model and local explanation.
+For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.9 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 254 seconds combined.
+PCA and statistics of the variables and attributions took 0.6 seconds.
+These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
+An increase in model complexity or data dimensionality will quickly become an obstacle.
+With its reduced computational complexity, this makes tree SHAP a good candidate to start with.
+Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 
-The above-described method and application are implemented as an
-open-source \textbf{R} package, \textbf{cheem} available on
-\href{https://CRAN.R-project.org/package=cheem}{CRAN}. Preprocessing was
-facilitated with models created via \textbf{randomForest} (Liaw and
-Wiener 2002), and explanations calculated with \textbf{treeshap}
-(Kominsarczyk et al. 2021). The application was made with \textbf{shiny}
-(Chang et al. 2021). The tour visual is built with \textbf{spinifex}
-(Spyrison and Cook 2020). Both views are created first with first with
-\textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML
-widgets with \textbf{plotly} (Sievert 2020). \textbf{DALEX} (Biecek
-2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and
-Burzykowski 2021) were a huge boon to understanding local explanations
-and how to apply them.
+The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
+Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002), and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
+2021).
+The application was made with \textbf{shiny} (Chang et al.
+2021).
+The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
+Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
+\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a huge boon to understanding local explanations and how to apply them.
 
 \hypertarget{installation-and-getting-started}{%
-\subsection{Installation and getting
-started}\label{installation-and-getting-started}}
+\subsection{Installation and getting started}\label{installation-and-getting-started}}
 
@@
 
-To illustrate the use of the cheem method, we apply it to modern
-datasets, two classification examples and then two of regression.
+To illustrate the use of the cheem method, we apply it to modern datasets, two classification examples and then two of regression.
 
 \hypertarget{penguin-species-classification}{%
-\subsection{1) Penguin, species
-classification}\label{penguin-species-classification}}
+\subsection{1) Penguin, species classification}\label{penguin-species-classification}}
 
-Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill,
-and Gorman 2020) consist of 330 observations across four physical
-measurements of three species of penguins foraging near Palmer Station,
-Antarctica. A random forest model was fit, classifying the species of
-the penguin given the physical measurements.
-
-(ref:casepenguins) Species classification of Palmer penguin data. We
-select a chinstrap penguin that is mislabeled as an adelie. By varying
-the contribution to bill length, we observe the explanation does not
-hold when bill length has a significant contribution. The .mp4 animation
-of this tour can be found at
-\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}
+Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) consist of 330 observations across four physical measurements of three species of penguins foraging near Palmer Station, Antarctica.
+A random forest model was fit, classifying the species of the penguin given the physical measurements.
 
@@
 
-\caption{(ref:casepenguins)}\label{fig:casepenguins}
+\caption{Species classification of Palmer penguin data.
+We select a chinstrap penguin that is mislabeled as an adelie.
+By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
+The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}}\label{fig:casepenguins}
 \end{figure}
 
-In figure @ref(fig:casepenguins), a misclassified point is contrasted
-with a correctly classified point of its observed class nearby in
-data-space. The attribution space from the tree SHAP local explanations
-is a more separable space, where the comparison is squarely in the
-middle of the orange distribution. The primary observation is between
-the predicted and observed clusters, a sign of uncertainty in the
-prediction. The tour varies the contribution of bill length (b\_l) as
-this variable differs most from the contribution of the comparison
-observation. Downplaying the contribution of bill length is crucial to
-the linear explanation of this observation being misclassified.
+In figure \ref{fig:casepenguins}, a misclassified point is contrasted with a correctly classified point of its observed class nearby in data-space.
+The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
+The primary observation is between the predicted and observed clusters, a sign of uncertainty in the prediction.
+The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
+Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.
 
 \hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{2) Chocolates, milk/dark chocolate
-classification}\label{chocolates-milkdark-chocolate-classification}}
+\subsection{2) Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
 
-The chocolates dataset consists of 88 observations of 10 nutritional
-measurements from their labels. Each of which was labeled as being
-either milk or dark chocolates. We can see if a manufacturer accurately
-portrays the chocolate with this data. We are curious to see if
-chocolates that nutritionally look like milk chocolates are labeled as
-dark chocolates, which may hold a higher market value. We should note
-that not all chocolates consist wholly of chocolate. The addition of
-other ingredients will decrease the predictive power of the model
-nutritional explanatory variable. A random forest model is fit
-classifying the type of chocolate. We selected a chocolate labeled dark,
-through predicted to be milk chocolate compared with a chocolate labeled
-85\% cocoa.
-
-(ref:casechocolates) Chocolates data type classification (milk or dark).
-We select a chocolate labeled as dark though a random forest model
-predicts it to be milk chocolate in light of the values on the
-nutritional label. We vary the contribution of calories from fat.
-Animated tour can be found at
-\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.
+The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
+Each of which was labeled as being either milk or dark chocolates.
+We can see if a manufacturer accurately portrays the chocolate with this data.
+We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value.
+We should note that not all chocolates consist wholly of chocolate.
+The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable.
+A random forest model is fit classifying the type of chocolate.
+We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85\% cocoa.
 
@@
 
-\caption{(ref:casechocolates)}\label{fig:casechocolates}
+\caption{Chocolates data type classification (milk or dark).
+We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
+We vary the contribution of calories from fat.
+Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.}\label{fig:casechocolates}
 \end{figure}
 
-Figure @ref(fig:casechocolates) similarly shows that attribution-space
-is more separable than data-space. Interestingly, the class imbalance
-that we suspected was not observed; there are only six chocolates
-labeled as dark and predicted as milk, while eight of the inverse case.
-Calories from fat is the variable with the largest difference in
-treeshap attribution between these points. In this case, it feels
-strange to call the selected observation a misclassification of the
-model. There are plausible reasons that a manufacturer has incentives to
-cut corners and label their products different than what they are. This
-feels more like a measurement theory-related problem. In this case, the
-candy is being sold as dark chocolate, while nutritional value more
-closely resembles milk chocolates.
+Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space.
+Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
+Calories from fat is the variable with the largest difference in treeshap attribution between these points.
+In this case, it feels strange to call the selected observation a misclassification of the model.
+There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+This feels more like a measurement theory-related problem.
+In this case, the candy is being sold as dark chocolate, while nutritional value more closely resembles milk chocolates.
 
@@
 
-The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill
-measurements of soccer/football players and wage information. After
-aggregation of the skill measurements, we regress the log wages {[}2020
-euros{]} given just the skill aggregates. The model was fit from 5000
-observations of the nine skill aggregates before being thinned to 500
-players to mitigate occlusion and render time. We compare a leading
-offensive fielder (L. Messi) with that of a top defensive fielder (V.
-van Dijk), the same observations were used in figure
-@ref(fig:shapdistrbd).
-
-(ref:casefifa) FIFA 2020, regressing log wages {[}2020 Euros{]} from
-aggregations of skill measurements. The primary observation is a star
-offensive player (L. Messi) compared with a top defensive player (V. van
-Dijk). The animate radial tour can be found at
-\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_fifa.mp4}
+The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
+After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
+The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
+We compare a leading offensive fielder (L.
+Messi) with that of a top defensive fielder (V.
+van Dijk), the same observations were used in figure \ref{fig:shapdistrbd}.
 
@@
 
-\caption{(ref:casefifa)}\label{fig:casefifa}
+\caption{FIFA 2020, regressing log wages {[}2020 Euros{]} from aggregations of skill measurements.
+The primary observation is a star offensive player (L.
+Messi) compared with a top defensive player (V.
+van Dijk).
+The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_fifa.mp4}}\label{fig:casefifa}
 \end{figure}
 
-With figure @ref(fig:casefifa), we will test the premise of the local
-explanation. If we remove reaction and movement skills from the basis,
-offense skills are almost singularly important for explaining the
-offensive player. We vary the contribution of offensive skills.
-Offensive skills are removed in the tour (panel b, frame 3), and Messi
-is no longer separated from the group. We also notice that accuracy has
-rotated into the frame, maintaining some separability.
+With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
+If we remove reaction and movement skills from the basis, offense skills are almost singularly important for explaining the offensive player.
+We vary the contribution of offensive skills.
+Offensive skills are removed in the tour (panel b, frame 3), and Messi is no longer separated from the group.
+We also notice that accuracy has rotated into the frame, maintaining some separability.
 
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{4) Ames housing 2018, sales price
-regression}\label{ames-housing-2018-sales-price-regression}}
+\subsection{4) Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
 
-Ames 2018, housing data was subset to North Ames (the neighborhood with
-the most house sales). The remaining are 338 house sales across nine
-variables. Using interaction from the global view, we select a house
-with an extreme negative residual and an accurate observation close to
-it in the data.
-
-(ref:caseames) Ames housing 2018 regressing log sales price {[}2018
-USD{]}. Because the SHAP values are relatively well distributed across
-the variables, there is a lot of redundant information to explain the
-gap between select observations. This case may not be ideal for the
-cheem methodology, though we did find a near singular frame in the
-manual tour not, something we came looking for, but certainly, an
-exciting feature to find. The corresponding animation is at
-\href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}
+Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
+The remaining are 338 house sales across nine variables.
+Using interaction from the global view, we select a house with an extreme negative residual and an accurate observation close to it in the data.
 
@@
 
-\caption{(ref:caseames)}\label{fig:caseames}
+\caption{Ames housing 2018 regressing log sales price {[}2018 USD{]}.
+Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
+This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
+The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}}\label{fig:caseames}
 \end{figure}
 
-Figure @ref(fig:caseames) shows the global view and extrema of the tour.
-The horizontal distance in the tour did not show a significant disparity
-between our selected points. This is not particularly surprising as most
-variables have a sizable contribution. Rotating any one variable out of
-the frame will rotate other vital variables into the frame, preserving
-most of the distance from intercept to prediction. However, the tour has
-revealed an interesting feature worth discussing. Notice that the
-observations pivot about the origin, the basis roughly halfway between
-bases in frames one and two of panel b) the data is near a singular
-profile. This means that there is a basis orthogonal to this point that
-describes sizable variation. Knowing these singular bases can point
-toward others with meaningful data variation.
+Figure \ref{fig:caseames} shows the global view and extrema of the tour.
+The horizontal distance in the tour did not show a significant disparity between our selected points.
+This is not particularly surprising as most variables have a sizable contribution.
+Rotating any one variable out of the frame will rotate other vital variables into the frame, preserving most of the distance from intercept to prediction.
+However, the tour has revealed an interesting feature worth discussing.
+Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames one and two of panel b) the data is near a singular profile.
+This means that there is a basis orthogonal to this point that describes sizable variation.
+Knowing these singular bases can point toward others with meaningful data variation.
 
@@
 
-The need to maintain the interpretability of black-box models is
-evident. One aspect uses local explanations of the model in the vicinity
-of an observation. Local explanations approximate the linear variable
-importance to the model. Our contribution is to assess explanations by
-examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data space,
-explanation space, model predictions side-by-side, using dynamic
-interaction to compare and contrast and identify primary and comparison
-observations of interest. The normalized linear importance from the
-explanation of the primary observation becomes the feature of interest
-to further explore with the radial tour. The tours explore the variable
-sensitivity to the structure identified in the explanation.
+The need to maintain the interpretability of black-box models is evident.
+One aspect uses local explanations of the model in the vicinity of an observation.
+Local explanations approximate the linear variable importance to the model.
+Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
+First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
+The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.
+The tours explore the variable sensitivity to the structure identified in the explanation.
 
-We have illustrated this method on random forest models using the tree
-SHAP local explanation, while it could be generally used with any
-compatible model-explanation pairing. We apply it to the classification
-and regression tasks. We have created an open-source \textbf{R} package
-\textbf{cheem}, available on
-\href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate
-preprocessing and exploration with the described interactive
-application. Toy and real data are provided, or upload your data after
-preprocessing.
+We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generally used with any compatible model-explanation pairing.
+We apply it to the classification and regression tasks.
+We have created an open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate preprocessing and exploration with the described interactive application.
+Toy and real data are provided, or upload your data after preprocessing.
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early
-in the project and to the broader \(\text{MI}^\text2\) lab group for the
-\textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
-This research was supported by Australian Government Research Training
-Program (RTP) scholarships. Thanks to Jieyang Chong for helping
-proofread this article.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
+This research was supported by Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
 
-The namesake, Cheem, refers to a fictional race of humanoid trees from
-Doctor Who lore. \textbf{DALEX} pulls on from that universe, and we
-initially apply tree SHAP explanations specific to tree-based models.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
-Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser,
-Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio
-Gil-López, Daniel Molina, and Richard Benjamins. 2020. {``Explainable
-{Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies,
-Opportunities and Challenges Toward Responsible {AI}.''}
-\emph{Information Fusion} 58: 82--115.
+Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
+2020.
+{``Explainable {Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies, Opportunities and Challenges Toward Responsible {AI}.''} \emph{Information Fusion} 58: 82--115.
 
 \leavevmode\vadjust pre{\hypertarget{ref-asimov_grand_1985}{}}%
-Asimov, Daniel. 1985. {``The {Grand} {Tour}: A {Tool} for {Viewing}
-{Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and
-Statistical Computing} 6 (1): 128--43.
+Asimov, Daniel.
+1985.
+{``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
 https://doi.org/\url{https://doi.org/10.1137/0906011}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
-Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive
-Models in {R}.''} \emph{The Journal of Machine Learning Research} 19
-(1): 3245--49.
+Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
 
 \leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
----------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus}
-{Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
+---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
-Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
-{Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive}
-{Models}}. CRC Press.
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
-Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with
-Comments and a Rejoinder by the Author).''} \emph{Statistical Science}
-16 (3): 199--231.
+Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
 
 \leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
-Buja, Andreas, and Daniel Asimov. 1986. {``Grand {Tour} {Methods}: {An}
-{Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on
-the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer}
-{Science} and {Statistics}}, 63--67. New York, NY, USA: Elsevier
-North-Holland, Inc. \url{http://dl.acm.org/citation.cfm?id=26036.26046}.
+Buja, Andreas, and Daniel Asimov.
+1986.
+{``Grand {Tour} {Methods}: {An} {Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer} {Science} and {Statistics}}, 63--67.
+New York, NY, USA: Elsevier North-Holland, Inc.
+\url{http://dl.acm.org/citation.cfm?id=26036.26046}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-chang_shiny_2021}{}}%
-Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke,
-Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara
-Borges. 2021. \emph{Shiny: Web Application Framework for r}.
+Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
+2021.
+\emph{Shiny: Web Application Framework for r}.
 \url{https://CRAN.R-project.org/package=shiny}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
-Cook, Dianne, and Andreas Buja. 1997. {``Manual {Controls} for
-{High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of
-Computational and Graphical Statistics} 6 (4): 464--80.
+Cook, Dianne, and Andreas Buja.
+1997.
+{``Manual {Controls} for {High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
 \url{https://doi.org/10.2307/1390747}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cook_grand_2008}{}}%
-Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham. 2008.
-{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual}
-{Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
+Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
+2008.
+{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual} {Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
 Berlin, Heidelberg: Springer Berlin Heidelberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-da_silva_projection_2021}{}}%
-da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee. 2021. {``A
-{Projection} {Pursuit} {Forest} {Algorithm} for {Supervised}
-{Classification}.''} \emph{Journal of Computational and Graphical
-Statistics}, 1--21.
+da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee.
+2021.
+{``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
 
 \leavevmode\vadjust pre{\hypertarget{ref-dastin_amazon_2018}{}}%
-Dastin, Jeffrey. 2018. {``Amazon Scraps Secret {AI} Recruiting Tool That
-Showed Bias Against Women.''} \emph{Reuters}, October.
+Dastin, Jeffrey.
+2018.
+{``Amazon Scraps Secret {AI} Recruiting Tool That Showed Bias Against Women.''} \emph{Reuters}, October.
 \url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-diaz_addressing_2018}{}}%
-Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren
-Gergle. 2018. {``Addressing Age-Related Bias in Sentiment Analysis.''}
-In \emph{Proceedings of the 2018 Chi Conference on Human Factors in
-Computing Systems}, 1--14.
+Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
+2018.
+{``Addressing Age-Related Bias in Sentiment Analysis.''} In \emph{Proceedings of the 2018 Chi Conference on Human Factors in Computing Systems}, 1--14.
 
 \leavevmode\vadjust pre{\hypertarget{ref-duffy_apple_2019}{}}%
-Duffy, Claire. 2019. {``Apple Co-Founder {Steve} {Wozniak} Says {Apple}
-{Card} Discriminated Against His Wife.''} \emph{CNN}, November.
+Duffy, Claire.
+2019.
+{``Apple Co-Founder {Steve} {Wozniak} Says {Apple} {Card} Discriminated Against His Wife.''} \emph{CNN}, November.
 \url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
-Gorman, Kristen B., Tony D. Williams, and William R. Fraser. 2014.
-{``Ecological Sexual Dimorphism and Environmental Variability Within a
-Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS
-One} 9 (3): e90081.
+Gorman, Kristen B., Tony D.
+Williams, and William R.
+Fraser.
+2014.
+{``Ecological Sexual Dimorphism and Environmental Variability Within a Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS One} 9 (3): e90081.
 
 \leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
-Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}:
-{Uncertainty} of Model Explanations for Non-Additive Predictive
-Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
+Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
-Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley}
-{Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
-Horst, Allison Marie, Alison Presmanes Hill, and Kristen B. Gorman.
-2020. {``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin
-Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
+Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
+Gorman.
+2020.
+{``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-karwowski_international_2006}{}}%
-Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of
-{Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
+Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-kodiyan_overview_2019}{}}%
-Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using
-{AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring
-Tool.''} \emph{Researchgate Preprint}.
+Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-kominsarczyk_treeshap_2021}{}}%
-Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw
-Biecek. 2021. {``Treeshap.''} Model Oriented.
-\url{https://github.com/ModelOriented/treeshap}.
+Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-larson_how_2016}{}}%
-Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin. 2016.
-{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''}
-\emph{ProPublica}, May.
+Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
+2016.
+{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''} \emph{ProPublica}, May.
 \url{https://www.propublica.org/article/how-we-analyzed-the-compas-recidivism-algorithm?token=RPR1E2qtzJltfJ0tS-gB_41kmfoWZAu4}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lee_state_2021}{}}%
-Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas
-Spyrison, Earo Wang, and H. Sherry Zhang. 2021. {``The State-of-the-Art
-on Tours for Dynamic Visualization of High-Dimensional Data.''}
-\emph{WIREs Computational Statistics} n/a (n/a): e1573.
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
+Sherry Zhang.
+2021.
+{``The State-of-the-Art on Tours for Dynamic Visualization of High-Dimensional Data.''} \emph{WIREs Computational Statistics} n/a (n/a): e1573.
 \url{https://doi.org/10.1002/wics.1573}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lee_pptree_2013}{}}%
-Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013.
-{``{PPtree}: {Projection} Pursuit Classification Tree.''}
-\emph{Electronic Journal of Statistics} 7: 1369--86.
+Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
 
 \leavevmode\vadjust pre{\hypertarget{ref-leone_fifa_2020}{}}%
-Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''}
-\url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
+Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-liaw_classification_2002}{}}%
-Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression
-by {randomForest}.''} \emph{R News} 2 (3): 18--22.
+Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
 
 \leavevmode\vadjust pre{\hypertarget{ref-lundberg_consistent_2018}{}}%
-Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent
-Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv
-Preprint arXiv:1802.03888}.
+Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-lundberg_unified_2017}{}}%
-Lundberg, Scott, and Su-In Lee. 2017. {``A Unified Approach to
-Interpreting Model Predictions.''} \emph{arXiv Preprint
-arXiv:1705.07874}.
+Lundberg, Scott, and Su-In Lee. 2017. {``A Unified Approach to Interpreting Model Predictions.''} \emph{arXiv Preprint arXiv:1705.07874}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-molnar_interpretable_2020}{}}%
+Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com.
 
 \leavevmode\vadjust pre{\hypertarget{ref-oneil_weapons_2016}{}}%
-O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data
-Increases Inequality and Threatens Democracy}. Crown.
+O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
 
 \leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
-Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin. 2016. {``"{Why}
-{Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any}
-{Classifier}.''} \emph{arXiv:1602.04938 {[}Cs, Stat{]}}, February.
+Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
+2016.
+{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} \emph{arXiv:1602.04938 {[}Cs, Stat{]}}, February.
 \url{http://arxiv.org/abs/1602.04938}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
-Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton
-University Press.
+Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
-Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical
-Science} 25 (3): 289--310.
+Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
 
 \leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
-Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje. 2017.
-{``Learning Important Features Through Propagating Activation
-Differences.''} In \emph{International {Conference} on {Machine}
-{Learning}}, 3145--53. PMLR.
+Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
+2017.
+{``Learning Important Features Through Propagating Activation Differences.''} In \emph{International {Conference} on {Machine} {Learning}}, 3145--53.
+PMLR.
 
 \leavevmode\vadjust pre{\hypertarget{ref-shrikumar_not_2016}{}}%
-Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul
-Kundaje. 2016. {``Not Just a Black Box: {Learning} Important Features
-Through Propagating Activation Differences.''} \emph{arXiv Preprint
-arXiv:1605.01713}.
+Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
+2016.
+{``Not Just a Black Box: {Learning} Important Features Through Propagating Activation Differences.''} \emph{arXiv Preprint arXiv:1605.01713}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-sievert_interactive_2020}{}}%
-Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data}
-{Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC.
-\url{https://plotly-r.com}.
+Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-simonyan_deep_2014}{}}%
-Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman. 2014. {``Deep
-Inside Convolutional Networks: {Visualising} Image Classification Models
-and Saliency Maps.''} In \emph{In {Workshop} at {International}
-{Conference} on {Learning} {Representations}}. Citeseer.
+Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
+2014.
+{``Deep Inside Convolutional Networks: {Visualising} Image Classification Models and Saliency Maps.''} In \emph{In {Workshop} at {International} {Conference} on {Learning} {Representations}}.
+Citeseer.
 
 \leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
-Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An {R} {Package}
-for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of
-{Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
+Spyrison, Nicholas, and Dianne Cook.
+2020.
+{``Spinifex: An {R} {Package} for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of {Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
 \url{https://doi.org/10.32614/RJ-2020-027}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
-Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation
-of Individual Classifications Using Game Theory.''} \emph{The Journal of
-Machine Learning Research} 11: 1--18.
+Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
 \leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
-Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and
-Damon Mayaffre. 2018. {``Textual {Deconvolution} {Saliency} ({TDS}): A
-Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the
-56th {Annual} {Meeting} of the {Association} for {Computational}
-{Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
+Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
+2018.
+{``Textual {Deconvolution} {Saliency} ({TDS}): A Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the 56th {Annual} {Meeting} of the {Association} for {Computational} {Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
 
 \leavevmode\vadjust pre{\hypertarget{ref-wickham_ggplot2_2016}{}}%
-Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data}
-{Analysis}}. Springer-Verlag New York.
-\url{https://ggplot2.tidyverse.org}.
+Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-wickham_visualizing_2015}{}}%
-Wickham, Hadley, Dianne Cook, and Heike Hofmann. 2015. {``Visualizing
-Statistical Models: {Removing} the Blindfold.''} \emph{Statistical
-Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
+Wickham, Hadley, Dianne Cook, and Heike Hofmann.
+2015.
+{``Visualizing Statistical Models: {Removing} the Blindfold.''} \emph{Statistical Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
 \url{https://doi.org/10.1002/sam.11271}.

### 2022-01-06 figures/_make_figures.r

@@ -182,7 +182,7 @@ dist_df$variable <- factor(dist_df$variable, levels = rev(.lvl_ord))
     scale_color_brewer(palette = "Dark2") +
     scale_fill_brewer(palette = "Dark2") +
     labs(title="SHAP distribution",
-         y = "Variable", x = "Normalized SHAP values\n the median of 25 permutations of the explanatory variables") +
+         y = "", x = "Normalized SHAP values\n the median of 25 orderings of the explanatory variables") +
     theme(legend.position = "off"))
 
 ## Breakdowns & plot ----
@@ -200,44 +200,47 @@ bd_dijk <- predict_parts(explainer       = rf_expl,
 bd_df_dijk <- my_bd_df(bd_dijk, "van Dijk (defense)")
 ## Bind, by row
 bd_df <- rbind(bd_df_messi, bd_df_dijk)
-bd_df <- bd_df[is.na(bd_df$variable) == FALSE, ]
+bd_df <- bd_df[!(bd_df$variable %in% c("intercept", "prediction")), ]
 (g_bd <- ggplot() + #scale_y_continuous(limits = c(0, 1)) +
-    # ## vertical "lines"
+    # ## vertical lines
     # geom_segment(aes(x=variable_num, xend=variable_num, y=variable, yend=last_cumulative),
     #              data = bd_df, color = "black", size = .5) +
-    ## horizontal "bars"
-    geom_segment(aes(x=cumulative, xend=last_cumulative, y=variable, yend=variable, color=player),
-                 data = bd_df, size = 1.5, alpha = .8) + 
-    facet_grid(col=vars(player)) +
+    ## horizontal bars
+    geom_segment(
+      aes(x = cumulative, xend = last_cumulative,
+          y = variable, yend = variable, color = player),
+      data = bd_df, size = 1.5, alpha = .8) +
+    facet_grid(col = vars(player)) +
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
     labs(title = "Breakdown plot", color = "Players",
-         y = "Variable", x = "Normalized contribution to prediction | variable order") +
-    theme(legend.margin = margin(0, 0, 0, 0),
+         y = "", x = "Normalized contribution to prediction | variable order") +
+    theme(legend.margin   = margin(0, 0, 0, 0),
           legend.position = "bottom",
[…]

### 2022-01-07 _paper.rmd
radial tour section plus update to figures

@@
 
-```{r shapdistrbd, echo=F, fig.align='center', fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries for FIFA 2020 players based on nine predictors corresponding to different skills.
+```{r shapdistrbd, echo=F, fig.align='center', fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
 Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
@@
 <!-- Tours intro -->
-A data visualization _tour_ animates many linear projections over small changes in the basis.
-One of the critical features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
-There are various types of tours that are distinguished by selecting their basis paths [@lee_state_2021; @cook_grand_2008].
+A _tour_ enables viewing of high-dimensional data by animating many linear projections with small incremental changes.
+It is achieved by following a path of linear projections (bases) of high-dimensional space.
+One of the features of the tour is the object permanence of the data points; one can track the relative change of observations in time, and as such gain information about the relationships between points across multiple variables.
+There are various types of tours that are distinguished by how the paths are generated [@lee_state_2021; @cook_grand_2008].
 
 <!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its basis path by manipulating a selected variable's contribution to the basis.
-A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
-The bases are then selected based on rotating this newly created manipulation space.
-A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
-Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis [@karwowski_international_2006].
-However, this navigation is relatively time-consuming due to the vast volume of the display space.
+The manual tour [@cook_manual_1997] defines its path by changing a selected variable's contribution to a basis, to allow the variable to contribute more or less to the projection.
+The contribution of all other variables is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
+It also lends itself to pre-computation to be queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
+<!-- However, this navigation is relatively time-consuming due to the vast volume of the display space.
 It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
-In this case, the projection of the normalized explanations is the feature of interest.
+In this case, the projection of the normalized explanations is the feature of interest.-->
 
 <!-- Radial tour -->
-More generally, the manual tour can change the contribution of a variable to the display dimensions.
-We will apply a more directed interaction, namely, a _radial_ tour.
+A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of the new work.
 In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and also interactive re-calculation to focus on a different variable.
 

### 2022-01-07 _paper.tex
radial tour section plus update to figures

@@
 ]{article}
-\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
-\author{}
-\date{\vspace{-2.5em}}
-
 \usepackage{amsmath,amssymb}
@@
 
+\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
+\author{}
+\date{\vspace{-2.5em}}
+
 \begin{document}
@@
 Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
+This phrase, Latin for ``other things held constant'' or ``all else unchanged'', shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
@@
 
-\caption{Illustration of the use of SHAP values for a random forest model for salaries for FIFA 2020 players based on nine predictors corresponding to different skills.
+\caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
 Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
@@
 
-A data visualization \emph{tour} animates many linear projections over small changes in the basis.
-One of the critical features of the tour is the object permanence of the data points; one can track the relative changes of observations as the basis moves, as opposed to discretely jumping to an orthogonal view with no intermediate information.
-There are various types of tours that are distinguished by selecting their basis paths (S.
+A \emph{tour} enables viewing of high-dimensional data by animating many linear projections with small incremental changes.
+It is achieved by following a path of linear projections (bases) of high-dimensional space.
+One of the features of the tour is the object permanence of the data points; one can track the relative change of observations in time, and as such gain information about the relationships between points across multiple variables.
+There are various types of tours that are distinguished by how the paths are generated (S.
 Lee et al.
@@
 
-The manual tour (Cook and Buja 1997) defines its basis path by manipulating a selected variable's contribution to the basis.
-A manipulation dimension is appended onto the projection plane, giving a full contribution to the chosen variable.
-The bases are then selected based on rotating this newly created manipulation space.
-A crucial feature of the manual tour is that it allows users to control the variable contributions of the basis.
-Such manipulations can be selected and queued in advance or selected on the spot for human-in-the-loop analysis (Karwowski 2006).
-However, this navigation is relatively time-consuming due to the vast volume of the display space.
-It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
-In this case, the projection of the normalized explanations is the feature of interest.
+The manual tour (Cook and Buja 1997) defines its path by changing a selected variable's contribution to a basis, to allow the variable to contribute more or less to the projection.
+The contribution of all other variables is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
+It also lends itself to pre-computation to be queued in advance or computed on-the-fly for human-in-the-loop analysis (Karwowski 2006).
 
-More generally, the manual tour can change the contribution of a variable to the display dimensions.
-We will apply a more directed interaction, namely, a \emph{radial} tour.
+A version of the manual tour called a \emph{radial tour} is implemented in Spyrison and Cook (2020) and forms the basis of the new work.
 In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and also interactive re-calculation to focus on a different variable.
 

### 2022-01-07 figures/_make_figures.r

@@ -182,7 +182,7 @@ dist_df$variable <- factor(dist_df$variable, levels = rev(.lvl_ord))
     scale_color_brewer(palette = "Dark2") +
     scale_fill_brewer(palette = "Dark2") +
     labs(title="SHAP distribution",
-         y = "", x = "Normalized SHAP values\n the median of 25 orderings of the explanatory variables") +
+         y = "", x = "Normalized SHAP values") +
     theme(legend.position = "off"))
 
 ## Breakdowns & plot ----
@@ -214,7 +214,7 @@ bd_df <- bd_df[!(bd_df$variable %in% c("intercept", "prediction")), ]
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
     labs(title = "Breakdown plot", color = "Players",
-         y = "", x = "Normalized contribution to prediction | variable order") +
+         y = "", x = "Normalized contribution to prediction") +
     theme(legend.margin   = margin(0, 0, 0, 0),
           legend.position = "bottom",
           axis.text.x     = element_blank(),

### 2022-01-07 _paper.rmd
global view section

@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
+The following section will use normalized explanations as the starting projection basis (call this the _attribution projection_) to further scrutinize the explanation.
+
+```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
+Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
+Plot b) is the break-down plot showing the SHAP values.
+It can be learned that reaction is important for explaining both players salaries.
+Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi."}
+knitr::include_graphics("./figures/shap_distr_bd.png")
+```
 
@@
 Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of variables is fixed, so the two observations can be more easily compared.
-
-```{r shapdistrbd, echo=F, fig.align='center', fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
-Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
-Plot b) is the break-down plot showing the SHAP values.
-It can be learned that reaction is important for explaining both players salaries.
-Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi."}
-knitr::include_graphics("./figures/shap_distr_bd.png")
-```
 
@@
 
-Below we illustrate the two primary displays of the cheem viewer application: the global view and the tour view.
-Then we cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing of the data before application runtime.
+To explore the local explanations, an ensemble of plots [@unwin] is provided, called the _cheem viewer_.
+There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
+In addition, there are numerous user inputs, including variable selection for the radial tour, and instance selection for making comparisons.
+Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) contain screenshots showing the cheem viewer for the two primary tasks classification (categorical response) and regression (quantitative response).
 
@@
 <!-- Purpose -->
-The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
-While the comparison of these spaces is interesting, the global view's purpose is to enable the selection of observations.
-The local explanation of these points will be explored in more detail.
+The global view provides the context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+Both of these spaces are of dimension $n\times p$, where $n$ is the number of observations and $p$ is the number of predictors.
+The attribution space corresponds to the local explanations for each observation, which will have $p$ values for each observation.
 
 <!-- Approximations of the spaces, PC1:2 -->
-An approximation of these spaces is given as the first two principal components of their respective spaces.
-Model information, the observed response by its prediction, is also provided.
-The orientation and magnitude of the variables are inscribed on a unit circle.
-<!-- Interactions -->Misclassified observations are circled in red if applicable.
-Linked brushing between the plots and tabular display of select points facilitates exploration of the spaces and the model.
-A single 2D projection will not encompass all of the structure of higher-dimensional space.
-However, it is a reasonable summarization given the real task at hand; the selection of observations to explore further.
+A visualisation of these spaces is provided by the first two principal components of their respective spaces.
+In addition, a plot observed by predicted response is also provided.
+In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot (XXX REF).
+A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
+<!-- Interactions --> For classification tasks, misclassified observations are circled in red if applicable.
+Linked brushing between the plots is provided and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
 
-## Radial cheem tour
+While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of observations, from which to explore the local explanations.
+
+## Radial tour
 
@@
 
-```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Display illustrating the classification case.
+```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem view for classification tasks.
 Plots are colored on predicted class, and red circles indicate misclassified observations.
 The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
-The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction.
-We probe the support of the variable contributions by selecting a variable to vary the contribution."}
+The first frame is the linear variable importances that best describe the observation's prediction.
+We probe the support of variable contributions by selecting a variable and use the radial tour to vary the coefficient."}
 knitr::include_graphics("./figures/app_classification.PNG")
@@
 
-```{r regressioncase, echo=F, out.width = "100%", fig.cap="Display of the regression task.
-The global view can be colored on the correlation of the attribution projection and observed response.
-In the tour, the horizontal values are the same as the classification case; the projection through the basis.
+```{r regressioncase, echo=F, out.width = "100%", fig.cap="Overview of the cheem view for tasks.
+The global view can be colored on the correlation between the projection of the data generated by the local explanation and the observed response.
+In the radial tour, the horizontal values are the same as the classification case; the projection through the basis.
 The vertical position is now mapped to the observed y and residuals."}

### 2022-01-07 _paper.tex
global view section

@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as the starting projection basis to further scrutinize the explanation.
+The following section will use normalized explanations as the starting projection basis (call this the \emph{attribution projection}) to further scrutinize the explanation.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.6\linewidth]{./figures/shap_distr_bd}
+
+}
+
+\caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
+Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
+Plot b) is the break-down plot showing the SHAP values.
+It can be learned that reaction is important for explaining both players salaries.
+Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
+\end{figure}
 
@@
 Plot (b) is a modified breakdown plot (Gosiewska and Biecek 2019) where the order of variables is fixed, so the two observations can be more easily compared.
-
-\begin{figure}
-
-{\centering \includegraphics[width=25in]{./figures/shap_distr_bd}
-
-}
-
-\caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
-Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
-Plot b) is the break-down plot showing the SHAP values.
-It can be learned that reaction is important for explaining both players salaries.
-Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
-\end{figure}
 
@@
 
-Below we illustrate the two primary displays of the cheem viewer application: the global view and the tour view.
-Then we cover what we take away from the classification and regression tasks.
-Lastly, we discuss the preprocessing of the data before application runtime.
+To explore the local explanations, an ensemble of plots (Unwin and Valero-Mora 2018) is provided, called the \emph{cheem viewer}.
+There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
+In addition, there are numerous user inputs, including variable selection for the radial tour, and instance selection for making comparisons.
+Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} contain screenshots showing the cheem viewer for the two primary tasks classification (categorical response) and regression (quantitative response).
 
@@
 
-The global view provides an essential context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
-While the comparison of these spaces is interesting, the global view's purpose is to enable the selection of observations.
-The local explanation of these points will be explored in more detail.
+The global view provides the context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of observations and \(p\) is the number of predictors.
+The attribution space corresponds to the local explanations for each observation, which will have \(p\) values for each observation.
 
-An approximation of these spaces is given as the first two principal components of their respective spaces.
-Model information, the observed response by its prediction, is also provided.
-The orientation and magnitude of the variables are inscribed on a unit circle.
-Misclassified observations are circled in red if applicable.
-Linked brushing between the plots and tabular display of select points facilitates exploration of the spaces and the model.
-A single 2D projection will not encompass all of the structure of higher-dimensional space.
-However, it is a reasonable summarization given the real task at hand; the selection of observations to explore further.
+A visualisation of these spaces is provided by the first two principal components of their respective spaces.
+In addition, a plot observed by predicted response is also provided.
+In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot (XXX REF).
+A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
+For classification tasks, misclassified observations are circled in red if applicable.
+Linked brushing between the plots is provided and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
 
-\hypertarget{radial-cheem-tour}{%
-\subsection{Radial cheem tour}\label{radial-cheem-tour}}
+While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of observations, from which to explore the local explanations.
+
+\hypertarget{radial-tour}{%
+\subsection{Radial tour}\label{radial-tour}}
 
@@
 
-\caption{Display illustrating the classification case.
+\caption{Overview of the cheem view for classification tasks.
 Plots are colored on predicted class, and red circles indicate misclassified observations.
 The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
-The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction.
-We probe the support of the variable contributions by selecting a variable to vary the contribution.}\label{fig:classificationcase}
+The first frame is the linear variable importances that best describe the observation's prediction.
+We probe the support of variable contributions by selecting a variable and use the radial tour to vary the coefficient.}\label{fig:classificationcase}
 \end{figure}
@@
 
-\caption{Display of the regression task.
-The global view can be colored on the correlation of the attribution projection and observed response.
-In the tour, the horizontal values are the same as the classification case; the projection through the basis.
+\caption{Overview of the cheem view for tasks.
+The global view can be colored on the correlation between the projection of the data generated by the local explanation and the observed response.
+In the radial tour, the horizontal values are the same as the classification case; the projection through the basis.
 The vertical position is now mapped to the observed y and residuals.}\label{fig:regressioncase}
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-unwin}{}}%
+Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%

### 2022-01-07 _paper.rmd
working on penguins example

@@
 
-## 1) Penguin, species classification
+## Palmer penguin, species classification
 
-Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] consist of 330 observations across four physical measurements of three species of penguins foraging near Palmer Station, Antarctica.
-A random forest model was fit, classifying the species of the penguin given the physical measurements.
+The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
+The data was provided publicly as a substitute for the over-used iris data, and is quite similar in form.
+There are 330 observations and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
+A random forest model was fitted with species as the response variable.
 
@@
 
-In figure \@ref(fig:casepenguins), a misclassified point is contrasted with a correctly classified point of its observed class nearby in data-space.
-The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
-The primary observation is between the predicted and observed clusters, a sign of uncertainty in the prediction.
-The tour varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
+Figure \@ref(fig:casepenguins) shows plots from the cheem viewer for exploring the random forest model on the penguins data.
+Plot (a) shows the global view, and plot (b) shows several plots generated with the radial tour.
+
+There is more separation visible in the attribution space than the data space, as would be expected.
+The observed vs predicted plot reveals a handful of misclassified observations.
+A misclassified point (represented by the asterisk in the global view, and as a solid vertical line in the tour view) is contrasted with a correctly classified point (represented by an $x$ and vertical dashed line).
+
+The radial tour starts from the attribution projection of the misclassified observation (left plot). The important variables identifed by SHAP are `b_d` and `wgt` in contrast to `f_l`.
+
+varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
 Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.

### 2022-01-07 _paper.tex
working on penguins example

@@
 
-\hypertarget{penguin-species-classification}{%
-\subsection{1) Penguin, species classification}\label{penguin-species-classification}}
+\hypertarget{palmer-penguin-species-classification}{%
+\subsection{Palmer penguin, species classification}\label{palmer-penguin-species-classification}}
 
-Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) consist of 330 observations across four physical measurements of three species of penguins foraging near Palmer Station, Antarctica.
-A random forest model was fit, classifying the species of the penguin given the physical measurements.
+The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
+The data was provided publicly as a substitute for the over-used iris data, and is quite similar in form.
+There are 330 observations and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
+A random forest model was fitted with species as the response variable.
 
@@
 
-In figure \ref{fig:casepenguins}, a misclassified point is contrasted with a correctly classified point of its observed class nearby in data-space.
-The attribution space from the tree SHAP local explanations is a more separable space, where the comparison is squarely in the middle of the orange distribution.
-The primary observation is between the predicted and observed clusters, a sign of uncertainty in the prediction.
-The tour varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
+Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
+Plot (a) shows the global view, and plot (b) shows several plots generated with the radial tour.
+
+There is more separation visible in the attribution space than the data space, as would be expected.
+The observed vs predicted plot reveals a handful of misclassified observations.
+A misclassified point (represented by the asterisk in the global view, and as a solid vertical line in the tour view) is contrasted with a correctly classified point (represented by an \(x\) and vertical dashed line).
+
+The radial tour starts from the attribution projection of the misclassified observation (left plot).
+The important variables identifed by SHAP are \texttt{b\_d} and \texttt{wgt} in contrast to \texttt{f\_l}.
+
+varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
 Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.

### 2022-01-07 _paper.rmd
typo

@@
 These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (footbal salaries, house prices) response variables.
+Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables.
 The methods are implemented in the __R__ package __cheem__, available on CRAN.

### 2022-01-07 _paper.tex
typo

@@
 These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (footbal salaries, house prices) response variables.
+Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables.
 The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.

### 2022-01-08 _paper.rmd
Saturday changes: - JMLR documents; .sty not effecting knit. - cheem scaling - Minor edits TODO: enlarge and clean up Case Studies & writing.

@@
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
-# documentclass: jdssv
+#documentclass: jmlr
 #classoption: article
 output:
+  ## pdf_document: requires \ref{} 
+  #pdf_document:
   ## bookdown::pdf_document2: allows \@ref()
   bookdown::pdf_document2:
-  ## pdf_document: requires \ref{} 
-  #pdf_document:
     keep_tex: true
     toc: false
-    #template: template/jdssv_template.tex
     number_sections: yes
     fig_caption: yes
+    #includes:
+    #  in_header: template/jmlr2e.sty
 bibliography: cheem_paper.bib
-#appendix:
-# title:
-#   formatted: "Methods for understanding the variable importance of local explanations of black-box models"
-#   plain:     "Methods for understanding the variable importance of local explanations of black-box models"
-#   short:     "cheem: radial tour scrutinizing local explanations"
 editor_options:
@@
 
-<!-- ## SHAP and history -->
+<!-- SHAP -->
 SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation conceptually by examining the effect of other variables on the contribution of the variable of interest on predicting the response.
@@
 @strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
-However, it is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+While the models themselves are invariant to the order of the predictors, the additive explanations are not.
+The SHAP values are the mean contributions over some subset these permutations.
+This is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+Partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one instance.
 It could also be considered to be similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each variable in the context of all other candidate variables.
 
+<!-- tree SHAP -->
 For our application, we use _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
+Instead of aggregating over permutations of the variable order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
@@
 ```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
-Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
-Plot b) is the break-down plot showing the SHAP values.
-It can be learned that reaction is important for explaining both players salaries.
+A star offensive and densive player, Messi and a defensive player are compared, L.
+Messi and V van Dijk respectively.
+Plot a) is the break-down plot showing one additive variable-attributution.
+These attributions are sensitive to the order of the variables.
+Plot b) shows the distribution of attribution for each variable across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi."}
@@
 There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
-A random forest model is fit regressing log wages [2020 Euros], from the skill measures.
+A random forest model is fit regressing wages [2020 Euros], from the skill measures.
 We then extract the SHAP values of a star offensive player (L.
@@
 In addition, a plot observed by predicted response is also provided.
-In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot (XXX REF).
+In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot [@gabriel_biplot_1971].
 A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
@@
 
-What information do we glean from using this method on a classification task?
 Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
-The initial frame is the linear attribution of that observation's local explanation.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation; we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+The model infomation in the global view is a depiction of the model conmfusion matrix.
+The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each variable to structure identified by the local explanation, evaluating the support or robustness of the prediction.
 
-```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem view for classification tasks.
+```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "The cheem view for classification tasks.
 Plots are colored on predicted class, and red circles indicate misclassified observations.
-The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
+Model visual is essentially a representation of the confusion matrix of the model.
+The radial tour is a 1D projection starting the attribution projection.
 The first frame is the linear variable importances that best describe the observation's prediction.
-We probe the support of variable contributions by selecting a variable and use the radial tour to vary the coefficient."}
+We probe the support of variable contributions by selecting a variable and use the radial tour to vary its coefficient."}
 knitr::include_graphics("./figures/app_classification.PNG")
@@
 \begin{itemize}
-	\item \textbf{Data:} a complete numerical matrix; explanatory and response variable.
-An optional categorical variable can be mapped to the color and shape of observations.
-Explanatory variables are scaled in visualization after modeling or creating local explanations.
-	\item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)] for compatibility with the local explanation, which requires tree-based models.
-	\item \textbf{Local explanation:} any model-compatible linear explanation could be used.
-We apply tree SHAP, a more computationally efficient variant of SHAP, compatible with tree-based models.
-Tree SHAP was calculated with the package \textbf{treeshap} [(ref:citeTs), hosted on GitHub only].
-The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
+	\item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale invariant.
+	\item \textbf{Model and explanation:} any model can be used with this method.
+Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)], compatibility tree SHAP.
+We use modest hyperparameters, namely: 125 trees, number variables randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} (ref:citeTs).
+This implementation aggregates over exhaustively over the attribution of all trees and we opt to not to fit interactions of features.
+	\item \textbf{Cheem view:} after the model and full explanation space are calculated we scale each feature by standard deviations away from the mean to achieve a common support for visuals.
+Statistics for mapping to color are calculated on the scaled spaces.
+Interactive tabular display reports the original values.
 \end{itemize}
@@
 The time to preprocess the data will vary significantly with the model and local explanation.
-For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.9 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 254 seconds combined.
-PCA and statistics of the variables and attributions took 0.6 seconds.
+For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 270 seconds combined.
+PCA and statistics of the variables and attributions took 2.8 seconds.
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
@@
 By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
-The .mp4 animation of this tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
+The .mp4 animation of this tour can be found at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
 ```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
@@
 We vary the contribution of calories from fat.
-Animated tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
+Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
 ```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
@@
 The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
-After aggregation of the skill measurements, we regress the log wages [2020 euros] given just the skill aggregates.
+After aggregation of the skill measurements, we regress player wages [2020 euros] given just the skill aggregates.
 The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
@@
 
-(ref:casefifa) FIFA 2020, regressing log wages [2020 Euros] from aggregations of skill measurements.
+(ref:casefifa) FIFA 2020 data, a random forest model regresses wages [2020 Euros] from aggregated of skill measurements.
 The primary observation is a star offensive player (L.
@@
 van Dijk).
-The animate radial tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4)
+The animate radial tour can be found at [shorturl.at/lCK26](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4)
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
@@
 
-(ref:caseames) Ames housing 2018 regressing log sales price [2018 USD].
+(ref:caseames) Ames housing 2018 regressing sales price [2018 USD].
 Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
 This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
-The corresponding animation is at [github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
+The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
 ```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}

### 2022-01-08 _paper.tex
Saturday changes: - JMLR documents; .sty not effecting knit. - cheem scaling - Minor edits TODO: enlarge and clean up Case Studies & writing.

@@
 
-\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
-\author{}
-\date{\vspace{-2.5em}}
-
 \begin{document}
@@
 Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-This phrase, Latin for ``other things held constant'' or ``all else unchanged'', shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
+This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
 It ignores correlations of the variables and imagines a case that was not observed.
@@
 2016; Shrikumar, Greenside, and Kundaje 2017), a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME (Ribeiro, Singh, and Guestrin 2016) SHAP (S.
-Lundberg and Lee 2017), and their variants are popular.
+There are fewer model-agnostic explanations, of which LIME (Ribeiro, Singh, and Guestrin 2016) SHAP (Lundberg and Lee 2017), and their variants are popular.
 
@@
 Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
-However, it is also related to partial dependence plots (Molnar 2020), used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+While the models themselves are invariant to the order of the predictors, the additive explanations are not.
+The SHAP values are the mean contributions over some subset these permutations.
+This is also related to partial dependence plots (Molnar 2020), used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
+Partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one instance.
 It could also be considered to be similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each variable in the context of all other candidate variables.
 
-For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (S.
-M.
-Lundberg, Erion, and Lee 2018).
+For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
+Instead of aggregating over permutations of the variable order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
@@
 \caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-Two observations (a star offensive player, Messi and a defensive player, van Dijk) are compared.
-Plot (a) shows the distribution of contributions for each variable across 25 permutations of predictors, with the median displayed as a dot, for both players.
-Plot b) is the break-down plot showing the SHAP values.
-It can be learned that reaction is important for explaining both players salaries.
+A star offensive and densive player, Messi and a defensive player are compared, L.
+Messi and V van Dijk respectively.
+Plot a) is the break-down plot showing one additive variable-attributution.
+These attributions are sensitive to the order of the variables.
+Plot b) shows the distribution of attribution for each variable across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
@@
 
-What information do we glean from using this method on a classification task?
 Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
-The initial frame is the linear attribution of that observation's local explanation.
-By default, the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation; we can test the sensitivity of each variable to structure identified by the local explanation; we are exploring the support of the explanation, evaluating the support or robustness of the prediction.
+The model infomation in the global view is a depiction of the model conmfusion matrix.
+The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each variable to structure identified by the local explanation, evaluating the support or robustness of the prediction.
 
@@
 
-\caption{Overview of the cheem view for classification tasks.
+\caption{The cheem view for classification tasks.
 Plots are colored on predicted class, and red circles indicate misclassified observations.
-The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point.
+Model visual is essentially a representation of the confusion matrix of the model.
+The radial tour is a 1D projection starting the attribution projection.
 The first frame is the linear variable importances that best describe the observation's prediction.
-We probe the support of variable contributions by selecting a variable and use the radial tour to vary the coefficient.}\label{fig:classificationcase}
+We probe the support of variable contributions by selecting a variable and use the radial tour to vary its coefficient.}\label{fig:classificationcase}
 \end{figure}
@@
 \begin{itemize}
-    \item \textbf{Data:} a complete numerical matrix; explanatory and response variable.
-An optional categorical variable can be mapped to the color and shape of observations.
-Explanatory variables are scaled in visualization after modeling or creating local explanations.
-    \item \textbf{Model:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)] for compatibility with the local explanation, which requires tree-based models.
-    \item \textbf{Local explanation:} any model-compatible linear explanation could be used.
-We apply tree SHAP, a more computationally efficient variant of SHAP, compatible with tree-based models.
-Tree SHAP was calculated with the package \textbf{treeshap} [Kominsarczyk et al.
-(2021), hosted on GitHub only].
-The global view shows all observations in attribution space, requiring the variable importance from \emph{all} observations rather than just one.
+    \item \textbf{Data:} predictors are unscaled complete numerical matrix; while the response variable is occasionally log scaled for a more Gaussian distribution.
+Most models and local explanations will be scale invariant.
+    \item \textbf{Model and explanation:} any model can be used with this method.
+Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)], compatibility tree SHAP.
+We use modest hyperparameters, namely: 125 trees, number variables randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} Kominsarczyk et al.
+(2021).
+This implementation aggregates over exhaustively over the attribution of all trees and we opt to not to fit interactions of features.
+    \item \textbf{Cheem view:} after the model and full explanation space are calculated we scale each feature by the standard deviation away from the mean to achieve a common support for visuals.
+Statistics are calculated on the scaled spaces.
+Interactive tabular display reports the original values.
 \end{itemize}
@@
 By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
-The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_penguins.mp4}}\label{fig:casepenguins}
+The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}}\label{fig:casepenguins}
 \end{figure}
@@
 We vary the contribution of calories from fat.
-Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_chocolates.mp4}.}\label{fig:casechocolates}
+Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolates}
 \end{figure}
@@
 van Dijk).
-The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_fifa.mp4}}\label{fig:casefifa}
+The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{shorturl.at/lCK26}}\label{fig:casefifa}
 \end{figure}
@@
 This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
-The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{github.com/nspyrison/cheem\_paper/blob/main/figures/case\_ames2018.mp4}}\label{fig:caseames}
+The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}}\label{fig:caseames}
 \end{figure}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lundberg_unified_2017}{}}%
-Lundberg, Scott, and Su-In Lee. 2017. {``A Unified Approach to Interpreting Model Predictions.''} \emph{arXiv Preprint arXiv:1705.07874}.
+Lundberg, Scott M., and Su-In Lee.
+2017.
+{``A Unified Approach to Interpreting Model Predictions.''} In \emph{Proceedings of the 31st International Conference on Neural Information Processing Systems}, 4768--77.
 
 \leavevmode\vadjust pre{\hypertarget{ref-molnar_interpretable_2020}{}}%
-Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com.
+Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
 
@@
 2016.
-{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} \emph{arXiv:1602.04938 {[}Cs, Stat{]}}, February.
-\url{http://arxiv.org/abs/1602.04938}.
+{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} In \emph{Proceedings of the 22nd {ACM} {SIGKDD} {International} {Conference} on {Knowledge} {Discovery} and {Data} {Mining}}, 1135--44.
+{KDD} '16.
+New York, NY, USA: Association for Computing Machinery.
+\url{https://doi.org/10.1145/2939672.2939778}.
 

### 2022-01-08 template/author_checklist.rmd
Saturday changes: - JMLR documents; .sty not effecting knit. - cheem scaling - Minor edits TODO: enlarge and clean up Case Studies & writing.

---
title: "Checklist of common JMLR formatting errors"
author: "Nick Spyrison"
date: "1/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Following: https://www.jmlr.org/format/formatting-errors.html

# Checklist of common JMLR formatting errors

## General

- One LaTeX file: The latex for the paper should be in one file. Otherwise it takes too much additional time for the production staff to perform all of the checks before publication.
- Make sure your abstract reads well. It will be used to advertise your paper.
- We recommend you use small caps for system names. (In LaTex that's \sc)
- Move file to have correct name, according to the JMLR naming conventions. See the authors' guide.
- Signed copyright form: Do not forget to provide this.
- Text should be in Computer Modern
- Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References.
- Acknowledgments should be a unnumbered section. You can use the \acks{} command for this.
- Appendix section number should be a letter (Should look like Appendix A. Title). You can use \appendix followed by regular \section command to format appendices

## Title Area

- Capitalization of paper title: In the title area, the paper title should be in title case. (First letter of each word capitalized except for little words like "and" and "or".)
- Author contact addresses: These should be complete enough that mail sent to that address will reach them
- Editor listed: Please include the name of the Action Editor who handled your submission in the \editor{} tag
- Keyword: Please add keywords to your article using \begin{keywords} ... \end{keywords} after your abstract. They should not be capitalised, they should be separated by commas, and they should not have a period at the end.
- Abstract: Do not delete the % at the end of \begin{abstract}% as it introduces an extra indent
- Title should be split across two lines, only when it is too long for one line.

## Headers and Footers
- Front page header: Use the \jmlrheading command to get a correct header and footer on the title page
- Copyright notice, bottom of front page uses authors full names: In the \jmlrheading command, please use the full names of all authors
- Running headers: Add running headers to your paper using the \ShortHeadings command. Please see the example paper on the web site for how to do this.
- Running headers: In the \ShortHeadings command, please capitalise the words in your paper title.
- Running headers: In the author list in the \ShortHeadings command, please use "and" before the last author
- Running headers: The ShortHeadings author list should have last names only, and should list all authors if possible. If you have many authors, "et al" is acceptable.

## Section Headers

- No unnumbered sections: We do not use \paragraph or \section*. We do not allow bold, unindented text to begin a paragraph that could be mistaken for a section header. If it looks like a section header it needs to be numbered like a section header, otherwise it needs to be formatted in some other way (e.g., by indentation, or by using italics instead of boldface).
- Section numbering: All sections, subsections, subsubsubsections must be numbered. Please use \subsubsection instead of \subsubsection*, etc
Capitalization of Section headings: All section headings should be title case. (First letter of each word capitalized except for little words like "and" and "or".) This includes subsections, sub subsections, etc.
- Section headers: There must be some text separating all section headers

## Tables and Figures

- Float placement: In general, floats should be at the top or bottom of the page (avoid using [h] option to \begin{figure}). Check for pages that contain several floats and only 2-3 lines of text. These will be confusing for a reader, who is likely to miss the text. When this happens, please put the floats on a floats page ([p] specifier) instead.
- Floats should also be placed as close as possible to the page at which they are first referenced. You should not, e.g., place all of the figures at the end of the paper.
Figures, tables should be centered
- Captions should be underneath figures and tables
- Captions NOT in italics, bold
- Tables and figures should not overfill margins
Multiline captions should have a hanging indent, that is, subsequent lines of the caption should be indented, aligning with the first word of the caption rather than the word "Figure". Usually this problem is caused by one of the other packages that you have included. Sometimes changing the order of includes will fix this (try moving the jmlr2e include farther down).

## Footnotes

- Footnote markers should follow punctuation
- Footnotes should be complete sentences

## Misc

- Dashes: When using dashes in text, please use --- instead of --, with no space before or after---like this.
- Data set: We use "data set" rather than "dataset".

## Citations
- Carefully read the section on citations in Instructions for Formatting JMLR Articles. A quick guideline is that the paper should still read correctly if all parenthesized text were omitted. Here are examples of BAD citations:
    - "See (Ginsberg, 1994) for an analysis..." [Uses citation as a noun]
    - "Drummond (Drummond, 1994) showed that..." [Should be "Drummond (1994)"]
    - "...the best method (see the ERT test (Davis, 1995))..." [double parens, should be   "...best method (see the ERT test, Davis, 1995)..."]
- Citations not nested in parenthesis. Parentheses should not occur within parentheses, even for citations. Within parentheses use \citealt or \citealp instead. Or you can use \citep[e.g.,][]{smith00}.

## Object References

- use \url commands to reference urls in papers. This causes hyperref to make the hyperlinks work nicely.
- Parenthesis: \eqref can sometimes result in parentheses within parentheses. When referencing an equation within a parenthetical statement, use Equation 1 rather than (1) or Equation (1).
- When referencing figures, sections, etc. ("Figure 1", "Section 2.1"), please make sure the words "Figure", "Section", etc., are uppercase

## Math

- Punctuation: Math in formal writing should be punctuated as if the extra space is not there, for example, displayed equations should usually be followed by a comma or a period, depending on the surrounding text.
- Punctuation: Before displayed equations, punctuation (e.g., ":") is not necessary unless you would use the same punctuation if the equation were not there.
- In general, only equations that are referenced should be numbered. We are willing to relax this for equations that are particularly important, as others may wish to discuss these in later correspondence.
- Proofs should end with a filled box.
- Equations should not run over the right hand margin. Appendices are especially prone to this.
- Theorems, Lemmas, etc., should be labeled like Theorem 1, Theorem 2, Theorem 3, not as Theorem 2.1, Theorem 2.2, Theorem 2.3

## References
- It is preferable to cite published papers, rather than tech reports.
- Author names: Make citations consistent. Either give the full first name of all cited authors, or give the first initials only for all cited authors.
- Capitalization in paper titles: Titles of conference papers should have the first word only titles. However, watch out for words like Bayes and Markov. These should still be capitalized, but BibTeX often leaves uncapitalised incorrectly.
- Capitalization of book titles, etc: The parts in italics (book names, journal names, conf names) should be title case. (First letter of each word capitalized except for little words like "and" and "or".)
- Please cite conference titles consistently. We do not have a enforced style for conference names, but whichever style you choose, be consistent. Don't refer to the same conference as "AAAI 2004" in one reference and the "The Nineteenth National Conference on Artificial Intelligence" in another one. Personally, I prefer the style "International Conference on Machine Learning (ICML)", leaving off the "Proceedings of the Twentieth..."

### 2022-01-08 template/author_guide.rmd
Saturday changes: - JMLR documents; .sty not effecting knit. - cheem scaling - Minor edits TODO: enlarge and clean up Case Studies & writing.

Following: https://www.jmlr.org/format/authors-guide.html

# Authors' Guide: Instructions for Final Preparation

## Journal of Machine Learning Research

Congratulations on the acceptance of your paper into JMLR. Once a paper has been accepted for publication, you should follow the procedure outlined below. The final version of your manuscript is due to us within three months of the acceptance date. If you require more than three months, please request an extension from the editor.

Because JMLR follows a low-cost, open access model, our production staff is all volunteer. Therefore, please follow the formatting instructions carefully, as this will save both your time and that of the production staff.

Please take special note of this common list of formatting errors.

Make sure that your article is formatted according the JMLR style. The JMLR style is described in Instructions for Formatting JMLR Articles. This page contains more information, including the official JMLR style file, and example papers. Please pay particular attention to the list of common formatting errors.

Your paper should be prepared in LaTeX, using the official JMLR style file, which you should not modify. We must be able to compile your paper using pdflatex. In particular, this means that if you use PostScript figures, you must use eps files rather than ps files.

Most likely, you will have done this already for the submission of your paper. However, you should go over the formatting rules and the checklist again carefully before submitting the camera ready copy. It is our experience that most authors break at least a dozen formatting rules. (Can you do better? :-)

Authors are responsible for correctly formatting their article. If you wish we can recommend professionals who will do this for you for a fee.

Make any final changes and have your article proofread. Because our production staff is all volunteer, we ourselves cannot copy edit your paper. Authors are notoriously bad proofreaders. Have the paper proofread by a skilled proofreader — NOT one of the authors.

If you wish, create an online appendix. Online appendices can contain source code, data, even quicktime demonstrations — anything that that will make life easier (or more interesting) for your readers and researchers who follow in your footsteps.

If there are online appendices containing source code, please sign the software release form (available in PDF format) and send it along with the Copyright form. The release form is intended to reduce the liability of JMLR.

Create a file archive (zip or tar.gz is file) with your paper. This archive should contain the latex source, the .bib file, figures, and everything needed to compile your document. The archive should contain only one .tex file (please do not use \include or \input.) We process your paper via an automatic system, so please follow these rules:

Your submission must be an archive that contains all of the tex files necessary to compile the paper. It must be compilable using pdflatex. We will not run any custom Makefiles.

Your main tex file should be named after the paper id you were assigned during the reviewing process, e.g., if your paper had id 12-150, then your tex file should be called 12-150.tex. It is OK if your archive file contains files within a subdirectory (of any name). If you wish to include online-only supplementary material, please place the final pdf for this material in a directory of your archive called supplementary.

The system will fail if it sees file types it does not recognize. The most common causes of this are .synctex.gz files.

Your page numbers should start at 1, both in the \jmlrheading command and in your actual paper.

You file should include an \editor{...} command that lists the action editor who has corresponded with you.

Your \jmlrheading command should contain the correct year, volume number, publication dates, etc. You must fill in all of the information. The format for this is:

   \jmlrheading{volume}{year}{pages}{dates when your paper submitted}{date published}{your paper id}{all authors full names}
For example, suppose that you submitted your paper to the production system in April 2017. The current volume at that time was 18 (you can find the current volume on the front page of jmlr.org)

       \jmlrheading{18}{2017}{1-36}{8/14; Revised 12/15}{4/17}{15-449}{John
Smith, Jane Doe, and Frank Jones}
Please use the current month and year as the publication date.

If you never needed to submit a revision (i.e. you submitted one version that was accepted with no changes), then you should leave off the "Revised" date. These dates will be verified against the paper database, so it is important that they be correct.

Our automated system does not run bibtex (this is similar to what arxiv.org does). Therefore if your paper uses bibtex, you should include the bbl file that is necessary to compile your final paper.
The scripts specifically check whether the paper headings are correct (e.g., page numbers, \jmlrheading, \ShortHeadings, \editor, and so on). To speed processing of your paper, please give special attention to make sure that these are correct and follow the lines of the example file.

In the past, we have spell checked all papers before publication, but we will no longer be doing this as part of the automated system. Fortunately, we are able to make available an aspell word list that we have been using for the past several years: https://github.com/JournalMLR/jmlr-aspell. This dictionary contains many technical terms that should make spell checking easier. Contributions are welcome; just create a pull request.

Print out, fill in, and sign the Permission to Publish form (PDF). The form agreement grants a CC-by license to the public, which allows for publication in JMLR online by JMLR, Inc. and in a print-archive edition by Microtome Publishing. The copyright remains with the author.

Upload your paper archive and the Permission to Publish form to the Production Editor of JMLR using the online web form. You should have received an email with the URLs of the automated system when your paper was accepted. If you have not received this email, please contact the production editor.

The production staff will review your submission. Most times there will be a few minor typographical changes that we will need from you. We will prepare a pdf from the files that you send us, and send it to you for final approval for publication.

After we receive the final version and release the article, we will post an announcement on the JMLR mailing list. The paper will also be stored at the JMLR distribution site and forwarded to ACM's computing repository (CORR) and the CiteSeer Index.

Based on the JAIR Author's Guide, gratefully reproduced by permission of JAIR. This file is available at http://www.jmlr.org/format/authors-guide.html. Last modified by Charles Sutton, 22 December 2014.

### 2022-01-08 figures/_make_figures.r

@@ -9,16 +9,18 @@
   my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
     ## Remade from: iBreakDown:::print.break_down_uncertainty
     data.frame(
-      player = player_tag,
-      label  = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+      player   = player_tag,
+      label    = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
       variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
-      value  = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
+      value    = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
       ## Of the distribution of local attributions:
-      min    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
-      q1     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
-      median = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), median, na.rm = TRUE),
-      q3     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.75, na.rm = TRUE),
-      max    = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), max, na.rm = TRUE))
+      # min      = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
+      # q1       = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
+      mean     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), mean, na.rm = TRUE),
[…]

### 2022-01-08 figures/_make_figures_case_study.r

@@ -18,7 +18,6 @@
   ames2018_ls   <- readRDS(paste0(fp, "preprocess_ames2018.rds"))
   chocolates_ls <- readRDS(paste0(fp, "preprocess_chocolates.rds"))
   
-
   .x_axis_title     <- "          x: PC1, y: PC2        x: PC1, y: PC2  x: predicted, y: observed"
   .x_axis_title_reg <- "          x: PC1, y: PC2             x: PC1, y: PC2     x: predicted, y: observed"
 }
@@ -28,17 +27,17 @@
   ## Data setup, spinifex::penguins
   names(penguins_ls)
   prim_obs <- 124L
-  comp_obs <- 86L
+  comp_obs <- 76L
   
   ### Global view and cheem tour stills
   .glob_view <- global_view(
     penguins_ls, prim_obs, comp_obs, as_ggplot = TRUE) +
     theme(legend.position = "right") +
     labs(color = 'Predicted class', shape = 'Predicted class',
-         x = .x_axis_title) + 
+         x = .x_axis_title) +
     ggtitle("Global view")
   .bas <- basis_attr_df(penguins_ls$attr_df, prim_obs)
-  .mv  <- 1 ## seems better than the 3 from
+  .mv  <- 1 ## seems better than the 3 returned from
   ##manip_var_of_attr_df(penguins_ls$attr_df, prim_obs, comp_obs)
   ## Cheem tour for stills
   .ggt <- radial_cheem_tour(
@@ -47,8 +46,7 @@
     do_add_pcp_segments = FALSE,
     pcp_shape = 124, angle = 6) + theme(legend.position = "")
   .cheem_stills <- spinifex::filmstrip(.ggt, nrow = 1) +
-    ggtitle("Cheem tour, extrema contributions")# +
-    #theme(plot.title = element_text(hjust = 0.05))
+    ggtitle("Radial tour, select frames")
   .cp <- cowplot::plot_grid(
     .glob_view, .cheem_stills,
     labels = c("a)", "b)"),
@@ -59,7 +57,7 @@
 ggplot2::ggsave(
   "./figures/case_penguins.png",
   plot = .cp, device = "png",
-  width = 6, height = 5, units = "in")
+  width = 7, height = 6, units = "in")
 .m <- gc()
 
 ### Save .mp4, add GitHub urls to paper
@@ -102,7 +100,7 @@ ggplot2::ggsave(
[…]

### 2022-01-08 figures/z_explore_SHAP_normalization.r

@@ -1,0 +1,157 @@
+# ch5_fig1_shap_distr_bd ----
+{
+  require("DALEX")
+  require("dplyr")
+  require("ggplot2")
+  set.seed(2022)
+  
+  ## Local func
+  my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
+    ## Remade from: iBreakDown:::print.break_down_uncertainty
+    data.frame(
+      player   = player_tag,
+      label    = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+      variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+      value    = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
+      ## Of the distribution of local attributions:
+      min      = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
+      q1       = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
+      median   = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), median, na.rm = TRUE),
+      q3       = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.75, na.rm = TRUE),
+      max      = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), max, na.rm = TRUE))
+  }
+  my_parts_distribution <- function(pred_parts, player_tag = "<tag unused>"){
+    df <- data.frame(
+      player       = player_tag,
+      label        = pred_parts$label,
+      variable     = pred_parts$variable_name,
+      value        = pred_parts$variable_value, ## Obs value of X
+      contribution = pred_parts$contribution,   ## SHAP contribution
+      B_perm_num   = pred_parts$B
+    )
+    rownames(df) <- paste(pred_parts$label, pred_parts$variable, pred_parts$B, sep = ": ")
+    return(df)
+  }
+  .lvl_ord <- c("reaction", "offense", "movement", "defense", "power", "accuracy", "BMI", "age", "goalkeeping")
[…]

### 2022-01-10 _paper.rmd
-updates to use cases, videos, captions, and text. TODO: regression Use cases (not faceting correctly) & videos.

@@
 (ref:casepenguins) Species classification of Palmer penguin data.
-We select a chinstrap penguin that is mislabeled as an adelie.
-By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
+We select a adelie penguin that is mislabeled as an chinstrap.
+Starting from the attribution projection we varying the contribution to `bill length`.
+This explanation is not supported sense by the time there is moderate contribution.
+The current basis is fairly close to the regularly attribution to other chinstrap penguins aside from the sign on flipper length being flipped.
+It feels odd that SHAP has identified a profile where chinstrap and adelie penguins are hard to distinguish when there are separating orientations that should be found within 125 trees.
 The .mp4 animation of this tour can be found at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
@@
 
-## 2) Chocolates, milk/dark chocolate classification
+## Chocolates, milk/dark chocolate classification
 
-The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
+The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels.
 Each of which was labeled as being either milk or dark chocolates.
@@
 (ref:casechocolates) Chocolates data type classification (milk or dark).
-We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
-We vary the contribution of calories from fat.
+We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate from the values on the nutritional label.
+The attribution projection already looks more like a dark chocolate than milk.
+We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `Sugar`.
+The misclassification of two instances seems improbable when sugar is near max contribution.
 Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
@@
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-Calories from fat is the variable with the largest difference in treeshap attribution between these points.
-In this case, it feels strange to call the selected observation a misclassification of the model.
-There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-This feels more like a measurement theory-related problem.
-In this case, the candy is being sold as dark chocolate, while nutritional value more closely resembles milk chocolates.
+We look at one extreme instance and vary the contribution of `sugar`.
+We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label.
+While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+This feels related a measurement theory problem.
 
-## 3) FIFA, wage regression
+## FIFA, wage regression
 
@@
 
-(ref:casefifa) FIFA 2020 data, a random forest model regresses wages [2020 Euros] from aggregated of skill measurements.
+(ref:casefifa) FIFA 2020 data, a random forest model regresses wages [2020 Euros] from nine aggregated of skill measurements.
 The primary observation is a star offensive player (L.
@@
 van Dijk).
+We remove three variables with low attribution from both players.
+The attribution projection starts with selected instance on the right.
+We vary the contribution from `defense`, the star offensive player is not distinguished in the horizontal direction.
+At this point defensive players have been rotated to the highest horizonatal value.
 The animate radial tour can be found at [shorturl.at/lCK26](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4)
@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, offense skills are almost singularly important for explaining the offensive player.
-We vary the contribution of offensive skills.
-Offensive skills are removed in the tour (panel b, frame 3), and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, maintaining some separability.
+While `offense` and `reaction` skills both crutial to the explaintion of a star offensive player.
+We increase the contribution of denfensive skills and Messi is no longer separated from the group while van Dijk and other defensive players are better explained in this case.
 
-## 4) Ames housing 2018, sales price regression
+## Ames housing 2018, sales price regression
 
@@
 (ref:caseames) Ames housing 2018 regressing sales price [2018 USD].
-Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
-This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
+The primary observation is a house that a large `lot area` (`LtA`), but small `living area` (`LvA`).
+The comparison instance is in the opposite; a large house on a relatively small lot.
+Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices.
+Varying the contribution `living area` the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in differet ways.
 The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)

### 2022-01-10 _paper.tex
-updates to use cases, videos, captions, and text. TODO: regression Use cases (not faceting correctly) & videos.

@@
 \PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
 %
@@
 ]{article}
-\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
-\author{}
-\date{\vspace{-2.5em}}
-
 \usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[T1]{fontenc}
@@
   colorlinks=true,
-  linkcolor={red},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={blue},
+  linkcolor=red,
+  filecolor=Maroon,
+  citecolor=Blue,
+  urlcolor=blue,
   pdfcreator={LaTeX via pandoc}}
@@
 \setcounter{secnumdepth}{5}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \usepackage{setspace}
@@
 \usepackage{xcolor}
-\ifLuaTeX
+\ifluatex
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
-A random forest model is fit regressing log wages {[}2020 Euros{]}, from the skill measures.
+A random forest model is fit regressing wages {[}2020 Euros{]}, from the skill measures.
 We then extract the SHAP values of a star offensive player (L.
@@
 In addition, a plot observed by predicted response is also provided.
-In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot (XXX REF).
+In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot (Gabriel 1971).
 A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
@@
 \begin{itemize}
-    \item \textbf{Data:} predictors are unscaled complete numerical matrix; while the response variable is occasionally log scaled for a more Gaussian distribution.
-Most models and local explanations will be scale invariant.
+    \item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale invariant.
     \item \textbf{Model and explanation:} any model can be used with this method.
@@
 This implementation aggregates over exhaustively over the attribution of all trees and we opt to not to fit interactions of features.
-    \item \textbf{Cheem view:} after the model and full explanation space are calculated we scale each feature by the standard deviation away from the mean to achieve a common support for visuals.
-Statistics are calculated on the scaled spaces.
+    \item \textbf{Cheem view:} after the model and full explanation space are calculated we scale each feature by standard deviations away from the mean to achieve a common support for visuals.
+Statistics for mapping to color are calculated on the scaled spaces.
 Interactive tabular display reports the original values.
@@
 \caption{Species classification of Palmer penguin data.
-We select a chinstrap penguin that is mislabeled as an adelie.
-By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
+We select a adelie penguin that is mislabeled as an chinstrap.
+Starting from the attribution projection we varying the contribution to \texttt{bill\ length}.
+This explanation is not supported sense by the time there is moderate contribution.
+The current basis is fairly close to the regularly attribution to other chinstrap penguins aside from the sign on flipper length being flipped.
+It feels odd that SHAP has identified a profile where chinstrap and adelie penguins are hard to distinguish when there are separating orientations that should be found within 125 trees.
 The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}}\label{fig:casepenguins}
@@
 \hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{2) Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
+\subsection{Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
 
-The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
+The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels.
 Each of which was labeled as being either milk or dark chocolates.
@@
 \caption{Chocolates data type classification (milk or dark).
-We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label.
-We vary the contribution of calories from fat.
+We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate from the values on the nutritional label.
+The attribution projection already looks more like a dark chocolate than milk.
+We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{Sugar}.
+The misclassification of two instances seems improbable when sugar is near max contribution.
 Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolates}
@@
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-Calories from fat is the variable with the largest difference in treeshap attribution between these points.
-In this case, it feels strange to call the selected observation a misclassification of the model.
-There are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-This feels more like a measurement theory-related problem.
-In this case, the candy is being sold as dark chocolate, while nutritional value more closely resembles milk chocolates.
+We look at one extreme instance and vary the contribution of \texttt{sugar}.
+We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label.
+While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+This feels related a measurement theory problem.
 
 \hypertarget{fifa-wage-regression}{%
-\subsection{3) FIFA, wage regression}\label{fifa-wage-regression}}
+\subsection{FIFA, wage regression}\label{fifa-wage-regression}}
 
 The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-After aggregation of the skill measurements, we regress the log wages {[}2020 euros{]} given just the skill aggregates.
+After aggregation of the skill measurements, we regress player wages {[}2020 euros{]} given just the skill aggregates.
 The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
@@
 
-\caption{FIFA 2020, regressing log wages {[}2020 Euros{]} from aggregations of skill measurements.
+\caption{FIFA 2020 data, a random forest model regresses wages {[}2020 Euros{]} from nine aggregated of skill measurements.
 The primary observation is a star offensive player (L.
@@
 van Dijk).
+We remove three variables with low attribution from both players.
+The attribution projection starts with selected instance on the right.
+We vary the contribution from \texttt{defense}, the star offensive player is not distinguished in the horizontal direction.
+At this point defensive players have been rotated to the highest horizonatal value.
 The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{shorturl.at/lCK26}}\label{fig:casefifa}
@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-If we remove reaction and movement skills from the basis, offense skills are almost singularly important for explaining the offensive player.
-We vary the contribution of offensive skills.
-Offensive skills are removed in the tour (panel b, frame 3), and Messi is no longer separated from the group.
-We also notice that accuracy has rotated into the frame, maintaining some separability.
+While \texttt{offense} and \texttt{reaction} skills both crutial to the explaintion of a star offensive player.
+We increase the contribution of denfensive skills and Messi is no longer separated from the group while van Dijk and other defensive players are better explained in this case.
 
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{4) Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
+\subsection{Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
 
@@
 
-\caption{Ames housing 2018 regressing log sales price {[}2018 USD{]}.
-Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations.
-This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find.
+\caption{Ames housing 2018 regressing sales price {[}2018 USD{]}.
+The primary observation is a house that a large \texttt{lot\ area} (\texttt{LtA}), but small \texttt{living\ area} (\texttt{LvA}).
+The comparison instance is in the opposite; a large house on a relatively small lot.
+Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices.
+Varying the contribution \texttt{living\ area} the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in differet ways.
 The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}}\label{fig:caseames}
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
+\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
 Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-asimov_grand_1985}{}}%
+\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
 Asimov, Daniel.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
+\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
 Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
 
-\leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
+\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
 ---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
+\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
 Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
+\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
 Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
 
-\leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
+\leavevmode\hypertarget{ref-buja_grand_1986}{}%
 Buja, Andreas, and Daniel Asimov.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chang_shiny_2021}{}}%
+\leavevmode\hypertarget{ref-chang_shiny_2021}{}%
 Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
+\leavevmode\hypertarget{ref-cook_manual_1997}{}%
 Cook, Dianne, and Andreas Buja.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cook_grand_2008}{}}%
+\leavevmode\hypertarget{ref-cook_grand_2008}{}%
 Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-da_silva_projection_2021}{}}%
+\leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
 da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dastin_amazon_2018}{}}%
+\leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
 Dastin, Jeffrey.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-diaz_addressing_2018}{}}%
+\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
 Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-duffy_apple_2019}{}}%
+\leavevmode\hypertarget{ref-duffy_apple_2019}{}%
 Duffy, Claire.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
+\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
+Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
+
+\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
 Gorman, Kristen B., Tony D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
+\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
 Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
+\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
 Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
+\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
 Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-karwowski_international_2006}{}}%
+\leavevmode\hypertarget{ref-karwowski_international_2006}{}%
 Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-kodiyan_overview_2019}{}}%
+\leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
 Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-kominsarczyk_treeshap_2021}{}}%
+\leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
 Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-larson_how_2016}{}}%
+\leavevmode\hypertarget{ref-larson_how_2016}{}%
 Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lee_state_2021}{}}%
+\leavevmode\hypertarget{ref-lee_state_2021}{}%
 Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lee_pptree_2013}{}}%
+\leavevmode\hypertarget{ref-lee_pptree_2013}{}%
 Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
 
-\leavevmode\vadjust pre{\hypertarget{ref-leone_fifa_2020}{}}%
+\leavevmode\hypertarget{ref-leone_fifa_2020}{}%
 Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-liaw_classification_2002}{}}%
+\leavevmode\hypertarget{ref-liaw_classification_2002}{}%
 Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lundberg_consistent_2018}{}}%
+\leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
 Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lundberg_unified_2017}{}}%
+\leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
 Lundberg, Scott M., and Su-In Lee.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-molnar_interpretable_2020}{}}%
+\leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
 Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-oneil_weapons_2016}{}}%
+\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
 O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
 
-\leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
+\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
 Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
+\leavevmode\hypertarget{ref-shapley_value_1953}{}%
 Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
+\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
 Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
 
-\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
+\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
 Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_not_2016}{}}%
+\leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
 Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sievert_interactive_2020}{}}%
+\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
 Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-simonyan_deep_2014}{}}%
+\leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
 Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
+\leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
 Spyrison, Nicholas, and Dianne Cook.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
+\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
 Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
-\leavevmode\vadjust pre{\hypertarget{ref-unwin}{}}%
+\leavevmode\hypertarget{ref-unwin}{}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
+\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
 Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wickham_ggplot2_2016}{}}%
+\leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
 Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-wickham_visualizing_2015}{}}%
+\leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
 Wickham, Hadley, Dianne Cook, and Heike Hofmann.

### 2022-01-10 figures/_make_figures_case_study.r

@@ -6,8 +6,11 @@
   require("dplyr")
   require("ggplot2")
   require("cowplot")
+  require("patchwork")
   
-  setwd("~/R/cheem_paper")
+  wd <- getwd()
+  if(substr(wd, nchar(wd)-10, nchar(wd)) != "cheem_paper")
+    warning("Expected work directory to be cheem_paper.")
   fp <- "../cheem/inst/shiny_apps/cheem_initial/data/"
   if(F)
     dir("../cheem/inst/shiny_apps/cheem_initial/data/")
@@ -18,8 +21,9 @@
   ames2018_ls   <- readRDS(paste0(fp, "preprocess_ames2018.rds"))
   chocolates_ls <- readRDS(paste0(fp, "preprocess_chocolates.rds"))
   
-  .x_axis_title     <- "          x: PC1, y: PC2        x: PC1, y: PC2  x: predicted, y: observed"
-  .x_axis_title_reg <- "          x: PC1, y: PC2             x: PC1, y: PC2     x: predicted, y: observed"
+  .x_title      <- "        x: PC1, y: PC2          x: PC1, y: PC2  x: predicted, y: observed"
+  .x_title_reg  <- "      x: PC1, y: PC2               x: PC1, y: PC2     x: predicted, y: observed"
+  .x_title_reg2 <- "        x: PC1, y: PC2        x: PC1, y: PC2    x: predicted, y: observed"
 }
 
 ## Penguins classification ------
@@ -32,44 +36,58 @@
   ### Global view and cheem tour stills
   .glob_view <- global_view(
     penguins_ls, prim_obs, comp_obs, as_ggplot = TRUE) +
-    theme(legend.position = "right") +
-    labs(color = 'Predicted class', shape = 'Predicted class',
-         x = .x_axis_title) +
-    ggtitle("Global view")
+    labs(color = 'Predicted class', shape = 'Predicted class',x = .x_title) +
+    ggtitle("Global view") + theme( 
+      plot.margin      = margin(0,0,0,0),
+      legend.margin    = margin(0,0,0,0),
+      legend.position  = "bottom",
+      legend.direction = "horizontal")
   .bas <- basis_attr_df(penguins_ls$attr_df, prim_obs)
   .mv  <- 1 ## seems better than the 3 returned from
[…]

### 2022-01-11 _paper.rmd
updated penguin case study but it will need more work

@@
 abstract: |
- Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models.
  The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
 XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation.
-These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
+These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
 Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables.
@@
 
-```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "The cheem view for classification tasks.
+```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks.
 Plots are colored on predicted class, and red circles indicate misclassified observations.
@@
 
-```{r regressioncase, echo=F, out.width = "100%", fig.cap="Overview of the cheem view for tasks.
+```{r regressioncase, echo=F, out.width = "100%", fig.cap="Overview of the cheem viewer for regression tasks.
 The global view can be colored on the correlation between the projection of the data generated by the local explanation and the observed response.
@@
 The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was provided publicly as a substitute for the over-used iris data, and is quite similar in form.
+The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
 There are 330 observations and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
@@
 
-(ref:casepenguins) Species classification of Palmer penguin data.
-We select a adelie penguin that is mislabeled as an chinstrap.
-Starting from the attribution projection we varying the contribution to `bill length`.
+(ref:casepenguins) Examining the SHAP values for a random forest model classifying Palmer penguin species.
+The observation of interest is an Adelie (green) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+XXX needs fixing Starting from the attribution projection the contribution of `b_l` is varied.
 This explanation is not supported sense by the time there is moderate contribution.
@@
 Figure \@ref(fig:casepenguins) shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Plot (a) shows the global view, and plot (b) shows several plots generated with the radial tour.
+Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
+Penguin 124, an Adelie (green), is the observation of interest because it has been misclassified as a Chinstrap (orange).
 
@@
 The observed vs predicted plot reveals a handful of misclassified observations.
-A misclassified point (represented by the asterisk in the global view, and as a solid vertical line in the tour view) is contrasted with a correctly classified point (represented by an $x$ and vertical dashed line).
+A misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view) is contrasted with a correctly classified point (represented by an $\times$ and vertical dotted line).
 
-The radial tour starts from the attribution projection of the misclassified observation (left plot). The important variables identifed by SHAP are `b_d` and `wgt` in contrast to `f_l`.
-
-varies the contribution of bill length (b_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.
+The radial tour starts from the attribution projection of the misclassified observation (left plot).
+The important variables identifed by SHAP in the (wrong) prediction for this observation is mostly `b_d` with smaller contributions of `f_l` and `wgt`.
+This projection is a view where the Gentoo (purple) penguins are most distinguished from Adelie and Chinstrap.
+For this observation we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of variables is not particularly useful.
+To explore this we use the radial tour to vary the contribution of bill length (`b_l`).
+The more `b_l` contributes to the the projection the more this penguin looks like a Chinstrap, that it is mislabelled as.
+This suggests that `b_l` should have been an important variable for explaining the (wrong) prediction.
 

### 2022-01-11 _paper.tex
updated penguin case study but it will need more work

--- 2022-01-08 _paper.tex
@@
 \begin{abstract}
-Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models.
 The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
 XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation.
-These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
+These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
 Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables.
@@
 
-\caption{The cheem view for classification tasks.
+\caption{Overview of the cheem viewer for classification tasks.
 Plots are colored on predicted class, and red circles indicate misclassified observations.
@@
 
-\caption{Overview of the cheem view for tasks.
+\caption{Overview of the cheem viewer for regression tasks.
 The global view can be colored on the correlation between the projection of the data generated by the local explanation and the observed response.
@@
 The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was provided publicly as a substitute for the over-used iris data, and is quite similar in form.
+The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
 There are 330 observations and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
@@
 
-\caption{Species classification of Palmer penguin data.
-We select a chinstrap penguin that is mislabeled as an adelie.
-By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution.
+\caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
+The observation of interest is an Adelie (green) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+XXX needs fixing Starting from the attribution projection the contribution of \texttt{b\_l} is varied.
+This explanation is not supported sense by the time there is moderate contribution.
+The current basis is fairly close to the regularly attribution to other chinstrap penguins aside from the sign on flipper length being flipped.
+It feels odd that SHAP has identified a profile where chinstrap and adelie penguins are hard to distinguish when there are separating orientations that should be found within 125 trees.
 The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}}\label{fig:casepenguins}
@@
 Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Plot (a) shows the global view, and plot (b) shows several plots generated with the radial tour.
+Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
+Penguin 124, an Adelie (green), is the observation of interest because it has been misclassified as a Chinstrap (orange).
 
@@
 The observed vs predicted plot reveals a handful of misclassified observations.
-A misclassified point (represented by the asterisk in the global view, and as a solid vertical line in the tour view) is contrasted with a correctly classified point (represented by an \(x\) and vertical dashed line).
+A misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view) is contrasted with a correctly classified point (represented by an \(\times\) and vertical dotted line).
 
 The radial tour starts from the attribution projection of the misclassified observation (left plot).
-The important variables identifed by SHAP are \texttt{b\_d} and \texttt{wgt} in contrast to \texttt{f\_l}.
-
-varies the contribution of bill length (b\_l) as this variable differs most from the contribution of the comparison observation.
-Downplaying the contribution of bill length is crucial to the linear explanation of this observation being misclassified.
+The important variables identifed by SHAP in the (wrong) prediction for this observation is mostly \texttt{b\_d} with smaller contributions of \texttt{f\_l} and \texttt{wgt}.
+This projection is a view where the Gentoo (purple) penguins are most distinguished from Adelie and Chinstrap.
+For this observation we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of variables is not particularly useful.
+To explore this we use the radial tour to vary the contribution of bill length (\texttt{b\_l}).
+The more \texttt{b\_l} contributes to the the projection the more this penguin looks like a Chinstrap, that it is mislabelled as.
+This suggests that \texttt{b\_l} should have been an important variable for explaining the (wrong) prediction.
 
 \hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{2) Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
+\subsection{Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
 
-The chocolates dataset consists of 88 observations of 10 nutritional measurements from their labels.
+The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels.
 Each of which was labeled as being either milk or dark chocolates.
@@
 \url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gabriel_biplot_1971}{}}%
+Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
 

### 2022-01-11 _paper.rmd
updated chocolate case study but it will need more work

@@
 
-The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels.
-Each of which was labeled as being either milk or dark chocolates.
-We can see if a manufacturer accurately portrays the chocolate with this data.
-We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value.
-We should note that not all chocolates consist wholly of chocolate.
-The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable.
-A random forest model is fit classifying the type of chocolate.
-We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85% cocoa.
+The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+Dark chocolates are typically considered healthier and often can be marketed at a higher price.
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
+<!--We should note that chocolates often have other ingredients will decrease the predictive power of the model nutritional explanatory variable.
+Hot sure that this is correct because these were all plain chocolate not caramel chocolate or with nuts.
+--> A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
+It is compared with a correctly classified dark chocolate.
 
@@
 While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-This feels related a measurement theory problem.
+
+XXX Attribution projection shows difference between milk and dark is mostly fiber.
+Is this one, considered similar to the small group of (predicted) milk chocolates that have high fiber and low sugar?
+there has to be another variable that is generating this little bubble.
+
+XXX Need to re-do.
+Not sure that sugar is the right variable to modify.
+Teh difference between milk and dark is typically fiber, and sugars.
+So maybe this is a low fiber dark chocolate?
+It is funny that in whatever view we look at it there is no doubt that this looks like a dark chocolate!
+Fiber has the most contribution
 

### 2022-01-11 _paper.tex
updated chocolate case study but it will need more work

@@
 
-The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels.
-Each of which was labeled as being either milk or dark chocolates.
-We can see if a manufacturer accurately portrays the chocolate with this data.
-We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value.
-We should note that not all chocolates consist wholly of chocolate.
-The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable.
-A random forest model is fit classifying the type of chocolate.
-We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85\% cocoa.
+The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+Dark chocolates are typically considered healthier and often can be marketed at a higher price.
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
+A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
+It is compared with a correctly classified dark chocolate.
 

### 2022-01-11 _paper.rmd
looked at the regression case studies but they need more work

@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-While `offense` and `reaction` skills both crutial to the explaintion of a star offensive player.
+While `offense` and `reaction` skills both crutial to the explanation of a star offensive player.
 We increase the contribution of denfensive skills and Messi is no longer separated from the group while van Dijk and other defensive players are better explained in this case.
+
+XXX think about the projection as a what if. So for Messi, if his new stats were .... (those shown in second projection) then is salary would be predicted to be halved.
 
@@
 
+XXX Observation 170, big residual; maybe look at obs 33
+
 # Discussion {#sec:cheemdiscussion}

### 2022-01-11 _paper.tex
looked at the regression case studies but they need more work

@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-While \texttt{offense} and \texttt{reaction} skills both crutial to the explaintion of a star offensive player.
+While \texttt{offense} and \texttt{reaction} skills both crutial to the explanation of a star offensive player.
 We increase the contribution of denfensive skills and Messi is no longer separated from the group while van Dijk and other defensive players are better explained in this case.
+
+XXX think about the projection as a what if. So for Messi, if his new stats were \ldots. (those shown in second projection) then is salary would be predicted to be halved.
 
@@
 Knowing these singular bases can point toward others with meaningful data variation.
+
+XXX Observation 170, big residual; maybe look at obs 33
 

### 2022-01-11 figures/_make_figures_case_study.r
Checking the fesibility pointing to .html animations.

@@ -9,7 +9,7 @@
   require("patchwork")
   
   wd <- getwd()
-  if(substr(wd, nchar(wd)-10, nchar(wd)) != "cheem_paper")
+  if(substr(wd, nchar(wd) - 10, nchar(wd)) != "cheem_paper")
     warning("Expected work directory to be cheem_paper.")
   fp <- "../cheem/inst/shiny_apps/cheem_initial/data/"
   if(F)
@@ -205,16 +205,28 @@ ggplot2::ggsave(
 .m <- gc()
 
 ### Save .mp4, add GitHub urls to paper
-.ggt <- radial_cheem_tour(
+.ggt <- cheem:::radial_cheem_tour_subplots(
   fifa_ls, basis = .bas, manip_var = .mv,
   primary_obs = prim_obs, comparison_obs = comp_obs,
   do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
-  pcp_shape = 124, angle = .15) + ## realistic angle
-  theme(legend.position = "off")
-.anim <- animate_gganimate(
-  .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom
-  height = 800, width = 800, units = "px", ## "px", "in", "cm", or "mm."
-  render = gganimate::av_renderer("./figures/case_fifa.mp4")) ## Alternative render
+  pcp_shape = 124, angle = .15)
+.anim <- animate_plotly(.ggt, fps = 6)
+htmlwidgets::saveWidget(widget = .anim,
+                        file = "./figures/case_fifa.html",
+                        selfcontained = TRUE)
+## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.html
+if(F){
+  .ggt <- radial_cheem_tour(
+    fifa_ls, basis = .bas, manip_var = .mv,
+    primary_obs = prim_obs, comparison_obs = comp_obs,
+    do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
+    pcp_shape = 124, angle = .15) + ## realistic angle
+    theme(legend.position = "off")
+  .anim <- animate_gganimate(
+    .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom
+    height = 800, width = 800, units = "px", ## "px", "in", "cm", or "mm."
[…]

### 2022-01-11 figures/_make_figures_case_study.r
Manually capture Ames and FIFA from app

@@ -213,7 +213,8 @@ ggplot2::ggsave(
 .anim <- animate_plotly(.ggt, fps = 6)
 htmlwidgets::saveWidget(widget = .anim,
                         file = "./figures/case_fifa.html",
-                        selfcontained = TRUE)
+                        selfcontained = FALSE,
+                        title = "FIFA 2020, cheem radial tour")
 ## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.html
 if(F){
   .ggt <- radial_cheem_tour(
@@ -293,7 +294,8 @@ ggplot2::ggsave(
 .anim <- animate_plotly(.ggt, fps = 6)
 htmlwidgets::saveWidget(widget = .anim,
                         file = "./figures/case_ames2018.html",
-                        selfcontained = TRUE)
+                        selfcontained = FALSE,
+                        title = "North Ames 2018, cheem radial tour")
 ## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.html
 
 .ggt <- radial_cheem_tour(

### 2022-01-11 figures/_make_figures_case_study.r
+ script

@@ -210,25 +210,9 @@ ggplot2::ggsave(
   primary_obs = prim_obs, comparison_obs = comp_obs,
   do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
   pcp_shape = 124, angle = .15)
-.anim <- animate_plotly(.ggt, fps = 6)
-htmlwidgets::saveWidget(widget = .anim,
-                        file = "./figures/case_fifa.html",
-                        selfcontained = FALSE,
-                        title = "FIFA 2020, cheem radial tour")
+animate_plotly(.ggt, fps = 6)
+messgae("Manual capturing this tour from app.")
 ## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.html
-if(F){
-  .ggt <- radial_cheem_tour(
-    fifa_ls, basis = .bas, manip_var = .mv,
-    primary_obs = prim_obs, comparison_obs = comp_obs,
-    do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
-    pcp_shape = 124, angle = .15) + ## realistic angle
-    theme(legend.position = "off")
-  .anim <- animate_gganimate(
-    .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom
-    height = 800, width = 800, units = "px", ## "px", "in", "cm", or "mm."
-    render = gganimate::av_renderer("./figures/case_fifa.mp4")) ## Alternative render
-}
-## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4
 
 
 ## Ames Housing 2018 (North Ames) ----
@@ -291,41 +275,8 @@ ggplot2::ggsave(
   primary_obs = prim_obs, comparison_obs = comp_obs,
   do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
   pcp_shape = 124, angle = .15)
-.anim <- animate_plotly(.ggt, fps = 6)
-htmlwidgets::saveWidget(widget = .anim,
-                        file = "./figures/case_ames2018.html",
-                        selfcontained = FALSE,
-                        title = "North Ames 2018, cheem radial tour")
-## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.html
-
-.ggt <- radial_cheem_tour(
-  ames2018_ls, basis = .bas, manip_var = .mv,
[…]

### 2022-01-11 _paper.rmd
chocolates changed to vary fiber from sugar.

@@
 
+XXX: NS: not as bad as I remember, but a couple, remove them?
+```{r}
+cheem::chocolates$Name[c(26, 74)]
+```
+
 (ref:casechocolates) Chocolates data type classification (milk or dark).
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `Sugar`.
+We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `fiber`.
 The misclassification of two instances seems improbable when sugar is near max contribution.
@@
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of `sugar`.
-We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label.
-While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+We look at one extreme instance and vary the contribution of `fiber`.
+We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
+While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
 
@@
 Fiber has the most contribution
+
+NS: Sugar is slightly better imo, but fiber is fine.
 
@@
 Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices.
-Varying the contribution `living area` the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in differet ways.
+Varying the contribution `living area` the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in different ways.
 The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)

### 2022-01-11 _paper.tex
chocolates changed to vary fiber from sugar.

@@
 \usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[T1]{fontenc}
@@
 \usepackage{xcolor}
-\ifLuaTeX
+\ifluatex
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
+XXX: NS: not as bad as I remember, but a couple, remove them?
+
+\begin{verbatim}
+## [1] "Dark w/ Honey and Almond Nougat" "Milk w/ Honey and Almond Nougat"
+\end{verbatim}
+
 \begin{figure}
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{Sugar}.
+We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
 The misclassification of two instances seems improbable when sugar is near max contribution.
@@
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of \texttt{sugar}.
-We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label.
-While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
+We look at one extreme instance and vary the contribution of \texttt{fiber}.
+We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
+While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
 
@@
 Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices.
-Varying the contribution \texttt{living\ area} the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in differet ways.
+Varying the contribution \texttt{living\ area} the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in different ways.
 The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}}\label{fig:caseames}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gabriel_biplot_1971}{}}%
+\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
 Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
 
-\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
+\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
 Gorman, Kristen B., Tony D.

### 2022-01-11 figures/_make_figures_case_study.r

@@ -114,7 +114,7 @@ ggplot2::ggsave(
   .inc_var_nms <- c("Calories", "SatFat", "Chol", "Na", "Fiber", "Sugars")
   ## Removed 4 with lowest contribution for the prim_obs.
   .bas <- basis_attr_df(chocolates_ls$attr_df[, .inc_var_nms], prim_obs)
-  .mv  <- which(.inc_var_nms == "Sugars")
+  .mv  <- which(.inc_var_nms == "Fiber")
   mt_interp <- manual_tour(.bas, .mv) %>%
     spinifex:::interpolate_manual_tour(.15) ## App angle.
   dim(mt_interp)

### 2022-01-12 _paper.rmd
- Text for chocolates and ames2018. - aspect ratio for chocholates/penguins.

@@
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
-<!--We should note that chocolates often have other ingredients will decrease the predictive power of the model nutritional explanatory variable.
-Hot sure that this is correct because these were all plain chocolate not caramel chocolate or with nuts.
---> A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
+A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
 It is compared with a correctly classified dark chocolate.
-
-XXX: NS: not as bad as I remember, but a couple, remove them?
-```{r}
-cheem::chocolates$Name[c(26, 74)]
-```
 
@@
 While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-
-XXX Attribution projection shows difference between milk and dark is mostly fiber.
-Is this one, considered similar to the small group of (predicted) milk chocolates that have high fiber and low sugar?
-there has to be another variable that is generating this little bubble.
-
-XXX Need to re-do.
-Not sure that sugar is the right variable to modify.
-Teh difference between milk and dark is typically fiber, and sugars.
-So maybe this is a low fiber dark chocolate?
-It is funny that in whatever view we look at it there is no doubt that this looks like a dark chocolate!
-Fiber has the most contribution
-
-NS: Sugar is slightly better imo, but fiber is fine.
+In the penguin case we explored how the model misclassified an observation from a type one error perspective.
+In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.
 
@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-While `offense` and `reaction` skills both crutial to the explanation of a star offensive player.
-We increase the contribution of denfensive skills and Messi is no longer separated from the group while van Dijk and other defensive players are better explained in this case.
-
-XXX think about the projection as a what if. So for Messi, if his new stats were .... (those shown in second projection) then is salary would be predicted to be halved.
+While `offense` and `reaction` skills both crucial to the explanation of a star offensive player.
+As the contribution of defensive skills increases Messi's is no longer separated from the group and other defensive players are better predicted in this attribution case.
+In terms of what-if analysis, were Messi's tree SHAP attributions at theses levels, his predicted wages would be halved.
 
@@
 
-Figure \@ref(fig:caseames) shows the global view and extrema of the tour.
-The horizontal distance in the tour did not show a significant disparity between our selected points.
-This is not particularly surprising as most variables have a sizable contribution.
-Rotating any one variable out of the frame will rotate other vital variables into the frame, preserving most of the distance from intercept to prediction.
-However, the tour has revealed an interesting feature worth discussing.
-Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames one and two of panel b) the data is near a singular profile.
-This means that there is a basis orthogonal to this point that describes sizable variation.
-Knowing these singular bases can point toward others with meaningful data variation.
+Figure \@ref(fig:caseames) selects a sale of with relatively large lot size and small house compared with the opposite.
+Contribution of `living area` is crucial to the predictions of each observation though the sign varies.
+The attribution of `lot area` is not particularity essential to either observation in isolation.
+Perhaps this is because this variable interacts with exogenous geometry and layout information, or with endogenous information maybe the year build and overall quality holds redundant information.
+As the contribution of `living area` is changed it crosses a basis where, if the explanation were this attribution, the predicted house sale would be equal.
 
 XXX Observation 170, big residual; maybe look at obs 33
+NS Looking  at old version? current Ames housing obs:
+prim_obs <- 311 ## Large lot area to living area ratio ## small house, big lot
+comp_obs <- 220 ## Small on that ratio ## large house, small lot
 

### 2022-01-12 _paper.tex
- Text for chocolates and ames2018. - aspect ratio for chocholates/penguins.

@@
 
-XXX: NS: not as bad as I remember, but a couple, remove them?
-
-\begin{verbatim}
-## [1] "Dark w/ Honey and Almond Nougat" "Milk w/ Honey and Almond Nougat"
-\end{verbatim}
-
 \begin{figure}
@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-While \texttt{offense} and \texttt{reaction} skills both crutial to the explanation of a star offensive player.
-We increase the contribution of denfensive skills and Messi is no longer separated from the group while van Dijk and other defensive players are better explained in this case.
-
-XXX think about the projection as a what if. So for Messi, if his new stats were \ldots. (those shown in second projection) then is salary would be predicted to be halved.
+While \texttt{offense} and \texttt{reaction} skills both crucial to the explanation of a star offensive player.
+As the contribution of defensive skills increases Messi's is no longer separated from the group and other defensive players are better predicted in this attribution case.
+In terms of what-if analysis, were Messi's tree SHAP attributions at theses levels, his predicted wages would be halved.
 
@@
 
-Figure \ref{fig:caseames} shows the global view and extrema of the tour.
-The horizontal distance in the tour did not show a significant disparity between our selected points.
-This is not particularly surprising as most variables have a sizable contribution.
-Rotating any one variable out of the frame will rotate other vital variables into the frame, preserving most of the distance from intercept to prediction.
-However, the tour has revealed an interesting feature worth discussing.
-Notice that the observations pivot about the origin, the basis roughly halfway between bases in frames one and two of panel b) the data is near a singular profile.
-This means that there is a basis orthogonal to this point that describes sizable variation.
-Knowing these singular bases can point toward others with meaningful data variation.
+Figure \ref{fig:caseames} selects a sale of with relatively large lot size and small house compared with the opposite.
+Contribution of \texttt{living\ area} is crucial to the predictions of each observation though the sign varies.
+The attribution of \texttt{lot\ area} is not particularity essential to either observation in isolation.
+Perhaps this is because this variable interacts with exogenous geometry and layout information, or with endogenous information maybe the year build and overall quality holds redundant information.
+As the contribution of \texttt{living\ area} is changed it crosses a basis where, if the explanation were this attribution, the predicted house sale would be equal.
 
 XXX Observation 170, big residual; maybe look at obs 33
+NS Looking at old version? current Ames housing obs:
+prim\_obs \textless- 311 \#\# Large lot area to living area ratio \#\# small house, big lot
+comp\_obs \textless- 220 \#\# Small on that ratio \#\# large house, small lot
 

### 2022-01-12 _paper.rmd
revised wording on penguins example to clarify subset

--- 2022-01-11 _paper.rmd
@@
 The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
-There are 330 observations and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
+The subset used here has missing values removed (as used in the `spinifex` package), and there are 333 observations and four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
 A random forest model was fitted with species as the response variable.

### 2022-01-12 _paper.tex
revised wording on penguins example to clarify subset

--- 2022-01-11 _paper.tex
@@
 The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
-There are 330 observations and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
+The subset used here has missing values removed (as used in the \texttt{spinifex} package), and there are 333 observations and four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
 A random forest model was fitted with species as the response variable.

### 2022-01-12 _paper.rmd
chocolates case study re-written

@@
 The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
-Dark chocolates are typically considered healthier and often can be marketed at a higher price.
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
-A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
-It is compared with a correctly classified dark chocolate.
+Dark chocolate is considered healthier than milk.
+The data was collected by students during the Iowa State University  class STAT503 from nutritional information from the manufacturer's web site, and normalized to 100g equivalents.
+The data is available in the `cheem` package.
+A random forest model is used for the classification of chocolate type.
+
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
+A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered as a healthy alternative.
+It is interesting to explore which of the nutritional variables contribute most to the misclassification.
 
@@
 
-Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space.
+This type of exploration is shown in \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
+It is compared with a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are clearer in the observed vs predicted plot, and can be seens to be mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+The opposite would be the interpretation of a point with high values in this plot.
+The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fibre and low sugar.
+The density for milk chocolates is further to the right, indicating that they generally have low fibre and high sugar.
+
+The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot which is the same projection of all observations as specifed by the SHAP values of the observation of interest.
+
+From the parallel coordinate plot we can quickly compare the variable contributions to the two different predictions.
+The observation of interest differs form the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding to the observation of interest, both observations are more like dark chocolates.
+If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
+When primarily Fiber is examined observation 22 looks more like a milk chocolate.
+
+<!-- Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space.
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
@@
 In the penguin case we explored how the model misclassified an observation from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.
+In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.-->
 

### 2022-01-12 _paper.tex
chocolates case study re-written

@@
 The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
-Dark chocolates are typically considered healthier and often can be marketed at a higher price.
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
-A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
-It is compared with a correctly classified dark chocolate.
+Dark chocolate is considered healthier than milk.
+The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's web site, and normalized to 100g equivalents.
+The data is available in the \texttt{cheem} package.
+A random forest model is used for the classification of chocolate type.
+
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
+A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered as a healthy alternative.
+It is interesting to explore which of the nutritional variables contribute most to the misclassification.
 
@@
 
-Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space.
-Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of \texttt{fiber}.
-We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
-While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an observation from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.
+This type of exploration is shown in \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
+It is compared with a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are clearer in the observed vs predicted plot, and can be seens to be mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+The opposite would be the interpretation of a point with high values in this plot.
+The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fibre and low sugar.
+The density for milk chocolates is further to the right, indicating that they generally have low fibre and high sugar.
+
+The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot which is the same projection of all observations as specifed by the SHAP values of the observation of interest.
+
+From the parallel coordinate plot we can quickly compare the variable contributions to the two different predictions.
+The observation of interest differs form the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding to the observation of interest, both observations are more like dark chocolates.
+If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
+When primarily Fiber is examined observation 22 looks more like a milk chocolate.
 

### 2022-01-12 _paper.rmd
ames housing needs to have variables explained
-Penguin, Chocolate, and Ames update -Introduce PI/CI, Obs->inst, var->feat

--- 2022-01-12 _paper.rmd
@@
  The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation.
-These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables.
+XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear feature importance in the vicinity of each instance.
+These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between features used to make predictions, across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any instance, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
@@
 There are different reasons and emphases to fit a model.
-@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomizes modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample observations.
+@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomizes modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample instances.
 The intended use has important implications for model selection and development.
@@
 sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]) in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) variable, which is then built into the model.
-Variable-level (feature-level) interpretability of models is essential in the evaluation models for such biases.
+For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
+Feature-level (variable-level) interpretability of models is essential in the evaluation models for such biases.
 It is also generally important for many problems, where it is not enough to accurately predict accurately but one must be able to explain which predictors are most responsible in generating a response value.
@@
 <!-- Interpretability & data drift -->
-Another concern is that of data drift, which is a shift in support or domain of the explanatory variables (feature or predictors).
+Another concern is that of data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
 Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
@@
 Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
-A common approach is to use _local explanations_, which attempt to approximate linear variable importance at the location of each observation (instance), or the predictions at a specific point in the data domain.
+A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
 Because these are point specific, a challenge is visualizing them to more comprehensively understand a model.
@@
 Tours are viewed as an animation over minor changes to the projection basis.
-Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
+Structure in a projection can then be explored visually to see which features contribute to the formation of that structure.
 The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information of the structure and features of the object.
@@
 There are various types of tours distinguished by the generation of projection bases.
-In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected variable.
+In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected feature.
 <!--tours and models --> Applying tours to models has been done in a couple of contexts.
@@
 In our proposed approach we use the radial manual tour to scrutinize a local explanation.
-Additional interactivity allows the user to identify an observation of interest, then explore its local explanation by changing variable contribution with the radial tour.
+Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in R package __cheem__.
@@
 Using a radial tour can be considered similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020].
-This phrase, Latin for "other things held constant" or "all else unchanged", shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
-It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
+This phrase, Latin for "other things held constant" or "all else unchanged", shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+It ignores correlations of the features and imagines a case that was not observed.
+In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object.
 A constraint in our approach is that the basis must remain orthonormal.
-That means, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+That means, when the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
 This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions, and is thus an interpretable visualisation.
@@
 Section \@ref(sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-Then Section \@ref(#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response variables.
+Then Section \@ref(#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
 We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explore in the future.
@@
 Consider a highly non-linear model.
-It can be hard to determine whether small changes in a variable's value will make a class prediction change group or identify which variables contribute to an extreme residual.
-Local explanations shed light on these situations by approximating linear variable importance in the vicinity of a single observation or point.
+It can be hard to determine whether small changes in a feature's value will make a class prediction change group or identify which features contribute to an extreme residual.
+Local explanations shed light on these situations by approximating linear feature importance in the vicinity of a single instance.
 
@@
 In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
-In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction [@ribeiro_why_2016].
+In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction [@ribeiro_why_2016].
 
 <!-- SHAP -->
-SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation conceptually by examining the effect of other variables on the contribution of the variable of interest on predicting the response.
+SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
 This explanations almost all point to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
@@
 The SHAP values are the mean contributions over some subset these permutations.
-This is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
-Partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one instance.
-It could also be considered to be similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each variable in the context of all other candidate variables.
+This is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
+It could also be considered to be similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 For our application, we use _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
-Instead of aggregating over permutations of the variable order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
+Instead of aggregating over permutations of the feature order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
@@
 Messi and V van Dijk respectively.
-Plot a) is the break-down plot showing one additive variable-attributution.
-These attributions are sensitive to the order of the variables.
-Plot b) shows the distribution of attribution for each variable across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Plot a) is the break-down plot showing one additive feature-attributution.
+These attributions are sensitive to the order of the features.
+Plot b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi."}
@@
 Consider soccer data from the FIFA 2020 season [@leone_fifa_2020].
-There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
+There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
 A random forest model is fit regressing wages [2020 Euros], from the skill measures.
@@
 The results are displayed in Figure \@ref(fig:shapdistrbd).
-We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
-Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of variables is fixed, so the two observations can be more easily compared.
+We expect to see a difference in the attribution of the feature importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
+Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of features is fixed, so the two instances can be more easily compared.
 
 <!-- Segue -->
-In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
+In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
+In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
 
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of observations in time, and as such gain information about the relationships between points across multiple variables.
+One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time, and as such gain information about the relationships between points across multiple features.
 There are various types of tours that are distinguished by how the paths are generated [@lee_state_2021; @cook_grand_2008].
@@
 <!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its path by changing a selected variable's contribution to a basis, to allow the variable to contribute more or less to the projection.
-The contribution of all other variables is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
+The manual tour [@cook_manual_1997] defines its path by changing a selected feature's contribution to a basis, to allow the feature to contribute more or less to the projection.
+The contribution of all other features is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
 It also lends itself to pre-computation to be queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
 <!-- However, this navigation is relatively time-consuming due to the vast volume of the display space.
-It is advisable to use this method to explore the sensitivity of the variable contribution to a previously identified feature of interest.
+It is advisable to use this method to explore the sensitivity of the feature contribution to a previously identified feature of interest.
 In this case, the projection of the normalized explanations is the feature of interest.-->
@@
 A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of the new work.
-In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and also interactive re-calculation to focus on a different variable.
+In a radial tour, the selected feature is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and also interactive re-calculation to focus on a different feature.
 
@@
 There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
-In addition, there are numerous user inputs, including variable selection for the radial tour, and instance selection for making comparisons.
-Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) contain screenshots showing the cheem viewer for the two primary tasks classification (categorical response) and regression (quantitative response).
+In addition, there are numerous user inputs, including feature selection for the radial tour, and instance selection for making comparisons.
+Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 <!-- Purpose -->
-The global view provides the context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
-Both of these spaces are of dimension $n\times p$, where $n$ is the number of observations and $p$ is the number of predictors.
-The attribution space corresponds to the local explanations for each observation, which will have $p$ values for each observation.
+The global view provides the context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
+Both of these spaces are of dimension $n\times p$, where $n$ is the number of instances and $p$ is the number of predictors.
+The attribution space corresponds to the local explanations for each instance, which will have $p$ values for each instance.
 
@@
 In addition, a plot observed by predicted response is also provided.
-In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot [@gabriel_biplot_1971].
+In both PCA plots the orientation and magnitude of the features are inscribed on a unit circle, similar to a biplot [@gabriel_biplot_1971].
 A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
-<!-- Interactions --> For classification tasks, misclassified observations are circled in red if applicable.
+<!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
 Linked brushing between the plots is provided and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
 
-While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of observations, from which to explore the local explanations.
+While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances, from which to explore the local explanations.
+The projection attribution of the Primary Instance (PI) is explored and typically viewed with an optional Comparison Instance (CI).
 
@@
 <!-- A function of the obs selected -->
-The global view facilitated the selection of a primary and optional comparison observation.
-The variable-level attribution of the primary observation is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observation's prediction, not the local shape of the model surface.
+The global view facilitated the selection instances.
+The feature-level attribution of the primary instance is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of the contributions of the linear features that best explain the difference between the model intercept and an instance's prediction, not the local shape of the model surface.
 
 <!-- Tour start frame -->
-The initial frame is the normalized SHAP values of the primary observation.
-The current projection basis is depicted as the width of a bar, the variable's contribution to the horizontal axis.
-The normalized values of all observations are shown as vertical parallel coordinate plots.
+The initial frame is the normalized SHAP values of the PI.
+The current projection basis is depicted as the width of a bar, the feature's contribution to the horizontal axis.
+The normalized values of all instances are shown as vertical parallel coordinate plots.
 
 <!-- Tour animation -->
-The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
-The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The radial tour creates a basis path by varying the contribution of a selected feature, fully into and out of a projection frame.
+Doing so tests an individual feature's sensitivity to the structure identified by the local explanation.
+The default feature selected has the largest discrepancy between the attribution of primary and comparison instances.
 The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
@@
 
-Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
+Typically we select a misclassified instance compared to a correctly classified point nearby in data space.
 The model infomation in the global view is a depiction of the model conmfusion matrix.
-The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each variable to structure identified by the local explanation, evaluating the support or robustness of the prediction.
+The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each feature to structure identified by the local explanation, evaluating the support or robustness of the prediction.
 
 ```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks.
-Plots are colored on predicted class, and red circles indicate misclassified observations.
+Plots are colored on predicted class, and red circles indicate misclassified instances.
 Model visual is essentially a representation of the confusion matrix of the model.
 The radial tour is a 1D projection starting the attribution projection.
-The first frame is the linear variable importances that best describe the observation's prediction.
-We probe the support of variable contributions by selecting a variable and use the radial tour to vary its coefficient."}
+The first frame is the linear feature importances that best describe the instance's prediction.
+We probe the support of feature contributions by selecting a feature and use the radial tour to vary its coefficient."}
 knitr::include_graphics("./figures/app_classification.PNG")
@@
 In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
-The vertical position is fixed to the observed response variable and residuals in the middle and right panels.
+The vertical position is fixed to the observed response feature and residuals in the middle and right panels.
 Correspondingly, the display changes from univariate density to 2D scatterplot.
@@
 <!-- Exploratory interactions -->
-A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+A tooltip displays instance number/name and classification information while the cursor hovers over a point.
 Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
 The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and, finally, identification of a primary and comparison observation.
+These interactions aid exploration of the spaces and, finally, identification of a primary and comparison instance.
 
 ```{r interactions, echo=F, out.width="100%", fig.cap = "Illustration of data explorations interactions in the global view.
-This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an observation."}
+This view has linked brushing, where instances selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
+Tooltips display when hovering over an instance."}
 knitr::include_graphics("./figures/app_interactions.PNG")
@@
 Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)], compatibility tree SHAP.
-We use modest hyperparameters, namely: 125 trees, number variables randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} (ref:citeTs).
+We use modest hyperparameters, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
 This implementation aggregates over exhaustively over the attribution of all trees and we opt to not to fit interactions of features.
@@
 The time to preprocess the data will vary significantly with the model and local explanation.
-For reference, the FIFA data, 5000 observations of nine explanatory variables, took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 270 seconds combined.
-PCA and statistics of the variables and attributions took 2.8 seconds.
+For reference, the FIFA data, 5000 instances of nine explanatory features, took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each instance took 270 seconds combined.
+PCA and statistics of the features and attributions took 2.8 seconds.
 These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
@@
 
+(ref:casepenguinsblfl) An orthogonal view of penguins data showing the PI, a Gentoo (purple) close to the distribution of Chinstrap (orange).
+```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl)"}
+knitr::include_graphics("./figures/case_penguins_BlFl.png")
+```
+
 The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
 The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
-The subset used here has missing values removed (as used in the `spinifex` package), and there are 333 observations and four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
-A random forest model was fitted with species as the response variable.
+After removing incomplete observations, there are 333 instances and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
+A random forest model was fitted with species as the response feature.
 
 (ref:casepenguins) Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The observation of interest is an Adelie (green) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
-XXX needs fixing Starting from the attribution projection the contribution of `b_l` is varied.
+The PI is an Chinstrap (orange) penguin that is misclassified as a Gentoo (purple), marked as an asterisk in (a), and the dashed vertical line in (b).
+Starting from the attribution projection the contribution of `b_l` is varied.
 This explanation is not supported sense by the time there is moderate contribution.
-The current basis is fairly close to the regularly attribution to other chinstrap penguins aside from the sign on flipper length being flipped.
-It feels odd that SHAP has identified a profile where chinstrap and adelie penguins are hard to distinguish when there are separating orientations that should be found within 125 trees.
+The basis showing the correct species separation is fairly close to the regularly attribution to other Gentoo penguins aside from the sign change on flipper length.
 The .mp4 animation of this tour can be found at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
@@
 Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
-Penguin 124, an Adelie (green), is the observation of interest because it has been misclassified as a Chinstrap (orange).
+Penguin 243, an Chinstrap (orange), is the PI because it has been misclassified as a Gentoo (purple).
 
 There is more separation visible in the attribution space than the data space, as would be expected.
-The observed vs predicted plot reveals a handful of misclassified observations.
+The observed vs predicted plot reveals a handful of misclassified instances.
 A misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view) is contrasted with a correctly classified point (represented by an $\times$ and vertical dotted line).
 
-The radial tour starts from the attribution projection of the misclassified observation (left plot).
-The important variables identifed by SHAP in the (wrong) prediction for this observation is mostly `b_d` with smaller contributions of `f_l` and `wgt`.
-This projection is a view where the Gentoo (purple) penguins are most distinguished from Adelie and Chinstrap.
-For this observation we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of variables is not particularly useful.
-To explore this we use the radial tour to vary the contribution of bill length (`b_l`).
-The more `b_l` contributes to the the projection the more this penguin looks like a Chinstrap, that it is mislabelled as.
-This suggests that `b_l` should have been an important variable for explaining the (wrong) prediction.
+The radial tour starts from the attribution projection of the misclassified instance (left plot).
+The important features identifed by SHAP in the (wrong) prediction for this instance is mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
+This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
+For this instance we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of features is not particularly useful.
+To explore this we use the radial tour to vary the contribution of flipper length (`f_l`).
+The more that `f_l` contributes to the the projection the more this penguin looks like a Gentoo, its observed species.
+This suggests that `f_l` should have been an important feature for explaining the (wrong) prediction.
 
@@
 
-The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
 Dark chocolates are typically considered healthier and often can be marketed at a higher price.
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
-A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22).
+A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the instance of interest (chocolate 22).
 It is compared with a correctly classified dark chocolate.
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `fiber`.
+We remove four features with the lowest contribution for the selected instance and vary the contribution of `sugar`.
 The misclassification of two instances seems improbable when sugar is near max contribution.
@@
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of `fiber`.
+We look at one extreme instance and vary the contribution of `sugar`.
 We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
 While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an observation from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.
+In the penguin case we explored how the model misclassified an instance from a type one error perspective.
+In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original instance.
 
@@
 After aggregation of the skill measurements, we regress player wages [2020 euros] given just the skill aggregates.
-The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
+The model was fit from 5000 instances of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
 We compare a leading offensive fielder (L.
 Messi) with that of a top defensive fielder (V.
-van Dijk), the same observations were used in figure \@ref(fig:shapdistrbd).
+van Dijk), the same instances were used in figure \@ref(fig:shapdistrbd).
 
 (ref:casefifa) FIFA 2020 data, a random forest model regresses wages [2020 Euros] from nine aggregated of skill measurements.
-The primary observation is a star offensive player (L.
+The PI is a star offensive player (L.
 Messi) compared with a top defensive player (V.
 van Dijk).
-We remove three variables with low attribution from both players.
+We remove three features with low attribution from both players.
 The attribution projection starts with selected instance on the right.
@@
 Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across nine variables.
-Using interaction from the global view, we select a house with an extreme negative residual and an accurate observation close to it in the data.
+The remaining are 338 house sales across nine features.
+Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
 
 (ref:caseames) Ames housing 2018 regressing sales price [2018 USD].
-The primary observation is a house that a large `lot area` (`LtA`), but small `living area` (`LvA`).
-The comparison instance is in the opposite; a large house on a relatively small lot.
-Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices.
-Varying the contribution `living area` the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in different ways.
+The PI sale price was under predicted, and had sizable attribution to lot area (`LtA`).
+The CI was predicted sales price was similar and much more accurate with its observed sales price while it has very little attribution to lot area.
+Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of `LtA`,  which is important to explaining the PI and near invariant to the sales price of the CI.
 The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
@@
 
-Figure \@ref(fig:caseames) selects a sale of with relatively large lot size and small house compared with the opposite.
-Contribution of `living area` is crucial to the predictions of each observation though the sign varies.
-The attribution of `lot area` is not particularity essential to either observation in isolation.
-Perhaps this is because this variable interacts with exogenous geometry and layout information, or with endogenous information maybe the year build and overall quality holds redundant information.
-As the contribution of `living area` is changed it crosses a basis where, if the explanation were this attribution, the predicted house sale would be equal.
-
-XXX Observation 170, big residual; maybe look at obs 33
-NS Looking  at old version? current Ames housing obs:
-prim_obs <- 311 ## Large lot area to living area ratio ## small house, big lot
-comp_obs <- 220 ## Small on that ratio ## large house, small lot
+Figure \@ref(fig:caseames) selects the house sale 74, a sizabile under prediction that has a large contribution to lot area.
+The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
+As the contribution of lot area decreases the predictive power decreases for the PI, while the CI remains stationary.
 
@@
 The need to maintain the interpretability of black-box models is evident.
-One aspect uses local explanations of the model in the vicinity of an observation.
-Local explanations approximate the linear variable importance to the model.
+One aspect uses local explanations of the model in the vicinity of an instance.
+Local explanations approximate the linear feature importance to the model.
 Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify primary and comparison observations of interest.
-The normalized linear importance from the explanation of the primary observation becomes the feature of interest to further explore with the radial tour.
-The tours explore the variable sensitivity to the structure identified in the explanation.
+First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify instances of interest.
+The normalized linear importance from the explanation of the PI becomes the feature of interest to further explore with the radial tour.
+The tours explore the feature sensitivity to the structure identified in the explanation.
 

### 2022-01-12 _paper.tex
-Penguin, Chocolate, and Ames update -Introduce PI/CI, Obs->inst, var->feat

--- 2022-01-12 _paper.tex
@@
 The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation.
-These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables.
+XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear feature importance in the vicinity of each instance.
+These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between features used to make predictions, across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any instance, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
 The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
@@
 There are different reasons and emphases to fit a model.
-Breiman (2001), reiterated by Shmueli (2010), taxonomizes modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample observations.
+Breiman (2001), reiterated by Shmueli (2010), taxonomizes modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample instances.
 The intended use has important implications for model selection and development.
@@
 2018)) in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) variable, which is then built into the model.
-Variable-level (feature-level) interpretability of models is essential in the evaluation models for such biases.
+For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
+Feature-level (variable-level) interpretability of models is essential in the evaluation models for such biases.
 It is also generally important for many problems, where it is not enough to accurately predict accurately but one must be able to explain which predictors are most responsible in generating a response value.
 
-Another concern is that of data drift, which is a shift in support or domain of the explanatory variables (feature or predictors).
+Another concern is that of data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
 Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
@@
 Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
-A common approach is to use \emph{local explanations}, which attempt to approximate linear variable importance at the location of each observation (instance), or the predictions at a specific point in the data domain.
+A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
 Because these are point specific, a challenge is visualizing them to more comprehensively understand a model.
@@
 There are various types of tours distinguished by the generation of projection bases.
-In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected variable.
+In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected feature.
 Applying tours to models has been done in a couple of contexts.
@@
 In our proposed approach we use the radial manual tour to scrutinize a local explanation.
-Additional interactivity allows the user to identify an observation of interest, then explore its local explanation by changing variable contribution with the radial tour.
+Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in R package \textbf{cheem}.
@@
 Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-This phrase, Latin for ``other things held constant'' or ``all else unchanged'', shows how an observation's prediction would change from a marginal change in one explanatory variable given that other variables are held constant.
-It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
+This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+It ignores correlations of the features and imagines a case that was not observed.
+In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object.
 A constraint in our approach is that the basis must remain orthonormal.
-That means, when the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+That means, when the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
 This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions, and is thus an interpretable visualisation.
@@
 Section \ref{sec:applicationdesign} discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-Then Section @ref(\#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response variables.
+Then Section @ref(\#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
 We conclude with Section \ref{sec:cheemdiscussion} of the insights gained and directions that might be explore in the future.
@@
 2018).
-In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
+In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction (Ribeiro, Singh, and Guestrin 2016).
 
-SHaply Additive exPlanations (SHAP) approximates the variable importance in the vicinity of one observation conceptually by examining the effect of other variables on the contribution of the variable of interest on predicting the response.
+SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
 This explanations almost all point to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
@@
 The SHAP values are the mean contributions over some subset these permutations.
-This is also related to partial dependence plots (Molnar 2020), used to explain the effect of a variable by predicting the response for a range of values on this variable, after fixing the value of all other variables to their mean.
-Partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one instance.
-It could also be considered to be similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each variable in the context of all other candidate variables.
+This is also related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
+It could also be considered to be similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each feature in the context of all other candidate features.
 
 For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
-Instead of aggregating over permutations of the variable order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
+Instead of aggregating over permutations of the feature order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
@@
 Messi and V van Dijk respectively.
-Plot a) is the break-down plot showing one additive variable-attributution.
-These attributions are sensitive to the order of the variables.
-Plot b) shows the distribution of attribution for each variable across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Plot a) is the break-down plot showing one additive feature-attributution.
+These attributions are sensitive to the order of the features.
+Plot b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
@@
 Consider soccer data from the FIFA 2020 season (Leone 2020).
-There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
+There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
 A random forest model is fit regressing wages {[}2020 Euros{]}, from the skill measures.
@@
 The results are displayed in Figure \ref{fig:shapdistrbd}.
-We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
-Plot (b) is a modified breakdown plot (Gosiewska and Biecek 2019) where the order of variables is fixed, so the two observations can be more easily compared.
+We expect to see a difference in the attribution of the feature importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
+Plot (b) is a modified breakdown plot (Gosiewska and Biecek 2019) where the order of features is fixed, so the two instances can be more easily compared.
 
-In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this instance, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
+In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
+In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
 
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of observations in time, and as such gain information about the relationships between points across multiple variables.
+One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time, and as such gain information about the relationships between points across multiple features.
 There are various types of tours that are distinguished by how the paths are generated (S.
@@
 
-The manual tour (Cook and Buja 1997) defines its path by changing a selected variable's contribution to a basis, to allow the variable to contribute more or less to the projection.
-The contribution of all other variables is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
+The manual tour (Cook and Buja 1997) defines its path by changing a selected feature's contribution to a basis, to allow the feature to contribute more or less to the projection.
+The contribution of all other features is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
 It also lends itself to pre-computation to be queued in advance or computed on-the-fly for human-in-the-loop analysis (Karwowski 2006).
@@
 A version of the manual tour called a \emph{radial tour} is implemented in Spyrison and Cook (2020) and forms the basis of the new work.
-In a radial tour, the selected variable is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and also interactive re-calculation to focus on a different variable.
+In a radial tour, the selected feature is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and also interactive re-calculation to focus on a different feature.
 
@@
 There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
-In addition, there are numerous user inputs, including variable selection for the radial tour, and instance selection for making comparisons.
-Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} contain screenshots showing the cheem viewer for the two primary tasks classification (categorical response) and regression (quantitative response).
+In addition, there are numerous user inputs, including feature selection for the radial tour, and instance selection for making comparisons.
+Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 
-The global view provides the context of all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
-Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of observations and \(p\) is the number of predictors.
-The attribution space corresponds to the local explanations for each observation, which will have \(p\) values for each observation.
+The global view provides the context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
+Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
+The attribution space corresponds to the local explanations for each instance, which will have \(p\) values for each instance.
 
@@
 In addition, a plot observed by predicted response is also provided.
-In both PCA plots the orientation and magnitude of the variables are inscribed on a unit circle, similar to a biplot (Gabriel 1971).
+In both PCA plots the orientation and magnitude of the features are inscribed on a unit circle, similar to a biplot (Gabriel 1971).
 A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
-For classification tasks, misclassified observations are circled in red if applicable.
+For classification tasks, misclassified instances are circled in red if applicable.
 Linked brushing between the plots is provided and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
 
-While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of observations, from which to explore the local explanations.
+While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances, from which to explore the local explanations.
+The projection attribution of the Primary Instance (PI) is explored and typically viewed with an optional Comparison Instance (CI).
 
@@
 
-The global view facilitated the selection of a primary and optional comparison observation.
-The variable-level attribution of the primary observation is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of the contributions of the linear variables that best explain the difference between the model intercept and an observation's prediction, not the local shape of the model surface.
+The global view facilitated the selection instances.
+The feature-level attribution of the primary instance is normalized and used as the initial 1D basis in a radial tour.
+This is an approximation of the contributions of the linear features that best explain the difference between the model intercept and an instance's prediction, not the local shape of the model surface.
 
-The initial frame is the normalized SHAP values of the primary observation.
-The current projection basis is depicted as the width of a bar, the variable's contribution to the horizontal axis.
-The normalized values of all observations are shown as vertical parallel coordinate plots.
+The initial frame is the normalized SHAP values of the PI.
+The current projection basis is depicted as the width of a bar, the feature's contribution to the horizontal axis.
+The normalized values of all instances are shown as vertical parallel coordinate plots.
 
-The radial tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame.
-Doing so tests an individual variable's sensitivity to the structure identified by the local explanation.
-The default variable selected has the largest discrepancy between the attribution of primary and comparison observations.
+The radial tour creates a basis path by varying the contribution of a selected feature, fully into and out of a projection frame.
+Doing so tests an individual feature's sensitivity to the structure identified by the local explanation.
+The default feature selected has the largest discrepancy between the attribution of primary and comparison instances.
 The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
@@
 
-Typically we select a misclassified observation compared to a correctly classified point nearby in data space.
+Typically we select a misclassified instance compared to a correctly classified point nearby in data space.
 The model infomation in the global view is a depiction of the model conmfusion matrix.
-The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each variable to structure identified by the local explanation, evaluating the support or robustness of the prediction.
+The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each feature to structure identified by the local explanation, evaluating the support or robustness of the prediction.
 
@@
 \caption{Overview of the cheem viewer for classification tasks.
-Plots are colored on predicted class, and red circles indicate misclassified observations.
+Plots are colored on predicted class, and red circles indicate misclassified instances.
 Model visual is essentially a representation of the confusion matrix of the model.
 The radial tour is a 1D projection starting the attribution projection.
-The first frame is the linear variable importances that best describe the observation's prediction.
-We probe the support of variable contributions by selecting a variable and use the radial tour to vary its coefficient.}\label{fig:classificationcase}
+The first frame is the linear feature importances that best describe the instance's prediction.
+We probe the support of feature contributions by selecting a feature and use the radial tour to vary its coefficient.}\label{fig:classificationcase}
 \end{figure}
@@
 
-A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+A tooltip displays instance number/name and classification information while the cursor hovers over a point.
 Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
 The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and, finally, identification of a primary and comparison observation.
+These interactions aid exploration of the spaces and, finally, identification of a primary and comparison instance.
 
@@
 \caption{Illustration of data explorations interactions in the global view.
-This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an observation.}\label{fig:interactions}
+This view has linked brushing, where instances selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
+Tooltips display when hovering over an instance.}\label{fig:interactions}
 \end{figure}
@@
 Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)], compatibility tree SHAP.
-We use modest hyperparameters, namely: 125 trees, number variables randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} Kominsarczyk et al.
+We use modest hyperparameters, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} Kominsarczyk et al.
 (2021).
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}
+
+}
+
+\caption{An orthogonal view of penguins data showing the PI, a Gentoo (purple) close to the distribution of Chinstrap (orange).}\label{fig:casepenguinsblfl}
+\end{figure}
+
 The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
 The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
-The subset used here has missing values removed (as used in the \texttt{spinifex} package), and there are 333 observations and four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
-A random forest model was fitted with species as the response variable.
+After removing incomplete observations, there are 333 instances and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
+A random forest model was fitted with species as the response feature.
 
@@
 \caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The observation of interest is an Adelie (green) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
-XXX needs fixing Starting from the attribution projection the contribution of \texttt{b\_l} is varied.
+The PI is an Chinstrap (orange) penguin that is misclassified as a Gentoo (purple), marked as an asterisk in (a), and the dashed vertical line in (b).
+Starting from the attribution projection the contribution of \texttt{b\_l} is varied.
 This explanation is not supported sense by the time there is moderate contribution.
-The current basis is fairly close to the regularly attribution to other chinstrap penguins aside from the sign on flipper length being flipped.
-It feels odd that SHAP has identified a profile where chinstrap and adelie penguins are hard to distinguish when there are separating orientations that should be found within 125 trees.
+The basis showing the correct species separation is fairly close to the regularly attribution to other Gentoo penguins aside from the sign change on flipper length.
 The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}}\label{fig:casepenguins}
@@
 Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
-Penguin 124, an Adelie (green), is the observation of interest because it has been misclassified as a Chinstrap (orange).
+Penguin 243, an Chinstrap (orange), is the PI because it has been misclassified as a Gentoo (purple).
 
 There is more separation visible in the attribution space than the data space, as would be expected.
-The observed vs predicted plot reveals a handful of misclassified observations.
+The observed vs predicted plot reveals a handful of misclassified instances.
 A misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view) is contrasted with a correctly classified point (represented by an \(\times\) and vertical dotted line).
 
-The radial tour starts from the attribution projection of the misclassified observation (left plot).
-The important variables identifed by SHAP in the (wrong) prediction for this observation is mostly \texttt{b\_d} with smaller contributions of \texttt{f\_l} and \texttt{wgt}.
-This projection is a view where the Gentoo (purple) penguins are most distinguished from Adelie and Chinstrap.
-For this observation we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of variables is not particularly useful.
-To explore this we use the radial tour to vary the contribution of bill length (\texttt{b\_l}).
-The more \texttt{b\_l} contributes to the the projection the more this penguin looks like a Chinstrap, that it is mislabelled as.
-This suggests that \texttt{b\_l} should have been an important variable for explaining the (wrong) prediction.
+The radial tour starts from the attribution projection of the misclassified instance (left plot).
+The important features identifed by SHAP in the (wrong) prediction for this instance is mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
+This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
+For this instance we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of features is not particularly useful.
+To explore this we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
+The more that \texttt{f\_l} contributes to the the projection the more this penguin looks like a Gentoo, its observed species.
+This suggests that \texttt{f\_l} should have been an important feature for explaining the (wrong) prediction.
 
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
+We remove four features with the lowest contribution for the selected instance and vary the contribution of \texttt{sugar}.
 The misclassification of two instances seems improbable when sugar is near max contribution.
@@
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of \texttt{fiber}.
+We look at one extreme instance and vary the contribution of \texttt{sugar}.
 We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
 While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an observation from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.
+In the penguin case we explored how the model misclassified an instance from a type one error perspective.
+In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original instance.
 
@@
 After aggregation of the skill measurements, we regress player wages {[}2020 euros{]} given just the skill aggregates.
-The model was fit from 5000 observations of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
+The model was fit from 5000 instances of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
 We compare a leading offensive fielder (L.
 Messi) with that of a top defensive fielder (V.
-van Dijk), the same observations were used in figure \ref{fig:shapdistrbd}.
+van Dijk), the same instances were used in figure \ref{fig:shapdistrbd}.
 
@@
 \caption{FIFA 2020 data, a random forest model regresses wages {[}2020 Euros{]} from nine aggregated of skill measurements.
-The primary observation is a star offensive player (L.
+The PI is a star offensive player (L.
 Messi) compared with a top defensive player (V.
 van Dijk).
-We remove three variables with low attribution from both players.
+We remove three features with low attribution from both players.
 The attribution projection starts with selected instance on the right.
@@
 Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across nine variables.
-Using interaction from the global view, we select a house with an extreme negative residual and an accurate observation close to it in the data.
+The remaining are 338 house sales across nine features.
+Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
 
@@
 \caption{Ames housing 2018 regressing sales price {[}2018 USD{]}.
-The primary observation is a house that a large \texttt{lot\ area} (\texttt{LtA}), but small \texttt{living\ area} (\texttt{LvA}).
-The comparison instance is in the opposite; a large house on a relatively small lot.
-Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices.
-Varying the contribution \texttt{living\ area} the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in different ways.
+The PI sale price was under predicted, and had sizable attribution to lot area (\texttt{LtA}).
+The CI was predicted sales price was similar and much more accurate with its observed sales price while it has very little attribution to lot area.
+Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of \texttt{LtA}, which is important to explaining the PI and near invariant to the sales price of the CI.
 The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}}\label{fig:caseames}
@@
 
-Figure \ref{fig:caseames} selects a sale of with relatively large lot size and small house compared with the opposite.
-Contribution of \texttt{living\ area} is crucial to the predictions of each observation though the sign varies.
-The attribution of \texttt{lot\ area} is not particularity essential to either observation in isolation.
-Perhaps this is because this variable interacts with exogenous geometry and layout information, or with endogenous information maybe the year build and overall quality holds redundant information.
-As the contribution of \texttt{living\ area} is changed it crosses a basis where, if the explanation were this attribution, the predicted house sale would be equal.
-
-XXX Observation 170, big residual; maybe look at obs 33
-NS Looking at old version? current Ames housing obs:
-prim\_obs \textless- 311 \#\# Large lot area to living area ratio \#\# small house, big lot
-comp\_obs \textless- 220 \#\# Small on that ratio \#\# large house, small lot
+Figure \ref{fig:caseames} selects the house sale 74, a sizabile under prediction that has a large contribution to lot area.
+The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
+As the contribution of lot area decreases the predictive power decreases for the PI, while the CI remains stationary.
 

### 2022-01-12 figures/_make_figures_case_study.r

@@ -27,11 +27,33 @@
 }
 
 ## Penguins classification ------
+### Save orthogonal display of penguins with
+(.g <- ggplot(penguins_na.rm, aes(x=bill_length_mm,
+                                  y=flipper_length_mm,
+                                  colour=species)) +
+   geom_point(size=1) +
+   ## Prim obs
+   geom_point(data=penguins_na.rm[243,], shape=8, size=5, alpha=0.8) +
+   ## Comparison obs
+   #geom_point(data=penguins_na.rm[169,], shape=4, size=3, alpha=0.6) +
+   theme_bw() +
+   theme(aspect.ratio = 1) +
+   #legend.margin   = margin(0, 0, 0, 0),
+   #legend.position = "bottom") +
+   scale_color_brewer(palette = "Dark2") +
+   labs(y = "Flipper length [mm]", x = "Bill length [mm]", color = "Observed species"))
+## Save
+ggplot2::ggsave(
+  "./figures/case_penguins_BlFl.png",
+  plot = .g, device = "png",
+  height = 4, units = "in")
+.m <- gc()
+
 {
   ## Data setup, spinifex::penguins
   names(penguins_ls)
-  prim_obs <- 124L
-  comp_obs <- 76L
+  prim_obs <- 243
+  comp_obs <- 169
   
   ### Global view and cheem tour stills
   .glob_view <- global_view(
@@ -43,8 +65,7 @@
       legend.position  = "bottom",
       legend.direction = "horizontal")
   .bas <- basis_attr_df(penguins_ls$attr_df, prim_obs)
-  .mv  <- 1 ## seems better than the 3 returned from
-  ##manip_var_of_attr_df(penguins_ls$attr_df, prim_obs, comp_obs)
+  .mv  <- which(colnames(penguins_ls$attr_df) == "f_l")
   ## Cheem tour for stills
   mt_interp <- manual_tour(basis = .bas, manip_var = .mv) %>% 
     spinifex:::interpolate_manual_tour(angle = .15) ## app is .15
@@ -57,7 +78,7 @@
     theme(legend.position = "off", aspect.ratio = 1.4) +
     ggtitle("Radial tour, select frames")
   .ggt2 <- radial_cheem_tour(
-    penguins_ls, basis = mt_interp[,,18, drop=FALSE], manip_var = .mv,
[…]

### 2022-01-12 _paper.rmd

@@
 
-The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
-Dark chocolates are typically considered healthier and often can be marketed at a higher price.
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
-A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the instance of interest (chocolate 22).
-It is compared with a correctly classified dark chocolate.
+The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+Dark chocolate is considered healthier than milk.
+The data was collected by students during the Iowa State University  class STAT503 from nutritional information from the manufacturer's web site, and normalized to 100g equivalents.
+The data is available in the `cheem` package.
+A random forest model is used for the classification of chocolate type.
+
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
+A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered as a healthy alternative.
+It is interesting to explore which of the nutritional variables contribute most to the misclassification.
 
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove four features with the lowest contribution for the selected instance and vary the contribution of `sugar`.
+We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `fiber`.
 The misclassification of two instances seems improbable when sugar is near max contribution.
@@
 
-Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space.
+This type of exploration is shown in \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
+It is compared with a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are clearer in the observed vs predicted plot, and can be seens to be mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+The opposite would be the interpretation of a point with high values in this plot.
+The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fibre and low sugar.
+The density for milk chocolates is further to the right, indicating that they generally have low fibre and high sugar.
+
+The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot which is the same projection of all observations as specifed by the SHAP values of the observation of interest.
+
+From the parallel coordinate plot we can quickly compare the variable contributions to the two different predictions.
+The observation of interest differs form the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding to the observation of interest, both observations are more like dark chocolates.
+If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
+When primarily Fiber is examined observation 22 looks more like a milk chocolate.
+
+<!-- Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space.
 Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of `sugar`.
+We look at one extreme instance and vary the contribution of `fiber`.
 We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
 While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an instance from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original instance.
+In the penguin case we explored how the model misclassified an observation from a type one error perspective.
+In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.-->
 
@@
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
+
+XXX You need to explain the variables.
 

### 2022-01-13 _paper.rmd
re-write of penguins

@@
 
-(ref:casepenguinsblfl) An orthogonal view of penguins data showing the PI, a Gentoo (purple) close to the distribution of Chinstrap (orange).
-```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl)"}
-knitr::include_graphics("./figures/case_penguins_BlFl.png")
-```
-
 The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
@@
 The PI is an Chinstrap (orange) penguin that is misclassified as a Gentoo (purple), marked as an asterisk in (a), and the dashed vertical line in (b).
-Starting from the attribution projection the contribution of `b_l` is varied.
-This explanation is not supported sense by the time there is moderate contribution.
-The basis showing the correct species separation is fairly close to the regularly attribution to other Gentoo penguins aside from the sign change on flipper length.
-The .mp4 animation of this tour can be found at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
+The radial view shows varying the contribution of `f_l` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
+(The .mp4 animation of the radial tour is at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4).)
 ```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
@@
 Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
-Penguin 243, an Chinstrap (orange), is the PI because it has been misclassified as a Gentoo (purple).
+Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a  Chinstrap (orange).
+
+(ref:casepenguinsblfl) Checking what is learned from the cheem viewer.
+This is a plot of flipper length (`f_l`) and bill length (`b_l`), where the PI is highlighted by an asterisk.
+It is a Gentoo (purple) misclassified as a Chinstrap (orange).
+The PI has an unusually small `f_l` length which is the primary reason it is confused with a Chinstrap.
+```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl)"}
+knitr::include_graphics("./figures/case_penguins_BlFl.png")
+```
 
@@
 The observed vs predicted plot reveals a handful of misclassified instances.
-A misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view) is contrasted with a correctly classified point (represented by an $\times$ and vertical dotted line).
+The PI is a misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view).
+The CI is a correctly classified point (represented by an $\times$ and a vertical dotted line).
 
@@
 This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
-For this instance we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of features is not particularly useful.
+For this instance we would like to know why a Gentoo has been wrongly labelled a Chinstrap.
+This combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
 To explore this we use the radial tour to vary the contribution of flipper length (`f_l`).
-The more that `f_l` contributes to the the projection the more this penguin looks like a Gentoo, its observed species.
+The more that `f_l` contributes to the the projection the more this penguin looks like a Chinstrap.
 This suggests that `f_l` should have been an important feature for explaining the (wrong) prediction.

### 2022-01-13 _paper.rmd
done with penguin case study

@@
 The PI has an unusually small `f_l` length which is the primary reason it is confused with a Chinstrap.
-```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl)"}
+```{r casepenguinsblfl, echo=F, out.width="60%", fig.cap = "(ref:casepenguinsblfl)"}
 knitr::include_graphics("./figures/case_penguins_BlFl.png")
@@
 The observed vs predicted plot reveals a handful of misclassified instances.
+For this illustration we are going to explore why a Gentoo has been wrongly labelled as a Chinstrap.
 The PI is a misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view).
@@
 
-The radial tour starts from the attribution projection of the misclassified instance (left plot).
+The radial tour starts from the attribution projection of the misclassified instance (b, left).
 The important features identifed by SHAP in the (wrong) prediction for this instance is mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
-For this instance we would like to know why a Gentoo has been wrongly labelled a Chinstrap.
-This combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
+This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap, that is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
 To explore this we use the radial tour to vary the contribution of flipper length (`f_l`).
-The more that `f_l` contributes to the the projection the more this penguin looks like a Chinstrap.
-This suggests that `f_l` should have been an important feature for explaining the (wrong) prediction.
+(In our exploreation, this was the third variable explored.
+It is typically useful to explore the variables with larger contributions, here `b_l` and `b_d`, but when doing this nothing was revealed about how the PI differed from other Gentoos.) On varying `f_l` as it contributes more to the projection (b, right) we see that more, and more, this penguin looks like a Chinstrap.
+This suggests that `f_l` should be considered to be an important feature for explaining the (wrong) prediction.
+
+Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`f_l`) is important for the confusion of the PI as a Chinstrap.
+Here, flipper length and body length are plotted and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
 

### 2022-01-13 _paper.rmd
- Updates to Di's slack points 1-4, + (6): - Glob view, sq asp ratio - Glob view, reduce jitter - regression facet issue; residual removed.

@@
 
-# Tours and the radial tour
+# Tours and the radial tour {#sec:tour}
 
@@
 Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across nine features.
+The remaining are 338 house sales.
+A random forest is regressed this price with the features `LotArea` (`LtA`), `OverallQual` (`Qlt`), `YearBuild` (`YrB`), `LivingArea` (`LvA`), `Bathrms` (`Bth`), `Bedrms` (`Bdr`), `TotRms` (`Rms`), `GarageYrBlt` (`GYB`), and `GarageArea` (`GrA`).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
 
-XXX You need to explain the variables.
-
-(ref:caseames) Ames housing 2018 regressing sales price [2018 USD].
+(ref:caseames) Ames housing 2018 regressing sales price [USD].
 The PI sale price was under predicted, and had sizable attribution to lot area (`LtA`).

### 2022-01-13 _paper.tex
- Updates to Di's slack points 1-4, + (6): - Glob view, sq asp ratio - Glob view, reduce jitter - regression facet issue; residual removed.

@@
 
-\hypertarget{tours-and-the-radial-tour}{%
-\section{Tours and the radial tour}\label{tours-and-the-radial-tour}}
+\hypertarget{sec:tour}{%
+\section{Tours and the radial tour}\label{sec:tour}}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}
-
-}
-
-\caption{An orthogonal view of penguins data showing the PI, a Gentoo (purple) close to the distribution of Chinstrap (orange).}\label{fig:casepenguinsblfl}
-\end{figure}
-
 The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
@@
 The PI is an Chinstrap (orange) penguin that is misclassified as a Gentoo (purple), marked as an asterisk in (a), and the dashed vertical line in (b).
-Starting from the attribution projection the contribution of \texttt{b\_l} is varied.
-This explanation is not supported sense by the time there is moderate contribution.
-The basis showing the correct species separation is fairly close to the regularly attribution to other Gentoo penguins aside from the sign change on flipper length.
-The .mp4 animation of this tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}}\label{fig:casepenguins}
+The radial view shows varying the contribution of \texttt{f\_l} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
+(The .mp4 animation of the radial tour is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}.)}\label{fig:casepenguins}
 \end{figure}
@@
 Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
-Penguin 243, an Chinstrap (orange), is the PI because it has been misclassified as a Gentoo (purple).
+Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.6\linewidth]{./figures/case_penguins_BlFl}
+
+}
+
+\caption{Checking what is learned from the cheem viewer.
+This is a plot of flipper length (\texttt{f\_l}) and bill length (\texttt{b\_l}), where the PI is highlighted by an asterisk.
+It is a Gentoo (purple) misclassified as a Chinstrap (orange).
+The PI has an unusually small \texttt{f\_l} length which is the primary reason it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
+\end{figure}
 
@@
 The observed vs predicted plot reveals a handful of misclassified instances.
-A misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view) is contrasted with a correctly classified point (represented by an \(\times\) and vertical dotted line).
+For this illustration we are going to explore why a Gentoo has been wrongly labelled as a Chinstrap.
+The PI is a misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view).
+The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
 
-The radial tour starts from the attribution projection of the misclassified instance (left plot).
+The radial tour starts from the attribution projection of the misclassified instance (b, left).
 The important features identifed by SHAP in the (wrong) prediction for this instance is mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
-For this instance we would like to know why an Adelie has been wrongly labelled a Chinstrap, so this combination of features is not particularly useful.
+This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap, that is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
 To explore this we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
-The more that \texttt{f\_l} contributes to the the projection the more this penguin looks like a Gentoo, its observed species.
-This suggests that \texttt{f\_l} should have been an important feature for explaining the (wrong) prediction.
+(In our exploreation, this was the third variable explored.
+It is typically useful to explore the variables with larger contributions, here \texttt{b\_l} and \texttt{b\_d}, but when doing this nothing was revealed about how the PI differed from other Gentoos.) On varying \texttt{f\_l} as it contributes more to the projection (b, right) we see that more, and more, this penguin looks like a Chinstrap.
+This suggests that \texttt{f\_l} should be considered to be an important feature for explaining the (wrong) prediction.
+
+Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{f\_l}) is important for the confusion of the PI as a Chinstrap.
+Here, flipper length and body length are plotted and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
 
@@
 
-The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
-Dark chocolates are typically considered healthier and often can be marketed at a higher price.
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest.
-A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the instance of interest (chocolate 22).
-It is compared with a correctly classified dark chocolate.
+The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+Dark chocolate is considered healthier than milk.
+The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's web site, and normalized to 100g equivalents.
+The data is available in the \texttt{cheem} package.
+A random forest model is used for the classification of chocolate type.
+
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
+A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered as a healthy alternative.
+It is interesting to explore which of the nutritional variables contribute most to the misclassification.
 
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove four features with the lowest contribution for the selected instance and vary the contribution of \texttt{sugar}.
+We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
 The misclassification of two instances seems improbable when sugar is near max contribution.
@@
 
-Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space.
-Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of \texttt{sugar}.
-We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
-While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an instance from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original instance.
+This type of exploration is shown in \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
+It is compared with a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are clearer in the observed vs predicted plot, and can be seens to be mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+The opposite would be the interpretation of a point with high values in this plot.
+The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fibre and low sugar.
+The density for milk chocolates is further to the right, indicating that they generally have low fibre and high sugar.
+
+The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot which is the same projection of all observations as specifed by the SHAP values of the observation of interest.
+
+From the parallel coordinate plot we can quickly compare the variable contributions to the two different predictions.
+The observation of interest differs form the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding to the observation of interest, both observations are more like dark chocolates.
+If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
+When primarily Fiber is examined observation 22 looks more like a milk chocolate.
 
@@
 Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales across nine features.
+The remaining are 338 house sales.
+A random forest is regressed this price with the features \texttt{LotArea} (\texttt{LtA}), \texttt{OverallQual} (\texttt{Qlt}), \texttt{YearBuild} (\texttt{YrB}), \texttt{LivingArea} (\texttt{LvA}), \texttt{Bathrms} (\texttt{Bth}), \texttt{Bedrms} (\texttt{Bdr}), \texttt{TotRms} (\texttt{Rms}), \texttt{GarageYrBlt} (\texttt{GYB}), and \texttt{GarageArea} (\texttt{GrA}).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 
-\caption{Ames housing 2018 regressing sales price {[}2018 USD{]}.
+\caption{Ames housing 2018 regressing sales price {[}USD{]}.
 The PI sale price was under predicted, and had sizable attribution to lot area (\texttt{LtA}).

### 2022-01-13 figures/_make_figures_case_study.r

@@ -47,7 +47,6 @@ ggplot2::ggsave(
   "./figures/case_penguins_BlFl.png",
   plot = .g, device = "png",
   height = 4, units = "in")
-.m <- gc()
 
 {
   ## Data setup, spinifex::penguins
@@ -119,7 +118,7 @@ ggplot2::ggsave(
 {
   names(chocolates_ls)
   prim_obs <- 22L #"Dark Chocolate Bar, Lindt, Switzerland"
-  comp_obs <- 7L #"85% Cocoa Dark French Chocolate, Thorntos, UK (2nd) 
+  comp_obs <- 7L  #"70% Cocoa, Columbia
   if(F)
     global_view(chocolates_ls,
                 prim_obs,
@@ -149,7 +148,7 @@ ggplot2::ggsave(
     pcp_shape = 124, angle = 0) + theme(legend.position = "off", aspect.ratio = 1.4) +
     ggtitle("Radial tour, select frames")
   .ggt2 <- radial_cheem_tour(
-    chocolates_ls, basis = mt_interp[,,18], manip_var = .mv,
+    chocolates_ls, basis = mt_interp[,,20], manip_var = .mv,
     primary_obs = prim_obs, comparison_obs = comp_obs,
     do_add_pcp_segments = FALSE, inc_var_nms = .inc_var_nms,
     pcp_shape = 124, angle = 0) + theme(legend.position = "off", aspect.ratio = 1.4)
@@ -172,13 +171,82 @@ ggplot2::ggsave(
   chocolates_ls, basis = .bas, manip_var = .mv,
   primary_obs = prim_obs, comparison_obs = comp_obs,
   do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
-  pcp_shape = 124, angle = .15) + ## realistic angle
+  pcp_shape = 124, angle = .15) + ## app angle
   theme(legend.position = "top", legend.direction = "horizontal")
 .anim <- animate_gganimate(
   .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom
   height = 800, width = 600, units = "px", ## "px", "in", "cm", or "mm."
   render = gganimate::av_renderer("./figures/case_chocolates.mp4")) ## Alternative render
 ## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4
+
+
+# ## Chocolates inverse case ---
+# {
+#   names(chocolates_ls)
[…]

### 2022-01-13 figures/_make_figures_case_study.r
Auto stash before merge of "main" and "origin/main"

@@ -180,7 +180,11 @@ ggplot2::ggsave(
 ## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4
 
 
+<<<<<<< Updated upstream
 # ## Chocolates inverse case ---
+=======
+# ## Chocolates inverse case -----
+>>>>>>> Stashed changes
 # {
 #   names(chocolates_ls)
 #   prim_obs <- 83L #"Silky smooth Milk chocolate - Extra Creamy, Dove, US"
@@ -237,7 +241,11 @@ ggplot2::ggsave(
 #   chocolates_ls, basis = .bas, manip_var = .mv,
 #   primary_obs = prim_obs, comparison_obs = comp_obs,
 #   do_add_pcp_segments = TRUE, inc_var_nms = .inc_var_nms,
+<<<<<<< Updated upstream
 #   pcp_shape = 124, angle = .15) + ## app angle
+=======
+#   pcp_shape = 124, angle = .15) + ## realistic angle
+>>>>>>> Stashed changes
 #   theme(legend.position = "top", legend.direction = "horizontal")
 # .anim <- animate_gganimate(
 #   .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom

### 2022-01-13 _paper.rmd
-Clarify PCP distribution - manual screen shots/annotation

@@
 The initial frame is the normalized SHAP values of the PI.
-The current projection basis is depicted as the width of a bar, the feature's contribution to the horizontal axis.
-The normalized values of all instances are shown as vertical parallel coordinate plots.
+The distribution normalized values from all instances are shown depicted in the basis as vertical parallel coordinate plots, where each line connects an instance's feature attribution.
+The PI is highlighted as larger dashed line and CI as  a dotted line.
+The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal axis.
 
@@
 We probe the support of feature contributions by selecting a feature and use the radial tour to vary its coefficient."}
-knitr::include_graphics("./figures/app_classification.PNG")
+knitr::include_graphics("./figures/app_classification_annotations.PNG")
 ```

### 2022-01-13 _paper.rmd
small changes i think

--- 2022-01-13 _paper.rmd
@@
 
-This type of exploration is shown in \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
-It is compared with a correctly classified dark chocolate (observation 7).
-The PCA plot, and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
-The misclassifications are clearer in the observed vs predicted plot, and can be seens to be mistaken in both ways: milk to dark and dark to milk.
+This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+It is compared with a CI that is a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
 
@@
 The opposite would be the interpretation of a point with high values in this plot.
-The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fibre and low sugar.
-The density for milk chocolates is further to the right, indicating that they generally have low fibre and high sugar.
+The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
+The density for milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
 

### 2022-01-13 _paper.tex

@@
 The initial frame is the normalized SHAP values of the PI.
-The current projection basis is depicted as the width of a bar, the feature's contribution to the horizontal axis.
-The normalized values of all instances are shown as vertical parallel coordinate plots.
+The distribution normalized values from all instances are shown depicted in the basis as vertical parallel coordinate plots, where each line connects an instance's feature attribution.
+The PI is highlighted as larger dashed line and CI as a dotted line.
+The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal axis.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}
+{\centering \includegraphics[width=1\linewidth]{./figures/app_classification_annotations}
 

### 2022-01-14 _paper.rmd
- New screen caps with lettering consolidating 3 into 2. - remove ")" from all panel labels - grammerly edits through-out.

@@
  The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear feature importance in the vicinity of each instance.
-These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between features used to make predictions, across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
+XAI attempts to shed light on how models use predictors to arrive at a prediction with a point estimate of the linear feature importance in the vicinity of each instance.
+These can be considered linear projections and can be further explored interactively to understand better the interaction between features used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
@@
 For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
-Feature-level (variable-level) interpretability of models is essential in the evaluation models for such biases.
-It is also generally important for many problems, where it is not enough to accurately predict accurately but one must be able to explain which predictors are most responsible in generating a response value.
+Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
+It is also generally important for many problems, where it is not enough to predict accurately, but one must be able to explain which predictors are most responsible for generating a response value.
 
 <!-- Interpretability & data drift -->
-Another concern is that of data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
-Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
-Better interpretability of the model means that there is more transparency where models' predictions may be plausible or completely unreliable.
+Another concern is data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
+Non-linear models are typically more sensitive and do not extrapolate well outside the training data domain.
+Better interpretability of the model means more transparency where models' predictions may be plausible or completely unreliable.
 
@@
 A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
-Because these are point specific, a challenge is visualizing them to more comprehensively understand a model.
+Because these are point-specific, it is challenging to visualize them to understand a model more comprehensively.
 
@@
 Structure in a projection can then be explored visually to see which features contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information of the structure and features of the object.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information about the structure and features of the object.
 
@@
 <!-- Purposed approach -->
-In our proposed approach we use the radial manual tour to scrutinize a local explanation.
+In our proposed approach, we use the radial manual tour to scrutinize a local explanation.
 Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
@@
 That means, when the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions, and is thus an interpretable visualisation.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
 
@@
 The remainder of this paper is organized as follows.
-The following Section, \@ref(sec:explanations) covers the background of the local explanation, and the traditional visuals produced.
+The following Section, \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced.
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
@@
 Then Section \@ref(#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
-We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explore in the future.
+We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explored in the future.
 
@@
 It includes a large number of model-specific explanations such as deepLIFT [@shrikumar_not_2016; @shrikumar_learning_2017], a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME [@ribeiro_why_2016] SHAP [@lundberg_unified_2017], and their variants are popular.
+There are fewer model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017], and their variants are popular.
 
@@
 These instance-level explanations are used in various ways depending on the data.
-In image classification, where pixels would correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels [@simonyan_deep_2014].
+In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels [@simonyan_deep_2014].
 For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
@@
 SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
-This explanations almost all point to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
-@strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
+These explanations almost all point to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by permuting the players that contributing to the score.
+@strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
 While the models themselves are invariant to the order of the predictors, the additive explanations are not.
-The SHAP values are the mean contributions over some subset these permutations.
+The SHAP values are the mean contributions over some subset of these permutations.
 This is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
@@
 <!-- tree SHAP -->
-For our application, we use _tree SHAP_, a variant of SHAP enjoys a lower computational complexity [@lundberg_consistent_2018].
-Instead of aggregating over permutations of the feature order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
+For our application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
+Instead of aggregating over permutations of the feature order, tree SHAP approximates instance-level feature importance by exploring the structure of the decision trees.
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as the starting projection basis (call this the _attribution projection_) to further scrutinize the explanation.
+The following section will use normalized explanations as to the starting projection basis (call this the _attribution projection_) to further scrutinize the explanation.
 
 ```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-A star offensive and densive player, Messi and a defensive player are compared, L.
+A star offensive and defensive player, Messi and a defensive player are compared, L.
 Messi and V van Dijk respectively.
-Plot a) is the break-down plot showing one additive feature-attributution.
+Panel (a) is the break-down plot showing one additive feature-attribution.
 These attributions are sensitive to the order of the features.
-Plot b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
-Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi."}
+Panel (b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
 knitr::include_graphics("./figures/shap_distr_bd.png")
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time, and as such gain information about the relationships between points across multiple features.
+One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
 There are various types of tours that are distinguished by how the paths are generated [@lee_state_2021; @cook_grand_2008].
@@
 <!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its path by changing a selected feature's contribution to a basis, to allow the feature to contribute more or less to the projection.
-The contribution of all other features is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour [@cook_manual_1997] defines its path by changing a selected feature's contribution to a basis to allow the feature to contribute more or less to the projection.
+The requirement constrains the contribution of all other features that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
 The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
-It also lends itself to pre-computation to be queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
+It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
 <!-- However, this navigation is relatively time-consuming due to the vast volume of the display space.
@@
 A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of the new work.
-In a radial tour, the selected feature is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and also interactive re-calculation to focus on a different feature.
+In a radial tour, the selected feature can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and interactive re-calculation to focus on a different feature.
 
@@
 There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
-In addition, there are numerous user inputs, including feature selection for the radial tour, and instance selection for making comparisons.
+In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
 Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
@@
 <!-- Approximations of the spaces, PC1:2 -->
-A visualisation of these spaces is provided by the first two principal components of their respective spaces.
+A visualization of these spaces is provided by the first two principal components of their respective spaces.
 In addition, a plot observed by predicted response is also provided.
-In both PCA plots the orientation and magnitude of the features are inscribed on a unit circle, similar to a biplot [@gabriel_biplot_1971].
 A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
 <!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
 
-While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances, from which to explore the local explanations.
-The projection attribution of the Primary Instance (PI) is explored and typically viewed with an optional Comparison Instance (CI).
+While the comparison of these spaces is interesting, the main purpose of the global view is to enable the selection of instances to explore the local explanations.
+The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
 
@@
 The distribution normalized values from all instances are shown depicted in the basis as vertical parallel coordinate plots, where each line connects an instance's feature attribution.
-The PI is highlighted as larger dashed line and CI as  a dotted line.
+The PI is highlighted as a dashed line and CI as  a dotted line.
 The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal axis.
@@
 Typically we select a misclassified instance compared to a correctly classified point nearby in data space.
-The model infomation in the global view is a depiction of the model conmfusion matrix.
+The model information in the global view depicts the model confusion matrix.
 The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each feature to structure identified by the local explanation, evaluating the support or robustness of the prediction.
@@
 ```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks.
-Plots are colored on predicted class, and red circles indicate misclassified instances.
-Model visual is essentially a representation of the confusion matrix of the model.
-The radial tour is a 1D projection starting the attribution projection.
-The first frame is the linear feature importances that best describe the instance's prediction.
-We probe the support of feature contributions by selecting a feature and use the radial tour to vary its coefficient."}
-knitr::include_graphics("./figures/app_classification_annotations.PNG")
+Global view inputs, (a), set the PI, CI, and color statistic.
+Global view, (b) PC1 by PC2 for data space, attribution space, and prediction by observed y (visual of the confusion matrix for classification).
+Points are colored by predicted class, and red circles indicate misclassified instances.
+Radial tour inputs (c) select features to include, which variable is changed in the tour, and the presence of the parrelle coordiate lines connecting attribution distribution.
+The visual depicting 1D basis (d) shows the distribution of the feature attribution, bars show the current projection with the black bar is being varied.
+Panel (e) is the resulting projection of the data shown as density in the classification case."}
+knitr::include_graphics("./figures/app_classification.PNG")
 ```
@@
 
-In the regression case, the global view can be colored on a statistic to highlight the explanation space's structure.
+Structure resolved in the attribution space can be highlighted by coloring points on a statistic.
 For this purpose, we include residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
@@
 
-```{r regressioncase, echo=F, out.width = "100%", fig.cap="Overview of the cheem viewer for regression tasks.
-The global view can be colored on the correlation between the projection of the data generated by the local explanation and the observed response.
-In the radial tour, the horizontal values are the same as the classification case; the projection through the basis.
-The vertical position is now mapped to the observed y and residuals."}
-knitr::include_graphics("./figures/app_regression.PNG")
+```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression task highlighting differences with classification and interaction.
+Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
+Coloring on a statistic (b) highlights structure organized in the attribution space.
+Interactive tabular display (d) populates when instances are selected.
+Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, left and right respectively."}
+knitr::include_graphics("./figures/app_regression_interactions.PNG")
 ```
@@
 
-```{r interactions, echo=F, out.width="100%", fig.cap = "Illustration of data explorations interactions in the global view.
-This view has linked brushing, where instances selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an instance."}
-knitr::include_graphics("./figures/app_interactions.PNG")
-```
-
 ## Preprocessing
@@
 \begin{itemize}
-	\item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale invariant.
+	\item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
 	\item \textbf{Model and explanation:} any model can be used with this method.
@@
 Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
-This implementation aggregates over exhaustively over the attribution of all trees and we opt to not to fit interactions of features.
-	\item \textbf{Cheem view:} after the model and full explanation space are calculated we scale each feature by standard deviations away from the mean to achieve a common support for visuals.
+This implementation aggregates over exhaustively over all trees' attribution and we opt not to fit interactions of features.
+	\item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
 Statistics for mapping to color are calculated on the scaled spaces.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-With its reduced computational complexity, this makes tree SHAP a good candidate to start with.
+Its reduced computational complexity makes tree SHAP a good candidate to start with.
 Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
@@
 The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on [CRAN](https://CRAN.R-project.org/package=cheem).
-Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002], and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
+Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002] and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
 The application was made with __shiny__ [@chang_shiny_2021].
@@
 
-The following __R__ code will help getting up and running:
+The following __R__ code will help with geting up and running:
 
@@
 install.packages("cheem", dependencies = TRUE)
-## Restart the R session so the IDE has the correct directory structure
+## Restart the R session, so the IDE has the correct directory structure
 restartSession()
@@
 # Processing your data
-## Install treeshap from github, to use as a local explainer
-remotes::install_github('ModelOriented/treeshap') ## Local 
+## Install treeshap from github to use as a local explainer
+remotes::install_github('ModelOriented/treeshap')
 ## Follow the examples in cheem_ls()
@@
 The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
+The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
 After removing incomplete observations, there are 333 instances and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
-A random forest model was fitted with species as the response feature.
+A random forest model was fit with species as the response feature.
 
@@
 Figure \@ref(fig:casepenguins) shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
+Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
 Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a  Chinstrap (orange).
@@
 This is a plot of flipper length (`f_l`) and bill length (`b_l`), where the PI is highlighted by an asterisk.
-It is a Gentoo (purple) misclassified as a Chinstrap (orange).
+A Gentoo (purple) misclassified as a Chinstrap (orange).
 The PI has an unusually small `f_l` length which is the primary reason it is confused with a Chinstrap.
@@
 There is more separation visible in the attribution space than the data space, as would be expected.
-The observed vs predicted plot reveals a handful of misclassified instances.
-For this illustration we are going to explore why a Gentoo has been wrongly labelled as a Chinstrap.
-The PI is a misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view).
+The predicted vs observed plot reveals a handful of misclassified instances.
+For this illustration, we will explore why a Gentoo has been wrongly labeled as a Chinstrap.
+The PI is a misclassified point (represented by the asterisk in the global view and as a dashed vertical line in the tour view).
 The CI is a correctly classified point (represented by an $\times$ and a vertical dotted line).
@@
 The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identifed by SHAP in the (wrong) prediction for this instance is mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
+The important features identified by SHAP in the (wrong) prediction for this instance is mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
 This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap, that is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this we use the radial tour to vary the contribution of flipper length (`f_l`).
-(In our exploreation, this was the third variable explored.
-It is typically useful to explore the variables with larger contributions, here `b_l` and `b_d`, but when doing this nothing was revealed about how the PI differed from other Gentoos.) On varying `f_l` as it contributes more to the projection (b, right) we see that more, and more, this penguin looks like a Chinstrap.
-This suggests that `f_l` should be considered to be an important feature for explaining the (wrong) prediction.
+To explore this, we use the radial tour to vary the contribution of flipper length (`f_l`).
+(In our exploration, this was the third variable explored.
+It is typically useful to explore the variables with larger contributions, here `b_l` and `b_d`, but when doing this, nothing was revealed about how the PI differed from other Gentoos.) On varying `f_l` as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
+This suggests that `f_l` should be considered an important feature for explaining the (wrong) prediction.
 
 Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`f_l`) is important for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
+Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
 
@@
 
-The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
-The data was collected by students during the Iowa State University  class STAT503 from nutritional information from the manufacturer's web site, and normalized to 100g equivalents.
+The data was collected by students during the Iowa State University  class STAT503 from nutritional information from the manufacturer's website, and normalized to 100g equivalents.
 The data is available in the `cheem` package.
@@
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered as a healthy alternative.
-It is interesting to explore which of the nutritional variables contribute most to the misclassification.
+A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered a healthy alternative.
+It is interesting to explore which of the nutritional variables contribute most to misclassification.
 
@@
 We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `fiber`.
-The misclassification of two instances seems improbable when sugar is near max contribution.
+The misclassification of two instances seems improbable when sugar is near the max contribution.
 Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
@@
 
-This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
-It is compared with a CI that is a correctly classified dark chocolate (observation 7).
-The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
+This type of exploration is shown in \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
+It is compared with a correctly classified dark chocolate (observation 7).
+The PCA plot and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
 The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
@@
 The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
 In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be the interpretation of a point with high values in this plot.
+The opposite would be interpreting a point with high values in this plot.
 The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
@@
 The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
-Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot which is the same projection of all observations as specifed by the SHAP values of the observation of interest.
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the observation of interest.
 
-From the parallel coordinate plot we can quickly compare the variable contributions to the two different predictions.
-The observation of interest differs form the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+We can quickly compare the variable contributions to the two different predictions from the parallel coordinate plot.
+The observation of interest differs with the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
 
@@
 If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
-When primarily Fiber is examined observation 22 looks more like a milk chocolate.
+When primarily Fiber is examined, observation 22 looks more like a milk chocolate.
 
@@
 
-(ref:casefifa) FIFA 2020 data, a random forest model regresses wages [2020 Euros] from nine aggregated of skill measurements.
+(ref:casefifa) FIFA 2020 data, a random forest model, regresses wages [2020 Euros] from nine aggregated skill measurements.
 The PI is a star offensive player (L.
@@
 We remove three features with low attribution from both players.
-The attribution projection starts with selected instance on the right.
+The attribution projection starts with the selected instance on the right.
 We vary the contribution from `defense`, the star offensive player is not distinguished in the horizontal direction.
-At this point defensive players have been rotated to the highest horizonatal value.
+At this point, defensive players have been rotated to the highest horizonatal value.
 The animate radial tour can be found at [shorturl.at/lCK26](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4)
@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-While `offense` and `reaction` skills both crucial to the explanation of a star offensive player.
-As the contribution of defensive skills increases Messi's is no longer separated from the group and other defensive players are better predicted in this attribution case.
-In terms of what-if analysis, were Messi's tree SHAP attributions at theses levels, his predicted wages would be halved.
+While `offense` and `reaction` skills are both crucial to explaining a star offensive player.
+As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
+In terms of what-if analysis, were Messi's tree SHAP attributions at these levels, his predicted wages would be halved.
 
@@
 The remaining are 338 house sales.
-A random forest is regressed this price with the features `LotArea` (`LtA`), `OverallQual` (`Qlt`), `YearBuild` (`YrB`), `LivingArea` (`LvA`), `Bathrms` (`Bth`), `Bedrms` (`Bdr`), `TotRms` (`Rms`), `GarageYrBlt` (`GYB`), and `GarageArea` (`GrA`).
+A random forest model has regressed this price with the features `LotArea` (`LtA`), `OverallQual` (`Qlt`), `YearBuild` (`YrB`), `LivingArea` (`LvA`), `Bathrms` (`Bth`), `Bedrms` (`Bdr`), `TotRms` (`Rms`), `GarageYrBlt` (`GYB`), and `GarageArea` (`GrA`).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 
-Figure \@ref(fig:caseames) selects the house sale 74, a sizabile under prediction that has a large contribution to lot area.
+Figure \@ref(fig:caseames) selects the house sale 74, a sizeable under prediction that has a large contribution to lot area.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-As the contribution of lot area decreases the predictive power decreases for the PI, while the CI remains stationary.
+As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
 

### 2022-01-14 _paper.tex
- New screen caps with lettering consolidating 3 into 2. - remove ")" from all panel labels - grammerly edits through-out.

@@
 The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear feature importance in the vicinity of each instance.
-These can be considered to be linear projections, and as such can be further explored interactively to better understand the interaction between features used to make predictions, across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
+XAI attempts to shed light on how models use predictors to arrive at a prediction with a point estimate of the linear feature importance in the vicinity of each instance.
+These can be considered linear projections and can be further explored interactively to understand better the interaction between features used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
 The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
@@
 For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
-Feature-level (variable-level) interpretability of models is essential in the evaluation models for such biases.
-It is also generally important for many problems, where it is not enough to accurately predict accurately but one must be able to explain which predictors are most responsible in generating a response value.
+Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
+It is also generally important for many problems, where it is not enough to predict accurately, but one must be able to explain which predictors are most responsible for generating a response value.
 
-Another concern is that of data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
-Non-linear models are typically more sensitive, and do not extrapolate well outside of the training data domain.
-Better interpretability of the model means that there is more transparency where models' predictions may be plausible or completely unreliable.
+Another concern is data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
+Non-linear models are typically more sensitive and do not extrapolate well outside the training data domain.
+Better interpretability of the model means more transparency where models' predictions may be plausible or completely unreliable.
 
@@
 A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
-Because these are point specific, a challenge is visualizing them to more comprehensively understand a model.
+Because these are point-specific, it is challenging to visualize them to understand a model more comprehensively.
 
@@
 
-In our proposed approach we use the radial manual tour to scrutinize a local explanation.
+In our proposed approach, we use the radial manual tour to scrutinize a local explanation.
 Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
@@
 That means, when the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions, and is thus an interpretable visualisation.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
 
 The remainder of this paper is organized as follows.
-The following Section, \ref{sec:explanations} covers the background of the local explanation, and the traditional visuals produced.
+The following Section, \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
 Section \ref{sec:tour} explains the animations of continuous linear projections.
@@
 Then Section @ref(\#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
-We conclude with Section \ref{sec:cheemdiscussion} of the insights gained and directions that might be explore in the future.
+We conclude with Section \ref{sec:cheemdiscussion} of the insights gained and directions that might be explored in the future.
 
@@
 2016; Shrikumar, Greenside, and Kundaje 2017), a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME (Ribeiro, Singh, and Guestrin 2016) SHAP (Lundberg and Lee 2017), and their variants are popular.
+There are fewer model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (Lundberg and Lee 2017), and their variants are popular.
 
 These instance-level explanations are used in various ways depending on the data.
-In image classification, where pixels would correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels (Simonyan, Vedaldi, and Zisserman 2014).
+In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels (Simonyan, Vedaldi, and Zisserman 2014).
 For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
@@
 SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
-This explanations almost all point to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by permuting the players that contribute to the score.
-Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
+These explanations almost all point to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by permuting the players that contributing to the score.
+Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
 While the models themselves are invariant to the order of the predictors, the additive explanations are not.
-The SHAP values are the mean contributions over some subset these permutations.
+The SHAP values are the mean contributions over some subset of these permutations.
 This is also related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
@@
 
-For our application, we use \emph{tree SHAP}, a variant of SHAP enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
-Instead of aggregating over permutations of the feature order, tree SHAP approximates instance level feature importance by exploring the stucture of the decision trees.
+For our application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
+Instead of aggregating over permutations of the feature order, tree SHAP approximates instance-level feature importance by exploring the structure of the decision trees.
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as the starting projection basis (call this the \emph{attribution projection}) to further scrutinize the explanation.
+The following section will use normalized explanations as to the starting projection basis (call this the \emph{attribution projection}) to further scrutinize the explanation.
 
@@
 \caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-A star offensive and densive player, Messi and a defensive player are compared, L.
+A star offensive and defensive player, Messi and a defensive player are compared, L.
 Messi and V van Dijk respectively.
-Plot a) is the break-down plot showing one additive feature-attributution.
+Panel (a) is the break-down plot showing one additive feature-attribution.
 These attributions are sensitive to the order of the features.
-Plot b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
-Offense and movement are important for Messi but not van Dijk, and conversely defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
+Panel (b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
 \end{figure}
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time, and as such gain information about the relationships between points across multiple features.
+One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
 There are various types of tours that are distinguished by how the paths are generated (S.
@@
 
-The manual tour (Cook and Buja 1997) defines its path by changing a selected feature's contribution to a basis, to allow the feature to contribute more or less to the projection.
-The contribution of all other features is constrained by the requirement that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour (Cook and Buja 1997) defines its path by changing a selected feature's contribution to a basis to allow the feature to contribute more or less to the projection.
+The requirement constrains the contribution of all other features that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
 The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
-It also lends itself to pre-computation to be queued in advance or computed on-the-fly for human-in-the-loop analysis (Karwowski 2006).
+It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis (Karwowski 2006).
 
 A version of the manual tour called a \emph{radial tour} is implemented in Spyrison and Cook (2020) and forms the basis of the new work.
-In a radial tour, the selected feature is allowed to change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and also interactive re-calculation to focus on a different feature.
+In a radial tour, the selected feature can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and interactive re-calculation to focus on a different feature.
 
@@
 There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
-In addition, there are numerous user inputs, including feature selection for the radial tour, and instance selection for making comparisons.
+In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
 Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
@@
 
-A visualisation of these spaces is provided by the first two principal components of their respective spaces.
+A visualization of these spaces is provided by the first two principal components of their respective spaces.
 In addition, a plot observed by predicted response is also provided.
-In both PCA plots the orientation and magnitude of the features are inscribed on a unit circle, similar to a biplot (Gabriel 1971).
 A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
 For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
 
-While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances, from which to explore the local explanations.
-The projection attribution of the Primary Instance (PI) is explored and typically viewed with an optional Comparison Instance (CI).
+While the comparison of these spaces is interesting, the main purpose of the global view is to enable the selection of instances to explore the local explanations.
+The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
 
@@
 The distribution normalized values from all instances are shown depicted in the basis as vertical parallel coordinate plots, where each line connects an instance's feature attribution.
-The PI is highlighted as larger dashed line and CI as a dotted line.
+The PI is highlighted as a dashed line and CI as a dotted line.
 The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal axis.
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/app_classification_annotations}
+{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}
 
@@
 \caption{Overview of the cheem viewer for classification tasks.
-Plots are colored on predicted class, and red circles indicate misclassified instances.
-Model visual is essentially a representation of the confusion matrix of the model.
-The radial tour is a 1D projection starting the attribution projection.
-The first frame is the linear feature importances that best describe the instance's prediction.
-We probe the support of feature contributions by selecting a feature and use the radial tour to vary its coefficient.}\label{fig:classificationcase}
+Global view inputs, (a), set the PI, CI, and color statistic.
+Global view, (b) PC1 by PC2 for data space, attribution space, and prediction by observed y (visual of the confusion matrix for classification).
+Points are colored by predicted class, and red circles indicate misclassified instances.
+Radial tour inputs (c) select features to include, which variable is changed in the tour, and the presence of the parrelle coordiate lines connecting attribution distribution.
+The visual depicting 1D basis (d) shows the distribution of the feature attribution, bars show the current projection with the black bar is being varied.
+Panel (e) is the resulting projection of the data shown as density in the classification case.}\label{fig:classificationcase}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/app_regression}
+{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions}
 
@@
 
-\caption{Overview of the cheem viewer for regression tasks.
-The global view can be colored on the correlation between the projection of the data generated by the local explanation and the observed response.
-In the radial tour, the horizontal values are the same as the classification case; the projection through the basis.
-The vertical position is now mapped to the observed y and residuals.}\label{fig:regressioncase}
+\caption{Overview of the cheem viewer for regression task highlighting differences with classification and interaction.
+Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
+Coloring on a statistic (b) highlights structure organized in the attribution space.
+Interactive tabular display (d) populates when instances are selected.
+Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, left and right respectively.}\label{fig:regressioncase}
 \end{figure}
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/app_interactions}
-
-}
-
-\caption{Illustration of data explorations interactions in the global view.
-This view has linked brushing, where instances selected in one facet are highlighted in the other facets and populate an interactive tabular display below.
-Tooltips display when hovering over an instance.}\label{fig:interactions}
-\end{figure}
-
 \hypertarget{preprocessing}{%
@@
 \begin{itemize}
-    \item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale invariant.
+    \item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
     \item \textbf{Model and explanation:} any model can be used with this method.
@@
 (2021).
-This implementation aggregates over exhaustively over the attribution of all trees and we opt to not to fit interactions of features.
-    \item \textbf{Cheem view:} after the model and full explanation space are calculated we scale each feature by standard deviations away from the mean to achieve a common support for visuals.
+This implementation aggregates over exhaustively over all trees' attribution and we opt not to fit interactions of features.
+    \item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
 Statistics for mapping to color are calculated on the scaled spaces.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-With its reduced computational complexity, this makes tree SHAP a good candidate to start with.
+Its reduced computational complexity makes tree SHAP a good candidate to start with.
 Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
@@
 The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002), and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
+Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002) and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
 2021).
@@
 
-The following \textbf{R} code will help getting up and running:
+The following \textbf{R} code will help with geting up and running:
 
@@
 \FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
-\DocumentationTok{\#\# Restart the R session so the IDE has the correct directory structure}
+\DocumentationTok{\#\# Restart the R session, so the IDE has the correct directory structure}
 \FunctionTok{restartSession}\NormalTok{()}
@@
 \CommentTok{\# Processing your data}
-\DocumentationTok{\#\# Install treeshap from github, to use as a local explainer}
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{) }\DocumentationTok{\#\# Local }
+\DocumentationTok{\#\# Install treeshap from github to use as a local explainer}
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
 \DocumentationTok{\#\# Follow the examples in cheem\_ls()}
@@
 The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was publicly available to be a substitute for the overly-used iris data, and is quite similar in form.
+The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
 After removing incomplete observations, there are 333 instances and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
-A random forest model was fitted with species as the response feature.
+A random forest model was fit with species as the response feature.
 
@@
 Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Plot (a) shows the global view, and plot (b) shows several 1D projections generated with the radial tour.
+Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
 Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
@@
 This is a plot of flipper length (\texttt{f\_l}) and bill length (\texttt{b\_l}), where the PI is highlighted by an asterisk.
-It is a Gentoo (purple) misclassified as a Chinstrap (orange).
+A Gentoo (purple) misclassified as a Chinstrap (orange).
 The PI has an unusually small \texttt{f\_l} length which is the primary reason it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
@@
 There is more separation visible in the attribution space than the data space, as would be expected.
-The observed vs predicted plot reveals a handful of misclassified instances.
-For this illustration we are going to explore why a Gentoo has been wrongly labelled as a Chinstrap.
-The PI is a misclassified point (represented by the asterisk in the global view, and as a dashed vertical line in the tour view).
+The predicted vs observed plot reveals a handful of misclassified instances.
+For this illustration, we will explore why a Gentoo has been wrongly labeled as a Chinstrap.
+The PI is a misclassified point (represented by the asterisk in the global view and as a dashed vertical line in the tour view).
 The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
@@
 The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identifed by SHAP in the (wrong) prediction for this instance is mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
+The important features identified by SHAP in the (wrong) prediction for this instance is mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
 This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap, that is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
-(In our exploreation, this was the third variable explored.
-It is typically useful to explore the variables with larger contributions, here \texttt{b\_l} and \texttt{b\_d}, but when doing this nothing was revealed about how the PI differed from other Gentoos.) On varying \texttt{f\_l} as it contributes more to the projection (b, right) we see that more, and more, this penguin looks like a Chinstrap.
-This suggests that \texttt{f\_l} should be considered to be an important feature for explaining the (wrong) prediction.
+To explore this, we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
+(In our exploration, this was the third variable explored.
+It is typically useful to explore the variables with larger contributions, here \texttt{b\_l} and \texttt{b\_d}, but when doing this, nothing was revealed about how the PI differed from other Gentoos.) On varying \texttt{f\_l} as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
+This suggests that \texttt{f\_l} should be considered an important feature for explaining the (wrong) prediction.
 
 Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{f\_l}) is important for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
+Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
 
@@
 
-The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark.
+The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
-The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's web site, and normalized to 100g equivalents.
+The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website, and normalized to 100g equivalents.
 The data is available in the \texttt{cheem} package.
@@
 We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
-The misclassification of two instances seems improbable when sugar is near max contribution.
+The misclassification of two instances seems improbable when sugar is near the max contribution.
 Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolates}
@@
 It is compared with a correctly classified dark chocolate (observation 7).
-The PCA plot, and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
-The misclassifications are clearer in the observed vs predicted plot, and can be seens to be mistaken in both ways: milk to dark and dark to milk.
+The PCA plot and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
 
 The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
 In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be the interpretation of a point with high values in this plot.
-The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fibre and low sugar.
-The density for milk chocolates is further to the right, indicating that they generally have low fibre and high sugar.
+The opposite would be interpreting a point with high values in this plot.
+The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
+The density for milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
 
 The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
-Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot which is the same projection of all observations as specifed by the SHAP values of the observation of interest.
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the observation of interest.
 
-From the parallel coordinate plot we can quickly compare the variable contributions to the two different predictions.
-The observation of interest differs form the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+We can quickly compare the variable contributions to the two different predictions from the parallel coordinate plot.
+The observation of interest differs with the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
 
@@
 
-\caption{FIFA 2020 data, a random forest model regresses wages {[}2020 Euros{]} from nine aggregated of skill measurements.
+\caption{FIFA 2020 data, a random forest model, regresses wages {[}2020 Euros{]} from nine aggregated skill measurements.
 The PI is a star offensive player (L.
@@
 We remove three features with low attribution from both players.
-The attribution projection starts with selected instance on the right.
+The attribution projection starts with the selected instance on the right.
 We vary the contribution from \texttt{defense}, the star offensive player is not distinguished in the horizontal direction.
-At this point defensive players have been rotated to the highest horizonatal value.
+At this point, defensive players have been rotated to the highest horizonatal value.
 The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{shorturl.at/lCK26}}\label{fig:casefifa}
@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-While \texttt{offense} and \texttt{reaction} skills both crucial to the explanation of a star offensive player.
-As the contribution of defensive skills increases Messi's is no longer separated from the group and other defensive players are better predicted in this attribution case.
-In terms of what-if analysis, were Messi's tree SHAP attributions at theses levels, his predicted wages would be halved.
+While \texttt{offense} and \texttt{reaction} skills are both crucial to explaining a star offensive player.
+As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
+In terms of what-if analysis, were Messi's tree SHAP attributions at these levels, his predicted wages would be halved.
 
@@
 The remaining are 338 house sales.
-A random forest is regressed this price with the features \texttt{LotArea} (\texttt{LtA}), \texttt{OverallQual} (\texttt{Qlt}), \texttt{YearBuild} (\texttt{YrB}), \texttt{LivingArea} (\texttt{LvA}), \texttt{Bathrms} (\texttt{Bth}), \texttt{Bedrms} (\texttt{Bdr}), \texttt{TotRms} (\texttt{Rms}), \texttt{GarageYrBlt} (\texttt{GYB}), and \texttt{GarageArea} (\texttt{GrA}).
+A random forest model has regressed this price with the features \texttt{LotArea} (\texttt{LtA}), \texttt{OverallQual} (\texttt{Qlt}), \texttt{YearBuild} (\texttt{YrB}), \texttt{LivingArea} (\texttt{LvA}), \texttt{Bathrms} (\texttt{Bth}), \texttt{Bedrms} (\texttt{Bdr}), \texttt{TotRms} (\texttt{Rms}), \texttt{GarageYrBlt} (\texttt{GYB}), and \texttt{GarageArea} (\texttt{GrA}).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizabile under prediction that has a large contribution to lot area.
+Figure \ref{fig:caseames} selects the house sale 74, a sizeable under prediction that has a large contribution to lot area.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-As the contribution of lot area decreases the predictive power decreases for the PI, while the CI remains stationary.
+As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
 
@@
 \url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
-
-\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
-Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
 

### 2022-01-14 _paper.rmd

@@
 
-This type of exploration is shown in \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
-It is compared with a correctly classified dark chocolate (observation 7).
-The PCA plot and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+It is compared with a CI that is a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
 The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
@@
 The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
+The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
 In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be interpreting a point with high values in this plot.
+The opposite would be the interpretation of a point with high values in this plot.
 The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.

### 2022-01-14 _paper.tex

@@
 \PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
@@
 \usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   colorlinks=true,
-  linkcolor=red,
-  filecolor=Maroon,
-  citecolor=Blue,
-  urlcolor=blue,
+  linkcolor={red},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={blue},
   pdfcreator={LaTeX via pandoc}}
@@
 \usepackage{xcolor}
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+This phrase, Latin for ``other things held constant'' or ``all else unchanged'', shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 
-This type of exploration is shown in \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (observation 22).
-It is compared with a correctly classified dark chocolate (observation 7).
-The PCA plot and the SHAP PCA plots show a big difference between the two chocolate types but with confusion for a handful of observations.
+This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+It is compared with a CI that is a correctly classified dark chocolate (observation 7).
+The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
 The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
 Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
@@
 
-\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-asimov_grand_1985}{}}%
 Asimov, Daniel.
@@
 
-\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
 Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
 
-\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
 ---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
-\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
 Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
 
-\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
 Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
 
-\leavevmode\hypertarget{ref-buja_grand_1986}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
 Buja, Andreas, and Daniel Asimov.
@@
 
-\leavevmode\hypertarget{ref-chang_shiny_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-chang_shiny_2021}{}}%
 Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
@@
 
-\leavevmode\hypertarget{ref-cook_manual_1997}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
 Cook, Dianne, and Andreas Buja.
@@
 
-\leavevmode\hypertarget{ref-cook_grand_2008}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cook_grand_2008}{}}%
 Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
@@
 
-\leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-da_silva_projection_2021}{}}%
 da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee.
@@
 
-\leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dastin_amazon_2018}{}}%
 Dastin, Jeffrey.
@@
 
-\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-diaz_addressing_2018}{}}%
 Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
@@
 
-\leavevmode\hypertarget{ref-duffy_apple_2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-duffy_apple_2019}{}}%
 Duffy, Claire.
@@
 
-\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
 Gorman, Kristen B., Tony D.
@@
 
-\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
 Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
-\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
 Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
-\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
 Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
@@
 
-\leavevmode\hypertarget{ref-karwowski_international_2006}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-karwowski_international_2006}{}}%
 Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
 
-\leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-kodiyan_overview_2019}{}}%
 Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
 
-\leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-kominsarczyk_treeshap_2021}{}}%
 Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
 
-\leavevmode\hypertarget{ref-larson_how_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-larson_how_2016}{}}%
 Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
@@
 
-\leavevmode\hypertarget{ref-lee_state_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lee_state_2021}{}}%
 Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
@@
 
-\leavevmode\hypertarget{ref-lee_pptree_2013}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lee_pptree_2013}{}}%
 Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
 
-\leavevmode\hypertarget{ref-leone_fifa_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-leone_fifa_2020}{}}%
 Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
 
-\leavevmode\hypertarget{ref-liaw_classification_2002}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-liaw_classification_2002}{}}%
 Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
 
-\leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lundberg_consistent_2018}{}}%
 Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
 
-\leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lundberg_unified_2017}{}}%
 Lundberg, Scott M., and Su-In Lee.
@@
 
-\leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-molnar_interpretable_2020}{}}%
 Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
 
-\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-oneil_weapons_2016}{}}%
 O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
 
-\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
 Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
@@
 
-\leavevmode\hypertarget{ref-shapley_value_1953}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
 Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
 
-\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
 Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
 
-\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
 Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
@@
 
-\leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_not_2016}{}}%
 Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
@@
 
-\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-sievert_interactive_2020}{}}%
 Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
 
-\leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-simonyan_deep_2014}{}}%
 Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
@@
 
-\leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
 Spyrison, Nicholas, and Dianne Cook.
@@
 
-\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
 Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
-\leavevmode\hypertarget{ref-unwin}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-unwin}{}}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
-\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
 Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
@@
 
-\leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-wickham_ggplot2_2016}{}}%
 Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
-\leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-wickham_visualizing_2015}{}}%
 Wickham, Hadley, Dianne Cook, and Heike Hofmann.

### 2022-01-14 figures/_make_figures_case_study.r
decrease height of fig7&8, overlapped page number.

@@ -296,7 +296,7 @@ ggplot2::ggsave(
 ggplot2::ggsave(
   "./figures/case_fifa.png",
   plot = .cp, device = "png",
-  width = 6, height = 9, units = "in")
+  width = 6, height = 8, units = "in")
 .m <- gc()
 
 ### Save .mp4, add GitHub urls to paper
@@ -361,7 +361,7 @@ messgae("Manual capturing this tour from app.")
 ggplot2::ggsave(
   "./figures/case_ames2018.png",
   plot = .cp, device = "png",
-  width = 6, height = 9, units = "in")
+  width = 6, height = 8, units = "in")
 .m <- gc()
 
 ### Save .mp4, add GitHub urls to paper

### 2022-01-14 _paper.rmd
revised SHAP explanation

@@
 <!-- SHAP -->
-SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
-These explanations almost all point to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by permuting the players that contributing to the score.
-@strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
-While the models themselves are invariant to the order of the predictors, the additive explanations are not.
-The SHAP values are the mean contributions over some subset of these permutations.
-This is also related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance, by examining the effect of other features on the predictions.
+The explanations of SHAP almost all refer to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
+@strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
+<!--While the models themselves are invariant to the order of the predictors, the additive explanations are not.--> The SHAP values are the mean contributions over some subset of permutations of features.
+The approach is  related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
 Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as to the starting projection basis (call this the _attribution projection_) to further scrutinize the explanation.
+The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.
 

### 2022-01-14 _paper.tex
revised SHAP explanation

--- 2022-01-14 _paper.tex
@@
 
-SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
-These explanations almost all point to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by permuting the players that contributing to the score.
-Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
-While the models themselves are invariant to the order of the predictors, the additive explanations are not.
-The SHAP values are the mean contributions over some subset of these permutations.
-This is also related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance, by examining the effect of other features on the predictions.
+The explanations of SHAP almost all refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
+Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
+The SHAP values are the mean contributions over some subset of permutations of features.
+The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
 Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as to the starting projection basis (call this the \emph{attribution projection}) to further scrutinize the explanation.
+The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) which will have coefficients varied in order to further scrutinize the feature contributions.
 

### 2022-01-14 _paper.rmd
attributuon example revised

@@
 ```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-A star offensive and defensive player, Messi and a defensive player are compared, L.
-Messi and V van Dijk respectively.
+A star offensive and defensive player are compared, L.
+Messi and V.
+van Dijk, respectively.
 Panel (a) is the break-down plot showing one additive feature-attribution.
-These attributions are sensitive to the order of the features.
-Panel (b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}

### 2022-01-14 _paper.tex
attributuon example revised

@@
 \caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-A star offensive and defensive player, Messi and a defensive player are compared, L.
-Messi and V van Dijk respectively.
+A star offensive and defensive player are compared, L.
+Messi and V.
+van Dijk, respectively.
 Panel (a) is the break-down plot showing one additive feature-attribution.
-These attributions are sensitive to the order of the features.
-Panel (b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}

### 2022-01-15 _paper.rmd
-Add ref to fig2/3 in sect 4 text. -obs-> inst, var -> feat - chocolate_inverse .png, and new .mp4 manual capture for all classification.

--- 2022-01-14 _paper.rmd
@@
 A visualization of these spaces is provided by the first two principal components of their respective spaces.
-In addition, a plot observed by predicted response is also provided.
-A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
+In addition, a plot observed by predicted response is also provided (Fig.
+\@ref(fig:classificationcase), b) A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
 <!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Fig.
+\@ref(fig:classificationcase) a and c).
 
@@
 The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
+These instances are are highlighted as asterisk and $\times$ in 2D spaces and long dashed and short dotted lines on 1D spaces.
 
@@
 
-<!-- A function of the obs selected -->
-The global view facilitated the selection instances.
-The feature-level attribution of the primary instance is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of the contributions of the linear features that best explain the difference between the model intercept and an instance's prediction, not the local shape of the model surface.
+<!-- Segue of PI to attribution projection -->
+The global view facilitated the selection of instances.
+The attribution projection of the PI is the initial 1D basis in a radial tour.
+This is an approximation of the feature importance for the prediction of the instance that best explain the difference between the data mean and an instance's prediction, not the local shape of the model surface.
 
 <!-- Tour start frame -->
-The initial frame is the normalized SHAP values of the PI.
-The distribution normalized values from all instances are shown depicted in the basis as vertical parallel coordinate plots, where each line connects an instance's feature attribution.
-The PI is highlighted as a dashed line and CI as  a dotted line.
-The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal axis.
+Normalized attributions of the PI and CI are shown as dashed and dotted lines against the distribution of the attribution from all instances.
+These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
+\@ref(fig:classificationcase), d).
+This is an important global context, the summarization of the attribution of all instances.
 
 <!-- Tour animation -->
-The radial tour creates a basis path by varying the contribution of a selected feature, fully into and out of a projection frame.
-Doing so tests an individual feature's sensitivity to the structure identified by the local explanation.
-The default feature selected has the largest discrepancy between the attribution of primary and comparison instances.
-The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
-<!-- Segue to classification -->Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
+The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal positions.
+The radial tour is going to vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
+Doing so tests a feature's sensitivity to the structure identified by the local explanation.
+The default feature selected has the largest discrepancy between the attribution of the PI and CI.
+The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizonal position of Fig.
+\@ref(fig:classificationcase) e and Fig.
+\@ref(fig:regressioncase) d.
+
+The following sections elaborate on the difference in displays from applying this approach in classification and regression tasks.
+Now that we have introduced the global view and corresponding radial tour, let us discuss the differences between the classification and regression cases.
 
@@
 Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (c) select features to include, which variable is changed in the tour, and the presence of the parrelle coordiate lines connecting attribution distribution.
+Radial tour inputs (c) select features to include, which feature is changed in the tour, and the presence of the parallel coordiate lines connecting attribution distribution.
 The visual depicting 1D basis (d) shows the distribution of the feature attribution, bars show the current projection with the black bar is being varied.
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete observations, there are 333 instances and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
+After removing incomplete instances, there are 333 instances and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
 A random forest model was fit with species as the response feature.
@@
 To explore this, we use the radial tour to vary the contribution of flipper length (`f_l`).
-(In our exploration, this was the third variable explored.
-It is typically useful to explore the variables with larger contributions, here `b_l` and `b_d`, but when doing this, nothing was revealed about how the PI differed from other Gentoos.) On varying `f_l` as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
+(In our exploration, this was the third feature explored.
+It is typically useful to explore the features with larger contributions, here `b_l` and `b_d`, but when doing this, nothing was revealed about how the PI differed from other Gentoos).
+On varying `f_l` as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
 This suggests that `f_l` should be considered an important feature for explaining the (wrong) prediction.
@@
 Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`f_l`) is important for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
+Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
 
@@
 
-The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
+The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
@@
 A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered a healthy alternative.
-It is interesting to explore which of the nutritional variables contribute most to misclassification.
+It is interesting to explore which of the nutritional features contribute most to misclassification.
 
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `fiber`.
+We remove four features with the lowest contribution for the selected instance and vary the contribution of `fiber`.
 The misclassification of two instances seems improbable when sugar is near the max contribution.
@@
 
-This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
-It is compared with a CI that is a correctly classified dark chocolate (observation 7).
-The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
+This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (instance 22).
+It is compared with a CI that is a correctly classified dark chocolate (instance 7).
+The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
 The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
@@
 The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
-In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+In the density plot, instances on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
 The opposite would be the interpretation of a point with high values in this plot.
@@
 
-The observation of interest (dashed line) can be viewed against the comparison observation (dotted line).
-Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the observation of interest.
+The instance of interest (dashed line) can be viewed against the comparison instance (dotted line).
+Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the instance of interest.
 
-We can quickly compare the variable contributions to the two different predictions from the parallel coordinate plot.
-The observation of interest differs with the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+We can quickly compare the feature contributions to the two different predictions from the parallel coordinate plot.
+The instance of interest differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
 
-From the density plot, which is the attribution projection corresponding to the observation of interest, both observations are more like dark chocolates.
+From the density plot, which is the attribution projection corresponding to the instance of interest, both instances are more like dark chocolates.
 If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
-When primarily Fiber is examined, observation 22 looks more like a milk chocolate.
+When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
 
@@
 While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an observation from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original observation.-->
+In the penguin case we explored how the model misclassified an instance from a type one error perspective.
+In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original instance.-->
 

### 2022-01-15 _paper.tex
-Add ref to fig2/3 in sect 4 text. -obs-> inst, var -> feat - chocolate_inverse .png, and new .mp4 manual capture for all classification.

@@
 
-SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance, by examining the effect of other features on the predictions.
-The explanations of SHAP almost all refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
-Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
-The SHAP values are the mean contributions over some subset of permutations of features.
-The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+SHaply Additive exPlanations (SHAP) approximates the feature importance in the vicinity of one instance conceptually by examining the effect of other features on the contribution of the feature of interest on predicting the response.
+These explanations almost all point to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by permuting the players that contributing to the score.
+Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
+While the models themselves are invariant to the order of the predictors, the additive explanations are not.
+The SHAP values are the mean contributions over some subset of these permutations.
+This is also related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
 Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) which will have coefficients varied in order to further scrutinize the feature contributions.
+The following section will use normalized explanations as to the starting projection basis (call this the \emph{attribution projection}) to further scrutinize the explanation.
 
@@
 \caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-A star offensive and defensive player are compared, L.
-Messi and V.
-van Dijk, respectively.
+A star offensive and defensive player, Messi and a defensive player are compared, L.
+Messi and V van Dijk respectively.
 Panel (a) is the break-down plot showing one additive feature-attribution.
-Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players.
+These attributions are sensitive to the order of the features.
+Panel (b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
@@
 A visualization of these spaces is provided by the first two principal components of their respective spaces.
-In addition, a plot observed by predicted response is also provided.
-A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
+In addition, a plot observed by predicted response is also provided (Fig.
+\ref{fig:classificationcase}, b) A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
 For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Fig.
+\ref{fig:classificationcase} a and c).
 
@@
 The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
+These instances are are highlighted as asterisk and \(\times\) in 2D spaces and long dashed and short dotted lines on 1D spaces.
 
@@
 
-The global view facilitated the selection instances.
-The feature-level attribution of the primary instance is normalized and used as the initial 1D basis in a radial tour.
-This is an approximation of the contributions of the linear features that best explain the difference between the model intercept and an instance's prediction, not the local shape of the model surface.
+The global view facilitated the selection of instances.
+The attribution projection of the PI is the initial 1D basis in a radial tour.
+This is an approximation of the feature importance for the prediction of the instance that best explain the difference between the data mean and an instance's prediction, not the local shape of the model surface.
 
-The initial frame is the normalized SHAP values of the PI.
-The distribution normalized values from all instances are shown depicted in the basis as vertical parallel coordinate plots, where each line connects an instance's feature attribution.
-The PI is highlighted as a dashed line and CI as a dotted line.
-The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal axis.
+Normalized attributions of the PI and CI are shown as dashed and dotted lines against the distribution of the attribution from all instances.
+These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
+\ref{fig:classificationcase}, d).
+This is an important global context, the summarization of the attribution of all instances.
 
-The radial tour creates a basis path by varying the contribution of a selected feature, fully into and out of a projection frame.
-Doing so tests an individual feature's sensitivity to the structure identified by the local explanation.
-The default feature selected has the largest discrepancy between the attribution of primary and comparison instances.
-The following sections elaborate on the takeaways from applying this approach in classification and regression tasks.
-Now that we have introduced the global view and corresponding cheem radial tour, let us discuss the differences between the classification and regression cases.
+The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal positions.
+The radial tour is going to vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
+Doing so tests a feature's sensitivity to the structure identified by the local explanation.
+The default feature selected has the largest discrepancy between the attribution of the PI and CI.
+The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizonal position of Fig.
+\ref{fig:classificationcase} e and Fig.
+\ref{fig:regressioncase} d.
+
+The following sections elaborate on the difference in displays from applying this approach in classification and regression tasks.
+Now that we have introduced the global view and corresponding radial tour, let us discuss the differences between the classification and regression cases.
 
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete observations, there are 333 instances and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
+After removing incomplete instances, there are 333 instances and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
 A random forest model was fit with species as the response feature.
@@
 To explore this, we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
-(In our exploration, this was the third variable explored.
-It is typically useful to explore the variables with larger contributions, here \texttt{b\_l} and \texttt{b\_d}, but when doing this, nothing was revealed about how the PI differed from other Gentoos.) On varying \texttt{f\_l} as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
+(In our exploration, this was the third feature explored.
+It is typically useful to explore the features with larger contributions, here \texttt{b\_l} and \texttt{b\_d}, but when doing this, nothing was revealed about how the PI differed from other Gentoos).
+On varying \texttt{f\_l} as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
 This suggests that \texttt{f\_l} should be considered an important feature for explaining the (wrong) prediction.
@@
 Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{f\_l}) is important for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two variables, mostly because it has an unusually low value of flipper length relative to other Gentoos.
+Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
 
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
+We remove four features with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
 The misclassification of two instances seems improbable when sugar is near the max contribution.
@@
 
-This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
-It is compared with a CI that is a correctly classified dark chocolate (observation 7).
-The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
+This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (instance 22).
+It is compared with a CI that is a correctly classified dark chocolate (instance 7).
+The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
 The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.

### 2022-01-15 figures/_make_figures_case_study.r

@@ -57,7 +57,7 @@ ggplot2::ggsave(
   ### Global view and cheem tour stills
   .glob_view <- global_view(
     penguins_ls, prim_obs, comp_obs, as_ggplot = TRUE) +
-    labs(color = 'Predicted class', shape = 'Predicted class',x = .x_title) +
+    labs(color = "Predicted class", shape = "Predicted class", x = .x_title) +
     ggtitle("Global view") + theme(
       plot.margin      = margin(0,0,0,0),
       legend.margin    = margin(0,0,0,0),
@@ -85,10 +85,9 @@ ggplot2::ggsave(
   ## Using patchwork:
   .pw <- .ggt1 + .ggt2
   .cp <- cowplot::plot_grid(
-    .glob_view, .pw,
-    labels = c("a", "b"),
+    .glob_view, .pw, labels = c("a", "b"),
     ncol = 1, rel_heights = c(1, 1.6),
-    align = 'v', axis = 'l')
+    align = "v", axis = "l")
 }
 
 ### Save still shots for paper
@@ -99,18 +98,18 @@ ggplot2::ggsave(
 .m <- gc()
 
 ### Save .mp4, add GitHub urls to paper
-.ggt <- radial_cheem_tour(
-  penguins_ls, basis = .bas, manip_var = .mv,
-  primary_obs = prim_obs, comparison_obs = comp_obs,
-  do_add_pcp_segments = TRUE,
-  pcp_shape = 124, angle = .15) + ## app angle
-  theme(legend.position = "top", legend.direction = "horizontal")
-.anim <- animate_gganimate(
-  .ggt, fps = 6, res = 100, ## resolution, not the same as dpi, 100 seems about 1x zoom
-  height = 800, width = 600, units = "px", ## "px", "in", "cm", or "mm."
-  render = gganimate::av_renderer("./figures/case_penguins.mp4")) ## Alternative render
-## https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4
-
+message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
+# .ggt <- radial_cheem_tour(
+#   penguins_ls, basis = .bas, manip_var = .mv,
+#   primary_obs = prim_obs, comparison_obs = comp_obs,
+#   do_add_pcp_segments = TRUE,
+#   pcp_shape = 124, angle = .15) + ## app angle
[…]

### 2022-01-15 _paper.rmd

@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized explanations as to the starting projection basis (call this the _attribution projection_) to further scrutinize the explanation.
+The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.
 
 ```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
-A star offensive and defensive player, Messi and a defensive player are compared, L.
-Messi and V van Dijk respectively.
+A star offensive and defensive player are compared, L.
+Messi and V.
+van Dijk, respectively.
 Panel (a) is the break-down plot showing one additive feature-attribution.
-These attributions are sensitive to the order of the features.
-Panel (b) shows the distribution of attribution for each feature across 25 such permutation of the order of predictors, with the mean displayed as a dot, for each players.
+Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}

### 2022-01-16 _paper.rmd
- Add chocolates inverse case. - ames explaintion++ - Comb over explanations and variable formating - checkin before changing SHAP_bd plot adding sequences

@@
 @strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
-<!--While the models themselves are invariant to the order of the predictors, the additive explanations are not.--> The SHAP values are the mean contributions over some subset of permutations of features.
+The attribution of feature importance depends on the seqence of the features already included.
+The SHAP values are the mean contributions over different sequences of the of features.
 The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
@@
 The radial view shows varying the contribution of `f_l` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
-(The .mp4 animation of the radial tour is at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4).)
+(The animation of the radial tour is at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4).)
 ```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
@@
 Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
+From this view is makes sense that is a hard instance to account for as decsion trees can only partition only vertical and horizontal lines.
 
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove four features with the lowest contribution for the selected instance and vary the contribution of `fiber`.
-The misclassification of two instances seems improbable when sugar is near the max contribution.
+We remove four features with the lowest contribution for the selected instance and vary the contribution of Fiber.
+The misclassification seems improbable when sugar is near the max contribution.
 Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
@@
 
-<!-- Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space.
-Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case.
-We look at one extreme instance and vary the contribution of `fiber`.
-We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label.
-While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.
-In the penguin case we explored how the model misclassified an instance from a type one error perspective.
-In contrast, in the chocolates cases shows a type two error view; looking for extreme support of the models classification to challenge the original instance.-->
+(ref:casechocolatesinverse) Chocolates data type classification (milk or dark).
+Looking at the inverse misclassification, we select a milk chocolate while the model predicts it to be dark chocolate.
+Sodium and Fiber (`Na` and `Fbr`) have the largest differences in attributed feature importance.
+We remove features with the lowest contributions and vary the contribution of Sodium.
+This misclassification is not supported when sodium has contribution close to the attribution alligned with the other milk chocolates.
+Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
+```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse)"}
+knitr::include_graphics("./figures/case_chocolates_inverse.png")
+```
+
+It would also be interesting to explore the inverse case; which features lead to a mild chocolate being misclassified as dark.
+Chocolate 84 is just this case and we compare with a correctly predicted milk chocolate (instance 71).
+This exploration is shown in Figure \@ref(fig:casechocolatesinverse).
+From the parallel coordinate lines, we identidy discrepancies in Sodium and Fiber.
+We opt to vary Sodium, and find the attribtuion of the other milk chocolates not to support the prediction of a dark chocolate.
 
@@
 The attribution projection starts with the selected instance on the right.
-We vary the contribution from `defense`, the star offensive player is not distinguished in the horizontal direction.
+We vary the contribution from defense (`def`), the star offensive player is not distinguished in the horizontal direction.
 At this point, defensive players have been rotated to the highest horizonatal value.
-The animate radial tour can be found at [shorturl.at/lCK26](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4)
+The animate radial tour can be found at [shorturl.at/lCK26](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4).
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
@@
 With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-While `offense` and `reaction` skills are both crucial to explaining a star offensive player.
+While offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
 As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features `LotArea` (`LtA`), `OverallQual` (`Qlt`), `YearBuild` (`YrB`), `LivingArea` (`LvA`), `Bathrms` (`Bth`), `Bedrms` (`Bdr`), `TotRms` (`Rms`), `GarageYrBlt` (`GYB`), and `GarageArea` (`GrA`).
+A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Qualality (`Qlt`), Year the house was Build (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Build (`GYB`), and Garage Area (`GrA`).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of `LtA`, which is important to explaining the PI and near invariant to the sales price of the CI.
-The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
+The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4).
 ```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
@@
 
-Figure \@ref(fig:caseames) selects the house sale 74, a sizeable under prediction that has a large contribution to lot area.
+Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction that has a large contribution to lot area.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
+The attribution projection is places instances with high living areas to the right.
+We control the contribution of this feature.
 As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
+This large of an importance is living area is relatively uncommon.
+Boosting tree models may be more resilient to such a under prediction as up-weighting this residual would force its inclusion in the final model.
 

### 2022-01-16 _paper.tex
- Add chocolates inverse case. - ames explaintion++ - Comb over explanations and variable formating - checkin before changing SHAP_bd plot adding sequences

@@
 Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
-The SHAP values are the mean contributions over some subset of permutations of features.
+The attribution of feature importance depends on the seqence of the features already included.
+The SHAP values are the mean contributions over different sequences of the of features.
 The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
@@
 The radial view shows varying the contribution of \texttt{f\_l} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
-(The .mp4 animation of the radial tour is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}.)}\label{fig:casepenguins}
+(The animation of the radial tour is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}.)}\label{fig:casepenguins}
 \end{figure}
@@
 The attribution projection already looks more like a dark chocolate than milk.
-We remove four features with the lowest contribution for the selected instance and vary the contribution of \texttt{fiber}.
-The misclassification of two instances seems improbable when sugar is near the max contribution.
+We remove four features with the lowest contribution for the selected instance and vary the contribution of Fiber.
+The misclassification seems improbable when sugar is near the max contribution.
 Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolates}
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates_inverse}
+
+}
+
+\caption{Chocolates data type classification (milk or dark).
+Looking at the inverse misclassification, we select a milk chocolate while the model predicts it to be dark chocolate.
+Sodium and Fiber (\texttt{Na} and \texttt{Fbr}) have the largest differences in attributed feature importance.
+We remove features with the lowest contributions and vary the contribution of Sodium.
+This misclassification is not supported when sodium has contribution close to the attribution alligned with the other milk chocolates.
+Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolatesinverse}
+\end{figure}
+
+It would also be interesting to explore the inverse case; which features lead to a mild chocolate being misclassified as dark.
+Chocolate 84 is just this case and we compare with a correctly predicted milk chocolate (instance 71).
+This exploration is shown in Figure \ref{fig:casechocolatesinverse}.
+From the parallel coordinate lines, we identidy discrepancies in Sodium and Fiber.
+We opt to vary Sodium, and find the attribtuion of the other milk chocolates not to support the prediction of a dark chocolate.
+
 \hypertarget{fifa-wage-regression}{%
@@
 The attribution projection starts with the selected instance on the right.
-We vary the contribution from \texttt{defense}, the star offensive player is not distinguished in the horizontal direction.
+We vary the contribution from defense (\texttt{def}), the star offensive player is not distinguished in the horizontal direction.
 At this point, defensive players have been rotated to the highest horizonatal value.
-The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{shorturl.at/lCK26}}\label{fig:casefifa}
+The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{shorturl.at/lCK26}.}\label{fig:casefifa}
 \end{figure}
@@
 With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-While \texttt{offense} and \texttt{reaction} skills are both crucial to explaining a star offensive player.
+While offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
 As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features \texttt{LotArea} (\texttt{LtA}), \texttt{OverallQual} (\texttt{Qlt}), \texttt{YearBuild} (\texttt{YrB}), \texttt{LivingArea} (\texttt{LvA}), \texttt{Bathrms} (\texttt{Bth}), \texttt{Bedrms} (\texttt{Bdr}), \texttt{TotRms} (\texttt{Rms}), \texttt{GarageYrBlt} (\texttt{GYB}), and \texttt{GarageArea} (\texttt{GrA}).
+A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Qualality (\texttt{Qlt}), Year the house was Build (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Build (\texttt{GYB}), and Garage Area (\texttt{GrA}).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of \texttt{LtA}, which is important to explaining the PI and near invariant to the sales price of the CI.
-The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}}\label{fig:caseames}
+The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}.}\label{fig:caseames}
 \end{figure}
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizeable under prediction that has a large contribution to lot area.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction that has a large contribution to lot area.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
+The attribution projection is places instances with high living areas to the right.
+We control the contribution of this feature.
 As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
+This large of an importance is living area is relatively uncommon.
+Boosting tree models may be more resilient to such a under prediction as up-weighting this residual would force its inclusion in the final model.
 

### 2022-01-16 figures/_make_figures_case_study.r

@@ -198,7 +198,7 @@ message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
   .inc_var_nms <- c("CalFat", "TotFat", "SatFat", "Na", "Fiber", "Sugars")
   ## Removed 4 with lowest contribution for the prim_obs.
   .bas <- basis_attr_df(chocolates_ls$attr_df[, .inc_var_nms], prim_obs)
-  .mv  <- which(.inc_var_nms == "Sugars")
+  .mv  <- which(.inc_var_nms == "Na")
   mt_interp <- manual_tour(.bas, .mv) %>%
     spinifex:::interpolate_manual_tour(.15) ## App angle.
   dim(mt_interp)
@@ -210,7 +210,7 @@ message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
     pcp_shape = 124, angle = 0) + theme(legend.position = "off", aspect.ratio = 1.4) +
     ggtitle("Radial tour, select frames")
   .ggt2 <- radial_cheem_tour(
-    chocolates_ls, basis = mt_interp[,,18], manip_var = .mv,
+    chocolates_ls, basis = mt_interp[,,19], manip_var = .mv,
     primary_obs = prim_obs, comparison_obs = comp_obs,
     do_add_pcp_segments = FALSE, inc_var_nms = .inc_var_nms,
     pcp_shape = 124, angle = 0) + theme(legend.position = "off", aspect.ratio = 1.4)
@@ -247,7 +247,7 @@ message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
     ggtitle("Global view") + theme(
       plot.margin      = margin(0,0,0,0),
       legend.margin    = margin(0,0,0,0),
-      legend.position  = "bottom",
+      legend.position  = "off",
       legend.direction = "horizontal")
   .inc_var_nms <- c("BMI", "react", "off", "def", "mvm", "pwr")
   ## Removed 4 with lowest contribution for the prim_obs.
@@ -309,7 +309,7 @@ message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
     ggtitle("Global view") + theme(
       plot.margin      = margin(0,0,0,0),
       legend.margin    = margin(0,0,0,0),
-      legend.position  = "bottom",
+      legend.position  = "off",
       legend.direction = "horizontal")
[…]

### 2022-01-16 _paper.rmd
- add 2 sequences to the SHAP_bd plot - sequences of featuers rather than permutations.

@@
 Section \@ref(sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-Then Section \@ref(#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
+Then Section \@ref(sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
 We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explored in the future.
@@
 
-<!-- tree SHAP -->
-For our application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
-Instead of aggregating over permutations of the feature order, tree SHAP approximates instance-level feature importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.
-
 ```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
@@
 van Dijk, respectively.
-Panel (a) is the break-down plot showing one additive feature-attribution.
-Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players.
+Panel (a) shows breakdown plot of three seqences of the features, order and magnitude change.
+Panel (b) shows the distribution of attribution for each feature across 25 seqences of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
@@
 ```
+
+<!-- tree SHAP -->
+For our application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
+Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
+Tree SHAP is only compatible with tree-based models; we illustrate random forests.
+The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.
 

### 2022-01-16 figures/_make_figures.r

@@ -35,9 +35,10 @@
     return(df)
   }
   .lvl_ord <- c("reaction", "offense", "movement", "defense", "power", "accuracy", "BMI", "age", "goalkeeping")
-  my_bd_df <- function(break_down, player_tag = "<tag unused>"){
+  my_bd_df <- function(break_down, player_tag = "<tag unused>", seq_name = "<seq_name unused>", seq_ord = .lvl_ord){
     df <- data.frame(
       player       = player_tag,
+      seq_name     = seq_name,
       label        = break_down$label,
       variable     = break_down$variable_name,
       contribution = break_down$contribution, ## SHAP contribution
@@ -47,83 +48,103 @@
     .n <- nrow(df)
     df$variable[is.na(df$variable)|df$variable==""] <- "prediction"
     df$variable <- factor(
-      df$variable, rev(c("intercept", .lvl_ord, "prediction")))
+      df$variable, rev(c("intercept", seq_ord, "prediction")))
     df$cumulative <- (df$cumulative - min(df$cumulative)) /
       (max(df$cumulative) - min(df$cumulative))
     df$last_cumulative <- c(NA, df$cumulative[-.n])
     df$variable_num <- 1:.n
     df$next_variable_num <- c(2:.n, NA)
     rownames(df) <- paste(break_down$label, break_down$variable, break_down$B, sep = ": ")
+    df <- df[!(df$variable %in% c("intercept", "prediction")), ]
     return(df)
+  }
+  my_bd_plot <- function(df){
+    ggplot() + 
+      ## horizontal bars
+      geom_segment(
+        aes(x = cumulative, xend = last_cumulative,
+            y = variable, yend = variable, color = player),
+        data = df, size = 1.5, alpha = .8) +
+      facet_grid(rows = vars(seq_name), col = vars(player)) +
+      theme_bw() +
+      scale_color_brewer(palette = "Dark2") +
+      labs(title = "Breakdown plot", color = "Players",
+           y = "", x = "Normalized contribution to prediction") +
+      theme(legend.margin   = margin(0, 0, 0, 0),
+            legend.position = "off",
+            axis.text.x     = element_blank(),
+            axis.ticks.x    = element_blank())
   }
 }
 
 ### Create FIFA x ------
-.raw <- DALEX::fifa
-.dat_less_ys <- .raw %>%
-  dplyr::select(-c(`nationality`, ## useless class
[…]

### 2022-01-16 _paper.tex
+knit

@@
 Section \ref{sec:applicationdesign} discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-Then Section @ref(\#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
+Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative response features.
 We conclude with Section \ref{sec:cheemdiscussion} of the insights gained and directions that might be explored in the future.
@@
 
-For our application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
-Instead of aggregating over permutations of the feature order, tree SHAP approximates instance-level feature importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) which will have coefficients varied in order to further scrutinize the feature contributions.
-
 \begin{figure}
@@
 van Dijk, respectively.
-Panel (a) is the break-down plot showing one additive feature-attribution.
-Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players.
+Panel (a) shows breakdown plot of three seqences of the features, order and magnitude change.
+Panel (b) shows the distribution of attribution for each feature across 25 seqences of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
 \end{figure}
+
+For our application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
+Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
+Tree SHAP is only compatible with tree-based models; we illustrate random forests.
+The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) which will have coefficients varied in order to further scrutinize the feature contributions.
 

### 2022-01-16 _paper.rmd
SHAP wordsmithing and plot size.

@@
 
-```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+```{r shapdistrbd, echo=F, fig.align='center', out.width="100%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
 A star offensive and defensive player are compared, L.

### 2022-01-16 figures/_make_figures.r

@@ -204,7 +204,7 @@ dist_df$variable <- factor(dist_df$variable, levels = rev(.lvl_ord))
     theme_bw() +
     scale_color_brewer(palette = "Dark2") +
     scale_fill_brewer(palette = "Dark2") +
-    labs(title="SHAP distribution",
+    labs(title = "SHAP",
          y = "", x = "Normalized SHAP values") +
     theme(legend.margin   = margin(0, 0, 0, 0),
           legend.position = "bottom"))

### 2022-01-16 _paper.rmd
-Grammerly edits -Vimeo links -clean up figure script.

--- 2022-01-16 _paper.rmd
@@
 <!-- Interpretability & baises -->
-Interpretability is vital for exploring and protecting against potential biases (e.g.
-sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]) in any model.
+Interpretability is vital for exploring and protecting against potential biases (e.g., sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]) in any model.
 For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
 Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
-It is also generally important for many problems, where it is not enough to predict accurately, but one must be able to explain which predictors are most responsible for generating a response value.
+It is also generally important for many problems, where it is not enough to predict accurately.
+Still, one must be able to explain which predictors are most responsible for generating a response value.
 
 <!-- Interpretability & data drift -->
-Another concern is data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
+Another concern is data drift, a shift in support or domain of the explanatory features (variable or predictors).
 Non-linear models are typically more sensitive and do not extrapolate well outside the training data domain.
@@
 A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to visualize them to understand a model more comprehensively.
+Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
 
@@
 <!-- Purposed approach -->
-In our proposed approach, we use the radial manual tour to scrutinize a local explanation.
+We use the radial manual tour to scrutinize a local explanation in our proposed approach.
 Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
@@
 A constraint in our approach is that the basis must remain orthonormal.
-That means, when the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
 This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
@@
 The remainder of this paper is organized as follows.
-The following Section, \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced.
+The following Section, \@ref(sec:explanations), covers the background of the local explanation and the traditional visuals produced.
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
@@
 <!-- SHAP -->
-SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance, by examining the effect of other features on the predictions.
+SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
 The explanations of SHAP almost all refer to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
@@
 The attribution of feature importance depends on the seqence of the features already included.
-The SHAP values are the mean contributions over different sequences of the of features.
-The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+The SHAP values are the mean contributions over different feature sequences.
+The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
 Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered to be similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each feature in the context of all other candidate features.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each feature in the context of all other candidate features.
 
-```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+```{r shapdistrbd, echo=F, fig.align='center', out.width="100%", fig.cap = "Illustration SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
 A star offensive and defensive player are compared, L.
@@
 van Dijk, respectively.
-Panel (a) shows breakdown plot of three seqences of the features, order and magnitude change.
+Panel (a) shows breakdown plots of three sequences of the features, order and magnitude change.
 Panel (b) shows the distribution of attribution for each feature across 25 seqences of predictors, with the mean displayed as a dot, for each players.
@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.
+The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) will have coefficients varied to further scrutinize the feature contributions.
 
@@
 The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
-These instances are are highlighted as asterisk and $\times$ in 2D spaces and long dashed and short dotted lines on 1D spaces.
+These instances are highlighted as asterisk and $\times$ in 2D spaces and long dashed and short dotted lines on 1D spaces.
 
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour.
-This is an approximation of the feature importance for the prediction of the instance that best explain the difference between the data mean and an instance's prediction, not the local shape of the model surface.
+This approximation of the feature importance for the prediction of the instance best explains the difference between the data mean and an instance’s prediction, not the local shape of the model surface.
 
 <!-- Tour start frame -->
-Normalized attributions of the PI and CI are shown as dashed and dotted lines against the distribution of the attribution from all instances.
+Normalized attributions of the PI and CI are shown as dashed and dotted lines against the attribution distribution from all instances.
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
@@
 The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal positions.
-The radial tour is going to vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
+The radial tour will vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
 Doing so tests a feature's sensitivity to the structure identified by the local explanation.
 The default feature selected has the largest discrepancy between the attribution of the PI and CI.
-The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizonal position of Fig.
+The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizontal position of Fig.
 \@ref(fig:classificationcase) e and Fig.
@@
 The model information in the global view depicts the model confusion matrix.
-The radial tour is 1D, with density display, while the goal of the tour; exploring the sensitivity of each feature to structure identified by the local explanation, evaluating the support or robustness of the prediction.
+The radial tour is 1D, with a density display, while the goal of the tour; exploring the sensitivity of each feature to structure identified by the local explanation, evaluating the support or robustness of the prediction.
 
@@
 Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (c) select features to include, which feature is changed in the tour, and the presence of the parallel coordiate lines connecting attribution distribution.
-The visual depicting 1D basis (d) shows the distribution of the feature attribution, bars show the current projection with the black bar is being varied.
-Panel (e) is the resulting projection of the data shown as density in the classification case."}
+Radial tour inputs (c) select features to include and  which feature is changed in the tour.
+The visual depicting 1D basis (d) shows the distribution of the feature attribution, and bars show the current basis.
+The black bar is being varied.
+Panel (e) is the resulting projection of the data indicated as density in the classification case."}
 knitr::include_graphics("./figures/app_classification.PNG")
@@
 
-```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression task highlighting differences with classification and interaction.
+```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression task highlighting the differences from the classification task and interactive features.
 Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
@@
 Interactive tabular display (d) populates when instances are selected.
-Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, left and right respectively."}
+Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right)."}
 knitr::include_graphics("./figures/app_regression_interactions.PNG")
@@
 All computational operations should be prepared before runtime.
-The work remaining when an application is ran is solely reacting to inputs and rendering of visuals and tables.
+The work remaining when an application is run solely reacts to inputs and rendering of visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
-This implementation aggregates over exhaustively over all trees' attribution and we opt not to fit interactions of features.
+This implementation aggregates over exhaustively over all trees' attribution, and we opt not to fit interactions of features.
 	\item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
@@
 PCA and statistics of the features and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
+These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
 Its reduced computational complexity makes tree SHAP a good candidate to start with.
-Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 The radial view shows varying the contribution of `f_l` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at [shorturl.at/hFHNS](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4).)
+(The animation of the radial tour is at [vimeo.com/666431172](https://vimeo.com/666431172).)
 ```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
@@
 (ref:casepenguinsblfl) Checking what is learned from the cheem viewer.
-This is a plot of flipper length (`f_l`) and bill length (`b_l`), where the PI is highlighted by an asterisk.
+This is a plot of flipper length (`f_l`) and bill length (`b_l`), where an asterisk highlights the PI.
 A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small `f_l` length which is the primary reason it is confused with a Chinstrap.
+The PI has an unusually small `f_l` length which is why it is confused with a Chinstrap.
 ```{r casepenguinsblfl, echo=F, out.width="60%", fig.cap = "(ref:casepenguinsblfl)"}
@@
 The predicted vs observed plot reveals a handful of misclassified instances.
-For this illustration, we will explore why a Gentoo has been wrongly labeled as a Chinstrap.
+We will explore why a Gentoo has been wrongly labeled as a Chinstrap for this illustration.
 The PI is a misclassified point (represented by the asterisk in the global view and as a dashed vertical line in the tour view).
@@
 The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance is mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap, that is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
+The important features identified by SHAP in the (wrong) prediction for this instance are mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
+This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
+That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
 To explore this, we use the radial tour to vary the contribution of flipper length (`f_l`).
@@
 Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view is makes sense that is a hard instance to account for as decsion trees can only partition only vertical and horizontal lines.
+From this view it makes sense that its a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
 
@@
 Dark chocolate is considered healthier than milk.
-The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website, and normalized to 100g equivalents.
+The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
 The data is available in the `cheem` package.
@@
 The misclassification seems improbable when sugar is near the max contribution.
-Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
+Animated tour can be found at [vimeo.com/666431143](https://vimeo.com/666431143).
 ```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
@@
 We remove features with the lowest contributions and vary the contribution of Sodium.
-This misclassification is not supported when sodium has contribution close to the attribution alligned with the other milk chocolates.
-Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
+This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
+Animated tour can be found at [vimeo.com/666431143](https://vimeo.com/666431143).
 ```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse)"}
@@
 It would also be interesting to explore the inverse case; which features lead to a mild chocolate being misclassified as dark.
-Chocolate 84 is just this case and we compare with a correctly predicted milk chocolate (instance 71).
+Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
 This exploration is shown in Figure \@ref(fig:casechocolatesinverse).
-From the parallel coordinate lines, we identidy discrepancies in Sodium and Fiber.
-We opt to vary Sodium, and find the attribtuion of the other milk chocolates not to support the prediction of a dark chocolate.
+From the parallel coordinate lines, we identify discrepancies in Sodium and Fiber.
+We opt to vary Sodium, and find the attribution of the other milk chocolates not to support the prediction of a dark chocolate.
 
@@
 Messi) with that of a top defensive fielder (V.
-van Dijk), the same instances were used in figure \@ref(fig:shapdistrbd).
+van Dijk).
+The same instances were used in figure \@ref(fig:shapdistrbd).
 
@@
 We vary the contribution from defense (`def`), the star offensive player is not distinguished in the horizontal direction.
-At this point, defensive players have been rotated to the highest horizonatal value.
-The animate radial tour can be found at [shorturl.at/lCK26](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4).
+At this point, defensive players have been rotated to the highest horizontal  value.
+The animate radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
@@
 As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
-In terms of what-if analysis, were Messi's tree SHAP attributions at these levels, his predicted wages would be halved.
+In terms of what-if analysis, his predicted wages would be halved if Messi’s tree SHAP attributions at these levels.
 
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Qualality (`Qlt`), Year the house was Build (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Build (`GYB`), and Garage Area (`GrA`).
+A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Build (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Build (`GYB`), and Garage Area (`GrA`).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 (ref:caseames) Ames housing 2018 regressing sales price [USD].
-The PI sale price was under predicted, and had sizable attribution to lot area (`LtA`).
+The PI sale price was under predicted and had sizable attribution to lot area (`LtA`).
 The CI was predicted sales price was similar and much more accurate with its observed sales price while it has very little attribution to lot area.
 Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of `LtA`, which is important to explaining the PI and near invariant to the sales price of the CI.
-The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4).
+The corresponding animation is at [vimeo.com/666431134](https://vimeo.com/666431134).
 ```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
@@
 
-Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction that has a large contribution to lot area.
+Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction that has a large contribution from lot area.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
@@
 This large of an importance is living area is relatively uncommon.
-Boosting tree models may be more resilient to such a under prediction as up-weighting this residual would force its inclusion in the final model.
+Boosting tree models may be more resilient to such an under prediction as up-weighting this residual would force its inclusion in the final model.
 

### 2022-01-16 _paper.tex
SHAP wordsmithing and plot size.
-Grammerly edits -Vimeo links -clean up figure script.

@@
 
-Interpretability is vital for exploring and protecting against potential biases (e.g.~sex (Dastin 2018; Duffy 2019), race (Larson et al.
+Interpretability is vital for exploring and protecting against potential biases (e.g., sex (Dastin 2018; Duffy 2019), race (Larson et al.
 2016), and age (Díaz et al.
@@
 Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
-It is also generally important for many problems, where it is not enough to predict accurately, but one must be able to explain which predictors are most responsible for generating a response value.
+It is also generally important for many problems, where it is not enough to predict accurately.
+Still, one must be able to explain which predictors are most responsible for generating a response value.
 
-Another concern is data drift, which is a shift in support or domain of the explanatory features (variable or predictors).
+Another concern is data drift, a shift in support or domain of the explanatory features (variable or predictors).
 Non-linear models are typically more sensitive and do not extrapolate well outside the training data domain.
@@
 A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to visualize them to understand a model more comprehensively.
+Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
 
@@
 
-In our proposed approach, we use the radial manual tour to scrutinize a local explanation.
+We use the radial manual tour to scrutinize a local explanation in our proposed approach.
 Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
@@
 The remainder of this paper is organized as follows.
-The following Section, \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
+The following Section, \ref{sec:explanations}, covers the background of the local explanation and the traditional visuals produced.
 Section \ref{sec:tour} explains the animations of continuous linear projections.
@@
 
-SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance, by examining the effect of other features on the predictions.
+SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
 The explanations of SHAP almost all refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
@@
 The attribution of feature importance depends on the seqence of the features already included.
-The SHAP values are the mean contributions over different sequences of the of features.
-The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature, after fixing the value of all other features to their mean.
+The SHAP values are the mean contributions over different feature sequences.
+The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
 Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered to be similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each feature in the context of all other candidate features.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{./figures/shap_distr_bd}
+{\centering \includegraphics[width=1\linewidth]{./figures/shap_distr_bd}
 
@@
 
-\caption{Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+\caption{Illustration SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
 A star offensive and defensive player are compared, L.
@@
 Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) which will have coefficients varied in order to further scrutinize the feature contributions.
+The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) will have coefficients varied to further scrutinize the feature contributions.
 
@@
 The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
-These instances are are highlighted as asterisk and \(\times\) in 2D spaces and long dashed and short dotted lines on 1D spaces.
+These instances are highlighted as asterisk and \(\times\) in 2D spaces and long dashed and short dotted lines on 1D spaces.
 
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour.
-This is an approximation of the feature importance for the prediction of the instance that best explain the difference between the data mean and an instance's prediction, not the local shape of the model surface.
+This approximation of the feature importance for the prediction of the instance best explains the difference between the data mean and an instance's prediction, not the local shape of the model surface.
 
-Normalized attributions of the PI and CI are shown as dashed and dotted lines against the distribution of the attribution from all instances.
+Normalized attributions of the PI and CI are shown as dashed and dotted lines against the attribution distribution from all instances.
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
@@
 The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal positions.
-The radial tour is going to vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
+The radial tour will vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
 Doing so tests a feature's sensitivity to the structure identified by the local explanation.
 The default feature selected has the largest discrepancy between the attribution of the PI and CI.
-The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizonal position of Fig.
+The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizontal position of Fig.
 \ref{fig:classificationcase} e and Fig.
@@
 Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (c) select features to include, which feature is changed in the tour, and the presence of the parallel coordiate lines connecting attribution distribution.
-The visual depicting 1D basis (d) shows the distribution of the feature attribution, bars show the current projection with the black bar is being varied.
-Panel (e) is the resulting projection of the data shown as density in the classification case.}\label{fig:classificationcase}
+Radial tour inputs (c) select features to include and  which feature is changed in the tour.
+The visual depicting 1D basis (d) shows the distribution of the feature attribution, and bars show the current basis.
+The black bar is being varied.
+Panel (e) is the resulting projection of the data indicated as density in the classification case.}\label{fig:classificationcase}
 \end{figure}
@@
 
-\caption{Overview of the cheem viewer for regression task highlighting differences with classification and interaction.
+\caption{Overview of the cheem viewer for regression task highlighting the differences from the classification task and interactive features.
 Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
@@
 Interactive tabular display (d) populates when instances are selected.
-Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, left and right respectively.}\label{fig:regressioncase}
+Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right).}\label{fig:regressioncase}
 \end{figure}
@@
 (2021).
-This implementation aggregates over exhaustively over all trees' attribution and we opt not to fit interactions of features.
+This implementation aggregates over exhaustively over all trees' attribution, and we opt not to fit interactions of features.
     \item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
@@
 PCA and statistics of the features and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that the bulk of the time will be spent on the local attribution.
+These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
 Its reduced computational complexity makes tree SHAP a good candidate to start with.
-Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, which are attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
 
@@
 The radial view shows varying the contribution of \texttt{f\_l} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4}{shorturl.at/hFHNS}.)}\label{fig:casepenguins}
+(The animation of the radial tour is at \href{https://vimeo.com/666431172}{vimeo.com/666431172}.)}\label{fig:casepenguins}
 \end{figure}
@@
 \caption{Checking what is learned from the cheem viewer.
-This is a plot of flipper length (\texttt{f\_l}) and bill length (\texttt{b\_l}), where the PI is highlighted by an asterisk.
+This is a plot of flipper length (\texttt{f\_l}) and bill length (\texttt{b\_l}), where an asterisk highlights the PI.
 A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small \texttt{f\_l} length which is the primary reason it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
+The PI has an unusually small \texttt{f\_l} length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
 \end{figure}
@@
 The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance is mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap, that is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
+The important features identified by SHAP in the (wrong) prediction for this instance are mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
+This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
+That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
 To explore this, we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
@@
 Dark chocolate is considered healthier than milk.
-The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website, and normalized to 100g equivalents.
+The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
 The data is available in the \texttt{cheem} package.
@@
 The misclassification seems improbable when sugar is near the max contribution.
-Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolates}
+Animated tour can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.}\label{fig:casechocolates}
 \end{figure}
@@
 We remove features with the lowest contributions and vary the contribution of Sodium.
-This misclassification is not supported when sodium has contribution close to the attribution alligned with the other milk chocolates.
-Animated tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4}{shorturl.at/dBE28}.}\label{fig:casechocolatesinverse}
+This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
+Animated tour can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.}\label{fig:casechocolatesinverse}
 \end{figure}
@@
 It would also be interesting to explore the inverse case; which features lead to a mild chocolate being misclassified as dark.
-Chocolate 84 is just this case and we compare with a correctly predicted milk chocolate (instance 71).
+Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
 This exploration is shown in Figure \ref{fig:casechocolatesinverse}.
-From the parallel coordinate lines, we identidy discrepancies in Sodium and Fiber.
-We opt to vary Sodium, and find the attribtuion of the other milk chocolates not to support the prediction of a dark chocolate.
+From the parallel coordinate lines, we identify discrepancies in Sodium and Fiber.
+We opt to vary Sodium, and find the attribution of the other milk chocolates not to support the prediction of a dark chocolate.
 
@@
 Messi) with that of a top defensive fielder (V.
-van Dijk), the same instances were used in figure \ref{fig:shapdistrbd}.
+van Dijk).
+The same instances were used in figure \ref{fig:shapdistrbd}.
 
@@
 We vary the contribution from defense (\texttt{def}), the star offensive player is not distinguished in the horizontal direction.
-At this point, defensive players have been rotated to the highest horizonatal value.
-The animate radial tour can be found at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_fifa.mp4}{shorturl.at/lCK26}.}\label{fig:casefifa}
+At this point, defensive players have been rotated to the highest horizontal value.
+The animate radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.}\label{fig:casefifa}
 \end{figure}
@@
 As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
-In terms of what-if analysis, were Messi's tree SHAP attributions at these levels, his predicted wages would be halved.
+In terms of what-if analysis, his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
 
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Qualality (\texttt{Qlt}), Year the house was Build (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Build (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Build (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Build (\texttt{GYB}), and Garage Area (\texttt{GrA}).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 \caption{Ames housing 2018 regressing sales price {[}USD{]}.
-The PI sale price was under predicted, and had sizable attribution to lot area (\texttt{LtA}).
+The PI sale price was under predicted and had sizable attribution to lot area (\texttt{LtA}).
 The CI was predicted sales price was similar and much more accurate with its observed sales price while it has very little attribution to lot area.
 Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of \texttt{LtA}, which is important to explaining the PI and near invariant to the sales price of the CI.
-The corresponding animation is at \href{https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4}{shorturl.at/oqyDU}.}\label{fig:caseames}
+The corresponding animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.}\label{fig:caseames}
 \end{figure}
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction that has a large contribution to lot area.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction that has a large contribution from lot area.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.

### 2022-01-16 figures/_make_figures_case_study.r

@@ -27,27 +27,6 @@
 }
 
 ## Penguins classification ------
-### Save orthogonal display of penguins with
-(.g <- ggplot(penguins_na.rm, aes(x=bill_length_mm,
-                                  y=flipper_length_mm,
-                                  colour=species)) +
-   geom_point(size=1) +
-   ## Prim obs
-   geom_point(data=penguins_na.rm[243,], shape=8, size=5, alpha=0.8) +
-   ## Comparison obs
-   #geom_point(data=penguins_na.rm[169,], shape=4, size=3, alpha=0.6) +
-   theme_bw() +
-   theme(aspect.ratio = 1) +
-   #legend.margin   = margin(0, 0, 0, 0),
-   #legend.position = "bottom") +
-   scale_color_brewer(palette = "Dark2") +
-   labs(y = "Flipper length [mm]", x = "Bill length [mm]", color = "Observed species"))
-## Save
-ggplot2::ggsave(
-  "./figures/case_penguins_BlFl.png",
-  plot = .g, device = "png",
-  height = 4, units = "in")
-
 {
   ## Data setup, spinifex::penguins
   names(penguins_ls)
@@ -90,6 +69,26 @@ ggplot2::ggsave(
     align = "v", axis = "l")
 }
 
+## Penguins orthogonal BlFl ----
+(.g <- ggplot(penguins_na.rm, aes(x=bill_length_mm,
+                                  y=flipper_length_mm,
+                                  colour=species)) +
+   geom_point(size=1) +
+   ## Prim obs
+   geom_point(data=penguins_na.rm[243,], shape=8, size=5, alpha=0.8) +
+   ## Comparison obs
+   #geom_point(data=penguins_na.rm[169,], shape=4, size=3, alpha=0.6) +
+   theme_bw() +
+   theme(aspect.ratio = 1) +
+   
+   scale_color_brewer(palette = "Dark2") +
+   labs(y = "Flipper length [mm]", x = "Bill length [mm]", color = "Observed species"))
+## Save
+ggplot2::ggsave(
+  "./figures/case_penguins_BlFl.png",
+  plot = .g, device = "png",
+  height = 4, units = "in")
+
 ### Save still shots for paper
 ggplot2::ggsave(
   "./figures/case_penguins.png",
@@ -99,17 +98,7 @@ ggplot2::ggsave(
 
 ### Save .mp4, add GitHub urls to paper
[…]

### 2022-01-16 _paper.rmd
Linting and touch ups from updating thesis.

@@
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
-Section \@ref(sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
+Section \@ref(sec:cheemviwer) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
 Then Section \@ref(sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
@@
 
-# The cheem viewer {#sec:applicationdesign}
+# The cheem viewer {#sec:cheemviwer}
 
@@
 
-The following __R__ code will help with geting up and running:
+The following __R__ code will help with getting up and running:
 
@@
 The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
-The data is available in the `cheem` package.
+The data is available in the __cheem__ package.
 A random forest model is used for the classification of chocolate type.

### 2022-01-16 _paper.tex
Linting and touch ups from updating thesis.

@@
 Section \ref{sec:tour} explains the animations of continuous linear projections.
-Section \ref{sec:applicationdesign} discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
+Section \ref{sec:cheemviwer} discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
 Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative response features.
@@
 
-\hypertarget{sec:applicationdesign}{%
-\section{The cheem viewer}\label{sec:applicationdesign}}
+\hypertarget{sec:cheemviwer}{%
+\section{The cheem viewer}\label{sec:cheemviwer}}
 
@@
 
-The following \textbf{R} code will help with geting up and running:
+The following \textbf{R} code will help with getting up and running:
 
@@
 The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
-The data is available in the \texttt{cheem} package.
+The data is available in the \textbf{cheem} package.
 A random forest model is used for the classification of chocolate type.

### 2022-01-17 _paper.rmd
bib consistency and a couple typos.

@@
 There are different reasons and emphases to fit a model.
-@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomizes modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample instances.
+@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomize modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample instances.
 The intended use has important implications for model selection and development.
@@
 @strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
-The attribution of feature importance depends on the seqence of the features already included.
+The attribution of feature importance depends on the sequence of the features already included.
 The SHAP values are the mean contributions over different feature sequences.
@@
 Panel (a) shows breakdown plots of three sequences of the features, order and magnitude change.
-Panel (b) shows the distribution of attribution for each feature across 25 seqences of predictors, with the mean displayed as a dot, for each players.
+Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
@@
 
-To explore the local explanations, an ensemble of plots [@unwin] is provided, called the _cheem viewer_.
+To explore the local explanations, an ensemble of plots [@unwin_ensemble_2018] is provided, called the _cheem viewer_.
 There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
@@
 As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
-In terms of what-if analysis, his predicted wages would be halved if Messi’s tree SHAP attributions at these levels.
+In terms of what-if analysis, his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
 

### 2022-01-17 _paper.tex
bib consistency and a couple typos.

@@
 There are different reasons and emphases to fit a model.
-Breiman (2001), reiterated by Shmueli (2010), taxonomizes modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample instances.
+Breiman (2001), reiterated by Shmueli (2010), taxonomize modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample instances.
 The intended use has important implications for model selection and development.
@@
 Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
-The attribution of feature importance depends on the seqence of the features already included.
+The attribution of feature importance depends on the sequence of the features already included.
 The SHAP values are the mean contributions over different feature sequences.
@@
 Panel (a) shows breakdown plots of three sequences of the features, order and magnitude change.
-Panel (b) shows the distribution of attribution for each feature across 25 seqences of predictors, with the mean displayed as a dot, for each players.
+Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot, for each players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
@@
 
-\leavevmode\hypertarget{ref-unwin}{}%
+\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.

### 2022-01-20 figures/_make_figures_case_study.r
x_axis positioning, recapture manual annoted screen shots.

@@ -21,9 +21,9 @@
   ames2018_ls   <- readRDS(paste0(fp, "preprocess_ames2018.rds"))
   chocolates_ls <- readRDS(paste0(fp, "preprocess_chocolates.rds"))
   
-  .x_title      <- "        x: PC1, y: PC2          x: PC1, y: PC2  x: predicted, y: observed"
-  .x_title_reg  <- "      x: PC1, y: PC2               x: PC1, y: PC2     x: predicted, y: observed"
-  .x_title_reg2 <- "        x: PC1, y: PC2        x: PC1, y: PC2    x: predicted, y: observed"
+  .x_title_peng <- "x: PC1, y: PC2          x: PC1, y: PC2        x: pred, y: obs"
+  .x_title_choc <- "x: PC1, y: PC2                x: PC1, y: PC2             x: pred, y: obs"
+  .x_title_reg  <- "x: PC1, y: PC2                       x: PC1, y: PC2                      x: pred, y: obs"
 }
 
 ## Penguins classification ------
@@ -36,7 +36,7 @@
   ### Global view and cheem tour stills
   .glob_view <- global_view(
     penguins_ls, prim_obs, comp_obs, as_ggplot = TRUE) +
-    labs(color = "Predicted class", shape = "Predicted class", x = .x_title) +
+    labs(color = "Predicted class", shape = "Predicted class", x = .x_title_peng) +
     ggtitle("Global view") + theme(
       plot.margin      = margin(0,0,0,0),
       legend.margin    = margin(0,0,0,0),
@@ -69,10 +69,18 @@
     align = "v", axis = "l")
 }
 
+### Save still shots for paper
+ggplot2::ggsave(
+  "./figures/case_penguins.png",
+  plot = .cp, device = "png",
+  width = 4.5, height = 7, units = "in")
+.m <- gc()
+
+
 ## Penguins orthogonal BlFl ----
-(.g <- ggplot(penguins_na.rm, aes(x=bill_length_mm,
-                                  y=flipper_length_mm,
-                                  colour=species)) +
+(.g <- ggplot(penguins_na.rm, aes(x = bill_length_mm,
+                                  y = flipper_length_mm,
+                                  colour = species)) +
    geom_point(size=1) +
    ## Prim obs
    geom_point(data=penguins_na.rm[243,], shape=8, size=5, alpha=0.8) +
@@ -88,13 +96,6 @@ ggplot2::ggsave(
   "./figures/case_penguins_BlFl.png",
[…]

### 2022-02-03 _paper.rmd
modifications to section 4

@@
 A visualization of these spaces is provided by the first two principal components of their respective spaces.
-In addition, a plot observed by predicted response is also provided (Fig.
-\@ref(fig:classificationcase), b) A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
+A single 2D projection will not encompass all of the structure of higher-dimensional space, but can still be a useful visual summary.
+In addition, a plot of the observed against predicted response values is also provided (Fig.
+\@ref(fig:classificationcase), b), to help identify instances poorly predicted by the model.
 <!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
@@
 While the comparison of these spaces is interesting, the main purpose of the global view is to enable the selection of instances to explore the local explanations.
-The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
-These instances are highlighted as asterisk and $\times$ in 2D spaces and long dashed and short dotted lines on 1D spaces.
+The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
+These instances are highlighted as asterisk and $\times$ respectively.
 
@@
 
+XXX Re-order these paragraphs
 <!-- Segue of PI to attribution projection -->
-The global view facilitated the selection of instances.
-The attribution projection of the PI is the initial 1D basis in a radial tour.
-This approximation of the feature importance for the prediction of the instance best explains the difference between the data mean and an instance’s prediction, not the local shape of the model surface.
+The attribution projection of the PI (selected from the global view) is the initial 1D basis in a radial tour (Fig.
+\@ref(fig:classificationcase), e).
+This is an  approximation of the feature importance, for the prediction of this instance.
+This combination of features best explains the difference between the mean response and an instance’s predicted value.
+It is not an indication of the local shape of the model surface, that is, some indication of the tangent to the curve at this point.
 
 <!-- Tour start frame -->
-Normalized attributions of the PI and CI are shown as dashed and dotted lines against the attribution distribution from all instances.
+The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively, overlaid on attribution projections from all instances.
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
@@
 
-The following sections elaborate on the difference in displays from applying this approach in classification and regression tasks.
+XXX Wrong place for this The following sections elaborate on the difference in displays from applying this approach in classification and regression tasks.
 Now that we have introduced the global view and corresponding radial tour, let us discuss the differences between the classification and regression cases.

### 2022-02-03 _paper.rmd
Rotate fifa and ames use cases: basis below scatterplot; and knit.

--- 2022-01-17 _paper.rmd
@@
 <!-- What-if & counterfactual analysis -->
-Using a radial tour can be considered similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020].
-This phrase, Latin for "other things held constant" or "all else unchanged", shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+Using a radial tour can be compared with counterfactual, what-if analysis, such as _ceteris paribus_ profiles [@biecek_ceterisparibus_2020].
+Ceteris paribus, is Latin for "other things held constant" or "all else unchanged".
+These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.

### 2022-02-03 _paper.tex
Rotate fifa and ames use cases: basis below scatterplot; and knit.

@@
 
-Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020).
-This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
+Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.

### 2022-02-03 figures/_make_figures_case_study.r

@@ -212,6 +212,136 @@ message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
 
 
 
+## ONE OFF FUNC; ----
+## regression case basis below scatterplot
+THIS_REG_radial_cheem_tour  <- function(
+  cheem_ls, basis, manip_var, 
+  primary_obs         = NULL,
+  comparison_obs      = NULL,
+  do_add_pcp_segments = TRUE,
+  pcp_shape           = c(142, 124, 3), ## '|' plotly and gganimate, or '+' respectively
+  angle               = .15,
+  row_index           = NULL,
+  inc_var_nms         = NULL,
+  do_center_frame     = TRUE,
+  do_add_residual     = FALSE
+){
+  if(is.null(row_index) == FALSE)
+    if(sum(row_index) == 0L) 
+      stop("radial_cheem_tour: sum of row_index was 0.")
+  
+  ## Initialize
+  x <- y <- NULL
+  decode_df <- cheem_ls$decode_df
+  .prim_obs <- primary_obs    ## Proto_basis1d_distribution EXPECTS NUMERIC INDEX;
+  .comp_obs <- comparison_obs ## Don't coerce to logical index
+  .n        <- nrow(decode_df)
+  ## column & row indexes
+  if(is.null(inc_var_nms)) inc_var_nms <- colnames(cheem_ls$attr_df)
+  .col_idx <- colnames(decode_df) %in% inc_var_nms
+  if(is.null(row_index) == FALSE){
+    ## Change row_index from numeric to logical if needed and replicate
+    row_index <- as_logical_index(row_index, .n)
+    row_index[c(.prim_obs, .comp_obs)] <- TRUE
+  }
+  ## Subset columns and scale plot data
+  .dat <- decode_df[, .col_idx] %>% spinifex::scale_sd() %>%
+    spinifex::scale_01() %>% as.data.frame()
+  ## Manual (radial) tour 1d
+  .mt_path <- spinifex::manual_tour(basis, manip_var)
+  
+  ## Problem type & aesthetics
+  .prob_type <- cheem_ls$type ## Either "classification" or "regression"
+  .alpha <- logistic_tform(.n)
+  
+  ### Classification case ---
+  if(.prob_type == "classification")
+    stop("NS: DO NOT USE THIS ONE OFF FUNCTION FOR CLASSIFICATION, 
+         it was made to make the regression case basis below the scatterplot")
+  
[…]

### 2022-02-03 _paper.tex

@@
 A visualization of these spaces is provided by the first two principal components of their respective spaces.
-In addition, a plot observed by predicted response is also provided (Fig.
-\ref{fig:classificationcase}, b) A single 2D projection will not encompass all of the structure of higher-dimensional space, but it is generally a useful visual summary.
+A single 2D projection will not encompass all of the structure of higher-dimensional space, but can still be a useful visual summary.
+In addition, a plot of the observed against predicted response values is also provided (Fig.
+\ref{fig:classificationcase}, b), to help identify instances poorly predicted by the model.
 For classification tasks, misclassified instances are circled in red if applicable.
@@
 While the comparison of these spaces is interesting, the main purpose of the global view is to enable the selection of instances to explore the local explanations.
-The projection attribution of the Primary Instance (PI) is examined and typically viewed with an optional Comparison Instance (CI).
-These instances are highlighted as asterisk and \(\times\) in 2D spaces and long dashed and short dotted lines on 1D spaces.
+The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
+These instances are highlighted as asterisk and \(\times\) respectively.
 
@@
 
-The global view facilitated the selection of instances.
-The attribution projection of the PI is the initial 1D basis in a radial tour.
-This approximation of the feature importance for the prediction of the instance best explains the difference between the data mean and an instance's prediction, not the local shape of the model surface.
+XXX Re-order these paragraphs
+The attribution projection of the PI (selected from the global view) is the initial 1D basis in a radial tour (Fig.
+\ref{fig:classificationcase}, e).
+This is an approximation of the feature importance, for the prediction of this instance.
+This combination of features best explains the difference between the mean response and an instance's predicted value.
+It is not an indication of the local shape of the model surface, that is, some indication of the tangent to the curve at this point.
 
-Normalized attributions of the PI and CI are shown as dashed and dotted lines against the attribution distribution from all instances.
+The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively, overlaid on attribution projections from all instances.
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.

### 2022-02-04 _paper.rmd
explaining the interface

@@
 To explore the local explanations, an ensemble of plots [@unwin_ensemble_2018] is provided, called the _cheem viewer_.
-There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
+There are two primary plots: the **global view** to give the context of all of the SHAP values, and the **radial tour view** to explore the local explanations with user-controlled rotation.
 In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
+There are different plots used for categorical and quantitative response.
 Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
@@
 
-While the comparison of these spaces is interesting, the main purpose of the global view is to enable the selection of instances to explore the local explanations.
+While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances to explore the local explanations.
 The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
@@
 
-XXX Re-order these paragraphs
+<!-- Local explanations as par coords -->
+The local explanations for all observations are normalized (squared sum of values adds to 1) and thus the relative importance of features can be compared across all instances.
+These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
+\@ref(fig:classificationcase) and \@ref(fig:regressioncase) d).
+The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
+From this plot, we would read the range of importances across all instances can be interpreted.
+For classification, one would look at differences between groups on any feature, for example, Fig.
+\@ref(fig:classificationcase) d suggests that `b_l` is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in \@ref(fig:regressioncase) d, and which have a range of high and low values (e.g.
+`off`, `def`) suggesting important for some instances and not important for other instances.
+
 <!-- Segue of PI to attribution projection -->
-The attribution projection of the PI (selected from the global view) is the initial 1D basis in a radial tour (Fig.
-\@ref(fig:classificationcase), e).
-This is an  approximation of the feature importance, for the prediction of this instance.
+The overlaid bars on the parallel coordinate plot represent the attribution projection of the PI.
+(Remember that the PI is interactively selected from the global view).
+The attribution projection is an  approximation of the feature importance, for the prediction of this instance.
 This combination of features best explains the difference between the mean response and an instance’s predicted value.
-It is not an indication of the local shape of the model surface, that is, some indication of the tangent to the curve at this point.
-
-<!-- Tour start frame -->
-The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively, overlaid on attribution projections from all instances.
-These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
-\@ref(fig:classificationcase), d).
-This is an important global context, the summarization of the attribution of all instances.
+It is not an indication of the local shape of the model surface, that is, it is not some indication of the tangent to the curve at this point.
 
 <!-- Tour animation -->
-The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal positions.
-The radial tour will vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
-Doing so tests a feature's sensitivity to the structure identified by the local explanation.
-The default feature selected has the largest discrepancy between the attribution of the PI and CI.
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Fig.
+\@ref(fig:classificationcase), e), and as scatterplots for a quantitative response ((Fig.
+\@ref(fig:regressioncase), e).
+The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The user uses the radial tour to vary contribution of the selected feature between 0-1.
+Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
+For classification, if the separation between classes diminishes when the feature contribution is reduced then this suggests that the feature is important for the class separation.
+For regression, if the relationship scatterplot weakens when the feature contribution is reduced then this suggests that the feature is important for accurately predicting the response.
+
+<!-- The default feature selected has the largest discrepancy between the attribution of the PI and CI.
 The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizontal position of Fig.
 \@ref(fig:classificationcase) e and Fig.
-\@ref(fig:regressioncase) d.
-
-XXX Wrong place for this The following sections elaborate on the difference in displays from applying this approach in classification and regression tasks.
-Now that we have introduced the global view and corresponding radial tour, let us discuss the differences between the classification and regression cases.
+\@ref(fig:regressioncase) d.-->
 

### 2022-02-04 _paper.tex
explaining the interface

@@
 Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged''.
+These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 To explore the local explanations, an ensemble of plots (Unwin and Valero-Mora 2018) is provided, called the \emph{cheem viewer}.
-There are two primary plots: the global view to give the context of all of the SHAP values, and the radial tour view to explore the local explanations with user-controlled rotation.
+There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values, and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
 In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
+There are different plots used for categorical and quantitative response.
 Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
@@
 
-XXX Re-order these paragraphs
-The attribution projection of the PI (selected from the global view) is the initial 1D basis in a radial tour (Fig.
-\ref{fig:classificationcase}, e).
-This is an approximation of the feature importance, for the prediction of this instance.
+The local explanations for all observations are normalized (squared sum of values adds to 1) and thus the relative importance of features can be compared across all instances.
+These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
+\ref{fig:classificationcase} and \ref{fig:regressioncase} d).
+The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
+From this plot, we would read the range of importances across all instances can be interpreted.
+For classification, one would look at differences between groups on any feature, for example, Fig.
+\ref{fig:classificationcase} d suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in \ref{fig:regressioncase} d, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
+
+The overlaid bars on the parallel coordinate plot represent the attribution projection of the PI.
+(Remember that the PI is interactively selected from the global view).
+The attribution projection is an approximation of the feature importance, for the prediction of this instance.
 This combination of features best explains the difference between the mean response and an instance's predicted value.
-It is not an indication of the local shape of the model surface, that is, some indication of the tangent to the curve at this point.
+It is not an indication of the local shape of the model surface, that is, it is not some indication of the tangent to the curve at this point.
 
-The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively, overlaid on attribution projections from all instances.
-These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
-\ref{fig:classificationcase}, d).
-This is an important global context, the summarization of the attribution of all instances.
-
-The current projection basis is depicted as the width of a bar, the feature contributions to the horizontal positions.
-The radial tour will vary one selected feature to a full contribution, zero contribution, and back to the attribution projection.
-Doing so tests a feature's sensitivity to the structure identified by the local explanation.
-The default feature selected has the largest discrepancy between the attribution of the PI and CI.
-The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizontal position of Fig.
-\ref{fig:classificationcase} e and Fig.
-\ref{fig:regressioncase} d.
-
-XXX Wrong place for this The following sections elaborate on the difference in displays from applying this approach in classification and regression tasks.
-Now that we have introduced the global view and corresponding radial tour, let us discuss the differences between the classification and regression cases.
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Fig.
+\ref{fig:classificationcase}, e), and as scatterplots for a quantitative response ((Fig.
+\ref{fig:regressioncase}, e).
+The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The user uses the radial tour to vary contribution of the selected feature between 0-1.
+Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
+For classification, if the separation between classes diminishes when the feature contribution is reduced then this suggests that the feature is important for the class separation.
+For regression, if the relationship scatterplot weakens when the feature contribution is reduced then this suggests that the feature is important for accurately predicting the response.
 
@@
 
-\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
 Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
-\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-unwin_ensemble_2018}{}}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
-\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
 Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.

### 2022-02-04 _paper.rmd
classification task

@@
 
-Typically we select a misclassified instance compared to a correctly classified point nearby in data space.
-The model information in the global view depicts the model confusion matrix.
-The radial tour is 1D, with a density display, while the goal of the tour; exploring the sensitivity of each feature to structure identified by the local explanation, evaluating the support or robustness of the prediction.
+Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
+The global view (right plot) displays the model confusion matrix.
+The radial tour is 1D, plotted as a density where colour indicates class.
+A scrollbar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
 

### 2022-02-04 _paper.rmd
regression task

@@
 
-Structure resolved in the attribution space can be highlighted by coloring points on a statistic.
-For this purpose, we include residuals, log Mahalanobis distance of data space (a measure of outlyingness), and the correlation of the attribution projection with the observed response.
-In the radial tour, the horizontal positions are the same, the basis projection of the radial tour.
-The vertical position is fixed to the observed response feature and residuals in the middle and right panels.
-Correspondingly, the display changes from univariate density to 2D scatterplot.
-The basis is still one component (horizontal) independent of the vertical position.
+Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
+The global view (right plot) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
+
+In the radial tour view, the observed response and the residuals (vertical) are plotted against the attribution projection of the PI (horizontal).
+The attribution projection can be interpreted similarly to the predicted value from the global view plot.
+It represents a linear combination of the features, and a good fit would be indicated when there is a strong relationship seen with the observed values.
+This can be viewed as a local linear approximation if the fitted model is nonlinear.
+As the contribution of a feature is varied, if the value of the PI doesn't change much it would indicate that the prediction for this instance is NOT sensitive to that feature.
+Conversely, if the predicted value varies substantially, the prediction is very sensitive to that feature, suggesting that the feature is very important for the PI's prediction.
 

### 2022-02-04 _paper.rmd
-regression use case: fix rotation of basis. -screen captures relabelled and text adjusted.

@@
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
-\@ref(fig:classificationcase) and \@ref(fig:regressioncase) d).
+\@ref(fig:classificationcase) e and \@ref(fig:regressioncase) e).
 The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
@@
 For classification, one would look at differences between groups on any feature, for example, Fig.
-\@ref(fig:classificationcase) d suggests that `b_l` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in \@ref(fig:regressioncase) d, and which have a range of high and low values (e.g.
+\@ref(fig:classificationcase) e suggests that `b_l` is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in \@ref(fig:regressioncase) e, and which have a range of high and low values (e.g.
 `off`, `def`) suggesting important for some instances and not important for other instances.
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Fig.
-\@ref(fig:classificationcase), e), and as scatterplots for a quantitative response ((Fig.
-\@ref(fig:regressioncase), e).
+\@ref(fig:classificationcase), f), and as scatterplots for a quantitative response (Fig.
+\@ref(fig:regressioncase), f).
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
@@
 For regression, if the relationship scatterplot weakens when the feature contribution is reduced then this suggests that the feature is important for accurately predicting the response.
-
 <!-- The default feature selected has the largest discrepancy between the attribution of the PI and CI.
 The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizontal position of Fig.
-\@ref(fig:classificationcase) e and Fig.
-\@ref(fig:regressioncase) d.-->
+\@ref(fig:classificationcase) f and Fig.
+\@ref(fig:regressioncase) f.-->
 
@@
 Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
-The global view (right plot) displays the model confusion matrix.
-The radial tour is 1D, plotted as a density where colour indicates class.
-A scrollbar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
+The global view (Fig.
+\@ref(fig:classificationcase) c) displays the model confusion matrix.
+The radial tour is 1D, plotted as a density where color indicates class.
+A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
 
@@
 Global view inputs, (a), set the PI, CI, and color statistic.
-Global view, (b) PC1 by PC2 for data space, attribution space, and prediction by observed y (visual of the confusion matrix for classification).
+Global view, (b) PC1 by PC2 approximations of the data space and attribution space.
+(c) prediction by observed y (visual of the confusion matrix for classification tasks).
 Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (c) select features to include and which feature is changed in the tour.
-The visual depicting 1D basis (d) shows the distribution of the feature attribution, and bars show the current basis.
-The black bar is being varied.
-Panel (e) is the resulting projection of the data indicated as density in the classification case."}
+Radial tour inputs (d) select features to include and which feature is changed in the tour.
+(e) shows parallel coordinate display of the distribution of the feature attributions while bars dipict contribution for the current basis.
+The black bar is the variable being changed in the radial tour.
+Panel (f) is the resulting projection of the data indicated as density in the classification case."}
 knitr::include_graphics("./figures/app_classification.PNG")
@@
 Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
-The global view (right plot) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The global view (Fig.
+\@ref(fig:regressioncase) a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
 The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
@@
 Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
-Coloring on a statistic (b) highlights structure organized in the attribution space.
+Applied linked brushing, (b), selected points are highlighted in the component spaces and tabularly displayed.
+Coloring on a statistic (c) highlights structure organized in the attribution space.
 Interactive tabular display (d) populates when instances are selected.
-Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right)."}
+Contribution of the 1D basis effecting the horizontal position (e) parallel coorindate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right)."}
 knitr::include_graphics("./figures/app_regression_interactions.PNG")
@@
 While offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
+As the contribution of defensive skills increases, Messi's is no longer separated from the group, and the defensive players are better predicted in this attribution case.
+Essentially, the further right an observation is the better the the basis does at describing its predicted value.
 In terms of what-if analysis, his predicted wages would be halved if Messi's tree SHAP attributions at these levels.

### 2022-02-04 _paper.tex
-regression use case: fix rotation of basis. -screen captures relabelled and text adjusted.

@@
 Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged''.
-These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
-\ref{fig:classificationcase} and \ref{fig:regressioncase} d).
+\ref{fig:classificationcase} e and \ref{fig:regressioncase} e).
 The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
@@
 For classification, one would look at differences between groups on any feature, for example, Fig.
-\ref{fig:classificationcase} d suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in \ref{fig:regressioncase} d, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
+\ref{fig:classificationcase} e suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in \ref{fig:regressioncase} e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
 
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Fig.
-\ref{fig:classificationcase}, e), and as scatterplots for a quantitative response ((Fig.
-\ref{fig:regressioncase}, e).
+\ref{fig:classificationcase}, f), and as scatterplots for a quantitative response (Fig.
+\ref{fig:regressioncase}, f).
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
@@
 Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
-The global view (right plot) displays the model confusion matrix.
-The radial tour is 1D, plotted as a density where colour indicates class.
-A scrollbar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
+The global view (Fig.
+\ref{fig:classificationcase} c) displays the model confusion matrix.
+The radial tour is 1D, plotted as a density where color indicates class.
+A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
 
@@
 Global view inputs, (a), set the PI, CI, and color statistic.
-Global view, (b) PC1 by PC2 for data space, attribution space, and prediction by observed y (visual of the confusion matrix for classification).
+Global view, (b) PC1 by PC2 approximations of the data space and attribution space.
+(c) prediction by observed y (visual of the confusion matrix for classification tasks).
 Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (c) select features to include and which feature is changed in the tour.
-The visual depicting 1D basis (d) shows the distribution of the feature attribution, and bars show the current basis.
-The black bar is being varied.
-Panel (e) is the resulting projection of the data indicated as density in the classification case.}\label{fig:classificationcase}
+Radial tour inputs (d) select features to include and which feature is changed in the tour.
+(e) shows parallel coordinate display of the distribution of the feature attributions while bars dipict contribution for the current basis.
+The black bar is the variable being changed in the radial tour.
+Panel (f) is the resulting projection of the data indicated as density in the classification case.}\label{fig:classificationcase}
 \end{figure}
@@
 Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
-The global view (right plot) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The global view (Fig.
+\ref{fig:regressioncase} a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
 The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
@@
 Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
-Coloring on a statistic (b) highlights structure organized in the attribution space.
+Applied linked brushing, (b), selected points are highlighted in the component spaces and tabularly displayed.
+Coloring on a statistic (c) highlights structure organized in the attribution space.
 Interactive tabular display (d) populates when instances are selected.
-Regression projection (e) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right).}\label{fig:regressioncase}
+Contribution of the 1D basis effecting the horizontal position (e) parallel coorindate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right).}\label{fig:regressioncase}
 \end{figure}
@@
 While offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group, and other defensive players are better predicted in this attribution case.
+As the contribution of defensive skills increases, Messi's is no longer separated from the group, and the defensive players are better predicted in this attribution case.
+Essentially, the further right an observation is the better the the basis does at describing its predicted value.
 In terms of what-if analysis, his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
+\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
 Gorman, Kristen B., Tony D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
+\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
 Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
-\leavevmode\vadjust pre{\hypertarget{ref-unwin_ensemble_2018}{}}%
+\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
+\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
 Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.

### 2022-02-04 figures/_make_figures_case_study.r

@@ -376,7 +376,7 @@ THIS_REG_radial_cheem_tour  <- function(
     primary_obs = prim_obs, comparison_obs = comp_obs,
     do_add_pcp_segments = FALSE, inc_var_nms = .inc_var_nms,
     pcp_shape = 124, angle = 0) + .t
-  .pw <- .ggt1 / .ggt2
+  .pw <- .ggt1 + .ggt2
   .cp <- cowplot::plot_grid(
     .glob_view, .pw, labels = c("a", "b"),
     rel_heights = c(1, 1.2), ncol = 1)
@@ -430,7 +430,7 @@ message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")
     primary_obs = prim_obs, comparison_obs = comp_obs,
     do_add_pcp_segments = FALSE, inc_var_nms = .inc_var_nms,
     pcp_shape = 124, angle = 0) + .t
-  .pw <- .ggt1 / .ggt2
+  .pw <- .ggt1 + .ggt2
   .cp <- cowplot::plot_grid(
     .glob_view, .pw, labels = c("a", "b"),
     rel_heights = c(1, 1.2), ncol = 1)

### 2022-02-04 _paper.rmd
extended conclusion.

@@
 ---
-title: "Exploring Local Explanations of Non-linear Models Using Animated Linear Projections"
+title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
 shorttitle: Cheem
@@
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate non-linear models.
+While predictive modeling may opt for more accurate nonlinear models.
 The use of black-box models is becoming increasingly common, but not without their share of controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
@@
 Another concern is data drift, a shift in support or domain of the explanatory features (variable or predictors).
-Non-linear models are typically more sensitive and do not extrapolate well outside the training data domain.
+Nonlinear models are typically more sensitive and do not extrapolate well outside the training data domain.
 Better interpretability of the model means more transparency where models' predictions may be plausible or completely unreliable.
@@
 <!-- Reminder of local explanation -->
-Consider a highly non-linear model.
+Consider a highly nonlinear model.
 It can be hard to determine whether small changes in a feature's value will make a class prediction change group or identify which features contribute to an extreme residual.
@@
 To explore the local explanations, an ensemble of plots [@unwin_ensemble_2018] is provided, called the _cheem viewer_.
-There are two primary plots: the **global view** to give the context of all of the SHAP values, and the **radial tour view** to explore the local explanations with user-controlled rotation.
+There are two primary plots: the __global view__ to give the context of all of the SHAP values, and the __radial tour view__ to explore the local explanations with user-controlled rotation.
 In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
@@
 
-We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generally used with any compatible model-explanation pairing.
-We apply it to the classification and regression tasks.
+We have illustrated this method on random forest models using the tree SHAP local explanation.
+While this analysis could be generally used with any compatible model-explanation pairing.
 We have created an open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem), to facilitate preprocessing and exploration with the described interactive application.
 Toy and real data are provided, or upload your data after preprocessing.
+
+One limitation of the __cheem__ packages is that it currently only calculates tree SHAP for tree-based models supported by __treeshap__ (tree based models from __randomForest__, __ranger__, __gbm__, __xgboost__, __lightgbm__, or __catboost__).
+The SHAP and oscillation explainations could be added with use of the `DALEX::explain()` seems to be an excellent direction to extend [@biecek_dalex_2018; @biecek_explanatory_2021].
+Tree SHAP was selected because of its reduced computational complexity; processing runtime of the explanation will be a looming concern.
+
+Attribution space does tend to be more separable than data space at least in the first couple principal components.
+Different statistics should be explored to convey a clearer understanding what structure the explanation distinguishes.
+It is curious that the attribution space the same dimensionality of the data with a seemingly better separation.
+It sounds like a candidate to fit another model.
+However, such an attempt should want to be vigilant not to over-fit the training data.
 

### 2022-02-04 _paper.tex
extended conclusion.

@@
 \hypersetup{
-  pdftitle={Exploring Local Explanations of Non-linear Models Using Animated Linear Projections},
+  pdftitle={Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections},
   pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
@@
 
-\title{Exploring Local Explanations of Non-linear Models Using Animated Linear Projections}
+\title{Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections}
 \author{}
@@
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate non-linear models.
+While predictive modeling may opt for more accurate nonlinear models.
 The use of black-box models is becoming increasingly common, but not without their share of controversy (O'Neil 2016; Kodiyan 2019).
@@
 
-Consider a highly non-linear model.
+Consider a highly nonlinear model.
 It can be hard to determine whether small changes in a feature's value will make a class prediction change group or identify which features contribute to an extreme residual.
@@
 
-We have illustrated this method on random forest models using the tree SHAP local explanation, while it could be generally used with any compatible model-explanation pairing.
-We apply it to the classification and regression tasks.
+We have illustrated this method on random forest models using the tree SHAP local explanation.
+While this analysis could be generally used with any compatible model-explanation pairing.
 We have created an open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate preprocessing and exploration with the described interactive application.
 Toy and real data are provided, or upload your data after preprocessing.
+
+One limitation of the \textbf{cheem} packages is that it currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{randomForest}, \textbf{ranger}, \textbf{gbm}, \textbf{xgboost}, \textbf{lightgbm}, or \textbf{catboost}).
+The SHAP and oscillation explainations could be added with use of the \texttt{DALEX::explain()} seems to be an excellent direction to extend (Biecek 2018; Biecek and Burzykowski 2021).
+Tree SHAP was selected because of its reduced computational complexity; processing runtime of the explanation will be a looming concern.
+
+Attribution space does tend to be more separable than data space at least in the first couple principal components.
+Different statistics should be explored to convey a clearer understanding what structure the explanation distinguishes.
+It is curious that the attribution space the same dimensionality of the data with a seemingly better separation.
+It sounds like a candidate to fit another model.
+However, such an attempt should want to be vigilant not to over-fit the training data.
 

### 2022-02-06 _paper.rmd
detail to fifa and chocolates inverse case.

@@
 
-(ref:citeRf) @liaw_classification_2002
-(ref:citeTs) @kominsarczyk_treeshap_2021
+(ref:citeRf) [@liaw_classification_2002]
+(ref:citeTs) [@kominsarczyk_treeshap_2021]
 \begin{itemize}
@@
 	\item \textbf{Model and explanation:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [(ref:citeRf)], compatibility tree SHAP.
+Currently, we apply random forest models via the package \textbf{randomForest} (ref:citeRf), compatibility tree SHAP.
 We use modest hyperparameters, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
@@
 Its reduced computational complexity makes tree SHAP a good candidate to start with.
-Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+(Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
 
@@
 It is compared with a CI that is a correctly classified dark chocolate (instance 7).
-The PCA plot, and the SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
+The PCA plot, and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
 The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
@@
 
-It would also be interesting to explore the inverse case; which features lead to a mild chocolate being misclassified as dark.
+<!-- segue and selection for chocolates inverse case -->
+It would also be interesting to explore the inverse case.
+This would describes which features lead to a milk chocolate being misclassified as dark and how those attributions differ from the previous misclassification.
 Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
 This exploration is shown in Figure \@ref(fig:casechocolatesinverse).
-From the parallel coordinate lines, we identify discrepancies in Sodium and Fiber.
-We opt to vary Sodium, and find the attribution of the other milk chocolates not to support the prediction of a dark chocolate.
+
+<!-- treeSHAP pca, PCP, tour -->
+The difference of position in the tree SHAP PCA with the previous case is quite large; this gives an approximate feel that the attribution should be quite different.
+Looking to the initial contribution we find this to be the case.
+Previously fiber was very important while it is absent from the attribution in this case.
+Conversely, calories from fat and total fat are highly attributed here, while unimportant in the preceding case.
+
+Comparing the attribution with the CI (dotted line) discrepancies in Sodium and Fiber are identified.
+We opt to vary Sodium.
+Even in the initial projection, the instance looks slightly more like its observered milk than predicted dark chocolate.
+The misclassification seem least supported when the basis reaches sodium attribution of typical dark chocolate.
 
@@
 The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
-After aggregation of the skill measurements, we regress player wages [2020 euros] given just the skill aggregates.
-The model was fit from 5000 instances of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
+From a plot of the correlation matrix, nine higher level skill groupings were identified and aggregated.
+A random forest model if fit from these aggregations and regress player wages [2020 euros].
+The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
+Continuing from the exploration in section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field position of players.
 We compare a leading offensive fielder (L.
@@
 At this point, defensive players have been rotated to the highest horizontal value.
-The animate radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).
+The animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
@@
 
-With figure \@ref(fig:casefifa), we will test the premise of the local explanation.
-While offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group, and the defensive players are better predicted in this attribution case.
-Essentially, the further right an observation is the better the the basis does at describing its predicted value.
-In terms of what-if analysis, his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
+With figure \@ref(fig:casefifa), we will test the sopport of the local explanation.
+Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
+If either of them are rotated out the other will be rotated into frame maintaining a far-right position.
+However, when we vary the defensive skills, the other skills are rotated out of the frame.
+
+As the contribution of defensive skills increases, Messi's is no longer separated from the group.
+Players with high values in defensive skills are now the right most points.
+In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
 

### 2022-02-06 _paper.tex
detail to fifa and chocolates inverse case.

@@
     \item \textbf{Model and explanation:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} [Liaw and Wiener (2002)], compatibility tree SHAP.
+Currently, we apply random forest models via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
 We use modest hyperparameters, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} Kominsarczyk et al.
-(2021).
+Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (Kominsarczyk et al.
+2021).
 This implementation aggregates over exhaustively over all trees' attribution, and we opt not to fit interactions of features.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP a good candidate to start with.
-Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.
+Its reduced computational complexity makes tree SHAP a good candidate to start with.\footnote{Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.}
 
@@
 
-It would also be interesting to explore the inverse case; which features lead to a mild chocolate being misclassified as dark.
+It would also be interesting to explore the inverse case.
+This would describes which features lead to a milk chocolate being misclassified as dark and how those attributions differ from the previous misclassification.
 Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
 This exploration is shown in Figure \ref{fig:casechocolatesinverse}.
-From the parallel coordinate lines, we identify discrepancies in Sodium and Fiber.
-We opt to vary Sodium, and find the attribution of the other milk chocolates not to support the prediction of a dark chocolate.
+
+The difference of position in the tree SHAP PCA with the previous case is quite large; this gives an approximate feel that the attribution should be quite different.
+Looking to the initial contribution we find this to be the case.
+Previously fiber was very important while it is absent from the attribution in this case.
+Conversely, calories from fat and total fat are highly attributed here, while unimportant in the preceding case.
+
+Comparing the attribution with the CI (dotted line) discrepancies in Sodium and Fiber are identified.
+We opt to vary Sodium.
+Even in the initial projection, the instance looks slightly more like its observered milk than predicted dark chocolate.
+The misclassification seem least supported when the basis reaches sodium attribution of typical dark chocolate.
 
@@
 The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-After aggregation of the skill measurements, we regress player wages {[}2020 euros{]} given just the skill aggregates.
-The model was fit from 5000 instances of the nine skill aggregates before being thinned to 500 players to mitigate occlusion and render time.
+From a plot of the correlation matrix, nine higher level skill groupings were identified and aggregated.
+A random forest model if fit from these aggregations and regress player wages {[}2020 euros{]}.
+The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
+Continuing from the exploration in section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field position of players.
 We compare a leading offensive fielder (L.
@@
 At this point, defensive players have been rotated to the highest horizontal value.
-The animate radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.}\label{fig:casefifa}
+The animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.}\label{fig:casefifa}
 \end{figure}
 
-With figure \ref{fig:casefifa}, we will test the premise of the local explanation.
-While offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group, and the defensive players are better predicted in this attribution case.
-Essentially, the further right an observation is the better the the basis does at describing its predicted value.
-In terms of what-if analysis, his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
+With figure \ref{fig:casefifa}, we will test the sopport of the local explanation.
+Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
+If either of them are rotated out the other will be rotated into frame maintaining a far-right position.
+However, when we vary the defensive skills, the other skills are rotated out of the frame.
+
+As the contribution of defensive skills increases, Messi's is no longer separated from the group.
+Players with high values in defensive skills are now the right most points.
+In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
 

### 2022-02-06 _paper.rmd
replaced Fig. with Figure and mapped out conclusion

@@
 There are different plots used for categorical and quantitative response.
-Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
+Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 <!-- Approximations of the spaces, PC1:2 -->
-A visualization of these spaces is provided by the first two principal components of their respective spaces.
-A single 2D projection will not encompass all of the structure of higher-dimensional space, but can still be a useful visual summary.
-In addition, a plot of the observed against predicted response values is also provided (Fig.
-\@ref(fig:classificationcase), b), to help identify instances poorly predicted by the model.
+A visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
+These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are useful visual summaries.
+In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:classificationcase) XXX), to help identify instances poorly predicted by the model.
 <!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Fig.
-\@ref(fig:classificationcase) a and c).
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \@ref(fig:classificationcase)a,c).
 
@@
 The local explanations for all observations are normalized (squared sum of values adds to 1) and thus the relative importance of features can be compared across all instances.
-These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
-\@ref(fig:classificationcase) e and \@ref(fig:regressioncase) e).
+These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
 The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
 From this plot, we would read the range of importances across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature, for example, Fig.
-\@ref(fig:classificationcase) e suggests that `b_l` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in \@ref(fig:regressioncase) e, and which have a range of high and low values (e.g.
+For classification, one would look at differences between groups on any feature, for example, Figure \@ref(fig:classificationcase)e suggests that `b_l` is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, and which have a range of high and low values (e.g.
 `off`, `def`) suggesting important for some instances and not important for other instances.
@@
 <!-- Tour animation -->
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Fig.
-\@ref(fig:classificationcase), f), and as scatterplots for a quantitative response (Fig.
-\@ref(fig:regressioncase), f).
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \@ref(fig:classificationcase)f), and as scatterplots for a quantitative response (Figure \@ref(fig:regressioncase)f).
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
@@
 Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
-The global view (Fig.
-\@ref(fig:classificationcase) c) displays the model confusion matrix.
+The global view (Figure \@ref(fig:classificationcase)c) displays the model confusion matrix.
 The radial tour is 1D, plotted as a density where color indicates class.
@@
 Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
-The global view (Fig.
-\@ref(fig:regressioncase) a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The global view (Figure \@ref(fig:regressioncase)a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
 The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
@@
 
+<!-- summarise methods -->
+<!-- summarise findings -->
+<!-- suggest next steps -->>
+
 The need to maintain the interpretability of black-box models is evident.

### 2022-02-06 _paper.tex
replaced Fig. with Figure and mapped out conclusion

@@
 There are different plots used for categorical and quantitative response.
-Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} contain screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
+Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 
-A visualization of these spaces is provided by the first two principal components of their respective spaces.
-A single 2D projection will not encompass all of the structure of higher-dimensional space, but can still be a useful visual summary.
-In addition, a plot of the observed against predicted response values is also provided (Fig.
-\ref{fig:classificationcase}, b), to help identify instances poorly predicted by the model.
+A visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
+These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are useful visual summaries.
+In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:classificationcase} XXX), to help identify instances poorly predicted by the model.
 For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Fig.
-\ref{fig:classificationcase} a and c).
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:classificationcase}a,c).
 
@@
 The local explanations for all observations are normalized (squared sum of values adds to 1) and thus the relative importance of features can be compared across all instances.
-These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Fig.
-\ref{fig:classificationcase} e and \ref{fig:regressioncase} e).
+These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
 The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
 From this plot, we would read the range of importances across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature, for example, Fig.
-\ref{fig:classificationcase} e suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in \ref{fig:regressioncase} e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
+For classification, one would look at differences between groups on any feature, for example, Figure \ref{fig:classificationcase}e suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
 
@@
 
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Fig.
-\ref{fig:classificationcase}, f), and as scatterplots for a quantitative response (Fig.
-\ref{fig:regressioncase}, f).
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f), and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
@@
 Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
-The global view (Fig.
-\ref{fig:classificationcase} c) displays the model confusion matrix.
+The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
 The radial tour is 1D, plotted as a density where color indicates class.
@@
 Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
-The global view (Fig.
-\ref{fig:regressioncase} a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
 The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP a good candidate to start with.\footnote{Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.}
+Its reduced computational complexity makes tree SHAP a good candidate to start with.
+(Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
 
@@
 
+\begin{quote}
+\end{quote}
+
 The need to maintain the interpretability of black-box models is evident.

### 2022-02-06 _paper.rmd
figure 3, labeling consistency, caption touch up

--- 2022-02-06 _paper.rmd
@@
 ```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression task highlighting the differences from the classification task and interactive features.
-Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
-Applied linked brushing, (b), selected points are highlighted in the component spaces and tabularly displayed.
+Panel (a) PCA of the data and attributions spaces, (b), residual plot, predictions by observed values.
+Four points areselected points and highlighted in the PC spaces and tabularly displayed.
 Coloring on a statistic (c) highlights structure organized in the attribution space.

### 2022-02-06 _paper.tex
figure 3, labeling consistency, caption touch up

--- 2022-02-06 _paper.tex
@@
 \caption{Overview of the cheem viewer for regression task highlighting the differences from the classification task and interactive features.
-Panel (a) shows linked brushing across the global view and the tooltip display when the cursor hovers over an instance.
-Applied linked brushing, (b), selected points are highlighted in the component spaces and tabularly displayed.
+Panel (a) PCA of the data and attributions spaces, (b), residual plot, predictions by observed values.
+Four points areselected points and highlighted in the PC spaces and tabularly displayed.
 Coloring on a statistic (c) highlights structure organized in the attribution space.

### 2022-02-06 _paper.rmd
installation

--- 2022-02-06 _paper.rmd
@@
 
-The following __R__ code will help with getting up and running:
+The package can be installed from GitHub using the following __R__ code:
 
 ```{r eval=FALSE, echo=TRUE}
-## Download the package
-install.packages("cheem", dependencies = TRUE)
-## Restart the R session, so the IDE has the correct directory structure
-restartSession()
-## Load cheem into session
+# install.packages("remotes") # if you don't have it
+remotes::install.packages("cheem", dependencies = TRUE)
 library("cheem")
-## Try the app
 run_app()
+```
 
-# Processing your data
-## Install treeshap from github to use as a local explainer
+To process your own data, you will need to use the `treeshap` package, which can be installed from GitHub:
+
+```{r eval=FALSE, echo=TRUE}
 remotes::install_github('ModelOriented/treeshap')
-## Follow the examples in cheem_ls()
-?cheem_ls
 ```
+
+and follow the examples provided with the package to compute the local explainers. XXX Add some information on what type of data object gets passed into cheem.
+
+Alternatively, the cheem viewer shiny app can be directly accessed at https://ebsmonash.shinyapps.io/cheem_initial/.
 

### 2022-02-06 _paper.tex
installation

--- 2022-02-06 _paper.tex
@@
 
-The following \textbf{R} code will help with getting up and running:
+The package can be installed from GitHub using the following \textbf{R} code:
 
@@
 \begin{Highlighting}[]
-\DocumentationTok{\#\# Download the package}
-\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
-\DocumentationTok{\#\# Restart the R session, so the IDE has the correct directory structure}
-\FunctionTok{restartSession}\NormalTok{()}
-\DocumentationTok{\#\# Load cheem into session}
+\CommentTok{\# install.packages("remotes") \# if you don\textquotesingle{}t have it}
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
 \FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
-\DocumentationTok{\#\# Try the app}
 \FunctionTok{run\_app}\NormalTok{()}
-
-\CommentTok{\# Processing your data}
-\DocumentationTok{\#\# Install treeshap from github to use as a local explainer}
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
-\DocumentationTok{\#\# Follow the examples in cheem\_ls()}
-\NormalTok{?cheem\_ls}
 \end{Highlighting}
 \end{Shaded}
+
+To process your own data, you will need to use the \texttt{treeshap} package, which can be installed from GitHub:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+and follow the examples provided with the package to compute the local explainers. XXX Add some information on what type of data object gets passed into cheem.
+
+Alternatively, the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
 

### 2022-02-06 _paper.rmd
installation

@@
 
-The package can be installed from GitHub using the following __R__ code:
+Version 0.2.0 of the package can be installed from CRAN using the following __R__ code:
 
 ```{r eval=FALSE, echo=TRUE}
-# install.packages("remotes") # if you don't have it
-remotes::install.packages("cheem", dependencies = TRUE)
+install.packages("cheem", dependencies = TRUE)
 library("cheem")
@@
 ```{r eval=FALSE, echo=TRUE}
+# install.packages("remotes") # if you don't have it
 remotes::install_github('ModelOriented/treeshap')

### 2022-02-06 _paper.tex
installation

@@
 
-The package can be installed from GitHub using the following \textbf{R} code:
+Version 0.2.0 of the package can be installed from CRAN using the following \textbf{R} code:
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# install.packages("remotes") \# if you don\textquotesingle{}t have it}
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
+\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
 \FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
@@
 \begin{Highlighting}[]
+\CommentTok{\# install.packages("remotes") \# if you don\textquotesingle{}t have it}
 \NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}

### 2022-02-06 _paper.rmd
installation cleanup

--- 2022-02-06 _paper.rmd
@@
 ```{r eval=FALSE, echo=TRUE}
-# install.packages("remotes") # if you don't have it
+# Install remotes if absent
+if(require("remotes") == FALSE) install.packages("remotes")
 remotes::install.packages("cheem", dependencies = TRUE)
@@
 
-and follow the examples provided with the package to compute the local explainers. XXX Add some information on what type of data object gets passed into cheem.
-
+Follow the examples provided with the package to compute the local explainers, (see `?cheem_ls`).
+The application expects the return of call from `cheem_ls()` which is then saved to an .rds file with `saveRDS()`.
 Alternatively, the cheem viewer shiny app can be directly accessed at https://ebsmonash.shinyapps.io/cheem_initial/.

### 2022-02-06 _paper.tex
installation cleanup

--- 2022-02-06 _paper.tex
@@
 \begin{Highlighting}[]
-\DocumentationTok{\#\# Download the package}
-\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
-\DocumentationTok{\#\# Restart the R session, so the IDE has the correct directory structure}
-\FunctionTok{restartSession}\NormalTok{()}
-\DocumentationTok{\#\# Load cheem into session}
+\CommentTok{\# Install remotes if absent}
+\ControlFlowTok{if}\NormalTok{(}\FunctionTok{require}\NormalTok{(}\StringTok{"remotes"}\NormalTok{) }\SpecialCharTok{==} \ConstantTok{FALSE}\NormalTok{) }\FunctionTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{)}
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
 \FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
-\DocumentationTok{\#\# Try the app}
 \FunctionTok{run\_app}\NormalTok{()}
-
-\CommentTok{\# Processing your data}
-\DocumentationTok{\#\# Install treeshap from github to use as a local explainer}
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
-\DocumentationTok{\#\# Follow the examples in cheem\_ls()}
-\NormalTok{?cheem\_ls}
 \end{Highlighting}
 \end{Shaded}
+
+To process your own data, you will need to use the \texttt{treeshap} package, which can be installed from GitHub:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Follow the examples provided with the package to compute the local explainers, (see \texttt{?cheem\_ls}).
+The application expects the return of call from \texttt{cheem\_ls()} which is then saved to an .rds file with \texttt{saveRDS()}.
+Alternatively, the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
 
@@
 \section{Discussion}\label{sec:cheemdiscussion}}
+
+\begin{quote}
+\end{quote}
 

### 2022-02-06 _paper.rmd
links to vimeo movies are all the same now

--- 2022-02-06 _paper.rmd
@@
 The misclassification seems improbable when sugar is near the max contribution.
-Animated tour can be found at [vimeo.com/666431143](https://vimeo.com/666431143).
+(A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
 ```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
@@
 This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
-Animated tour can be found at [vimeo.com/666431143](https://vimeo.com/666431143).
+(A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
 ```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse)"}
@@
 At this point, defensive players have been rotated to the highest horizontal value.
-The animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).
+(A video of the animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).)
 ```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
@@
 Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of `LtA`, which is important to explaining the PI and near invariant to the sales price of the CI.
-The corresponding animation is at [vimeo.com/666431134](https://vimeo.com/666431134).
+(A video showing the animation is at [vimeo.com/666431134](https://vimeo.com/666431134).)
 ```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}

### 2022-02-06 _paper.tex
links to vimeo movies are all the same now

--- 2022-02-06 _paper.tex
@@
 The misclassification seems improbable when sugar is near the max contribution.
-Animated tour can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.}\label{fig:casechocolates}
+(A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolates}
 \end{figure}
@@
 This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
-Animated tour can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.}\label{fig:casechocolatesinverse}
+(A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolatesinverse}
 \end{figure}
@@
 At this point, defensive players have been rotated to the highest horizontal value.
-The animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.}\label{fig:casefifa}
+(A video of the animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.)}\label{fig:casefifa}
 \end{figure}
@@
 Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of \texttt{LtA}, which is important to explaining the PI and near invariant to the sales price of the CI.
-The corresponding animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.}\label{fig:caseames}
+(A video showing the animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.)}\label{fig:caseames}
 \end{figure}

### 2022-02-06 _paper.rmd

--- 2022-02-06 _paper.rmd
@@
 This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
-Animated tour can be found at [vimeo.com/666431143](https://vimeo.com/666431143).
+(A video of the tour animation can be found at [vimeo.com/666431148](https://vimeo.com/666431148).)
 ```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse)"}

### 2022-02-06 _paper.tex

--- 2022-02-06 _paper.tex
@@
 This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
-Animated tour can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.}\label{fig:casechocolatesinverse}
+(A video of the tour animation can be found at \href{https://vimeo.com/666431148}{vimeo.com/666431148}.)}\label{fig:casechocolatesinverse}
 \end{figure}

### 2022-02-06 _paper.rmd
cleaning captions for chocolates

--- 2022-02-06 _paper.rmd
@@
 
-(ref:casechocolates) Chocolates data type classification (milk or dark).
-We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate from the values on the nutritional label.
-The attribution projection already looks more like a dark chocolate than milk.
-We remove four features with the lowest contribution for the selected instance and vary the contribution of Fiber.
-The misclassification seems improbable when sugar is near the max contribution.
+(ref:casechocolates) Examining the local interpretation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection this chocolate correctly dark more than milk, which suggests that the local explanation does not help to understand the prediction for this instance.
+So, we vary the contribution of Sugar -- reducing it corresponds primarily with increasing Fiber.
+When Sugar is zero, Fiber is contributing strongly towards the left.
+In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
+This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
 (A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
@@
 
-(ref:casechocolatesinverse) Chocolates data type classification (milk or dark).
-Looking at the inverse misclassification, we select a milk chocolate while the model predicts it to be dark chocolate.
-Sodium and Fiber (`Na` and `Fbr`) have the largest differences in attributed feature importance.
-We remove features with the lowest contributions and vary the contribution of Sodium.
-This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
+(ref:casechocolatesinverse) Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+In the attribution projection the PI could be either milk or dark.
+Sodium and Fiber have the largest differences in attributed feature importance, both with low values relative to other milk chocolates.
+We suspect the lack of use of these variables is what has contributed to the mistake, so we vary the contribution of Sodium.
+If sodium had a larger contribution to the prediction (like in this view) the PI would look more like other milk chocolates.
 (A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)

### 2022-02-06 _paper.tex
cleaning captions for chocolates

--- 2022-02-06 _paper.tex
@@
 
-\caption{Chocolates data type classification (milk or dark).
-We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate from the values on the nutritional label.
-The attribution projection already looks more like a dark chocolate than milk.
-We remove four features with the lowest contribution for the selected instance and vary the contribution of Fiber.
-The misclassification seems improbable when sugar is near the max contribution.
+\caption{Examining the local interpretation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection this chocolate correctly dark more than milk, which suggests that the local explanation does not help to understand the prediction for this instance.
+So, we vary the contribution of Sugar -- reducing it corresponds primarily with increasing Fiber.
+When Sugar is zero, Fiber is contributing strongly towards the left.
+In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
+This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
 (A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolates}
@@
 
-\caption{Chocolates data type classification (milk or dark).
-Looking at the inverse misclassification, we select a milk chocolate while the model predicts it to be dark chocolate.
-Sodium and Fiber (\texttt{Na} and \texttt{Fbr}) have the largest differences in attributed feature importance.
-We remove features with the lowest contributions and vary the contribution of Sodium.
-This misclassification is not supported when sodium has contribution close to the attribution aligned with the other milk chocolates.
+\caption{Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+In the attribution projection the PI could be either milk or dark.
+Sodium and Fiber have the largest differences in attributed feature importance, both with low values relative to other milk chocolates.
+We suspect the lack of use of these variables is what has contributed to the mistake, so we vary the contribution of Sodium.
+If sodium had a larger contribution to the prediction (like in this view) the PI would look more like other milk chocolates.
 (A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolatesinverse}

### 2022-02-06 _paper.rmd

@@
 
-Version 0.2.0 of the package can be installed from CRAN using the following __R__ code:
+The package can be installed from GitHub using the following __R__ code:
 
 ```{r eval=FALSE, echo=TRUE}
-install.packages("cheem", dependencies = TRUE)
+# Install remotes if absent
+if(require("remotes") == FALSE) install.packages("remotes")
+remotes::install.packages("cheem", dependencies = TRUE)
 library("cheem")
@@
 ```{r eval=FALSE, echo=TRUE}
-# install.packages("remotes") # if you don't have it
 remotes::install_github('ModelOriented/treeshap')

### 2022-02-06 _paper.tex

@@
 
-Version 0.2.0 of the package can be installed from CRAN using the following \textbf{R} code:
+The package can be installed from GitHub using the following \textbf{R} code:
 
@@
 \begin{Highlighting}[]
-\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
+\CommentTok{\# Install remotes if absent}
+\ControlFlowTok{if}\NormalTok{(}\FunctionTok{require}\NormalTok{(}\StringTok{"remotes"}\NormalTok{) }\SpecialCharTok{==} \ConstantTok{FALSE}\NormalTok{) }\FunctionTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{)}
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
 \FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
@@
 \begin{Highlighting}[]
-\CommentTok{\# install.packages("remotes") \# if you don\textquotesingle{}t have it}
 \NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
@@
 
-and follow the examples provided with the package to compute the local explainers. XXX Add some information on what type of data object gets passed into cheem.
-
+Follow the examples provided with the package to compute the local explainers, (see \texttt{?cheem\_ls}).
+The application expects the return of call from \texttt{cheem\_ls()} which is then saved to an .rds file with \texttt{saveRDS()}.
 Alternatively, the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.

### 2022-02-07 _paper.rmd
house prices and conclusion

@@
 (ref:casepenguins) Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is an Chinstrap (orange) penguin that is misclassified as a Gentoo (purple), marked as an asterisk in (a), and the dashed vertical line in (b).
-The radial view shows varying the contribution of `f_l` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
+The PI is a Gentoo (purple) Chinstrap (orange) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+The radial view shows varying the contribution of `f_l` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
 (The animation of the radial tour is at [vimeo.com/666431172](https://vimeo.com/666431172).)
@@
 
-(ref:caseames) Ames housing 2018 regressing sales price [USD].
-The PI sale price was under predicted and had sizable attribution to lot area (`LtA`).
-The CI was predicted sales price was similar and much more accurate with its observed sales price while it has very little attribution to lot area.
-Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of `LtA`, which is important to explaining the PI and near invariant to the sales price of the CI.
+(ref:caseames) Exploring an instance with a large residual as the PI from fitting sales price [USD] to other variables in the Ames housing 2018.
+The sale price of the PI was under-predicted.
+The local explanation indicates a sizable attribution to lot area (`LtA`).
+The CI has a similar predicted sales price and smaller residual, and has very little attribution to lot area.
+In the attribution projection, the PI has a higher sales price than the CI.
+Reducing the contribution of lot area brings these two prices in line.
+This suggests if lot area was not considered by the model, then the two houses would be quite similar.
+That is, the large residual is due to a lack of factoring in the lot area for the prediction of PI's sales price.
 (A video showing the animation is at [vimeo.com/666431134](https://vimeo.com/666431134).)
@@
 
-The need to maintain the interpretability of black-box models is evident.
-One aspect uses local explanations of the model in the vicinity of an instance.
-Local explanations approximate the linear feature importance to the model.
-Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify instances of interest.
-The normalized linear importance from the explanation of the PI becomes the feature of interest to further explore with the radial tour.
-The tours explore the feature sensitivity to the structure identified in the explanation.
+New tools to assist with the interpretability of black-box models is increasingly important.
+This paper provides a technique that builds on local interpretations to explore the feature importance local to an instance.
+The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
+Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
+A global view shows the data space, explanation space, model fit.
+The user can interactively select instances to compare and contrast, and study further.
 
-We have illustrated this method on random forest models using the tree SHAP local explanation.
-While this analysis could be generally used with any compatible model-explanation pairing.
-We have created an open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem), to facilitate preprocessing and exploration with the described interactive application.
-Toy and real data are provided, or upload your data after preprocessing.
+The approach has been illustrated on four data examples using random forest models and using the tree SHAP local explanation.
+In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of it's species.
+This was verified by making a follow-up plot of the data including this variable.
+In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate misclassified as dark due to it's lowish Sodium value.
+For the FIFA example, we show how low Messi's salary would go if it depended on defensive skills.
+In the Ames housing data, an inaccurate prediction for a house was likely due to lot area not being effectively used.
 
-One limitation of the __cheem__ packages is that it currently only calculates tree SHAP for tree-based models supported by __treeshap__ (tree based models from __randomForest__, __ranger__, __gbm__, __xgboost__, __lightgbm__, or __catboost__).
-The SHAP and oscillation explainations could be added with use of the `DALEX::explain()` seems to be an excellent direction to extend [@biecek_dalex_2018; @biecek_explanatory_2021].
-Tree SHAP was selected because of its reduced computational complexity; processing runtime of the explanation will be a looming concern.
+The approach is manually intensive, and thus only feasible for investigating a few instances.
+The approach is to select instances which the model hasn't done well and compare it with an instance where the model did fit well.
+The radial tour launches from the attribution projection, to enable exploration of the sensitivity of the prediction to any feature.
+It can be helpful to make additional plots of the features and response in order to cross-check interpretations made from the cheem viewer.
+This methodology provides an additional tool in the box for studying model fitting.
 
-Attribution space does tend to be more separable than data space at least in the first couple principal components.
+An implemenation is provided in the open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem).
+Example data sets are provided, and you can upload your own data after model fitting and computing the local explanaitons.
+In theory this approach would work with any black box model but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ (tree based models from __randomForest__, __ranger__, __gbm__, __xgboost__, __lightgbm__, or __catboost__).
+Tree SHAP was selected because of its computational efficiency.
+The SHAP and oscillation explanations could be added with use of the `DALEX::explain()` and would be an excellent direction to extend the work [@biecek_dalex_2018; @biecek_explanatory_2021].
+
+<!-- Attribution space does tend to be more separable than data space at least in the first couple principal components.
 Different statistics should be explored to convey a clearer understanding what structure the explanation distinguishes.
@@
 It sounds like a candidate to fit another model.
-However, such an attempt should want to be vigilant not to over-fit the training data.
+However, such an attempt should want to be vigilant not to over-fit the training data.-->
 

### 2022-02-07 _paper.tex
house prices and conclusion

@@
 \caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is an Chinstrap (orange) penguin that is misclassified as a Gentoo (purple), marked as an asterisk in (a), and the dashed vertical line in (b).
-The radial view shows varying the contribution of \texttt{f\_l} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably a Chinstrap than a Gentoo (b, right).
+The PI is a Gentoo (purple) Chinstrap (orange) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+The radial view shows varying the contribution of \texttt{f\_l} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
 (The animation of the radial tour is at \href{https://vimeo.com/666431172}{vimeo.com/666431172}.)}\label{fig:casepenguins}
@@
 
-\caption{Ames housing 2018 regressing sales price {[}USD{]}.
-The PI sale price was under predicted and had sizable attribution to lot area (\texttt{LtA}).
-The CI was predicted sales price was similar and much more accurate with its observed sales price while it has very little attribution to lot area.
-Varying the contribution lot area the separation between these house sales crosses when there is a low contribution of \texttt{LtA}, which is important to explaining the PI and near invariant to the sales price of the CI.
+\caption{Exploring an instance with a large residual as the PI from fitting sales price {[}USD{]} to other variables in the Ames housing 2018.
+The sale price of the PI was under-predicted.
+The local explanation indicates a sizable attribution to lot area (\texttt{LtA}).
+The CI has a similar predicted sales price and smaller residual, and has very little attribution to lot area.
+In the attribution projection, the PI has a higher sales price than the CI.
+Reducing the contribution of lot area brings these two prices in line.
+This suggests if lot area was not considered by the model, then the two houses would be quite similar.
+That is, the large residual is due to a lack of factoring in the lot area for the prediction of PI's sales price.
 (A video showing the animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.)}\label{fig:caseames}
@@
 
-The need to maintain the interpretability of black-box models is evident.
-One aspect uses local explanations of the model in the vicinity of an instance.
-Local explanations approximate the linear feature importance to the model.
-Our contribution is to assess explanations by examining the support by varying the contributions with a radial tour.
-First, a global view visualizes approximations of the data space, explanation space, model predictions side-by-side, using dynamic interaction to compare and contrast and identify instances of interest.
-The normalized linear importance from the explanation of the PI becomes the feature of interest to further explore with the radial tour.
-The tours explore the feature sensitivity to the structure identified in the explanation.
+New tools to assist with the interpretability of black-box models is increasingly important.
+This paper provides a technique that builds on local interpretations to explore the feature importance local to an instance.
+The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
+Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
+A global view shows the data space, explanation space, model fit.
+The user can interactively select instances to compare and contrast, and study further.
 
-We have illustrated this method on random forest models using the tree SHAP local explanation.
-While this analysis could be generally used with any compatible model-explanation pairing.
-We have created an open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}, to facilitate preprocessing and exploration with the described interactive application.
-Toy and real data are provided, or upload your data after preprocessing.
+The approach has been illustrated on four data examples using random forest models and using the tree SHAP local explanation.
+In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of it's species.
+This was verified by making a follow-up plot of the data including this variable.
+In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate misclassified as dark due to it's lowish Sodium value.
+For the FIFA example, we show how low Messi's salary would go if it depended on defensive skills.
+In the Ames housing data, an inaccurate prediction for a house was likely due to lot area not being effectively used.
 
-One limitation of the \textbf{cheem} packages is that it currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{randomForest}, \textbf{ranger}, \textbf{gbm}, \textbf{xgboost}, \textbf{lightgbm}, or \textbf{catboost}).
-The SHAP and oscillation explainations could be added with use of the \texttt{DALEX::explain()} seems to be an excellent direction to extend (Biecek 2018; Biecek and Burzykowski 2021).
-Tree SHAP was selected because of its reduced computational complexity; processing runtime of the explanation will be a looming concern.
+The approach is manually intensive, and thus only feasible for investigating a few instances.
+The approach is to select instances which the model hasn't done well and compare it with an instance where the model did fit well.
+The radial tour launches from the attribution projection, to enable exploration of the sensitivity of the prediction to any feature.
+It can be helpful to make additional plots of the features and response in order to cross-check interpretations made from the cheem viewer.
+This methodology provides an additional tool in the box for studying model fitting.
 
-Attribution space does tend to be more separable than data space at least in the first couple principal components.
-Different statistics should be explored to convey a clearer understanding what structure the explanation distinguishes.
-It is curious that the attribution space the same dimensionality of the data with a seemingly better separation.
-It sounds like a candidate to fit another model.
-However, such an attempt should want to be vigilant not to over-fit the training data.
+An implemenation is provided in the open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
+Example data sets are provided, and you can upload your own data after model fitting and computing the local explanaitons.
+In theory this approach would work with any black box model but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{randomForest}, \textbf{ranger}, \textbf{gbm}, \textbf{xgboost}, \textbf{lightgbm}, or \textbf{catboost}).
+Tree SHAP was selected because of its computational efficiency.
+The SHAP and oscillation explanations could be added with use of the \texttt{DALEX::explain()} and would be an excellent direction to extend the work (Biecek 2018; Biecek and Burzykowski 2021).
 

### 2022-02-07 _paper.rmd
-Grammerly edits -Variable (caps, not `code`) consistency -Migrate to thesis

@@
 Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
-A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
+A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
 Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
@@
 We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
+Additional interactivity allows the user to identify an instance of interest then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in R package __cheem__.
@@
 @strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
-The attribution of feature importance depends on the sequence of the features already included.
+The attribution of feature importance depends on the sequence of the included features.
 The SHAP values are the mean contributions over different feature sequences.
@@
 van Dijk, respectively.
-Panel (a) shows breakdown plots of three sequences of the features, order and magnitude change.
-Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot, for each players.
+Panel (a) shows breakdown plots of three sequences of the features, order, and magnitude change.
+Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
@@
 <!-- Fifa example -->
-Following the use case _Explanatory Model Analysis_ [@biecek_explanatory_2021], we use FIFA data to illustrate the use of SHAP.
+Following the use case _Explanatory Model Analysis_ [@biecek_explanatory_2021], we use FIFA data to illustrate SHAP use.
 Consider soccer data from the FIFA 2020 season [@leone_fifa_2020].
 There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing wages [2020 Euros], from the skill measures.
+A random forest model is fit regressing wages [2020 Euros] from the skill measures.
 We then extract the SHAP values of a star offensive player (L.
@@
 In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
-There are different plots used for categorical and quantitative response.
+There are different plots used for the categorical and quantitative responses.
 Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
@@
 Both of these spaces are of dimension $n\times p$, where $n$ is the number of instances and $p$ is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, which will have $p$ values for each instance.
+The attribution space corresponds to the local explanations for each instance, each a vector of $p$ values.
 
@@
 These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are useful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:classificationcase) XXX), to help identify instances poorly predicted by the model.
+In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:regressioncase)a), to help identify instances poorly predicted by the model.
 <!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
@@
 The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
-These instances are highlighted as asterisk and $\times$ respectively.
+These instances are highlighted as asterisk and $\times$, respectively.
 
@@
 <!-- Local explanations as par coords -->
-The local explanations for all observations are normalized (squared sum of values adds to 1) and thus the relative importance of features can be compared across all instances.
+The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
@@
 From this plot, we would read the range of importances across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature, for example, Figure \@ref(fig:classificationcase)e suggests that `b_l` is important for distinguishing the green class from the other two.
+For classification, one would look at differences between groups on any feature.
+For example, Figure \@ref(fig:classificationcase)e suggests that `b_l` is important for distinguishing the green class from the other two.
 For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, and which have a range of high and low values (e.g.
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection is an  approximation of the feature importance, for the prediction of this instance.
+The attribution projection is an  approximation of the feature importance for the prediction of this instance.
 This combination of features best explains the difference between the mean response and an instance’s predicted value.
-It is not an indication of the local shape of the model surface, that is, it is not some indication of the tangent to the curve at this point.
+It is not an indication of the local shape of the model surface.
+That is, it is not some indication of the tangent to the curve at this point.
 
@@
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The user uses the radial tour to vary contribution of the selected feature between 0-1.
+The user uses the radial tour to vary the contribution of the selected feature between 0-1.
 Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
@@
 
-Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
+Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI makes it easier to examine the features most responsible for the error.
 The global view (Figure \@ref(fig:classificationcase)c) displays the model confusion matrix.
 The radial tour is 1D, plotted as a density where color indicates class.
-A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
+A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
 
@@
 Radial tour inputs (d) select features to include and which feature is changed in the tour.
-(e) shows parallel coordinate display of the distribution of the feature attributions while bars dipict contribution for the current basis.
+(e) shows a parallel coordinate display of the distribution of the feature attributions while bars depict contribution for the current basis.
 The black bar is the variable being changed in the radial tour.
@@
 
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
+Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing or not.
 The global view (Figure \@ref(fig:regressioncase)a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation, to help with understand where the model fits better or worse.
 
@@
 This can be viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a feature is varied, if the value of the PI doesn't change much it would indicate that the prediction for this instance is NOT sensitive to that feature.
+As the contribution of a feature is varied, if the value of the PI does not change much, it would indicate that the prediction for this instance is NOT sensitive to that feature.
 Conversely, if the predicted value varies substantially, the prediction is very sensitive to that feature, suggesting that the feature is very important for the PI's prediction.
@@
 Panel (a) PCA of the data and attributions spaces, (b), residual plot, predictions by observed values.
-Four points areselected points and highlighted in the PC spaces and tabularly displayed.
+Four points are selected points and highlighted in the PC spaces and tabularly displayed.
 Coloring on a statistic (c) highlights structure organized in the attribution space.
 Interactive tabular display (d) populates when instances are selected.
-Contribution of the 1D basis effecting the horizontal position (e) parallel coorindate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right)."}
+Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (left and right)."}
 knitr::include_graphics("./figures/app_regression_interactions.PNG")
@@
 All computational operations should be prepared before runtime.
-The work remaining when an application is run solely reacts to inputs and rendering of visuals and tables.
+The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
 Below we discuss the steps and details of the reprocessing.
@@
 Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
-This implementation aggregates over exhaustively over all trees' attribution, and we opt not to fit interactions of features.
+This implementation aggregates exhaustively over all trees' attribution, and we opt not to fit interactions of features.
 	\item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
-Statistics for mapping to color are calculated on the scaled spaces.
+Statistics for mapping to color are computed on the scaled spaces.
 Interactive tabular display reports the original values.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP a good candidate to start with.
+Its reduced computational complexity makes tree SHAP a good candidate to start.
 (Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
@@
 The application expects the return of call from `cheem_ls()` which is then saved to an .rds file with `saveRDS()`.
-Alternatively, the cheem viewer shiny app can be directly accessed at https://ebsmonash.shinyapps.io/cheem_initial/.
+Alternatively, the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/).
 
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances and we will use the four physical measurements, `bill_length_mm` (`b_l`), `bill_depth_mm` (`b_d`), `flipper_length_mm` (`f_l`), `body_mass_g` (`wgt`), for this illustration.
+After removing incomplete instances, there are 333 instances, and we will use the four physical measurements, bill length (`b_l`), bill depth (`b_d`), flipper length (`f_l`), body mass (`wgt`), for this illustration.
 A random forest model was fit with species as the response feature.
@@
 We will explore why a Gentoo has been wrongly labeled as a Chinstrap for this illustration.
-The PI is a misclassified point (represented by the asterisk in the global view and as a dashed vertical line in the tour view).
+The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
 The CI is a correctly classified point (represented by an $\times$ and a vertical dotted line).
@@
 Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view it makes sense that its a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
+From this view, it makes sense that its a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
 
@@
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate that is nutritionally more like milk should not be considered a healthy alternative.
+A reason to do this is that a dark chocolate nutritionally more like milk should not be considered a healthy alternative.
 It is interesting to explore which of the nutritional features contribute most to misclassification.
@@
 (ref:casechocolates) Examining the local interpretation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection this chocolate correctly dark more than milk, which suggests that the local explanation does not help to understand the prediction for this instance.
+From the attribution projection, this chocolate correctly looks more like dark more than milk, which suggests that the local explanation does not help understand the prediction for this instance.
 So, we vary the contribution of Sugar -- reducing it corresponds primarily with increasing Fiber.
@@
 The PCA plot, and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
-The misclassifications are clearer in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
+The misclassifications are more apparent in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
 
-The attribution projection for chocolate 22 suggests that Fiber, Sugars and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value, while Sugars and Calories have reasonably large positive values.
+The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
 In the density plot, instances on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be the interpretation of a point with high values in this plot.
+The opposite would be interpreting a point with high values in this plot.
 The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
-The density for milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
+The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
 
@@
 From the density plot, which is the attribution projection corresponding to the instance of interest, both instances are more like dark chocolates.
-If we vary the contribution of Sugars, and completely remove Sugars from the projection, this is where the difference becomes apparent.
+If we vary the contribution of Sugars and altogether remove Sugars from the projection, this is where the difference becomes apparent.
 When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
@@
 (ref:casechocolatesinverse) Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
-In the attribution projection the PI could be either milk or dark.
+In the attribution projection, the PI could be either milk or dark.
 Sodium and Fiber have the largest differences in attributed feature importance, both with low values relative to other milk chocolates.
 We suspect the lack of use of these variables is what has contributed to the mistake, so we vary the contribution of Sodium.
-If sodium had a larger contribution to the prediction (like in this view) the PI would look more like other milk chocolates.
+If Sodium had a larger contribution to the prediction (like in this view), the PI would look more like other milk chocolates.
 (A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
@@
 It would also be interesting to explore the inverse case.
-This would describes which features lead to a milk chocolate being misclassified as dark and how those attributions differ from the previous misclassification.
+This would describe which features lead to a milk chocolate being misclassified as dark and how those attributions differ from the previous misclassification.
 Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
@@
 The difference of position in the tree SHAP PCA with the previous case is quite large; this gives an approximate feel that the attribution should be quite different.
-Looking to the initial contribution we find this to be the case.
-Previously fiber was very important while it is absent from the attribution in this case.
-Conversely, calories from fat and total fat are highly attributed here, while unimportant in the preceding case.
+Looking at the initial contribution, we find this to be the case.
+Previously Fiber was very important while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat are highly attributed here, while unimportant in the preceding case.
 
-Comparing the attribution with the CI (dotted line) discrepancies in Sodium and Fiber are identified.
+Comparing the attribution with the CI (dotted line), discrepancies in Sodium and Fiber are identified.
 We opt to vary Sodium.
-Even in the initial projection, the instance looks slightly more like its observered milk than predicted dark chocolate.
-The misclassification seem least supported when the basis reaches sodium attribution of typical dark chocolate.
+The instance looks slightly more like its observed milk than predicted dark chocolate even in the initial projection.
+The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
 
@@
 The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
-From a plot of the correlation matrix, nine higher level skill groupings were identified and aggregated.
+From a correlation matrix plot, nine higher-level skill groupings were identified and aggregated.
 A random forest model if fit from these aggregations and regress player wages [2020 euros].
@@
 Continuing from the exploration in section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field position of players.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players.
 We compare a leading offensive fielder (L.
@@
 The attribution projection starts with the selected instance on the right.
-We vary the contribution from defense (`def`), the star offensive player is not distinguished in the horizontal direction.
+We vary the contribution from defense (`def`).
+The star offensive player is not distinguished in the horizontal direction.
 At this point, defensive players have been rotated to the highest horizontal value.
@@
 
-With figure \@ref(fig:casefifa), we will test the sopport of the local explanation.
+With figure \@ref(fig:casefifa), we will test the support of the local explanation.
 Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
-If either of them are rotated out the other will be rotated into frame maintaining a far-right position.
+If either of them were rotated out, the other will be rotated into frame, maintaining a far-right position.
 However, when we vary the defensive skills, the other skills are rotated out of the frame.
@@
 As the contribution of defensive skills increases, Messi's is no longer separated from the group.
-Players with high values in defensive skills are now the right most points.
+Players with high values in defensive skills are now the rightmost points.
 In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Build (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Build (`GYB`), and Garage Area (`GrA`).
+A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 The sale price of the PI was under-predicted.
-The local explanation indicates a sizable attribution to lot area (`LtA`).
-The CI has a similar predicted sales price and smaller residual, and has very little attribution to lot area.
+The local explanation indicates a sizable attribution to Lot Area (`LtA`).
+The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
 In the attribution projection, the PI has a higher sales price than the CI.
 Reducing the contribution of lot area brings these two prices in line.
-This suggests if lot area was not considered by the model, then the two houses would be quite similar.
+This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
 That is, the large residual is due to a lack of factoring in the lot area for the prediction of PI's sales price.
@@
 
-Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction that has a large contribution from lot area.
+Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction with a large Lot Area contribution.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
@@
 As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
-This large of an importance is living area is relatively uncommon.
-Boosting tree models may be more resilient to such an under prediction as up-weighting this residual would force its inclusion in the final model.
+This large of importance in the Living Area is relatively uncommon.
+Boosting tree models may be more resilient to such an under-prediction as up-weighting this residual would force its inclusion in the final model.
 
@@
 <!-- summarise findings -->
-<!-- suggest next steps -->>
+<!-- suggest next steps -->
 
@@
 A global view shows the data space, explanation space, model fit.
-The user can interactively select instances to compare and contrast, and study further.
+The user can interactively select instances to compare, contrast, and study further.
 
 The approach has been illustrated on four data examples using random forest models and using the tree SHAP local explanation.
-In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of it's species.
-This was verified by making a follow-up plot of the data including this variable.
-In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate misclassified as dark due to it's lowish Sodium value.
+In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
+This was verified by making a follow-up plot of the data, including this variable.
+In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
 For the FIFA example, we show how low Messi's salary would go if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to lot area not being effectively used.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used.
 
-The approach is manually intensive, and thus only feasible for investigating a few instances.
-The approach is to select instances which the model hasn't done well and compare it with an instance where the model did fit well.
-The radial tour launches from the attribution projection, to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and response in order to cross-check interpretations made from the cheem viewer.
+The approach is manually intensive and thus only feasible for investigating a few instances.
+The approach is to select instances that the model has not done well and compare it with an instance where it did fit well.
+The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
+It can be helpful to make additional plots of the features and responses in order to cross-check interpretations made from the cheem viewer.
 This methodology provides an additional tool in the box for studying model fitting.
 
-An implemenation is provided in the open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem).
-Example data sets are provided, and you can upload your own data after model fitting and computing the local explanaitons.
-In theory this approach would work with any black box model but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ (tree based models from __randomForest__, __ranger__, __gbm__, __xgboost__, __lightgbm__, or __catboost__).
+An implementation is provided in the open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem).
+Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
+In theory, this approach would work with any black-box model but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ (tree based models from __randomForest__, __ranger__, __gbm__, __xgboost__, __lightgbm__, or __catboost__).
 Tree SHAP was selected because of its computational efficiency.
-The SHAP and oscillation explanations could be added with use of the `DALEX::explain()` and would be an excellent direction to extend the work [@biecek_dalex_2018; @biecek_explanatory_2021].
+The SHAP and oscillation explanations could be added with the use of the `DALEX::explain()` and would be an excellent direction to extend the work [@biecek_dalex_2018; @biecek_explanatory_2021].
 

### 2022-02-07 _paper.tex
-Grammerly edits -Variable (caps, not `code`) consistency -Migrate to thesis

@@
 Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
-A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation), or the predictions at a specific point in the data domain.
+A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
 Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
@@
 We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest, then explore its local explanation by changing feature contribution with the radial tour.
+Additional interactivity allows the user to identify an instance of interest then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in R package \textbf{cheem}.
@@
 Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
-The attribution of feature importance depends on the sequence of the features already included.
+The attribution of feature importance depends on the sequence of the included features.
 The SHAP values are the mean contributions over different feature sequences.
@@
 van Dijk, respectively.
-Panel (a) shows breakdown plots of three sequences of the features, order and magnitude change.
-Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot, for each players.
+Panel (a) shows breakdown plots of three sequences of the features, order, and magnitude change.
+Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
@@
 
-Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), we use FIFA data to illustrate the use of SHAP.
+Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), we use FIFA data to illustrate SHAP use.
 Consider soccer data from the FIFA 2020 season (Leone 2020).
 There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing wages {[}2020 Euros{]}, from the skill measures.
+A random forest model is fit regressing wages {[}2020 Euros{]} from the skill measures.
 We then extract the SHAP values of a star offensive player (L.
@@
 In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
-There are different plots used for categorical and quantitative response.
+There are different plots used for the categorical and quantitative responses.
 Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
@@
 Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, which will have \(p\) values for each instance.
+The attribution space corresponds to the local explanations for each instance, each a vector of \(p\) values.
 
@@
 These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are useful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:classificationcase} XXX), to help identify instances poorly predicted by the model.
+In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a), to help identify instances poorly predicted by the model.
 For classification tasks, misclassified instances are circled in red if applicable.
@@
 The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
-These instances are highlighted as asterisk and \(\times\) respectively.
+These instances are highlighted as asterisk and \(\times\), respectively.
 
@@
 
-The local explanations for all observations are normalized (squared sum of values adds to 1) and thus the relative importance of features can be compared across all instances.
+The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
 These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
@@
 From this plot, we would read the range of importances across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature, for example, Figure \ref{fig:classificationcase}e suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
+For classification, one would look at differences between groups on any feature.
+For example, Figure \ref{fig:classificationcase}e suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
 For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection is an approximation of the feature importance, for the prediction of this instance.
+The attribution projection is an approximation of the feature importance for the prediction of this instance.
 This combination of features best explains the difference between the mean response and an instance's predicted value.
-It is not an indication of the local shape of the model surface, that is, it is not some indication of the tangent to the curve at this point.
+It is not an indication of the local shape of the model surface.
+That is, it is not some indication of the tangent to the curve at this point.
 
@@
 
-Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI, makes it easier to examine the features most responsible for the error.
+Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI makes it easier to examine the features most responsible for the error.
 The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
 The radial tour is 1D, plotted as a density where color indicates class.
-A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity the separation to that feature.
+A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
 
@@
 
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing, or not.
+Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing or not.
 The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance) or correlation, to help with understand where the model fits better or worse.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation, to help with understand where the model fits better or worse.
 
@@
 Panel (a) PCA of the data and attributions spaces, (b), residual plot, predictions by observed values.
-Four points areselected points and highlighted in the PC spaces and tabularly displayed.
+Four points are selected points and highlighted in the PC spaces and tabularly displayed.
 Coloring on a statistic (c) highlights structure organized in the attribution space.
 Interactive tabular display (d) populates when instances are selected.
-Contribution of the 1D basis effecting the horizontal position (e) parallel coorindate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals, (left and right).}\label{fig:regressioncase}
+Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (left and right).}\label{fig:regressioncase}
 \end{figure}
@@
 2021).
-This implementation aggregates over exhaustively over all trees' attribution, and we opt not to fit interactions of features.
+This implementation aggregates exhaustively over all trees' attribution, and we opt not to fit interactions of features.
     \item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
-Statistics for mapping to color are calculated on the scaled spaces.
+Statistics for mapping to color are computed on the scaled spaces.
 Interactive tabular display reports the original values.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP a good candidate to start with.
+Its reduced computational complexity makes tree SHAP a good candidate to start.
 (Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
@@
 The application expects the return of call from \texttt{cheem\_ls()} which is then saved to an .rds file with \texttt{saveRDS()}.
-Alternatively, the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
+Alternatively, the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
 
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances and we will use the four physical measurements, \texttt{bill\_length\_mm} (\texttt{b\_l}), \texttt{bill\_depth\_mm} (\texttt{b\_d}), \texttt{flipper\_length\_mm} (\texttt{f\_l}), \texttt{body\_mass\_g} (\texttt{wgt}), for this illustration.
+After removing incomplete instances, there are 333 instances, and we will use the four physical measurements, bill length (\texttt{b\_l}), bill depth (\texttt{b\_d}), flipper length (\texttt{f\_l}), body mass (\texttt{wgt}), for this illustration.
 A random forest model was fit with species as the response feature.
@@
 We will explore why a Gentoo has been wrongly labeled as a Chinstrap for this illustration.
-The PI is a misclassified point (represented by the asterisk in the global view and as a dashed vertical line in the tour view).
+The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
 The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
@@
 \caption{Examining the local interpretation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection this chocolate correctly dark more than milk, which suggests that the local explanation does not help to understand the prediction for this instance.
+From the attribution projection, this chocolate correctly looks more like dark more than milk, which suggests that the local explanation does not help understand the prediction for this instance.
 So, we vary the contribution of Sugar -- reducing it corresponds primarily with increasing Fiber.
@@
 \caption{Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
-In the attribution projection the PI could be either milk or dark.
+In the attribution projection, the PI could be either milk or dark.
 Sodium and Fiber have the largest differences in attributed feature importance, both with low values relative to other milk chocolates.
 We suspect the lack of use of these variables is what has contributed to the mistake, so we vary the contribution of Sodium.
-If sodium had a larger contribution to the prediction (like in this view) the PI would look more like other milk chocolates.
+If Sodium had a larger contribution to the prediction (like in this view), the PI would look more like other milk chocolates.
 (A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolatesinverse}
@@
 The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-From a plot of the correlation matrix, nine higher level skill groupings were identified and aggregated.
+From a correlation matrix plot, nine higher-level skill groupings were identified and aggregated.
 A random forest model if fit from these aggregations and regress player wages {[}2020 euros{]}.
@@
 Continuing from the exploration in section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field position of players.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players.
 We compare a leading offensive fielder (L.
@@
 The attribution projection starts with the selected instance on the right.
-We vary the contribution from defense (\texttt{def}), the star offensive player is not distinguished in the horizontal direction.
+We vary the contribution from defense (\texttt{def}).
+The star offensive player is not distinguished in the horizontal direction.
 At this point, defensive players have been rotated to the highest horizontal value.
@@
 
-With figure \ref{fig:casefifa}, we will test the sopport of the local explanation.
+With figure \ref{fig:casefifa}, we will test the support of the local explanation.
 Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
-If either of them are rotated out the other will be rotated into frame maintaining a far-right position.
+If either of them were rotated out, the other will be rotated into frame, maintaining a far-right position.
 However, when we vary the defensive skills, the other skills are rotated out of the frame.
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Build (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Build (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 The sale price of the PI was under-predicted.
-The local explanation indicates a sizable attribution to lot area (\texttt{LtA}).
-The CI has a similar predicted sales price and smaller residual, and has very little attribution to lot area.
+The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}).
+The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
 In the attribution projection, the PI has a higher sales price than the CI.
 Reducing the contribution of lot area brings these two prices in line.
-This suggests if lot area was not considered by the model, then the two houses would be quite similar.
+This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
 That is, the large residual is due to a lack of factoring in the lot area for the prediction of PI's sales price.
@@
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction that has a large contribution from lot area.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with a large Lot Area contribution.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
@@
 \section{Discussion}\label{sec:cheemdiscussion}}
-
-\begin{quote}
-\end{quote}
 
@@
 A global view shows the data space, explanation space, model fit.
-The user can interactively select instances to compare and contrast, and study further.
+The user can interactively select instances to compare, contrast, and study further.
 
 The approach has been illustrated on four data examples using random forest models and using the tree SHAP local explanation.
-In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of it's species.
-This was verified by making a follow-up plot of the data including this variable.
-In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate misclassified as dark due to it's lowish Sodium value.
+In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
+This was verified by making a follow-up plot of the data, including this variable.
+In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
 For the FIFA example, we show how low Messi's salary would go if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to lot area not being effectively used.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used.
 
-The approach is manually intensive, and thus only feasible for investigating a few instances.
-The approach is to select instances which the model hasn't done well and compare it with an instance where the model did fit well.
-The radial tour launches from the attribution projection, to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and response in order to cross-check interpretations made from the cheem viewer.
+The approach is manually intensive and thus only feasible for investigating a few instances.
+The approach is to select instances that the model has not done well and compare it with an instance where it did fit well.
+The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
+It can be helpful to make additional plots of the features and responses in order to cross-check interpretations made from the cheem viewer.
 This methodology provides an additional tool in the box for studying model fitting.
 
-An implemenation is provided in the open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Example data sets are provided, and you can upload your own data after model fitting and computing the local explanaitons.
-In theory this approach would work with any black box model but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{randomForest}, \textbf{ranger}, \textbf{gbm}, \textbf{xgboost}, \textbf{lightgbm}, or \textbf{catboost}).
+An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
+Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
+In theory, this approach would work with any black-box model but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{randomForest}, \textbf{ranger}, \textbf{gbm}, \textbf{xgboost}, \textbf{lightgbm}, or \textbf{catboost}).
 Tree SHAP was selected because of its computational efficiency.
-The SHAP and oscillation explanations could be added with use of the \texttt{DALEX::explain()} and would be an excellent direction to extend the work (Biecek 2018; Biecek and Burzykowski 2021).
+The SHAP and oscillation explanations could be added with the use of the \texttt{DALEX::explain()} and would be an excellent direction to extend the work (Biecek 2018; Biecek and Burzykowski 2021).
 

### 2022-02-11 _paper.rmd
installation instructions

@@
 
-The package can be installed from GitHub using the following __R__ code:
+The package can be installed from CRAN, and the shiny app can be run, using the following __R__ code:
 
 ```{r eval=FALSE, echo=TRUE}
-# Install remotes if absent
-if(require("remotes") == FALSE) install.packages("remotes")
-remotes::install.packages("cheem", dependencies = TRUE)
+install.packages("cheem", dependencies = TRUE)
 library("cheem")
@@
 
-To process your own data, you will need to use the `treeshap` package, which can be installed from GitHub:
+Alternatively,
+
+- a version of the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/). 
+- the development version of the package is available at https://github.com/nspyrison/cheem, and 
+- documentation of the package can be found at https://nspyrison.github.io/cheem/.
+
+To explore your own data, you will need to use pre-process it using the `treeshap` package, which can be installed from GitHub:
 
@@
 
-Follow the examples provided with the package to compute the local explainers, (see `?cheem_ls`).
-The application expects the return of call from `cheem_ls()` which is then saved to an .rds file with `saveRDS()`.
-Alternatively, the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/).
+Follow the examples provided with the package to compute the local explainers (using `?cheem_ls`).
+The application expects the output returned by `cheem_ls()`, saved to an .rds file with `saveRDS()`, to be uploaded.
 

### 2022-02-11 _paper.tex
installation instructions

@@
 
-The package can be installed from GitHub using the following \textbf{R} code:
+The package can be installed from CRAN, and the shiny app can be run, using the following \textbf{R} code:
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Install remotes if absent}
-\ControlFlowTok{if}\NormalTok{(}\FunctionTok{require}\NormalTok{(}\StringTok{"remotes"}\NormalTok{) }\SpecialCharTok{==} \ConstantTok{FALSE}\NormalTok{) }\FunctionTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{)}
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
+\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
 \FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
@@
 
-To process your own data, you will need to use the \texttt{treeshap} package, which can be installed from GitHub:
+Alternatively,
+
+\begin{itemize}
+\tightlist
+\item
+  a version of the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
+\item
+  the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
+\item
+  documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+\end{itemize}
+
+To explore your own data, you will need to use pre-process it using the \texttt{treeshap} package, which can be installed from GitHub:
 
@@
 
-Follow the examples provided with the package to compute the local explainers, (see \texttt{?cheem\_ls}).
-The application expects the return of call from \texttt{cheem\_ls()} which is then saved to an .rds file with \texttt{saveRDS()}.
-Alternatively, the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
+Follow the examples provided with the package to compute the local explainers (using \texttt{?cheem\_ls}).
+The application expects the output returned by \texttt{cheem\_ls()}, saved to an .rds file with \texttt{saveRDS()}, to be uploaded.
 

### 2022-02-11 _paper.rmd
figure 8 caption

@@
 
-(ref:casefifa) FIFA 2020 data, a random forest model, regresses wages [2020 Euros] from nine aggregated skill measurements.
-The PI is a star offensive player (L.
-Messi) compared with a top defensive player (V.
+(ref:casefifa) Exploring the wages (euros) relative to skill measurements in the FIFA 2020 data.
+Star offensive player (L.
+Messi) is the PI and he is compared with a top defensive player (V.
 van Dijk).
-We remove three features with low attribution from both players.
-The attribution projection starts with the selected instance on the right.
-We vary the contribution from defense (`def`).
-The star offensive player is not distinguished in the horizontal direction.
-At this point, defensive players have been rotated to the highest horizontal value.
+The attribution projection is shown at left, and it can  be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
+Defense (`def`) is the chosen variable to vary.
+It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
+The increased contribution in defense comes at the expense of offensive and reaction skills.
+The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
 (A video of the animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).)
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+A random forest model has fitted price (USD) to the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
 
-(ref:caseames) Exploring an instance with a large residual as the PI from fitting sales price [USD] to other variables in the Ames housing 2018.
+(ref:caseames) Exploring an instance with a large residual as the PI from fitting sales price (USD) to other variables in the Ames housing 2018.
 The sale price of the PI was under-predicted.

### 2022-02-11 _paper.tex
figure 8 caption

@@
 
-\caption{FIFA 2020 data, a random forest model, regresses wages {[}2020 Euros{]} from nine aggregated skill measurements.
-The PI is a star offensive player (L.
-Messi) compared with a top defensive player (V.
+\caption{Exploring the wages (euros) relative to skill measurements in the FIFA 2020 data.
+Star offensive player (L.
+Messi) is the PI and he is compared with a top defensive player (V.
 van Dijk).
-We remove three features with low attribution from both players.
-The attribution projection starts with the selected instance on the right.
-We vary the contribution from defense (\texttt{def}).
-The star offensive player is not distinguished in the horizontal direction.
-At this point, defensive players have been rotated to the highest horizontal value.
+The attribution projection is shown at left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
+Defense (\texttt{def}) is the chosen variable to vary.
+It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
+The increased contribution in defense comes at the expense of offensive and reaction skills.
+The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
 (A video of the animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.)}\label{fig:casefifa}
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+A random forest model has fitted price (USD) to the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
 Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
@@
 
-\caption{Exploring an instance with a large residual as the PI from fitting sales price {[}USD{]} to other variables in the Ames housing 2018.
+\caption{Exploring an instance with a large residual as the PI from fitting sales price (USD) to other variables in the Ames housing 2018.
 The sale price of the PI was under-predicted.

### 2022-02-27 _paper.rmd
text changes from thesis, Needed: references and figure changes

--- 2022-02-07 _paper.rmd
@@
 abstract: |
- The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction with a point estimate of the linear feature importance in the vicinity of each instance.
-These can be considered linear projections and can be further explored interactively to understand better the interaction between features used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
+ The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
+This trade-off has led to the emergence of eXplainable AI (XAI).
+XAI attempts to shed light on how models use predictors to arrive at a prediction with _local explanations_, a point estimate of the linear feature importance in the vicinity of one instance.
+These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
@@
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
-Section \@ref(sec:cheemviwer) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-Then Section \@ref(sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features.
-We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explored in the future.
+Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
+Then Section \@ref(sec:casestudies) illustrates the application to supervised learning with categorical and quantitative output.
+Section \@ref(sec:cheemdiscussion) concludes with the insights gained and possible directions to explore in the future.
 
@@
 SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The explanations of SHAP almost all refer to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
-@strumbelj_efficient_2010 introduced the use of SHAP for local explanations in ML models.
+The explanations of SHAP refer to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
+@strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
 The attribution of feature importance depends on the sequence of the included features.
@@
 The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
-Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, which helps to understand the relative importance of each feature in the context of all other candidate features.
+Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, helps to understand the relative importance of each feature in the context of all other candidate features.
 
-```{r shapdistrbd, echo=F, fig.align='center', out.width="100%", fig.cap = "Illustration SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+```{r shapdistrbd, echo=F, fig.align='center', out.width="100%", fig.cap = "Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
 A star offensive and defensive player are compared, L.
@@
 van Dijk, respectively.
-Panel (a) shows breakdown plots of three sequences of the features, order, and magnitude change.
+Panel (a) shows breakdown plots of three sequences of the features.
+The sequence of the variables impacts the magnitude of their attribution.
 Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player.
+Reaction skills are important for both players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
@@
 
-<!-- tree SHAP -->
-For our application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
-Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) will have coefficients varied to further scrutinize the feature contributions.
-
 <!-- Fifa example -->
-Following the use case _Explanatory Model Analysis_ [@biecek_explanatory_2021], we use FIFA data to illustrate SHAP use.
+Following the use case _Explanatory Model Analysis_ [@biecek_explanatory_2021], FIFA data is used to  illustrate SHAP.
 Consider soccer data from the FIFA 2020 season [@leone_fifa_2020].
 There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing wages [2020 Euros] from the skill measures.
-We then extract the SHAP values of a star offensive player (L.
+A random forest model is fit regressing player's wages [2020 Euros] from their skill measurements.
+The SHAP values are compared for a star offensive player (L.
 Messi) and defensive player (V.
@@
 The results are displayed in Figure \@ref(fig:shapdistrbd).
-We expect to see a difference in the attribution of the feature importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
-Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of features is fixed, so the two instances can be more easily compared.
+A difference in the attribution of the feature importance across the two positions of the players can be expected.
+This would be interpreted as how a player's salary depends on this combination of skill sets.
+Panel (a) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where three sequences of features are presented, so the two instances can be more easily compared.
+The magnitude of the contributions depends on the sequence in which they appear.
+Panel (b) shows the differences of the player's median values of over 25 such sequences.<!-- Segue --> In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
+In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
 
-<!-- Segue -->
-In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
-In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
+<!-- tree SHAP -->
+For the application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
+Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
+Tree SHAP is only compatible with tree-based models; random forests are used for illustration.
+The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) that will be used to explore the sensitivity of the feature contributions.
 
@@
 
-To explore the local explanations, an ensemble of plots [@unwin_ensemble_2018] is provided, called the _cheem viewer_.
-There are two primary plots: the __global view__ to give the context of all of the SHAP values, and the __radial tour view__ to explore the local explanations with user-controlled rotation.
-In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
+To explore the local explanations, coordinated views [@roberts_state_2007] [also known as ensemble graphics, @unwin_ensemble_2018] are provided in the _cheem viewer_ application.
+There are two primary plots: the __global view__ to give the context of all of the SHAP values and the __radial tour view__ to explore the local explanations with user-controlled rotation.
+There are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
 There are different plots used for the categorical and quantitative responses.
@@
 <!-- Purpose -->
-The global view provides the context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
+The global view provides a context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
 Both of these spaces are of dimension $n\times p$, where $n$ is the number of instances and $p$ is the number of predictors.
@@
 <!-- Approximations of the spaces, PC1:2 -->
-A visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
-These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are useful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:regressioncase)a), to help identify instances poorly predicted by the model.
-<!-- Interactions --> For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \@ref(fig:classificationcase)a,c).
+Visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
+These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
+In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:regressioncase)a) to help identify instances poorly predicted by the model.
+<!-- Interactions --> For classification tasks, misclassified instances are circled in red.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \@ref(fig:regressioncase)d).
 
-While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances to explore the local explanations.
+While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of instances to explore the local explanations.
 The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
@@
 The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
-These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
-The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
-From this plot, we would read the range of importances across all instances can be interpreted.
+These are depicted as vertical parallel coordinate plots [@ocagne_coordonnees_1885] on the basis biplot [@gabriel_biplot_1971].
+1D biplot display depicts the values of the current basis as bars.
+The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
+The attribution projections of the PI and CI are shown as dashed and dotted lines.
+From this plot, the range and density of the importance across all instances can be interpreted.
 For classification, one would look at differences between groups on any feature.
-For example, Figure \@ref(fig:classificationcase)e suggests that `b_l` is important for distinguishing the green class from the other two.
+For example, Figure \@ref(fig:classificationcase)e suggests that `bl` is important for distinguishing the green class from the other two.
 For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, and which have a range of high and low values (e.g.
-`off`, `def`) suggesting important for some instances and not important for other instances.
+`off`, `def`) suggesting they are important for some instances and not important for other instances.
 
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection is an  approximation of the feature importance for the prediction of this instance.
+The attribution projection is approximates the feature importance for predicting this instance.
 This combination of features best explains the difference between the mean response and an instance’s predicted value.
@@
 <!-- Tour animation -->
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \@ref(fig:classificationcase)f), and as scatterplots for a quantitative response (Figure \@ref(fig:regressioncase)f).
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \@ref(fig:classificationcase)f) and as scatterplots for a quantitative response (Figure \@ref(fig:regressioncase)f).
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The user uses the radial tour to vary the contribution of the selected feature between 0-1.
+The radial tour varies the contribution of the selected feature between 0-1.
+This is viewed as an animation of the projections from many intermediate bases.
 Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
-For classification, if the separation between classes diminishes when the feature contribution is reduced then this suggests that the feature is important for the class separation.
-For regression, if the relationship scatterplot weakens when the feature contribution is reduced then this suggests that the feature is important for accurately predicting the response.
-<!-- The default feature selected has the largest discrepancy between the attribution of the PI and CI.
-The data (scaled by standard deviation away from the mean) is projected through the current basis --- the horizontal position of Fig.
-\@ref(fig:classificationcase) f and Fig.
-\@ref(fig:regressioncase) f.-->
+For classification, if the separation between classes diminishes when the feature contribution is reduced, this suggests that the feature is important for class separation.
+For regression, if the relationship scatterplot weakens when the feature contribution is reduced, indicating that the feature is important for accurately predicting the response.
 
@@
 The global view (Figure \@ref(fig:classificationcase)c) displays the model confusion matrix.
-The radial tour is 1D, plotted as a density where color indicates class.
-A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
+The radial tour is 1D and displays as density where color indicates class.
+An animation slider enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
 
@@
 The black bar is the variable being changed in the radial tour.
-Panel (f) is the resulting projection of the data indicated as density in the classification case."}
+Panel (f) is the resulting data projection indicated as density in the classification case."}
 knitr::include_graphics("./figures/app_classification.PNG")
@@
 
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing or not.
+Selecting an inaccurately predicted instance as PI and an accurately predicted instance with similar feature values as CI is a helpful way to understand how the model is failing or not.
 The global view (Figure \@ref(fig:regressioncase)a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation, to help with understand where the model fits better or worse.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation aid in understanding the structure identified in these spaces.
 
@@
 The attribution projection can be interpreted similarly to the predicted value from the global view plot.
-It represents a linear combination of the features, and a good fit would be indicated when there is a strong relationship seen with the observed values.
+It represents a linear combination of the features, and a good fit would be indicated when there is a strong relationship with the observed values.
 This can be viewed as a local linear approximation if the fitted model is nonlinear.
@@
 
-```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression task highlighting the differences from the classification task and interactive features.
-Panel (a) PCA of the data and attributions spaces, (b), residual plot, predictions by observed values.
-Four points are selected points and highlighted in the PC spaces and tabularly displayed.
+```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression  and illustration of interactive features.
+Panel (a) PCA of the data and attributions spaces, (b) residual plot, predictions by observed values.
+Four selected points are highlighted in the PC spaces and tabularly displayed.
 Coloring on a statistic (c) highlights structure organized in the attribution space.
@@
 Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (left and right)."}
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (middle and right)."}
 knitr::include_graphics("./figures/app_regression_interactions.PNG")
@@
 The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible.
-The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.
+These reactive inputs make the software flexible and extensible (Figure \@ref(fig:classificationcase)a & d).
+The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
 
@@
 A tooltip displays instance number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
-The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and, finally, identification of a primary and comparison instance.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \@ref(fig:classificationcase)a & b).
+The information corresponding to the selected points is populated on a dynamic table (Figure \@ref(fig:classificationcase)d).
+These interactions aid exploration of the spaces and, finally, the identification of primary and comparison instances.
 
@@
 
-It is vital to mitigate the render time of visuals, especially when users may want to iterate many times.
+It is vital to mitigate the render time of visuals, especially when users may want to iterate many explorations.
 All computational operations should be prepared before runtime.
 The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
-Below we discuss the steps and details of the reprocessing.
+Below discusses the steps and details of the reprocessing.
 
@@
 	\item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
-	\item \textbf{Model and explanation:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} (ref:citeRf), compatibility tree SHAP.
-We use modest hyperparameters, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
-This implementation aggregates exhaustively over all trees' attribution, and we opt not to fit interactions of features.
-	\item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
+	\item \textbf{Model:} any model and compatible explanation could be explored with this method.
+Currently, random forest models are applied via the package \textbf{randomForest} (ref:citeRf), compatibility tree SHAP.
+Modest hyperparameters are used, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+	\item \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
+This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of features.
+	\item \textbf{Cheem viewer:} after the model and full explanation space are calculated, each feature is scaled by standard deviations away from the mean to achieve common support for visuals.
 Statistics for mapping to color are computed on the scaled spaces.
-Interactive tabular display reports the original values.
 \end{itemize}
@@
 <!-- Note on time of execution -->
-The time to preprocess the data will vary significantly with the model and local explanation.
-For reference, the FIFA data, 5000 instances of nine explanatory features, took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each instance took 270 seconds combined.
+The time to preprocess the data will vary significantly with the complexity of the  model and local explanation.
+For reference, the FIFA data contained 5000 instances of nine explanatory features took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each instance took 270 seconds total.
 PCA and statistics of the features and attributions took 2.8 seconds.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP a good candidate to start.
+Its reduced computational complexity makes tree SHAP an excellent candidate to start.
 (Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
@@
 The tour visual is built with __spinifex__ [@spyrison_spinifex_2020].
-Both views are created first with first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive HTML widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a huge boon to understanding local explanations and how to apply them.
+Both views are created first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive `html` widgets with __plotly__ [@sievert_interactive_2020].
+__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a boon to understanding local explanations and how to apply them.
 
@@
 
-To process your own data, you will need to use the `treeshap` package, which can be installed from GitHub:
-
-```{r eval=FALSE, echo=TRUE}
-remotes::install_github('ModelOriented/treeshap')
-```
-
-Follow the examples provided with the package to compute the local explainers, (see `?cheem_ls`).
-The application expects the return of call from `cheem_ls()` which is then saved to an .rds file with `saveRDS()`.
+Follow the examples provided with the package to compute the local explainers (see `?cheem_ls`).
+The application uses the returned objects from a from `cheem_ls()` call.
 Alternatively, the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/).
@@
 
-To illustrate the use of the cheem method, we apply it to modern datasets, two classification examples and then two of regression.
+To illustrate the cheem method, it is applied to modern datasets, two classification examples and then two of regression.
 
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances, and we will use the four physical measurements, bill length (`b_l`), bill depth (`b_d`), flipper length (`f_l`), body mass (`wgt`), for this illustration.
+After removing incomplete instances, there are 333 instances of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), body mass (`bm`), for this illustration.
 A random forest model was fit with species as the response feature.
 
-(ref:casepenguins) Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) Chinstrap (orange) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
-The radial view shows varying the contribution of `f_l` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
+(ref:casepenguins-cap) Examining the SHAP values for a random forest model classifying Palmer penguin species.
+The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
 (The animation of the radial tour is at [vimeo.com/666431172](https://vimeo.com/666431172).)
-```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
+```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins-cap)"}
 knitr::include_graphics("./figures/case_penguins.png")
@@
 
-(ref:casepenguinsblfl) Checking what is learned from the cheem viewer.
-This is a plot of flipper length (`f_l`) and bill length (`b_l`), where an asterisk highlights the PI.
+(ref:casepenguinsblfl-cap) Checking what is learned from the cheem viewer.
+This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI.
 A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small `f_l` length which is why it is confused with a Chinstrap.
-```{r casepenguinsblfl, echo=F, out.width="60%", fig.cap = "(ref:casepenguinsblfl)"}
+The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.
+```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl-cap)"}
 knitr::include_graphics("./figures/case_penguins_BlFl.png")
@@
 The predicted vs observed plot reveals a handful of misclassified instances.
-We will explore why a Gentoo has been wrongly labeled as a Chinstrap for this illustration.
+A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
 The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
@@
 The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance are mostly `b_l` and `b_d` with small contributions of `f_l` and `wgt`.
+The important features identified by SHAP in the (wrong) prediction for this instance are mostly `bl` and `bd` with small contributions of `fl` and `bm`.
 This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
 That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this, we use the radial tour to vary the contribution of flipper length (`f_l`).
+To explore this, the radial tour is used to vary the contribution of flipper length (`fl`).
 (In our exploration, this was the third feature explored.
-It is typically useful to explore the features with larger contributions, here `b_l` and `b_d`, but when doing this, nothing was revealed about how the PI differed from other Gentoos).
-On varying `f_l` as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
-This suggests that `f_l` should be considered an important feature for explaining the (wrong) prediction.
+It is typically helpful to explore the features with more significant contributions, here `bl` and `bd`.
+Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
+On varying `fl` as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
+This suggests that `fl` should be considered an important feature for explaining the (wrong) prediction.
 
-Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`f_l`) is important for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that its a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
+Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`fl`) is vital for the confusion of the PI as a Chinstrap.
+Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two features, mainly because it has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
 
@@
 
+<!-- Data description -->
 The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
-The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
+The data were collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
 The data is available in the __cheem__ package.
-A random forest model is used for the classification of chocolate type.
+A random forest model is used for the classification of chocolate types.
 
+<!-- Dark misclassified as milk, rational -->
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate nutritionally more like milk should not be considered a healthy alternative.
-It is interesting to explore which of the nutritional features contribute most to misclassification.
+A reason to do this is that a dark chocolate, nutritionally more like milk, should not be considered a healthy alternative.
+It is interesting to explore which nutritional features contribute most to misclassification.
 
-(ref:casechocolates) Examining the local interpretation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark more than milk, which suggests that the local explanation does not help understand the prediction for this instance.
-So, we vary the contribution of Sugar -- reducing it corresponds primarily with increasing Fiber.
-When Sugar is zero, Fiber is contributing strongly towards the left.
+(ref:casechocolates-cap) Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
+So, the contribution of Sugar is varied --- reducing it corresponds primarily with increasing Fiber.
+When Sugar is zero, Fiber contributes strongly towards the left.
 In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
@@
 (A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
-```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
+```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates-cap)"}
 knitr::include_graphics("./figures/case_chocolates.png")
@@
 
+<!-- Dark misclassified as milk, description -->
 This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (instance 22).
 It is compared with a CI that is a correctly classified dark chocolate (instance 7).
-The PCA plot, and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
-The misclassifications are more apparent in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
+The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
+The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
 
+<!-- Dark misclassified as milk,  -->
 The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
@@
 The opposite would be interpreting a point with high values in this plot.
-The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
+The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
 The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
 
-The instance of interest (dashed line) can be viewed against the comparison instance (dotted line).
-Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the instance of interest.
+<!-- Attention to PCP -->
+The PI (dashed line) can be viewed against the CI (dotted line).
+Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the PI.
+The feature contributions to the two different predictions can be quickly compared in the parallel coordinate plot.
+The PI differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
 
-We can quickly compare the feature contributions to the two different predictions from the parallel coordinate plot.
-The instance of interest differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
-
-From the density plot, which is the attribution projection corresponding to the instance of interest, both instances are more like dark chocolates.
-If we vary the contribution of Sugars and altogether remove Sugars from the projection, this is where the difference becomes apparent.
+<!-- Attention to Density, end of first case -->
+From the density plot, which is the attribution projection corresponding to the PI, both instances are more like dark chocolates.
+Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
 When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
 
-(ref:casechocolatesinverse) Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+<!-- segue and selection for chocolates inverse case -->
+It would also be interesting to explore an inverse misclassification.
+In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
+Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (instance 71).
+The corresponding global view and radial tour frames are shown in Figure \@ref(fig:casechocolatesinverse).
+
+(ref:casechocolatesinverse-cap) Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed feature importance, both with low values relative to other milk chocolates.
-We suspect the lack of use of these variables is what has contributed to the mistake, so we vary the contribution of Sodium.
-If Sodium had a larger contribution to the prediction (like in this view), the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
-```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse)"}
+Sodium and Fiber have the largest differences in attributed feature importance, with low values relative to other milk chocolates.
+The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
+If Sodium had a larger contribution to the prediction (like in this view).
+the PI would look more like other milk chocolates.
+(A video of the tour animation can be found at [vimeo.com/666431148](https://vimeo.com/666431148).)
+```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse-cap)"}
 knitr::include_graphics("./figures/case_chocolates_inverse.png")
@@
 
-<!-- segue and selection for chocolates inverse case -->
-It would also be interesting to explore the inverse case.
-This would describe which features lead to a milk chocolate being misclassified as dark and how those attributions differ from the previous misclassification.
-Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
-This exploration is shown in Figure \@ref(fig:casechocolatesinverse).
+<!-- treeSHAP pca, for inverse case -->
+The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
+Looking at the attribution projection, this is found to be the case.
+Previously, Fiber was essential while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
 
-<!-- treeSHAP pca, PCP, tour -->
-The difference of position in the tree SHAP PCA with the previous case is quite large; this gives an approximate feel that the attribution should be quite different.
-Looking at the initial contribution, we find this to be the case.
-Previously Fiber was very important while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat are highly attributed here, while unimportant in the preceding case.
-
-Comparing the attribution with the CI (dotted line), discrepancies in Sodium and Fiber are identified.
-We opt to vary Sodium.
-The instance looks slightly more like its observed milk than predicted dark chocolate even in the initial projection.
+<!-- PCP, for inverse case -->
+Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
+The contribution of Sodium is selected to be varied.
+Even in the initial projection, the instance looks slightly more like its observed milk than predicted dark chocolate.
 The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
@@
 
+<!-- Data description -->
 The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
-From a correlation matrix plot, nine higher-level skill groupings were identified and aggregated.
-A random forest model if fit from these aggregations and regress player wages [2020 euros].
+Nine higher-level skill groupings were identified and aggregated from highly correlated features.
+A random forest model is fit from these predictors, regressing player wages [2020 euros].
 The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
 Continuing from the exploration in section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players.
-We compare a leading offensive fielder (L.
-Messi) with that of a top defensive fielder (V.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
+A leading offensive fielder (L.
+Messi) is compared with a top defensive fielder (V.
 van Dijk).
@@
 
-(ref:casefifa) FIFA 2020 data, a random forest model, regresses wages [2020 Euros] from nine aggregated skill measurements.
+(ref:casefifa-cap) FIFA 2020 data, a random forest model, regresses wages [2020 Euros] from nine aggregated skill measurements.
 The PI is a star offensive player (L.
@@
 van Dijk).
-We remove three features with low attribution from both players.
+Three features with low attribution from both players are removed.
 The attribution projection starts with the selected instance on the right.
-We vary the contribution from defense (`def`).
+The contribution to defense (`def`) is varied.
 The star offensive player is not distinguished in the horizontal direction.
@@
 (A video of the animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).)
-```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa)"}
+```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa-cap)"}
 knitr::include_graphics("./figures/case_fifa.png")
@@
 
-With figure \@ref(fig:casefifa), we will test the support of the local explanation.
+Figure \@ref(fig:casefifa), tests the support of the local explanation.
 Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
-If either of them were rotated out, the other will be rotated into frame, maintaining a far-right position.
-However, when we vary the defensive skills, the other skills are rotated out of the frame.
+If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
+However, Increasing the contribution of a variable with low importance would rotate both features out of the frame.
 
+The contribution from `def` will be varied to contrast with offensive skills.
 As the contribution of defensive skills increases, Messi's is no longer separated from the group.
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
-Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
+A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
 
-(ref:caseames) Exploring an instance with a large residual as the PI from fitting sales price [USD] to other variables in the Ames housing 2018.
+(ref:caseames-cap) Exploring an instance with a large residual as the PI from fitting sales price [USD] to other variables in the Ames housing 2018.
 The sale price of the PI was under-predicted.
@@
 In the attribution projection, the PI has a higher sales price than the CI.
-Reducing the contribution of lot area brings these two prices in line.
+Reducing the contribution of Lot Area brings these two prices in line.
 This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
-That is, the large residual is due to a lack of factoring in the lot area for the prediction of PI's sales price.
+That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
 (A video showing the animation is at [vimeo.com/666431134](https://vimeo.com/666431134).)
-```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
+```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames-cap)"}
 knitr::include_graphics("./figures/case_ames2018.png")
@@
 
-Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction with a large Lot Area contribution.
+Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-The attribution projection is places instances with high living areas to the right.
-We control the contribution of this feature.
-As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
+The attribution projection places instances with high Living Areas to the right.
+The contribution of Living Area contrasts the contribution of this feature.
+As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
 This large of importance in the Living Area is relatively uncommon.
-Boosting tree models may be more resilient to such an under-prediction as up-weighting this residual would force its inclusion in the final model.
+Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
 
@@
 
-<!-- summarise methods -->
-<!-- summarise findings -->
-<!-- suggest next steps -->
-
-New tools to assist with the interpretability of black-box models is increasingly important.
-This paper provides a technique that builds on local interpretations to explore the feature importance local to an instance.
+There is a clear need to extend the interpretability of black-box models.
+This chapter provides a technique that builds on local interpretations to explore the feature importance local to an instance.
 The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
 Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
-A global view shows the data space, explanation space, model fit.
+A global view shows the data space, explanation space, residual plot.
 The user can interactively select instances to compare, contrast, and study further.
+Then the radial tour is used to explore the feature sensitivity identified by the attribution projection.
 
-The approach has been illustrated on four data examples using random forest models and using the tree SHAP local explanation.
+This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
 In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
-This was verified by making a follow-up plot of the data, including this variable.
-In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
-For the FIFA example, we show how low Messi's salary would go if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used.
+This was verified by making a follow-up plot of the data.
+The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
+In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used by the random forest model.
 
-The approach is manually intensive and thus only feasible for investigating a few instances.
-The approach is to select instances that the model has not done well and compare it with an instance where it did fit well.
+This analysis is manually intensive and thus only feasible for investigating a few instances.
+The approach selects instances that the model has not done well to predict and compare it with an instance where it did fit well.
 The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and responses in order to cross-check interpretations made from the cheem viewer.
+It can be helpful to make additional plots of the features and responses to cross-check interpretations made from the cheem viewer.
 This methodology provides an additional tool in the box for studying model fitting.
@@
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ (tree based models from __randomForest__, __ranger__, __gbm__, __xgboost__, __lightgbm__, or __catboost__).
+In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ [tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively].
 Tree SHAP was selected because of its computational efficiency.
 The SHAP and oscillation explanations could be added with the use of the `DALEX::explain()` and would be an excellent direction to extend the work [@biecek_dalex_2018; @biecek_explanatory_2021].
-
-<!-- Attribution space does tend to be more separable than data space at least in the first couple principal components.
-Different statistics should be explored to convey a clearer understanding what structure the explanation distinguishes.
-It is curious that the attribution space the same dimensionality of the data with a seemingly better separation.
-It sounds like a candidate to fit another model.
-However, such an attempt should want to be vigilant not to over-fit the training data.-->
 

### 2022-02-27 _paper.tex
text changes from thesis, Needed: references and figure changes

--- 2022-02-07 _paper.tex
@@
 \begin{abstract}
-The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction with a point estimate of the linear feature importance in the vicinity of each instance.
-These can be considered linear projections and can be further explored interactively to understand better the interaction between features used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response features.
+The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
+This trade-off has led to the emergence of eXplainable AI (XAI).
+XAI attempts to shed light on how models use predictors to arrive at a prediction with \emph{local explanations}, a point estimate of the linear feature importance in the vicinity of one instance.
+These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
 The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
@@
 Section \ref{sec:tour} explains the animations of continuous linear projections.
-Section \ref{sec:cheemviwer} discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure.
-Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative response features.
-We conclude with Section \ref{sec:cheemdiscussion} of the insights gained and directions that might be explored in the future.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
+Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative output.
+Section \ref{sec:cheemdiscussion} concludes with the insights gained and possible directions to explore in the future.
 
@@
 SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The explanations of SHAP almost all refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing the performance of this player in the presence or absence of other players.
-Strumbelj and Kononenko (2010) introduced the use of SHAP for local explanations in ML models.
+The explanations of SHAP refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
+Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
 The attribution of feature importance depends on the sequence of the included features.
@@
 The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
-Partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), which helps to understand the relative importance of each feature in the context of all other candidate features.
+Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), helps to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 
-\caption{Illustration SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills.
+\caption{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
 A star offensive and defensive player are compared, L.
@@
 van Dijk, respectively.
-Panel (a) shows breakdown plots of three sequences of the features, order, and magnitude change.
+Panel (a) shows breakdown plots of three sequences of the features.
+The sequence of the variables impacts the magnitude of their attribution.
 Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player.
+Reaction skills are important for both players.
 Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
@@
 
-For our application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
-Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; we illustrate random forests.
-The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) will have coefficients varied to further scrutinize the feature contributions.
-
-Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), we use FIFA data to illustrate SHAP use.
+Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), FIFA data is used to illustrate SHAP.
 Consider soccer data from the FIFA 2020 season (Leone 2020).
 There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing wages {[}2020 Euros{]} from the skill measures.
-We then extract the SHAP values of a star offensive player (L.
+A random forest model is fit regressing player's wages {[}2020 Euros{]} from their skill measurements.
+The SHAP values are compared for a star offensive player (L.
 Messi) and defensive player (V.
@@
 The results are displayed in Figure \ref{fig:shapdistrbd}.
-We expect to see a difference in the attribution of the feature importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets.
-Plot (b) is a modified breakdown plot (Gosiewska and Biecek 2019) where the order of features is fixed, so the two instances can be more easily compared.
-
+A difference in the attribution of the feature importance across the two positions of the players can be expected.
+This would be interpreted as how a player's salary depends on this combination of skill sets.
+Panel (a) is a modified breakdown plot (Gosiewska and Biecek 2019) where three sequences of features are presented, so the two instances can be more easily compared.
+The magnitude of the contributions depends on the sequence in which they appear.
+Panel (b) shows the differences of the player's median values of over 25 such sequences.
 In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
 In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
+
+For the application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
+Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
+Tree SHAP is only compatible with tree-based models; random forests are used for illustration.
+The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) that will be used to explore the sensitivity of the feature contributions.
 
@@
 
-To explore the local explanations, an ensemble of plots (Unwin and Valero-Mora 2018) is provided, called the \emph{cheem viewer}.
-There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values, and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
-In addition, there are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
+To explore the local explanations, coordinated views (\textbf{roberts\_state\_2007?}) (also known as ensemble graphics, Unwin and Valero-Mora 2018) are provided in the \emph{cheem viewer} application.
+There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
+There are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
 There are different plots used for the categorical and quantitative responses.
@@
 
-The global view provides the context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
+The global view provides a context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
 Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
@@
 
-A visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
-These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are useful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a), to help identify instances poorly predicted by the model.
-For classification tasks, misclassified instances are circled in red if applicable.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:classificationcase}a,c).
+Visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
+These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
+In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify instances poorly predicted by the model.
+For classification tasks, misclassified instances are circled in red.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
 
-While the comparison of these spaces is interesting, a main purpose of the global view is to enable the selection of instances to explore the local explanations.
+While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of instances to explore the local explanations.
 The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
@@
 The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
-These are depicted as a vertical parallel coordinate plot, where each line connects one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
-The attribution projections of the PI and CI are shown as dashed and dotted lines, respectively.
-From this plot, we would read the range of importances across all instances can be interpreted.
+These are depicted as vertical parallel coordinate plots (Ocagne 1885) on the basis biplot (Gabriel 1971).
+1D biplot display depicts the values of the current basis as bars.
+The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
+The attribution projections of the PI and CI are shown as dashed and dotted lines.
+From this plot, the range and density of the importance across all instances can be interpreted.
 For classification, one would look at differences between groups on any feature.
-For example, Figure \ref{fig:classificationcase}e suggests that \texttt{b\_l} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting important for some instances and not important for other instances.
+For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
+For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting they are important for some instances and not important for other instances.
 
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection is an approximation of the feature importance for the prediction of this instance.
+The attribution projection is approximates the feature importance for predicting this instance.
 This combination of features best explains the difference between the mean response and an instance's predicted value.
@@
 
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f), and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The user uses the radial tour to vary the contribution of the selected feature between 0-1.
+The radial tour varies the contribution of the selected feature between 0-1.
+This is viewed as an animation of the projections from many intermediate bases.
 Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
-For classification, if the separation between classes diminishes when the feature contribution is reduced then this suggests that the feature is important for the class separation.
-For regression, if the relationship scatterplot weakens when the feature contribution is reduced then this suggests that the feature is important for accurately predicting the response.
+For classification, if the separation between classes diminishes when the feature contribution is reduced, this suggests that the feature is important for class separation.
+For regression, if the relationship scatterplot weakens when the feature contribution is reduced, indicating that the feature is important for accurately predicting the response.
 
@@
 The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
-The radial tour is 1D, plotted as a density where color indicates class.
-A scroll bar here enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
+The radial tour is 1D and displays as density where color indicates class.
+An animation slider enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
 
@@
 The black bar is the variable being changed in the radial tour.
-Panel (f) is the resulting projection of the data indicated as density in the classification case.}\label{fig:classificationcase}
+Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
 \end{figure}
@@
 
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance, with similar feature values, as CI is a useful way to understand how the model is failing or not.
+Selecting an inaccurately predicted instance as PI and an accurately predicted instance with similar feature values as CI is a helpful way to understand how the model is failing or not.
 The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation, to help with understand where the model fits better or worse.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation aid in understanding the structure identified in these spaces.
 
@@
 
-\caption{Overview of the cheem viewer for regression task highlighting the differences from the classification task and interactive features.
-Panel (a) PCA of the data and attributions spaces, (b), residual plot, predictions by observed values.
-Four points are selected points and highlighted in the PC spaces and tabularly displayed.
+\caption{Overview of the cheem viewer for regression  and illustration of interactive features.
+Panel (a) PCA of the data and attributions spaces, (b) residual plot, predictions by observed values.
+Four selected points are highlighted in the PC spaces and tabularly displayed.
 Coloring on a statistic (c) highlights structure organized in the attribution space.
@@
 Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (left and right).}\label{fig:regressioncase}
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (middle and right).}\label{fig:regressioncase}
 \end{figure}
@@
 The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible.
-The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.
+These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a \& d).
+The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
 
 A tooltip displays instance number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots.
-The information corresponding to the selected points is populated on a dynamic table.
-These interactions aid exploration of the spaces and, finally, identification of a primary and comparison instance.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
+The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
+These interactions aid exploration of the spaces and, finally, the identification of primary and comparison instances.
 
@@
     \item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
-    \item \textbf{Model and explanation:} any model can be used with this method.
-Currently, we apply random forest models via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
-We use modest hyperparameters, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (Kominsarczyk et al.
+    \item \textbf{Model:} any model and compatible explanation could be explored with this method.
+Currently, random forest models are applied via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
+Modest hyperparameters are used, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+    \item \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (Kominsarczyk et al.
 2021).
-This implementation aggregates exhaustively over all trees' attribution, and we opt not to fit interactions of features.
-    \item \textbf{Cheem view:} after the model and full explanation space are calculated, we scale each feature by standard deviations away from the mean to achieve common support for visuals.
+This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of features.
+    \item \textbf{Cheem viewer:} after the model and full explanation space are calculated, each feature is scaled by standard deviations away from the mean to achieve common support for visuals.
 Statistics for mapping to color are computed on the scaled spaces.
-Interactive tabular display reports the original values.
 \end{itemize}
 
-The time to preprocess the data will vary significantly with the model and local explanation.
-For reference, the FIFA data, 5000 instances of nine explanatory features, took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each instance took 270 seconds combined.
+The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
+For reference, the FIFA data contained 5000 instances of nine explanatory features took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each instance took 270 seconds total.
 PCA and statistics of the features and attributions took 2.8 seconds.
@@
 An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP a good candidate to start.
+Its reduced computational complexity makes tree SHAP an excellent candidate to start.
 (Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
@@
 The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
-Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive HTML widgets with \textbf{plotly} (Sievert 2020).
-\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a huge boon to understanding local explanations and how to apply them.
+Both views are created first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive \texttt{html} widgets with \textbf{plotly} (Sievert 2020).
+\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a boon to understanding local explanations and how to apply them.
 
@@
 
-To process your own data, you will need to use the \texttt{treeshap} package, which can be installed from GitHub:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}ModelOriented/treeshap\textquotesingle{}}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Follow the examples provided with the package to compute the local explainers, (see \texttt{?cheem\_ls}).
-The application expects the return of call from \texttt{cheem\_ls()} which is then saved to an .rds file with \texttt{saveRDS()}.
+Follow the examples provided with the package to compute the local explainers (see \texttt{?cheem\_ls}).
+The application uses the returned objects from a from \texttt{cheem\_ls()} call.
 Alternatively, the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances, and we will use the four physical measurements, bill length (\texttt{b\_l}), bill depth (\texttt{b\_d}), flipper length (\texttt{f\_l}), body mass (\texttt{wgt}), for this illustration.
+After removing incomplete instances, there are 333 instances of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), body mass (\texttt{bm}), for this illustration.
 A random forest model was fit with species as the response feature.
@@
 \caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) Chinstrap (orange) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
-The radial view shows varying the contribution of \texttt{f\_l} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
+The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+The radial view shows varying the contribution of \texttt{fl} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
 (The animation of the radial tour is at \href{https://vimeo.com/666431172}{vimeo.com/666431172}.)}\label{fig:casepenguins}
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{./figures/case_penguins_BlFl}
+{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}
 
@@
 \caption{Checking what is learned from the cheem viewer.
-This is a plot of flipper length (\texttt{f\_l}) and bill length (\texttt{b\_l}), where an asterisk highlights the PI.
+This is a plot of flipper length (\texttt{fl}) and bill length (\texttt{bl}), where an asterisk highlights the PI.
 A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small \texttt{f\_l} length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
+The PI has an unusually small \texttt{fl} length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
 \end{figure}
@@
 The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance are mostly \texttt{b\_l} and \texttt{b\_d} with small contributions of \texttt{f\_l} and \texttt{wgt}.
+The important features identified by SHAP in the (wrong) prediction for this instance are mostly \texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
 This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
 That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this, we use the radial tour to vary the contribution of flipper length (\texttt{f\_l}).
+To explore this, the radial tour is used to vary the contribution of flipper length (\texttt{fl}).
 (In our exploration, this was the third feature explored.
-It is typically useful to explore the features with larger contributions, here \texttt{b\_l} and \texttt{b\_d}, but when doing this, nothing was revealed about how the PI differed from other Gentoos).
-On varying \texttt{f\_l} as it contributes more to the projection (b, right), we see that more, and more, this penguin looks like a Chinstrap.
-This suggests that \texttt{f\_l} should be considered an important feature for explaining the (wrong) prediction.
+It is typically helpful to explore the features with more significant contributions, here \texttt{bl} and \texttt{bd}.
+Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
+On varying \texttt{fl} as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
+This suggests that \texttt{fl} should be considered an important feature for explaining the (wrong) prediction.
 
-Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{f\_l}) is important for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and we can see that the PI is closer to the Chinstrap group in these two features, mostly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that its a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
+Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{fl}) is vital for the confusion of the PI as a Chinstrap.
+Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two features, mainly because it has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
 
@@
 Dark chocolate is considered healthier than milk.
-The data was collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
+The data were collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
 The data is available in the \textbf{cheem} package.
-A random forest model is used for the classification of chocolate type.
+A random forest model is used for the classification of chocolate types.
 
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate nutritionally more like milk should not be considered a healthy alternative.
-It is interesting to explore which of the nutritional features contribute most to misclassification.
+A reason to do this is that a dark chocolate, nutritionally more like milk, should not be considered a healthy alternative.
+It is interesting to explore which nutritional features contribute most to misclassification.
 
@@
 
-\caption{Examining the local interpretation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark more than milk, which suggests that the local explanation does not help understand the prediction for this instance.
-So, we vary the contribution of Sugar -- reducing it corresponds primarily with increasing Fiber.
-When Sugar is zero, Fiber is contributing strongly towards the left.
+\caption{Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
+So, the contribution of Sugar is varied --- reducing it corresponds primarily with increasing Fiber.
+When Sugar is zero, Fiber contributes strongly towards the left.
 In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
@@
 It is compared with a CI that is a correctly classified dark chocolate (instance 7).
-The PCA plot, and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
-The misclassifications are more apparent in the observed vs predicted plot, and can be seen to be mistaken in both ways: milk to dark and dark to milk.
+The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
+The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
 
@@
 The opposite would be interpreting a point with high values in this plot.
-The dark chocolates (orange) are mostly on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
+The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
 The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
 
-The instance of interest (dashed line) can be viewed against the comparison instance (dotted line).
-Now one needs to pay different attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the instance of interest.
+The PI (dashed line) can be viewed against the CI (dotted line).
+Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the PI.
+The feature contributions to the two different predictions can be quickly compared in the parallel coordinate plot.
+The PI differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
 
-We can quickly compare the feature contributions to the two different predictions from the parallel coordinate plot.
-The instance of interest differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
+From the density plot, which is the attribution projection corresponding to the PI, both instances are more like dark chocolates.
+Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
+When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
 
-From the density plot, which is the attribution projection corresponding to the instance of interest, both instances are more like dark chocolates.
-If we vary the contribution of Sugars and altogether remove Sugars from the projection, this is where the difference becomes apparent.
-When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
+It would also be interesting to explore an inverse misclassification.
+In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
+Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (instance 71).
+The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
 
@@
 In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed feature importance, both with low values relative to other milk chocolates.
-We suspect the lack of use of these variables is what has contributed to the mistake, so we vary the contribution of Sodium.
-If Sodium had a larger contribution to the prediction (like in this view), the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolatesinverse}
+Sodium and Fiber have the largest differences in attributed feature importance, with low values relative to other milk chocolates.
+The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
+If Sodium had a larger contribution to the prediction (like in this view).
+the PI would look more like other milk chocolates.
+(A video of the tour animation can be found at \href{https://vimeo.com/666431148}{vimeo.com/666431148}.)}\label{fig:casechocolatesinverse}
 \end{figure}
 
-It would also be interesting to explore the inverse case.
-This would describe which features lead to a milk chocolate being misclassified as dark and how those attributions differ from the previous misclassification.
-Chocolate 84 is just this case, and we compare it with a correctly predicted milk chocolate (instance 71).
-This exploration is shown in Figure \ref{fig:casechocolatesinverse}.
+The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
+Looking at the attribution projection, this is found to be the case.
+Previously, Fiber was essential while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
 
-The difference of position in the tree SHAP PCA with the previous case is quite large; this gives an approximate feel that the attribution should be quite different.
-Looking at the initial contribution, we find this to be the case.
-Previously Fiber was very important while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat are highly attributed here, while unimportant in the preceding case.
-
-Comparing the attribution with the CI (dotted line), discrepancies in Sodium and Fiber are identified.
-We opt to vary Sodium.
-The instance looks slightly more like its observed milk than predicted dark chocolate even in the initial projection.
+Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
+The contribution of Sodium is selected to be varied.
+Even in the initial projection, the instance looks slightly more like its observed milk than predicted dark chocolate.
 The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
@@
 The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-From a correlation matrix plot, nine higher-level skill groupings were identified and aggregated.
-A random forest model if fit from these aggregations and regress player wages {[}2020 euros{]}.
+Nine higher-level skill groupings were identified and aggregated from highly correlated features.
+A random forest model is fit from these predictors, regressing player wages {[}2020 euros{]}.
 The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
 Continuing from the exploration in section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players.
-We compare a leading offensive fielder (L.
-Messi) with that of a top defensive fielder (V.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
+A leading offensive fielder (L.
+Messi) is compared with a top defensive fielder (V.
 van Dijk).
@@
 van Dijk).
-We remove three features with low attribution from both players.
+Three features with low attribution from both players are removed.
 The attribution projection starts with the selected instance on the right.
-We vary the contribution from defense (\texttt{def}).
+The contribution to defense (\texttt{def}) is varied.
 The star offensive player is not distinguished in the horizontal direction.
@@
 
-With figure \ref{fig:casefifa}, we will test the support of the local explanation.
+Figure \ref{fig:casefifa}, tests the support of the local explanation.
 Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
-If either of them were rotated out, the other will be rotated into frame, maintaining a far-right position.
-However, when we vary the defensive skills, the other skills are rotated out of the frame.
+If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
+However, Increasing the contribution of a variable with low importance would rotate both features out of the frame.
 
+The contribution from \texttt{def} will be varied to contrast with offensive skills.
 As the contribution of defensive skills increases, Messi's is no longer separated from the group.
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
-Using interaction from the global view, we select a house with an extreme negative residual and an accurate instance close to it in the data.
+A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
 
@@
 In the attribution projection, the PI has a higher sales price than the CI.
-Reducing the contribution of lot area brings these two prices in line.
+Reducing the contribution of Lot Area brings these two prices in line.
 This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
-That is, the large residual is due to a lack of factoring in the lot area for the prediction of PI's sales price.
+That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
 (A video showing the animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.)}\label{fig:caseames}
@@
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with a large Lot Area contribution.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-The attribution projection is places instances with high living areas to the right.
-We control the contribution of this feature.
-As the contribution of lot area decreases, the predictive power decreases for the PI, while the CI remains stationary.
+The attribution projection places instances with high Living Areas to the right.
+The contribution of Living Area contrasts the contribution of this feature.
+As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
 This large of importance in the Living Area is relatively uncommon.
-Boosting tree models may be more resilient to such an under-prediction as up-weighting this residual would force its inclusion in the final model.
+Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
 
@@
 
-New tools to assist with the interpretability of black-box models is increasingly important.
-This paper provides a technique that builds on local interpretations to explore the feature importance local to an instance.
+There is a clear need to extend the interpretability of black-box models.
+This chapter provides a technique that builds on local interpretations to explore the feature importance local to an instance.
 The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
 Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
-A global view shows the data space, explanation space, model fit.
+A global view shows the data space, explanation space, residual plot.
 The user can interactively select instances to compare, contrast, and study further.
+Then the radial tour is used to explore the feature sensitivity identified by the attribution projection.
 
-The approach has been illustrated on four data examples using random forest models and using the tree SHAP local explanation.
+This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
 In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
-This was verified by making a follow-up plot of the data, including this variable.
-In the chocolates example, we showed how a dark chocolate was misclassified primarily due to it value on Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
-For the FIFA example, we show how low Messi's salary would go if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used.
+This was verified by making a follow-up plot of the data.
+The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
+In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used by the random forest model.
 
-The approach is manually intensive and thus only feasible for investigating a few instances.
-The approach is to select instances that the model has not done well and compare it with an instance where it did fit well.
+This analysis is manually intensive and thus only feasible for investigating a few instances.
+The approach selects instances that the model has not done well to predict and compare it with an instance where it did fit well.
 The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and responses in order to cross-check interpretations made from the cheem viewer.
+It can be helpful to make additional plots of the features and responses to cross-check interpretations made from the cheem viewer.
 This methodology provides an additional tool in the box for studying model fitting.
@@
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{randomForest}, \textbf{ranger}, \textbf{gbm}, \textbf{xgboost}, \textbf{lightgbm}, or \textbf{catboost}).
+In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} \textbf{greenwell\_gbm\_2020?}; \textbf{shi\_lightgbm\_2022?}; Liaw and Wiener 2002; \textbf{wright\_ranger\_2017?}; \textbf{chen\_xgboost\_2021?}, respectively).
 Tree SHAP was selected because of its computational efficiency.
@@
 
+\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
+Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
+
 \leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
@@
 
+\leavevmode\hypertarget{ref-ocagne_coordonnees_1885}{}%
+Ocagne, Maurice d'.
+1885.
+\emph{Coordonnées Parallèles Et Axiales.
+{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul Graphique déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
+
 \leavevmode\hypertarget{ref-ribeiro_why_2016}{}%

### 2022-02-28 _paper.rmd

@@
 
-# The cheem viewer {#sec:cheemviwer}
+# The cheem viewer {#sec:cheemviewer}
 
@@
 
-Follow the examples provided with the package to compute the local explainers (see `?cheem_ls`).
-The application uses the returned objects from a from `cheem_ls()` call.
-Alternatively, the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/).
+Alternatively,
+
+- a version of the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/). 
+- the development version of the package is available at https://github.com/nspyrison/cheem, and 
+- documentation of the package can be found at https://nspyrison.github.io/cheem/.
+
+Follow the examples provided with the package to compute the local explainers (using `?cheem_ls`).
+The application expects the output returned by `cheem_ls()`, saved to an .rds file with `saveRDS()`, to be uploaded.
 
@@
 
-(ref:casefifa-cap) FIFA 2020 data, a random forest model, regresses wages [2020 Euros] from nine aggregated skill measurements.
-The PI is a star offensive player (L.
-Messi) compared with a top defensive player (V.
+(ref:casefifa-cap) Exploring the wages [euros] relative to skill measurements in the FIFA 2020 data.
+Star offensive player (L.
+Messi) is the PI and he is compared with a top defensive player (V.
 van Dijk).
-Three features with low attribution from both players are removed.
-The attribution projection starts with the selected instance on the right.
-The contribution to defense (`def`) is varied.
-The star offensive player is not distinguished in the horizontal direction.
-At this point, defensive players have been rotated to the highest horizontal value.
+The attribution projection is shown at left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
+Defense (`def`) is the chosen feature to vary.
+It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
+The increased contribution in defense comes at the expense of offensive and reaction skills.
+The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
 (A video of the animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).)
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+A random forest model has fitted this price [USD] with the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
 Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.

### 2022-02-28 _paper.tex

@@
 
-\hypertarget{sec:cheemviwer}{%
-\section{The cheem viewer}\label{sec:cheemviwer}}
+\hypertarget{sec:cheemviewer}{%
+\section{The cheem viewer}\label{sec:cheemviewer}}
 
-To explore the local explanations, coordinated views (\textbf{roberts\_state\_2007?}) (also known as ensemble graphics, Unwin and Valero-Mora 2018) are provided in the \emph{cheem viewer} application.
+To explore the local explanations, coordinated views (Roberts 2007) (also known as ensemble graphics, Unwin and Valero-Mora 2018) are provided in the \emph{cheem viewer} application.
 There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
@@
 
-Follow the examples provided with the package to compute the local explainers (see \texttt{?cheem\_ls}).
-The application uses the returned objects from a from \texttt{cheem\_ls()} call.
-Alternatively, the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
+Alternatively,
+
+\begin{itemize}
+\tightlist
+\item
+  a version of the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
+\item
+  the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
+\item
+  documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+\end{itemize}
+
+Follow the examples provided with the package to compute the local explainers (using \texttt{?cheem\_ls}).
+The application expects the output returned by \texttt{cheem\_ls()}, saved to an .rds file with \texttt{saveRDS()}, to be uploaded.
 
@@
 
-\caption{FIFA 2020 data, a random forest model, regresses wages {[}2020 Euros{]} from nine aggregated skill measurements.
-The PI is a star offensive player (L.
-Messi) compared with a top defensive player (V.
+\caption{Exploring the wages {[}euros{]} relative to skill measurements in the FIFA 2020 data.
+Star offensive player (L.
+Messi) is the PI and he is compared with a top defensive player (V.
 van Dijk).
-Three features with low attribution from both players are removed.
-The attribution projection starts with the selected instance on the right.
-The contribution to defense (\texttt{def}) is varied.
-The star offensive player is not distinguished in the horizontal direction.
-At this point, defensive players have been rotated to the highest horizontal value.
+The attribution projection is shown at left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
+Defense (\texttt{def}) is the chosen feature to vary.
+It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
+The increased contribution in defense comes at the expense of offensive and reaction skills.
+The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
 (A video of the animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.)}\label{fig:casefifa}
@@
 The remaining are 338 house sales.
-A random forest model has regressed this price with the features Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+A random forest model has fitted this price {[}USD{]} with the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
 Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
@@
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} \textbf{greenwell\_gbm\_2020?}; \textbf{shi\_lightgbm\_2022?}; Liaw and Wiener 2002; \textbf{wright\_ranger\_2017?}; \textbf{chen\_xgboost\_2021?}, respectively).
+In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} Greenwell et al.
+2020; Shi et al.
+2022; Liaw and Wiener 2002; Wright and Ziegler 2017; Chen et al.
+2021, respectively).
 Tree SHAP was selected because of its computational efficiency.
@@
 
+\leavevmode\hypertarget{ref-chen_xgboost_2021}{}%
+Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al.
+2021.
+{``Xgboost: {Extreme} {Gradient} {Boosting}.''} \url{https://CRAN.R-project.org/package=xgboost}.
+
 \leavevmode\hypertarget{ref-cook_manual_1997}{}%
@@
 Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+
+\leavevmode\hypertarget{ref-greenwell_gbm_2020}{}%
+Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M. Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression} {Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
 
@@
 
+\leavevmode\hypertarget{ref-roberts_state_2007}{}%
+Roberts, Jonathan C.
+2007.
+{``State of the Art: {Coordinated} \& Multiple Views in Exploratory Visualization.''} In \emph{Fifth International Conference on Coordinated and Multiple Views in Exploratory Visualization ({CMV} 2007)}, 61--71.
+IEEE.
+
 \leavevmode\hypertarget{ref-shapley_value_1953}{}%
 Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
+
+\leavevmode\hypertarget{ref-shi_lightgbm_2022}{}%
+Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas Finley, Taifeng Wang, et al.
+2022.
+{``Lightgbm: {Light} {Gradient} {Boosting} {Machine}.''} \url{https://CRAN.R-project.org/package=lightgbm}.
 
@@
 \leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
-Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble Graphics.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
+Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
@@
 
+\leavevmode\hypertarget{ref-wright_ranger_2017}{}%
+Wright, Marvin N., and Andreas Ziegler.
+2017.
+{``Ranger: {A} {Fast} {Implementation} of {Random} {Forests} for {High} {Dimensional} {Data} in {C}++ and {R}.''} \emph{Journal of Statistical Software} 77 (1): 1--17.
+\url{https://doi.org/10.18637/jss.v077.i01}.
+
 \end{CSLReferences}

### 2022-03-01 _paper.rmd
latex bullets to rmarkdown bullets for consistency with thesis.

@@
 
-(ref:citeRf) [@liaw_classification_2002]
-(ref:citeTs) [@kominsarczyk_treeshap_2021]
-\begin{itemize}
-	\item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
-	\item \textbf{Model:} any model and compatible explanation could be explored with this method.
-Currently, random forest models are applied via the package \textbf{randomForest} (ref:citeRf), compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-	\item \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (ref:citeTs).
-This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of features.
-	\item \textbf{Cheem viewer:} after the model and full explanation space are calculated, each feature is scaled by standard deviations away from the mean to achieve common support for visuals.
+<!-- Switch to rmd bullets, consistent with thesis -->
+- __Data:__ predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
+- __Model:__ any model and compatible explanation could be explored with this method.
+Currently, random forest models are applied via the package __randomForest__ [@liaw_classification_2002], compatibility tree SHAP.
+Modest hyperparameters are used, namely: 125 trees, number variables randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+- __Local explanation:__ Tree SHAP is calculated for _each_ observation using the package __treeshap__ [@kominsarczyk_treeshap_2021].
+This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of variables.
+- __Cheem viewer:__ after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
 Statistics for mapping to color are computed on the scaled spaces.
-\end{itemize}
 

### 2022-03-01 _paper.tex
latex bullets to rmarkdown bullets for consistency with thesis.

@@
 \begin{itemize}
-    \item \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
-    \item \textbf{Model:} any model and compatible explanation could be explored with this method.
+\tightlist
+\item
+  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
+\item
+  \textbf{Model:} any model and compatible explanation could be explored with this method.
 Currently, random forest models are applied via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, number features randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-    \item \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} instance using the package \textbf{treeshap} (Kominsarczyk et al.
+Modest hyperparameters are used, namely: 125 trees, number variables randomly sampled at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
+\item
+  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} (Kominsarczyk et al.
 2021).
-This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of features.
-    \item \textbf{Cheem viewer:} after the model and full explanation space are calculated, each feature is scaled by standard deviations away from the mean to achieve common support for visuals.
+This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of variables.
+\item
+  \textbf{Cheem viewer:} after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
 Statistics for mapping to color are computed on the scaled spaces.

### 2022-03-14 _paper.rmd
wresting jmlr2e.sty

@@
 author:
+- Nicholas Spyrison 
+#^[Monash University, nicholas.spyrison@monash.edu, ORCiD: 0000-0002-8417-0212]
+- Dianne Cook
+- Kimbal Marriott
+#^[ORCiD: 0000-0002-9813-0377]
   #- name: "Nicholas Spyrison"
+  #   affiliation: Monash University
   #   address: |
@@
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
-#documentclass: jmlr
+#documentclass: jmlr2e
 #classoption: article
@@
   ## pdf_document: requires \ref{} 
-  #pdf_document:
-  ## bookdown::pdf_document2: allows \@ref()
-  bookdown::pdf_document2:
+  pdf_document:
+  ## bookdown::pdf_document2: allows \@ref(), but doesn't add authors....
+  #bookdown::pdf_document2:
+    #includes:
+    #  in_header: template/jmlr2e.sty
     keep_tex: true
@@
     fig_caption: yes
-    #includes:
-    #  in_header: template/jmlr2e.sty
 bibliography: cheem_paper.bib
@@
   error      = FALSE,
-  cache      = TRUE,
+  cache      = FALSE,
   cache.lazy = FALSE)
@@
 <!-- Introduce explanatory vs predictive modeling -->
-There are different reasons and emphases to fit a model.
-@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomize modeling based on its purpose; _explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more on the predictions of out-of-sample instances.
+There are different reasons and emphases when considering to fit a model.
+@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomize models based on their purpose.
+_Explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more narrowly on the performance of some objective function.
 The intended use has important implications for model selection and development.
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate nonlinear models.
-The use of black-box models is becoming increasingly common, but not without their share of controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
-However, the loss of interpretation presents a challenge.
+Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network to name a few [@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995, receptively].
+
+<!-- nonlinear leading to interpretability issues -->
+Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
+This difficulty in interpreting the terms in complex nonlinear models sometimes leads to them being referred to as black box models.
+Despite the potentially better performance of nonlinear models, their use is not without controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
+And the loss of interpretation presents a challenge.
 
 <!-- Interpretability & baises -->
-Interpretability is vital for exploring and protecting against potential biases (e.g., sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]) in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
-Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
-It is also generally important for many problems, where it is not enough to predict accurately.
-Still, one must be able to explain which predictors are most responsible for generating a response value.
+Interpretability is vital for exploring and protecting against potential biases in any model (e.g.
+sex -- @dastin_amazon_2018; @duffy_apple_2019, race -- @larson_how_2016, and age -- @diaz_addressing_2018).
+For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
+This bias is then built into the model.
+Variable-level (feature-level) interpretability of models is essential in evaluating and addressing such biases.
 
 <!-- Interpretability & data drift -->
-Another concern is data drift, a shift in support or domain of the explanatory features (variable or predictors).
-Nonlinear models are typically more sensitive and do not extrapolate well outside the training data domain.
-Better interpretability of the model means more transparency where models' predictions may be plausible or completely unreliable.
+Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors).
+Some nonlinear models are sensitive to this and do not extrapolate well outside the support of the training data [@espadoto_toward_2021].
+Maintaining variable interpretability is also essential to address issues arising from data drift.
 
 <!-- Local explanations -->
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
+Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models [@adadi_peeking_2018; @arrieta_explainable_2020].
 A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
@@
 It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
-<!-- However, this navigation is relatively time-consuming due to the vast volume of the display space.
-It is advisable to use this method to explore the sensitivity of the feature contribution to a previously identified feature of interest.
-In this case, the projection of the normalized explanations is the feature of interest.-->
 

### 2022-03-14 _paper.tex
wresting jmlr2e.sty

@@
   pdftitle={Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections},
-  pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
+  pdfauthor={Nicholas Spyrison; Dianne Cook; Kimbal Marriott},
+  pdfkeywords={R, XAI, explanable artificial intelligence, black-box
+models, local explanations, grand tour, manual tour, radial tour},
   colorlinks=true,
@@
 
-\title{Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections}
-\author{}
-\date{\vspace{-2.5em}}
+\title{Exploring Local Explanations of Nonlinear Models Using Animated
+Linear Projections}
+\author{Nicholas Spyrison \and Dianne Cook \and Kimbal Marriott}
+\date{}
 
@@
 \begin{abstract}
-The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
-This trade-off has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction with \emph{local explanations}, a point estimate of the linear feature importance in the vicinity of one instance.
-These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
-The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
+The increased predictive power of nonlinear models comes at the cost of
+interpretability of its terms. This trade-off has led to the emergence
+of eXplainable AI (XAI). XAI attempts to shed light on how models use
+predictors to arrive at a prediction with \emph{local explanations}, a
+point estimate of the linear feature importance in the vicinity of one
+instance. These can be considered linear projections and can be further
+explored to understand better the interactions between features used to
+make predictions across the predictive model surface. Here we describe
+interactive linear interpolation used for exploration at any instance
+and illustrate with examples with categorical (penguin species,
+chocolate types) and quantitative (soccer/football salaries, house
+prices) output. The methods are implemented in the \textbf{R} package
+\textbf{cheem}, available on CRAN.
 \end{abstract}
@@
 
-There are different reasons and emphases to fit a model.
-Breiman (2001), reiterated by Shmueli (2010), taxonomize modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample instances.
-The intended use has important implications for model selection and development.
-In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate nonlinear models.
-The use of black-box models is becoming increasingly common, but not without their share of controversy (O'Neil 2016; Kodiyan 2019).
-However, the loss of interpretation presents a challenge.
-
-Interpretability is vital for exploring and protecting against potential biases (e.g., sex (Dastin 2018; Duffy 2019), race (Larson et al.
-2016), and age (Díaz et al.
-2018)) in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
-Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
-It is also generally important for many problems, where it is not enough to predict accurately.
-Still, one must be able to explain which predictors are most responsible for generating a response value.
-
-Another concern is data drift, a shift in support or domain of the explanatory features (variable or predictors).
-Nonlinear models are typically more sensitive and do not extrapolate well outside the training data domain.
-Better interpretability of the model means more transparency where models' predictions may be plausible or completely unreliable.
-
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
-A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
-
-In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
-Lee et al.
-2021) is a sequence of linear projections of data onto a lower-dimensional space.
-Tours are viewed as an animation over minor changes to the projection basis.
-Structure in a projection can then be explored visually to see which features contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information about the structure and features of the object.
-
-There are various types of tours distinguished by the generation of projection bases.
-In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected feature.
-Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths (Y.
-D.
-Lee et al.
-2013; da Silva, Cook, and Lee 2021).
-
-We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in R package \textbf{cheem}.
-Example datasets are provided to illustrate usage for classification and regression tasks.
-
-Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
-It ignores correlations of the features and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object.
-A constraint in our approach is that the basis must remain orthonormal.
-When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
-
-The remainder of this paper is organized as follows.
-The following Section, \ref{sec:explanations}, covers the background of the local explanation and the traditional visuals produced.
-Section \ref{sec:tour} explains the animations of continuous linear projections.
-Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
-Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative output.
-Section \ref{sec:cheemdiscussion} concludes with the insights gained and possible directions to explore in the future.
+There are different reasons and emphases when considering to fit a
+model. Breiman (2001b), reiterated by Shmueli (2010), taxonomize models
+based on their purpose. \emph{Explanatory} modeling is done for some
+inferential purpose, while \emph{predictive} modeling focuses more
+narrowly on the performance of some objective function. The intended use
+has important implications for model selection and development. In
+explanatory modeling, interpretability is vital for drawing inferential
+conclusions. Nonlinear models range from additive models with at least
+one polynomial term to more complex machine learning models such as
+random forest, support-vector machine, or neural network to name a few
+(Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995,
+receptively).
+
+Nonlinear models have many or complex interactive terms, which cause an
+opaqueness to the interpretation of the variables. This difficulty in
+interpreting the terms in complex nonlinear models sometimes leads to
+them being referred to as black box models. Despite the potentially
+better performance of nonlinear models, their use is not without
+controversy (O'Neil 2016; Kodiyan 2019). And the loss of interpretation
+presents a challenge.
+
+Interpretability is vital for exploring and protecting against potential
+biases in any model (e.g.~sex -- Dastin (2018); Duffy (2019), race --
+Larson et al. (2016), and age -- Díaz et al. (2018)). For instance,
+models regularly pick up on biases in the training data where such
+classes correlate with changes in the response variable. This bias is
+then built into the model. Variable-level (feature-level)
+interpretability of models is essential in evaluating and addressing
+such biases.
+
+Another concern is data drift, where a shift in the range of the
+explanatory variables (features or predictors). Some nonlinear models
+are sensitive to this and do not extrapolate well outside the support of
+the training data (Espadoto et al. 2021). Maintaining variable
+interpretability is also essential to address issues arising from data
+drift.
+
+Explainable Artificial Intelligence (XAI) is an emerging field of
+research that tries to increase the interpretability of black-box models
+(Adadi and Berrada 2018; Arrieta et al. 2020). A common approach is to
+use \emph{local explanations}, which attempt to approximate linear
+feature importance at the location of each instance (observation) or the
+predictions at a specific point in the data domain. Because these are
+point-specific, it is challenging to visualize them to comprehensively
+understand a model.
+
+In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and
+Asimov 1986; S. Lee et al. 2021) is a sequence of linear projections of
+data onto a lower-dimensional space. Tours are viewed as an animation
+over minor changes to the projection basis. Structure in a projection
+can then be explored visually to see which features contribute to the
+formation of that structure. The intuition is similar to watching the
+shadow of a hidden 3D object change as the object is rotated; watching
+the shape of the shadow change conveys information about the structure
+and features of the object.
+
+There are various types of tours distinguished by the generation of
+projection bases. In a \emph{manual} tour (Cook and Buja 1997; Spyrison
+and Cook 2020), the path is defined by changing the contribution of a
+selected feature. Applying tours to models has been done in a couple of
+contexts. Specifically for exploring various statistical model fits and
+classification boundaries (Wickham, Cook, and Hofmann 2015), and using
+tree- and forest-based approaches as a projection pursuit index to
+generate a tour basis paths (Y. D. Lee et al. 2013; da Silva, Cook, and
+Lee 2021).
+
+We use the radial manual tour to scrutinize a local explanation in our
+proposed approach. Additional interactivity allows the user to identify
+an instance of interest then explore its local explanation by changing
+feature contribution with the radial tour. The methods are implemented
+in R package \textbf{cheem}. Example datasets are provided to illustrate
+usage for classification and regression tasks.
+
+Using a radial tour can be compared with counterfactual, what-if
+analysis, such as \emph{ceteris paribus} profiles (Biecek 2020). Ceteris
+paribus, is Latin for ``other things held constant'' or ``all else
+unchanged.'' These profiles show how an instance's prediction would
+change from a marginal change in one explanatory feature given that
+other features are held constant. It ignores correlations of the
+features and imagines a case that was not observed. In contrast, our
+approach is a geometric explanation of the factual; it varies
+contributions of the features by rotating the basis, a reorientation of
+the data object. A constraint in our approach is that the basis must
+remain orthonormal. When the contribution of one feature decreases, the
+contributions of others necessarily increase such that there is a
+complete component in that direction. This also ensures that what is
+seen is strictly a low-dimensional projection from high-dimensions and
+is thus an interpretable visualization.
+
+The remainder of this paper is organized as follows. The following
+Section, @ref(sec:explanations), covers the background of the local
+explanation and the traditional visuals produced. Section @ref(sec:tour)
+explains the animations of continuous linear projections. Section
+@ref(sec:cheemviewer) discusses the visual layout in the graphical user
+interface, how it facilitates analysis, data preprocessing, and package
+infrastructure. Then Section @ref(sec:casestudies) illustrates the
+application to supervised learning with categorical and quantitative
+output. Section @ref(sec:cheemdiscussion) concludes with the insights
+gained and possible directions to explore in the future.
 
@@
 
-Consider a highly nonlinear model.
-It can be hard to determine whether small changes in a feature's value will make a class prediction change group or identify which features contribute to an extreme residual.
-Local explanations shed light on these situations by approximating linear feature importance in the vicinity of a single instance.
-
-A comprehensive summary of the taxonomy and literature of explanation techniques is provided in Figure 6 of Arrieta et al.
-(2020).
-It includes a large number of model-specific explanations such as deepLIFT (Shrikumar et al.
-2016; Shrikumar, Greenside, and Kundaje 2017), a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (Lundberg and Lee 2017), and their variants are popular.
-
-These instance-level explanations are used in various ways depending on the data.
-In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels (Simonyan, Vedaldi, and Zisserman 2014).
-For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
-2018).
-In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction (Ribeiro, Singh, and Guestrin 2016).
-
-SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The explanations of SHAP refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
-Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
-The attribution of feature importance depends on the sequence of the included features.
-The SHAP values are the mean contributions over different feature sequences.
-The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
-Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), helps to understand the relative importance of each feature in the context of all other candidate features.
+Consider a highly nonlinear model. It can be hard to determine whether
+small changes in a feature's value will make a class prediction change
+group or identify which features contribute to an extreme residual.
+Local explanations shed light on these situations by approximating
+linear feature importance in the vicinity of a single instance.
+
+A comprehensive summary of the taxonomy and literature of explanation
+techniques is provided in Figure 6 of Arrieta et al. (2020). It includes
+a large number of model-specific explanations such as deepLIFT
+(Shrikumar et al. 2016; Shrikumar, Greenside, and Kundaje 2017), a
+popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic explanations, of which LIME, (Ribeiro,
+Singh, and Guestrin 2016) SHAP, (Lundberg and Lee 2017), and their
+variants are popular.
+
+These instance-level explanations are used in various ways depending on
+the data. In image classification, where pixels correspond to
+predictors, saliency maps overlay or offset a heatmap indicating
+important pixels (Simonyan, Vedaldi, and Zisserman 2014). For instance,
+pixels corresponding to snow may be highlighted when distinguishing if a
+picture contains a wolf or husky. In text analysis, word-level
+contextual sentiment analysis can be used to highlight the sentiment and
+magnitude of influential words (Vanni et al. 2018). In the case of
+numeric regression, they are used to explain feature additive
+contributions from the model intercept to the instance's prediction
+(Ribeiro, Singh, and Guestrin 2016).
+
+SHaply Additive exPlanations (SHAP) quantifies the feature contributions
+of one instance by examining the effect of other features on the
+predictions. The explanations of SHAP refer to Shapley (1953)'s method
+to evaluate an individual's contribution to cooperative games by
+assessing this player's performance in the presence or absence of other
+players. Strumbelj and Kononenko (2010) introduced SHAP for local
+explanations in ML models. The attribution of feature importance depends
+on the sequence of the included features. The SHAP values are the mean
+contributions over different feature sequences. The approach is related
+to partial dependence plots (Molnar 2020), used to explain the effect of
+a feature by predicting the response for a range of values on this
+feature after fixing the value of all other features to their mean.
+Though partial dependence plots are a global approximation of the
+feature importance, while SHAP is specific to one instance. It could
+also be considered similar to examining the coefficients from all
+subsets regression, as described in Wickham, Cook, and Hofmann (2015),
+helps to understand the relative importance of each feature in the
+context of all other candidate features.
 
@@
 
-Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), FIFA data is used to illustrate SHAP.
-Consider soccer data from the FIFA 2020 season (Leone 2020).
-There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing player's wages {[}2020 Euros{]} from their skill measurements.
-The SHAP values are compared for a star offensive player (L.
-Messi) and defensive player (V.
-van Dijk).
-The results are displayed in Figure \ref{fig:shapdistrbd}.
-A difference in the attribution of the feature importance across the two positions of the players can be expected.
-This would be interpreted as how a player's salary depends on this combination of skill sets.
-Panel (a) is a modified breakdown plot (Gosiewska and Biecek 2019) where three sequences of features are presented, so the two instances can be more easily compared.
-The magnitude of the contributions depends on the sequence in which they appear.
-Panel (b) shows the differences of the player's median values of over 25 such sequences.
-In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
-In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
-
-For the application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
-Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; random forests are used for illustration.
-The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) that will be used to explore the sensitivity of the feature contributions.
+Following the use case \emph{Explanatory Model Analysis} (Biecek and
+Burzykowski 2021), FIFA data is used to illustrate SHAP. Consider soccer
+data from the FIFA 2020 season (Leone 2020). There are 5000 instances of
+9 skill measures (after aggregating highly correlated features). A
+random forest model is fit regressing player's wages {[}2020 Euros{]}
+from their skill measurements. The SHAP values are compared for a star
+offensive player (L. Messi) and defensive player (V. van Dijk). The
+results are displayed in Figure @ref(fig:shapdistrbd). A difference in
+the attribution of the feature importance across the two positions of
+the players can be expected. This would be interpreted as how a player's
+salary depends on this combination of skill sets. Panel (a) is a
+modified breakdown plot (Gosiewska and Biecek 2019) where three
+sequences of features are presented, so the two instances can be more
+easily compared. The magnitude of the contributions depends on the
+sequence in which they appear. Panel (b) shows the differences of the
+player's median values of over 25 such sequences. In summary, these
+plots highlight how local explanations bring interpretability to a
+model, at least in the vicinity of their instances. In this instance,
+two players with different positions receive different profiles of
+feature importance to explain the prediction of their wages.
+
+For the application, we use \emph{tree SHAP}, a variant of SHAP that
+enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
+Instead of aggregating over sequences of the features, tree SHAP
+calculates instance-level feature importance by exploring the structure
+of the decision trees. Tree SHAP is only compatible with tree-based
+models; random forests are used for illustration. The following section
+will use normalized SHAP values as a projection basis (call this the
+\emph{attribution projection}) that will be used to explore the
+sensitivity of the feature contributions.
 
@@
 
-A \emph{tour} enables viewing of high-dimensional data by animating many linear projections with small incremental changes.
-It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
-There are various types of tours that are distinguished by how the paths are generated (S.
-Lee et al.
-2021; Cook et al.
-2008).
-
-The manual tour (Cook and Buja 1997) defines its path by changing a selected feature's contribution to a basis to allow the feature to contribute more or less to the projection.
-The requirement constrains the contribution of all other features that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
-It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis (Karwowski 2006).
-
-A version of the manual tour called a \emph{radial tour} is implemented in Spyrison and Cook (2020) and forms the basis of the new work.
-In a radial tour, the selected feature can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and interactive re-calculation to focus on a different feature.
+A \emph{tour} enables viewing of high-dimensional data by animating many
+linear projections with small incremental changes. It is achieved by
+following a path of linear projections (bases) of high-dimensional
+space. One of the features of the tour is the object permanence of the
+data points; one can track the relative change of instances in time and
+gain information about the relationships between points across multiple
+features. There are various types of tours that are distinguished by how
+the paths are generated (S. Lee et al. 2021; Cook et al. 2008).
+
+The manual tour (Cook and Buja 1997) defines its path by changing a
+selected feature's contribution to a basis to allow the feature to
+contribute more or less to the projection. The requirement constrains
+the contribution of all other features that a basis needs to be
+orthonormal (column correspond to vectors, with unit length, and
+orthogonal to each other). The manual tour is primarily used to assess
+the importance of a feature to structure visible in a projection. It
+also lends itself to pre-computation queued in advance or computed
+on-the-fly for human-in-the-loop analysis (Karwowski 2006).
+
+A version of the manual tour called a \emph{radial tour} is implemented
+in Spyrison and Cook (2020) and forms the basis of the new work. In a
+radial tour, the selected feature can change its magnitude of
+contribution but not its angle; it must move along the direction of its
+original contribution. The implementation allows for pre-computation and
+interactive re-calculation to focus on a different feature.
 
@@
 
-To explore the local explanations, coordinated views (Roberts 2007) (also known as ensemble graphics, Unwin and Valero-Mora 2018) are provided in the \emph{cheem viewer} application.
-There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
-There are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
-There are different plots used for the categorical and quantitative responses.
-Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
+To explore the local explanations, coordinated views (Roberts 2007)
+(also known as ensemble graphics, Unwin and Valero-Mora 2018) are
+provided in the \emph{cheem viewer} application. There are two primary
+plots: the \textbf{global view} to give the context of all of the SHAP
+values and the \textbf{radial tour view} to explore the local
+explanations with user-controlled rotation. There are numerous user
+inputs, including feature selection for the radial tour and instance
+selection for making comparisons. There are different plots used for the
+categorical and quantitative responses. Figures
+@ref(fig:classificationcase) and @ref(fig:regressioncase) are
+screenshots showing the cheem viewer for the two primary tasks:
+classification (categorical response) and regression (quantitative
+response).
 
@@
 
-The global view provides a context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, each a vector of \(p\) values.
-
-Visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
-These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify instances poorly predicted by the model.
-For classification tasks, misclassified instances are circled in red.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
-
-While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of instances to explore the local explanations.
-The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
-These instances are highlighted as asterisk and \(\times\), respectively.
+The global view provides a context of all instances and facilitates the
+exploration of the separability of the data- and attribution-spaces.
+Both of these spaces are of dimension \(n\times p\), where \(n\) is the
+number of instances and \(p\) is the number of predictors. The
+attribution space corresponds to the local explanations for each
+instance, each a vector of \(p\) values.
+
+Visualization is provided by the first two principal components of the
+data (left) and the attribution (middle) spaces. These single 2D
+projections will not reveal all of the structure of higher-dimensional
+space, but they are helpful visual summaries. In addition, a plot of the
+observed against predicted response values is also provided (Figures
+@ref(fig:classificationcase)b, @ref(fig:regressioncase)a) to help
+identify instances poorly predicted by the model. For classification
+tasks, misclassified instances are circled in red. Linked brushing
+between the plots is provided, and a tabular display of selected points
+helps to facilitate exploration of the spaces and the model (shown in
+Figures @ref(fig:regressioncase)d).
+
+While the comparison of these spaces is interesting, the primary purpose
+of the global view is to enable the selection of instances to explore
+the local explanations. The projection attribution of the primary
+instance (PI) is examined and typically viewed with an optional
+comparison instance (CI). These instances are highlighted as asterisk
+and \(\times\), respectively.
 
@@
 
-The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
-These are depicted as vertical parallel coordinate plots (Ocagne 1885) on the basis biplot (Gabriel 1971).
-1D biplot display depicts the values of the current basis as bars.
-The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
-The attribution projections of the PI and CI are shown as dashed and dotted lines.
-From this plot, the range and density of the importance across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature.
-For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting they are important for some instances and not important for other instances.
-
-The overlaid bars on the parallel coordinate plot represent the attribution projection of the PI.
-(Remember that the PI is interactively selected from the global view).
-The attribution projection is approximates the feature importance for predicting this instance.
-This combination of features best explains the difference between the mean response and an instance's predicted value.
-It is not an indication of the local shape of the model surface.
-That is, it is not some indication of the tangent to the curve at this point.
-
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
-The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The radial tour varies the contribution of the selected feature between 0-1.
-This is viewed as an animation of the projections from many intermediate bases.
-Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
-For classification, if the separation between classes diminishes when the feature contribution is reduced, this suggests that the feature is important for class separation.
-For regression, if the relationship scatterplot weakens when the feature contribution is reduced, indicating that the feature is important for accurately predicting the response.
+The local explanations for all observations are normalized (squared sum
+of values adds to 1), and thus, the relative importance of features can
+be compared across all instances. These are depicted as vertical
+parallel coordinate plots (Ocagne 1885) on the basis biplot (Gabriel
+1971). 1D biplot display depicts the values of the current basis as
+bars. The parallel coordinate overlays lines connecting one instance's
+feature attribution (Figures @ref(fig:classificationcase)e and
+@ref(fig:regressioncase)e). The attribution projections of the PI and CI
+are shown as dashed and dotted lines. From this plot, the range and
+density of the importance across all instances can be interpreted. For
+classification, one would look at differences between groups on any
+feature. For example, Figure @ref(fig:classificationcase)e suggests that
+\texttt{bl} is important for distinguishing the green class from the
+other two. For regression, one might generally observe which features
+have low values for all instances (not important), for example,
+\texttt{BMI} and \texttt{pwr} in Figure @ref(fig:regressioncase)e, and
+which have a range of high and low values (e.g.~\texttt{off},
+\texttt{def}) suggesting they are important for some instances and not
+important for other instances.
+
+The overlaid bars on the parallel coordinate plot represent the
+attribution projection of the PI. (Remember that the PI is interactively
+selected from the global view). The attribution projection is
+approximates the feature importance for predicting this instance. This
+combination of features best explains the difference between the mean
+response and an instance's predicted value. It is not an indication of
+the local shape of the model surface. That is, it is not some indication
+of the tangent to the curve at this point.
+
+The attribution projection of the PI is the initial 1D basis in a radial
+tour, displayed as a density plot for a categorical response (Figure
+@ref(fig:classificationcase)f) and as scatterplots for a quantitative
+response (Figure @ref(fig:regressioncase)f). The PI and CI are indicated
+by vertical dashed and dotted lines, respectively. The radial tour
+varies the contribution of the selected feature between 0-1. This is
+viewed as an animation of the projections from many intermediate bases.
+Doing so tests the sensitivity of structure (class separation or
+strength of relationship) to the feature's contribution. For
+classification, if the separation between classes diminishes when the
+feature contribution is reduced, this suggests that the feature is
+important for class separation. For regression, if the relationship
+scatterplot weakens when the feature contribution is reduced, indicating
+that the feature is important for accurately predicting the response.
 
@@
 
-Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI makes it easier to examine the features most responsible for the error.
-The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
-The radial tour is 1D and displays as density where color indicates class.
-An animation slider enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
+Selecting a misclassified instance as PI and a correctly classified
+point nearby in data space as CI makes it easier to examine the features
+most responsible for the error. The global view (Figure
+@ref(fig:classificationcase)c) displays the model confusion matrix. The
+radial tour is 1D and displays as density where color indicates class.
+An animation slider enables the user to vary the contribution of each
+feature to explore the sensitivity of the separation to that feature.
 
@@
 
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance with similar feature values as CI is a helpful way to understand how the model is failing or not.
-The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation aid in understanding the structure identified in these spaces.
-
-In the radial tour view, the observed response and the residuals (vertical) are plotted against the attribution projection of the PI (horizontal).
-The attribution projection can be interpreted similarly to the predicted value from the global view plot.
-It represents a linear combination of the features, and a good fit would be indicated when there is a strong relationship with the observed values.
-This can be viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a feature is varied, if the value of the PI does not change much, it would indicate that the prediction for this instance is NOT sensitive to that feature.
-Conversely, if the predicted value varies substantially, the prediction is very sensitive to that feature, suggesting that the feature is very important for the PI's prediction.
+Selecting an inaccurately predicted instance as PI and an accurately
+predicted instance with similar feature values as CI is a helpful way to
+understand how the model is failing or not. The global view (Figure
+@ref(fig:regressioncase)a) shows a scatterplot of the observed vs
+predicted values, which should exhibit a strong relationship if the
+model is a good fit. The points can be colored by a statistic, residual,
+a measure of outlyingness (log Mahalanobis distance), or correlation aid
+in understanding the structure identified in these spaces.
+
+In the radial tour view, the observed response and the residuals
+(vertical) are plotted against the attribution projection of the PI
+(horizontal). The attribution projection can be interpreted similarly to
+the predicted value from the global view plot. It represents a linear
+combination of the features, and a good fit would be indicated when
+there is a strong relationship with the observed values. This can be
+viewed as a local linear approximation if the fitted model is nonlinear.
+As the contribution of a feature is varied, if the value of the PI does
+not change much, it would indicate that the prediction for this instance
+is NOT sensitive to that feature. Conversely, if the predicted value
+varies substantially, the prediction is very sensitive to that feature,
+suggesting that the feature is very important for the PI's prediction.
 
@@
 
-The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a \& d).
-The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
-
-A tooltip displays instance number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
-The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
-These interactions aid exploration of the spaces and, finally, the identification of primary and comparison instances.
+The application has several reactive inputs that affect the data used,
+aesthetic display, and tour manipulation. These reactive inputs make the
+software flexible and extensible (Figure @ref(fig:classificationcase)a
+\& d). The application also has more exploratory interactions to help
+link points across displays, reveal structures found in different
+spaces, and access the original data.
+
+A tooltip displays instance number/name and classification information
+while the cursor hovers over a point. Linked brushing allows the
+selection of points (left click and drag) where those points will be
+highlighted across plots (Figure @ref(fig:classificationcase)a \& b).
+The information corresponding to the selected points is populated on a
+dynamic table (Figure @ref(fig:classificationcase)d). These interactions
+aid exploration of the spaces and, finally, the identification of
+primary and comparison instances.
 
@@
 
-It is vital to mitigate the render time of visuals, especially when users may want to iterate many explorations.
-All computational operations should be prepared before runtime.
-The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
-Below discusses the steps and details of the reprocessing.
+It is vital to mitigate the render time of visuals, especially when
+users may want to iterate many explorations. All computational
+operations should be prepared before runtime. The work remaining when an
+application is run solely reacts to inputs and rendering visuals and
+tables. Below discusses the steps and details of the reprocessing.
 
@@
 \item
-  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
+  \textbf{Data:} predictors and response are unscaled complete numerical
+  matrix. Most models and local explanations are scale-invariant.
 \item
-  \textbf{Model:} any model and compatible explanation could be explored with this method.
-Currently, random forest models are applied via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, number variables randomly sampled at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
+  \textbf{Model:} any model and compatible explanation could be explored
+  with this method. Currently, random forest models are applied via the
+  package \textbf{randomForest} (Liaw and Wiener 2002), compatibility
+  tree SHAP. Modest hyperparameters are used, namely: 125 trees, number
+  variables randomly sampled at each split, mtry = \(\sqrt{p}\) or
+  \(p/3\) for classification and regression, and minimum size of
+  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
+  classification and regression.
 \item
-  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} (Kominsarczyk et al.
-2021).
-This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of variables.
+  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
+  observation using the package \textbf{treeshap} (Kominsarczyk et al.
+  2021). This implementation aggregates exhaustively over all trees'
+  attribution and does not fit interactions of variables.
 \item
-  \textbf{Cheem viewer:} after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
-Statistics for mapping to color are computed on the scaled spaces.
+  \textbf{Cheem viewer:} after the model and full explanation space are
+  calculated, each variable is scaled by standard deviations away from
+  the mean to achieve common support for visuals. Statistics for mapping
+  to color are computed on the scaled spaces.
 \end{itemize}
 
-The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
-For reference, the FIFA data contained 5000 instances of nine explanatory features took 2.5 seconds to fit a random forest model of modest hyperparameters.
+The time to preprocess the data will vary significantly with the
+complexity of the model and local explanation. For reference, the FIFA
+data contained 5000 instances of nine explanatory features took 2.5
+seconds to fit a random forest model of modest hyperparameters.
 Extracting the tree SHAP values of each instance took 270 seconds total.
 PCA and statistics of the features and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
-An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-(Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
+These runtimes were from a non-parallelized R session on a modern
+laptop, but suffice to say that most of the time will be spent on the
+local attribution. An increase in model complexity or data
+dimensionality will quickly become an obstacle. Its reduced
+computational complexity makes tree SHAP an excellent candidate to
+start. (Alternatively, the package \textbf{fastshap} (Greenwell 2020)
+claims extremely low runtimes, attributed to fewer calls to the
+prediction function, partial implementation in C++, and efficient use of
+logical subsetting.)
 
@@
 
-The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002) and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
-2021).
-The application was made with \textbf{shiny} (Chang et al.
-2021).
-The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
-Both views are created first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive \texttt{html} widgets with \textbf{plotly} (Sievert 2020).
-\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a boon to understanding local explanations and how to apply them.
+The above-described method and application are implemented as an
+open-source \textbf{R} package, \textbf{cheem} available on
+\href{https://CRAN.R-project.org/package=cheem}{CRAN}. Preprocessing was
+facilitated with models created via \textbf{randomForest} (Liaw and
+Wiener 2002) and explanations calculated with \textbf{treeshap}
+(Kominsarczyk et al. 2021). The application was made with \textbf{shiny}
+(Chang et al. 2021). The tour visual is built with \textbf{spinifex}
+(Spyrison and Cook 2020). Both views are created first with
+\textbf{ggplot2} (Wickham 2016) and then rendered as interactive
+\texttt{html} widgets with \textbf{plotly} (Sievert 2020).
+\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model
+Analysis} (Biecek and Burzykowski 2021) were a boon to understanding
+local explanations and how to apply them.
 
 \hypertarget{installation-and-getting-started}{%
-\subsection{Installation and getting started}\label{installation-and-getting-started}}
-
-The package can be installed from CRAN, and the shiny app can be run, using the following \textbf{R} code:
+\subsection{Installation and getting
+started}\label{installation-and-getting-started}}
+
+The package can be installed from CRAN, and the shiny app can be run,
+using the following \textbf{R} code:
 
@@
 \item
-  a version of the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
+  a version of the cheem viewer shiny app can be directly accessed at
+  \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
 \item
-  the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
+  the development version of the package is available at
+  \url{https://github.com/nspyrison/cheem}, and
 \item
-  documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+  documentation of the package can be found at
+  \url{https://nspyrison.github.io/cheem/}.
 \end{itemize}
 
-Follow the examples provided with the package to compute the local explainers (using \texttt{?cheem\_ls}).
-The application expects the output returned by \texttt{cheem\_ls()}, saved to an .rds file with \texttt{saveRDS()}, to be uploaded.
+Follow the examples provided with the package to compute the local
+explainers (using \texttt{?cheem\_ls}). The application expects the
+output returned by \texttt{cheem\_ls()}, saved to an .rds file with
+\texttt{saveRDS()}, to be uploaded.
 
@@
 
-To illustrate the cheem method, it is applied to modern datasets, two classification examples and then two of regression.
+To illustrate the cheem method, it is applied to modern datasets, two
+classification examples and then two of regression.
 
 \hypertarget{palmer-penguin-species-classification}{%
-\subsection{Palmer penguin, species classification}\label{palmer-penguin-species-classification}}
-
-The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), body mass (\texttt{bm}), for this illustration.
-A random forest model was fit with species as the response feature.
+\subsection{Palmer penguin, species
+classification}\label{palmer-penguin-species-classification}}
+
+The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst,
+Hill, and Gorman 2020) was collected on three species of penguins
+foraging near Palmer Station, Antarctica. The data was publicly
+available to substitute for the overly-used iris data and is quite
+similar in form. After removing incomplete instances, there are 333
+instances of four physical measurements, bill length (\texttt{bl}), bill
+depth (\texttt{bd}), flipper length (\texttt{fl}), body mass
+(\texttt{bm}), for this illustration. A random forest model was fit with
+species as the response feature.
+
+(ref:casepenguins-cap) Examining the SHAP values for a random forest
+model classifying Palmer penguin species. The PI is a Gentoo (purple)
+penguin that is misclassified as a Chinstrap (orange), marked as an
+asterisk in (a), and the dashed vertical line in (b). The radial view
+shows varying the contribution of \texttt{fl} from the initial
+attribution projection (b, left), which produces a linear combination
+where the PI is more probably (higher density value) a Chinstrap than a
+Gentoo (b, right). (The animation of the radial tour is at
+\href{https://vimeo.com/666431172}{vimeo.com/666431172}.)
 
@@
 
-\caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
-The radial view shows varying the contribution of \texttt{fl} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at \href{https://vimeo.com/666431172}{vimeo.com/666431172}.)}\label{fig:casepenguins}
+\caption{(ref:casepenguins-cap)}\label{fig:casepenguins}
 \end{figure}
 
-Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
-Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
+Figure @ref(fig:casepenguins) shows plots from the cheem viewer for
+exploring the random forest model on the penguins data. Panel (a) shows
+the global view, and panel (b) shows several 1D projections generated
+with the radial tour. Penguin 243, a Gentoo (purple), is the PI because
+it has been misclassified as a Chinstrap (orange).
+
+(ref:casepenguinsblfl-cap) Checking what is learned from the cheem
+viewer. This is a plot of flipper length (\texttt{fl}) and bill length
+(\texttt{bl}), where an asterisk highlights the PI. A Gentoo (purple)
+misclassified as a Chinstrap (orange). The PI has an unusually small
+\texttt{fl} length which is why it is confused with a Chinstrap.
 
@@
 
-\caption{Checking what is learned from the cheem viewer.
-This is a plot of flipper length (\texttt{fl}) and bill length (\texttt{bl}), where an asterisk highlights the PI.
-A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small \texttt{fl} length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
+\caption{(ref:casepenguinsblfl-cap)}\label{fig:casepenguinsblfl}
 \end{figure}
 
-There is more separation visible in the attribution space than the data space, as would be expected.
-The predicted vs observed plot reveals a handful of misclassified instances.
-A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
-The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
-The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
-
-The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance are mostly \texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
-That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this, the radial tour is used to vary the contribution of flipper length (\texttt{fl}).
-(In our exploration, this was the third feature explored.
-It is typically helpful to explore the features with more significant contributions, here \texttt{bl} and \texttt{bd}.
-Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
-On varying \texttt{fl} as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
-This suggests that \texttt{fl} should be considered an important feature for explaining the (wrong) prediction.
-
-Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{fl}) is vital for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two features, mainly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
+There is more separation visible in the attribution space than the data
+space, as would be expected. The predicted vs observed plot reveals a
+handful of misclassified instances. A Gentoo that has been wrongly
+labeled as a Chinstrap is selected for illustration. The PI is a
+misclassified point (represented by the asterisk in the global view and
+a dashed vertical line in the tour view). The CI is a correctly
+classified point (represented by an \(\times\) and a vertical dotted
+line).
+
+The radial tour starts from the attribution projection of the
+misclassified instance (b, left). The important features identified by
+SHAP in the (wrong) prediction for this instance are mostly \texttt{bl}
+and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
+This projection is a view where the Gentoo (purple) looks much more
+likely for this instance than Chinstrap. That is, this combination of
+features is not particularly useful because the PI looks very much like
+other Gentoo penguins. To explore this, the radial tour is used to vary
+the contribution of flipper length (\texttt{fl}). (In our exploration,
+this was the third feature explored. It is typically helpful to explore
+the features with more significant contributions, here \texttt{bl} and
+\texttt{bd}. Still, when doing this, nothing was revealed about how the
+PI differed from other Gentoos). On varying \texttt{fl} as it
+contributes increasingly to the projection (b, right), more and more,
+this penguin looks like a Chinstrap. This suggests that \texttt{fl}
+should be considered an important feature for explaining the (wrong)
+prediction.
+
+Figure @ref(fig:casepenguinsblfl) confirms that flipper length
+(\texttt{fl}) is vital for the confusion of the PI as a Chinstrap. Here,
+flipper length and body length are plotted, and the PI can be seen to be
+closer to the Chinstrap group in these two features, mainly because it
+has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard instance to account for
+as decision trees can only partition only vertical and horizontal lines.
 
 \hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
-
-The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
-Dark chocolate is considered healthier than milk.
-The data were collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
-The data is available in the \textbf{cheem} package.
-A random forest model is used for the classification of chocolate types.
-
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate, nutritionally more like milk, should not be considered a healthy alternative.
-It is interesting to explore which nutritional features contribute most to misclassification.
+\subsection{Chocolates, milk/dark chocolate
+classification}\label{chocolates-milkdark-chocolate-classification}}
+
+The chocolates dataset consists of 88 instances of ten nutritional
+measurements determined from their labels and labeled as either milk or
+dark. Dark chocolate is considered healthier than milk. The data were
+collected by students during the Iowa State University class STAT503
+from nutritional information from the manufacturer's website and
+normalized to 100g equivalents. The data is available in the
+\textbf{cheem} package. A random forest model is used for the
+classification of chocolate types.
+
+It could be interesting to examine the nutritional properties of any
+dark chocolates that have been misclassified as milk. A reason to do
+this is that a dark chocolate, nutritionally more like milk, should not
+be considered a healthy alternative. It is interesting to explore which
+nutritional features contribute most to misclassification.
+
+(ref:casechocolates-cap) Examining the local explanation for a PI which
+is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more
+like dark than milk, which suggests that the local explanation does not
+help understand the prediction for this instance. So, the contribution
+of Sugar is varied --- reducing it corresponds primarily with increasing
+Fiber. When Sugar is zero, Fiber contributes strongly towards the left.
+In this particular view, the PI is closer to the bulk of the milk
+chocolates, suggesting that the prediction put a lot of importance on
+Fiber. This chocolate is a rare dark chocolate without any Fiber leading
+to it being mistaken for a milk chocolate. (A video of the tour
+animation can be found at
+\href{https://vimeo.com/666431143}{vimeo.com/666431143}.)
 
@@
 
-\caption{Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
-So, the contribution of Sugar is varied --- reducing it corresponds primarily with increasing Fiber.
-When Sugar is zero, Fiber contributes strongly towards the left.
-In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
-This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolates}
+\caption{(ref:casechocolates-cap)}\label{fig:casechocolates}
 \end{figure}
 
-This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (instance 22).
-It is compared with a CI that is a correctly classified dark chocolate (instance 7).
-The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
-The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
-
-The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
-In the density plot, instances on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be interpreting a point with high values in this plot.
-The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
-The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
-
-The PI (dashed line) can be viewed against the CI (dotted line).
-Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the PI.
-The feature contributions to the two different predictions can be quickly compared in the parallel coordinate plot.
-The PI differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
-
-From the density plot, which is the attribution projection corresponding to the PI, both instances are more like dark chocolates.
-Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
-When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
-
-It would also be interesting to explore an inverse misclassification.
-In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
-Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (instance 71).
-The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
+This type of exploration is shown in Figure @ref(fig:casechocolates),
+where a chocolate labeled dark but predicted to be milk is chosen as the
+PI (instance 22). It is compared with a CI that is a correctly
+classified dark chocolate (instance 7). The PCA plot and the tree SHAP
+PCA plots (a) show a big difference between the two chocolate types but
+with confusion for a handful of instances. The misclassifications are
+more apparent in the observed vs predicted plot and can be seen to be
+mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars,
+and Calories are most responsible for its incorrect prediction. The way
+to read this plot is to see that Fiber has a large negative value while
+Sugars and Calories have reasonably large positive values. In the
+density plot, instances on the very left of the display would have high
+values of Fiber (matching the negative projection coefficient) and low
+values of Sugars and Calories. The opposite would be interpreting a
+point with high values in this plot. The dark chocolates (orange) are
+primarily on the left, and this is a reason why they are considered to
+be healthier: high fiber and low sugar. The density of milk chocolates
+is further to the right, indicating that they generally have low fiber
+and high sugar.
+
+The PI (dashed line) can be viewed against the CI (dotted line). Now,
+one needs to pay attention to the parallel plot of the SHAP values,
+which are local to a particular instance, and the density plot, which is
+the same projection of all instances as specified by the SHAP values of
+the PI. The feature contributions to the two different predictions can
+be quickly compared in the parallel coordinate plot. The PI differs with
+the comparison primarily on the Fiber feature, which suggests that this
+is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding
+to the PI, both instances are more like dark chocolates. Varying the
+contribution of Sugars and altogether removing it from the projection is
+where the difference becomes apparent. When primarily Fiber is examined,
+instance 22 looks more like a milk chocolate.
+
+It would also be interesting to explore an inverse misclassification. In
+this case, a milk chocolate is selected while it was misclassified as a
+dark chocolate. Chocolate 84 is just this case and is compared with a
+correctly predicted milk chocolate (instance 71). The corresponding
+global view and radial tour frames are shown in Figure
+@ref(fig:casechocolatesinverse).
+
+(ref:casechocolatesinverse-cap) Examining the local interpretation for a
+PI which is milk (green) chocolate incorrectly predicted to be dark
+(orange). In the attribution projection, the PI could be either milk or
+dark. Sodium and Fiber have the largest differences in attributed
+feature importance, with low values relative to other milk chocolates.
+The lack of importance attributed to these variables is suspected of
+contributing to the mistake, so the contribution of Sodium is varied. If
+Sodium had a larger contribution to the prediction (like in this view).
+the PI would look more like other milk chocolates. (A video of the tour
+animation can be found at
+\href{https://vimeo.com/666431148}{vimeo.com/666431148}.)
 
@@
 
-\caption{Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
-In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed feature importance, with low values relative to other milk chocolates.
-The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
-If Sodium had a larger contribution to the prediction (like in this view).
-the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at \href{https://vimeo.com/666431148}{vimeo.com/666431148}.)}\label{fig:casechocolatesinverse}
+\caption{(ref:casechocolatesinverse-cap)}\label{fig:casechocolatesinverse}
 \end{figure}
 
-The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
-Looking at the attribution projection, this is found to be the case.
-Previously, Fiber was essential while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
-
-Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
-The contribution of Sodium is selected to be varied.
-Even in the initial projection, the instance looks slightly more like its observed milk than predicted dark chocolate.
-The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
+The difference of position in the tree SHAP PCA with the previous case
+is quite significant; this gives a higher-level sense that the
+attributions should be quite different. Looking at the attribution
+projection, this is found to be the case. Previously, Fiber was
+essential while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat have high attributions here,
+while they were unimportant in the preceding case.
+
+Comparing the attribution with the CI (dotted line), large discrepancies
+in Sodium and Fiber are identified. The contribution of Sodium is
+selected to be varied. Even in the initial projection, the instance
+looks slightly more like its observed milk than predicted dark
+chocolate. The misclassification appears least supported when the basis
+reaches sodium attribution of typical dark chocolate.
 
@@
 
-The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-Nine higher-level skill groupings were identified and aggregated from highly correlated features.
-A random forest model is fit from these predictors, regressing player wages {[}2020 euros{]}.
-The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
-Continuing from the exploration in section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
-A leading offensive fielder (L.
-Messi) is compared with a top defensive fielder (V.
-van Dijk).
-The same instances were used in figure \ref{fig:shapdistrbd}.
+The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill
+measurements of soccer/football players and wage information. Nine
+higher-level skill groupings were identified and aggregated from highly
+correlated features. A random forest model is fit from these predictors,
+regressing player wages {[}2020 euros{]}. The model was fit from 5000
+instances before being thinned to 500 players to mitigate occlusion and
+render time. Continuing from the exploration in section
+@ref(sec:explanations), we are interested to see the difference in
+attribution based on the exogenous player position. That is, the model
+should be able to use multiple linear profiles to better predict the
+wages from different field positions of players despite not having this
+information. A leading offensive fielder (L. Messi) is compared with a
+top defensive fielder (V. van Dijk). The same instances were used in
+figure @ref(fig:shapdistrbd).
+
+(ref:casefifa-cap) Exploring the wages {[}euros{]} relative to skill
+measurements in the FIFA 2020 data. Star offensive player (L. Messi) is
+the PI and he is compared with a top defensive player (V. van Dijk). The
+attribution projection is shown at left, and it can be seen that this
+combination of features produces a view where Messi has very high
+predicted (and observed) wages. Defense (\texttt{def}) is the chosen
+feature to vary. It starts off very low, and as its contribution is
+increased (right plot) it can be seen that Messi's predicted wages
+decrease dramatically. The increased contribution in defense comes at
+the expense of offensive and reaction skills. The interpretation is that
+indeed, Messi's high wages are most attributable to his offensive and
+reaction skills, as initially provided by the local explanation. (A
+video of the animated radial tour can be found at
+\href{https://vimeo.com/666431163}{vimeo.com/666431163}.)
 
@@
 
-\caption{Exploring the wages {[}euros{]} relative to skill measurements in the FIFA 2020 data.
-Star offensive player (L.
-Messi) is the PI and he is compared with a top defensive player (V.
-van Dijk).
-The attribution projection is shown at left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
-Defense (\texttt{def}) is the chosen feature to vary.
-It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
-The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.)}\label{fig:casefifa}
+\caption{(ref:casefifa-cap)}\label{fig:casefifa}
 \end{figure}
 
-Figure \ref{fig:casefifa}, tests the support of the local explanation.
-Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
-If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
-However, Increasing the contribution of a variable with low importance would rotate both features out of the frame.
-
-The contribution from \texttt{def} will be varied to contrast with offensive skills.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group.
-Players with high values in defensive skills are now the rightmost points.
-In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
+Figure @ref(fig:casefifa), tests the support of the local explanation.
+Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both
+crucial to explaining a star offensive player. If either of them were
+rotated out, the other would be rotated into the frame, maintaining a
+far-right position. However, Increasing the contribution of a variable
+with low importance would rotate both features out of the frame.
+
+The contribution from \texttt{def} will be varied to contrast with
+offensive skills. As the contribution of defensive skills increases,
+Messi's is no longer separated from the group. Players with high values
+in defensive skills are now the rightmost points. In terms of what-if
+analysis, the difference between the data mean and his predicted wages
+would be halved if Messi's tree SHAP attributions at these levels.
 
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
-
-Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales.
-A random forest model has fitted this price {[}USD{]} with the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
-Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
+\subsection{Ames housing 2018, sales price
+regression}\label{ames-housing-2018-sales-price-regression}}
+
+Ames 2018, housing data was subset to North Ames (the neighborhood with
+the most house sales). The remaining are 338 house sales. A random
+forest model has fitted this price {[}USD{]} with the property features:
+Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house
+was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of
+Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total
+number of Rooms (\texttt{Rms}), Year the Garage was Built
+(\texttt{GYB}), and Garage Area (\texttt{GrA}). Using interaction from
+the global view, a house with an extreme negative residual and an
+accurate instance with a similar prediction are selected.
+
+(ref:caseames-cap) Exploring an instance with a large residual as the PI
+from fitting sales price {[}USD{]} to other variables in the Ames
+housing 2018. The sale price of the PI was under-predicted. The local
+explanation indicates a sizable attribution to Lot Area (\texttt{LtA}).
+The CI has a similar predicted sales price and smaller residual and has
+minimal attribution to Lot Area. In the attribution projection, the PI
+has a higher sales price than the CI. Reducing the contribution of Lot
+Area brings these two prices in line. This suggests if the model did not
+consider Lot Area, then the two houses would be quite similar. That is,
+the large residual is due to a lack of factoring in the Lot Area for the
+prediction of PI's sales price. (A video showing the animation is at
+\href{https://vimeo.com/666431134}{vimeo.com/666431134}.)
 
@@
 
-\caption{Exploring an instance with a large residual as the PI from fitting sales price {[}USD{]} to other variables in the Ames housing 2018.
-The sale price of the PI was under-predicted.
-The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}).
-The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
-In the attribution projection, the PI has a higher sales price than the CI.
-Reducing the contribution of Lot Area brings these two prices in line.
-This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
-That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.)}\label{fig:caseames}
+\caption{(ref:caseames-cap)}\label{fig:caseames}
 \end{figure}
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
-The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-The attribution projection places instances with high Living Areas to the right.
-The contribution of Living Area contrasts the contribution of this feature.
-As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
-This large of importance in the Living Area is relatively uncommon.
-Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
+Figure @ref(fig:caseames) selects the house sale 74, a sizable under
+prediction with an enormous Lot Area contribution. The CI has a similar
+predicted price though the prediction was accurate and gives almost no
+attribution to lot size. The attribution projection places instances
+with high Living Areas to the right. The contribution of Living Area
+contrasts the contribution of this feature. As the contribution of Lot
+Area decreases, the predictive power decreases for the PI, while the CI
+remains stationary. This large of importance in the Living Area is
+relatively uncommon. Boosting tree models may be more resilient to such
+an under-prediction as they would up-weighting this residual and force
+its inclusion in the final model.
 
@@
 
-There is a clear need to extend the interpretability of black-box models.
-This chapter provides a technique that builds on local interpretations to explore the feature importance local to an instance.
-The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
-Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
-A global view shows the data space, explanation space, residual plot.
-The user can interactively select instances to compare, contrast, and study further.
-Then the radial tour is used to explore the feature sensitivity identified by the attribution projection.
-
-This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
-In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
-This was verified by making a follow-up plot of the data.
-The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
-In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used by the random forest model.
-
-This analysis is manually intensive and thus only feasible for investigating a few instances.
-The approach selects instances that the model has not done well to predict and compare it with an instance where it did fit well.
-The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and responses to cross-check interpretations made from the cheem viewer.
-This methodology provides an additional tool in the box for studying model fitting.
-
-An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} Greenwell et al.
-2020; Shi et al.
-2022; Liaw and Wiener 2002; Wright and Ziegler 2017; Chen et al.
-2021, respectively).
-Tree SHAP was selected because of its computational efficiency.
-The SHAP and oscillation explanations could be added with the use of the \texttt{DALEX::explain()} and would be an excellent direction to extend the work (Biecek 2018; Biecek and Burzykowski 2021).
+There is a clear need to extend the interpretability of black-box
+models. This chapter provides a technique that builds on local
+interpretations to explore the feature importance local to an instance.
+The local interpretations form an attribution projection from which
+feature contributions are varied using a radial tour. Several diagnostic
+plots are provided to assist with understanding the sensitivity of the
+prediction to particular features. A global view shows the data space,
+explanation space, residual plot. The user can interactively select
+instances to compare, contrast, and study further. Then the radial tour
+is used to explore the feature sensitivity identified by the attribution
+projection.
+
+This approach has been illustrated using four data examples of random
+forest models with the tree SHAP local explanation. In the penguins
+example, we showed how the misclassification of a penguin arose due to
+it having an unusually small flipper size compared to others of its
+species. This was verified by making a follow-up plot of the data. The
+chocolates example shows how a dark chocolate was misclassified
+primarily due to its attribution to Fiber, and a milk chocolate was
+misclassified as dark due to its lowish Sodium value. In the FIFA
+example, we show how low Messi's salary would be if it depended on
+defensive skills. In the Ames housing data, an inaccurate prediction for
+a house was likely due to the Lot Area not being effectively used by the
+random forest model.
+
+This analysis is manually intensive and thus only feasible for
+investigating a few instances. The approach selects instances that the
+model has not done well to predict and compare it with an instance where
+it did fit well. The radial tour launches from the attribution
+projection to enable exploration of the sensitivity of the prediction to
+any feature. It can be helpful to make additional plots of the features
+and responses to cross-check interpretations made from the cheem viewer.
+This methodology provides an additional tool in the box for studying
+model fitting.
+
+An implementation is provided in the open-source \textbf{R} package
+\textbf{cheem}, available on
+\href{https://CRAN.R-project.org/package=cheem}{CRAN}. Example data sets
+are provided, and you can upload your data after model fitting and
+computing the local explanations. In theory, this approach would work
+with any black-box model, but the implementation currently only
+calculates tree SHAP for tree-based models supported by
+\textbf{treeshap} (tree based models from \textbf{gbm},
+\textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or
+\textbf{xgboost} Greenwell et al. 2020; Shi et al. 2022; Liaw and Wiener
+2002; Wright and Ziegler 2017; Chen et al. 2021, respectively). Tree
+SHAP was selected because of its computational efficiency. The SHAP and
+oscillation explanations could be added with the use of the
+\texttt{DALEX::explain()} and would be an excellent direction to extend
+the work (Biecek 2018; Biecek and Burzykowski 2021).
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+We would like to thank Professor Przemyslaw Biecek for his input early
+in the project and to the broader \(\text{MI}^\text2\) lab group for the
+\textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
+This research was supported by Australian Government Research Training
+Program (RTP) scholarships. Thanks to Jieyang Chong for helping
+proofread this article.
+
+The namesake, Cheem, refers to a fictional race of humanoid trees from
+Doctor Who lore. \textbf{DALEX} pulls on from that universe, and we
+initially apply tree SHAP explanations specific to tree-based models.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
+Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the
+Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''}
+\emph{IEEE Access} 6: 52138--60.
+
+\leavevmode\hypertarget{ref-anderson_introduction_1995}{}%
+Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT
+press.
+
 \leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
-Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
-2020.
-{``Explainable {Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies, Opportunities and Challenges Toward Responsible {AI}.''} \emph{Information Fusion} 58: 82--115.
+Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser,
+Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio
+Gil-López, Daniel Molina, and Richard Benjamins. 2020. {``Explainable
+{Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies,
+Opportunities and Challenges Toward Responsible {AI}.''}
+\emph{Information Fusion} 58: 82--115.
 
 \leavevmode\hypertarget{ref-asimov_grand_1985}{}%
-Asimov, Daniel.
-1985.
-{``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
+Asimov, Daniel. 1985. {``The {Grand} {Tour}: A {Tool} for {Viewing}
+{Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and
+Statistical Computing} 6 (1): 128--43.
 https://doi.org/\url{https://doi.org/10.1137/0906011}.
@@
 \leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
-Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
+Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive
+Models in {R}.''} \emph{The Journal of Machine Learning Research} 19
+(1): 3245--49.
 
 \leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
----------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
+---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus}
+{Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
 \leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
-Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
+{Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive}
+{Models}}. CRC Press.
+
+\leavevmode\hypertarget{ref-boser_training_1992}{}%
+Boser, Bernhard E., Isabelle M. Guyon, and Vladimir N. Vapnik. 1992.
+{``A Training Algorithm for Optimal Margin Classifiers.''} In
+\emph{Proceedings of the Fifth Annual Workshop on {Computational}
+Learning Theory}, 144--52.
+
+\leavevmode\hypertarget{ref-breiman_random_2001}{}%
+Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45
+(1): 5--32.
 
 \leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
-Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
+---------. 2001b. {``Statistical Modeling: {The} Two Cultures (with
+Comments and a Rejoinder by the Author).''} \emph{Statistical Science}
+16 (3): 199--231.
 
 \leavevmode\hypertarget{ref-buja_grand_1986}{}%
-Buja, Andreas, and Daniel Asimov.
-1986.
-{``Grand {Tour} {Methods}: {An} {Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer} {Science} and {Statistics}}, 63--67.
-New York, NY, USA: Elsevier North-Holland, Inc.
-\url{http://dl.acm.org/citation.cfm?id=26036.26046}.
+Buja, Andreas, and Daniel Asimov. 1986. {``Grand {Tour} {Methods}: {An}
+{Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on
+the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer}
+{Science} and {Statistics}}, 63--67. New York, NY, USA: Elsevier
+North-Holland, Inc. \url{http://dl.acm.org/citation.cfm?id=26036.26046}.
 
 \leavevmode\hypertarget{ref-chang_shiny_2021}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
-2021.
-\emph{Shiny: Web Application Framework for r}.
+Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke,
+Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara
+Borges. 2021. \emph{Shiny: Web Application Framework for r}.
 \url{https://CRAN.R-project.org/package=shiny}.
@@
 \leavevmode\hypertarget{ref-chen_xgboost_2021}{}%
-Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al.
-2021.
-{``Xgboost: {Extreme} {Gradient} {Boosting}.''} \url{https://CRAN.R-project.org/package=xgboost}.
+Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang,
+Hyunsu Cho, Kailong Chen, et al. 2021. {``Xgboost: {Extreme} {Gradient}
+{Boosting}.''} \url{https://CRAN.R-project.org/package=xgboost}.
 
 \leavevmode\hypertarget{ref-cook_manual_1997}{}%
-Cook, Dianne, and Andreas Buja.
-1997.
-{``Manual {Controls} for {High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
+Cook, Dianne, and Andreas Buja. 1997. {``Manual {Controls} for
+{High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of
+Computational and Graphical Statistics} 6 (4): 464--80.
 \url{https://doi.org/10.2307/1390747}.
@@
 \leavevmode\hypertarget{ref-cook_grand_2008}{}%
-Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
-2008.
-{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual} {Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
+Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham. 2008.
+{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual}
+{Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
 Berlin, Heidelberg: Springer Berlin Heidelberg.
@@
 \leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
-da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee.
-2021.
-{``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
+da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee. 2021. {``A
+{Projection} {Pursuit} {Forest} {Algorithm} for {Supervised}
+{Classification}.''} \emph{Journal of Computational and Graphical
+Statistics}, 1--21.
 
 \leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
-Dastin, Jeffrey.
-2018.
-{``Amazon Scraps Secret {AI} Recruiting Tool That Showed Bias Against Women.''} \emph{Reuters}, October.
+Dastin, Jeffrey. 2018. {``Amazon Scraps Secret {AI} Recruiting Tool That
+Showed Bias Against Women.''} \emph{Reuters}, October.
 \url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.
@@
 \leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
-Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
-2018.
-{``Addressing Age-Related Bias in Sentiment Analysis.''} In \emph{Proceedings of the 2018 Chi Conference on Human Factors in Computing Systems}, 1--14.
+Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren
+Gergle. 2018. {``Addressing Age-Related Bias in Sentiment Analysis.''}
+In \emph{Proceedings of the 2018 Chi Conference on Human Factors in
+Computing Systems}, 1--14.
 
 \leavevmode\hypertarget{ref-duffy_apple_2019}{}%
-Duffy, Claire.
-2019.
-{``Apple Co-Founder {Steve} {Wozniak} Says {Apple} {Card} Discriminated Against His Wife.''} \emph{CNN}, November.
+Duffy, Claire. 2019. {``Apple Co-Founder {Steve} {Wozniak} Says {Apple}
+{Card} Discriminated Against His Wife.''} \emph{CNN}, November.
 \url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
 
+\leavevmode\hypertarget{ref-espadoto_toward_2021}{}%
+Espadoto, Mateus, Rafael M. Martins, Andreas Kerren, Nina S. T. Hirata,
+and Alexandru C. Telea. 2021. {``Toward a {Quantitative} {Survey} of
+{Dimension} {Reduction} {Techniques}.''} \emph{IEEE Transactions on
+Visualization and Computer Graphics} 27 (3): 2153--73.
+\url{https://doi.org/10.1109/TVCG.2019.2944182}.
+
 \leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
-Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
+Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices
+with Application to Principal Component Analysis.''} \emph{Biometrika}
+58 (3): 453--67.
 
 \leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
-Gorman, Kristen B., Tony D.
-Williams, and William R.
-Fraser.
-2014.
-{``Ecological Sexual Dimorphism and Environmental Variability Within a Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS One} 9 (3): e90081.
+Gorman, Kristen B., Tony D. Williams, and William R. Fraser. 2014.
+{``Ecological Sexual Dimorphism and Environmental Variability Within a
+Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS
+One} 9 (3): e90081.
 
 \leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
-Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
+Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}:
+{Uncertainty} of Model Explanations for Non-Additive Predictive
+Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
 \leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
-Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley}
+{Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
 \leavevmode\hypertarget{ref-greenwell_gbm_2020}{}%
-Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M. Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression} {Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
+Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M.
+Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression}
+{Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
 
 \leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
-Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
-Gorman.
-2020.
-{``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
+Horst, Allison Marie, Alison Presmanes Hill, and Kristen B. Gorman.
+2020. {``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin
+Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
 
 \leavevmode\hypertarget{ref-karwowski_international_2006}{}%
-Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
+Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of
+{Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
 
 \leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
-Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
+Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using
+{AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring
+Tool.''} \emph{Researchgate Preprint}.
 
 \leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
-Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
+Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw
+Biecek. 2021. {``Treeshap.''} Model Oriented.
+\url{https://github.com/ModelOriented/treeshap}.
 
 \leavevmode\hypertarget{ref-larson_how_2016}{}%
-Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
-2016.
-{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''} \emph{ProPublica}, May.
+Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin. 2016.
+{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''}
+\emph{ProPublica}, May.
 \url{https://www.propublica.org/article/how-we-analyzed-the-compas-recidivism-algorithm?token=RPR1E2qtzJltfJ0tS-gB_41kmfoWZAu4}.
@@
 \leavevmode\hypertarget{ref-lee_state_2021}{}%
-Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
-Sherry Zhang.
-2021.
-{``The State-of-the-Art on Tours for Dynamic Visualization of High-Dimensional Data.''} \emph{WIREs Computational Statistics} n/a (n/a): e1573.
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas
+Spyrison, Earo Wang, and H. Sherry Zhang. 2021. {``The State-of-the-Art
+on Tours for Dynamic Visualization of High-Dimensional Data.''}
+\emph{WIREs Computational Statistics} n/a (n/a): e1573.
 \url{https://doi.org/10.1002/wics.1573}.
@@
 \leavevmode\hypertarget{ref-lee_pptree_2013}{}%
-Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
+Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013.
+{``{PPtree}: {Projection} Pursuit Classification Tree.''}
+\emph{Electronic Journal of Statistics} 7: 1369--86.
 
 \leavevmode\hypertarget{ref-leone_fifa_2020}{}%
-Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
+Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''}
+\url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
 
 \leavevmode\hypertarget{ref-liaw_classification_2002}{}%
-Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
+Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression
+by {randomForest}.''} \emph{R News} 2 (3): 18--22.
 
 \leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
-Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
+Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent
+Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv
+Preprint arXiv:1802.03888}.
 
 \leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
-Lundberg, Scott M., and Su-In Lee.
-2017.
-{``A Unified Approach to Interpreting Model Predictions.''} In \emph{Proceedings of the 31st International Conference on Neural Information Processing Systems}, 4768--77.
+Lundberg, Scott M., and Su-In Lee. 2017. {``A Unified Approach to
+Interpreting Model Predictions.''} In \emph{Proceedings of the 31st
+International Conference on Neural Information Processing Systems},
+4768--77.
 
 \leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
-Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
+Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu.
+com.
+\href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
 
 \leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
-O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
+O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data
+Increases Inequality and Threatens Democracy}. Crown.
 
 \leavevmode\hypertarget{ref-ocagne_coordonnees_1885}{}%
-Ocagne, Maurice d'.
-1885.
-\emph{Coordonnées Parallèles Et Axiales.
-{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul Graphique déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
+Ocagne, Maurice d'. 1885. \emph{Coordonnées Parallèles Et Axiales.
+{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul
+Graphique déduits de La Considération Des Coordonnées Parallèles, Par
+{Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
 
 \leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
-Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
-2016.
-{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} In \emph{Proceedings of the 22nd {ACM} {SIGKDD} {International} {Conference} on {Knowledge} {Discovery} and {Data} {Mining}}, 1135--44.
-{KDD} '16.
-New York, NY, USA: Association for Computing Machinery.
-\url{https://doi.org/10.1145/2939672.2939778}.
+Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin. 2016. {``"{Why}
+{Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any}
+{Classifier}.''} In \emph{Proceedings of the 22nd {ACM} {SIGKDD}
+{International} {Conference} on {Knowledge} {Discovery} and {Data}
+{Mining}}, 1135--44. {KDD} '16. New York, NY, USA: Association for
+Computing Machinery. \url{https://doi.org/10.1145/2939672.2939778}.
 
 \leavevmode\hypertarget{ref-roberts_state_2007}{}%
-Roberts, Jonathan C.
-2007.
-{``State of the Art: {Coordinated} \& Multiple Views in Exploratory Visualization.''} In \emph{Fifth International Conference on Coordinated and Multiple Views in Exploratory Visualization ({CMV} 2007)}, 61--71.
-IEEE.
+Roberts, Jonathan C. 2007. {``State of the Art: {Coordinated} \&
+Multiple Views in Exploratory Visualization.''} In \emph{Fifth
+International Conference on Coordinated and Multiple Views in
+Exploratory Visualization ({CMV} 2007)}, 61--71. IEEE.
 
 \leavevmode\hypertarget{ref-shapley_value_1953}{}%
-Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
+Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton
+University Press.
 
 \leavevmode\hypertarget{ref-shi_lightgbm_2022}{}%
-Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas Finley, Taifeng Wang, et al.
-2022.
-{``Lightgbm: {Light} {Gradient} {Boosting} {Machine}.''} \url{https://CRAN.R-project.org/package=lightgbm}.
+Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas
+Finley, Taifeng Wang, et al. 2022. {``Lightgbm: {Light} {Gradient}
+{Boosting} {Machine}.''}
+\url{https://CRAN.R-project.org/package=lightgbm}.
 
 \leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
-Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
+Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical
+Science} 25 (3): 289--310.
 
 \leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
-Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
-2017.
-{``Learning Important Features Through Propagating Activation Differences.''} In \emph{International {Conference} on {Machine} {Learning}}, 3145--53.
-PMLR.
+Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje. 2017.
+{``Learning Important Features Through Propagating Activation
+Differences.''} In \emph{International {Conference} on {Machine}
+{Learning}}, 3145--53. PMLR.
 
 \leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
-Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
-2016.
-{``Not Just a Black Box: {Learning} Important Features Through Propagating Activation Differences.''} \emph{arXiv Preprint arXiv:1605.01713}.
+Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul
+Kundaje. 2016. {``Not Just a Black Box: {Learning} Important Features
+Through Propagating Activation Differences.''} \emph{arXiv Preprint
+arXiv:1605.01713}.
 
 \leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
-Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
+Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data}
+{Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC.
+\url{https://plotly-r.com}.
 
 \leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
-Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
-2014.
-{``Deep Inside Convolutional Networks: {Visualising} Image Classification Models and Saliency Maps.''} In \emph{In {Workshop} at {International} {Conference} on {Learning} {Representations}}.
-Citeseer.
+Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman. 2014. {``Deep
+Inside Convolutional Networks: {Visualising} Image Classification Models
+and Saliency Maps.''} In \emph{In {Workshop} at {International}
+{Conference} on {Learning} {Representations}}. Citeseer.
 
 \leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
-Spyrison, Nicholas, and Dianne Cook.
-2020.
-{``Spinifex: An {R} {Package} for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of {Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
+Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An {R} {Package}
+for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of
+{Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
 \url{https://doi.org/10.32614/RJ-2020-027}.
@@
 \leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
-Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
+Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation
+of Individual Classifications Using Game Theory.''} \emph{The Journal of
+Machine Learning Research} 11: 1--18.
 
 \leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
-Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
+Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''}
+\emph{Journal of Computational and Graphical Statistics} 27 (1):
+157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
 \leavevmode\hypertarget{ref-vanni_textual_2018}{}%
-Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
-2018.
-{``Textual {Deconvolution} {Saliency} ({TDS}): A Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the 56th {Annual} {Meeting} of the {Association} for {Computational} {Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
+Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and
+Damon Mayaffre. 2018. {``Textual {Deconvolution} {Saliency} ({TDS}): A
+Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the
+56th {Annual} {Meeting} of the {Association} for {Computational}
+{Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
 
 \leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
-Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data}
+{Analysis}}. Springer-Verlag New York.
+\url{https://ggplot2.tidyverse.org}.
 
 \leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
-Wickham, Hadley, Dianne Cook, and Heike Hofmann.
-2015.
-{``Visualizing Statistical Models: {Removing} the Blindfold.''} \emph{Statistical Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
+Wickham, Hadley, Dianne Cook, and Heike Hofmann. 2015. {``Visualizing
+Statistical Models: {Removing} the Blindfold.''} \emph{Statistical
+Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
 \url{https://doi.org/10.1002/sam.11271}.
@@
 \leavevmode\hypertarget{ref-wright_ranger_2017}{}%
-Wright, Marvin N., and Andreas Ziegler.
-2017.
-{``Ranger: {A} {Fast} {Implementation} of {Random} {Forests} for {High} {Dimensional} {Data} in {C}++ and {R}.''} \emph{Journal of Statistical Software} 77 (1): 1--17.
-\url{https://doi.org/10.18637/jss.v077.i01}.
+Wright, Marvin N., and Andreas Ziegler. 2017. {``Ranger: {A} {Fast}
+{Implementation} of {Random} {Forests} for {High} {Dimensional} {Data}
+in {C}++ and {R}.''} \emph{Journal of Statistical Software} 77 (1):
+1--17. \url{https://doi.org/10.18637/jss.v077.i01}.
 

### 2022-03-14 z_sty_reprex/sty_reprex.rmd
wresting jmlr2e.sty

---
title: "title"
author: 
  - author
  - author2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header:
        #- ../template/jmlr2e.sty ## Doesn't work
        #- tmp_jmlr.tex ## Doesn't work
        - mystyles.sty ## Works
---

A minimal reprex following https://stackoverflow.com/questions/56174593/how-to-add-sty-file-to-rmarkdown-pdf-output.

# Section 1

Hello world

### 2022-03-14 _paper.rmd
partially working .sty (z_sty_reprex/)

@@
 - Dianne Cook
+#^[ORCiD: 000-0002-3813-7155]
 - Kimbal Marriott

### 2022-03-14 z_sty_reprex/sty_reprex.rmd
partially working .sty (z_sty_reprex/)

@@
       in_header:
-        #- ../template/jmlr2e.sty ## Doesn't work
-        #- tmp_jmlr.tex ## Doesn't work
-        - mystyles.sty ## Works
+        - jmlr2e.sty ## Doesn't work
+        #- mystyles.sty ## Works (trivial \usepackage{})
 ---
@@
 Hello world
+
+# Attempts and notes
+
+- Trivial sty files work (mystyles.sty)
+- Trivial tex file didn't work.
+- jmlr2e.sty doesn't work wholesale, hence:
+- Changing to *.tex didn't work.
+- Trying to pare back and to isolate the issue:
+    - adding \makeatletter and \makeatother to change the order of operations handling @ in marco assignment. _partial success!!_
+    - uncommenting .sty file, keeping an eye out for @ in macronames. 
+    - \makeatletter and \makeatother at top and bottom is back to original error... (! LaTeX Error: Missing \begin{document}.); try changing back and forth

### 2022-03-15 z_sty_reprex/sty_reprex.rmd
sty reprex progress atm.

@@
       in_header:
-        - jmlr2e.sty ## Doesn't work
-        #- mystyles.sty ## Works (trivial \usepackage{})
+        - jmlr2e.sty          ## Applying comments and changes here
+        #- jmlr2eoriginal.sty ## Not working; full original .sty
+        #- mystyles.sty       ## Works (with trivial \usepackage{})
 ---
@@
 
-# Section 1
+# Section title
 
@@
 # Attempts and notes
+
+Primary errors:
+
+- original error: `! LaTeX Error: Missing \begin{document}.`
+- alternative error: `! You can't use \spacefactor in vertical mode.`
+
+Attempt notes:
 
@@
     - uncommenting .sty file, keeping an eye out for @ in macronames.
-    - \makeatletter and \makeatother at top and bottom is back to original error... (! LaTeX Error: Missing \begin{document}.); try changing back and forth
+    - \makeatletter and \makeatother at top and bottom is back to original error... (); try changing back and forth

### 2022-03-16 _paper.rmd
Use RJH's jmlr template. Figure sizes and html consistency.

--- 2022-03-14 _paper.rmd
@@
 title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
-shorttitle: Cheem
-keywords:
-  - R
-  - XAI
-  - explanable artificial intelligence
-  - black-box models
-  - local explanations
-  - grand tour
-  - manual tour
-  - radial tour
+shorttitle: "Cheem, Exploring Local Explanations"
 author:
-- Nicholas Spyrison 
-#^[Monash University, nicholas.spyrison@monash.edu, ORCiD: 0000-0002-8417-0212]
-- Dianne Cook
-- Kimbal Marriott
-#^[ORCiD: 0000-0002-9813-0377]
-  #- name: "Nicholas Spyrison"
-  #   affiliation: Monash University
-  #   address: |
-  #     | ORCiD: 0000-0002-8417-0212
-  #   email: \email{nicholas.spyrison@monash.edu}
-  #   orcid: 0000-0002-8417-0212
-  #- name: "Dianne Cook"
-  #   affiliation: Monash University
-  #   address: |
-  #     | ORCiD: 000-0002-3813-7155
-  #   orcid: 000-0002-3813-7155
-  #- name: "Kimbal Marriott"
-  #   affiliation: Monash University
-  #   address: |
-  #     | ORCiD: 0000-0002-9813-0377
-  #   orcid: 0000-0002-9813-0377
-abstract: |
- The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
+- familyname: Spyrison
+  othernames: Nicholas
+  address: Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212
+  email: Nicholas.Spyrison@monash.edu
+  correspondingauthor: true
+- familyname: Cook
+  othernames: Dianne
+  address: Monash University\newline ORCiD 0000-0002-3813-7155
+  email: DiCook@monash.edu
+- familyname: Marriott
+  othernames: Kimbal
+  address: Monash University\newline ORCiD 0000-0002-9813-0377
+  email: Kim.Marriott@monash.edu
+# Nick ORCiD 0000-0002-8417-0212
+# Di   ORCiD 0000-0002-3813-7155
+# Kim  ORCiD 0000-0002-9813-0377
+year: 2022
+date_submitted: "MM/DD"
+abstract: "The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
 This trade-off has led to the emergence of eXplainable AI (XAI).
@@
 Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
-The methods are implemented in the __R__ package __cheem__, available on CRAN.
-#documentclass: jmlr2e
-#classoption: article
+The methods are implemented in the __R__ package __cheem__, available on CRAN."
+keywords: "explainable artificial intelligence, nonlinear model interpretability, local explanations, grand tour, radial tour"
+acknowledgment: "We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
+This research was supported by Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models."
+bibliography: cheem_paper.bib
+classoption: 11pt,twoside
 output:
-  ## pdf_document: requires \ref{} 
-  pdf_document:
-  ## bookdown::pdf_document2: allows \@ref(), but doesn't add authors....
-  #bookdown::pdf_document2:
-    #includes:
-    #  in_header: template/jmlr2e.sty
-    keep_tex: true
-    toc: false
+  bookdown::pdf_document2:
+    template: template_rjh/jmlr_template.tex
+    fig_caption: yes
+    fig_height: 5
+    fig_width: 8
+    # includes:
+    #   in_header: template_rjh/preamble.tex
+    keep_tex: yes
     number_sections: yes
-    fig_caption: yes
-bibliography: cheem_paper.bib
-editor_options:
-  chunk_output_type: console
-urlcolor: blue
-linkcolor: red
-header-includes:
-  - \usepackage{setspace} ## for title page spacing
-  - \usepackage{hyperref} ## for all sorts of linking
+    #citation_package: natbib
+    latex_engine: pdflatex
 ---
@@
 
-```{r shapdistrbd, echo=F, fig.align='center', out.width="100%", fig.cap = "Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
+```{r shapdistrbd, echo=F, fig.align='center', out.width="85%", fig.cap = "Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
 A star offensive and defensive player are compared, L.
@@
 
-- a version of the cheem viewer shiny app can be directly accessed at [ebsmonash.shinyapps.io/cheem_initial/](https://ebsmonash.shinyapps.io/cheem_initial/). 
+- a version of the cheem viewer shiny app can be directly accessed at https://ebsmonash.shinyapps.io/cheem_initial/. 
 - the development version of the package is available at https://github.com/nspyrison/cheem, and 
@@
 The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at [vimeo.com/666431172](https://vimeo.com/666431172).)
+(The animation of the radial tour is at https://vimeo.com/666431172.)
 ```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins-cap)"}
@@
 This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at [vimeo.com/666431143](https://vimeo.com/666431143).)
+(A video of the tour animation can be found at https://vimeo.com/666431143.)
 ```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates-cap)"}
@@
 the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at [vimeo.com/666431148](https://vimeo.com/666431148).)
+(A video of the tour animation can be found at https://vimeo.com/666431148.)
 ```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse-cap)"}
@@
 The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at [vimeo.com/666431163](https://vimeo.com/666431163).)
-```{r casefifa, echo=F, out.width = "100%", fig.cap = "(ref:casefifa-cap)"}
+(A video of the animated radial tour can be found at https://vimeo.com/666431163.)
+```{r casefifa, echo=F, out.width = "90%", fig.cap = "(ref:casefifa-cap)"}
 knitr::include_graphics("./figures/case_fifa.png")
@@
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at [vimeo.com/666431134](https://vimeo.com/666431134).)
+(A video showing the animation is at https://vimeo.com/666431134.)
 ```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames-cap)"}
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI$^2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages.
 This research was supported by Australian Government Research Training Program (RTP) scholarships.

### 2022-03-16 _paper.tex
Use RJH's jmlr template. Figure sizes and html consistency.

--- 2022-03-01 _paper.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+\documentclass[11pt,twoside]{article}
+
+% Any additional packages needed should be included after jmlr2e.
+% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
+% and defines many common macros, such as 'proof' and 'example'.
 %
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections},
-  pdfkeywords={R, XAI, explanable artificial intelligence, black-box models, local explanations, grand tour, manual tour, radial tour},
-  colorlinks=true,
-  linkcolor=red,
-  filecolor=Maroon,
-  citecolor=Blue,
-  urlcolor=blue,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
+% It also sets the bibliographystyle to plainnat; for more information on
+% natbib citation styles, see the natbib documentation, a copy of which
+% is archived at http://www.jmlr.org/format/natbib.pdf
+
+% Available options for package jmlr2e are:
+%
+%   - abbrvbib : use abbrvnat for the bibliography style
+%   - nohyperref : do not load the hyperref package
+%   - preprint : remove JMLR specific information from the template,
+%         useful for example for posting to preprint servers.
+%
+% Example of using the package with custom options:
+%
+% \usepackage[abbrvbib, preprint]{jmlr2e}
+
+\usepackage[]{jmlr2e}
+
+% Pandoc syntax highlighting
 \usepackage{color}
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% From pandoc table feature
 \usepackage{longtable,booktabs,array}
@@
 \makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\usepackage{setspace}
-\usepackage{hyperref}
+
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% Definitions of handy macros can go here
+
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
-\ifluatex
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% Heading arguments are {volume}{year}{pages}{date submitted}{date published}{paper id}{author-full-names}
+
+\jmlrheading{}{2022}{}{MM/DD}{--}{}{%
+Nicholas Spyrison, Dianne Cook and Kimbal Marriott}
+
+% Short headings should be running head and authors last names
+
+\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison, Cook and Marriott}
+\firstpageno{1}
+
+\begin{document}
+% Add back JMLR proof environment which is over-written by amsthm
+\def\proof{\par\noindent{\bf Proof\ }}
+\def\endproof{\hfill\BlackBox\\[2mm]}
 
 \title{Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections}
-\author{}
-\date{\vspace{-2.5em}}
-
-\begin{document}
+
+\author{%
+\name Nicholas Spyrison \email \href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
+%
+\name Dianne Cook \email \href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
+%
+\name Kimbal Marriott \email \href{mailto:Kim.Marriott@monash.edu}{\nolinkurl{Kim.Marriott@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-9813-0377 \AND
+}
+
+\editor{}
+
 \maketitle
+
 \begin{abstract}
+% Put your abstract here. Abstracts are limited to 200 words for
+% regular articles and 100 words for Letters to the Editor. Please no
+% personal pronouns, also please do not use the words ``new'' and/or
+% ``novel'' in the abstract. An article usually includes an abstract, a
+% concise summary of the work covered at length in the main body of the
+% article.
 The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
@@
 
+\begin{keywords}
+explainable artificial intelligence, nonlinear model interpretability, local explanations, grand tour, radial tour
+\end{keywords}
+
+%  Body of the article
 \hypertarget{sec:intro}{%
@@
 
-There are different reasons and emphases to fit a model.
-Breiman (2001), reiterated by Shmueli (2010), taxonomize modeling based on its purpose; \emph{explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more on the predictions of out-of-sample instances.
+There are different reasons and emphases when considering to fit a model.
+Breiman (2001b), reiterated by Shmueli (2010), taxonomize models based on their purpose.
+\emph{Explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more narrowly on the performance of some objective function.
 The intended use has important implications for model selection and development.
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-While predictive modeling may opt for more accurate nonlinear models.
-The use of black-box models is becoming increasingly common, but not without their share of controversy (O'Neil 2016; Kodiyan 2019).
-However, the loss of interpretation presents a challenge.
-
-Interpretability is vital for exploring and protecting against potential biases (e.g., sex (Dastin 2018; Duffy 2019), race (Larson et al.
-2016), and age (Díaz et al.
-2018)) in any model.
-For instance, models regularly pick up on biases in the training data that have observed influence on the response (output) feature, which is then built into the model.
-Feature-level (variable-level) interpretability of models is essential in evaluating such biases.
-It is also generally important for many problems, where it is not enough to predict accurately.
-Still, one must be able to explain which predictors are most responsible for generating a response value.
-
-Another concern is data drift, a shift in support or domain of the explanatory features (variable or predictors).
-Nonlinear models are typically more sensitive and do not extrapolate well outside the training data domain.
-Better interpretability of the model means more transparency where models' predictions may be plausible or completely unreliable.
-
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models.
+Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network to name a few (Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995, receptively).
+
+Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
+This difficulty in interpreting the terms in complex nonlinear models sometimes leads to them being referred to as black box models.
+Despite the potentially better performance of nonlinear models, their use is not without controversy (O'Neil 2016; Kodiyan 2019).
+And the loss of interpretation presents a challenge.
+
+Interpretability is vital for exploring and protecting against potential biases in any model (e.g.~sex -- Dastin (2018); Duffy (2019), race -- Larson et al.
+(2016), and age -- Díaz et al.
+(2018)).
+For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
+This bias is then built into the model.
+Variable-level (feature-level) interpretability of models is essential in evaluating and addressing such biases.
+
+Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors).
+Some nonlinear models are sensitive to this and do not extrapolate well outside the support of the training data (Espadoto et al.
+2021).
+Maintaining variable interpretability is also essential to address issues arising from data drift.
+
+Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models (Adadi and Berrada 2018; Arrieta et al.
+2020).
 A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/shap_distr_bd}
+{\centering \includegraphics[width=0.85\linewidth]{./figures/shap_distr_bd}
 
@@
 \item
-  a version of the cheem viewer shiny app can be directly accessed at \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
+  a version of the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
 \item
@@
 The radial view shows varying the contribution of \texttt{fl} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at \href{https://vimeo.com/666431172}{vimeo.com/666431172}.)}\label{fig:casepenguins}
+(The animation of the radial tour is at \url{https://vimeo.com/666431172}.)}\label{fig:casepenguins}
 \end{figure}
@@
 This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at \href{https://vimeo.com/666431143}{vimeo.com/666431143}.)}\label{fig:casechocolates}
+(A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)}\label{fig:casechocolates}
 \end{figure}
@@
 the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at \href{https://vimeo.com/666431148}{vimeo.com/666431148}.)}\label{fig:casechocolatesinverse}
+(A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)}\label{fig:casechocolatesinverse}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/case_fifa}
+{\centering \includegraphics[width=0.9\linewidth]{./figures/case_fifa}
 
@@
 The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at \href{https://vimeo.com/666431163}{vimeo.com/666431163}.)}\label{fig:casefifa}
+(A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)}\label{fig:casefifa}
 \end{figure}
@@
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at \href{https://vimeo.com/666431134}{vimeo.com/666431134}.)}\label{fig:caseames}
+(A video showing the animation is at \url{https://vimeo.com/666431134}.)}\label{fig:caseames}
 \end{figure}
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\(^2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
 This research was supported by Australian Government Research Training Program (RTP) scholarships.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
+Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
+
+\leavevmode\hypertarget{ref-anderson_introduction_1995}{}%
+Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT press.
+
 \leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
@@
 
+\leavevmode\hypertarget{ref-boser_training_1992}{}%
+Boser, Bernhard E., Isabelle M.
+Guyon, and Vladimir N.
+Vapnik.
+1992.
+{``A Training Algorithm for Optimal Margin Classifiers.''} In \emph{Proceedings of the Fifth Annual Workshop on {Computational} Learning Theory}, 144--52.
+
+\leavevmode\hypertarget{ref-breiman_random_2001}{}%
+Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45 (1): 5--32.
+
 \leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
-Breiman, Leo. 2001. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
+---------. 2001b. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
 
@@
 \url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
+
+\leavevmode\hypertarget{ref-espadoto_toward_2021}{}%
+Espadoto, Mateus, Rafael M.
+Martins, Andreas Kerren, Nina S.
+T.
+Hirata, and Alexandru C.
+Telea.
+2021.
+{``Toward a {Quantitative} {Survey} of {Dimension} {Reduction} {Techniques}.''} \emph{IEEE Transactions on Visualization and Computer Graphics} 27 (3): 2153--73.
+\url{https://doi.org/10.1109/TVCG.2019.2944182}.
 
@@
 
+% Acknowledgements should go at the end, before appendices and references
+
+\acks{We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
+This research was supported by Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.}
+
+% Appendices to go in include-after
+
+\vskip 0.2in
+\bibliography{cheem\_paper.bib}
+
 \end{document}

### 2022-03-16 jmlr.tex
Use RJH's jmlr template. Figure sizes and html consistency.

--- 2022-03-16 _paper.tex
@@
 
-% Pandoc syntax highlighting
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
 % tightlist command for lists without linebreak
@@
 
-% Pandoc citation processing
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
 % Definitions of handy macros can go here
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
+%% Any special functions or other packages can be loaded here.
 
@@
 
-\jmlrheading{}{2022}{}{MM/DD}{--}{}{%
-Nicholas Spyrison, Dianne Cook and Kimbal Marriott}
+\jmlrheading{}{2022}{}{03/22}{--}{}{%
+Fan Cheng, Anastasios Panagiotelis and Rob J Hyndman}
 
@@
 
-\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison, Cook and Marriott}
+\ShortHeadings{Shorter title}{Cheng, Panagiotelis and Hyndman}
 \firstpageno{1}
@@
 
-\title{Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections}
+\title{Title}
 
 \author{%
-\name Nicholas Spyrison \email \href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
+\name Fan Cheng \email \href{mailto:Fan.Cheng@monash.edu}{\nolinkurl{Fan.Cheng@monash.edu}} \\\addr Monash University\newline Australia \AND
 %
-\name Dianne Cook \email \href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
+\name Anastasios Panagiotelis \email \href{mailto:Anastasios.Panagiotelis@sydney.edu.au}{\nolinkurl{Anastasios.Panagiotelis@sydney.edu.au}} \\\addr University of Sydney\newline Australia \AND
 %
-\name Kimbal Marriott \email \href{mailto:Kim.Marriott@monash.edu}{\nolinkurl{Kim.Marriott@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-9813-0377 \AND
+\name Rob J Hyndman \email \href{mailto:Rob.Hyndman@monash.edu}{\nolinkurl{Rob.Hyndman@monash.edu}} \\\addr Monash University\newline Australia \AND
 }
@@
 % article.
-The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
-This trade-off has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction with \emph{local explanations}, a point estimate of the linear feature importance in the vicinity of one instance.
-These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
-The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
+Blah blah blah
 \end{abstract}
@@
 \begin{keywords}
-explainable artificial intelligence, nonlinear model interpretability, local explanations, grand tour, radial tour
+dimension reduction, divergence estimation, Hellinger distance, anomaly detection, k-d trees
 \end{keywords}
@@
 %  Body of the article
-\hypertarget{sec:intro}{%
-\section{Introduction}\label{sec:intro}}
+\newpage
 
-There are different reasons and emphases when considering to fit a model.
-Breiman (2001b), reiterated by Shmueli (2010), taxonomize models based on their purpose.
-\emph{Explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more narrowly on the performance of some objective function.
-The intended use has important implications for model selection and development.
-In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network to name a few (Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995, receptively).
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
-This difficulty in interpreting the terms in complex nonlinear models sometimes leads to them being referred to as black box models.
-Despite the potentially better performance of nonlinear models, their use is not without controversy (O'Neil 2016; Kodiyan 2019).
-And the loss of interpretation presents a challenge.
+Blah blah. You can refer to papers \citep{Witten2011} or cite them directly as in \citet{AriasCastro2020}.
 
-Interpretability is vital for exploring and protecting against potential biases in any model (e.g.~sex -- Dastin (2018); Duffy (2019), race -- Larson et al.
-(2016), and age -- Díaz et al.
-(2018)).
-For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
-This bias is then built into the model.
-Variable-level (feature-level) interpretability of models is essential in evaluating and addressing such biases.
+Add the following to the yaml header if you want to remove the JMLR details.
 
-Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors).
-Some nonlinear models are sensitive to this and do not extrapolate well outside the support of the training data (Espadoto et al.
-2021).
-Maintaining variable interpretability is also essential to address issues arising from data drift.
-
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models (Adadi and Berrada 2018; Arrieta et al.
-2020).
-A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
-
-In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
-Lee et al.
-2021) is a sequence of linear projections of data onto a lower-dimensional space.
-Tours are viewed as an animation over minor changes to the projection basis.
-Structure in a projection can then be explored visually to see which features contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information about the structure and features of the object.
-
-There are various types of tours distinguished by the generation of projection bases.
-In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected feature.
-Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths (Y.
-D.
-Lee et al.
-2013; da Silva, Cook, and Lee 2021).
-
-We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in R package \textbf{cheem}.
-Example datasets are provided to illustrate usage for classification and regression tasks.
-
-Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
-It ignores correlations of the features and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object.
-A constraint in our approach is that the basis must remain orthonormal.
-When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
-
-The remainder of this paper is organized as follows.
-The following Section, \ref{sec:explanations}, covers the background of the local explanation and the traditional visuals produced.
-Section \ref{sec:tour} explains the animations of continuous linear projections.
-Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
-Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative output.
-Section \ref{sec:cheemdiscussion} concludes with the insights gained and possible directions to explore in the future.
-
-\hypertarget{sec:explanations}{%
-\section{Local explanations}\label{sec:explanations}}
-
-Consider a highly nonlinear model.
-It can be hard to determine whether small changes in a feature's value will make a class prediction change group or identify which features contribute to an extreme residual.
-Local explanations shed light on these situations by approximating linear feature importance in the vicinity of a single instance.
-
-A comprehensive summary of the taxonomy and literature of explanation techniques is provided in Figure 6 of Arrieta et al.
-(2020).
-It includes a large number of model-specific explanations such as deepLIFT (Shrikumar et al.
-2016; Shrikumar, Greenside, and Kundaje 2017), a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME, (Ribeiro, Singh, and Guestrin 2016) SHAP, (Lundberg and Lee 2017), and their variants are popular.
-
-These instance-level explanations are used in various ways depending on the data.
-In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels (Simonyan, Vedaldi, and Zisserman 2014).
-For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
-2018).
-In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction (Ribeiro, Singh, and Guestrin 2016).
-
-SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The explanations of SHAP refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
-Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
-The attribution of feature importance depends on the sequence of the included features.
-The SHAP values are the mean contributions over different feature sequences.
-The approach is related to partial dependence plots (Molnar 2020), used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
-Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), helps to understand the relative importance of each feature in the context of all other candidate features.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.85\linewidth]{./figures/shap_distr_bd}
-
-}
-
-\caption{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
-A star offensive and defensive player are compared, L.
-Messi and V.
-van Dijk, respectively.
-Panel (a) shows breakdown plots of three sequences of the features.
-The sequence of the variables impacts the magnitude of their attribution.
-Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player.
-Reaction skills are important for both players.
-Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
-\end{figure}
-
-Following the use case \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021), FIFA data is used to illustrate SHAP.
-Consider soccer data from the FIFA 2020 season (Leone 2020).
-There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing player's wages {[}2020 Euros{]} from their skill measurements.
-The SHAP values are compared for a star offensive player (L.
-Messi) and defensive player (V.
-van Dijk).
-The results are displayed in Figure \ref{fig:shapdistrbd}.
-A difference in the attribution of the feature importance across the two positions of the players can be expected.
-This would be interpreted as how a player's salary depends on this combination of skill sets.
-Panel (a) is a modified breakdown plot (Gosiewska and Biecek 2019) where three sequences of features are presented, so the two instances can be more easily compared.
-The magnitude of the contributions depends on the sequence in which they appear.
-Panel (b) shows the differences of the player's median values of over 25 such sequences.
-In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
-In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
-
-For the application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
-Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; random forests are used for illustration.
-The following section will use normalized SHAP values as a projection basis (call this the \emph{attribution projection}) that will be used to explore the sensitivity of the feature contributions.
-
-\hypertarget{sec:tour}{%
-\section{Tours and the radial tour}\label{sec:tour}}
-
-A \emph{tour} enables viewing of high-dimensional data by animating many linear projections with small incremental changes.
-It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
-There are various types of tours that are distinguished by how the paths are generated (S.
-Lee et al.
-2021; Cook et al.
-2008).
-
-The manual tour (Cook and Buja 1997) defines its path by changing a selected feature's contribution to a basis to allow the feature to contribute more or less to the projection.
-The requirement constrains the contribution of all other features that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
-It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis (Karwowski 2006).
-
-A version of the manual tour called a \emph{radial tour} is implemented in Spyrison and Cook (2020) and forms the basis of the new work.
-In a radial tour, the selected feature can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and interactive re-calculation to focus on a different feature.
-
-\hypertarget{sec:cheemviewer}{%
-\section{The cheem viewer}\label{sec:cheemviewer}}
-
-To explore the local explanations, coordinated views (Roberts 2007) (also known as ensemble graphics, Unwin and Valero-Mora 2018) are provided in the \emph{cheem viewer} application.
-There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
-There are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
-There are different plots used for the categorical and quantitative responses.
-Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
-
-\hypertarget{global-view}{%
-\subsection{Global view}\label{global-view}}
-
-The global view provides a context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, each a vector of \(p\) values.
-
-Visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
-These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify instances poorly predicted by the model.
-For classification tasks, misclassified instances are circled in red.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
-
-While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of instances to explore the local explanations.
-The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
-These instances are highlighted as asterisk and \(\times\), respectively.
-
-\hypertarget{radial-tour}{%
-\subsection{Radial tour}\label{radial-tour}}
-
-The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
-These are depicted as vertical parallel coordinate plots (Ocagne 1885) on the basis biplot (Gabriel 1971).
-1D biplot display depicts the values of the current basis as bars.
-The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
-The attribution projections of the PI and CI are shown as dashed and dotted lines.
-From this plot, the range and density of the importance across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature.
-For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting they are important for some instances and not important for other instances.
-
-The overlaid bars on the parallel coordinate plot represent the attribution projection of the PI.
-(Remember that the PI is interactively selected from the global view).
-The attribution projection is approximates the feature importance for predicting this instance.
-This combination of features best explains the difference between the mean response and an instance's predicted value.
-It is not an indication of the local shape of the model surface.
-That is, it is not some indication of the tangent to the curve at this point.
-
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
-The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The radial tour varies the contribution of the selected feature between 0-1.
-This is viewed as an animation of the projections from many intermediate bases.
-Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
-For classification, if the separation between classes diminishes when the feature contribution is reduced, this suggests that the feature is important for class separation.
-For regression, if the relationship scatterplot weakens when the feature contribution is reduced, indicating that the feature is important for accurately predicting the response.
-
-\hypertarget{classification-task}{%
-\subsection{Classification task}\label{classification-task}}
-
-Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI makes it easier to examine the features most responsible for the error.
-The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
-The radial tour is 1D and displays as density where color indicates class.
-An animation slider enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}
-
-}
-
-\caption{Overview of the cheem viewer for classification tasks.
-Global view inputs, (a), set the PI, CI, and color statistic.
-Global view, (b) PC1 by PC2 approximations of the data space and attribution space.
-(c) prediction by observed y (visual of the confusion matrix for classification tasks).
-Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (d) select features to include and which feature is changed in the tour.
-(e) shows a parallel coordinate display of the distribution of the feature attributions while bars depict contribution for the current basis.
-The black bar is the variable being changed in the radial tour.
-Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
-\end{figure}
-
-\hypertarget{regression-task}{%
-\subsection{Regression task}\label{regression-task}}
-
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance with similar feature values as CI is a helpful way to understand how the model is failing or not.
-The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation aid in understanding the structure identified in these spaces.
-
-In the radial tour view, the observed response and the residuals (vertical) are plotted against the attribution projection of the PI (horizontal).
-The attribution projection can be interpreted similarly to the predicted value from the global view plot.
-It represents a linear combination of the features, and a good fit would be indicated when there is a strong relationship with the observed values.
-This can be viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a feature is varied, if the value of the PI does not change much, it would indicate that the prediction for this instance is NOT sensitive to that feature.
-Conversely, if the predicted value varies substantially, the prediction is very sensitive to that feature, suggesting that the feature is very important for the PI's prediction.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions}
-
-}
-
-\caption{Overview of the cheem viewer for regression  and illustration of interactive features.
-Panel (a) PCA of the data and attributions spaces, (b) residual plot, predictions by observed values.
-Four selected points are highlighted in the PC spaces and tabularly displayed.
-Coloring on a statistic (c) highlights structure organized in the attribution space.
-Interactive tabular display (d) populates when instances are selected.
-Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (middle and right).}\label{fig:regressioncase}
-\end{figure}
-
-\hypertarget{interactive-features}{%
-\subsection{Interactive features}\label{interactive-features}}
-
-The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a \& d).
-The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
-
-A tooltip displays instance number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
-The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
-These interactions aid exploration of the spaces and, finally, the identification of primary and comparison instances.
-
-\hypertarget{preprocessing}{%
-\subsection{Preprocessing}\label{preprocessing}}
-
-It is vital to mitigate the render time of visuals, especially when users may want to iterate many explorations.
-All computational operations should be prepared before runtime.
-The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
-Below discusses the steps and details of the reprocessing.
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
-\item
-  \textbf{Model:} any model and compatible explanation could be explored with this method.
-Currently, random forest models are applied via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, number variables randomly sampled at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
-\item
-  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} (Kominsarczyk et al.
-2021).
-This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of variables.
-\item
-  \textbf{Cheem viewer:} after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
-Statistics for mapping to color are computed on the scaled spaces.
-\end{itemize}
-
-The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
-For reference, the FIFA data contained 5000 instances of nine explanatory features took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each instance took 270 seconds total.
-PCA and statistics of the features and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
-An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-(Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
-
-\hypertarget{sec:infrastructure}{%
-\subsection{Package infrastructure}\label{sec:infrastructure}}
-
-The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002) and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
-2021).
-The application was made with \textbf{shiny} (Chang et al.
-2021).
-The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
-Both views are created first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive \texttt{html} widgets with \textbf{plotly} (Sievert 2020).
-\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) were a boon to understanding local explanations and how to apply them.
-
-\hypertarget{installation-and-getting-started}{%
-\subsection{Installation and getting started}\label{installation-and-getting-started}}
-
-The package can be installed from CRAN, and the shiny app can be run, using the following \textbf{R} code:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
-\FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
-\FunctionTok{run\_app}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-Alternatively,
-
-\begin{itemize}
-\tightlist
-\item
-  a version of the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
-\item
-  the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
-\item
-  documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
-\end{itemize}
-
-Follow the examples provided with the package to compute the local explainers (using \texttt{?cheem\_ls}).
-The application expects the output returned by \texttt{cheem\_ls()}, saved to an .rds file with \texttt{saveRDS()}, to be uploaded.
-
-\hypertarget{sec:casestudies}{%
-\section{Case studies}\label{sec:casestudies}}
-
-To illustrate the cheem method, it is applied to modern datasets, two classification examples and then two of regression.
-
-\hypertarget{palmer-penguin-species-classification}{%
-\subsection{Palmer penguin, species classification}\label{palmer-penguin-species-classification}}
-
-The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), body mass (\texttt{bm}), for this illustration.
-A random forest model was fit with species as the response feature.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins}
-
-}
-
-\caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
-The radial view shows varying the contribution of \texttt{fl} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at \url{https://vimeo.com/666431172}.)}\label{fig:casepenguins}
-\end{figure}
-
-Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
-Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}
-
-}
-
-\caption{Checking what is learned from the cheem viewer.
-This is a plot of flipper length (\texttt{fl}) and bill length (\texttt{bl}), where an asterisk highlights the PI.
-A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small \texttt{fl} length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
-\end{figure}
-
-There is more separation visible in the attribution space than the data space, as would be expected.
-The predicted vs observed plot reveals a handful of misclassified instances.
-A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
-The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
-The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
-
-The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance are mostly \texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
-That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this, the radial tour is used to vary the contribution of flipper length (\texttt{fl}).
-(In our exploration, this was the third feature explored.
-It is typically helpful to explore the features with more significant contributions, here \texttt{bl} and \texttt{bd}.
-Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
-On varying \texttt{fl} as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
-This suggests that \texttt{fl} should be considered an important feature for explaining the (wrong) prediction.
-
-Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{fl}) is vital for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two features, mainly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
-
-\hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
-
-The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
-Dark chocolate is considered healthier than milk.
-The data were collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
-The data is available in the \textbf{cheem} package.
-A random forest model is used for the classification of chocolate types.
-
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate, nutritionally more like milk, should not be considered a healthy alternative.
-It is interesting to explore which nutritional features contribute most to misclassification.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates}
-
-}
-
-\caption{Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
-So, the contribution of Sugar is varied --- reducing it corresponds primarily with increasing Fiber.
-When Sugar is zero, Fiber contributes strongly towards the left.
-In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
-This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)}\label{fig:casechocolates}
-\end{figure}
-
-This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (instance 22).
-It is compared with a CI that is a correctly classified dark chocolate (instance 7).
-The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
-The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
-
-The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
-In the density plot, instances on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be interpreting a point with high values in this plot.
-The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
-The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
-
-The PI (dashed line) can be viewed against the CI (dotted line).
-Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the PI.
-The feature contributions to the two different predictions can be quickly compared in the parallel coordinate plot.
-The PI differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
-
-From the density plot, which is the attribution projection corresponding to the PI, both instances are more like dark chocolates.
-Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
-When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
-
-It would also be interesting to explore an inverse misclassification.
-In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
-Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (instance 71).
-The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates_inverse}
-
-}
-
-\caption{Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
-In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed feature importance, with low values relative to other milk chocolates.
-The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
-If Sodium had a larger contribution to the prediction (like in this view).
-the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)}\label{fig:casechocolatesinverse}
-\end{figure}
-
-The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
-Looking at the attribution projection, this is found to be the case.
-Previously, Fiber was essential while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
-
-Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
-The contribution of Sodium is selected to be varied.
-Even in the initial projection, the instance looks slightly more like its observed milk than predicted dark chocolate.
-The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
-
-\hypertarget{fifa-wage-regression}{%
-\subsection{FIFA, wage regression}\label{fifa-wage-regression}}
-
-The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
-Nine higher-level skill groupings were identified and aggregated from highly correlated features.
-A random forest model is fit from these predictors, regressing player wages {[}2020 euros{]}.
-The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
-Continuing from the exploration in section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
-A leading offensive fielder (L.
-Messi) is compared with a top defensive fielder (V.
-van Dijk).
-The same instances were used in figure \ref{fig:shapdistrbd}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{./figures/case_fifa}
-
-}
-
-\caption{Exploring the wages {[}euros{]} relative to skill measurements in the FIFA 2020 data.
-Star offensive player (L.
-Messi) is the PI and he is compared with a top defensive player (V.
-van Dijk).
-The attribution projection is shown at left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
-Defense (\texttt{def}) is the chosen feature to vary.
-It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
-The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)}\label{fig:casefifa}
-\end{figure}
-
-Figure \ref{fig:casefifa}, tests the support of the local explanation.
-Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
-If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
-However, Increasing the contribution of a variable with low importance would rotate both features out of the frame.
-
-The contribution from \texttt{def} will be varied to contrast with offensive skills.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group.
-Players with high values in defensive skills are now the rightmost points.
-In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
-
-\hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
-
-Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales.
-A random forest model has fitted this price {[}USD{]} with the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
-Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{./figures/case_ames2018}
-
-}
-
-\caption{Exploring an instance with a large residual as the PI from fitting sales price {[}USD{]} to other variables in the Ames housing 2018.
-The sale price of the PI was under-predicted.
-The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}).
-The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
-In the attribution projection, the PI has a higher sales price than the CI.
-Reducing the contribution of Lot Area brings these two prices in line.
-This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
-That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at \url{https://vimeo.com/666431134}.)}\label{fig:caseames}
-\end{figure}
-
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
-The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-The attribution projection places instances with high Living Areas to the right.
-The contribution of Living Area contrasts the contribution of this feature.
-As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
-This large of importance in the Living Area is relatively uncommon.
-Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
-
-\hypertarget{sec:cheemdiscussion}{%
-\section{Discussion}\label{sec:cheemdiscussion}}
-
-There is a clear need to extend the interpretability of black-box models.
-This chapter provides a technique that builds on local interpretations to explore the feature importance local to an instance.
-The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
-Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
-A global view shows the data space, explanation space, residual plot.
-The user can interactively select instances to compare, contrast, and study further.
-Then the radial tour is used to explore the feature sensitivity identified by the attribution projection.
-
-This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
-In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
-This was verified by making a follow-up plot of the data.
-The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
-In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used by the random forest model.
-
-This analysis is manually intensive and thus only feasible for investigating a few instances.
-The approach selects instances that the model has not done well to predict and compare it with an instance where it did fit well.
-The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and responses to cross-check interpretations made from the cheem viewer.
-This methodology provides an additional tool in the box for studying model fitting.
-
-An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
-Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} Greenwell et al.
-2020; Shi et al.
-2022; Liaw and Wiener 2002; Wright and Ziegler 2017; Chen et al.
-2021, respectively).
-Tree SHAP was selected because of its computational efficiency.
-The SHAP and oscillation explanations could be added with the use of the \texttt{DALEX::explain()} and would be an excellent direction to extend the work (Biecek 2018; Biecek and Burzykowski 2021).
-
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\(^2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
-Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
-
-\leavevmode\hypertarget{ref-anderson_introduction_1995}{}%
-Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT press.
-
-\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
-Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
-2020.
-{``Explainable {Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies, Opportunities and Challenges Toward Responsible {AI}.''} \emph{Information Fusion} 58: 82--115.
-
-\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
-Asimov, Daniel.
-1985.
-{``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
-https://doi.org/\url{https://doi.org/10.1137/0906011}.
-
-\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
-Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
-
-\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
----------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
-
-\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
-Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
-
-\leavevmode\hypertarget{ref-boser_training_1992}{}%
-Boser, Bernhard E., Isabelle M.
-Guyon, and Vladimir N.
-Vapnik.
-1992.
-{``A Training Algorithm for Optimal Margin Classifiers.''} In \emph{Proceedings of the Fifth Annual Workshop on {Computational} Learning Theory}, 144--52.
-
-\leavevmode\hypertarget{ref-breiman_random_2001}{}%
-Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45 (1): 5--32.
-
-\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
----------. 2001b. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
-
-\leavevmode\hypertarget{ref-buja_grand_1986}{}%
-Buja, Andreas, and Daniel Asimov.
-1986.
-{``Grand {Tour} {Methods}: {An} {Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer} {Science} and {Statistics}}, 63--67.
-New York, NY, USA: Elsevier North-Holland, Inc.
-\url{http://dl.acm.org/citation.cfm?id=26036.26046}.
-
-\leavevmode\hypertarget{ref-chang_shiny_2021}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
-2021.
-\emph{Shiny: Web Application Framework for r}.
-\url{https://CRAN.R-project.org/package=shiny}.
-
-\leavevmode\hypertarget{ref-chen_xgboost_2021}{}%
-Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al.
-2021.
-{``Xgboost: {Extreme} {Gradient} {Boosting}.''} \url{https://CRAN.R-project.org/package=xgboost}.
-
-\leavevmode\hypertarget{ref-cook_manual_1997}{}%
-Cook, Dianne, and Andreas Buja.
-1997.
-{``Manual {Controls} for {High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
-\url{https://doi.org/10.2307/1390747}.
-
-\leavevmode\hypertarget{ref-cook_grand_2008}{}%
-Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
-2008.
-{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual} {Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
-Berlin, Heidelberg: Springer Berlin Heidelberg.
-\url{https://doi.org/10.1007/978-3-540-33037-0_13}.
-
-\leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
-da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee.
-2021.
-{``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
-
-\leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
-Dastin, Jeffrey.
-2018.
-{``Amazon Scraps Secret {AI} Recruiting Tool That Showed Bias Against Women.''} \emph{Reuters}, October.
-\url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.
-
-\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
-Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
-2018.
-{``Addressing Age-Related Bias in Sentiment Analysis.''} In \emph{Proceedings of the 2018 Chi Conference on Human Factors in Computing Systems}, 1--14.
-
-\leavevmode\hypertarget{ref-duffy_apple_2019}{}%
-Duffy, Claire.
-2019.
-{``Apple Co-Founder {Steve} {Wozniak} Says {Apple} {Card} Discriminated Against His Wife.''} \emph{CNN}, November.
-\url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.
-
-\leavevmode\hypertarget{ref-espadoto_toward_2021}{}%
-Espadoto, Mateus, Rafael M.
-Martins, Andreas Kerren, Nina S.
-T.
-Hirata, and Alexandru C.
-Telea.
-2021.
-{``Toward a {Quantitative} {Survey} of {Dimension} {Reduction} {Techniques}.''} \emph{IEEE Transactions on Visualization and Computer Graphics} 27 (3): 2153--73.
-\url{https://doi.org/10.1109/TVCG.2019.2944182}.
-
-\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
-Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
-
-\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
-Gorman, Kristen B., Tony D.
-Williams, and William R.
-Fraser.
-2014.
-{``Ecological Sexual Dimorphism and Environmental Variability Within a Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS One} 9 (3): e90081.
-
-\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
-Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
-
-\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
-Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
-
-\leavevmode\hypertarget{ref-greenwell_gbm_2020}{}%
-Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M. Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression} {Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
-
-\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
-Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
-Gorman.
-2020.
-{``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
-
-\leavevmode\hypertarget{ref-karwowski_international_2006}{}%
-Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
-
-\leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
-Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using {AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring Tool.''} \emph{Researchgate Preprint}.
-
-\leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
-Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
-
-\leavevmode\hypertarget{ref-larson_how_2016}{}%
-Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin.
-2016.
-{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''} \emph{ProPublica}, May.
-\url{https://www.propublica.org/article/how-we-analyzed-the-compas-recidivism-algorithm?token=RPR1E2qtzJltfJ0tS-gB_41kmfoWZAu4}.
-
-\leavevmode\hypertarget{ref-lee_state_2021}{}%
-Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
-Sherry Zhang.
-2021.
-{``The State-of-the-Art on Tours for Dynamic Visualization of High-Dimensional Data.''} \emph{WIREs Computational Statistics} n/a (n/a): e1573.
-\url{https://doi.org/10.1002/wics.1573}.
-
-\leavevmode\hypertarget{ref-lee_pptree_2013}{}%
-Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
-
-\leavevmode\hypertarget{ref-leone_fifa_2020}{}%
-Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
-
-\leavevmode\hypertarget{ref-liaw_classification_2002}{}%
-Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
-
-\leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
-Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
-
-\leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
-Lundberg, Scott M., and Su-In Lee.
-2017.
-{``A Unified Approach to Interpreting Model Predictions.''} In \emph{Proceedings of the 31st International Conference on Neural Information Processing Systems}, 4768--77.
-
-\leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
-Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
-
-\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
-O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
-
-\leavevmode\hypertarget{ref-ocagne_coordonnees_1885}{}%
-Ocagne, Maurice d'.
-1885.
-\emph{Coordonnées Parallèles Et Axiales.
-{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul Graphique déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
-
-\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
-Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
-2016.
-{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} In \emph{Proceedings of the 22nd {ACM} {SIGKDD} {International} {Conference} on {Knowledge} {Discovery} and {Data} {Mining}}, 1135--44.
-{KDD} '16.
-New York, NY, USA: Association for Computing Machinery.
-\url{https://doi.org/10.1145/2939672.2939778}.
-
-\leavevmode\hypertarget{ref-roberts_state_2007}{}%
-Roberts, Jonathan C.
-2007.
-{``State of the Art: {Coordinated} \& Multiple Views in Exploratory Visualization.''} In \emph{Fifth International Conference on Coordinated and Multiple Views in Exploratory Visualization ({CMV} 2007)}, 61--71.
-IEEE.
-
-\leavevmode\hypertarget{ref-shapley_value_1953}{}%
-Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
-
-\leavevmode\hypertarget{ref-shi_lightgbm_2022}{}%
-Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas Finley, Taifeng Wang, et al.
-2022.
-{``Lightgbm: {Light} {Gradient} {Boosting} {Machine}.''} \url{https://CRAN.R-project.org/package=lightgbm}.
-
-\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
-Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
-
-\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
-Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
-2017.
-{``Learning Important Features Through Propagating Activation Differences.''} In \emph{International {Conference} on {Machine} {Learning}}, 3145--53.
-PMLR.
-
-\leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
-Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
-2016.
-{``Not Just a Black Box: {Learning} Important Features Through Propagating Activation Differences.''} \emph{arXiv Preprint arXiv:1605.01713}.
-
-\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
-Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
-
-\leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
-Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
-2014.
-{``Deep Inside Convolutional Networks: {Visualising} Image Classification Models and Saliency Maps.''} In \emph{In {Workshop} at {International} {Conference} on {Learning} {Representations}}.
-Citeseer.
-
-\leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
-Spyrison, Nicholas, and Dianne Cook.
-2020.
-{``Spinifex: An {R} {Package} for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of {Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
-\url{https://doi.org/10.32614/RJ-2020-027}.
-
-\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
-Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
-
-\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
-Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
-
-\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
-Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
-2018.
-{``Textual {Deconvolution} {Saliency} ({TDS}): A Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the 56th {Annual} {Meeting} of the {Association} for {Computational} {Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
-
-\leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
-Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
-
-\leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
-Wickham, Hadley, Dianne Cook, and Heike Hofmann.
-2015.
-{``Visualizing Statistical Models: {Removing} the Blindfold.''} \emph{Statistical Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
-\url{https://doi.org/10.1002/sam.11271}.
-
-\leavevmode\hypertarget{ref-wright_ranger_2017}{}%
-Wright, Marvin N., and Andreas Ziegler.
-2017.
-{``Ranger: {A} {Fast} {Implementation} of {Random} {Forests} for {High} {Dimensional} {Data} in {C}++ and {R}.''} \emph{Journal of Statistical Software} 77 (1): 1--17.
-\url{https://doi.org/10.18637/jss.v077.i01}.
-
-\end{CSLReferences}
+\begin{verbatim}
+jmlroption: preprint
+\end{verbatim}
 
@@
 
-\acks{We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.}
+\acks{This research was supported in part by the\ldots{}}
 
@@
 \vskip 0.2in
-\bibliography{cheem\_paper.bib}
+\bibliography{references.bib}
 

### 2022-03-16 template_rjh/jmlr.Rmd
Use RJH's jmlr template. Figure sizes and html consistency.

---
title: "Title"
shorttitle: "Shorter title"
author:
- familyname: Cheng
  othernames: Fan
  address: Monash University\newline Australia
  email: Fan.Cheng@monash.edu
  correspondingauthor: true
- familyname: Panagiotelis
  othernames: Anastasios
  address: University of Sydney\newline Australia
  email: Anastasios.Panagiotelis@sydney.edu.au
- familyname: Hyndman
  othernames: Rob J
  address: Monash University\newline Australia
  email: Rob.Hyndman@monash.edu
year: 2022
date_submitted: "03/22"
abstract: "Blah blah blah"
keywords: "dimension reduction, divergence estimation, Hellinger distance, anomaly detection, k-d trees"
acknowledgment: "This research was supported in part by the..."
bibliography: references.bib
classoption: 11pt,twoside
output:
  bookdown::pdf_document2:
    template: jmlr_template.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: natbib
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  messages = FALSE,
  warning = FALSE,
  eval = TRUE
  # include = FALSE
)
library(tidyverse)
```

\newpage

# Introduction

Blah blah. You can refer to papers [@Witten2011] or cite them directly as in @AriasCastro2020.

Add the following to the yaml header if you want to remove the JMLR details.

```
jmlroption: preprint
```

### 2022-03-26 _paper.rmd
guidelines, checklist & acknowledgments cleanup

@@
   email: Kim.Marriott@monash.edu
-# Nick ORCiD 0000-0002-8417-0212
-# Di   ORCiD 0000-0002-3813-7155
-# Kim  ORCiD 0000-0002-9813-0377
 year: 2022
@@
 The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models."
+DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.\\newline The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models."
 bibliography: cheem_paper.bib
@@
     fig_width: 8
-    # includes:
-    #   in_header: template_rjh/preamble.tex
     keep_tex: yes
@@
 
-```{r loadlibraries, include=FALSE, cache=FALSE}
+```{r include=FALSE, cache=FALSE}
+## Paper compilation dependancies
 library("knitr")
@@
 library("magrittr")
-## Work packages
-library("cheem")
-library("spinifex")
+
 ## chunk options
@@
 
-# Acknowledgments
-
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI$^2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
-
 # References

### 2022-03-26 _paper.tex
guidelines, checklist & acknowledgments cleanup

@@
 
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI\(^2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
-
 \hypertarget{references}{%
@@
 The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.}
+DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.\newline The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.}
 

### 2022-03-26 _paper.rmd
Grammerly edits. figure size. checklist & guideline escaping.

@@
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network to name a few [@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995, receptively].
+Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network, to name a few [@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995, receptively].
 
@@
 Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
-This difficulty in interpreting the terms in complex nonlinear models sometimes leads to them being referred to as black box models.
+This difficulty in interpreting the terms in complex nonlinear models sometimes lead to them being referred to as black box models.
 Despite the potentially better performance of nonlinear models, their use is not without controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
@@
 <!-- Interpretability & baises -->
-Interpretability is vital for exploring and protecting against potential biases in any model (e.g.
-sex -- @dastin_amazon_2018; @duffy_apple_2019, race -- @larson_how_2016, and age -- @diaz_addressing_2018).
+Interpretability is vital for exploring and protecting against potential biases in any model (e.g., sex -- @dastin_amazon_2018; @duffy_apple_2019, race -- @larson_how_2016, and age -- @diaz_addressing_2018).
 For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
@@
 <!-- Local explanations -->
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models [@adadi_peeking_2018; @arrieta_explainable_2020].
+Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black box models [@adadi_peeking_2018; @arrieta_explainable_2020].
 A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
@@
 <!-- Data visualization tours -->
-In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear projections of data onto a lower-dimensional space.
+In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear data projections onto a lower-dimensional space.
 Tours are viewed as an animation over minor changes to the projection basis.
@@
 <!--tours and models --> Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015], and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
+Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015], and using tree-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
 
@@
 Ceteris paribus, is Latin for "other things held constant" or "all else unchanged".
-These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The explanations of SHAP refer to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
+The SHAP explanation refers to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
 @strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
@@
 Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, helps to understand the relative importance of each feature in the context of all other candidate features.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, help to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 For example, Figure \@ref(fig:classificationcase)e suggests that `bl` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, and which have a range of high and low values (e.g.
-`off`, `def`) suggesting they are important for some instances and not important for other instances.
+For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, and which have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some instances and not important for other instances.
 
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection is approximates the feature importance for predicting this instance.
+The attribution projection approximates the feature importance for predicting this instance.
 This combination of features best explains the difference between the mean response and an instance’s predicted value.
@@
 
-The package can be installed from CRAN, and the shiny app can be run, using the following __R__ code:
+The package can be installed from CRAN, and the shiny app can be run using the following __R__ code:
 
@@
 Follow the examples provided with the package to compute the local explainers (using `?cheem_ls`).
-The application expects the output returned by `cheem_ls()`, saved to an .rds file with `saveRDS()`, to be uploaded.
+The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()`, to be uploaded.
 
@@
 Defense (`def`) is the chosen feature to vary.
-It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
+It starts off very low, and as its contribution increases (right plot) Messi's predicted wages decrease dramatically.
 The increased contribution in defense comes at the expense of offensive and reaction skills.
@@
 (A video showing the animation is at https://vimeo.com/666431134.)
-```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames-cap)"}
+```{r caseames, echo=F, out.width="90%", fig.cap = "(ref:caseames-cap)"}
 knitr::include_graphics("./figures/case_ames2018.png")
@@
 
-There is a clear need to extend the interpretability of black-box models.
+There is a clear need to extend the interpretability of black box models.
 This chapter provides a technique that builds on local interpretations to explore the feature importance local to an instance.
@@
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ [tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively].
+In theory, this approach would work with any black box model, but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ [tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively].
 Tree SHAP was selected because of its computational efficiency.

### 2022-03-26 _paper.tex
Grammerly edits. figure size. checklist & guideline escaping.

@@
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network to name a few (Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995, receptively).
+Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network, to name a few (Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995, receptively).
 
 Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
-This difficulty in interpreting the terms in complex nonlinear models sometimes leads to them being referred to as black box models.
+This difficulty in interpreting the terms in complex nonlinear models sometimes lead to them being referred to as black box models.
 Despite the potentially better performance of nonlinear models, their use is not without controversy (O'Neil 2016; Kodiyan 2019).
@@
 
-Interpretability is vital for exploring and protecting against potential biases in any model (e.g.~sex -- Dastin (2018); Duffy (2019), race -- Larson et al.
+Interpretability is vital for exploring and protecting against potential biases in any model (e.g., sex -- Dastin (2018); Duffy (2019), race -- Larson et al.
 (2016), and age -- Díaz et al.
@@
 
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black-box models (Adadi and Berrada 2018; Arrieta et al.
+Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black box models (Adadi and Berrada 2018; Arrieta et al.
 2020).
@@
 Lee et al.
-2021) is a sequence of linear projections of data onto a lower-dimensional space.
+2021) is a sequence of linear data projections onto a lower-dimensional space.
 Tours are viewed as an animation over minor changes to the projection basis.
@@
 Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree- and forest-based approaches as a projection pursuit index to generate a tour basis paths (Y.
+Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree-based approaches as a projection pursuit index to generate a tour basis paths (Y.
 D.
@@
 Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant.
+Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The explanations of SHAP refer to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
+The SHAP explanation refers to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
 Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
@@
 Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), helps to understand the relative importance of each feature in the context of all other candidate features.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), help to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g.~\texttt{off}, \texttt{def}) suggesting they are important for some instances and not important for other instances.
+For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some instances and not important for other instances.
 
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection is approximates the feature importance for predicting this instance.
+The attribution projection approximates the feature importance for predicting this instance.
 This combination of features best explains the difference between the mean response and an instance's predicted value.
@@
 
-The package can be installed from CRAN, and the shiny app can be run, using the following \textbf{R} code:
+The package can be installed from CRAN, and the shiny app can be run using the following \textbf{R} code:
 
@@
 Follow the examples provided with the package to compute the local explainers (using \texttt{?cheem\_ls}).
-The application expects the output returned by \texttt{cheem\_ls()}, saved to an .rds file with \texttt{saveRDS()}, to be uploaded.
+The application expects the output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file with \texttt{saveRDS()}, to be uploaded.
 
@@
 Defense (\texttt{def}) is the chosen feature to vary.
-It starts off very low, and as its contribution is increased (right plot) it can be seen that Messi's predicted wages decrease dramatically.
+It starts off very low, and as its contribution increases (right plot) Messi's predicted wages decrease dramatically.
 The increased contribution in defense comes at the expense of offensive and reaction skills.
@@
 
-{\centering \includegraphics[width=1\linewidth]{./figures/case_ames2018}
+{\centering \includegraphics[width=0.9\linewidth]{./figures/case_ames2018}
 
@@
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black-box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} Greenwell et al.
+In theory, this approach would work with any black box model, but the implementation currently only calculates tree SHAP for tree-based models supported by \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} Greenwell et al.
 2020; Shi et al.

### 2022-03-26 template_rjh/author_checklist.rmd
guidelines, checklist & acknowledgments cleanup
Grammerly edits. figure size. checklist & guideline escaping.

--- 2022-01-08 template/author_checklist.rmd
@@
 author: "Nick Spyrison"
-date: "1/8/2022"
+date: "8/1/2022"
 output: html_document
@@
 - Make sure your abstract reads well. It will be used to advertise your paper.
-- We recommend you use small caps for system names. (In LaTex that's \sc)
+- We recommend you use small caps for system names. (In LaTex that's \\sc)
 - Move file to have correct name, according to the JMLR naming conventions. See the authors' guide.
@@
 - Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References.
-- Acknowledgments should be a unnumbered section. You can use the \acks{} command for this.
-- Appendix section number should be a letter (Should look like Appendix A. Title). You can use \appendix followed by regular \section command to format appendices
+- Acknowledgments should be a unnumbered section. You can use the \\acks{} command for this.
+- Appendix section number should be a letter (Should look like Appendix A. Title). You can use \\appendix followed by regular \\section command to format appendices
 
@@
 - Author contact addresses: These should be complete enough that mail sent to that address will reach them
-- Editor listed: Please include the name of the Action Editor who handled your submission in the \editor{} tag
-- Keyword: Please add keywords to your article using \begin{keywords} ...
-\end{keywords} after your abstract.
+- Editor listed: Please include the name of the Action Editor who handled your submission in the \\editor{} tag
+- Keyword: Please add keywords to your article using \\begin{keywords} ...
+\\end{keywords} after your abstract.
 They should not be capitalised, they should be separated by commas, and they should not have a period at the end.
-- Abstract: Do not delete the % at the end of \begin{abstract}% as it introduces an extra indent
+- Abstract: Do not delete the % at the end of \\begin{abstract}% as it introduces an extra indent
 - Title should be split across two lines, only when it is too long for one line.
@@
 ## Headers and Footers
-- Front page header: Use the \jmlrheading command to get a correct header and footer on the title page
-- Copyright notice, bottom of front page uses authors full names: In the \jmlrheading command, please use the full names of all authors
-- Running headers: Add running headers to your paper using the \ShortHeadings command. Please see the example paper on the web site for how to do this.
-- Running headers: In the \ShortHeadings command, please capitalise the words in your paper title.
-- Running headers: In the author list in the \ShortHeadings command, please use "and" before the last author
+- Front page header: Use the \\jmlrheading command to get a correct header and footer on the title page
+- Copyright notice, bottom of front page uses authors full names: In the \\jmlrheading command, please use the full names of all authors
+- Running headers: Add running headers to your paper using the \\ShortHeadings command. Please see the example paper on the web site for how to do this.
+- Running headers: In the \\ShortHeadings command, please capitalise the words in your paper title.
+- Running headers: In the author list in the \\ShortHeadings command, please use "and" before the last author
 - Running headers: The ShortHeadings author list should have last names only, and should list all authors if possible. If you have many authors, "et al" is acceptable.
@@
 
-- No unnumbered sections: We do not use \paragraph or \section*.
+- No unnumbered sections: We do not use \\paragraph or \\section*.
 We do not allow bold, unindented text to begin a paragraph that could be mistaken for a section header.
 If it looks like a section header it needs to be numbered like a section header, otherwise it needs to be formatted in some other way (e.g., by indentation, or by using italics instead of boldface).
-- Section numbering: All sections, subsections, subsubsubsections must be numbered. Please use \subsubsection instead of \subsubsection*, etc
+- Section numbering: All sections, subsections, subsubsubsections must be numbered. Please use \\subsubsection instead of \\subsubsection*, etc
 Capitalization of Section headings: All section headings should be title case.
@@
 
-- Float placement: In general, floats should be at the top or bottom of the page (avoid using [h] option to \begin{figure}).
+- Float placement: In general, floats should be at the top or bottom of the page (avoid using [h] option to \\begin{figure}).
 Check for pages that contain several floats and only 2-3 lines of text.
@@
 
-- use \url commands to reference urls in papers. This causes hyperref to make the hyperlinks work nicely.
-- Parenthesis: \eqref can sometimes result in parentheses within parentheses. When referencing an equation within a parenthetical statement, use Equation 1 rather than (1) or Equation (1).
+- use \\url commands to reference urls in papers. This causes hyperref to make the hyperlinks work nicely.
+- Parenthesis: \\eqref can sometimes result in parentheses within parentheses. When referencing an equation within a parenthetical statement, use Equation 1 rather than (1) or Equation (1).
 - When referencing figures, sections, etc. ("Figure 1", "Section 2.1"), please make sure the words "Figure", "Section", etc., are uppercase

### 2022-03-26 template_rjh/author_guide.rmd
guidelines, checklist & acknowledgments cleanup
Grammerly edits. figure size. checklist & guideline escaping.

--- 2022-01-08 template/author_guide.rmd
@@
+---
+title: "Author guide notes"
+author: "Nick Spyrison"
+date: "8/1/2022"
+output: html_document
+---
+
 Following: https://www.jmlr.org/format/authors-guide.html
@@
 This archive should contain the latex source, the .bib file, figures, and everything needed to compile your document.
-The archive should contain only one .tex file (please do not use \include or \input.) We process your paper via an automatic system, so please follow these rules:
+The archive should contain only one .tex file (please do not use \\include or \\input.) We process your paper via an automatic system, so please follow these rules:
 
@@
 
-Your page numbers should start at 1, both in the \jmlrheading command and in your actual paper.
+Your page numbers should start at 1, both in the \\jmlrheading command and in your actual paper.
 
-You file should include an \editor{...} command that lists the action editor who has corresponded with you.
+You file should include an \\editor{...} command that lists the action editor who has corresponded with you.
 
-Your \jmlrheading command should contain the correct year, volume number, publication dates, etc. You must fill in all of the information. The format for this is:
+Your \\jmlrheading command should contain the correct year, volume number, publication dates, etc. You must fill in all of the information. The format for this is:
 
-   \jmlrheading{volume}{year}{pages}{dates when your paper submitted}{date published}{your paper id}{all authors full names}
+   \\jmlrheading{volume}{year}{pages}{dates when your paper submitted}{date published}{your paper id}{all authors full names}
 For example, suppose that you submitted your paper to the production system in April 2017. The current volume at that time was 18 (you can find the current volume on the front page of jmlr.org)
 
-       \jmlrheading{18}{2017}{1-36}{8/14; Revised 12/15}{4/17}{15-449}{John
-Smith, Jane Doe, and Frank Jones}
+       \\jmlrheading{18}{2017}{1-36}{8/14; Revised 12/15}{4/17}{15-449}{JohnSmith, Jane Doe, and Frank Jones}
 Please use the current month and year as the publication date.
@@
 Our automated system does not run bibtex (this is similar to what arxiv.org does). Therefore if your paper uses bibtex, you should include the bbl file that is necessary to compile your final paper.
-The scripts specifically check whether the paper headings are correct (e.g., page numbers, \jmlrheading, \ShortHeadings, \editor, and so on).
+The scripts specifically check whether the paper headings are correct (e.g., page numbers, \\jmlrheading, \\ShortHeadings, \\editor, and so on).
 To speed processing of your paper, please give special attention to make sure that these are correct and follow the lines of the example file.

### 2022-03-28 _paper.rmd
Working through checklist, Monday.

@@
 The methods are implemented in the __R__ package __cheem__, available on CRAN."
-keywords: "explainable artificial intelligence, nonlinear model interpretability, local explanations, grand tour, radial tour"
+keywords: "explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour"
 acknowledgment: "We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
@@
 There are different reasons and emphases when considering to fit a model.
-@breiman_statistical_2001, reiterated by @shmueli_explain_2010, taxonomize models based on their purpose.
+@breiman_statistical_2001 and @shmueli_explain_2010, taxonomize models based on their purpose.
 _Explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more narrowly on the performance of some objective function.
@@
 <!-- Interpretability & baises -->
-Interpretability is vital for exploring and protecting against potential biases in any model (e.g., sex -- @dastin_amazon_2018; @duffy_apple_2019, race -- @larson_how_2016, and age -- @diaz_addressing_2018).
+Interpretability is vital for exploring and protecting against potential biases in any model, e.g., sex\-\-\-@dastin_amazon_2018; @duffy_apple_2019, race\-\-\-@larson_how_2016, and age\-\-\-@diaz_addressing_2018.
 For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
@@
 The methods are implemented in R package __cheem__.
-Example datasets are provided to illustrate usage for classification and regression tasks.
+Example data sets are provided to illustrate usage for classification and regression tasks.
 
@@
 
-# Local explanations {#sec:explanations}
+# Local Explanations {#sec:explanations}
 
@@
 
-# Tours and the radial tour {#sec:tour}
+# Tours and the Radial Tour {#sec:tour}
 
@@
 
-# The cheem viewer {#sec:cheemviewer}
+# The Cheem Viewer {#sec:cheemviewer}
 
@@
 
-## Global view
+## Global View
 
@@
 
-## Radial tour
+## Radial Tour
 
@@
 
-## Classification task
+## Classification Task
 
@@
 
-## Regression task
+## Regression Task
 
@@
 
-## Interactive features
+## Interactive Features
 
@@
 
-## Package infrastructure {#sec:infrastructure}
+## Package Infrastructure {#sec:infrastructure}
 
@@
 
-## Installation and getting started
+## Installation and Getting Started
 
@@
 
-- a version of the cheem viewer shiny app can be directly accessed at https://ebsmonash.shinyapps.io/cheem_initial/. 
-- the development version of the package is available at https://github.com/nspyrison/cheem, and 
-- documentation of the package can be found at https://nspyrison.github.io/cheem/.
+- A version of the cheem viewer shiny app can be directly accessed at
+\url{https://ebsmonash.shinyapps.io/cheem_initial/}. 
+- The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and 
+- documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
 
@@
 
-# Case studies {#sec:casestudies}
+# Case Studies {#sec:casestudies}
 
-To illustrate the cheem method, it is applied to modern datasets, two classification examples and then two of regression.
+To illustrate the cheem method, it is applied to modern data sets, two classification examples and then two of regression.
 
-## Palmer penguin, species classification
+## Palmer Penguin, Species Classification
 
@@
 The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at https://vimeo.com/666431172.)
+(The animation of the radial tour is at \url{https://vimeo.com/666431172}.)
 ```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins-cap)"}
@@
 
-## Chocolates, milk/dark chocolate classification
+## Chocolates, Milk/Dark Chocolate Classification
 
 <!-- Data description -->
-The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
+The chocolates data set consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
@@
 From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
-So, the contribution of Sugar is varied --- reducing it corresponds primarily with increasing Fiber.
+So, the contribution of Sugar is varied---reducing it corresponds primarily with increasing Fiber.
 When Sugar is zero, Fiber contributes strongly towards the left.
@@
 This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at https://vimeo.com/666431143.)
+(A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)
 ```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates-cap)"}
@@
 the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at https://vimeo.com/666431148.)
+(A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)
 ```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse-cap)"}
@@
 
-## FIFA, wage regression
+## FIFA, Wage Regression
 
@@
 The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
-Continuing from the exploration in section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
+Continuing from the exploration in Section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
 That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
@@
 van Dijk).
-The same instances were used in figure \@ref(fig:shapdistrbd).
+The same instances were used in Figure \@ref(fig:shapdistrbd).
 
@@
 The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at https://vimeo.com/666431163.)
+(A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)
 ```{r casefifa, echo=F, out.width = "90%", fig.cap = "(ref:casefifa-cap)"}
@@
 
-## Ames housing 2018, sales price regression
+## Ames Housing 2018, Sales Price Regression
 
@@
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at https://vimeo.com/666431134.)
+(A video showing the animation is at \url{https://vimeo.com/666431134}.)
 ```{r caseames, echo=F, out.width="90%", fig.cap = "(ref:caseames-cap)"}
@@
 
-An implementation is provided in the open-source __R__ package __cheem__, available on [CRAN](https://CRAN.R-project.org/package=cheem).
+An implementation is provided in the open-source __R__ package __cheem__, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.

### 2022-03-28 _paper.tex
Working through checklist, Monday.

@@
 There are different reasons and emphases when considering to fit a model.
-Breiman (2001b), reiterated by Shmueli (2010), taxonomize models based on their purpose.
+Breiman (2001b) and Shmueli (2010), taxonomize models based on their purpose.
 \emph{Explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more narrowly on the performance of some objective function.
@@
 
-Interpretability is vital for exploring and protecting against potential biases in any model (e.g., sex -- Dastin (2018); Duffy (2019), race -- Larson et al.
-(2016), and age -- Díaz et al.
-(2018)).
+Interpretability is vital for exploring and protecting against potential biases in any model, e.g., sex-\/-\/-Dastin (2018); Duffy (2019), race-\/-\/-Larson et al.
+(2016), and age-\/-\/-Díaz et al.
+(2018).
 For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
@@
 The methods are implemented in R package \textbf{cheem}.
-Example datasets are provided to illustrate usage for classification and regression tasks.
+Example data sets are provided to illustrate usage for classification and regression tasks.
 
@@
 \hypertarget{sec:explanations}{%
-\section{Local explanations}\label{sec:explanations}}
+\section{Local Explanations}\label{sec:explanations}}
 
@@
 \hypertarget{sec:tour}{%
-\section{Tours and the radial tour}\label{sec:tour}}
+\section{Tours and the Radial Tour}\label{sec:tour}}
 
@@
 \hypertarget{sec:cheemviewer}{%
-\section{The cheem viewer}\label{sec:cheemviewer}}
+\section{The Cheem Viewer}\label{sec:cheemviewer}}
 
@@
 \hypertarget{global-view}{%
-\subsection{Global view}\label{global-view}}
+\subsection{Global View}\label{global-view}}
 
@@
 \hypertarget{radial-tour}{%
-\subsection{Radial tour}\label{radial-tour}}
+\subsection{Radial Tour}\label{radial-tour}}
 
@@
 \hypertarget{classification-task}{%
-\subsection{Classification task}\label{classification-task}}
+\subsection{Classification Task}\label{classification-task}}
 
@@
 \hypertarget{regression-task}{%
-\subsection{Regression task}\label{regression-task}}
+\subsection{Regression Task}\label{regression-task}}
 
@@
 \hypertarget{interactive-features}{%
-\subsection{Interactive features}\label{interactive-features}}
+\subsection{Interactive Features}\label{interactive-features}}
 
@@
 \hypertarget{sec:infrastructure}{%
-\subsection{Package infrastructure}\label{sec:infrastructure}}
+\subsection{Package Infrastructure}\label{sec:infrastructure}}
 
@@
 \hypertarget{installation-and-getting-started}{%
-\subsection{Installation and getting started}\label{installation-and-getting-started}}
+\subsection{Installation and Getting Started}\label{installation-and-getting-started}}
 
@@
 \item
-  a version of the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
+  A version of the cheem viewer shiny app can be directly accessed at
+  \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
 \item
-  the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
+  The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
 \item
@@
 \hypertarget{sec:casestudies}{%
-\section{Case studies}\label{sec:casestudies}}
-
-To illustrate the cheem method, it is applied to modern datasets, two classification examples and then two of regression.
+\section{Case Studies}\label{sec:casestudies}}
+
+To illustrate the cheem method, it is applied to modern data sets, two classification examples and then two of regression.
 
 \hypertarget{palmer-penguin-species-classification}{%
-\subsection{Palmer penguin, species classification}\label{palmer-penguin-species-classification}}
+\subsection{Palmer Penguin, Species Classification}\label{palmer-penguin-species-classification}}
 
@@
 \hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}
-
-The chocolates dataset consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
+\subsection{Chocolates, Milk/Dark Chocolate Classification}\label{chocolates-milkdark-chocolate-classification}}
+
+The chocolates data set consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
@@
 \hypertarget{fifa-wage-regression}{%
-\subsection{FIFA, wage regression}\label{fifa-wage-regression}}
+\subsection{FIFA, Wage Regression}\label{fifa-wage-regression}}
 
@@
 The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
-Continuing from the exploration in section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
+Continuing from the exploration in Section \ref{sec:explanations}, we are interested to see the difference in attribution based on the exogenous player position.
 That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
@@
 van Dijk).
-The same instances were used in figure \ref{fig:shapdistrbd}.
+The same instances were used in Figure \ref{fig:shapdistrbd}.
 
@@
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{Ames housing 2018, sales price regression}\label{ames-housing-2018-sales-price-regression}}
+\subsection{Ames Housing 2018, Sales Price Regression}\label{ames-housing-2018-sales-price-regression}}
 
@@
 
-An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
+An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
@@
 {``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
-https://doi.org/\url{https://doi.org/10.1137/0906011}.
+\url{https://doi.org/10.1137/0906011}.
 

### 2022-03-28 template_rjh/author_checklist.rmd
Working through checklist, Monday.

@@
 - We recommend you use small caps for system names. (In LaTex that's \\sc)
-- Move file to have correct name, according to the JMLR naming conventions. See the authors' guide.
-- Signed copyright form: Do not forget to provide this.
-- Text should be in Computer Modern
-- Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References.
-- Acknowledgments should be a unnumbered section. You can use the \\acks{} command for this.
-- Appendix section number should be a letter (Should look like Appendix A. Title). You can use \\appendix followed by regular \\section command to format appendices
+- Move file to have correct name, according to the JMLR naming conventions. See the authors' guide. _NEEDED_
+- Signed copyright form: Do not forget to provide this. _NEEDED_
+- Text should be in Computer Modern.
+- Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References. _Not the order of the template, may need an eye on this._
+- Acknowledgments should be a unnumbered section. You can use the \\acks{} command for this. NS
+- Appendix section number should be a letter (Should look like Appendix A. Title). You can use \\appendix followed by regular \\section command to format appendices. NS
 
@@
 
-- Capitalization of paper title: In the title area, the paper title should be in title case. (First letter of each word capitalized except for little words like "and" and "or".)
-- Author contact addresses: These should be complete enough that mail sent to that address will reach them
-- Editor listed: Please include the name of the Action Editor who handled your submission in the \\editor{} tag
+- Capitalization of paper title: In the title area, the paper title should be in title case. (First letter of each word capitalized except for little words like "and" and "or".). NS
+- Author contact addresses: These should be complete enough that mail sent to that address will reach them. NS
+- Editor listed: Please include the name of the Action Editor who handled your submission in the \\editor{} tag _NS--- todo_
 - Keyword: Please add keywords to your article using \\begin{keywords} ...
@@
 They should not be capitalised, they should be separated by commas, and they should not have a period at the end.
-- Abstract: Do not delete the % at the end of \\begin{abstract}% as it introduces an extra indent
-- Title should be split across two lines, only when it is too long for one line.
+NS
+- Abstract: Do not delete the % at the end of \\begin{abstract}% as it introduces an extra indent. _NS---doesn't get added in .rmd, may need to watch submission_
+- Title should be split across two lines, only when it is too long for one line. NS
 
 ## Headers and Footers
-- Front page header: Use the \\jmlrheading command to get a correct header and footer on the title page
-- Copyright notice, bottom of front page uses authors full names: In the \\jmlrheading command, please use the full names of all authors
-- Running headers: Add running headers to your paper using the \\ShortHeadings command. Please see the example paper on the web site for how to do this.
-- Running headers: In the \\ShortHeadings command, please capitalise the words in your paper title.
-- Running headers: In the author list in the \\ShortHeadings command, please use "and" before the last author
-- Running headers: The ShortHeadings author list should have last names only, and should list all authors if possible. If you have many authors, "et al" is acceptable.
+- Front page header: Use the \\jmlrheading command to get a correct header and footer on the title page. NA
+- Copyright notice, bottom of front page uses authors full names: In the \\jmlrheading command, please use the full names of all authors. NS
+- Running headers: Add running headers to your paper using the \\ShortHeadings command. Please see the example paper on the web site for how to do this. NS
+- Running headers: In the \\ShortHeadings command, please capitalise the words in your paper title. NS
+- Running headers: In the author list in the \\ShortHeadings command, please use "and" before the last author. NS
+- Running headers: The ShortHeadings author list should have last names only, and should list all authors if possible. If you have many authors, "et al" is acceptable. NS
 
@@
 If it looks like a section header it needs to be numbered like a section header, otherwise it needs to be formatted in some other way (e.g., by indentation, or by using italics instead of boldface).
-- Section numbering: All sections, subsections, subsubsubsections must be numbered. Please use \\subsubsection instead of \\subsubsection*, etc
+NS
+- Section numbering: All sections, subsections, subsubsubsections must be numbered. Please use \\subsubsection instead of \\subsubsection*, etc. NS
 Capitalization of Section headings: All section headings should be title case.
 (First letter of each word capitalized except for little words like "and" and "or".) This includes subsections, sub subsections, etc.
-- Section headers: There must be some text separating all section headers
+NS, done
+- Section headers: There must be some text separating all section headers. NS
 
@@
 When this happens, please put the floats on a floats page ([p] specifier) instead.
+NA
 - Floats should also be placed as close as possible to the page at which they are first referenced. You should not, e.g., place all of the figures at the end of the paper.
-Figures, tables should be centered
-- Captions should be underneath figures and tables
-- Captions NOT in italics, bold
-- Tables and figures should not overfill margins
+Figures, tables should be centered. NS
+- Captions should be underneath figures and tables. NS
+- Captions NOT in italics, bold. NS `codeface` and \\url{}, but not egregious.
+- Tables and figures should not overfill margins. NA (manual check, reference urls run over)
 Multiline captions should have a hanging indent, that is, subsequent lines of the caption should be indented, aligning with the first word of the caption rather than the word "Figure".
@@
 Sometimes changing the order of includes will fix this (try moving the jmlr2e include farther down).
+NS
 
@@
 
-- Footnote markers should follow punctuation
-- Footnotes should be complete sentences
+- Footnote markers should follow punctuation. NA
+- Footnotes should be complete sentences. NA
 
@@
 
-- Dashes: When using dashes in text, please use --- instead of --, with no space before or after---like this.
-- Data set: We use "data set" rather than "dataset".
+- Dashes: When using dashes in text, please use --- instead of --, with no space before or after---like this. NS
+- Data set: We use "data set" rather than "dataset". NS
 
@@
 Here are examples of BAD citations:
-    - "See (Ginsberg, 1994) for an analysis..." [Uses citation as a noun]
-    - "Drummond (Drummond, 1994) showed that..." [Should be "Drummond (1994)"]
-    - "...the best method (see the ERT test (Davis, 1995))..." [double parens, should be   "...best method (see the ERT test, Davis, 1995)..."]
+    - "See (Ginsberg, 1994) for an analysis..." [Uses citation as a noun]. NS, looks good
+    - "Drummond (Drummond, 1994) showed that..." [Should be "Drummond (1994)"]. NS, looks good
+    - "...the best method (see the ERT test (Davis, 1995))..." [double parens, should be   "...best method (see the ERT test, Davis, 1995)..."] NA
 - Citations not nested in parenthesis.
@@
 - use \\url commands to reference urls in papers. This causes hyperref to make the hyperlinks work nicely.
-- Parenthesis: \\eqref can sometimes result in parentheses within parentheses. When referencing an equation within a parenthetical statement, use Equation 1 rather than (1) or Equation (1).
-- When referencing figures, sections, etc. ("Figure 1", "Section 2.1"), please make sure the words "Figure", "Section", etc., are uppercase
+    - NS-- makes no difference in text, if fully qualified with https:// at the start.
+- Parenthesis: \\eqref can sometimes result in parentheses within parentheses. When referencing an equation within a parenthetical statement, use Equation 1 rather than (1) or Equation (1). NA
+- When referencing figures, sections, etc. ("Figure 1", "Section 2.1"), please make sure the words "Figure", "Section", etc., are uppercase. NS
 
@@
 - Punctuation: Math in formal writing should be punctuated as if the extra space is not there, for example, displayed equations should usually be followed by a comma or a period, depending on the surrounding text.
-- Punctuation: Before displayed equations, punctuation (e.g., ":") is not necessary unless you would use the same punctuation if the equation were not there.
+NA
+- Punctuation: Before displayed equations, punctuation (e.g., ":") is not necessary unless you would use the same punctuation if the equation were not there. NA
 - In general, only equations that are referenced should be numbered.
 We are willing to relax this for equations that are particularly important, as others may wish to discuss these in later correspondence.
-- Proofs should end with a filled box.
-- Equations should not run over the right hand margin. Appendices are especially prone to this.
-- Theorems, Lemmas, etc., should be labeled like Theorem 1, Theorem 2, Theorem 3, not as Theorem 2.1, Theorem 2.2, Theorem 2.3
+NA
+- Proofs should end with a filled box. NA
+- Equations should not run over the right hand margin. Appendices are especially prone to this. NA
+- Theorems, Lemmas, etc., should be labeled like Theorem 1, Theorem 2, Theorem 3, not as Theorem 2.1, Theorem 2.2, Theorem 2.3 NA
 
 ## References
-- It is preferable to cite published papers, rather than tech reports.
+- It is preferable to cite published papers, rather than tech reports. NS
 - Author names: Make citations consistent. Either give the full first name of all cited authors, or give the first initials only for all cited authors. 
+    - NS: I think rmd defaults to three names before et al. see thesis for .cls (.csl? from rjournal changing this)
 - Capitalization in paper titles: Titles of conference papers should have the first word only titles.
@@
 These should still be capitalized, but BibTeX often leaves uncapitalised incorrectly.
+    - NS: I think this is a bit unrealistic. They may push back about this one.
 - Capitalization of book titles, etc: The parts in italics (book names, journal names, conf names) should be title case.
-(First letter of each word capitalized except for little words like "and" and "or".)
+(First letter of each word capitalized except for little words like "and" and "or".) NS See above subitem comment/.
 - Please cite conference titles consistently.

### 2022-03-30 _paper.rmd
Checklist cleanup & grammerly

@@
   address: Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212
-  email: Nicholas.Spyrison@monash.edu
+  email: nicholas.spyrison@monash.edu
   correspondingauthor: true
@@
   address: Monash University\newline ORCiD 0000-0002-3813-7155
-  email: DiCook@monash.edu
+  email: dicook@monash.edu
 - familyname: Marriott
@@
   address: Monash University\newline ORCiD 0000-0002-9813-0377
-  email: Kim.Marriott@monash.edu
+  email: kim.marriott@monash.edu
 year: 2022
@@
   bookdown::pdf_document2:
-    template: template_rjh/jmlr_template.tex
+    template: jmlr_template.tex
     fig_caption: yes
@@
 When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable vizualization.
 
@@
 The remainder of this paper is organized as follows.
-The following Section, \@ref(sec:explanations), covers the background of the local explanation and the traditional visuals produced.
+Section \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced.
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
 Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
-Then Section \@ref(sec:casestudies) illustrates the application to supervised learning with categorical and quantitative output.
+Then Section \@ref(sec:casestudies) illustrates the application of supervised learning with categorical and quantitative output.
 Section \@ref(sec:cheemdiscussion) concludes with the insights gained and possible directions to explore in the future.
@@
 <!-- Taxonomy of local explanations -->
-A comprehensive summary of the taxonomy and literature of explanation techniques is provided in Figure 6 of @arrieta_explainable_2020.
+A comprehensive summary of the taxonomy and literature on explanation techniques is provided in Figure 6 by @arrieta_explainable_2020.
 It includes a large number of model-specific explanations such as deepLIFT [@shrikumar_not_2016; @shrikumar_learning_2017], a popular recursive method for estimating importance in neural networks.
@@
 For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
+In text analysis, word-level contextual sentiment analysis is used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
 In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction [@ribeiro_why_2016].
@@
 The magnitude of the contributions depends on the sequence in which they appear.
-Panel (b) shows the differences of the player's median values of over 25 such sequences.<!-- Segue --> In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
+Panel (b) shows the differences in the player's median values of over 25 such sequences.<!-- Segue --> In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
 In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
@@
 <!-- Tours intro -->
-A _tour_ enables viewing of high-dimensional data by animating many linear projections with small incremental changes.
+A _tour_ enables the  viewing of high-dimensional data by animating many linear projections with small incremental changes.
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
@@
 The global view (Figure \@ref(fig:regressioncase)a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation aid in understanding the structure identified in these spaces.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation to aid in understanding the structure identified in these spaces.
 
@@
 
-The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on [CRAN](https://CRAN.R-project.org/package=cheem).
+The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
 Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002] and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
@@
 (ref:casepenguins-cap) Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
 The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
@@
 
-There is more separation visible in the attribution space than the data space, as would be expected.
+There is more separation visible in the attribution space than in  the data space, as would be expected.
 The predicted vs observed plot reveals a handful of misclassified instances.
@@
 Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two features, mainly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard instance to account for as decision trees can only partition only vertical and horizontal lines.
+From this view, it makes sense that it is a hard instance to account for, as decision trees can only partition only vertical and horizontal lines.
 
@@
 Dark chocolate is considered healthier than milk.
-The data were collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
+Students collected the data during the Iowa State University class STAT503 from nutritional information on the manufacturer's websites and were normalized to 100g equivalents.
 The data is available in the __cheem__ package.
@@
 From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
-So, the contribution of Sugar is varied---reducing it corresponds primarily with increasing Fiber.
+So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
 When Sugar is zero, Fiber contributes strongly towards the left.
@@
 Players with high values in defensive skills are now the rightmost points.
-In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions at these levels.
+In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions were at these levels.
 
@@
 As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
-This large of importance in the Living Area is relatively uncommon.
+This large importance in the Living Area is relatively uncommon.
 Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.

### 2022-03-30 _paper.tex
Checklist cleanup & grammerly

@@
 \author{%
-\name Nicholas Spyrison \email \href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
+\name Nicholas Spyrison \email \href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
 %
-\name Dianne Cook \email \href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
+\name Dianne Cook \email \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
 %
-\name Kimbal Marriott \email \href{mailto:Kim.Marriott@monash.edu}{\nolinkurl{Kim.Marriott@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-9813-0377 \AND
+\name Kimbal Marriott \email \href{mailto:kim.marriott@monash.edu}{\nolinkurl{kim.marriott@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-9813-0377 \AND
 }
@@
 \begin{keywords}
-explainable artificial intelligence, nonlinear model interpretability, local explanations, grand tour, radial tour
+explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour
 \end{keywords}
@@
 When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable vizualization.
 
 The remainder of this paper is organized as follows.
-The following Section, \ref{sec:explanations}, covers the background of the local explanation and the traditional visuals produced.
+Section \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
 Section \ref{sec:tour} explains the animations of continuous linear projections.
 Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
-Then Section \ref{sec:casestudies} illustrates the application to supervised learning with categorical and quantitative output.
+Then Section \ref{sec:casestudies} illustrates the application of supervised learning with categorical and quantitative output.
 Section \ref{sec:cheemdiscussion} concludes with the insights gained and possible directions to explore in the future.
@@
 
-A comprehensive summary of the taxonomy and literature of explanation techniques is provided in Figure 6 of Arrieta et al.
+A comprehensive summary of the taxonomy and literature on explanation techniques is provided in Figure 6 by Arrieta et al.
 (2020).
@@
 For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis can be used to highlight the sentiment and magnitude of influential words (Vanni et al.
+In text analysis, word-level contextual sentiment analysis is used to highlight the sentiment and magnitude of influential words (Vanni et al.
 2018).
@@
 The magnitude of the contributions depends on the sequence in which they appear.
-Panel (b) shows the differences of the player's median values of over 25 such sequences.
+Panel (b) shows the differences in the player's median values of over 25 such sequences.
 In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
@@
 
-A \emph{tour} enables viewing of high-dimensional data by animating many linear projections with small incremental changes.
+A \emph{tour} enables the viewing of high-dimensional data by animating many linear projections with small incremental changes.
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
@@
 The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation aid in understanding the structure identified in these spaces.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation to aid in understanding the structure identified in these spaces.
 
@@
 
-The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on \href{https://CRAN.R-project.org/package=cheem}{CRAN}.
+The above-described method and application are implemented as an open-source \textbf{R} package, \textbf{cheem} available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
 Preprocessing was facilitated with models created via \textbf{randomForest} (Liaw and Wiener 2002) and explanations calculated with \textbf{treeshap} (Kominsarczyk et al.
@@
 \caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a), and the dashed vertical line in (b).
+The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
 The radial view shows varying the contribution of \texttt{fl} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
@@
 
-There is more separation visible in the attribution space than the data space, as would be expected.
+There is more separation visible in the attribution space than in the data space, as would be expected.
 The predicted vs observed plot reveals a handful of misclassified instances.
@@
 Dark chocolate is considered healthier than milk.
-The data were collected by students during the Iowa State University class STAT503 from nutritional information from the manufacturer's website and normalized to 100g equivalents.
+Students collected the data during the Iowa State University class STAT503 from nutritional information on the manufacturer's websites and were normalized to 100g equivalents.
 The data is available in the \textbf{cheem} package.

### 2022-03-30 template_rjh/author_checklist.rmd
Checklist cleanup & grammerly

@@
 
+Transactions JMLR - TMLR submissions: https://openreview.net/group?id=TMLR, submission
+
 # Checklist of common JMLR formatting errors
@@
 
-- One LaTeX file: The latex for the paper should be in one file. Otherwise it takes too much additional time for the production staff to perform all of the checks before publication.
+- One LaTeX file: The latex for the paper should be in one file. Otherwise it takes too much additional time for the production staff to perform all of the checks before publication. NS
 - Make sure your abstract reads well. It will be used to advertise your paper. 
-- We recommend you use small caps for system names. (In LaTex that's \\sc)
+- We recommend you use small caps for system names. (In LaTex that's \\sc). NA?
 - Move file to have correct name, according to the JMLR naming conventions. See the authors' guide. _NEEDED_
 - Signed copyright form: Do not forget to provide this. _NEEDED_
-- Text should be in Computer Modern.
+- Text should be in Computer Modern. _Don't see where to set_
 - Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References. _Not the order of the template, may need an eye on this._
@@
 - Author contact addresses: These should be complete enough that mail sent to that address will reach them. NS
-- Editor listed: Please include the name of the Action Editor who handled your submission in the \\editor{} tag _NS--- todo_
+- Editor listed: Please include the name of the Action Editor who handled your submission in the \\editor{} tag _todo after inital submission_
 - Keyword: Please add keywords to your article using \\begin{keywords} ...
@@
 NA
-- Floats should also be placed as close as possible to the page at which they are first referenced. You should not, e.g., place all of the figures at the end of the paper.
+- Floats should also be placed as close as possible to the page at which they are first referenced.
+You should not, e.g., place all of the figures at the end of the paper.
+NS, pretty good, def not at end.
 Figures, tables should be centered. NS
 - Captions should be underneath figures and tables. NS
-- Captions NOT in italics, bold. NS `codeface` and \\url{}, but not egregious.
-- Tables and figures should not overfill margins. NA (manual check, reference urls run over)
-Multiline captions should have a hanging indent, that is, subsequent lines of the caption should be indented, aligning with the first word of the caption rather than the word "Figure".
+- Captions NOT in italics, bold. NS `codeface` and \\url{}, but not egregious formatting.
+- Tables and figures should not overfill margins. NA (manual check, reference urls run over, but correctly formatted)
+- Multiline captions should have a hanging indent, that is, subsequent lines of the caption should be indented, aligning with the first word of the caption rather than the word "Figure".
 Usually this problem is caused by one of the other packages that you have included.
@@
 - Equations should not run over the right hand margin. Appendices are especially prone to this. NA
-- Theorems, Lemmas, etc., should be labeled like Theorem 1, Theorem 2, Theorem 3, not as Theorem 2.1, Theorem 2.2, Theorem 2.3 NA
+- Theorems, Lemmas, etc., should be labeled like Theorem 1, Theorem 2, Theorem 3, not as Theorem 2.1, Theorem 2.2, Theorem 2.3. NA
 
@@
 However, watch out for words like Bayes and Markov.
-These should still be capitalized, but BibTeX often leaves uncapitalised incorrectly.
+These should still be capitalized, but BibTeX often leaves capitalized incorrectly.
     - NS: I think this is a bit unrealistic. They may push back about this one.
 - Capitalization of book titles, etc: The parts in italics (book names, journal names, conf names) should be title case.
-(First letter of each word capitalized except for little words like "and" and "or".) NS See above subitem comment/.
+(First letter of each word capitalized except for little words like "and" and "or".)
+    - NS: I think this is a bit unrealistic. They may push back about this one.
 - Please cite conference titles consistently.
@@
 Personally, I prefer the style "International Conference on Machine Learning (ICML)", leaving off the "Proceedings of the Twentieth..."
+    - NS: I think this is a bit unrealistic. They may push back about this one.

### 2022-04-11 _paper.rmd
-author checklists: acknoledgements placement, bib cleanup -grammarly edits.

@@
 keywords: "explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour"
-acknowledgment: "We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.\\newline The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models."
 bibliography: cheem_paper.bib
@@
 There are different reasons and emphases when considering to fit a model.
-@breiman_statistical_2001 and @shmueli_explain_2010, taxonomize models based on their purpose.
+@breiman_statistical_2001 and @shmueli_explain_2010 taxonomize models based on their purpose.
 _Explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more narrowly on the performance of some objective function.
@@
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network, to name a few [@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995, receptively].
+Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forests, support-vector machines, or neural networks, to name a few [@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995, receptively].
 
 <!-- nonlinear leading to interpretability issues -->
-Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
+Nonlinear models have many or complexly interacting terms, which cause an opaqueness to the interpretation of the variables.
 This difficulty in interpreting the terms in complex nonlinear models sometimes lead to them being referred to as black box models.
@@
 A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
+Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
 
@@
 <!--tours and models --> Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015], and using tree-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
+Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015] and using tree-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
 
@@
 We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest then explore its local explanation by changing feature contribution with the radial tour.
+Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in R package __cheem__.
@@
 When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable vizualization.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable visualization.
 
@@
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
-Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
+Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface, and how it facilitates analysis, data preprocessing, and package infrastructure.
 Then Section \@ref(sec:casestudies) illustrates the application of supervised learning with categorical and quantitative output.
@@
 For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis is used to highlight the sentiment and magnitude of influential words [@vanni_textual_2018].
+In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words [@vanni_textual_2018].
 In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction [@ribeiro_why_2016].
@@
 SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The SHAP explanation refers to @shapley_value_1953's method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
+The SHAP explanation refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
 @strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
@@
 Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015, help to understand the relative importance of each feature in the context of all other candidate features.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015 help to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing player's wages [2020 Euros] from their skill measurements.
+A random forest model is fit regressing players' wages [2020 Euros] from their skill measurements.
 The SHAP values are compared for a star offensive player (L.
@@
 A difference in the attribution of the feature importance across the two positions of the players can be expected.
-This would be interpreted as how a player's salary depends on this combination of skill sets.
+This would be interpreted as how a player's salary depends on this combination of skills.
 Panel (a) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where three sequences of features are presented, so the two instances can be more easily compared.
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
+One key feature of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
 There are various types of tours that are distinguished by how the paths are generated [@lee_state_2021; @cook_grand_2008].
@@
 
+>> CONTINUE HERE.
+
 <!-- Purpose -->
-The global view provides a context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
+The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
 Both of these spaces are of dimension $n\times p$, where $n$ is the number of instances and $p$ is the number of predictors.
@@
 <!-- Local explanations as par coords -->
-The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
+The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of features can be compared across all instances.
 These are depicted as vertical parallel coordinate plots [@ocagne_coordonnees_1885] on the basis biplot [@gabriel_biplot_1971].
-1D biplot display depicts the values of the current basis as bars.
+1D biplot displays the values of the current basis as bars.
 The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
@@
 For example, Figure \@ref(fig:classificationcase)e suggests that `bl` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, and which have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some instances and not important for other instances.
+For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some instances and not important for other instances.
 
@@
 The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The radial tour varies the contribution of the selected feature between 0-1.
+The radial tour varies the contribution of the selected feature between 0 and 1.
 This is viewed as an animation of the projections from many intermediate bases.
@@
 The radial tour is 1D and displays as density where color indicates class.
-An animation slider enables the user to vary the contribution of each feature to explore the sensitivity of the separation to that feature.
+An animation slider enables users to vary the contribution of features to explore the sensitivity of the separation to that feature.
 
@@
 Global view inputs, (a), set the PI, CI, and color statistic.
-Global view, (b) PC1 by PC2 approximations of the data space and attribution space.
+Global view, (b) PC1 by PC2 approximations of the data- and attribution-space.
 (c) prediction by observed y (visual of the confusion matrix for classification tasks).
@@
 
-To illustrate the cheem method, it is applied to modern data sets, two classification examples and then two of regression.
+To illustrate the cheem method it is applied to modern data sets, two classification examples and then two of regression.
 
@@
 That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this, the radial tour is used to vary the contribution of flipper length (`fl`).
+The radial tour is used to vary the contribution of flipper length (`fl`) to explore this.
 (In our exploration, this was the third feature explored.
@@
 When Sugar is zero, Fiber contributes strongly towards the left.
-In this particular view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
+In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
 This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
@@
 Star offensive player (L.
-Messi) is the PI and he is compared with a top defensive player (V.
+Messi) is the PI, and he is compared with a top defensive player (V.
 van Dijk).
-The attribution projection is shown at left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
+The attribution projection is shown on the left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
 Defense (`def`) is the chosen feature to vary.
@@
 A random forest model has fitted this price [USD] with the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
-Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
+Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
 
@@
 The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
-In the attribution projection, the PI has a higher sales price than the CI.
+The PI has
+a higher sales price than the CI in the attribution projection.
 Reducing the contribution of Lot Area brings these two prices in line.
-This suggests if the model did not consider Lot Area, then the two houses would be quite similar.
+This suggests that if the model did not consider Lot Area, then the two houses would be quite similar.
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
@@
 Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
-A global view shows the data space, explanation space, residual plot.
+A global view shows the data space, explanation space, and residual plot.
 The user can interactively select instances to compare, contrast, and study further.
@@
 In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area not being effectively used by the random forest model.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area is not being effectively used by the random forest model.
 
@@
 
+<!--checklist asks for: Acknowledgments, Appendices, then References-->
+## Acknowledgments {-}
+
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
+This research was supported by the Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+
 # References

### 2022-04-11 _paper.tex
-author checklists: acknoledgements placement, bib cleanup -grammarly edits.

@@
 There are different reasons and emphases when considering to fit a model.
-Breiman (2001b) and Shmueli (2010), taxonomize models based on their purpose.
+Breiman (2001b) and Shmueli (2010) taxonomize models based on their purpose.
 \emph{Explanatory} modeling is done for some inferential purpose, while \emph{predictive} modeling focuses more narrowly on the performance of some objective function.
@@
 In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forest, support-vector machine, or neural network, to name a few (Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995, receptively).
-
-Nonlinear models have many or complex interactive terms, which cause an opaqueness to the interpretation of the variables.
+Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forests, support-vector machines, or neural networks, to name a few (Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995, receptively).
+
+Nonlinear models have many or complexly interacting terms, which cause an opaqueness to the interpretation of the variables.
 This difficulty in interpreting the terms in complex nonlinear models sometimes lead to them being referred to as black box models.
@@
 A common approach is to use \emph{local explanations}, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to visualize them to comprehensively understand a model.
+Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
 
@@
 Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015), and using tree-based approaches as a projection pursuit index to generate a tour basis paths (Y.
+Specifically for exploring various statistical model fits and classification boundaries (Wickham, Cook, and Hofmann 2015) and using tree-based approaches as a projection pursuit index to generate a tour basis paths (Y.
 D.
@@
 We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest then explore its local explanation by changing feature contribution with the radial tour.
+Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in R package \textbf{cheem}.
@@
 Section \ref{sec:tour} explains the animations of continuous linear projections.
-Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, how it facilitates analysis, data preprocessing, and package infrastructure.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, and how it facilitates analysis, data preprocessing, and package infrastructure.
 Then Section \ref{sec:casestudies} illustrates the application of supervised learning with categorical and quantitative output.
@@
 For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
-In text analysis, word-level contextual sentiment analysis is used to highlight the sentiment and magnitude of influential words (Vanni et al.
+In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words (Vanni et al.
 2018).
@@
 SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
-The SHAP explanation refers to Shapley (1953)'s method to evaluate an individual's contribution to cooperative games by assessing this player's performance in the presence or absence of other players.
+The SHAP explanation refers to Shapley (1953)'s method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
 Strumbelj and Kononenko (2010) introduced SHAP for local explanations in ML models.
@@
 Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015), help to understand the relative importance of each feature in the context of all other candidate features.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015) help to understand the relative importance of each feature in the context of all other candidate features.
 
@@
 There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
-A random forest model is fit regressing player's wages {[}2020 Euros{]} from their skill measurements.
+A random forest model is fit regressing players' wages {[}2020 Euros{]} from their skill measurements.
 The SHAP values are compared for a star offensive player (L.
@@
 A difference in the attribution of the feature importance across the two positions of the players can be expected.
-This would be interpreted as how a player's salary depends on this combination of skill sets.
+This would be interpreted as how a player's salary depends on this combination of skills.
 Panel (a) is a modified breakdown plot (Gosiewska and Biecek 2019) where three sequences of features are presented, so the two instances can be more easily compared.
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One of the features of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
+One key feature of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
 There are various types of tours that are distinguished by how the paths are generated (S.
@@
 
-The global view provides a context of all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
+\begin{quote}
+\begin{quote}
+CONTINUE HERE.
+\end{quote}
+\end{quote}
+
+The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
 Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
@@
 
-The local explanations for all observations are normalized (squared sum of values adds to 1), and thus, the relative importance of features can be compared across all instances.
+The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of features can be compared across all instances.
 These are depicted as vertical parallel coordinate plots (Ocagne 1885) on the basis biplot (Gabriel 1971).
-1D biplot display depicts the values of the current basis as bars.
+1D biplot displays the values of the current basis as bars.
 The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
@@
 For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, and which have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some instances and not important for other instances.
+For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some instances and not important for other instances.
 
@@
 That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
-To explore this, the radial tour is used to vary the contribution of flipper length (\texttt{fl}).
+The radial tour is used to vary the contribution of flipper length (\texttt{fl}) to explore this.
 (In our exploration, this was the third feature explored.
@@
 Star offensive player (L.
-Messi) is the PI and he is compared with a top defensive player (V.
+Messi) is the PI, and he is compared with a top defensive player (V.
 van Dijk).
-The attribution projection is shown at left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
+The attribution projection is shown on the left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
 Defense (\texttt{def}) is the chosen feature to vary.
@@
 A random forest model has fitted this price {[}USD{]} with the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
-Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction are selected.
+Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
 
@@
 The SHAP and oscillation explanations could be added with the use of the \texttt{DALEX::explain()} and would be an excellent direction to extend the work (Biecek 2018; Biecek and Burzykowski 2021).
+
+\hypertarget{acknowledgments}{%
+\subsection*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{subsection}{Acknowledgments}
+
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
+This research was supported by the Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
 
@@
 \leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
-Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.
+Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \url{https://christophm.github.io/interpretable-ml-book/}.
 
@@
 
-\acks{We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
-This research was supported by Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.\newline The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.}
-
 % Appendices to go in include-after

### 2022-04-11 template_rjh/author_checklist.rmd
-author checklists: acknoledgements placement, bib cleanup -grammarly edits.

@@
 - Signed copyright form: Do not forget to provide this. _NEEDED_
-- Text should be in Computer Modern. _Don't see where to set_
-- Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References. _Not the order of the template, may need an eye on this._
+- Text should be in Computer Modern. _Don't see option in .sty, .tex_
+- Order of paper: At the end of your paper, the order of sections should be: Acknowledgments, Appendices, then References. _Not the order of the template; manually unnumbered section_
 - Acknowledgments should be a unnumbered section. You can use the \\acks{} command for this. NS
@@
 NS
-- Abstract: Do not delete the % at the end of \\begin{abstract}% as it introduces an extra indent. _NS---doesn't get added in .rmd, may need to watch submission_
+- Abstract: Do not delete the % at the end of \\begin{abstract}% as it introduces an extra indent. _NS---handled in .sty I think. don't worry what .tex shows_
 - Title should be split across two lines, only when it is too long for one line. NS
@@
 Here are examples of BAD citations:
-    - "See (Ginsberg, 1994) for an analysis..." [Uses citation as a noun]. NS, looks good
+    - "See (Ginsberg, 1994) for an analysis..." [Uses citation as a noun]. NS, looks good, _use citation as people rather than noun (the work)_
     - "Drummond (Drummond, 1994) showed that..." [Should be "Drummond (1994)"]. NS, looks good
@@
 Or you can use \citep[e.g.,][]{smith00}.
+NS, no `))` will, keep an eye out.
 

### 2022-04-21 _paper.rmd
grammarly round

@@
 Nonlinear models have many or complexly interacting terms, which cause an opaqueness to the interpretation of the variables.
-This difficulty in interpreting the terms in complex nonlinear models sometimes lead to them being referred to as black box models.
+This difficulty in interpreting the terms of complex nonlinear models sometimes lead to them being referred to as black box models.
 Despite the potentially better performance of nonlinear models, their use is not without controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
@@
 Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in R package __cheem__.
+The methods are implemented in the R package __cheem__.
 Example data sets are provided to illustrate usage for classification and regression tasks.
@@
 The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-Both of these spaces are of dimension $n\times p$, where $n$ is the number of instances and $p$ is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, each a vector of $p$ values.
+These spaces both have dimensionality $n \times p$, where $n$ is the number of instances and $p$ is the number of predictors.
+The attribution space corresponds to the local explanations for each instance, each a vector of $p$ predictor importance in the vicinity of the observation.
 
 <!-- Approximations of the spaces, PC1:2 -->
-Visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
+The visualization is composed of the first two principal components of the data (left) and the attribution (middle) spaces.
 These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
@@
 For example, Figure \@ref(fig:classificationcase)e suggests that `bl` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some instances and not important for other instances.
+For regression, one might generally observe which features have low values for all instances (not important).
+For example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some instances and not important for others.
 
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \@ref(fig:classificationcase)f) and as scatterplots for a quantitative response (Figure \@ref(fig:regressioncase)f).
-The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The PI and CI are indicated by vertical dashed and dotted lines.
 The radial tour varies the contribution of the selected feature between 0 and 1.
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), body mass (`bm`), for this illustration.
+After removing incomplete instances, there are 333 instances of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), and body mass (`bm`) for this illustration.
 A random forest model was fit with species as the response feature.
@@
 The feature contributions to the two different predictions can be quickly compared in the parallel coordinate plot.
-The PI differs with the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
+The PI differs from the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
 
@@
 Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
-When primarily Fiber is examined, instance 22 looks more like a milk chocolate.
+When a frame with contribution primarily of Fiber is examined instance 22 looks more like a milk chocolate.
 
@@
 Defense (`def`) is the chosen feature to vary.
-It starts off very low, and as its contribution increases (right plot) Messi's predicted wages decrease dramatically.
+It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
 The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
+The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
 (A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)

### 2022-04-21 _paper.tex
grammarly round

@@
 Nonlinear models have many or complexly interacting terms, which cause an opaqueness to the interpretation of the variables.
-This difficulty in interpreting the terms in complex nonlinear models sometimes lead to them being referred to as black box models.
+This difficulty in interpreting the terms of complex nonlinear models sometimes lead to them being referred to as black box models.
 Despite the potentially better performance of nonlinear models, their use is not without controversy (O'Neil 2016; Kodiyan 2019).
@@
 Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in R package \textbf{cheem}.
+The methods are implemented in the R package \textbf{cheem}.
 Example data sets are provided to illustrate usage for classification and regression tasks.
@@
 The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-Both of these spaces are of dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
+These spaces each are have dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
 The attribution space corresponds to the local explanations for each instance, each a vector of \(p\) values.
 
-Visualization is provided by the first two principal components of the data (left) and the attribution (middle) spaces.
+The visualization is composed of the first two principal components of the data (left) and the attribution (middle) spaces.
 These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
@@
 For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important), for example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some instances and not important for other instances.
+For regression, one might generally observe which features have low values for all instances (not important).
+For example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some instances and not important for others.
 
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
-The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The PI and CI are indicated by vertical dashed and dotted lines.
 The radial tour varies the contribution of the selected feature between 0 and 1.
@@
 The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), body mass (\texttt{bm}), for this illustration.
+After removing incomplete instances, there are 333 instances of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), and body mass (\texttt{bm}) for this illustration.
 A random forest model was fit with species as the response feature.
@@
 Defense (\texttt{def}) is the chosen feature to vary.
-It starts off very low, and as its contribution increases (right plot) Messi's predicted wages decrease dramatically.
+It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
 The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that indeed, Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
+The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
 (A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)}\label{fig:casefifa}

### 2022-04-22 _paper.rmd
thru the end of preprocessing.

@@
 <!-- A figure \@ref(fig:crest) reference (with @; \\@ref(fig:crest). -->
-<!-- (ref:baseplotcap) Separate caption created above the R chunk -->
+<!-- (ref:baseplotcap) Separate caption created above the __R__ chunk -->
 <!-- ```{r baseplot, echo=F, fig.cap = "(ref:baseplotcap)"} -->
@@
 Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in the R package __cheem__.
+The methods are implemented in the __R__ package __cheem__.
 Example data sets are provided to illustrate usage for classification and regression tasks.
@@
 The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-These spaces both have dimensionality $n \times p$, where $n$ is the number of instances and $p$ is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, each a vector of $p$ predictor importance in the vicinity of the observation.
+These spaces both have dimensionality $n \times p$, where $n$ is the number of instances and $p$ is the number of featuress.
+The attribution space corresponds to the local explanations for each instance; feature importance in the vicinity of the instance.
 
@@
 
-While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of instances to explore the local explanations.
+While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular instances to explore in detail.
 The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
@@
 The attribution projection approximates the feature importance for predicting this instance.
-This combination of features best explains the difference between the mean response and an instance’s predicted value.
+The combination of features best explains the difference between the mean response and an instance’s predicted value.
 It is not an indication of the local shape of the model surface.
@@
 Global view, (b) PC1 by PC2 approximations of the data- and attribution-space.
-(c) prediction by observed y (visual of the confusion matrix for classification tasks).
+(c) prediction by observed $y$ (visual of the confusion matrix for classification tasks).
 Points are colored by predicted class, and red circles indicate misclassified instances.
@@
 ```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression  and illustration of interactive features.
-Panel (a) PCA of the data and attributions spaces, (b) residual plot, predictions by observed values.
+Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
 Four selected points are highlighted in the PC spaces and tabularly displayed.
@@
 Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (middle and right)."}
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right)."}
 knitr::include_graphics("./figures/app_regression_interactions.PNG")
@@
 <!-- Switch to rmd bullets, consistent with thesis -->
-- __Data:__ predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
+- __Data:__ predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep normality assumptions of the model in mind. 
 - __Model:__ any model and compatible explanation could be explored with this method.
 Currently, random forest models are applied via the package __randomForest__ [@liaw_classification_2002], compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, number variables randomly sampled at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
+Modest hyperparameters are used, namely: 125 trees, the number of variables at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
 - __Local explanation:__ Tree SHAP is calculated for _each_ observation using the package __treeshap__ [@kominsarczyk_treeshap_2021].
-This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of variables.
+We opt to find the attribution of each instance in the training data and not fit to fit feature interactions.
 - __Cheem viewer:__ after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
@@
 PCA and statistics of the features and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
+These runtimes were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
 Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-(Alternatively, the package __fastshap__ [@greenwell_fastshap_2020] claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
+[Alternatively, the package __fastshap__ claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.]
 
@@
 
-The package can be installed from CRAN, and the shiny app can be run using the following __R__ code:
+The package can be installed from CRAN, and the application can be run using the following __R__ code:
 
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of __R__ and Python packages.
 This research was supported by the Australian Government Research Training Program (RTP) scholarships.

### 2022-04-22 _paper.tex
thru the end of preprocessing.

@@
 The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-These spaces each are have dimension \(n\times p\), where \(n\) is the number of instances and \(p\) is the number of predictors.
-The attribution space corresponds to the local explanations for each instance, each a vector of \(p\) values.
+These spaces both have dimensionality \(n \times p\), where \(n\) is the number of instances and \(p\) is the number of featuress.
+The attribution space corresponds to the local explanations for each instance; feature importance in the vicinity of the instance.
 
@@
 The attribution projection approximates the feature importance for predicting this instance.
-This combination of features best explains the difference between the mean response and an instance's predicted value.
+The combination of features best explains the difference between the mean response and an instance's predicted value.
 It is not an indication of the local shape of the model surface.
@@
 \caption{Overview of the cheem viewer for regression  and illustration of interactive features.
-Panel (a) PCA of the data and attributions spaces, (b) residual plot, predictions by observed values.
+Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
 Four selected points are highlighted in the PC spaces and tabularly displayed.
@@
 Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (middle and right).}\label{fig:regressioncase}
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}\label{fig:regressioncase}
 \end{figure}
@@
 Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-(Alternatively, the package \textbf{fastshap} (Greenwell 2020) claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting.)
+(Alternatively, the package \textbf{fastshap} claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, Greenwell 2020.)
 

### 2022-04-22 _paper.rmd
Polish and demphasize the figure 5, Penguins validation

@@
 - The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
-- documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+- Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
 
-Follow the examples provided with the package to compute the local explainers (using `?cheem_ls`).
-The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()`, to be uploaded.
+Follow the examples provided with the package to compute the local explaination (using `?cheem_ls`).
+The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()` to be uploaded.
 
@@
 The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
+The data is publicly available to substitute for the overly-used iris data and is quite similar in form.
 After removing incomplete instances, there are 333 instances of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), and body mass (`bm`) for this illustration.

### 2022-04-22 _paper.tex
Polish and demphasize the figure 5, Penguins validation

@@
 Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in the R package \textbf{cheem}.
+The methods are implemented in the \textbf{R} package \textbf{cheem}.
 Example data sets are provided to illustrate usage for classification and regression tasks.
@@
 \item
-  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant.
+  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep normality assumptions of the model in mind.
 \item
@@
 Currently, random forest models are applied via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, number variables randomly sampled at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
+Modest hyperparameters are used, namely: 125 trees, the number of variables at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
 \item
@@
 2021).
-This implementation aggregates exhaustively over all trees' attribution and does not fit interactions of variables.
+We opt to find the attribution of each instance in the training data and not fit to fit feature interactions.
 \item
@@
 PCA and statistics of the features and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized R session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
+These runtimes were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
@@
 
-The package can be installed from CRAN, and the shiny app can be run using the following \textbf{R} code:
+The package can be installed from CRAN, and the application can be run using the following \textbf{R} code:
 
@@
 \item
-  documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+  Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
 \end{itemize}
 
-Follow the examples provided with the package to compute the local explainers (using \texttt{?cheem\_ls}).
-The application expects the output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file with \texttt{saveRDS()}, to be uploaded.
+Follow the examples provided with the package to compute the local explaination (using \texttt{?cheem\_ls}).
+The application expects the output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file with \texttt{saveRDS()} to be uploaded.
 
@@
 The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data was publicly available to substitute for the overly-used iris data and is quite similar in form.
+The data is publicly available to substitute for the overly-used iris data and is quite similar in form.
 After removing incomplete instances, there are 333 instances of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), and body mass (\texttt{bm}) for this illustration.
@@
 
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of R and Python packages.
+We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of \textbf{R} and Python packages.
 This research was supported by the Australian Government Research Training Program (RTP) scholarships.

### 2022-04-22 figures/_make_figures_case_study.r

@@ -75,24 +75,25 @@ ggplot2::ggsave(
 
 
 ## Penguins orthogonal BlFl ----
-(.g <- ggplot(penguins_na.rm, aes(x = bill_length_mm,
-                                  y = flipper_length_mm,
-                                  colour = species)) +
-   geom_point(size=1) +
-   ## Prim obs
-   geom_point(data=penguins_na.rm[243,], shape=8, size=5, alpha=0.8) +
-   ## Comparison obs
-   #geom_point(data=penguins_na.rm[169,], shape=4, size=3, alpha=0.6) +
-   theme_bw() +
-   theme(aspect.ratio = 1) +
-   
-   scale_color_brewer(palette = "Dark2") +
-   labs(y = "Flipper length [mm]", x = "Bill length [mm]", color = "Observed species"))
+(
+  .g <- ggplot(penguins_na.rm, aes(x = bill_length_mm,
+                                   y = flipper_length_mm),
+               alpha = logistic_tform(nrow(penguins_na.rm))) +
+    geom_point(aes(colour = species), size = 1) +
+    ## Prim obs
+    geom_point(data=penguins_na.rm[243,], shape=8, size=5, alpha=0.8) +
+    ## Comparison obs
+    geom_point(data=penguins_na.rm[169,], shape=4, size=3, alpha=0.6) +
+    theme_bw() +
+    theme(aspect.ratio = 1) +
+    
+    scale_color_brewer(palette = "Dark2") +
+    labs(y = "Flipper length [mm]", x = "Bill length [mm]", color = "Observed species"))
 ## Save
 ggplot2::ggsave(
   "./figures/case_penguins_BlFl.png",
-  plot = .g, device = "png",
-  height = 4, units = "in")
+  plot = .g, device = "pdf",
+  height = 2.1, units = "in")
 
 ### Save .mp4, add GitHub urls to paper
 message("NOTE: Manually capturing view from app with Screen to GIF (.mp4)")

### 2022-04-25 _paper.rmd
Edits for FIFA thru end.

@@
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate, nutritionally more like milk, should not be considered a healthy alternative.
+A reason to do this is that a dark chocolate nutritionally more like milk should not be considered a healthy alternative.
 It is interesting to explore which nutritional features contribute most to misclassification.
@@
 Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the PI.
-The feature contributions to the two different predictions can be quickly compared in the parallel coordinate plot.
+The feature contribution of the two different predictions can be quickly compared in the parallel coordinate plot.
 The PI differs from the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
@@
 
-(ref:casechocolatesinverse-cap) Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+(ref:casechocolatesinverse-cap) Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
@@
 If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
-However, Increasing the contribution of a variable with low importance would rotate both features out of the frame.
+However, increasing the contribution of a variable with low importance would rotate both features out of the frame.
 
@@
 
-Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
+Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
 The remaining are 338 house sales.
-A random forest model has fitted this price [USD] with the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
-Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
+A random forest model was fit predicting the sale price [USD] from the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+Using  interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
 
-(ref:caseames-cap) Exploring an instance with a large residual as the PI from fitting sales price [USD] to other variables in the Ames housing 2018.
-The sale price of the PI was under-predicted.
-The local explanation indicates a sizable attribution to Lot Area (`LtA`).
-The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
-The PI has
-a higher sales price than the CI in the attribution projection.
+(ref:caseames-cap) The largest neighborhood of Ames house price data is fit with a random forest predicting the sales price [USD].
+Exploring an instance with a extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
+The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this feature.
+The PI has a higher predicted value than the CI in the attribution projection.
 Reducing the contribution of Lot Area brings these two prices in line.
-This suggests that if the model did not consider Lot Area, then the two houses would be quite similar.
+This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar.
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
@@
 There is a clear need to extend the interpretability of black box models.
-This chapter provides a technique that builds on local interpretations to explore the feature importance local to an instance.
-The local interpretations form an attribution projection from which feature contributions are varied using a radial tour.
+This chapter provides a technique that builds on local explanations to explore the feature importance local to an instance.
+The local explanations of an attribution projection from which feature contributions are varied using a radial tour.
 Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
@@
 The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
-
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
 __DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.

### 2022-04-25 _paper.tex
Edits for FIFA thru end.

@@
 
-\caption{Examining the local interpretation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+\caption{Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
@@
 
-Ames 2018, housing data was subset to North Ames (the neighborhood with the most house sales).
+Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
 The remaining are 338 house sales.
-A random forest model has fitted this price {[}USD{]} with the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
-Using interaction from the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
+A random forest model was fit predicting the sale price {[}USD{]} from the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+Using interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
 
@@
 
-\caption{Exploring an instance with a large residual as the PI from fitting sales price {[}USD{]} to other variables in the Ames housing 2018.
-The sale price of the PI was under-predicted.
-The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}).
-The CI has a similar predicted sales price and smaller residual and has minimal attribution to Lot Area.
-The PI has
-a higher sales price than the CI in the attribution projection.
+\caption{The largest neighborhood of Ames house price data is fit with a random forest predicting the sales price {[}USD{]}.
+Exploring an instance with a extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
+The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}), while the CI has minimal attribution to this feature.
+The PI has a higher predicted value than the CI in the attribution projection.
 Reducing the contribution of Lot Area brings these two prices in line.
-This suggests that if the model did not consider Lot Area, then the two houses would be quite similar.
+This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar.
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
@@
 Thanks to Jieyang Chong for helping proofread this article.
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-DALEX pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
-
 The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.

### 2022-04-25 _paper.rmd
few grammerly edits

@@
 Section \@ref(sec:tour) explains the animations of continuous linear projections.
-Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface, and how it facilitates analysis, data preprocessing, and package infrastructure.
+Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
 Then Section \@ref(sec:casestudies) illustrates the application of supervised learning with categorical and quantitative output.
@@
 The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-These spaces both have dimensionality $n \times p$, where $n$ is the number of instances and $p$ is the number of featuress.
+These spaces both have dimensionality $n \times p$, where $n$ is the number of instances and $p$ is the number of features.
 The attribution space corresponds to the local explanations for each instance; feature importance in the vicinity of the instance.
@@
 
-Follow the examples provided with the package to compute the local explaination (using `?cheem_ls`).
+Follow the examples provided with the package to compute the local explanation (using `?cheem_ls`).
 The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()` to be uploaded.
@@
 It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate nutritionally more like milk should not be considered a healthy alternative.
+A reason to do this is that a dark chocolate, nutritionally more like milk should not be considered a healthy alternative.
 It is interesting to explore which nutritional features contribute most to misclassification.
@@
 Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
-When a frame with contribution primarily of Fiber is examined instance 22 looks more like a milk chocolate.
+When a frame with contribution primarily from Fiber is examined instance 22 looks more like a milk chocolate.
 
@@
 The remaining are 338 house sales.
-A random forest model was fit predicting the sale price [USD] from the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+A random forest model was fit, predicting the sale price [USD] from the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
 Using  interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
@@
 (ref:caseames-cap) The largest neighborhood of Ames house price data is fit with a random forest predicting the sales price [USD].
-Exploring an instance with a extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
+Exploring an instance with an extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
 The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this feature.

### 2022-04-25 _paper.tex
few grammerly edits

@@
 Section \ref{sec:tour} explains the animations of continuous linear projections.
-Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface, and how it facilitates analysis, data preprocessing, and package infrastructure.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
 Then Section \ref{sec:casestudies} illustrates the application of supervised learning with categorical and quantitative output.
@@
 The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-These spaces both have dimensionality \(n \times p\), where \(n\) is the number of instances and \(p\) is the number of featuress.
+These spaces both have dimensionality \(n \times p\), where \(n\) is the number of instances and \(p\) is the number of features.
 The attribution space corresponds to the local explanations for each instance; feature importance in the vicinity of the instance.
@@
 
-Follow the examples provided with the package to compute the local explaination (using \texttt{?cheem\_ls}).
+Follow the examples provided with the package to compute the local explanation (using \texttt{?cheem\_ls}).
 The application expects the output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file with \texttt{saveRDS()} to be uploaded.
@@
 The remaining are 338 house sales.
-A random forest model was fit predicting the sale price {[}USD{]} from the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+A random forest model was fit, predicting the sale price {[}USD{]} from the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
 Using interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
@@
 \caption{The largest neighborhood of Ames house price data is fit with a random forest predicting the sales price {[}USD{]}.
-Exploring an instance with a extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
+Exploring an instance with an extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
 The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}), while the CI has minimal attribution to this feature.

### 2022-04-29 _paper.rmd
missed a flag.

@@
 ```{r include=FALSE, cache=FALSE}
-## Paper compilation dependancies
+## Compilation dependencies
 library("knitr")
@@
   cache      = FALSE,
-  cache.lazy = FALSE)
+  cache.lazy = FALSE
+)
 ```
@@
 ## Global View
-
->> CONTINUE HERE.
 

### 2022-05-10 _paper.rmd
Kim asked to be removed from cheem paper.

@@
   email: dicook@monash.edu
-- familyname: Marriott
-  othernames: Kimbal
-  address: Monash University\newline ORCiD 0000-0002-9813-0377
-  email: kim.marriott@monash.edu
+# - familyname: Marriott
+#   othernames: Kim
+#   address: Monash University\newline ORCiD 0000-0002-9813-0377
+#   email: kim.marriott@monash.edu
 year: 2022

### 2022-05-10 _paper.tex
missed a flag.
Kim asked to be removed from cheem paper.

@@
 \jmlrheading{}{2022}{}{MM/DD}{--}{}{%
-Nicholas Spyrison, Dianne Cook and Kimbal Marriott}
+Nicholas Spyrison and Dianne Cook}
 
@@
 
-\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison, Cook and Marriott}
+\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison and Cook}
 \firstpageno{1}
@@
 \name Dianne Cook \email \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
-%
-\name Kimbal Marriott \email \href{mailto:kim.marriott@monash.edu}{\nolinkurl{kim.marriott@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-9813-0377 \AND
 }
@@
 \subsection{Global View}\label{global-view}}
-
-\begin{quote}
-\begin{quote}
-CONTINUE HERE.
-\end{quote}
-\end{quote}
 

### 2022-05-11 _paper.rmd
tweaks to intro

@@
 <!-- Interpretability & data drift -->
-Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors).
+Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors) between training and test sets.
 Some nonlinear models are sensitive to this and do not extrapolate well outside the support of the training data [@espadoto_toward_2021].
@@
 <!-- Local explanations -->
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black box models [@adadi_peeking_2018; @arrieta_explainable_2020].
+Explainable Artificial Intelligence (XAI) is an emerging field of research that provides methods for the interpreting of black box models [@adadi_peeking_2018; @arrieta_explainable_2020].
 A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
 Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
+There are common approaches for visualising high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
 
 <!-- Data visualization tours -->
-In multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] is a sequence of linear data projections onto a lower-dimensional space.
+For multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] of linear data projections onto a lower-dimensional space, can help.
 Tours are viewed as an animation over minor changes to the projection basis.
@@
 
-<!-- Manual tours --> 
-There are various types of tours distinguished by the generation of projection bases.
-In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected feature.
-<!--tours and models --> Applying tours to models has been done in a couple of contexts.
+<!-- tours and models -->
+Applying tours to models has been done in a couple of contexts.
 Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015] and using tree-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
 
-<!-- Purposed approach -->
-We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
+<!-- Manual tours --> 
+There are various types of tours distinguished by method of  generating the sequence of projection bases.
+In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected feature.
+We propose a radial manual tour can be used to scrutinize a local explanation.
+Additional interactive elements in a graphical user interface should allow the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in the __R__ package __cheem__.
@@
 Section \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced.
-Section \@ref(sec:tour) explains the animations of continuous linear projections.
+Section \@ref(sec:tour) explains the tours and particularly the radial manual tour.
 Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
-Then Section \@ref(sec:casestudies) illustrates the application of supervised learning with categorical and quantitative output.
+Illustrations are provided in Section \@ref(sec:casestudies) for a range of supervised learning tasks with categorical and quantitative outputs.
 Section \@ref(sec:cheemdiscussion) concludes with the insights gained and possible directions to explore in the future.
@@
 
-```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks.
+```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks (categorical outputs).
 Global view inputs, (a), set the PI, CI, and color statistic.
@@
 
-```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression  and illustration of interactive features.
+```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression tasks (quantitative outputs) and illustration of interactive features.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.

### 2022-05-11 _paper.tex
tweaks to intro

@@
 
-Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors).
+Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors) between training and test sets.
 Some nonlinear models are sensitive to this and do not extrapolate well outside the support of the training data (Espadoto et al.
@@
 
-Explainable Artificial Intelligence (XAI) is an emerging field of research that tries to increase the interpretability of black box models (Adadi and Berrada 2018; Arrieta et al.
+Explainable Artificial Intelligence (XAI) is an emerging field of research that provides methods for the interpreting of black box models (Adadi and Berrada 2018; Arrieta et al.
 2020).
@@
 Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
-
-In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
+There are common approaches for visualising high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
+
+For multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
 Lee et al.
-2021) is a sequence of linear data projections onto a lower-dimensional space.
+2021) of linear data projections onto a lower-dimensional space, can help.
 Tours are viewed as an animation over minor changes to the projection basis.
@@
 
-There are various types of tours distinguished by the generation of projection bases.
-In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected feature.
 Applying tours to models has been done in a couple of contexts.
@@
 
-We use the radial manual tour to scrutinize a local explanation in our proposed approach.
-Additional interactivity allows the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
+There are various types of tours distinguished by method of generating the sequence of projection bases.
+In a \emph{manual} tour (Cook and Buja 1997; Spyrison and Cook 2020), the path is defined by changing the contribution of a selected feature.
+We propose a radial manual tour can be used to scrutinize a local explanation.
+Additional interactive elements in a graphical user interface should allow the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
 The methods are implemented in the \textbf{R} package \textbf{cheem}.
@@
 Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
+Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged''.
+These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 Section \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
-Section \ref{sec:tour} explains the animations of continuous linear projections.
+Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
 Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
-Then Section \ref{sec:casestudies} illustrates the application of supervised learning with categorical and quantitative output.
+Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative outputs.
 Section \ref{sec:cheemdiscussion} concludes with the insights gained and possible directions to explore in the future.
@@
 
-\caption{Overview of the cheem viewer for classification tasks.
+\caption{Overview of the cheem viewer for classification tasks (categorical outputs).
 Global view inputs, (a), set the PI, CI, and color statistic.
@@
 
-\caption{Overview of the cheem viewer for regression  and illustration of interactive features.
+\caption{Overview of the cheem viewer for regression tasks (quantitative outputs) and illustration of interactive features.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-adadi_peeking_2018}{}}%
 Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
 
-\leavevmode\hypertarget{ref-anderson_introduction_1995}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-anderson_introduction_1995}{}}%
 Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT press.
 
-\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
 Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
@@
 
-\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
 Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
 
-\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
 ---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
-\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
 Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
 
-\leavevmode\hypertarget{ref-boser_training_1992}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-boser_training_1992}{}}%
 Boser, Bernhard E., Isabelle M.
@@
 
-\leavevmode\hypertarget{ref-breiman_random_2001}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-breiman_random_2001}{}}%
 Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45 (1): 5--32.
 
-\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
 ---------. 2001b. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
 
-\leavevmode\hypertarget{ref-buja_grand_1986}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
 Buja, Andreas, and Daniel Asimov.
@@
 
-\leavevmode\hypertarget{ref-chen_xgboost_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-chen_xgboost_2021}{}}%
 Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al.
@@
 
-\leavevmode\hypertarget{ref-espadoto_toward_2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-espadoto_toward_2021}{}}%
 Espadoto, Mateus, Rafael M.
@@
 
-\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gabriel_biplot_1971}{}}%
 Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices with Application to Principal Component Analysis.''} \emph{Biometrika} 58 (3): 453--67.
 
-\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
 Gorman, Kristen B., Tony D.
@@
 
-\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
 Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
-\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
 Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
-\leavevmode\hypertarget{ref-greenwell_gbm_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-greenwell_gbm_2020}{}}%
 Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M. Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression} {Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
 
-\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
 Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
@@
 
-\leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-molnar_interpretable_2020}{}}%
 Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \url{https://christophm.github.io/interpretable-ml-book/}.
 
-\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-oneil_weapons_2016}{}}%
 O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
 
-\leavevmode\hypertarget{ref-ocagne_coordonnees_1885}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ocagne_coordonnees_1885}{}}%
 Ocagne, Maurice d'.
@@
 \emph{Coordonnées Parallèles Et Axiales.
-{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul Graphique déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
-
-\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
+{Méthode} de Transformation Géométrique Et Procédé Nouveau de Calcul Graphique Déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
 Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
@@
 
-\leavevmode\hypertarget{ref-roberts_state_2007}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-roberts_state_2007}{}}%
 Roberts, Jonathan C.
@@
 
-\leavevmode\hypertarget{ref-shapley_value_1953}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
 Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
 
-\leavevmode\hypertarget{ref-shi_lightgbm_2022}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shi_lightgbm_2022}{}}%
 Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas Finley, Taifeng Wang, et al.
@@
 
-\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
 Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
 
-\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
 Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
@@
 
-\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
 Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
-\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-unwin_ensemble_2018}{}}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
-\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
 Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
@@
 
-\leavevmode\hypertarget{ref-wright_ranger_2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-wright_ranger_2017}{}}%
 Wright, Marvin N., and Andreas Ziegler.

### 2022-05-11 _paper.rmd
ames housing figure caption

@@
 
-## Ames Housing 2018, Sales Price Regression
-
-Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales.
-A random forest model was fit, predicting the sale price [USD] from the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
-Using  interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
-
-(ref:caseames-cap) The largest neighborhood of Ames house price data is fit with a random forest predicting the sales price [USD].
-Exploring an instance with an extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
+(ref:caseames-cap) Exploring an instance with an extreme residual as the PI in relation to an instance with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
 The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this feature.
@@
 ```
+
+## Ames Housing 2018, Sales Price Regression
+
+Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
+The remaining are 338 house sales.
+A random forest model was fit, predicting the sale price [USD] from the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+Using  interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
 

### 2022-05-11 _paper.tex
ames housing figure caption

@@
 
-\hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{Ames Housing 2018, Sales Price Regression}\label{ames-housing-2018-sales-price-regression}}
-
-Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales.
-A random forest model was fit, predicting the sale price {[}USD{]} from the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
-Using interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
-
 \begin{figure}
@@
 
-\caption{The largest neighborhood of Ames house price data is fit with a random forest predicting the sales price {[}USD{]}.
-Exploring an instance with an extreme residual as the PI and an accurate prediction for a similar price is used for comparison.
+\caption{Exploring an instance with an extreme residual as the PI in relation to an instance with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
 The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}), while the CI has minimal attribution to this feature.
@@
 \end{figure}
+
+\hypertarget{ames-housing-2018-sales-price-regression}{%
+\subsection{Ames Housing 2018, Sales Price Regression}\label{ames-housing-2018-sales-price-regression}}
+
+Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
+The remaining are 338 house sales.
+A random forest model was fit, predicting the sale price {[}USD{]} from the property features: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+Using interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
 

### 2022-05-11 _paper.rmd
conclusions

@@
 There is a clear need to extend the interpretability of black box models.
-This chapter provides a technique that builds on local explanations to explore the feature importance local to an instance.
+This paper provides a technique that builds on local explanations to explore the feature importance local to an instance.
 The local explanations of an attribution projection from which feature contributions are varied using a radial tour.
@@
 This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
+Local explanations focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
+They can also form the basis of learning how the model has got it wrong, with when the instance is misclassified or has a large residual.
+
 In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.

### 2022-05-11 _paper.rmd
conclusions - more changes

@@
 This analysis is manually intensive and thus only feasible for investigating a few instances.
-The approach selects instances that the model has not done well to predict and compare it with an instance where it did fit well.
+The recommended approach is to investigate an instance where the model has not predicted accurately and compare it with an instance with similar predictor values where model fitted well.
 The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.

### 2022-05-11 _paper.rmd
outline of paper tweaked
euros in figure caption unnecessary

@@
 Illustrations are provided in Section \@ref(sec:casestudies) for a range of supervised learning tasks with categorical and quantitative outputs.
-Section \@ref(sec:cheemdiscussion) concludes with the insights gained and possible directions to explore in the future.
+Section \@ref(sec:cheemdiscussion) concludes with a summary of the insights gained.
 
@@
 
-(ref:casefifa-cap) Exploring the wages [euros] relative to skill measurements in the FIFA 2020 data.
+(ref:casefifa-cap) Exploring the wages relative to skill measurements in the FIFA 2020 data.
 Star offensive player (L.

### 2022-05-11 _paper.tex
outline of paper tweaked
euros in figure caption unnecessary

@@
 Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative outputs.
-Section \ref{sec:cheemdiscussion} concludes with the insights gained and possible directions to explore in the future.
+Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
 
@@
 
-\caption{Exploring the wages {[}euros{]} relative to skill measurements in the FIFA 2020 data.
+\caption{Exploring the wages relative to skill measurements in the FIFA 2020 data.
 Star offensive player (L.

### 2022-05-11 _paper.tex
extra comma removed
milk/dark classification title

@@
 
-\hypertarget{chocolates-milkdark-chocolate-classification}{%
-\subsection{Chocolates, Milk/Dark Chocolate Classification}\label{chocolates-milkdark-chocolate-classification}}
+\hypertarget{chocolates-milkdark-classification}{%
+\subsection{Chocolates, Milk/Dark Classification}\label{chocolates-milkdark-classification}}
 
@@
 
-Figure \ref{fig:casefifa}, tests the support of the local explanation.
+Figure \ref{fig:casefifa} tests the support of the local explanation.
 Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.

### 2022-05-11 _paper.rmd
extra comma removed
milk/dark classification title
ames housing citation

@@
 
-## Chocolates, Milk/Dark Chocolate Classification
+## Chocolates, Milk/Dark Classification
 
@@
 
-Figure \@ref(fig:casefifa), tests the support of the local explanation.
+Figure \@ref(fig:casefifa) tests the support of the local explanation.
 Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
@@
 
-Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
+Ames housing data 2018 [@de_cock_ames_2011; @prevek18_ames_2018] was subset to North Ames (the neighborhood with the most house sales).
 The remaining are 338 house sales.

### 2022-05-11 _paper.tex
Kim to acknowledgements and knit

@@
 Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020).
-Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged''.
-These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
+Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
 It ignores correlations of the features and imagines a case that was not observed.
@@
 
-Ames housing data 2018 was subset to North Ames (the neighborhood with the most house sales).
+Ames housing data 2018 (De Cock 2011; prevek18 2018) was subset to North Ames (the neighborhood with the most house sales).
 The remaining are 338 house sales.
@@
 
+Kim Marriott provided advice on many aspects of this work, especially on the explanations in the applications section.
 We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of \textbf{R} and Python packages.
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-adadi_peeking_2018}{}}%
+\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
 Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
 
-\leavevmode\vadjust pre{\hypertarget{ref-anderson_introduction_1995}{}}%
+\leavevmode\hypertarget{ref-anderson_introduction_1995}{}%
 Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
+\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
 Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio Gil-López, Daniel Molina, and Richard Benjamins.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
+\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
 Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
 
-\leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
+\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
 ---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
+\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
 Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-boser_training_1992}{}}%
+\leavevmode\hypertarget{ref-boser_training_1992}{}%
 Boser, Bernhard E., Isabelle M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-breiman_random_2001}{}}%
+\leavevmode\hypertarget{ref-breiman_random_2001}{}%
 Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45 (1): 5--32.
 
-\leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
+\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
 ---------. 2001b. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
 
-\leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
+\leavevmode\hypertarget{ref-buja_grand_1986}{}%
 Buja, Andreas, and Daniel Asimov.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chen_xgboost_2021}{}}%
+\leavevmode\hypertarget{ref-chen_xgboost_2021}{}%
 Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-diaz_addressing_2018}{}}%
+\leavevmode\hypertarget{ref-de_cock_ames_2011}{}%
+De Cock, Dean. 2011. {``Ames, {Iowa}: {Alternative} to the {Boston} Housing Data as an End of Semester Regression Project.''} \emph{Journal of Statistics Education} 19 (3).
+
+\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
 Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren Gergle.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-espadoto_toward_2021}{}}%
+\leavevmode\hypertarget{ref-espadoto_toward_2021}{}%
 Espadoto, Mateus, Rafael M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
+\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
 Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
+\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
 Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-greenwell_gbm_2020}{}}%
+\leavevmode\hypertarget{ref-greenwell_gbm_2020}{}%
 Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M. Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression} {Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
+\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
 Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-molnar_interpretable_2020}{}}%
+\leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
 Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu. com. \url{https://christophm.github.io/interpretable-ml-book/}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-oneil_weapons_2016}{}}%
+\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
 O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data Increases Inequality and Threatens Democracy}. Crown.
 
-\leavevmode\vadjust pre{\hypertarget{ref-ocagne_coordonnees_1885}{}}%
+\leavevmode\hypertarget{ref-ocagne_coordonnees_1885}{}%
 Ocagne, Maurice d'.
@@
 \emph{Coordonnées Parallèles Et Axiales.
-{Méthode} de Transformation Géométrique Et Procédé Nouveau de Calcul Graphique Déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
+{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul Graphique déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
+
+\leavevmode\hypertarget{ref-prevek18_ames_2018}{}%
+prevek18. 2018. {``Ames {Housing} {Dataset}.''} \emph{Kaggle.com}. \url{https://kaggle.com/prevek18/ames-housing-dataset}.
+
+\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
 Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-roberts_state_2007}{}}%
+\leavevmode\hypertarget{ref-roberts_state_2007}{}%
 Roberts, Jonathan C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
+\leavevmode\hypertarget{ref-shapley_value_1953}{}%
 Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-shi_lightgbm_2022}{}}%
+\leavevmode\hypertarget{ref-shi_lightgbm_2022}{}%
 Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas Finley, Taifeng Wang, et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
+\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
 Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
 
-\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
+\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
 Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
+\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
 Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
 
-\leavevmode\vadjust pre{\hypertarget{ref-unwin_ensemble_2018}{}}%
+\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
 Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
+\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
 Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wright_ranger_2017}{}}%
+\leavevmode\hypertarget{ref-wright_ranger_2017}{}%
 Wright, Marvin N., and Andreas Ziegler.

### 2022-05-11 _paper.rmd
Kim to acknowledgements and knit
JMLR and arXiv submission,

@@
 year: 2022
-date_submitted: "MM/DD"
+date_submitted: "05/11"
 abstract: "The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
@@
 
+Kim Marriott provided advice on many aspects of this work, especially on the explanations in the applications section.
 We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of __R__ and Python packages.

### 2023-03-14 _paper.rmd
-Attempt to add tag for release. -Added Przemek.

@@
   address: Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212
-  email: nicholas.spyrison@monash.edu
+  email: spyrison@gmail.com
   correspondingauthor: true
@@
   email: dicook@monash.edu
+- familyname: Biecek
+  othernames: Przemyslaw
+  address:  Warsaw University of Technology and University of Warsaw\newline ORCiD 0000-0001-8423-1823
+  email: przemyslaw.biecek@gmail.com
 # - familyname: Marriott

### 2023-03-14 _paper.tex
JMLR and arXiv submission,
-Attempt to add tag for release. -Added Przemek.

--- 2022-05-11 _paper.tex
@@
 
-\jmlrheading{}{2022}{}{MM/DD}{--}{}{%
-Nicholas Spyrison and Dianne Cook}
+\jmlrheading{}{2022}{}{05/11}{--}{}{%
+Nicholas Spyrison, Dianne Cook and Przemyslaw Biecek}
 
@@
 
-\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison and Cook}
+\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison, Cook and Biecek}
 \firstpageno{1}
@@
 \author{%
-\name Nicholas Spyrison \email \href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
+\name Nicholas Spyrison \email \href{mailto:spyrison@gmail.com}{\nolinkurl{spyrison@gmail.com}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
 %
 \name Dianne Cook \email \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
+%
+\name Przemyslaw Biecek \email \href{mailto:przemyslaw.biecek@gmail.com}{\nolinkurl{przemyslaw.biecek@gmail.com}} \\\addr Warsaw University of Technology and University of Warsaw\newline ORCiD 0000-0001-8423-1823 \AND
 }
@@
 \url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de_cock_ames_2011}{}}%
+De Cock, Dean. 2011. {``Ames, {Iowa}: {Alternative} to the {Boston} Housing Data as an End of Semester Regression Project.''} \emph{Journal of Statistics Education} 19 (3).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-prevek18_ames_2018}{}}%
+prevek18. 2018. {``Ames {Housing} {Dataset}.''} \emph{Kaggle.com}. \url{https://kaggle.com/prevek18/ames-housing-dataset}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%

### 2023-03-16 motivating_letter.Rmd
name change for files, and formatted for JCGS plus motivating letter

---
title: ''
output: pdf_document
date: "2023-03-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editors,

Please consider our paper titled "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections" for publication in the Journal of Computational and Graphical Statistics. We'd like to acknowledge that it was previously submitted to the Journal of Machine Learning Research. The primary reason was that the editor felt it had too much visualisation to be a good fit for the journal. However, some useful feedback was provided, and below we have described how we have improved the paper based on their reviews.

We feel that JCGS is a good fit for this work because it describes useful new visual approaches for understanding model fits based on the relatively new local explanations diagnostics. This work can also be used to assess the reliability of the information reported by these diagnostics.

Here is an explanation of how the paper was improved based on the review from JMLR:

*(a) The writing of the paper is informal, lacks clarity and depth in many parts that are critical to understanding the method proposed, and requires significant improvement. The method of radial tour, which (apparently?) seems to be the most important contribution of the paper, is practically not explained at all. Up to Section 4, when the description of the interactive visualization tool starts, all we get for the explanation of the radial tour is one sentence, which is drastically not enough. This is observable throughout the text as a whole, and it is very hard to point exactly where and when it happens. In general, the paper's writing is informal, lacking the depth necessary to make it reproducible and auditable. Without clear, detailed explanations of the methods used, it becomes hard to accept it in a deeply technical journal such as JMLR.*

*(b) The paper does not seem to fit the scope of the journal. Even though the idea behind the radial tour seems interesting (although it is hard to know because the authors have skipped explaining it in details), it seems that one of the most important contributions of the paper is actually an interactive visualization tool. As such, I believe that JMLR is not the right venue for this paper. The authors should've chosen a visualization venue for this, where they'll be able to get a much more interested audience and much more relevant feedback too. I'd recommend IEEE VIS, IEEE TVCG, EuroVis, Computer Graphics Forum, Computers & Graphics, Information Visualization journal, or the PacificVis conference.*

* *Abstract: While it is not wrong, not all XAI techniques use local explanations; some definitely do. The wording here could be a bit more accurate.*

* *The introduction seems to end a bit abruptly. Maybe a couple of sentences about the results and outcomes of the research?*

* *Page 3: "Consider a highly nonlinear model." I feel like this (and the reasoning that follows) is a bit of an informal statement. Although intuitively I agree with the authors, is there any objective (or at least semi-formal) way to explain this problem? What is highly non-linear (vs. just linear)? Why, exactly, is it that it is so hard to interpret nonlinear models? That would improve the argumentation here quite a bit.*

* *"The attribution of feature importance depends on the sequence of the included features." This sentence feels out of place.*

* *"modified breakdown plots" -- Since this is taking an important place in the paper, with a figure attached to it, I think it needs to be better explained. What are exactly modified breakdown plots?*

* *"The magnitude of the contributions depends on the sequence in which they appear." -- This is unclear. The figure has no label for the horizontal axes, which makes it very hard to interpret. And, in case, the sequence alters the magnitude of the contributions, then it should be explained how the sequence is determined. Right now, this explanation does not help to make the visualization or the method presented in Figure 1 easier to understand.*

* *"The attribution space corresponds to the local explanations for each instance; feature importance in the
vicinity of the instance." -- I think this needs to be better explained. The concept of the attribution space remains obscure to me. It seems to be a very interesting concept and very important to the paper.*

* *"The projection attribution of the primary instance (PI) is examined and typically viewed with an optional
comparison instance (CI)." -- The concept of PI and CI remains confusing to me. Needs a clearer explanation.*

* *Section 4.2 as a whole -- This explanation is overly confusing and informal/subjective. It needs to be significantly revised to become more fluid and clear.*

* *I don't believe that sections 4.7 and 4.8 are needed.*

Regards,

Nick Spyrison

### 2023-03-16 paper.Rmd
name change for files, and formatted for JCGS plus motivating letter

--- 2022-05-11 _paper.rmd
@@
 title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
-shorttitle: "Cheem, Exploring Local Explanations"
+type: ARTICLE TEMPLATE
 author:
-- familyname: Spyrison
-  othernames: Nicholas
-  address: Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212
-  email: nicholas.spyrison@monash.edu
-  correspondingauthor: true
-- familyname: Cook
-  othernames: Dianne
-  address: Monash University\newline ORCiD 0000-0002-3813-7155
+  - name: Nicholas Spyrison
+    affiliation: a
+    orcid: 0000-0002-8417-0212
+    email: spyrison@gmail.com
+  - name: Dianne Cook
+    affiliation: a
+    orcid: 0000-0002-3813-7155
     email: dicook@monash.edu
-# - familyname: Marriott
-#   othernames: Kim
-#   address: Monash University\newline ORCiD 0000-0002-9813-0377
-#   email: kim.marriott@monash.edu
-year: 2022
-date_submitted: "05/11"
-abstract: "The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
+  - name: Przemyslaw Biecek
+    affiliation: b
+    orcid: 0000-0001-8423-1823
+    email: przemyslaw.biecek@gmail.com
+affiliation:
+  - num: a
+    address: |
+      Monash University, Melbourne, Australia
+  - num: b
+    address: |
+      Warsaw University of Technology and University, Warsaw, Poland
+abstract: |
+  The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
 This trade-off has led to the emergence of eXplainable AI (XAI).
@@
 Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
-The methods are implemented in the __R__ package __cheem__, available on CRAN."
-keywords: "explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour"
-bibliography: cheem_paper.bib
-classoption: 11pt,twoside
-output:
-  bookdown::pdf_document2:
-    template: jmlr_template.tex
-    fig_caption: yes
-    fig_height: 5
-    fig_width: 8
-    keep_tex: yes
-    number_sections: yes
-    #citation_package: natbib
-    latex_engine: pdflatex
+The methods are implemented in the __R__ package __cheem__, available on CRAN.
+keywords: | 
+  explainable artificial intelligence; nonlinear model interpretability; visual analytics; local explanations; grand tour; radial tour
+bibliography: paper.bib
+header-includes: |
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \def\tightlist{}
+output: rticles::tf_article
 ---
@@
   cache      = FALSE,
-  cache.lazy = FALSE
+  cache.lazy = FALSE,
+  fig.height = 5,
+  fig.width = 8,
+  out.width = "100%"
 )

### 2023-03-26 paper.Rmd
suggestion to refer to shapley's lens

@@
 The implementation allows for pre-computation and interactive re-calculation to focus on a different feature.
+
+# Shapley's Lenses {#sec:shapleylenses}
+
+Shapley values are a useful technique for understanding the nature of the relationship between the input variables and the target variable.
+The data exploration process is navigated by a trained predictive model.
+
+Thus, local attributions can be viewed as a new representation of data (commonly referred as data embeddings by machine learning community) with certain desired properties, e.g., the unit is standardized between variables because it reflects the relation with target variable, so it is expressed on the scale of the target variable.
+
+But different models can produce different local attributions.
+This phenomenon, known as the multiplicity of good models or as the Rashomon set of equally good models, means that models that fit the data equally well can describe the relationships present in the data in different ways.
+
+The response to the challenge in the characterization of Rashomon sets may be Shapley's Lenses - the opportunity to compare these models from the perspective of exploring the local attributions generated by the Shapley values for different models.
 

### 2023-04-03 paper.Rmd
uodated introduction

@@
 This trade-off has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction with _local explanations_, a point estimate of the linear feature importance in the vicinity of one instance.
-These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
+XAI attempts to shed light on how models use predictors to arrive at a prediction with _local explanations_, a point estimate of the linear variable importance in the vicinity of one observation.
+These can be considered linear projections and can be further explored to understand better the interactions between variables used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
@@
 <!-- Introduce explanatory vs predictive modeling -->
-There are different reasons and emphases when considering to fit a model.
-@breiman_statistical_2001 and @shmueli_explain_2010 taxonomize models based on their purpose.
-_Explanatory_ modeling is done for some inferential purpose, while _predictive_ modeling focuses more narrowly on the performance of some objective function.
-The intended use has important implications for model selection and development.
-In explanatory modeling, interpretability is vital for drawing inferential conclusions.
-Nonlinear models range from additive models with at least one polynomial term to more complex machine learning models such as random forests, support-vector machines, or neural networks, to name a few [@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995, receptively].
+There are different reasons and purposes for fitting a model.
+According to the taxonomies of @breiman_statistical_2001 and @shmueli_explain_2010, it can be useful to group models into two types: explanatory and predictive.
+Explanatory modeling is used for inferential purposes, while predictive modeling focuses solely on the performance of an objective function.
+The intended use of the model has important implications for its selection and development.
+Interpretability is critical in explanatory modeling to draw meaningful inferential conclusions, such as which variables most contribute to a prediction or whether some observations are less well fit.
+Interpretability becomes more difficult when the model is nonlinear.
+Nonlinear models occur in statistical models with polynomial or interaction terms between quantitative predictors, and almost all computational models such as random forests, support-vector machines, or neural networks [e.g.
+@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995].
 
 <!-- nonlinear leading to interpretability issues -->
-Nonlinear models have many or complexly interacting terms, which cause an opaqueness to the interpretation of the variables.
-This difficulty in interpreting the terms of complex nonlinear models sometimes lead to them being referred to as black box models.
-Despite the potentially better performance of nonlinear models, their use is not without controversy [@oneil_weapons_2016; @kodiyan_overview_2019].
-And the loss of interpretation presents a challenge.
+In linear models interpretation of the importance of variables is relatively straight forward, one adjusts for the covariance of multiple variables when examining the relationship with the response.
+The interpretation is valid for the full domain of the predictors.
+In nonlinear models one needs to consider the model in small neighbourhoods of the domain to make any assessment of variable importance.
+Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues decribed in @stahl-ethics.
+Understanding how nonlinear models behave when usage extrapolates outside the domain of predictors, either in subspaces where few samples were provided in the training set, or extending outside the domain.
+It is especially important because nonlinear models can vary wildly and predictions can be dramatically wrong in these areas.
 
-<!-- Interpretability & baises -->
+<!-- Interpretability & baises 
 Interpretability is vital for exploring and protecting against potential biases in any model, e.g., sex\-\-\-@dastin_amazon_2018; @duffy_apple_2019, race\-\-\-@larson_how_2016, and age\-\-\-@diaz_addressing_2018.
-For instance, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
+For observation, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
 This bias is then built into the model.
-Variable-level (feature-level) interpretability of models is essential in evaluating and addressing such biases.
+Variable-level interpretability of models is essential in evaluating and addressing such biases.
+-->
 
-<!-- Interpretability & data drift -->
+<!-- Interpretability & data drift 
 Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors) between training and test sets.
@@
 Maintaining variable interpretability is also essential to address issues arising from data drift.
+-->
 
 <!-- Local explanations -->
-Explainable Artificial Intelligence (XAI) is an emerging field of research that provides methods for the interpreting of black box models [@adadi_peeking_2018; @arrieta_explainable_2020].
-A common approach is to use _local explanations_, which attempt to approximate linear feature importance at the location of each instance (observation) or the predictions at a specific point in the data domain.
+Explainable Artificial Intelligence (XAI) is an emerging field of research focused on methods for the interpreting of models [@adadi_peeking_2018; @arrieta_explainable_2020].
+A class of techniques called _local explanations_, provide methods to approximate linear variable importance at the location of each observation or the predictions at a specific point in the data domain.
 Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
@@
 <!-- Data visualization tours -->
-For multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] of linear data projections onto a lower-dimensional space, can help.
-Tours are viewed as an animation over minor changes to the projection basis.
-Structure in a projection can then be explored visually to see which features contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information about the structure and features of the object.
-
+For multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] of linear data projections onto a lower-dimensional space, could be an element of XAI, complementing local explanations.
+<!--Tours are viewed as an animation over minor changes to the projection basis.
+Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
+The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information about the structure and variables of the object.-->
 <!-- tours and models -->
-Applying tours to models has been done in a couple of contexts.
-Specifically for exploring various statistical model fits and classification boundaries [@wickham_visualizing_2015] and using tree-based approaches as a projection pursuit index to generate a tour basis paths [@lee_pptree_2013; @da_silva_projection_2021].
-
+Applying tours to model interpretation is recommended in @wickham_visualizing_2015 primarily to examine the fitted model in the space of the data.
+@cook_interactive_2007 describe the use of tours for exploring classification boundaries and model diagnostics [@Caragea2008; @lee_pptree_2013; @da_silva_projection_2021].
 <!-- Manual tours --> 
-There are various types of tours distinguished by method of  generating the sequence of projection bases.
-In a _manual_ tour [@cook_manual_1997; @spyrison_spinifex_2020], the path is defined by changing the contribution of a selected feature.
-We propose a radial manual tour can be used to scrutinize a local explanation.
-Additional interactive elements in a graphical user interface should allow the user to identify an instance of interest and then explore its local explanation by changing feature contribution with the radial tour.
-The methods are implemented in the __R__ package __cheem__.
-Example data sets are provided to illustrate usage for classification and regression tasks.
-
+There are various types of tours.
+In a _manual_ or radial tour [@cook_manual_1997; @spyrison_spinifex_2020], the path of linear projections is defined by changing the contribution of a selected variable.
+We propose to use this to scrutinize a local explanation.
+<!-- Additional interactive elements in a graphical user interface should allow the user to identify an observation of interest and then explore its local explanation by changing variable contribution with the radial tour.
+-->
 <!-- What-if & counterfactual analysis -->
-Using a radial tour can be compared with counterfactual, what-if analysis, such as _ceteris paribus_ profiles [@biecek_ceterisparibus_2020].
-Ceteris paribus, is Latin for "other things held constant" or "all else unchanged".
-These profiles show how an instance's prediction would change from a marginal change in one explanatory feature, given that other features are held constant.
-It ignores correlations of the features and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object.
+This approach could be considered to be a counterfactual, what-if analysis, such as _ceteris paribus_ ("other things held constant") profiles [@biecek_ceterisparibus_2020].
+<!-- These profiles show how an observation's prediction would change from a marginal change in one explanatory variable, given that other variables are held constant.
+It ignores correlations of the variables and imagines a case that was not observed.
+In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
 A constraint in our approach is that the basis must remain orthonormal.
-When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable visualization.
+When the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
+This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable visualization.-->
 
@@
 The remainder of this paper is organized as follows.
-Section \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced.
-Section \@ref(sec:tour) explains the tours and particularly the radial manual tour.
-Section \@ref(sec:cheemviewer) discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
-Illustrations are provided in Section \@ref(sec:casestudies) for a range of supervised learning tasks with categorical and quantitative outputs.
-Section \@ref(sec:cheemdiscussion) concludes with a summary of the insights gained.
+Section \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
+Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
+Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative outputs.
+Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
+The methods are implemented in the __R__ package __cheem__.
+Example data sets are provided to illustrate usage for classification and regression tasks.
 
@@
 Consider a highly nonlinear model.
-It can be hard to determine whether small changes in a feature's value will make a class prediction change group or identify which features contribute to an extreme residual.
-Local explanations shed light on these situations by approximating linear feature importance in the vicinity of a single instance.
+It can be hard to determine whether small changes in a variable's value will make a class prediction change group or identify which variables contribute to an extreme residual.
+Local explanations shed light on these situations by approximating linear variable importance in the vicinity of a single observation.
 
@@
 <!-- Uses of local explanations -->
-These instance-level explanations are used in various ways depending on the data.
+These observation-level explanations are used in various ways depending on the data.
 In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels [@simonyan_deep_2014].
-For instance, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+For observation, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
 In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words [@vanni_textual_2018].
-In the case of numeric regression, they are used to explain feature additive contributions from the model intercept to the instance's prediction [@ribeiro_why_2016].
+In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction [@ribeiro_why_2016].
 
 <!-- SHAP -->
-SHaply Additive exPlanations (SHAP) quantifies the feature contributions of one instance by examining the effect of other features on the predictions.
+SHaply Additive exPlanations (SHAP) quantifies the variable contributions of one observation by examining the effect of other variables on the predictions.
 The SHAP explanation refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
 @strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
-The attribution of feature importance depends on the sequence of the included features.
-The SHAP values are the mean contributions over different feature sequences.
-The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a feature by predicting the response for a range of values on this feature after fixing the value of all other features to their mean.
-Though partial dependence plots are a global approximation of the feature importance, while SHAP is specific to one instance.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015 help to understand the relative importance of each feature in the context of all other candidate features.
+The attribution of variable importance depends on the sequence of the included variables.
+The SHAP values are the mean contributions over different variable sequences.
+The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
+Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015 help to understand the relative importance of each variable in the context of all other candidate variables.
 
@@
 van Dijk, respectively.
-Panel (a) shows breakdown plots of three sequences of the features.
+Panel (a) shows breakdown plots of three sequences of the variables.
 The sequence of the variables impacts the magnitude of their attribution.
-Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player.
+Panel (b) shows the distribution of attribution for each variable across 25 sequences of predictors, with the mean displayed as a dot for each player.
 Reaction skills are important for both players.
@@
 Consider soccer data from the FIFA 2020 season [@leone_fifa_2020].
-There are 5000 instances of 9 skill measures (after aggregating highly correlated features).
+There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
 A random forest model is fit regressing players' wages [2020 Euros] from their skill measurements.
@@
 The results are displayed in Figure \@ref(fig:shapdistrbd).
-A difference in the attribution of the feature importance across the two positions of the players can be expected.
+A difference in the attribution of the variable importance across the two positions of the players can be expected.
 This would be interpreted as how a player's salary depends on this combination of skills.
-Panel (a) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where three sequences of features are presented, so the two instances can be more easily compared.
+Panel (a) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where three sequences of variables are presented, so the two observations can be more easily compared.
 The magnitude of the contributions depends on the sequence in which they appear.
-Panel (b) shows the differences in the player's median values of over 25 such sequences.<!-- Segue --> In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their instances.
-In this instance, two players with different positions receive different profiles of feature importance to explain the prediction of their wages.
+Panel (b) shows the differences in the player's median values of over 25 such sequences.<!-- Segue --> In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this observation, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
 
@@
 For the application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
-Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees.
+Instead of aggregating over sequences of the variables, tree SHAP calculates observation-level variable importance by exploring the structure of the decision trees.
 Tree SHAP is only compatible with tree-based models; random forests are used for illustration.
-The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) that will be used to explore the sensitivity of the feature contributions.
+The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) that will be used to explore the sensitivity of the variable contributions.
 
@@
 It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One key feature of the tour is the object permanence of the data points; one can track the relative change of instances in time and gain information about the relationships between points across multiple features.
+One key variable of the tour is the object permanence of the data points; one can track the relative change of observations in time and gain information about the relationships between points across multiple variables.
 There are various types of tours that are distinguished by how the paths are generated [@lee_state_2021; @cook_grand_2008].
@@
 <!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its path by changing a selected feature's contribution to a basis to allow the feature to contribute more or less to the projection.
-The requirement constrains the contribution of all other features that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a feature to structure visible in a projection.
+The manual tour [@cook_manual_1997] defines its path by changing a selected variable's contribution to a basis to allow the variable to contribute more or less to the projection.
+The requirement constrains the contribution of all other variables that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
 It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
@@
 A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of the new work.
-In a radial tour, the selected feature can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and interactive re-calculation to focus on a different feature.
+In a radial tour, the selected variable can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and interactive re-calculation to focus on a different variable.
 
@@
 There are two primary plots: the __global view__ to give the context of all of the SHAP values and the __radial tour view__ to explore the local explanations with user-controlled rotation.
-There are numerous user inputs, including feature selection for the radial tour and instance selection for making comparisons.
+There are numerous user inputs, including variable selection for the radial tour and observation selection for making comparisons.
 There are different plots used for the categorical and quantitative responses.
@@
 <!-- Purpose -->
-The global view provides context for all instances and facilitates the exploration of the separability of the data- and attribution-spaces.
-These spaces both have dimensionality $n \times p$, where $n$ is the number of instances and $p$ is the number of features.
-The attribution space corresponds to the local explanations for each instance; feature importance in the vicinity of the instance.
+The global view provides context for all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+These spaces both have dimensionality $n \times p$, where $n$ is the number of observations and $p$ is the number of variables.
+The attribution space corresponds to the local explanations for each observation; variable importance in the vicinity of the observation.
 
@@
 These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:regressioncase)a) to help identify instances poorly predicted by the model.
-<!-- Interactions --> For classification tasks, misclassified instances are circled in red.
+In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:regressioncase)a) to help identify observations poorly predicted by the model.
+<!-- Interactions --> For classification tasks, misclassified observations are circled in red.
 Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \@ref(fig:regressioncase)d).
 
-While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular instances to explore in detail.
-The projection attribution of the primary instance (PI) is examined and typically viewed with an optional comparison instance (CI).
-These instances are highlighted as asterisk and $\times$, respectively.
+While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular observations to explore in detail.
+The projection attribution of the primary observation (PI) is examined and typically viewed with an optional comparison observation (CI).
+These observations are highlighted as asterisk and $\times$, respectively.
 
@@
 <!-- Local explanations as par coords -->
-The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of features can be compared across all instances.
+The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
 These are depicted as vertical parallel coordinate plots [@ocagne_coordonnees_1885] on the basis biplot [@gabriel_biplot_1971].
 1D biplot displays the values of the current basis as bars.
-The parallel coordinate overlays lines connecting one instance's feature attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
+The parallel coordinate overlays lines connecting one observation's variable attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
 The attribution projections of the PI and CI are shown as dashed and dotted lines.
-From this plot, the range and density of the importance across all instances can be interpreted.
-For classification, one would look at differences between groups on any feature.
+From this plot, the range and density of the importance across all observations can be interpreted.
+For classification, one would look at differences between groups on any variable.
 For example, Figure \@ref(fig:classificationcase)e suggests that `bl` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which features have low values for all instances (not important).
-For example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some instances and not important for others.
+For regression, one might generally observe which variables have low values for all observations (not important).
+For example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some observations and not important for others.
 
@@
 (Remember that the PI is interactively selected from the global view).
-The attribution projection approximates the feature importance for predicting this instance.
-The combination of features best explains the difference between the mean response and an instance’s predicted value.
+The attribution projection approximates the variable importance for predicting this observation.
+The combination of variables best explains the difference between the mean response and an observation’s predicted value.
 It is not an indication of the local shape of the model surface.
@@
 The PI and CI are indicated by vertical dashed and dotted lines.
-The radial tour varies the contribution of the selected feature between 0 and 1.
+The radial tour varies the contribution of the selected variable between 0 and 1.
 This is viewed as an animation of the projections from many intermediate bases.
-Doing so tests the sensitivity of structure (class separation or strength of relationship) to the feature's contribution.
-For classification, if the separation between classes diminishes when the feature contribution is reduced, this suggests that the feature is important for class separation.
-For regression, if the relationship scatterplot weakens when the feature contribution is reduced, indicating that the feature is important for accurately predicting the response.
+Doing so tests the sensitivity of structure (class separation or strength of relationship) to the variable's contribution.
+For classification, if the separation between classes diminishes when the variable contribution is reduced, this suggests that the variable is important for class separation.
+For regression, if the relationship scatterplot weakens when the variable contribution is reduced, indicating that the variable is important for accurately predicting the response.
 
@@
 
-Selecting a misclassified instance as PI and a correctly classified point nearby in data space as CI makes it easier to examine the features most responsible for the error.
+Selecting a misclassified observation as PI and a correctly classified point nearby in data space as CI makes it easier to examine the variables most responsible for the error.
 The global view (Figure \@ref(fig:classificationcase)c) displays the model confusion matrix.
 The radial tour is 1D and displays as density where color indicates class.
-An animation slider enables users to vary the contribution of features to explore the sensitivity of the separation to that feature.
+An animation slider enables users to vary the contribution of variables to explore the sensitivity of the separation to that variable.
 
@@
 (c) prediction by observed $y$ (visual of the confusion matrix for classification tasks).
-Points are colored by predicted class, and red circles indicate misclassified instances.
-Radial tour inputs (d) select features to include and which feature is changed in the tour.
-(e) shows a parallel coordinate display of the distribution of the feature attributions while bars depict contribution for the current basis.
+Points are colored by predicted class, and red circles indicate misclassified observations.
+Radial tour inputs (d) select variables to include and which variable is changed in the tour.
+(e) shows a parallel coordinate display of the distribution of the variable attributions while bars depict contribution for the current basis.
 The black bar is the variable being changed in the radial tour.
@@
 
-Selecting an inaccurately predicted instance as PI and an accurately predicted instance with similar feature values as CI is a helpful way to understand how the model is failing or not.
+Selecting an inaccurately predicted observation as PI and an accurately predicted observation with similar variable values as CI is a helpful way to understand how the model is failing or not.
 The global view (Figure \@ref(fig:regressioncase)a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
@@
 The attribution projection can be interpreted similarly to the predicted value from the global view plot.
-It represents a linear combination of the features, and a good fit would be indicated when there is a strong relationship with the observed values.
+It represents a linear combination of the variables, and a good fit would be indicated when there is a strong relationship with the observed values.
 This can be viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a feature is varied, if the value of the PI does not change much, it would indicate that the prediction for this instance is NOT sensitive to that feature.
-Conversely, if the predicted value varies substantially, the prediction is very sensitive to that feature, suggesting that the feature is very important for the PI's prediction.
+As the contribution of a variable is varied, if the value of the PI does not change much, it would indicate that the prediction for this observation is NOT sensitive to that variable.
+Conversely, if the predicted value varies substantially, the prediction is very sensitive to that variable, suggesting that the variable is very important for the PI's prediction.
 
-```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression tasks (quantitative outputs) and illustration of interactive features.
+```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression tasks (quantitative outputs) and illustration of interactive variables.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
@@
 Coloring on a statistic (c) highlights structure organized in the attribution space.
-Interactive tabular display (d) populates when instances are selected.
-Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis.
+Interactive tabular display (d) populates when observations are selected.
+Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the variable attribution from all observations, and horizontal bars show the contribution to the current basis.
 Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right)."}
@@
 
-## Interactive Features
+## Interactive variables
 
@@
 <!-- Exploratory interactions -->
-A tooltip displays instance number/name and classification information while the cursor hovers over a point.
+A tooltip displays observation number/name and classification information while the cursor hovers over a point.
 Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \@ref(fig:classificationcase)a & b).
 The information corresponding to the selected points is populated on a dynamic table (Figure \@ref(fig:classificationcase)d).
-These interactions aid exploration of the spaces and, finally, the identification of primary and comparison instances.
+These interactions aid exploration of the spaces and, finally, the identification of primary and comparison observations.
 
@@
 - __Local explanation:__ Tree SHAP is calculated for _each_ observation using the package __treeshap__ [@kominsarczyk_treeshap_2021].
-We opt to find the attribution of each instance in the training data and not fit to fit feature interactions.
+We opt to find the attribution of each observation in the training data and not fit to fit variable interactions.
 - __Cheem viewer:__ after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
@@
 The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
-For reference, the FIFA data contained 5000 instances of nine explanatory features took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each instance took 270 seconds total.
-PCA and statistics of the features and attributions took 2.8 seconds.
+For reference, the FIFA data contained 5000 observations of nine explanatory variables took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 270 seconds total.
+PCA and statistics of the variables and attributions took 2.8 seconds.
 These runtimes were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
@@
 The data is publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete instances, there are 333 instances of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), and body mass (`bm`) for this illustration.
-A random forest model was fit with species as the response feature.
+After removing incomplete observations, there are 333 observations of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), and body mass (`bm`) for this illustration.
+A random forest model was fit with species as the response variable.
 
@@
 There is more separation visible in the attribution space than in the data space, as would be expected.
-The predicted vs observed plot reveals a handful of misclassified instances.
+The predicted vs observed plot reveals a handful of misclassified observations.
 A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
@@
 
-The radial tour starts from the attribution projection of the misclassified instance (b, left).
-The important features identified by SHAP in the (wrong) prediction for this instance are mostly `bl` and `bd` with small contributions of `fl` and `bm`.
-This projection is a view where the Gentoo (purple) looks much more likely for this instance than Chinstrap.
-That is, this combination of features is not particularly useful because the PI looks very much like other Gentoo penguins.
+The radial tour starts from the attribution projection of the misclassified observation (b, left).
+The important variables identified by SHAP in the (wrong) prediction for this observation are mostly `bl` and `bd` with small contributions of `fl` and `bm`.
+This projection is a view where the Gentoo (purple) looks much more likely for this observation than Chinstrap.
+That is, this combination of variables is not particularly useful because the PI looks very much like other Gentoo penguins.
 The radial tour is used to vary the contribution of flipper length (`fl`) to explore this.
-(In our exploration, this was the third feature explored.
-It is typically helpful to explore the features with more significant contributions, here `bl` and `bd`.
+(In our exploration, this was the third variable explored.
+It is typically helpful to explore the variables with more significant contributions, here `bl` and `bd`.
 Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
 On varying `fl` as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
-This suggests that `fl` should be considered an important feature for explaining the (wrong) prediction.
+This suggests that `fl` should be considered an important variable for explaining the (wrong) prediction.
 
 Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`fl`) is vital for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two features, mainly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard instance to account for, as decision trees can only partition only vertical and horizontal lines.
+Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two variables, mainly because it has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard observation to account for, as decision trees can only partition only vertical and horizontal lines.
 
@@
 <!-- Data description -->
-The chocolates data set consists of 88 instances of ten nutritional measurements determined from their labels and labeled as either milk or dark.
+The chocolates data set consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
 Dark chocolate is considered healthier than milk.
@@
 A reason to do this is that a dark chocolate, nutritionally more like milk should not be considered a healthy alternative.
-It is interesting to explore which nutritional features contribute most to misclassification.
+It is interesting to explore which nutritional variables contribute most to misclassification.
 
 (ref:casechocolates-cap) Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this instance.
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
 So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
@@
 <!-- Dark misclassified as milk, description -->
-This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (instance 22).
-It is compared with a CI that is a correctly classified dark chocolate (instance 7).
-The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of instances.
+This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+It is compared with a CI that is a correctly classified dark chocolate (observation 7).
+The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
 The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
@@
 The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
-In the density plot, instances on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
 The opposite would be interpreting a point with high values in this plot.
@@
 The PI (dashed line) can be viewed against the CI (dotted line).
-Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular instance, and the density plot, which is the same projection of all instances as specified by the SHAP values of the PI.
-The feature contribution of the two different predictions can be quickly compared in the parallel coordinate plot.
-The PI differs from the comparison primarily on the Fiber feature, which suggests that this is the reason for the incorrect prediction.
+Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the PI.
+The variable contribution of the two different predictions can be quickly compared in the parallel coordinate plot.
+The PI differs from the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
 
 <!-- Attention to Density, end of first case -->
-From the density plot, which is the attribution projection corresponding to the PI, both instances are more like dark chocolates.
+From the density plot, which is the attribution projection corresponding to the PI, both observations are more like dark chocolates.
 Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
-When a frame with contribution primarily from Fiber is examined instance 22 looks more like a milk chocolate.
+When a frame with contribution primarily from Fiber is examined observation 22 looks more like a milk chocolate.
 
@@
 In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
-Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (instance 71).
+Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (observation 71).
 The corresponding global view and radial tour frames are shown in Figure \@ref(fig:casechocolatesinverse).
@@
 In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed feature importance, with low values relative to other milk chocolates.
+Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates.
 The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
@@
 The contribution of Sodium is selected to be varied.
-Even in the initial projection, the instance looks slightly more like its observed milk than predicted dark chocolate.
+Even in the initial projection, the observation looks slightly more like its observed milk than predicted dark chocolate.
 The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
@@
 The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
-Nine higher-level skill groupings were identified and aggregated from highly correlated features.
+Nine higher-level skill groupings were identified and aggregated from highly correlated variables.
 A random forest model is fit from these predictors, regressing player wages [2020 euros].
-The model was fit from 5000 instances before being thinned to 500 players to mitigate occlusion and render time.
+The model was fit from 5000 observations before being thinned to 500 players to mitigate occlusion and render time.
 Continuing from the exploration in Section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
@@
 van Dijk).
-The same instances were used in Figure \@ref(fig:shapdistrbd).
+The same observations were used in Figure \@ref(fig:shapdistrbd).
 
@@
 van Dijk).
-The attribution projection is shown on the left, and it can be seen that this combination of features produces a view where Messi has very high predicted (and observed) wages.
-Defense (`def`) is the chosen feature to vary.
+The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
+Defense (`def`) is the chosen variable to vary.
 It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
@@
 If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
-However, increasing the contribution of a variable with low importance would rotate both features out of the frame.
+However, increasing the contribution of a variable with low importance would rotate both variables out of the frame.
 
@@
 
-(ref:caseames-cap) Exploring an instance with an extreme residual as the PI in relation to an instance with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
-The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this feature.
+(ref:caseames-cap) Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
+The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
 The PI has a higher predicted value than the CI in the attribution projection.
@@
 The remaining are 338 house sales.
-A random forest model was fit, predicting the sale price [USD] from the property features: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
-Using  interactions with the global view, a house with an extreme negative residual and an accurate instance with a similar prediction is selected.
+A random forest model was fit, predicting the sale price [USD] from the property variables: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
+Using  interactions with the global view, a house with an extreme negative residual and an accurate observation with a similar prediction is selected.
 
@@
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-The attribution projection places instances with high Living Areas to the right.
-The contribution of Living Area contrasts the contribution of this feature.
+The attribution projection places observations with high Living Areas to the right.
+The contribution of Living Area contrasts the contribution of this variable.
 As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
@@
 There is a clear need to extend the interpretability of black box models.
-This paper provides a technique that builds on local explanations to explore the feature importance local to an instance.
-The local explanations of an attribution projection from which feature contributions are varied using a radial tour.
-Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular features.
+This paper provides a technique that builds on local explanations to explore the variable importance local to an observation.
+The local explanations of an attribution projection from which variable contributions are varied using a radial tour.
+Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular variables.
 A global view shows the data space, explanation space, and residual plot.
-The user can interactively select instances to compare, contrast, and study further.
-Then the radial tour is used to explore the feature sensitivity identified by the attribution projection.
+The user can interactively select observations to compare, contrast, and study further.
+Then the radial tour is used to explore the variable sensitivity identified by the attribution projection.
 
@@
 Local explanations focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
-They can also form the basis of learning how the model has got it wrong, with when the instance is misclassified or has a large residual.
+They can also form the basis of learning how the model has got it wrong, with when the observation is misclassified or has a large residual.
 
@@
 
-This analysis is manually intensive and thus only feasible for investigating a few instances.
-The recommended approach is to investigate an instance where the model has not predicted accurately and compare it with an instance with similar predictor values where model fitted well.
-The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any feature.
-It can be helpful to make additional plots of the features and responses to cross-check interpretations made from the cheem viewer.
+This analysis is manually intensive and thus only feasible for investigating a few observations.
+The recommended approach is to investigate an observation where the model has not predicted accurately and compare it with an observation with similar predictor values where model fitted well.
+The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any variable.
+It can be helpful to make additional plots of the variables and responses to cross-check interpretations made from the cheem viewer.
 This methodology provides an additional tool in the box for studying model fitting.

### 2023-05-04 paper.Rmd
corrected abstract

@@
 This trade-off has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction with _local explanations_, a point estimate of the linear variable importance in the vicinity of one observation.
-These can be considered linear projections and can be further explored to understand better the interactions between variables used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output.
+XAI attempts to shed light on how models use predictors to arrive at a prediction.
+_Local explanations_, which provide a point estimate of the linear variable importance in the vicinity of one observation is one XAI method.
+These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
 The methods are implemented in the __R__ package __cheem__, available on CRAN.

### 2023-05-05 paper.Rmd
revised introduction

@@
 XAI attempts to shed light on how models use predictors to arrive at a prediction.
-_Local explanations_, which provide a point estimate of the linear variable importance in the vicinity of one observation is one XAI method.
+"Local explanations", which provide a point estimate of the linear variable importance in the vicinity of one observation, are one XAI method.
 These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface.
@@
 Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
-Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data preprocessing, and package infrastructure.
-Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative outputs.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data pre-processing, and package infrastructure.
+Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative response variables.
+These show how the local explanations can be used to get an overview of the model's use of predictors, and to investigate errors in the model predictions.
 Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
 The methods are implemented in the __R__ package __cheem__.
-Example data sets are provided to illustrate usage for classification and regression tasks.
 
@@
 It can be hard to determine whether small changes in a variable's value will make a class prediction change group or identify which variables contribute to an extreme residual.
-Local explanations shed light on these situations by approximating linear variable importance in the vicinity of a single observation.
+Local explanations shed light on these situations by estimate linear variable importance in the vicinity of a single observation.
+There are many approaches for calculating local explanations.
 
 <!-- Taxonomy of local explanations -->
-A comprehensive summary of the taxonomy and literature on explanation techniques is provided in Figure 6 by @arrieta_explainable_2020.
+A comprehensive summary of the taxonomy of currently available methods is provided in Figure 6 by @arrieta_explainable_2020.
 It includes a large number of model-specific explanations such as deepLIFT [@shrikumar_not_2016; @shrikumar_learning_2017], a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic explanations, of which LIME, [@ribeiro_why_2016] SHAP, [@lundberg_unified_2017], and their variants are popular.
+There are fewer model-agnostic methods, of which LIME, [@ribeiro_why_2016] SHaply Additive exPlanations (SHAP), [@lundberg_unified_2017], are popular.
 
@@
 These observation-level explanations are used in various ways depending on the data.
-In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap indicating important pixels [@simonyan_deep_2014].
-For observation, pixels corresponding to snow may be highlighted when distinguishing if a picture contains a wolf or husky.
+In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap to indicate important pixels [@simonyan_deep_2014].
+For example, pixels corresponding to snow may be highlighted as important contributors when distinguishing if a picture contains a coyote or husky.
 In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words [@vanni_textual_2018].
-In the case of numeric regression, they are used to explain variable additive contributions from the model intercept to the observation's prediction [@ribeiro_why_2016].
+In the case of numeric regression, they are used to explain additive contributions of variables from the model intercept to the observation's prediction [@ribeiro_why_2016].
 
 <!-- SHAP -->
-SHaply Additive exPlanations (SHAP) quantifies the variable contributions of one observation by examining the effect of other variables on the predictions.
+We will be focusing on SHAP values in this this paper, but the approach is applicable for any methods of calculating the local explanations.
+SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
 The SHAP explanation refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
-@strumbelj_efficient_2010 introduced SHAP for local explanations in ML models.
-The attribution of variable importance depends on the sequence of the included variables.
-The SHAP values are the mean contributions over different variable sequences.
-The approach is related to partial dependence plots [@molnar_interpretable_2020], used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
+@strumbelj_efficient_2010 introduced SHAP for local explanations in machine learning models.
+The attribution of variable importance depends on the sequence being the included variables.
+The SHAP value for an observation are the average contributions of each variable, computed for different sequences, or permutations, of the variables.
+The approach is related to partial dependence plots (see for example @molnar_interpretable_2020), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
 Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015 help to understand the relative importance of each variable in the context of all other candidate variables.
+It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015 which help to understand the relative importance of each variable in the context of all other candidate variables.
 
@@
 <!-- Fifa example -->
-Following the use case _Explanatory Model Analysis_ [@biecek_explanatory_2021], FIFA data is used to  illustrate SHAP.
-Consider soccer data from the FIFA 2020 season [@leone_fifa_2020].
-There are 5000 observations of 9 skill measures (after aggregating highly correlated variables).
-A random forest model is fit regressing players' wages [2020 Euros] from their skill measurements.
-The SHAP values are compared for a star offensive player (L.
-Messi) and defensive player (V.
+We use 2020 season FIFA data [@leone_fifa_2020] to illustrate SHAP following the procedures described in @biecek_explanatory_2021.
+There are 5000 observations of nine predictor variables measuring players' skills and one response variable, wages (in euros).
+A random forest model is fit regressing players' wages on the skill variables.
+In this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are compared for a star offensive player (L.
+Messi) and a prominent defensive player (V.
 van Dijk).
-The results are displayed in Figure \@ref(fig:shapdistrbd).
+What we are interested to know is how the skill variables locally contribute to the wage prediction of each player.
 A difference in the attribution of the variable importance across the two positions of the players can be expected.
-This would be interpreted as how a player's salary depends on this combination of skills.
-Panel (a) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where three sequences of variables are presented, so the two observations can be more easily compared.
-The magnitude of the contributions depends on the sequence in which they appear.
-Panel (b) shows the differences in the player's median values of over 25 such sequences.<!-- Segue --> In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+This would be interpreted as how a player's salary depends on which combination of skills.
+Panel (a) is a version of a breakdown plot [@gosiewska_ibreakdown_2019] where just three sequences of variables are shown.
+Here we can see how the variable contribution can change depending on sequence, relative to both players.
+(Note that the order of the variables is different in each plot, because they have been sorted by biggest average contribution across both players.) For all sequences, and for both players `reaction` has the strongest contribution, with perhaps more importance for the defensive player.
+Then it differs by player: for Messi `offense` and `movement` have the strongest contributions, and for van Dijk it is `defense` and `power`, regardless of the variable sequence.
+
+Panel (b) shows the differences in the player's median values (large dots) for 25 such sequences (tick marks).
+We can see that the wage predictions for the two players come from different combinations of skills sets, as might be expected for players who's value on the team depends on their offensive or defensive prowess.
+It is also interesting to see from the distribution of values across the different sequence of variables, that there is some multimodality.
+For example, look at the SAHP values for `reaction` for Messi, and in some sequences reaction has a much lower contribution than others.
+This suggests that other variables (`offense`, `movement` probably) can substitute for `reaction` in the wage prediction.
+
+In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
 In this observation, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
@@
 Instead of aggregating over sequences of the variables, tree SHAP calculates observation-level variable importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models; random forests are used for illustration.
-The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) that will be used to explore the sensitivity of the variable contributions.
+Tree SHAP is only compatible with tree-based models.
+so random forests are used for illustration.
+<!--The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) that will be used to explore the sensitivity of the variable contributions.-->
 
@@
 There are different plots used for the categorical and quantitative responses.
-Figures \@ref(fig:classificationcase) and \@ref(fig:regressioncase) are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
+Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \@ref(fig:classificationcase)b, \@ref(fig:regressioncase)a) to help identify observations poorly predicted by the model.
+In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify observations poorly predicted by the model.
 <!-- Interactions --> For classification tasks, misclassified observations are circled in red.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \@ref(fig:regressioncase)d).
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
 
@@
 1D biplot displays the values of the current basis as bars.
-The parallel coordinate overlays lines connecting one observation's variable attribution (Figures \@ref(fig:classificationcase)e and \@ref(fig:regressioncase)e).
+The parallel coordinate overlays lines connecting one observation's variable attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
 The attribution projections of the PI and CI are shown as dashed and dotted lines.
@@
 For classification, one would look at differences between groups on any variable.
-For example, Figure \@ref(fig:classificationcase)e suggests that `bl` is important for distinguishing the green class from the other two.
+For example, Figure \ref{fig:classificationcase}e suggests that `bl` is important for distinguishing the green class from the other two.
 For regression, one might generally observe which variables have low values for all observations (not important).
-For example, `BMI` and `pwr` in Figure \@ref(fig:regressioncase)e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some observations and not important for others.
+For example, `BMI` and `pwr` in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some observations and not important for others.
 
@@
 <!-- Tour animation -->
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \@ref(fig:classificationcase)f) and as scatterplots for a quantitative response (Figure \@ref(fig:regressioncase)f).
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
 The PI and CI are indicated by vertical dashed and dotted lines.
@@
 Selecting a misclassified observation as PI and a correctly classified point nearby in data space as CI makes it easier to examine the variables most responsible for the error.
-The global view (Figure \@ref(fig:classificationcase)c) displays the model confusion matrix.
+The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
 The radial tour is 1D and displays as density where color indicates class.
@@
 
-```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks (categorical outputs).
+```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks (categorical response).
 Global view inputs, (a), set the PI, CI, and color statistic.
@@
 Selecting an inaccurately predicted observation as PI and an accurately predicted observation with similar variable values as CI is a helpful way to understand how the model is failing or not.
-The global view (Figure \@ref(fig:regressioncase)a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
 The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation to aid in understanding the structure identified in these spaces.
@@
 
-```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression tasks (quantitative outputs) and illustration of interactive variables.
+```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
@@
 The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible (Figure \@ref(fig:classificationcase)a & d).
+These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a & d).
 The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
@@
 A tooltip displays observation number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \@ref(fig:classificationcase)a & b).
-The information corresponding to the selected points is populated on a dynamic table (Figure \@ref(fig:classificationcase)d).
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a & b).
+The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
 These interactions aid exploration of the spaces and, finally, the identification of primary and comparison observations.
@@
 
-Figure \@ref(fig:casepenguins) shows plots from the cheem viewer for exploring the random forest model on the penguins data.
+Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
 Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
@@
 
-Figure \@ref(fig:casepenguinsblfl) confirms that flipper length (`fl`) is vital for the confusion of the PI as a Chinstrap.
+Figure \ref{fig:casepenguinsblfl} confirms that flipper length (`fl`) is vital for the confusion of the PI as a Chinstrap.
 Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two variables, mainly because it has an unusually low value of flipper length relative to other Gentoos.
@@
 <!-- Dark misclassified as milk, description -->
-This type of exploration is shown in Figure \@ref(fig:casechocolates), where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
 It is compared with a CI that is a correctly classified dark chocolate (observation 7).
@@
 Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (observation 71).
-The corresponding global view and radial tour frames are shown in Figure \@ref(fig:casechocolatesinverse).
+The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
 
@@
 The model was fit from 5000 observations before being thinned to 500 players to mitigate occlusion and render time.
-Continuing from the exploration in Section \@ref(sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
+Continuing from the exploration in Section \ref{sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
 That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
@@
 van Dijk).
-The same observations were used in Figure \@ref(fig:shapdistrbd).
+The same observations were used in Figure \ref{fig:shapdistrbd}.
 
@@
 
-Figure \@ref(fig:casefifa) tests the support of the local explanation.
+Figure \ref{fig:casefifa} tests the support of the local explanation.
 Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
@@
 
-Figure \@ref(fig:caseames) selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.

### 2023-05-05 figures/_make_figures_case_study.r
data should be available with the paper

@@ -8,18 +8,11 @@
   require("cowplot")
   require("patchwork")
   
-  wd <- getwd()
-  if(substr(wd, nchar(wd) - 10, nchar(wd)) != "cheem_paper")
-    warning("Expected work directory to be cheem_paper.")
-  fp <- "../cheem/inst/shiny_apps/cheem_initial/data/"
-  if(F)
-    dir("../cheem/inst/shiny_apps/cheem_initial/data/")
-  
   ## Load data:
-  penguins_ls   <- readRDS(paste0(fp, "preprocess_penguins.rds"))
-  fifa_ls       <- readRDS(paste0(fp, "preprocess_fifa.rds"))
-  ames2018_ls   <- readRDS(paste0(fp, "preprocess_ames2018.rds"))
-  chocolates_ls <- readRDS(paste0(fp, "preprocess_chocolates.rds"))
+  penguins_ls   <- readRDS("data/preprocess_penguins.rds")
+  fifa_ls       <- readRDS("data/preprocess_fifa.rds")
+  ames2018_ls   <- readRDS("data/preprocess_ames2018.rds")
+  chocolates_ls <- readRDS("data/preprocess_chocolates.rds")
   
   .t <- theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
               axis.text   = element_blank(),

### 2023-05-05 paper.Rmd
added preprocessing code and tweaked intro more

@@
 The interpretation is valid for the full domain of the predictors.
-In nonlinear models one needs to consider the model in small neighbourhoods of the domain to make any assessment of variable importance.
-Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues decribed in @stahl-ethics.
-Understanding how nonlinear models behave when usage extrapolates outside the domain of predictors, either in subspaces where few samples were provided in the training set, or extending outside the domain.
+In nonlinear models one needs to consider the model in small neighborhoods of the domain to make any assessment of variable importance.
+Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues described in @stahl-ethics.
+Understanding how nonlinear models behave when usage extrapolates outside the domain of predictors, either in sub-spaces where few samples were provided in the training set, or extending outside the domain.
 It is especially important because nonlinear models can vary wildly and predictions can be dramatically wrong in these areas.
@@
 <!-- Reminder of local explanation -->
-Consider a highly nonlinear model.
-It can be hard to determine whether small changes in a variable's value will make a class prediction change group or identify which variables contribute to an extreme residual.
-Local explanations shed light on these situations by estimate linear variable importance in the vicinity of a single observation.
-There are many approaches for calculating local explanations.
-
+Local explanations shed light on nonlinear model fits by estimating linear variable importance in the vicinity of a single observation.  There are many approaches for calculating local explanations.
 <!-- Taxonomy of local explanations -->
@@
 <!-- SHAP -->
-We will be focusing on SHAP values in this this paper, but the approach is applicable for any methods of calculating the local explanations.
+We will be focusing on SHAP values in this this paper, but the approach is applicable for any method used to calculate the local explanations.
 SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
-The SHAP explanation refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
+The term "SHAP" refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
 @strumbelj_efficient_2010 introduced SHAP for local explanations in machine learning models.
-The attribution of variable importance depends on the sequence being the included variables.
-The SHAP value for an observation are the average contributions of each variable, computed for different sequences, or permutations, of the variables.
+Variable importance can depend on the sequence in which variables are entered into the model fitting process, thus for any sequence we get a set of variable contribution values for a single observation.
+These values will add up to the difference between the fitted value for the observation, and the average fitted value for all observations.
+Using all possible sequence, or permutation, gives multiple values for each variable, which are averaged to get the SHAP value for an observation.
+It can be helpful to standardize variables prior to computing SHAP values, if they have been measured on different scales.
+
 The approach is related to partial dependence plots (see for example @molnar_interpretable_2020), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
@@
 van Dijk).
-What we are interested to know is how the skill variables locally contribute to the wage prediction of each player.
+We are interested in knowing how the skill variables locally contribute to the wage prediction of each player.
 A difference in the attribution of the variable importance across the two positions of the players can be expected.
 This would be interpreted as how a player's salary depends on which combination of skills.
-Panel (a) is a version of a breakdown plot [@gosiewska_ibreakdown_2019] where just three sequences of variables are shown.
+Panel (a) is a version of a breakdown plot [@gosiewska_ibreakdown_2019] where just three sequences of variables are shown, for two observations.
+A breakdown plot shows the absolute values of the variable attribution for an observation, usually sorted from highest value to the lowest.
+There is no scale on the horizontal axis here because values are considered relative to each other.
 Here we can see how the variable contribution can change depending on sequence, relative to both players.

### 2023-05-05 data/preprocess_ames2018.r

@@ -1,0 +1,34 @@
+{
+  dat  <- amesHousing2018_NorthAmes
+  X    <- dat[, 1:9]
+  colnames(X) <- c("LtA", "Qlt", "YrB", "LvA", "Bth", "Bdr", "Rms", "GYB", "GrA")
+  Y    <- dat$SalePrice
+  clas <- amesHousing2018_NorthAmes$SubclassMS ## class is a zone subclass
+}
+
+rf_fit <- randomForest::randomForest(
+  X, Y, ntree = 125,
+  mtry = ifelse(is_discrete(Y), sqrt(ncol(X)), ncol(X) / 3),
+  nodesize = max(ifelse(is_discrete(Y), 1, 5), nrow(X) / 500))
+rf_pred <- predict(rf_fit, X)
+rf_shap <- treeshap::treeshap(
+  treeshap::randomForest.unify(rf_fit, X), X, FALSE, FALSE)
+rf_shap <- rf_shap$shaps
+
+chm <- cheem_ls(X, Y, rf_shap, rf_pred, clas,
+                label = "North Ames, RF, treeshap")
+
+## Export ----
+NM <- "preprocess_ames2018.rds"
+saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+cat("Saved", NM, "\n")
+
+if(F){
+  ## Don't run load cheem list
+  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  lapply(chm, object.size)
+  
+  ## Don't run manual check
+  names(chm)
+  global_view(chm)
+}

### 2023-05-05 data/preprocess_chocolates.r

@@ -1,0 +1,37 @@
+{
+  X <- chocolates[, 5:14] %>% as.data.frame()
+  Y <- clas <- chocolates$Type
+  colnames(X) <- gsub("\\_.*", "", colnames(X))
+  nm_imc <- paste(chocolates$Name, chocolates$MFR, chocolates$Country, sep = ", ")
+  r_idx <- which(nm_imc == "85% Cocoa Dark French Chocolate, Thorntons, UK")[2]
+  nm_imc[r_idx] <- paste0(nm_imc[r_idx], " (2nd)")
+  row.names(X) <- nm_imc
+}
+
+## Model and predict
+train    <- data.matrix(X) %>% xgb.DMatrix(label = Y)
+xgb_fit  <- xgboost(data = train, max.depth = 3, nrounds = 25)
+xgb_pred <- predict(xgb_fit, newdata = train)
+
+## shapviz
+xgb_shap <- shapviz(xgb_fit, X_pred = train, X = X)
+xgb_shap <- xgb_shap$S
+
+## Cheem
+chm <- cheem_ls(X, Y, xgb_shap, xgb_pred, clas,
+                label = "Chocolates, xgb, shapviz")
+
+## Export ----
+NM <- "preprocess_chocolates.rds"
+saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+cat("Saved", NM, "\n")
+
+if(F){
+  ## Don't run load cheem list
+  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  lapply(chm, object.size)
+  
+  ## Don't run manual check
+  names(chm)
+  global_view(chm)
+}

### 2023-05-05 data/preprocess_fifa.r

@@ -1,0 +1,81 @@
+## Setup ------
+{
+  .raw <- DALEX::fifa
+  .dat_less_ys <- .raw %>%
+    dplyr::select(
+      -c(`nationality`, ## useless class
+         ## potential target vars:
+         `overall`, `potential`, `value_eur`, `wage_eur`)) %>%
+    as.data.frame()
+  
+  if(F) ## Don't run, View corrplot of the groupings
+    corrplot::corrplot(cor(.dat_less_ys),
+                       method = "circle", ## geom
+                       type = "upper", ## only upper triangle
+                       diag = F, ## remove auto correlation
+                       order = "FPC", ## First principal component
+                       tl.col = "black", tl.srt = 90, ## Text label color and rotation
+                       tl.pos = "td")
+  
+  ## Aggregate some highly correlated vars.
+  dat <- .dat_less_ys %>%
+    dplyr::mutate(
+      .keep = "none",
+      BMI = (weight_kg+(height_cm / 100)^2)/2,
+      age = age,
+      react = movement_reactions,
+      off = (attacking_finishing+skill_long_passing+attacking_volleys+
+               power_long_shots+skill_curve+mentality_positioning+attacking_crossing+
+               attacking_short_passing+skill_dribbling+skill_ball_control)/10,
+      def = (defending_sliding_tackle+mentality_interceptions+
+               defending_standing_tackle+defending_marking+mentality_aggression)/5,
+      acc = (attacking_heading_accuracy+power_shot_power)/2,
+      mvm = (movement_sprint_speed+movement_balance+movement_acceleration+
+               mentality_vision+mentality_composure+movement_agility+
+               mentality_penalties+skill_fk_accuracy+power_stamina+movement_reactions)/10L,
+      pwr = (power_strength+power_jumping)/2,
+      gk  = (goalkeeping_diving+goalkeeping_positioning+goalkeeping_reflexes+
+               goalkeeping_handling+goalkeeping_kicking)/5
+    )
+  
+  ## Starting with 42 variables, we remove `nationality`, and some potential Y vars,
[…]

### 2023-05-05 data/preprocess_penguins.r

@@ -1,0 +1,35 @@
+## Setup ------
+{
+  raw <- spinifex::penguins_na.rm
+  X <- raw[, 1:4] %>% as.data.frame()
+  Y <- clas <- raw$species
+  colnames(X) <- c("bl", "bd", "fl", "bm")
+}
+
+## Model and predict
+train    <- data.matrix(X) %>% xgb.DMatrix(label = Y)
+xgb_fit  <- xgboost(data = train, max.depth = 3, nrounds = 25)
+xgb_pred <- predict(xgb_fit, newdata = train)
+
+## shapviz
+xgb_shap <- shapviz(xgb_fit, X_pred = train, X = X)
+xgb_shap <- xgb_shap$S
+
+## Cheem
+chm <- cheem_ls(X, Y, xgb_shap, xgb_pred, clas,
+                label = "Penguins, xgb, shapviz")
+
+## Export ----
+NM <- "preprocess_penguins.rds"
+saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+cat("Saved", NM, "\n")
+
+if(F){
+  ## Don't run load cheem list
+  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  lapply(chm, object.size)
+  
+  ## Don't run manual check
+  names(chm)
+  global_view(chm)
+}

### 2023-05-05 data/preprocess_ames2018.r
added raw data and fixed preprocessing code

@@ -1,3 +1,4 @@
+load("data/amesHousing2018_NorthAmes.rda")
 {
   dat  <- amesHousing2018_NorthAmes
   X    <- dat[, 1:9]
@@ -20,12 +21,12 @@ chm <- cheem_ls(X, Y, rf_shap, rf_pred, clas,
 
 ## Export ----
 NM <- "preprocess_ames2018.rds"
-saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+saveRDS(chm, file = paste0("data/", NM))
 cat("Saved", NM, "\n")
 
 if(F){
   ## Don't run load cheem list
-  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  chm <- readRDS(paste0("data/", NM))
   lapply(chm, object.size)
   
   ## Don't run manual check

### 2023-05-05 data/preprocess_chocolates.r

@@ -1,3 +1,4 @@
+load("data/chocolates.rda")
 {
   X <- chocolates[, 5:14] %>% as.data.frame()
   Y <- clas <- chocolates$Type
@@ -23,12 +24,12 @@ chm <- cheem_ls(X, Y, xgb_shap, xgb_pred, clas,
 
 ## Export ----
 NM <- "preprocess_chocolates.rds"
-saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+saveRDS(chm, file = paste0("data/", NM))
 cat("Saved", NM, "\n")
 
 if(F){
   ## Don't run load cheem list
-  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  chm <- readRDS(paste0("data/", NM))
   lapply(chm, object.size)
   
   ## Don't run manual check

### 2023-05-05 data/preprocess_fifa.r

@@ -67,12 +67,12 @@ chm <- subset_cheem(chm, 1:500)
 
 ## Export ----
 NM <- "preprocess_fifa.rds"
-saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+saveRDS(chm, file = paste0("data/", NM))
 cat("Saved", NM, "\n")
 
 if(F){
   ## Don't run load cheem list
-  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  chm <- readRDS(paste0("data/", NM))
   lapply(chm, object.size)
   
   ## Don't run manual check

### 2023-05-05 data/preprocess_penguins.r

@@ -21,12 +21,12 @@ chm <- cheem_ls(X, Y, xgb_shap, xgb_pred, clas,
 
 ## Export ----
 NM <- "preprocess_penguins.rds"
-saveRDS(chm, file = paste0("~/R/cheem/inst/shiny_apps/cheem/data/", NM))
+saveRDS(chm, file = paste0("data/", NM))
 cat("Saved", NM, "\n")
 
 if(F){
   ## Don't run load cheem list
-  chm <- readRDS(paste0("./inst/shiny_apps/cheem/data/", NM))
+  chm <- readRDS(paste0("data/", NM))
   lapply(chm, object.size)
   
   ## Don't run manual check

### 2023-05-05 paper.Rmd
added figure explaining radial tour, and removed shapley lenses section, but made some of this point in the second last paragraph of section 2

@@
 Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
-It could also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015 which help to understand the relative importance of each variable in the context of all other candidate variables.
 
@@
 It is also interesting to see from the distribution of values across the different sequence of variables, that there is some multimodality.
-For example, look at the SAHP values for `reaction` for Messi, and in some sequences reaction has a much lower contribution than others.
+For example, look at the SHAP values for `reaction` for Messi, and in some sequences reaction has a much lower contribution than others.
 This suggests that other variables (`offense`, `movement` probably) can substitute for `reaction` in the wage prediction.
 
-In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this observation, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.
+This can also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015.
+Various models that are similarly good might use different combinations of the variables.
+Examining the coefficients from multiple models helps to understand the relative importance of each variable in the context of all other variables.
+This is similar to the approach here with SHAP values, that by examining the variation in values across different permutations of variables, we can gain more understanding of the relationship between the response and predictors.
+
+<!--In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
+In this observation, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.-->
 
@@
 <!-- Radial tour -->
-A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of the new work.
+A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of this new work.
 In a radial tour, the selected variable can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
@@
 
+```{r radialtour, echo=F, fig.align='center', out.width="99%", fig.cap = "The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to structure in the plot.
+Here we have a 1D projection of the penguins data displayed as a density plot.
+The line segments on the bottom correspond to the coefficients of the variables making up the projection.
+The structure in the plot is bimodality (left), and the importance of the variable \\textsf{bd} is being explored.
+As this variable contribution is reduced in the plot (middle, right) we can see that the bimodality decreases.
+Thus \\textsf{bd} is an important variable contributing to the bimodal structure."}
+knitr::include_graphics("./figures/radial_tour.png")
+```
+
+<!--
 # Shapley's Lenses {#sec:shapleylenses}
@@
 The response to the challenge in the characterization of Rashomon sets may be Shapley's Lenses - the opportunity to compare these models from the perspective of exploring the local attributions generated by the Shapley values for different models.
+
+-->
 

### 2023-05-05 paper.Rmd
revised section 4.2

@@
 <!-- Purpose -->
-The global view provides context for all observations and facilitates the exploration of the separability of the data- and attribution-spaces.
+The global view provides context for all observations and facilitates the exploration of the separability of the data and attribution spaces.
+The attribution space refers to the SHAP values for each observation.
 These spaces both have dimensionality $n \times p$, where $n$ is the number of observations and $p$ is the number of variables.
-The attribution space corresponds to the local explanations for each observation; variable importance in the vicinity of the observation.
 
@@
 In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify observations poorly predicted by the model.
-<!-- Interactions --> For classification tasks, misclassified observations are circled in red.
+<!-- Interactions --> For classification tasks, color indicates the predicted class, and misclassified observations are circled in red.
 Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
@@
 While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular observations to explore in detail.
-The projection attribution of the primary observation (PI) is examined and typically viewed with an optional comparison observation (CI).
+We have designed it to enable a comparison between an observation that is interesting is some way, perhaps misclassified, or poorly predicted, relative to an observation with similar predictor values but more expected prediction.
+For brevity, we call the interesting observation the primary investigation (PI) and the other is the comparison investigation (CI).
 These observations are highlighted as asterisk and $\times$, respectively.
@@
 
+There are two plots in this part of the interface.
+The first (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a display of the SHAP values for all observations.
+This will generally give the global view of variables important for the fit as a whole, but it will also highlight observations which have different patterns.
+The second plot is the radial tour, which for classification is a density plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for regression are scatterplots of the observed response values, and residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
+
 <!-- Local explanations as par coords -->
 The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
-These are depicted as vertical parallel coordinate plots [@ocagne_coordonnees_1885] on the basis biplot [@gabriel_biplot_1971].
-1D biplot displays the values of the current basis as bars.
-The parallel coordinate overlays lines connecting one observation's variable attribution (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e).
-The attribution projections of the PI and CI are shown as dashed and dotted lines.
-From this plot, the range and density of the importance across all observations can be interpreted.
-For classification, one would look at differences between groups on any variable.
+These are depicted as a vertical parallel coordinate plot [@ocagne_coordonnees_1885].
+(The SHAP values of the PI and CI are shown as dashed and dotted lines, respectively.) One should obtain a sense of the overall importance of variables from this plot.
+The more important variables will have larger values, and in the case of classification tasks variables which have different magnitude for different classes are more globally important.
 For example, Figure \ref{fig:classificationcase}e suggests that `bl` is important for distinguishing the green class from the other two.
@@
 <!-- Segue of PI to attribution projection -->
-The overlaid bars on the parallel coordinate plot represent the attribution projection of the PI.
-(Remember that the PI is interactively selected from the global view).
-The attribution projection approximates the variable importance for predicting this observation.
+A barchart is overlaid to represent the projection shown in the radial tour at right.
+It starts from the SHAP values of the PI, but if the user changes the projection the length of these bars will reflect this change.
+(The PI is interactively selected by clicking on a point in the global view).
+By scaling the SHAP value it becomes an (attribution) projection.
+<!-- approximates the variable importance for predicting this observation.
 The combination of variables best explains the difference between the mean response and an observation’s predicted value.
 It is not an indication of the local shape of the model surface.
-That is, it is not some indication of the tangent to the curve at this point.
+That is, it is not some indication of the tangent to the curve at this point.-->
 
@@
 The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
-The PI and CI are indicated by vertical dashed and dotted lines.
-The radial tour varies the contribution of the selected variable between 0 and 1.
+The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The radial tour varies the contribution of the selected variable.
 This is viewed as an animation of the projections from many intermediate bases.
@@
 [Alternatively, the package __fastshap__ claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.]
-
-## Package Infrastructure {#sec:infrastructure}
-
-The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
-Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002] and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
-The application was made with __shiny__ [@chang_shiny_2021].
-The tour visual is built with __spinifex__ [@spyrison_spinifex_2020].
-Both views are created first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive `html` widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a boon to understanding local explanations and how to apply them.
-
-## Installation and Getting Started
-
-The package can be installed from CRAN, and the application can be run using the following __R__ code:
-
-```{r eval=FALSE, echo=TRUE}
-install.packages("cheem", dependencies = TRUE)
-library("cheem")
-run_app()
-```
-
-Alternatively,
-
-- A version of the cheem viewer shiny app can be directly accessed at
-\url{https://ebsmonash.shinyapps.io/cheem_initial/}.
-- The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
-- Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
-
-Follow the examples provided with the package to compute the local explanation (using `?cheem_ls`).
-The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()` to be uploaded.
 
@@
 <!--checklist asks for: Acknowledgments, Appendices, then References-->
+## Package Infrastructure {#sec:infrastructure}
+
+The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
+Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002] and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
+The application was made with __shiny__ [@chang_shiny_2021].
+The tour visual is built with __spinifex__ [@spyrison_spinifex_2020].
+Both views are created first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive `html` widgets with __plotly__ [@sievert_interactive_2020].
+__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a boon to understanding local explanations and how to apply them.
+
+## Installation and Getting Started
+
+The package can be installed from CRAN, and the application can be run using the following __R__ code:
+
+```{r eval=FALSE, echo=TRUE}
+install.packages("cheem", dependencies = TRUE)
+library("cheem")
+run_app()
+```
+
+Alternatively,
+
+- A version of the cheem viewer shiny app can be directly accessed at
+\url{https://ebsmonash.shinyapps.io/cheem_initial/}.
+- The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
+- Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+
+Follow the examples provided with the package to compute the local explanation (using `?cheem_ls`).
+The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()` to be uploaded.
+
 ## Acknowledgments {-}
@@
 Kim Marriott provided advice on many aspects of this work, especially on the explanations in the applications section.
-We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader MI 2 lab group for the DALEX ecosystem of __R__ and Python packages.
 This research was supported by the Australian Government Research Training Program (RTP) scholarships.

### 2023-05-05 paper.Rmd
finished revising

@@
 Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-[Alternatively, the package __fastshap__ claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.]
+Alternatively, the package __fastshap__ claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.
 
@@
 
+<!--checklist asks for: Acknowledgments, Appendices, then References-->
+
+# Package Infrastructure {#sec:infrastructure}
+
 An implementation is provided in the open-source __R__ package __cheem__, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-In theory, this approach would work with any black box model, but the implementation currently only calculates tree SHAP for tree-based models supported by __treeshap__ [tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively].
-Tree SHAP was selected because of its computational efficiency.
-The SHAP and oscillation explanations could be added with the use of the `DALEX::explain()` and would be an excellent direction to extend the work [@biecek_dalex_2018; @biecek_explanatory_2021].
+The local explanations need to be pre-computed and uploaded.
+Examples show how to do this for tree SHAP values, using  __treeshap__ (tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively).
+The SHAP and oscillation explanations could be easily added using `DALEX::explain()` [@biecek_dalex_2018; @biecek_explanatory_2021].
 
-<!--checklist asks for: Acknowledgments, Appendices, then References-->
-## Package Infrastructure {#sec:infrastructure}
-
-The above-described method and application are implemented as an open-source __R__ package, __cheem__ available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
-Preprocessing was facilitated with models created via __randomForest__ [@liaw_classification_2002] and explanations calculated with __treeshap__ [@kominsarczyk_treeshap_2021].
 The application was made with __shiny__ [@chang_shiny_2021].
@@
 __DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a boon to understanding local explanations and how to apply them.
-
-## Installation and Getting Started
 

### 2023-05-05 paper.Rmd
spell check

@@
 Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
-There are common approaches for visualising high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
+There are common approaches for visualizing high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
 
@@
 <!-- What-if & counterfactual analysis -->
-This approach could be considered to be a counterfactual, what-if analysis, such as _ceteris paribus_ ("other things held constant") profiles [@biecek_ceterisparibus_2020].
+This approach could be considered to be a counter-factual, what-if analysis, such as _ceteris paribus_ ("other things held constant") profiles [@biecek_ceterisparibus_2020].
 <!-- These profiles show how an observation's prediction would change from a marginal change in one explanatory variable, given that other variables are held constant.
@@
 <!-- Segue of PI to attribution projection -->
-A barchart is overlaid to represent the projection shown in the radial tour at right.
+A bar chart is overlaid to represent the projection shown in the radial tour at right.
 It starts from the SHAP values of the PI, but if the user changes the projection the length of these bars will reflect this change.
@@
 It is vital to mitigate the render time of visuals, especially when users may want to iterate many explorations.
-All computational operations should be prepared before runtime.
+All computational operations should be prepared before run time.
 The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
@@
 PCA and statistics of the variables and attributions took 2.8 seconds.
-These runtimes were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
+These run times were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
 Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-Alternatively, the package __fastshap__ claims extremely low runtimes, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.
+Alternatively, the package __fastshap__ claims extremely low run times, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.
 

### 2023-05-08 jss/cheem_paper.Rmd
JSS formatted paper

--- 2023-05-05 paper.Rmd
@@
 ---
-title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
-type: ARTICLE TEMPLATE
+documentclass: jss
 author:
   - name: Nicholas Spyrison
-    affiliation: a
     orcid: 0000-0002-8417-0212
-    email: spyrison@gmail.com
+    affiliation: Monash University
+    # use this syntax to add text on several lines
+    address: |
+      | Dept of Human Centred Computing
+      | Faculty of Information Technology
+    email: \email{spyrison@gmail.com}
+    url: https://nspyrison.netlify.app
   - name: Dianne Cook
-    affiliation: a
     orcid: 0000-0002-3813-7155
-    email: dicook@monash.edu
+    address: |
+      | Dept of Econometrics and Business Statistics
+    affiliation: Monash University
   - name: Przemyslaw Biecek
-    affiliation: b
     orcid: 0000-0001-8423-1823
-    email: przemyslaw.biecek@gmail.com
-affiliation:
-  - num: a
-    address: |
-      Monash University, Melbourne, Australia
-  - num: b
-    address: |
-      Warsaw University of Technology and University, Warsaw, Poland
-abstract: |
+    affiliation: |
+      | Warsaw University of 
+      | Technology and University
+title:
+  formatted: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
+  plain:     "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
+  # For running headers, if needed
+  short:     "Exploring Local Explanations"
+abstract: >
     The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
@@
 Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
-The methods are implemented in the __R__ package __cheem__, available on CRAN.
-keywords: | 
-  explainable artificial intelligence; nonlinear model interpretability; visual analytics; local explanations; grand tour; radial tour
+The methods are implemented in the R package cheem, available on CRAN.
+keywords:
+  # at least one keyword must be supplied
+  formatted: [explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour]
+  plain:     [explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour]
+preamble: >
+  \usepackage{amsmath}
+output: rticles::jss_article
 bibliography: paper.bib
-header-includes: |
-  \usepackage{hyperref}
-  \usepackage[utf8]{inputenc}
-  \def\tightlist{}
-output: rticles::tf_article
 ---
 
-```{r include=FALSE, cache=FALSE}
+```{r, setup, include=FALSE}
 ## Compilation dependencies
 library("knitr")
-library("kableExtra")
+#library("kableExtra")
 library("magrittr")
@@
 ```
+
 <!-- bookdown::pdf_document2: allows \@ref(), pdf_document: requires \ref{} -->

### 2023-05-08 jss/cheem_paper.tex
JSS formatted paper

--- 2023-05-05 paper.Rmd
@@
----
-title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
-type: ARTICLE TEMPLATE
-author:
-  - name: Nicholas Spyrison
-    affiliation: a
-    orcid: 0000-0002-8417-0212
-    email: spyrison@gmail.com
-  - name: Dianne Cook
-    affiliation: a
-    orcid: 0000-0002-3813-7155
-    email: dicook@monash.edu
-  - name: Przemyslaw Biecek
-    affiliation: b
-    orcid: 0000-0001-8423-1823
-    email: przemyslaw.biecek@gmail.com
-affiliation:
-  - num: a
-    address: |
-      Monash University, Melbourne, Australia
-  - num: b
-    address: |
-      Warsaw University of Technology and University, Warsaw, Poland
-abstract: |
-  The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
-This trade-off has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction.
-"Local explanations", which provide a point estimate of the linear variable importance in the vicinity of one observation, are one XAI method.
-These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface.
-Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
-The methods are implemented in the __R__ package __cheem__, available on CRAN.
-keywords: | 
-  explainable artificial intelligence; nonlinear model interpretability; visual analytics; local explanations; grand tour; radial tour
-bibliography: paper.bib
-header-includes: |
-  \usepackage{hyperref}
+\documentclass[
+]{jss}
+
+%% recommended packages
+\usepackage{orcidlink,thumbpdf,lmodern}
+
 \usepackage[utf8]{inputenc}
-  \def\tightlist{}
-output: rticles::tf_article
----
 
-```{r include=FALSE, cache=FALSE}
-## Compilation dependencies
-library("knitr")
-library("kableExtra")
-library("magrittr")
+\author{
+Nicholas Spyrison~\orcidlink{0000-0002-8417-0212}\\Monash
+University \And Dianne Cook~\orcidlink{0000-0002-3813-7155}\\Monash
+University \And Przemyslaw
+Biecek~\orcidlink{0000-0001-8423-1823}\\Warsaw University of\\
+Technology and University
+}
+\title{Exploring Local Explanations of Nonlinear Models Using Animated
+Linear Projections}
 
-## chunk options
-knitr::opts_chunk$set(
-  fig.align  = "center",
-  echo       = FALSE,
-  collapse   = TRUE,
-  message    = FALSE,
-  warning    = FALSE,
-  error      = FALSE,
-  cache      = FALSE,
-  cache.lazy = FALSE,
-  fig.height = 5,
-  fig.width = 8,
-  out.width = "100%"
-)
-```
-<!-- bookdown::pdf_document2: allows \@ref(), pdf_document: requires \ref{} -->
-<!-- # cheat sheet {#sec:cheatsheet} -->
-<!-- A bib reference [@wickham_visualizing_2015]. -->
-<!-- A [Section intro](#sec:cheatsheet) reference, alternatively, section \@ref(sec:intro) (with @; \\@ref(sec:intro)). -->
-<!-- ```{r crest, echo=FALSE, out.height = "10%", out.width = "10%", fig.cap = "A caption for crest figure"} -->
-<!-- knitr::include_graphics("./figures/crest.jpg") -->
-<!-- ``` -->
-<!-- A figure \@ref(fig:crest) reference (with @; \\@ref(fig:crest). -->
-<!-- (ref:baseplotcap) Separate caption created above the __R__ chunk -->
-<!-- ```{r baseplot, echo=F, fig.cap = "(ref:baseplotcap)"} -->
-<!-- plot(1,2) -->
-<!-- ``` -->
+\Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}
+\Plaintitle{Exploring Local Explanations of Nonlinear Models Using
+Animated Linear Projections}
+\Shorttitle{Exploring Local Explanations}
 
-# Introduction {#sec:intro}
+\Abstract{
+The increased predictive power of nonlinear models comes at the cost of
+interpretability of its terms. This trade-off has led to the emergence
+of eXplainable AI (XAI). XAI attempts to shed light on how models use
+predictors to arrive at a prediction. ``Local explanations'', which
+provide a point estimate of the linear variable importance in the
+vicinity of one observation, are one XAI method. These can be considered
+to be linear projections and can be further explored to understand the
+interactions between variables used to make predictions across the
+predictive model surface. Here we describe interactive linear
+interpolation used for exploration at any observation and illustrate
+with examples from categorical (penguin species, chocolate types) and
+quantitative (soccer/football salaries, house prices) response models.
+The methods are implemented in the R package cheem, available on CRAN.
+}
 
-<!-- Introduce explanatory vs predictive modeling -->
-There are different reasons and purposes for fitting a model.
-According to the taxonomies of @breiman_statistical_2001 and @shmueli_explain_2010, it can be useful to group models into two types: explanatory and predictive.
-Explanatory modeling is used for inferential purposes, while predictive modeling focuses solely on the performance of an objective function.
-The intended use of the model has important implications for its selection and development.
-Interpretability is critical in explanatory modeling to draw meaningful inferential conclusions, such as which variables most contribute to a prediction or whether some observations are less well fit.
+\Keywords{explainable artificial intelligence, nonlinear model
+interpretability, visual analytics, local explanations, grand
+tour, radial tour}
+\Plainkeywords{explainable artificial intelligence, nonlinear model
+interpretability, visual analytics, local explanations, grand
+tour, radial tour}
+
+%% publication information
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{}
+%% \Acceptdate{2012-06-04}
+
+\Address{
+    Nicholas Spyrison\\
+    Monash University\\
+    Dept of Human Centred Computing\\
+Faculty of Information Technology\\
+  E-mail: \email{spyrison@gmail.com}\\
+  URL: \url{https://nspyrison.netlify.app}\\~\\
+      Dianne Cook\\
+    Monash University\\
+    Dept of Econometrics and Business Statistics\\
+
+    }
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage{amsmath}
+
+\begin{document}
+
+\hypertarget{sec:intro}{%
+\section{Introduction}\label{sec:intro}}
+
+There are different reasons and purposes for fitting a model. According
+to the taxonomies of \citet{breiman_statistical_2001} and
+\citet{shmueli_explain_2010}, it can be useful to group models into two
+types: explanatory and predictive. Explanatory modeling is used for
+inferential purposes, while predictive modeling focuses solely on the
+performance of an objective function. The intended use of the model has
+important implications for its selection and development.
+Interpretability is critical in explanatory modeling to draw meaningful
+inferential conclusions, such as which variables most contribute to a
+prediction or whether some observations are less well fit.
 Interpretability becomes more difficult when the model is nonlinear.
-Nonlinear models occur in statistical models with polynomial or interaction terms between quantitative predictors, and almost all computational models such as random forests, support-vector machines, or neural networks [e.g.
-@breiman_random_2001; @boser_training_1992; @anderson_introduction_1995].
+Nonlinear models occur in statistical models with polynomial or
+interaction terms between quantitative predictors, and almost all
+computational models such as random forests, support-vector machines, or
+neural networks
+\citep[e.g.][]{breiman_random_2001, boser_training_1992, anderson_introduction_1995}.
 
-<!-- nonlinear leading to interpretability issues -->
-In linear models interpretation of the importance of variables is relatively straight forward, one adjusts for the covariance of multiple variables when examining the relationship with the response.
-The interpretation is valid for the full domain of the predictors.
-In nonlinear models one needs to consider the model in small neighborhoods of the domain to make any assessment of variable importance.
-Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues described in @stahl-ethics.
-Understanding how nonlinear models behave when usage extrapolates outside the domain of predictors, either in sub-spaces where few samples were provided in the training set, or extending outside the domain.
-It is especially important because nonlinear models can vary wildly and predictions can be dramatically wrong in these areas.
+In linear models interpretation of the importance of variables is
+relatively straight forward, one adjusts for the covariance of multiple
+variables when examining the relationship with the response. The
+interpretation is valid for the full domain of the predictors. In
+nonlinear models one needs to consider the model in small neighborhoods
+of the domain to make any assessment of variable importance. Even though
+this is difficult, it is especially important to interpret model fits as
+we become more dependent on nonlinear models for routine aspects of life
+to avoid issues described in \citet{stahl-ethics}. Understanding how
+nonlinear models behave when usage extrapolates outside the domain of
+predictors, either in sub-spaces where few samples were provided in the
+training set, or extending outside the domain. It is especially
+important because nonlinear models can vary wildly and predictions can
+be dramatically wrong in these areas.
 
-<!-- Interpretability & baises 
-Interpretability is vital for exploring and protecting against potential biases in any model, e.g., sex\-\-\-@dastin_amazon_2018; @duffy_apple_2019, race\-\-\-@larson_how_2016, and age\-\-\-@diaz_addressing_2018.
-For observation, models regularly pick up on biases in the training data where such classes correlate with changes in the response variable.
-This bias is then built into the model.
-Variable-level interpretability of models is essential in evaluating and addressing such biases.
--->
+Explainable Artificial Intelligence (XAI) is an emerging field of
+research focused on methods for the interpreting of models
+\citep{adadi_peeking_2018, arrieta_explainable_2020}. A class of
+techniques called \emph{local explanations}, provide methods to
+approximate linear variable importance at the location of each
+observation or the predictions at a specific point in the data domain.
+Because these are point-specific, it is challenging to comprehensively
+visualize them to understand a model. There are common approaches for
+visualizing high-dimensional data as a whole, but what is needed are new
+approaches for viewing these individual local explanations, in relation
+to the whole.
 
-<!-- Interpretability & data drift 
-Another concern is data drift, where a shift in the range of the explanatory variables (features or predictors) between training and test sets.
-Some nonlinear models are sensitive to this and do not extrapolate well outside the support of the training data [@espadoto_toward_2021].
-Maintaining variable interpretability is also essential to address issues arising from data drift.
--->
+For multivariate data visualization, a \emph{tour}
+\citep{asimov_grand_1985, buja_grand_1986, lee_state_2021} of linear
+data projections onto a lower-dimensional space, could be an element of
+XAI, complementing local explanations. Applying tours to model
+interpretation is recommended in \citet{wickham_visualizing_2015}
+primarily to examine the fitted model in the space of the data.
+\citet{cook_interactive_2007} describe the use of tours for exploring
+classification boundaries and model diagnostics
+\citep{Caragea2008, lee_pptree_2013, da_silva_projection_2021}. There
+are various types of tours. In a \emph{manual} or radial tour
+\citep{cook_manual_1997, spyrison_spinifex_2020}, the path of linear
+projections is defined by changing the contribution of a selected
+variable. We propose to use this to scrutinize a local explanation. This
+approach could be considered to be a counter-factual, what-if analysis,
+such as \emph{ceteris paribus} (``other things held constant'') profiles
+\citep{biecek_ceterisparibus_2020}.
 
-<!-- Local explanations -->
-Explainable Artificial Intelligence (XAI) is an emerging field of research focused on methods for the interpreting of models [@adadi_peeking_2018; @arrieta_explainable_2020].
-A class of techniques called _local explanations_, provide methods to approximate linear variable importance at the location of each observation or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
-There are common approaches for visualizing high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
+The remainder of this paper is organized as follows. Section
+\ref{sec:explanations} covers the background of the local explanation
+and the traditional visuals produced. Section \ref{sec:tour} explains
+the tours and particularly the radial manual tour. Section
+\ref{sec:cheemviewer} discusses the visual layout in the graphical user
+interface and how it facilitates analysis, data pre-processing, and
+package infrastructure. Illustrations are provided in Section
+\ref{sec:casestudies} for a range of supervised learning tasks with
+categorical and quantitative response variables. These show how the
+local explanations can be used to get an overview of the model's use of
+predictors, and to investigate errors in the model predictions. Section
+\ref{sec:cheemdiscussion} concludes with a summary of the insights
+gained. The methods are implemented in the \textbf{R} package
+\textbf{cheem}.
 
-<!-- Data visualization tours -->
-For multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] of linear data projections onto a lower-dimensional space, could be an element of XAI, complementing local explanations.
-<!--Tours are viewed as an animation over minor changes to the projection basis.
-Structure in a projection can then be explored visually to see which variables contribute to the formation of that structure.
-The intuition is similar to watching the shadow of a hidden 3D object change as the object is rotated; watching the shape of the shadow change conveys information about the structure and variables of the object.-->
-<!-- tours and models -->
-Applying tours to model interpretation is recommended in @wickham_visualizing_2015 primarily to examine the fitted model in the space of the data.
-@cook_interactive_2007 describe the use of tours for exploring classification boundaries and model diagnostics [@Caragea2008; @lee_pptree_2013; @da_silva_projection_2021].
-<!-- Manual tours --> 
-There are various types of tours.
-In a _manual_ or radial tour [@cook_manual_1997; @spyrison_spinifex_2020], the path of linear projections is defined by changing the contribution of a selected variable.
-We propose to use this to scrutinize a local explanation.
-<!-- Additional interactive elements in a graphical user interface should allow the user to identify an observation of interest and then explore its local explanation by changing variable contribution with the radial tour.
--->
-<!-- What-if & counterfactual analysis -->
-This approach could be considered to be a counter-factual, what-if analysis, such as _ceteris paribus_ ("other things held constant") profiles [@biecek_ceterisparibus_2020].
-<!-- These profiles show how an observation's prediction would change from a marginal change in one explanatory variable, given that other variables are held constant.
-It ignores correlations of the variables and imagines a case that was not observed.
-In contrast, our approach is a geometric explanation of the factual; it varies contributions of the variables by rotating the basis, a reorientation of the data object.
-A constraint in our approach is that the basis must remain orthonormal.
-When the contribution of one variable decreases, the contributions of others necessarily increase such that there is a complete component in that direction.
-This also ensures that what is seen is strictly a low-dimensional projection from high-dimensional space and is thus an interpretable visualization.-->
+\hypertarget{sec:explanations}{%
+\section{Local Explanations}\label{sec:explanations}}
 
-<!-- Paper structure -->
-The remainder of this paper is organized as follows.
-Section \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
-Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
-Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data pre-processing, and package infrastructure.
-Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative response variables.
-These show how the local explanations can be used to get an overview of the model's use of predictors, and to investigate errors in the model predictions.
-Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
-The methods are implemented in the __R__ package __cheem__.
+Local explanations shed light on nonlinear model fits by estimating
+linear variable importance in the vicinity of a single observation.
+There are many approaches for calculating local explanations. A
+comprehensive summary of the taxonomy of currently available methods is
+provided in Figure 6 by \citet{arrieta_explainable_2020}. It includes a
+large number of model-specific explanations such as deepLIFT
+\citep{shrikumar_not_2016, shrikumar_learning_2017}, a popular recursive
+method for estimating importance in neural networks. There are fewer
+model-agnostic methods, of which LIME, \citep{ribeiro_why_2016} SHaply
+Additive exPlanations (SHAP), \citep{lundberg_unified_2017}, are
+popular.
 
-# Local Explanations {#sec:explanations}
+These observation-level explanations are used in various ways depending
+on the data. In image classification, where pixels correspond to
+predictors, saliency maps overlay or offset a heatmap to indicate
+important pixels \citep{simonyan_deep_2014}. For example, pixels
+corresponding to snow may be highlighted as important contributors when
+distinguishing if a picture contains a coyote or husky. In text
+analysis, word-level contextual sentiment analysis highlights the
+sentiment and magnitude of influential words \citep{vanni_textual_2018}.
+In the case of numeric regression, they are used to explain additive
+contributions of variables from the model intercept to the observation's
+prediction \citep{ribeiro_why_2016}.
 
-<!-- Reminder of local explanation -->
-Local explanations shed light on nonlinear model fits by estimating linear variable importance in the vicinity of a single observation.  There are many approaches for calculating local explanations.
-<!-- Taxonomy of local explanations -->
-A comprehensive summary of the taxonomy of currently available methods is provided in Figure 6 by @arrieta_explainable_2020.
-It includes a large number of model-specific explanations such as deepLIFT [@shrikumar_not_2016; @shrikumar_learning_2017], a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic methods, of which LIME, [@ribeiro_why_2016] SHaply Additive exPlanations (SHAP), [@lundberg_unified_2017], are popular.
+We will be focusing on SHAP values in this this paper, but the approach
+is applicable for any method used to calculate the local explanations.
+SHAP calculates the variable contributions of one observation by
+examining the effect of other variables on the predictions. The term
+``SHAP'' refers to \citet{shapley_value_1953}'s method to evaluate an
+individual's contribution in cooperative games by assessing this
+player's performance in the presence or absence of other players.
+\citet{strumbelj_efficient_2010} introduced SHAP for local explanations
+in machine learning models. Variable importance can depend on the
+sequence in which variables are entered into the model fitting process,
+thus for any sequence we get a set of variable contribution values for a
+single observation. These values will add up to the difference between
+the fitted value for the observation, and the average fitted value for
+all observations. Using all possible sequence, or permutation, gives
+multiple values for each variable, which are averaged to get the SHAP
+value for an observation. It can be helpful to standardize variables
+prior to computing SHAP values, if they have been measured on different
+scales.
 
-<!-- Uses of local explanations -->
-These observation-level explanations are used in various ways depending on the data.
-In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap to indicate important pixels [@simonyan_deep_2014].
-For example, pixels corresponding to snow may be highlighted as important contributors when distinguishing if a picture contains a coyote or husky.
-In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words [@vanni_textual_2018].
-In the case of numeric regression, they are used to explain additive contributions of variables from the model intercept to the observation's prediction [@ribeiro_why_2016].
+The approach is related to partial dependence plots (see for example
+\citet{molnar_interpretable_2020}), used to explain the effect of a
+variable by predicting the response for a range of values on this
+variable after fixing the value of all other variables to their mean.
+Though partial dependence plots are a global approximation of the
+variable importance, while SHAP is specific to one observation.
 
-<!-- SHAP -->
-We will be focusing on SHAP values in this this paper, but the approach is applicable for any method used to calculate the local explanations.
-SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
-The term "SHAP" refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
-@strumbelj_efficient_2010 introduced SHAP for local explanations in machine learning models.
-Variable importance can depend on the sequence in which variables are entered into the model fitting process, thus for any sequence we get a set of variable contribution values for a single observation.
-These values will add up to the difference between the fitted value for the observation, and the average fitted value for all observations.
-Using all possible sequence, or permutation, gives multiple values for each variable, which are averaged to get the SHAP value for an observation.
-It can be helpful to standardize variables prior to computing SHAP values, if they have been measured on different scales.
+\begin{CodeChunk}
+\begin{figure}
 
-The approach is related to partial dependence plots (see for example @molnar_interpretable_2020), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
-Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
+{\centering \includegraphics[width=0.85\linewidth]{./figures/shap_distr_bd}
 
-```{r shapdistrbd, echo=F, fig.align='center', out.width="85%", fig.cap = "Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
+}
+
+\caption[Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors]{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
 A star offensive and defensive player are compared, L.
@@
 Reaction skills are important for both players.
-Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
-knitr::include_graphics("./figures/shap_distr_bd.png")
-```
+Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
+\end{figure}
+\end{CodeChunk}
 
-<!-- Fifa example -->
-We use 2020 season FIFA data [@leone_fifa_2020] to illustrate SHAP following the procedures described in @biecek_explanatory_2021.
-There are 5000 observations of nine predictor variables measuring players' skills and one response variable, wages (in euros).
-A random forest model is fit regressing players' wages on the skill variables.
-In this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are compared for a star offensive player (L.
-Messi) and a prominent defensive player (V.
-van Dijk).
-We are interested in knowing how the skill variables locally contribute to the wage prediction of each player.
-A difference in the attribution of the variable importance across the two positions of the players can be expected.
-This would be interpreted as how a player's salary depends on which combination of skills.
-Panel (a) is a version of a breakdown plot [@gosiewska_ibreakdown_2019] where just three sequences of variables are shown, for two observations.
-A breakdown plot shows the absolute values of the variable attribution for an observation, usually sorted from highest value to the lowest.
-There is no scale on the horizontal axis here because values are considered relative to each other.
-Here we can see how the variable contribution can change depending on sequence, relative to both players.
-(Note that the order of the variables is different in each plot, because they have been sorted by biggest average contribution across both players.) For all sequences, and for both players `reaction` has the strongest contribution, with perhaps more importance for the defensive player.
-Then it differs by player: for Messi `offense` and `movement` have the strongest contributions, and for van Dijk it is `defense` and `power`, regardless of the variable sequence.
+We use 2020 season FIFA data \citep{leone_fifa_2020} to illustrate SHAP
+following the procedures described in \citet{biecek_explanatory_2021}.
+There are 5000 observations of nine predictor variables measuring
+players' skills and one response variable, wages (in euros). A random
+forest model is fit regressing players' wages on the skill variables. In
+this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are
+compared for a star offensive player (L. Messi) and a prominent
+defensive player (V. van Dijk). We are interested in knowing how the
+skill variables locally contribute to the wage prediction of each
+player. A difference in the attribution of the variable importance
+across the two positions of the players can be expected. This would be
+interpreted as how a player's salary depends on which combination of
+skills. Panel (a) is a version of a breakdown plot
+\citep{gosiewska_ibreakdown_2019} where just three sequences of
+variables are shown, for two observations. A breakdown plot shows the
+absolute values of the variable attribution for an observation, usually
+sorted from highest value to the lowest. There is no scale on the
+horizontal axis here because values are considered relative to each
+other. Here we can see how the variable contribution can change
+depending on sequence, relative to both players. (Note that the order of
+the variables is different in each plot, because they have been sorted
+by biggest average contribution across both players.) For all sequences,
+and for both players \texttt{reaction} has the strongest contribution,
+with perhaps more importance for the defensive player. Then it differs
+by player: for Messi \texttt{offense} and \texttt{movement} have the
+strongest contributions, and for van Dijk it is \texttt{defense} and
+\texttt{power}, regardless of the variable sequence.
 
-Panel (b) shows the differences in the player's median values (large dots) for 25 such sequences (tick marks).
-We can see that the wage predictions for the two players come from different combinations of skills sets, as might be expected for players who's value on the team depends on their offensive or defensive prowess.
-It is also interesting to see from the distribution of values across the different sequence of variables, that there is some multimodality.
-For example, look at the SHAP values for `reaction` for Messi, and in some sequences reaction has a much lower contribution than others.
-This suggests that other variables (`offense`, `movement` probably) can substitute for `reaction` in the wage prediction.
+Panel (b) shows the differences in the player's median values (large
+dots) for 25 such sequences (tick marks). We can see that the wage
+predictions for the two players come from different combinations of
+skills sets, as might be expected for players who's value on the team
+depends on their offensive or defensive prowess. It is also interesting
+to see from the distribution of values across the different sequence of
+variables, that there is some multimodality. For example, look at the
+SHAP values for \texttt{reaction} for Messi, and in some sequences
+reaction has a much lower contribution than others. This suggests that
+other variables (\texttt{offense}, \texttt{movement} probably) can
+substitute for \texttt{reaction} in the wage prediction.
 
-This can also be considered similar to examining the coefficients from all subsets regression, as described in @wickham_visualizing_2015.
-Various models that are similarly good might use different combinations of the variables.
-Examining the coefficients from multiple models helps to understand the relative importance of each variable in the context of all other variables.
-This is similar to the approach here with SHAP values, that by examining the variation in values across different permutations of variables, we can gain more understanding of the relationship between the response and predictors.
+This can also be considered similar to examining the coefficients from
+all subsets regression, as described in
+\citet{wickham_visualizing_2015}. Various models that are similarly good
+might use different combinations of the variables. Examining the
+coefficients from multiple models helps to understand the relative
+importance of each variable in the context of all other variables. This
+is similar to the approach here with SHAP values, that by examining the
+variation in values across different permutations of variables, we can
+gain more understanding of the relationship between the response and
+predictors.
 
-<!--In summary, these plots highlight how local explanations bring interpretability to a model, at least in the vicinity of their observations.
-In this observation, two players with different positions receive different profiles of variable importance to explain the prediction of their wages.-->
+For the application, we use \emph{tree SHAP}, a variant of SHAP that
+enjoys a lower computational complexity
+\citep{lundberg_consistent_2018}. Instead of aggregating over sequences
+of the variables, tree SHAP calculates observation-level variable
+importance by exploring the structure of the decision trees. Tree SHAP
+is only compatible with tree-based models. so random forests are used
+for illustration.
 
-<!-- tree SHAP -->
-For the application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018].
-Instead of aggregating over sequences of the variables, tree SHAP calculates observation-level variable importance by exploring the structure of the decision trees.
-Tree SHAP is only compatible with tree-based models.
-so random forests are used for illustration.
-<!--The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) that will be used to explore the sensitivity of the variable contributions.-->
+\hypertarget{sec:tour}{%
+\section{Tours and the Radial Tour}\label{sec:tour}}
 
-# Tours and the Radial Tour {#sec:tour}
+A \emph{tour} enables the viewing of high-dimensional data by animating
+many linear projections with small incremental changes. It is achieved
+by following a path of linear projections (bases) of high-dimensional
+space. One key variable of the tour is the object permanence of the data
+points; one can track the relative change of observations in time and
+gain information about the relationships between points across multiple
+variables. There are various types of tours that are distinguished by
+how the paths are generated \citep{lee_state_2021, cook_grand_2008}.
 
-<!-- Tours intro -->
-A _tour_ enables the  viewing of high-dimensional data by animating many linear projections with small incremental changes.
-It is achieved by following a path of linear projections (bases) of high-dimensional space.
-One key variable of the tour is the object permanence of the data points; one can track the relative change of observations in time and gain information about the relationships between points across multiple variables.
-There are various types of tours that are distinguished by how the paths are generated [@lee_state_2021; @cook_grand_2008].
+The manual tour \citep{cook_manual_1997} defines its path by changing a
+selected variable's contribution to a basis to allow the variable to
+contribute more or less to the projection. The requirement constrains
+the contribution of all other variables that a basis needs to be
+orthonormal (column correspond to vectors, with unit length, and
+orthogonal to each other). The manual tour is primarily used to assess
+the importance of a variable to structure visible in a projection. It
+also lends itself to pre-computation queued in advance or computed
+on-the-fly for human-in-the-loop analysis
+\citep{karwowski_international_2006}.
 
-<!-- Manual tour -->
-The manual tour [@cook_manual_1997] defines its path by changing a selected variable's contribution to a basis to allow the variable to contribute more or less to the projection.
-The requirement constrains the contribution of all other variables that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
-It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
+A version of the manual tour called a \emph{radial tour} is implemented
+in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
+In a radial tour, the selected variable can change its magnitude of
+contribution but not its angle; it must move along the direction of its
+original contribution. The implementation allows for pre-computation and
+interactive re-calculation to focus on a different variable.
 
-<!-- Radial tour -->
-A version of the manual tour called a _radial tour_ is implemented in @spyrison_spinifex_2020 and forms the basis of this new work.
-In a radial tour, the selected variable can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
-The implementation allows for pre-computation and interactive re-calculation to focus on a different variable.
+\begin{CodeChunk}
+\begin{figure}
 
-```{r radialtour, echo=F, fig.align='center', out.width="99%", fig.cap = "The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to structure in the plot.
+{\centering \includegraphics[width=0.99\linewidth]{./figures/radial_tour}
+
+}
+
+\caption{The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to structure in the plot.
 Here we have a 1D projection of the penguins data displayed as a density plot.
 The line segments on the bottom correspond to the coefficients of the variables making up the projection.
-The structure in the plot is bimodality (left), and the importance of the variable \\textsf{bd} is being explored.
+The structure in the plot is bimodality (left), and the importance of the variable \textsf{bd} is being explored.
 As this variable contribution is reduced in the plot (middle, right) we can see that the bimodality decreases.
-Thus \\textsf{bd} is an important variable contributing to the bimodal structure."}
-knitr::include_graphics("./figures/radial_tour.png")
-```
+Thus \textsf{bd} is an important variable contributing to the bimodal structure.}\label{fig:radialtour}
+\end{figure}
+\end{CodeChunk}
 
-<!--
-# Shapley's Lenses {#sec:shapleylenses}
+\hypertarget{sec:cheemviewer}{%
+\section{The Cheem Viewer}\label{sec:cheemviewer}}
 
-Shapley values are a useful technique for understanding the nature of the relationship between the input variables and the target variable.
-The data exploration process is navigated by a trained predictive model.
+To explore the local explanations, coordinated views
+\citep{roberts_state_2007} \citep[also known as ensemble
+graphics,][]{unwin_ensemble_2018} are provided in the \emph{cheem
+viewer} application. There are two primary plots: the \textbf{global
+view} to give the context of all of the SHAP values and the
+\textbf{radial tour view} to explore the local explanations with
+user-controlled rotation. There are numerous user inputs, including
+variable selection for the radial tour and observation selection for
+making comparisons. There are different plots used for the categorical
+and quantitative responses. Figures \ref{fig:classificationcase} and
+\ref{fig:regressioncase} are screenshots showing the cheem viewer for
+the two primary tasks: classification (categorical response) and
+regression (quantitative response).
 
-Thus, local attributions can be viewed as a new representation of data (commonly referred as data embeddings by machine learning community) with certain desired properties, e.g., the unit is standardized between variables because it reflects the relation with target variable, so it is expressed on the scale of the target variable.
+\hypertarget{global-view}{%
+\subsection{Global View}\label{global-view}}
 
-But different models can produce different local attributions.
-This phenomenon, known as the multiplicity of good models or as the Rashomon set of equally good models, means that models that fit the data equally well can describe the relationships present in the data in different ways.
+The global view provides context for all observations and facilitates
+the exploration of the separability of the data and attribution spaces.
+The attribution space refers to the SHAP values for each observation.
+These spaces both have dimensionality \(n \times p\), where \(n\) is the
+number of observations and \(p\) is the number of variables.
 
-The response to the challenge in the characterization of Rashomon sets may be Shapley's Lenses - the opportunity to compare these models from the perspective of exploring the local attributions generated by the Shapley values for different models.
+The visualization is composed of the first two principal components of
+the data (left) and the attribution (middle) spaces. These single 2D
+projections will not reveal all of the structure of higher-dimensional
+space, but they are helpful visual summaries. In addition, a plot of the
+observed against predicted response values is also provided (Figures
+\ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help
+identify observations poorly predicted by the model. For classification
+tasks, color indicates the predicted class, and misclassified
+observations are circled in red. Linked brushing between the plots is
+provided, and a tabular display of selected points helps to facilitate
+exploration of the spaces and the model (shown in Figures
+\ref{fig:regressioncase}d).
 
--->
+While the comparison of these spaces is interesting, the primary purpose
+of the global view is to enable the selection of particular observations
+to explore in detail. We have designed it to enable a comparison between
+an observation that is interesting is some way, perhaps misclassified,
+or poorly predicted, relative to an observation with similar predictor
+values but more expected prediction. For brevity, we call the
+interesting observation the primary investigation (PI) and the other is
+the comparison investigation (CI). These observations are highlighted as
+asterisk and \(\times\), respectively.
 
-# The Cheem Viewer {#sec:cheemviewer}
+\hypertarget{radial-tour}{%
+\subsection{Radial Tour}\label{radial-tour}}
 
-To explore the local explanations, coordinated views [@roberts_state_2007] [also known as ensemble graphics, @unwin_ensemble_2018] are provided in the _cheem viewer_ application.
-There are two primary plots: the __global view__ to give the context of all of the SHAP values and the __radial tour view__ to explore the local explanations with user-controlled rotation.
-There are numerous user inputs, including variable selection for the radial tour and observation selection for making comparisons.
-There are different plots used for the categorical and quantitative responses.
-Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
+There are two plots in this part of the interface. The first (Figures
+\ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a
+display of the SHAP values for all observations. This will generally
+give the global view of variables important for the fit as a whole, but
+it will also highlight observations which have different patterns. The
+second plot is the radial tour, which for classification is a density
+plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for
+regression are scatterplots of the observed response values, and
+residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
 
-## Global View
+The local explanations for all observations are normalized (sum of
+squares equals 1), and thus, the relative importance of variables can be
+compared across all observations. These are depicted as a vertical
+parallel coordinate plot \citep{ocagne_coordonnees_1885}. (The SHAP
+values of the PI and CI are shown as dashed and dotted lines,
+respectively.) One should obtain a sense of the overall importance of
+variables from this plot. The more important variables will have larger
+values, and in the case of classification tasks variables which have
+different magnitude for different classes are more globally important.
+For example, Figure \ref{fig:classificationcase}e suggests that
+\texttt{bl} is important for distinguishing the green class from the
+other two. For regression, one might generally observe which variables
+have low values for all observations (not important). For example,
+\texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have
+a range of high and low values (e.g., \texttt{off}, \texttt{def}),
+suggesting they are important for some observations and not important
+for others.
 
-<!-- Purpose -->
-The global view provides context for all observations and facilitates the exploration of the separability of the data and attribution spaces.
-The attribution space refers to the SHAP values for each observation.
-These spaces both have dimensionality $n \times p$, where $n$ is the number of observations and $p$ is the number of variables.
+A bar chart is overlaid to represent the projection shown in the radial
+tour at right. It starts from the SHAP values of the PI, but if the user
+changes the projection the length of these bars will reflect this
+change. (The PI is interactively selected by clicking on a point in the
+global view). By scaling the SHAP value it becomes an (attribution)
+projection.
 
-<!-- Approximations of the spaces, PC1:2 -->
-The visualization is composed of the first two principal components of the data (left) and the attribution (middle) spaces.
-These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
-In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify observations poorly predicted by the model.
-<!-- Interactions --> For classification tasks, color indicates the predicted class, and misclassified observations are circled in red.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
+The attribution projection of the PI is the initial 1D basis in a radial
+tour, displayed as a density plot for a categorical response (Figure
+\ref{fig:classificationcase}f) and as scatterplots for a quantitative
+response (Figure \ref{fig:regressioncase}f). The PI and CI are indicated
+by vertical dashed and dotted lines, respectively. The radial tour
+varies the contribution of the selected variable. This is viewed as an
+animation of the projections from many intermediate bases. Doing so
+tests the sensitivity of structure (class separation or strength of
+relationship) to the variable's contribution. For classification, if the
+separation between classes diminishes when the variable contribution is
+reduced, this suggests that the variable is important for class
+separation. For regression, if the relationship scatterplot weakens when
+the variable contribution is reduced, indicating that the variable is
+important for accurately predicting the response.
 
-While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular observations to explore in detail.
-We have designed it to enable a comparison between an observation that is interesting is some way, perhaps misclassified, or poorly predicted, relative to an observation with similar predictor values but more expected prediction.
-For brevity, we call the interesting observation the primary investigation (PI) and the other is the comparison investigation (CI).
-These observations are highlighted as asterisk and $\times$, respectively.
+\hypertarget{classification-task}{%
+\subsection{Classification Task}\label{classification-task}}
 
-## Radial Tour
+Selecting a misclassified observation as PI and a correctly classified
+point nearby in data space as CI makes it easier to examine the
+variables most responsible for the error. The global view (Figure
+\ref{fig:classificationcase}c) displays the model confusion matrix. The
+radial tour is 1D and displays as density where color indicates class.
+An animation slider enables users to vary the contribution of variables
+to explore the sensitivity of the separation to that variable.
 
-There are two plots in this part of the interface.
-The first (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a display of the SHAP values for all observations.
-This will generally give the global view of variables important for the fit as a whole, but it will also highlight observations which have different patterns.
-The second plot is the radial tour, which for classification is a density plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for regression are scatterplots of the observed response values, and residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
+\begin{CodeChunk}
+\begin{figure}
 
-<!-- Local explanations as par coords -->
-The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
-These are depicted as a vertical parallel coordinate plot [@ocagne_coordonnees_1885].
-(The SHAP values of the PI and CI are shown as dashed and dotted lines, respectively.) One should obtain a sense of the overall importance of variables from this plot.
-The more important variables will have larger values, and in the case of classification tasks variables which have different magnitude for different classes are more globally important.
-For example, Figure \ref{fig:classificationcase}e suggests that `bl` is important for distinguishing the green class from the other two.
-For regression, one might generally observe which variables have low values for all observations (not important).
-For example, `BMI` and `pwr` in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., `off`, `def`), suggesting they are important for some observations and not important for others.
+{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}
 
-<!-- Segue of PI to attribution projection -->
-A bar chart is overlaid to represent the projection shown in the radial tour at right.
-It starts from the SHAP values of the PI, but if the user changes the projection the length of these bars will reflect this change.
-(The PI is interactively selected by clicking on a point in the global view).
-By scaling the SHAP value it becomes an (attribution) projection.
-<!-- approximates the variable importance for predicting this observation.
-The combination of variables best explains the difference between the mean response and an observation’s predicted value.
-It is not an indication of the local shape of the model surface.
-That is, it is not some indication of the tangent to the curve at this point.-->
+}
 
-<!-- Tour animation -->
-The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
-The PI and CI are indicated by vertical dashed and dotted lines, respectively.
-The radial tour varies the contribution of the selected variable.
-This is viewed as an animation of the projections from many intermediate bases.
-Doing so tests the sensitivity of structure (class separation or strength of relationship) to the variable's contribution.
-For classification, if the separation between classes diminishes when the variable contribution is reduced, this suggests that the variable is important for class separation.
-For regression, if the relationship scatterplot weakens when the variable contribution is reduced, indicating that the variable is important for accurately predicting the response.
-
-## Classification Task
-
-Selecting a misclassified observation as PI and a correctly classified point nearby in data space as CI makes it easier to examine the variables most responsible for the error.
-The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
-The radial tour is 1D and displays as density where color indicates class.
-An animation slider enables users to vary the contribution of variables to explore the sensitivity of the separation to that variable.
-
-```{r classificationcase, echo=F, out.width = "100%", fig.align="center", fig.cap = "Overview of the cheem viewer for classification tasks (categorical response).
+\caption[Overview of the cheem viewer for classification tasks (categorical response)]{Overview of the cheem viewer for classification tasks (categorical response).
 Global view inputs, (a), set the PI, CI, and color statistic.
@@
 The black bar is the variable being changed in the radial tour.
-Panel (f) is the resulting data projection indicated as density in the classification case."}
-knitr::include_graphics("./figures/app_classification.PNG")
-```
+Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
+\end{figure}
+\end{CodeChunk}
 
-## Regression Task
+\hypertarget{regression-task}{%
+\subsection{Regression Task}\label{regression-task}}
 
-Selecting an inaccurately predicted observation as PI and an accurately predicted observation with similar variable values as CI is a helpful way to understand how the model is failing or not.
-The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
-The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation to aid in understanding the structure identified in these spaces.
+Selecting an inaccurately predicted observation as PI and an accurately
+predicted observation with similar variable values as CI is a helpful
+way to understand how the model is failing or not. The global view
+(Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed
+vs predicted values, which should exhibit a strong relationship if the
+model is a good fit. The points can be colored by a statistic, residual,
+a measure of outlyingness (log Mahalanobis distance), or correlation to
+aid in understanding the structure identified in these spaces.
 
-In the radial tour view, the observed response and the residuals (vertical) are plotted against the attribution projection of the PI (horizontal).
-The attribution projection can be interpreted similarly to the predicted value from the global view plot.
-It represents a linear combination of the variables, and a good fit would be indicated when there is a strong relationship with the observed values.
-This can be viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a variable is varied, if the value of the PI does not change much, it would indicate that the prediction for this observation is NOT sensitive to that variable.
-Conversely, if the predicted value varies substantially, the prediction is very sensitive to that variable, suggesting that the variable is very important for the PI's prediction.
+In the radial tour view, the observed response and the residuals
+(vertical) are plotted against the attribution projection of the PI
+(horizontal). The attribution projection can be interpreted similarly to
+the predicted value from the global view plot. It represents a linear
+combination of the variables, and a good fit would be indicated when
+there is a strong relationship with the observed values. This can be
+viewed as a local linear approximation if the fitted model is nonlinear.
+As the contribution of a variable is varied, if the value of the PI does
+not change much, it would indicate that the prediction for this
+observation is NOT sensitive to that variable. Conversely, if the
+predicted value varies substantially, the prediction is very sensitive
+to that variable, suggesting that the variable is very important for the
+PI's prediction.
 
-```{r regressioncase, echo=F, out.width="100%", fig.cap="Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions}
+
+}
+
+\caption[Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables]{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
@@
 Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the variable attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right)."}
-knitr::include_graphics("./figures/app_regression_interactions.PNG")
-```
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}\label{fig:regressioncase}
+\end{figure}
+\end{CodeChunk}
 
-## Interactive variables
+\hypertarget{interactive-variables}{%
+\subsection{Interactive variables}\label{interactive-variables}}
 
-<!-- Reactive vs exploration interactions -->
-The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
-These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a & d).
-The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
+The application has several reactive inputs that affect the data used,
+aesthetic display, and tour manipulation. These reactive inputs make the
+software flexible and extensible (Figure \ref{fig:classificationcase}a
+\& d). The application also has more exploratory interactions to help
+link points across displays, reveal structures found in different
+spaces, and access the original data.
 
-<!-- Exploratory interactions -->
-A tooltip displays observation number/name and classification information while the cursor hovers over a point.
-Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a & b).
-The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
-These interactions aid exploration of the spaces and, finally, the identification of primary and comparison observations.
+A tooltip displays observation number/name and classification
+information while the cursor hovers over a point. Linked brushing allows
+the selection of points (left click and drag) where those points will be
+highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
+The information corresponding to the selected points is populated on a
+dynamic table (Figure \ref{fig:classificationcase}d). These interactions
+aid exploration of the spaces and, finally, the identification of
+primary and comparison observations.
 
-## Preprocessing
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
 
-It is vital to mitigate the render time of visuals, especially when users may want to iterate many explorations.
-All computational operations should be prepared before run time.
-The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
-Below discusses the steps and details of the reprocessing.
+It is vital to mitigate the render time of visuals, especially when
+users may want to iterate many explorations. All computational
+operations should be prepared before run time. The work remaining when
+an application is run solely reacts to inputs and rendering visuals and
+tables. Below discusses the steps and details of the reprocessing.
 
-<!-- Switch to rmd bullets, consistent with thesis -->
-- __Data:__ predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep normality assumptions of the model in mind. 
-- __Model:__ any model and compatible explanation could be explored with this method.
-Currently, random forest models are applied via the package __randomForest__ [@liaw_classification_2002], compatibility tree SHAP.
-Modest hyperparameters are used, namely: 125 trees, the number of variables at each split, mtry = $\sqrt{p}$ or $p/3$ for classification and regression, and minimum size of terminal nodes $max(1, n/500)$ or $max(5, n/500)$ for classification and regression.
-- __Local explanation:__ Tree SHAP is calculated for _each_ observation using the package __treeshap__ [@kominsarczyk_treeshap_2021].
-We opt to find the attribution of each observation in the training data and not fit to fit variable interactions.
-- __Cheem viewer:__ after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
-Statistics for mapping to color are computed on the scaled spaces.
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Data:} predictors and response are unscaled complete numerical
+  matrix. Most models and local explanations are scale-invariant. Keep
+  normality assumptions of the model in mind.
+\item
+  \textbf{Model:} any model and compatible explanation could be explored
+  with this method. Currently, random forest models are applied via the
+  package \textbf{randomForest} \citep{liaw_classification_2002},
+  compatibility tree SHAP. Modest hyperparameters are used, namely: 125
+  trees, the number of variables at each split, mtry = \(\sqrt{p}\) or
+  \(p/3\) for classification and regression, and minimum size of
+  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
+  classification and regression.
+\item
+  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
+  observation using the package \textbf{treeshap}
+  \citep{kominsarczyk_treeshap_2021}. We opt to find the attribution of
+  each observation in the training data and not fit to fit variable
+  interactions.
+\item
+  \textbf{Cheem viewer:} after the model and full explanation space are
+  calculated, each variable is scaled by standard deviations away from
+  the mean to achieve common support for visuals. Statistics for mapping
+  to color are computed on the scaled spaces.
+\end{itemize}
 
-<!-- Note on time of execution -->
-The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
-For reference, the FIFA data contained 5000 observations of nine explanatory variables took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 270 seconds total.
-PCA and statistics of the variables and attributions took 2.8 seconds.
-These run times were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
-An increase in model complexity or data dimensionality will quickly become an obstacle.
-Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-Alternatively, the package __fastshap__ claims extremely low run times, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.
+The time to preprocess the data will vary significantly with the
+complexity of the model and local explanation. For reference, the FIFA
+data contained 5000 observations of nine explanatory variables took 2.5
+seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 270 seconds
+total. PCA and statistics of the variables and attributions took 2.8
+seconds. These run times were from a non-parallelized session on a
+modern laptop, but suffice to say that most of the time will be spent on
+the local attribution. An increase in model complexity or data
+dimensionality will quickly become an obstacle. Its reduced
+computational complexity makes tree SHAP an excellent candidate to
+start. Alternatively, the package \textbf{fastshap} claims extremely low
+run times, attributed to fewer calls to the prediction function, partial
+implementation in C++, and efficient use of logical subsetting,
+\citet{greenwell_fastshap_2020}.
 
-# Case Studies {#sec:casestudies}
+\hypertarget{sec:casestudies}{%
+\section{Case Studies}\label{sec:casestudies}}
 
-To illustrate the cheem method it is applied to modern data sets, two classification examples and then two of regression.
+To illustrate the cheem method it is applied to modern data sets, two
+classification examples and then two of regression.
 
-## Palmer Penguin, Species Classification
+\hypertarget{palmer-penguin-species-classification}{%
+\subsection{Palmer Penguin, Species
+Classification}\label{palmer-penguin-species-classification}}
 
-The Palmer penguins data [@gorman_ecological_2014; @horst_palmerpenguins_2020] was collected on three species of penguins foraging near Palmer Station, Antarctica.
-The data is publicly available to substitute for the overly-used iris data and is quite similar in form.
-After removing incomplete observations, there are 333 observations of four physical measurements, bill length (`bl`), bill depth (`bd`), flipper length (`fl`), and body mass (`bm`) for this illustration.
-A random forest model was fit with species as the response variable.
+The Palmer penguins data
+\citep{gorman_ecological_2014, horst_palmerpenguins_2020} was collected
+on three species of penguins foraging near Palmer Station, Antarctica.
+The data is publicly available to substitute for the overly-used iris
+data and is quite similar in form. After removing incomplete
+observations, there are 333 observations of four physical measurements,
+bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length
+(\texttt{fl}), and body mass (\texttt{bm}) for this illustration. A
+random forest model was fit with species as the response variable.
 
-(ref:casepenguins-cap) Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
-The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at \url{https://vimeo.com/666431172}.)
-```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins-cap)"}
-knitr::include_graphics("./figures/case_penguins.png")
-```
+(ref:casepenguins-cap) Examining the SHAP values for a random forest
+model classifying Palmer penguin species. The PI is a Gentoo (purple)
+penguin that is misclassified as a Chinstrap (orange), marked as an
+asterisk in (a) and the dashed vertical line in (b). The radial view
+shows varying the contribution of \texttt{fl} from the initial
+attribution projection (b, left), which produces a linear combination
+where the PI is more probably (higher density value) a Chinstrap than a
+Gentoo (b, right). (The animation of the radial tour is at
+\url{https://vimeo.com/666431172}.)
 
-Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
-Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
-Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
+\begin{CodeChunk}
+\begin{figure}
 
-(ref:casepenguinsblfl-cap) Checking what is learned from the cheem viewer.
-This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI.
-A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.
-```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl-cap)"}
-knitr::include_graphics("./figures/case_penguins_BlFl.png")
-```
+{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins}
 
-There is more separation visible in the attribution space than in the data space, as would be expected.
-The predicted vs observed plot reveals a handful of misclassified observations.
-A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
-The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
-The CI is a correctly classified point (represented by an $\times$ and a vertical dotted line).
+}
 
-The radial tour starts from the attribution projection of the misclassified observation (b, left).
-The important variables identified by SHAP in the (wrong) prediction for this observation are mostly `bl` and `bd` with small contributions of `fl` and `bm`.
-This projection is a view where the Gentoo (purple) looks much more likely for this observation than Chinstrap.
-That is, this combination of variables is not particularly useful because the PI looks very much like other Gentoo penguins.
-The radial tour is used to vary the contribution of flipper length (`fl`) to explore this.
-(In our exploration, this was the third variable explored.
-It is typically helpful to explore the variables with more significant contributions, here `bl` and `bd`.
-Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
-On varying `fl` as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
-This suggests that `fl` should be considered an important variable for explaining the (wrong) prediction.
+\caption[(ref:casepenguins-cap)]{(ref:casepenguins-cap)}\label{fig:casepenguins}
+\end{figure}
+\end{CodeChunk}
 
-Figure \ref{fig:casepenguinsblfl} confirms that flipper length (`fl`) is vital for the confusion of the PI as a Chinstrap.
-Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two variables, mainly because it has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard observation to account for, as decision trees can only partition only vertical and horizontal lines.
+Figure \ref{fig:casepenguins} shows plots from the cheem viewer for
+exploring the random forest model on the penguins data. Panel (a) shows
+the global view, and panel (b) shows several 1D projections generated
+with the radial tour. Penguin 243, a Gentoo (purple), is the PI because
+it has been misclassified as a Chinstrap (orange).
 
-## Chocolates, Milk/Dark Classification
+(ref:casepenguinsblfl-cap) Checking what is learned from the cheem
+viewer. This is a plot of flipper length (\texttt{fl}) and bill length
+(\texttt{bl}), where an asterisk highlights the PI. A Gentoo (purple)
+misclassified as a Chinstrap (orange). The PI has an unusually small
+\texttt{fl} length which is why it is confused with a Chinstrap.
 
-<!-- Data description -->
-The chocolates data set consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
-Dark chocolate is considered healthier than milk.
-Students collected the data during the Iowa State University class STAT503 from nutritional information on the manufacturer's websites and were normalized to 100g equivalents.
-The data is available in the __cheem__ package.
-A random forest model is used for the classification of chocolate types.
+\begin{CodeChunk}
+\begin{figure}
 
-<!-- Dark misclassified as milk, rational -->
-It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
-A reason to do this is that a dark chocolate, nutritionally more like milk should not be considered a healthy alternative.
-It is interesting to explore which nutritional variables contribute most to misclassification.
+{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}
 
-(ref:casechocolates-cap) Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
-So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
-When Sugar is zero, Fiber contributes strongly towards the left.
-In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
-This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)
-```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates-cap)"}
-knitr::include_graphics("./figures/case_chocolates.png")
-```
+}
 
-<!-- Dark misclassified as milk, description -->
-This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
-It is compared with a CI that is a correctly classified dark chocolate (observation 7).
-The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
-The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
+\caption[(ref:casepenguinsblfl-cap)]{(ref:casepenguinsblfl-cap)}\label{fig:casepenguinsblfl}
+\end{figure}
+\end{CodeChunk}
 
-<!-- Dark misclassified as milk,  -->
-The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
-The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
-In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
-The opposite would be interpreting a point with high values in this plot.
-The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
-The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
+There is more separation visible in the attribution space than in the
+data space, as would be expected. The predicted vs observed plot reveals
+a handful of misclassified observations. A Gentoo that has been wrongly
+labeled as a Chinstrap is selected for illustration. The PI is a
+misclassified point (represented by the asterisk in the global view and
+a dashed vertical line in the tour view). The CI is a correctly
+classified point (represented by an \(\times\) and a vertical dotted
+line).
 
-<!-- Attention to PCP -->
-The PI (dashed line) can be viewed against the CI (dotted line).
-Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the PI.
-The variable contribution of the two different predictions can be quickly compared in the parallel coordinate plot.
-The PI differs from the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+The radial tour starts from the attribution projection of the
+misclassified observation (b, left). The important variables identified
+by SHAP in the (wrong) prediction for this observation are mostly
+\texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and
+\texttt{bm}. This projection is a view where the Gentoo (purple) looks
+much more likely for this observation than Chinstrap. That is, this
+combination of variables is not particularly useful because the PI looks
+very much like other Gentoo penguins. The radial tour is used to vary
+the contribution of flipper length (\texttt{fl}) to explore this. (In
+our exploration, this was the third variable explored. It is typically
+helpful to explore the variables with more significant contributions,
+here \texttt{bl} and \texttt{bd}. Still, when doing this, nothing was
+revealed about how the PI differed from other Gentoos). On varying
+\texttt{fl} as it contributes increasingly to the projection (b, right),
+more and more, this penguin looks like a Chinstrap. This suggests that
+\texttt{fl} should be considered an important variable for explaining
+the (wrong) prediction.
 
-<!-- Attention to Density, end of first case -->
-From the density plot, which is the attribution projection corresponding to the PI, both observations are more like dark chocolates.
-Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
-When a frame with contribution primarily from Fiber is examined observation 22 looks more like a milk chocolate.
+Figure \ref{fig:casepenguinsblfl} confirms that flipper length
+(\texttt{fl}) is vital for the confusion of the PI as a Chinstrap. Here,
+flipper length and body length are plotted, and the PI can be seen to be
+closer to the Chinstrap group in these two variables, mainly because it
+has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard observation to account
+for, as decision trees can only partition only vertical and horizontal
+lines.
 
-<!-- segue and selection for chocolates inverse case -->
-It would also be interesting to explore an inverse misclassification.
-In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
-Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (observation 71).
-The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
+\hypertarget{chocolates-milkdark-classification}{%
+\subsection{Chocolates, Milk/Dark
+Classification}\label{chocolates-milkdark-classification}}
 
-(ref:casechocolatesinverse-cap) Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
-In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates.
-The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
-If Sodium had a larger contribution to the prediction (like in this view).
-the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)
-```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse-cap)"}
-knitr::include_graphics("./figures/case_chocolates_inverse.png")
-```
+The chocolates data set consists of 88 observations of ten nutritional
+measurements determined from their labels and labeled as either milk or
+dark. Dark chocolate is considered healthier than milk. Students
+collected the data during the Iowa State University class STAT503 from
+nutritional information on the manufacturer's websites and were
+normalized to 100g equivalents. The data is available in the
+\textbf{cheem} package. A random forest model is used for the
+classification of chocolate types.
 
-<!-- treeSHAP pca, for inverse case -->
-The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
-Looking at the attribution projection, this is found to be the case.
-Previously, Fiber was essential while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
+It could be interesting to examine the nutritional properties of any
+dark chocolates that have been misclassified as milk. A reason to do
+this is that a dark chocolate, nutritionally more like milk should not
+be considered a healthy alternative. It is interesting to explore which
+nutritional variables contribute most to misclassification.
 
-<!-- PCP, for inverse case -->
-Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
-The contribution of Sodium is selected to be varied.
-Even in the initial projection, the observation looks slightly more like its observed milk than predicted dark chocolate.
-The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
+(ref:casechocolates-cap) Examining the local explanation for a PI which
+is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more
+like dark than milk, which suggests that the local explanation does not
+help understand the prediction for this observation. So, the
+contribution of Sugar is varied---reducing it corresponds primarily with
+an increased magnitude from Fiber. When Sugar is zero, Fiber contributes
+strongly towards the left. In this view, the PI is closer to the bulk of
+the milk chocolates, suggesting that the prediction put a lot of
+importance on Fiber. This chocolate is a rare dark chocolate without any
+Fiber leading to it being mistaken for a milk chocolate. (A video of the
+tour animation can be found at \url{https://vimeo.com/666431143}.)
 
-## FIFA, Wage Regression
+\begin{CodeChunk}
+\begin{figure}
 
-<!-- Data description -->
-The 2020 season FIFA data [@leone_fifa_2020; @biecek_dalex_2018] contains many skill measurements of soccer/football players and wage information.
-Nine higher-level skill groupings were identified and aggregated from highly correlated variables.
-A random forest model is fit from these predictors, regressing player wages [2020 euros].
-The model was fit from 5000 observations before being thinned to 500 players to mitigate occlusion and render time.
-Continuing from the exploration in Section \ref{sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
-That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
-A leading offensive fielder (L.
-Messi) is compared with a top defensive fielder (V.
-van Dijk).
-The same observations were used in Figure \ref{fig:shapdistrbd}.
+{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates}
 
-(ref:casefifa-cap) Exploring the wages relative to skill measurements in the FIFA 2020 data.
-Star offensive player (L.
-Messi) is the PI, and he is compared with a top defensive player (V.
-van Dijk).
-The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
-Defense (`def`) is the chosen variable to vary.
-It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
-The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)
-```{r casefifa, echo=F, out.width = "90%", fig.cap = "(ref:casefifa-cap)"}
-knitr::include_graphics("./figures/case_fifa.png")
-```
+}
+
+\caption[(ref:casechocolates-cap)]{(ref:casechocolates-cap)}\label{fig:casechocolates}
+\end{figure}
+\end{CodeChunk}
+
+This type of exploration is shown in Figure \ref{fig:casechocolates},
+where a chocolate labeled dark but predicted to be milk is chosen as the
+PI (observation 22). It is compared with a CI that is a correctly
+classified dark chocolate (observation 7). The PCA plot and the tree
+SHAP PCA plots (a) show a big difference between the two chocolate types
+but with confusion for a handful of observations. The misclassifications
+are more apparent in the observed vs predicted plot and can be seen to
+be mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars,
+and Calories are most responsible for its incorrect prediction. The way
+to read this plot is to see that Fiber has a large negative value while
+Sugars and Calories have reasonably large positive values. In the
+density plot, observations on the very left of the display would have
+high values of Fiber (matching the negative projection coefficient) and
+low values of Sugars and Calories. The opposite would be interpreting a
+point with high values in this plot. The dark chocolates (orange) are
+primarily on the left, and this is a reason why they are considered to
+be healthier: high fiber and low sugar. The density of milk chocolates
+is further to the right, indicating that they generally have low fiber
+and high sugar.
+
+The PI (dashed line) can be viewed against the CI (dotted line). Now,
+one needs to pay attention to the parallel plot of the SHAP values,
+which are local to a particular observation, and the density plot, which
+is the same projection of all observations as specified by the SHAP
+values of the PI. The variable contribution of the two different
+predictions can be quickly compared in the parallel coordinate plot. The
+PI differs from the comparison primarily on the Fiber variable, which
+suggests that this is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding
+to the PI, both observations are more like dark chocolates. Varying the
+contribution of Sugars and altogether removing it from the projection is
+where the difference becomes apparent. When a frame with contribution
+primarily from Fiber is examined observation 22 looks more like a milk
+chocolate.
+
+It would also be interesting to explore an inverse misclassification. In
+this case, a milk chocolate is selected while it was misclassified as a
+dark chocolate. Chocolate 84 is just this case and is compared with a
+correctly predicted milk chocolate (observation 71). The corresponding
+global view and radial tour frames are shown in Figure
+\ref{fig:casechocolatesinverse}.
+
+(ref:casechocolatesinverse-cap) Examining the local explanation for a PI
+which is milk (green) chocolate incorrectly predicted to be dark
+(orange). In the attribution projection, the PI could be either milk or
+dark. Sodium and Fiber have the largest differences in attributed
+variable importance, with low values relative to other milk chocolates.
+The lack of importance attributed to these variables is suspected of
+contributing to the mistake, so the contribution of Sodium is varied. If
+Sodium had a larger contribution to the prediction (like in this view).
+the PI would look more like other milk chocolates. (A video of the tour
+animation can be found at \url{https://vimeo.com/666431148}.)
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates_inverse}
+
+}
+
+\caption[(ref:casechocolatesinverse-cap)]{(ref:casechocolatesinverse-cap)}\label{fig:casechocolatesinverse}
+\end{figure}
+\end{CodeChunk}
+
+The difference of position in the tree SHAP PCA with the previous case
+is quite significant; this gives a higher-level sense that the
+attributions should be quite different. Looking at the attribution
+projection, this is found to be the case. Previously, Fiber was
+essential while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat have high attributions here,
+while they were unimportant in the preceding case.
+
+Comparing the attribution with the CI (dotted line), large discrepancies
+in Sodium and Fiber are identified. The contribution of Sodium is
+selected to be varied. Even in the initial projection, the observation
+looks slightly more like its observed milk than predicted dark
+chocolate. The misclassification appears least supported when the basis
+reaches sodium attribution of typical dark chocolate.
+
+\hypertarget{fifa-wage-regression}{%
+\subsection{FIFA, Wage Regression}\label{fifa-wage-regression}}
+
+The 2020 season FIFA data \citep{leone_fifa_2020, biecek_dalex_2018}
+contains many skill measurements of soccer/football players and wage
+information. Nine higher-level skill groupings were identified and
+aggregated from highly correlated variables. A random forest model is
+fit from these predictors, regressing player wages {[}2020 euros{]}. The
+model was fit from 5000 observations before being thinned to 500 players
+to mitigate occlusion and render time. Continuing from the exploration
+in Section \textbackslash ref\{sec:explanations), we are interested to
+see the difference in attribution based on the exogenous player
+position. That is, the model should be able to use multiple linear
+profiles to better predict the wages from different field positions of
+players despite not having this information. A leading offensive fielder
+(L. Messi) is compared with a top defensive fielder (V. van Dijk). The
+same observations were used in Figure \ref{fig:shapdistrbd}.
+
+(ref:casefifa-cap) Exploring the wages relative to skill measurements in
+the FIFA 2020 data. Star offensive player (L. Messi) is the PI, and he
+is compared with a top defensive player (V. van Dijk). The attribution
+projection is shown on the left, and it can be seen that this
+combination of variables produces a view where Messi has very high
+predicted (and observed) wages. Defense (\texttt{def}) is the chosen
+variable to vary. It starts very low, and Messi's predicted wages
+decrease dramatically as its contribution increases (right plot). The
+increased contribution in defense comes at the expense of offensive and
+reaction skills. The interpretation is that Messi's high wages are most
+attributable to his offensive and reaction skills, as initially provided
+by the local explanation. (A video of the animated radial tour can be
+found at \url{https://vimeo.com/666431163}.)
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{./figures/case_fifa}
+
+}
+
+\caption[(ref:casefifa-cap)]{(ref:casefifa-cap)}\label{fig:casefifa}
+\end{figure}
+\end{CodeChunk}
 
 Figure \ref{fig:casefifa} tests the support of the local explanation.
-Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
-If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
-However, increasing the contribution of a variable with low importance would rotate both variables out of the frame.
+Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both
+crucial to explaining a star offensive player. If either of them were
+rotated out, the other would be rotated into the frame, maintaining a
+far-right position. However, increasing the contribution of a variable
+with low importance would rotate both variables out of the frame.
 
-The contribution from `def` will be varied to contrast with offensive skills.
-As the contribution of defensive skills increases, Messi's is no longer separated from the group.
-Players with high values in defensive skills are now the rightmost points.
-In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions were at these levels.
+The contribution from \texttt{def} will be varied to contrast with
+offensive skills. As the contribution of defensive skills increases,
+Messi's is no longer separated from the group. Players with high values
+in defensive skills are now the rightmost points. In terms of what-if
+analysis, the difference between the data mean and his predicted wages
+would be halved if Messi's tree SHAP attributions were at these levels.
 
-(ref:caseames-cap) Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
-The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
-The PI has a higher predicted value than the CI in the attribution projection.
-Reducing the contribution of Lot Area brings these two prices in line.
-This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar.
-That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at \url{https://vimeo.com/666431134}.)
-```{r caseames, echo=F, out.width="90%", fig.cap = "(ref:caseames-cap)"}
-knitr::include_graphics("./figures/case_ames2018.png")
-```
+(ref:caseames-cap) Exploring an observation with an extreme residual as
+the PI in relation to an observation with an accurate prediction for a
+similarly priced house in a random forest fit to the Ames housing data.
+The local explanation indicates a sizable attribution to Lot Area
+(\texttt{LtA}), while the CI has minimal attribution to this variable.
+The PI has a higher predicted value than the CI in the attribution
+projection. Reducing the contribution of Lot Area brings these two
+prices in line. This suggests that if the model did not value Lot Area
+so highly for this observation, then the observed sales price would be
+quite similar. That is, the large residual is due to a lack of factoring
+in the Lot Area for the prediction of PI's sales price. (A video showing
+the animation is at \url{https://vimeo.com/666431134}.)
 
-## Ames Housing 2018, Sales Price Regression
+\begin{CodeChunk}
+\begin{figure}
 
-Ames housing data 2018 [@de_cock_ames_2011; @prevek18_ames_2018] was subset to North Ames (the neighborhood with the most house sales).
-The remaining are 338 house sales.
-A random forest model was fit, predicting the sale price [USD] from the property variables: Lot Area (`LtA`), Overall Quality (`Qlt`), Year the house was Built (`YrB`), Living Area (`LvA`), number of Bathrooms (`Bth`), number of Bedrooms (`Bdr`), the total number of Rooms (`Rms`), Year the Garage was Built (`GYB`), and Garage Area (`GrA`).
-Using  interactions with the global view, a house with an extreme negative residual and an accurate observation with a similar prediction is selected.
+{\centering \includegraphics[width=0.9\linewidth]{./figures/case_ames2018}
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
-The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
-The attribution projection places observations with high Living Areas to the right.
-The contribution of Living Area contrasts the contribution of this variable.
-As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
-This large importance in the Living Area is relatively uncommon.
-Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
+}
 
-# Discussion {#sec:cheemdiscussion}
+\caption[(ref:caseames-cap)]{(ref:caseames-cap)}\label{fig:caseames}
+\end{figure}
+\end{CodeChunk}
 
-There is a clear need to extend the interpretability of black box models.
-This paper provides a technique that builds on local explanations to explore the variable importance local to an observation.
-The local explanations of an attribution projection from which variable contributions are varied using a radial tour.
-Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular variables.
-A global view shows the data space, explanation space, and residual plot.
-The user can interactively select observations to compare, contrast, and study further.
-Then the radial tour is used to explore the variable sensitivity identified by the attribution projection.
+\hypertarget{ames-housing-2018-sales-price-regression}{%
+\subsection{Ames Housing 2018, Sales Price
+Regression}\label{ames-housing-2018-sales-price-regression}}
 
-This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
-Local explanations focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
-They can also form the basis of learning how the model has got it wrong, with when the observation is misclassified or has a large residual.
+Ames housing data 2018 \citep{de_cock_ames_2011, prevek18_ames_2018} was
+subset to North Ames (the neighborhood with the most house sales). The
+remaining are 338 house sales. A random forest model was fit, predicting
+the sale price {[}USD{]} from the property variables: Lot Area
+(\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built
+(\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms
+(\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of
+Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and
+Garage Area (\texttt{GrA}). Using interactions with the global view, a
+house with an extreme negative residual and an accurate observation with
+a similar prediction is selected.
 
-In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
-This was verified by making a follow-up plot of the data.
-The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
-In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area is not being effectively used by the random forest model.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under
+prediction with an enormous Lot Area contribution. The CI has a similar
+predicted price though the prediction was accurate and gives almost no
+attribution to lot size. The attribution projection places observations
+with high Living Areas to the right. The contribution of Living Area
+contrasts the contribution of this variable. As the contribution of Lot
+Area decreases, the predictive power decreases for the PI, while the CI
+remains stationary. This large importance in the Living Area is
+relatively uncommon. Boosting tree models may be more resilient to such
+an under-prediction as they would up-weighting this residual and force
+its inclusion in the final model.
 
-This analysis is manually intensive and thus only feasible for investigating a few observations.
-The recommended approach is to investigate an observation where the model has not predicted accurately and compare it with an observation with similar predictor values where model fitted well.
-The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any variable.
-It can be helpful to make additional plots of the variables and responses to cross-check interpretations made from the cheem viewer.
-This methodology provides an additional tool in the box for studying model fitting.
+\hypertarget{sec:cheemdiscussion}{%
+\section{Discussion}\label{sec:cheemdiscussion}}
 
-<!--checklist asks for: Acknowledgments, Appendices, then References-->
+There is a clear need to extend the interpretability of black box
+models. This paper provides a technique that builds on local
+explanations to explore the variable importance local to an observation.
+The local explanations of an attribution projection from which variable
+contributions are varied using a radial tour. Several diagnostic plots
+are provided to assist with understanding the sensitivity of the
+prediction to particular variables. A global view shows the data space,
+explanation space, and residual plot. The user can interactively select
+observations to compare, contrast, and study further. Then the radial
+tour is used to explore the variable sensitivity identified by the
+attribution projection.
 
-# Package Infrastructure {#sec:infrastructure}
+This approach has been illustrated using four data examples of random
+forest models with the tree SHAP local explanation. Local explanations
+focus on the model fit, and help to dissect which variables are most
+responsible for the fitted value. They can also form the basis of
+learning how the model has got it wrong, with when the observation is
+misclassified or has a large residual.
 
-An implementation is provided in the open-source __R__ package __cheem__, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
-Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-The local explanations need to be pre-computed and uploaded.
-Examples show how to do this for tree SHAP values, using  __treeshap__ (tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively).
-The SHAP and oscillation explanations could be easily added using `DALEX::explain()` [@biecek_dalex_2018; @biecek_explanatory_2021].
+In the penguins example, we showed how the misclassification of a
+penguin arose due to it having an unusually small flipper size compared
+to others of its species. This was verified by making a follow-up plot
+of the data. The chocolates example shows how a dark chocolate was
+misclassified primarily due to its attribution to Fiber, and a milk
+chocolate was misclassified as dark due to its lowish Sodium value. In
+the FIFA example, we show how low Messi's salary would be if it depended
+on defensive skills. In the Ames housing data, an inaccurate prediction
+for a house was likely due to the Lot Area is not being effectively used
+by the random forest model.
 
-The application was made with __shiny__ [@chang_shiny_2021].
-The tour visual is built with __spinifex__ [@spyrison_spinifex_2020].
-Both views are created first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive `html` widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a boon to understanding local explanations and how to apply them.
+This analysis is manually intensive and thus only feasible for
+investigating a few observations. The recommended approach is to
+investigate an observation where the model has not predicted accurately
+and compare it with an observation with similar predictor values where
+model fitted well. The radial tour launches from the attribution
+projection to enable exploration of the sensitivity of the prediction to
+any variable. It can be helpful to make additional plots of the
+variables and responses to cross-check interpretations made from the
+cheem viewer. This methodology provides an additional tool in the box
+for studying model fitting.
 
-The package can be installed from CRAN, and the application can be run using the following __R__ code:
+\hypertarget{sec:infrastructure}{%
+\section{Package Infrastructure}\label{sec:infrastructure}}
 
-```{r eval=FALSE, echo=TRUE}
-install.packages("cheem", dependencies = TRUE)
-library("cheem")
-run_app()
-```
+An implementation is provided in the open-source \textbf{R} package
+\textbf{cheem}, available on CRAN at
+\url{https://CRAN.R-project.org/package=cheem}. Example data sets are
+provided, and you can upload your data after model fitting and computing
+the local explanations. The local explanations need to be pre-computed
+and uploaded. Examples show how to do this for tree SHAP values, using
+\textbf{treeshap} (tree based models from \textbf{gbm},
+\textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or
+\textbf{xgboost} \citet{greenwell_gbm_2020}; \citet{shi_lightgbm_2022};
+\citet{liaw_classification_2002}; \citet{wright_ranger_2017};
+\citet{chen_xgboost_2021}, respectively). The SHAP and oscillation
+explanations could be easily added using \texttt{DALEX::explain()}
+\citep{biecek_dalex_2018, biecek_explanatory_2021}.
+
+The application was made with \textbf{shiny} \citep{chang_shiny_2021}.
+The tour visual is built with \textbf{spinifex}
+\citep{spyrison_spinifex_2020}. Both views are created first with
+\textbf{ggplot2} \citep{wickham_ggplot2_2016} and then rendered as
+interactive \texttt{html} widgets with \textbf{plotly}
+\citep{sievert_interactive_2020}. \textbf{DALEX}
+\citep{biecek_dalex_2018} and the free ebook, \emph{Explanatory Model
+Analysis} \citep{biecek_explanatory_2021} were a boon to understanding
+local explanations and how to apply them.
+
+The package can be installed from CRAN, and the application can be run
+using the following \textbf{R} code:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> install.packages("cheem", dependencies = TRUE)
+R> library("cheem")
+R> run_app()
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-- A version of the cheem viewer shiny app can be directly accessed at
+\begin{itemize}
+\tightlist
+\item
+  A version of the cheem viewer shiny app can be directly accessed at
   \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
-- The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
-- Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+\item
+  The development version of the package is available at
+  \url{https://github.com/nspyrison/cheem}, and
+\item
+  Documentation of the package can be found at
+  \url{https://nspyrison.github.io/cheem/}.
+\end{itemize}
 
-Follow the examples provided with the package to compute the local explanation (using `?cheem_ls`).
-The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()` to be uploaded.
+Follow the examples provided with the package to compute the local
+explanation (using \texttt{?cheem\_ls}). The application expects the
+output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file
+with \texttt{saveRDS()} to be uploaded.
 
-## Acknowledgments {-}
+\hypertarget{acknowledgments}{%
+\subsection*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{subsection}{Acknowledgments}
 
-Kim Marriott provided advice on many aspects of this work, especially on the explanations in the applications section.
-This research was supported by the Australian Government Research Training Program (RTP) scholarships.
-Thanks to Jieyang Chong for helping proofread this article.
-The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
-__DALEX__ pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+Kim Marriott provided advice on many aspects of this work, especially on
+the explanations in the applications section. This research was
+supported by the Australian Government Research Training Program (RTP)
+scholarships. Thanks to Jieyang Chong for helping proofread this
+article. The namesake, Cheem, refers to a fictional race of humanoid
+trees from Doctor Who lore. \textbf{DALEX} pulls on from that universe,
+and we initially apply tree SHAP explanations specific to tree-based
+models.
 
-# References
+\renewcommand\refname{References}
+\bibliography{paper.bib}
+
+\end{document}

### 2023-05-08 jss/data/preprocess_ames2018.r

@@ -1,0 +1,35 @@
+load("data/amesHousing2018_NorthAmes.rda")
+{
+  dat  <- amesHousing2018_NorthAmes
+  X    <- dat[, 1:9]
+  colnames(X) <- c("LtA", "Qlt", "YrB", "LvA", "Bth", "Bdr", "Rms", "GYB", "GrA")
+  Y    <- dat$SalePrice
+  clas <- amesHousing2018_NorthAmes$SubclassMS ## class is a zone subclass
+}
+
+rf_fit <- randomForest::randomForest(
+  X, Y, ntree = 125,
+  mtry = ifelse(is_discrete(Y), sqrt(ncol(X)), ncol(X) / 3),
+  nodesize = max(ifelse(is_discrete(Y), 1, 5), nrow(X) / 500))
+rf_pred <- predict(rf_fit, X)
+rf_shap <- treeshap::treeshap(
+  treeshap::randomForest.unify(rf_fit, X), X, FALSE, FALSE)
+rf_shap <- rf_shap$shaps
+
+chm <- cheem_ls(X, Y, rf_shap, rf_pred, clas,
+                label = "North Ames, RF, treeshap")
+
+## Export ----
+NM <- "preprocess_ames2018.rds"
+saveRDS(chm, file = paste0("data/", NM))
+cat("Saved", NM, "\n")
+
+if(F){
+  ## Don't run load cheem list
+  chm <- readRDS(paste0("data/", NM))
+  lapply(chm, object.size)
+  
+  ## Don't run manual check
+  names(chm)
+  global_view(chm)
+}

### 2023-05-08 jss/data/preprocess_chocolates.r

@@ -1,0 +1,38 @@
+load("data/chocolates.rda")
+{
+  X <- chocolates[, 5:14] %>% as.data.frame()
+  Y <- clas <- chocolates$Type
+  colnames(X) <- gsub("\\_.*", "", colnames(X))
+  nm_imc <- paste(chocolates$Name, chocolates$MFR, chocolates$Country, sep = ", ")
+  r_idx <- which(nm_imc == "85% Cocoa Dark French Chocolate, Thorntons, UK")[2]
+  nm_imc[r_idx] <- paste0(nm_imc[r_idx], " (2nd)")
+  row.names(X) <- nm_imc
+}
+
+## Model and predict
+train    <- data.matrix(X) %>% xgb.DMatrix(label = Y)
+xgb_fit  <- xgboost(data = train, max.depth = 3, nrounds = 25)
+xgb_pred <- predict(xgb_fit, newdata = train)
+
+## shapviz
+xgb_shap <- shapviz(xgb_fit, X_pred = train, X = X)
+xgb_shap <- xgb_shap$S
+
+## Cheem
+chm <- cheem_ls(X, Y, xgb_shap, xgb_pred, clas,
+                label = "Chocolates, xgb, shapviz")
+
+## Export ----
+NM <- "preprocess_chocolates.rds"
+saveRDS(chm, file = paste0("data/", NM))
+cat("Saved", NM, "\n")
+
+if(F){
+  ## Don't run load cheem list
+  chm <- readRDS(paste0("data/", NM))
+  lapply(chm, object.size)
+  
+  ## Don't run manual check
+  names(chm)
+  global_view(chm)
+}

### 2023-05-08 jss/data/preprocess_fifa.r

@@ -1,0 +1,81 @@
+## Setup ------
+{
+  .raw <- DALEX::fifa
+  .dat_less_ys <- .raw %>%
+    dplyr::select(
+      -c(`nationality`, ## useless class
+         ## potential target vars:
+         `overall`, `potential`, `value_eur`, `wage_eur`)) %>%
+    as.data.frame()
+  
+  if(F) ## Don't run, View corrplot of the groupings
+    corrplot::corrplot(cor(.dat_less_ys),
+                       method = "circle", ## geom
+                       type = "upper", ## only upper triangle
+                       diag = F, ## remove auto correlation
+                       order = "FPC", ## First principal component
+                       tl.col = "black", tl.srt = 90, ## Text label color and rotation
+                       tl.pos = "td")
+  
+  ## Aggregate some highly correlated vars.
+  dat <- .dat_less_ys %>%
+    dplyr::mutate(
+      .keep = "none",
+      BMI = (weight_kg+(height_cm / 100)^2)/2,
+      age = age,
+      react = movement_reactions,
+      off = (attacking_finishing+skill_long_passing+attacking_volleys+
+               power_long_shots+skill_curve+mentality_positioning+attacking_crossing+
+               attacking_short_passing+skill_dribbling+skill_ball_control)/10,
+      def = (defending_sliding_tackle+mentality_interceptions+
+               defending_standing_tackle+defending_marking+mentality_aggression)/5,
+      acc = (attacking_heading_accuracy+power_shot_power)/2,
+      mvm = (movement_sprint_speed+movement_balance+movement_acceleration+
+               mentality_vision+mentality_composure+movement_agility+
+               mentality_penalties+skill_fk_accuracy+power_stamina+movement_reactions)/10L,
+      pwr = (power_strength+power_jumping)/2,
+      gk  = (goalkeeping_diving+goalkeeping_positioning+goalkeeping_reflexes+
+               goalkeeping_handling+goalkeeping_kicking)/5
+    )
+  
+  ## Starting with 42 variables, we remove `nationality`, and some potential Y vars,
+  #### and aggregate into 9 aggregate 'aspect' dimensions based on var correlation 
[…]

### 2023-05-08 jss/data/preprocess_penguins.r

@@ -1,0 +1,35 @@
+## Setup ------
+{
+  raw <- spinifex::penguins_na.rm
+  X <- raw[, 1:4] %>% as.data.frame()
+  Y <- clas <- raw$species
+  colnames(X) <- c("bl", "bd", "fl", "bm")
+}
+
+## Model and predict
+train    <- data.matrix(X) %>% xgb.DMatrix(label = Y)
+xgb_fit  <- xgboost(data = train, max.depth = 3, nrounds = 25)
+xgb_pred <- predict(xgb_fit, newdata = train)
+
+## shapviz
+xgb_shap <- shapviz(xgb_fit, X_pred = train, X = X)
+xgb_shap <- xgb_shap$S
+
+## Cheem
+chm <- cheem_ls(X, Y, xgb_shap, xgb_pred, clas,
+                label = "Penguins, xgb, shapviz")
+
+## Export ----
+NM <- "preprocess_penguins.rds"
+saveRDS(chm, file = paste0("data/", NM))
+cat("Saved", NM, "\n")
+
+if(F){
+  ## Don't run load cheem list
+  chm <- readRDS(paste0("data/", NM))
+  lapply(chm, object.size)
+  
+  ## Don't run manual check
+  names(chm)
+  global_view(chm)
+}

### 2023-05-08 jss/figures/_make_figures.r

@@ -1,0 +1,278 @@
+# ch5_fig1_shap_distr_bd ----
+{
+  require("DALEX")
+  require("dplyr")
+  require("ggplot2")
+  set.seed(2022)
+  
+  ## Local func
+  my_parts_boxplot_df <- function(pred_parts, player_tag = "<tag unused>"){
+    ## Remade from: iBreakDown:::print.break_down_uncertainty
+    data.frame(
+      player   = player_tag,
+      label    = tapply(pred_parts$label, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+      variable = tapply(pred_parts$variable_name, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE),
+      value    = tapply(pred_parts$variable_value, paste(pred_parts$label, pred_parts$variable, sep = ": "), unique, na.rm = TRUE), ## oos variable value
+      ## Of the distribution of local attributions:
+      # min      = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), min, na.rm = TRUE),
+      # q1       = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.25, na.rm = TRUE),
+      mean     = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), mean, na.rm = TRUE),
+      median   = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), median, na.rm = TRUE)#,
+      # q3       = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), quantile, 0.75, na.rm = TRUE),
+      # max      = tapply(pred_parts$contribution, paste(pred_parts$label, pred_parts$variable, sep = ": "), max, na.rm = TRUE)
+    )
+  }
+  my_parts_distribution <- function(pred_parts, player_tag = "<tag unused>"){
+    df <- data.frame(
+      player       = player_tag,
+      label        = pred_parts$label,
+      variable     = pred_parts$variable_name,
+      value        = pred_parts$variable_value, ## Obs value of X
+      contribution = pred_parts$contribution,   ## SHAP contribution
+      B_perm_num   = pred_parts$B
+    )
+    rownames(df) <- paste(pred_parts$label, pred_parts$variable, pred_parts$B, sep = ": ")
+    return(df)
+  }
[…]

### 2023-05-08 jss/figures/_make_figures_case_study.r

@@ -1,0 +1,443 @@
+# CASE STUDIES FOR CHEEM ----
+## Setup ----
+{
+  require("cheem")
+  require("spinifex")
+  require("dplyr")
+  require("ggplot2")
+  require("cowplot")
+  require("patchwork")
+  
+  ## Load data:
+  penguins_ls   <- readRDS("data/preprocess_penguins.rds")
+  fifa_ls       <- readRDS("data/preprocess_fifa.rds")
+  ames2018_ls   <- readRDS("data/preprocess_ames2018.rds")
+  chocolates_ls <- readRDS("data/preprocess_chocolates.rds")
+  
+  .t <- theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
+              axis.text   = element_blank(),
+              axis.title  = element_blank(),
+              axis.ticks  = element_blank(),
+              legend.position = "off", aspect.ratio = 1)
+}
+
+## Penguins classification ------
+{
+  ## Data setup, spinifex::penguins
+  names(penguins_ls)
+  prim_obs <- 243
+  comp_obs <- 169
+  
+  ### Global view and cheem tour stills
+  .glob_view <- global_view(
+    penguins_ls, prim_obs, comp_obs, as_ggplot = TRUE) + .t +
+    labs(color = "Predicted class", shape = "Predicted class", x = element_blank()) +
+    ggtitle("Global view") + theme(legend.position  = "bottom",
+                                   legend.direction = "horizontal")
+  .bas <- basis_attr_df(penguins_ls$attr_df, prim_obs)
+  .mv  <- which(colnames(penguins_ls$attr_df) == "fl")
+  ## Cheem tour for stills
+  mt_interp <- manual_tour(basis = .bas, manip_var = .mv) %>%
+    spinifex:::interpolate_manual_tour(angle = .15) ## app is .15
+  dim(mt_interp)
+  .ggt1 <- radial_cheem_tour(
+    penguins_ls, basis = mt_interp[,,1, drop=FALSE], manip_var = .mv,
+    primary_obs = prim_obs, comparison_obs = comp_obs,
+    do_add_pcp_segments = FALSE,
+    pcp_shape = 124, angle = 0) + .t +
+    theme(plot.title = element_text(hjust = 0.22)) +
+    ggtitle("Radial tour, select frames")
+  .ggt2 <- radial_cheem_tour(
[…]

### 2023-05-16 jss/cheem_paper.Rmd
extended discussion related to unsupervised analysis

@@
 (The animation of the radial tour is at \url{https://vimeo.com/666431172}.)
-```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins-cap)"}
+
+```{r casepenguins, echo=F, out.width="100%", fig.cap = '(ref:casepenguins-cap)'}
 knitr::include_graphics("./figures/case_penguins.png")
@@
 The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.
-```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = "(ref:casepenguinsblfl-cap)"}
+
+```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = '(ref:casepenguinsblfl-cap)'}
 knitr::include_graphics("./figures/case_penguins_BlFl.png")
@@
 (A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)
-```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates-cap)"}
+
+```{r casechocolates, echo=F, out.width="100%", fig.cap = '(ref:casechocolates-cap)'}
 knitr::include_graphics("./figures/case_chocolates.png")
@@
 (A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)
-```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "(ref:casechocolatesinverse-cap)"}
+
+```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = '(ref:casechocolatesinverse-cap)'}
 knitr::include_graphics("./figures/case_chocolates_inverse.png")
@@
 (A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)
-```{r casefifa, echo=F, out.width = "90%", fig.cap = "(ref:casefifa-cap)"}
+
+```{r casefifa, echo=F, out.width = "90%", fig.cap = '(ref:casefifa-cap)'}
 knitr::include_graphics("./figures/case_fifa.png")
@@
 (A video showing the animation is at \url{https://vimeo.com/666431134}.)
-```{r caseames, echo=F, out.width="90%", fig.cap = "(ref:caseames-cap)"}
+```{r caseames, echo=F, out.width="90%", fig.cap = '(ref:caseames-cap)'}
 knitr::include_graphics("./figures/case_ames2018.png")
@@
 There is a clear need to extend the interpretability of black box models.
-This paper provides a technique that builds on local explanations to explore the variable importance local to an observation.
+With techniques such as SHAP, LIME, Break-down, one can calculate local explanations, i.e.
+for every observation in the data.
+These techniques quantify for each observation how strongly particular variables affect the model's predictions.
+Surprisingly few techniques allow us to understand the global distribution of these local explanations.
+Unsupervised data exploration techniques applied to data show how useful they are for identifying outliers, identifying segments of observations or discovering correlations between variables.
+All of these tasks can be performed for a set of explanations.
+
+To address this challenge in this paper provides a technique that builds on local explanations to explore the variable importance local to an observation.
 The local explanations of an attribution projection from which variable contributions are varied using a radial tour.

### 2023-05-16 jss/cheem_paper.tex
extended discussion related to unsupervised analysis

@@
 There is a clear need to extend the interpretability of black box
-models. This paper provides a technique that builds on local
-explanations to explore the variable importance local to an observation.
-The local explanations of an attribution projection from which variable
+models. With techniques such as SHAP, LIME, Break-down, one can
+calculate local explanations, i.e.~for every observation how strongly
+variables affect the model's predictions. Surprisingly few techniques
+allow us to understand the global distribution of these local
+explanations. Unsupervised data exploration techniques applied to data
+show how useful they are for identifying outliers, identifying segments
+of observations or discovering correlations between variables. All of
+these tasks can be performed for a set of explanations.
+
+This paper provides a technique that builds on local explanations to
+explore the variable importance local to an observation. The local
+explanations of an attribution projection from which variable
 contributions are varied using a radial tour. Several diagnostic plots

### 2023-05-16 jss/cheem_paper.Rmd
small update in the abstract

@@
 XAI attempts to shed light on how models use predictors to arrive at a prediction.
-"Local explanations", which provide a point estimate of the linear variable importance in the vicinity of one observation, are one XAI method.
-These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface.
+"Local variable attributions", which provide a point estimate of the linear variable importance in the vicinity of one observation, are one local XAI method.
+These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface as well as understand the structure of local explanations for observations, whether outlier are present, whether segments can be identified.
 Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.

### 2023-05-16 jss/cheem_paper.tex
small update in the abstract

@@
 of eXplainable AI (XAI). XAI attempts to shed light on how models use
-predictors to arrive at a prediction. ``Local explanations'', which
-provide a point estimate of the linear variable importance in the
-vicinity of one observation, are one XAI method. These can be considered
-to be linear projections and can be further explored to understand the
-interactions between variables used to make predictions across the
-predictive model surface. Here we describe interactive linear
+predictors to arrive at a prediction. ``Local variable attributions'',
+which provide a point estimate of the linear variable importance in the
+vicinity of one observation, are one local XAI method. These can be
+considered to be linear projections and can be further explored to
+understand the interactions between variables used to make predictions
+across the predictive model surface as well as understand the structure
+of local explanations for observations, whether outlier are present,
+whether segments can be identified. Here we describe interactive linear
 interpolation used for exploration at any observation and illustrate
@@
 models. With techniques such as SHAP, LIME, Break-down, one can
-calculate local explanations, i.e.~for every observation how strongly
+calculate local explanations, i.e.~for every observation in the data.
+These techniques quantify for each observation how strongly particular
 variables affect the model's predictions. Surprisingly few techniques
@@
 
-This paper provides a technique that builds on local explanations to
-explore the variable importance local to an observation. The local
-explanations of an attribution projection from which variable
-contributions are varied using a radial tour. Several diagnostic plots
-are provided to assist with understanding the sensitivity of the
-prediction to particular variables. A global view shows the data space,
-explanation space, and residual plot. The user can interactively select
-observations to compare, contrast, and study further. Then the radial
-tour is used to explore the variable sensitivity identified by the
-attribution projection.
+To address this challenge in this paper provides a technique that builds
+on local explanations to explore the variable importance local to an
+observation. The local explanations of an attribution projection from
+which variable contributions are varied using a radial tour. Several
+diagnostic plots are provided to assist with understanding the
+sensitivity of the prediction to particular variables. A global view
+shows the data space, explanation space, and residual plot. The user can
+interactively select observations to compare, contrast, and study
+further. Then the radial tour is used to explore the variable
+sensitivity identified by the attribution projection.
 

### 2023-05-16 jss/cheem_paper.Rmd
figure captions fixed

@@
 
-(ref:casepenguins-cap) Examining the SHAP values for a random forest model classifying Palmer penguin species.
+```{r casepenguins, echo=F, out.width="100%", fig.cap = "Examining the SHAP values for a random forest model classifying Palmer penguin species.
 The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
 The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at \url{https://vimeo.com/666431172}.)
-
-```{r casepenguins, echo=F, out.width="100%", fig.cap = '(ref:casepenguins-cap)'}
+(The animation of the radial tour is at https://vimeo.com/666431172.)"}
 knitr::include_graphics("./figures/case_penguins.png")
@@
 
-(ref:casepenguinsblfl-cap) Checking what is learned from the cheem viewer.
+```{r casepenguinsblfl, echo=F, out.width="100%", fig.align='center', fig.cap = "Checking what is learned from the cheem viewer.
 This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI.
 A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.
-
-```{r casepenguinsblfl, echo=F, out.width="100%", fig.cap = '(ref:casepenguinsblfl-cap)'}
+The PI has an unusually small `fl` length which is why it is confused with a Chinstrap."}
 knitr::include_graphics("./figures/case_penguins_BlFl.png")
@@
 
-(ref:casechocolates-cap) Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+```{r casechocolates, echo=F, out.width="100%", fig.cap = "Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
 From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
@@
 This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)
-
-```{r casechocolates, echo=F, out.width="100%", fig.cap = '(ref:casechocolates-cap)'}
+(A video of the tour animation can be found at https://vimeo.com/666431143.)"}
 knitr::include_graphics("./figures/case_chocolates.png")
@@
 
-(ref:casechocolatesinverse-cap) Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
@@
 the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)
-
-```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = '(ref:casechocolatesinverse-cap)'}
+(A video of the tour animation can be found at https://vimeo.com/666431148.)"}
 knitr::include_graphics("./figures/case_chocolates_inverse.png")
@@
 
-(ref:casefifa-cap) Exploring the wages relative to skill measurements in the FIFA 2020 data.
+```{r casefifa, echo=F, out.width = "90%", fig.cap = "Exploring the wages relative to skill measurements in the FIFA 2020 data.
 Star offensive player (L.
@@
 The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)
-
-```{r casefifa, echo=F, out.width = "90%", fig.cap = '(ref:casefifa-cap)'}
+(A video of the animated radial tour can be found at https://vimeo.com/666431163.)"}
 knitr::include_graphics("./figures/case_fifa.png")
@@
 
-(ref:caseames-cap) Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
+```{r caseames, echo=F, out.width="90%", fig.cap = "Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
 The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
@@
 That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at \url{https://vimeo.com/666431134}.)
-```{r caseames, echo=F, out.width="90%", fig.cap = '(ref:caseames-cap)'}
+(A video showing the animation is at https://vimeo.com/666431134.)"}
 knitr::include_graphics("./figures/case_ames2018.png")
@@
 
-An implementation is provided in the open-source __R__ package __cheem__, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
+An implementation is provided in the open-source __R__ package __cheem__, available on CRAN at https://CRAN.R-project.org/package=cheem.
 Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
@@
 - A version of the cheem viewer shiny app can be directly accessed at
-\url{https://ebsmonash.shinyapps.io/cheem_initial/}.
-- The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
-- Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
+https://ebsmonash.shinyapps.io/cheem_initial/.
+- The development version of the package is available at https://github.com/nspyrison/cheem, and
+- Documentation of the package can be found at https://nspyrison.github.io/cheem/.
 

### 2023-05-16 jss/cheem_paper.tex
figure captions fixed

@@
 
-(ref:casepenguins-cap) Examining the SHAP values for a random forest
-model classifying Palmer penguin species. The PI is a Gentoo (purple)
-penguin that is misclassified as a Chinstrap (orange), marked as an
-asterisk in (a) and the dashed vertical line in (b). The radial view
-shows varying the contribution of \texttt{fl} from the initial
-attribution projection (b, left), which produces a linear combination
-where the PI is more probably (higher density value) a Chinstrap than a
-Gentoo (b, right). (The animation of the radial tour is at
-\url{https://vimeo.com/666431172}.)
-
 \begin{CodeChunk}
@@
 
-\caption[(ref:casepenguins-cap)]{(ref:casepenguins-cap)}\label{fig:casepenguins}
+\caption[Examining the SHAP values for a random forest model classifying Palmer penguin species]{Examining the SHAP values for a random forest model classifying Palmer penguin species.
+The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
+The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
+(The animation of the radial tour is at https://vimeo.com/666431172.)}\label{fig:casepenguins}
 \end{figure}
@@
 
-(ref:casepenguinsblfl-cap) Checking what is learned from the cheem
-viewer. This is a plot of flipper length (\texttt{fl}) and bill length
-(\texttt{bl}), where an asterisk highlights the PI. A Gentoo (purple)
-misclassified as a Chinstrap (orange). The PI has an unusually small
-\texttt{fl} length which is why it is confused with a Chinstrap.
-
 \begin{CodeChunk}
@@
 
-\caption[(ref:casepenguinsblfl-cap)]{(ref:casepenguinsblfl-cap)}\label{fig:casepenguinsblfl}
+\caption[Checking what is learned from the cheem viewer]{Checking what is learned from the cheem viewer.
+This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI.
+A Gentoo (purple) misclassified as a Chinstrap (orange).
+The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
 \end{figure}
@@
 
-(ref:casechocolates-cap) Examining the local explanation for a PI which
-is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more
-like dark than milk, which suggests that the local explanation does not
-help understand the prediction for this observation. So, the
-contribution of Sugar is varied---reducing it corresponds primarily with
-an increased magnitude from Fiber. When Sugar is zero, Fiber contributes
-strongly towards the left. In this view, the PI is closer to the bulk of
-the milk chocolates, suggesting that the prediction put a lot of
-importance on Fiber. This chocolate is a rare dark chocolate without any
-Fiber leading to it being mistaken for a milk chocolate. (A video of the
-tour animation can be found at \url{https://vimeo.com/666431143}.)
-
 \begin{CodeChunk}
@@
 
-\caption[(ref:casechocolates-cap)]{(ref:casechocolates-cap)}\label{fig:casechocolates}
+\caption[Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
+So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
+When Sugar is zero, Fiber contributes strongly towards the left.
+In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
+This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
+(A video of the tour animation can be found at https://vimeo.com/666431143.)}\label{fig:casechocolates}
 \end{figure}
@@
 
-(ref:casechocolatesinverse-cap) Examining the local explanation for a PI
-which is milk (green) chocolate incorrectly predicted to be dark
-(orange). In the attribution projection, the PI could be either milk or
-dark. Sodium and Fiber have the largest differences in attributed
-variable importance, with low values relative to other milk chocolates.
-The lack of importance attributed to these variables is suspected of
-contributing to the mistake, so the contribution of Sodium is varied. If
-Sodium had a larger contribution to the prediction (like in this view).
-the PI would look more like other milk chocolates. (A video of the tour
-animation can be found at \url{https://vimeo.com/666431148}.)
-
 \begin{CodeChunk}
@@
 
-\caption[(ref:casechocolatesinverse-cap)]{(ref:casechocolatesinverse-cap)}\label{fig:casechocolatesinverse}
+\caption[Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+In the attribution projection, the PI could be either milk or dark.
+Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates.
+The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
+If Sodium had a larger contribution to the prediction (like in this view).
+the PI would look more like other milk chocolates.
+(A video of the tour animation can be found at https://vimeo.com/666431148.)}\label{fig:casechocolatesinverse}
 \end{figure}
@@
 
-(ref:casefifa-cap) Exploring the wages relative to skill measurements in
-the FIFA 2020 data. Star offensive player (L. Messi) is the PI, and he
-is compared with a top defensive player (V. van Dijk). The attribution
-projection is shown on the left, and it can be seen that this
-combination of variables produces a view where Messi has very high
-predicted (and observed) wages. Defense (\texttt{def}) is the chosen
-variable to vary. It starts very low, and Messi's predicted wages
-decrease dramatically as its contribution increases (right plot). The
-increased contribution in defense comes at the expense of offensive and
-reaction skills. The interpretation is that Messi's high wages are most
-attributable to his offensive and reaction skills, as initially provided
-by the local explanation. (A video of the animated radial tour can be
-found at \url{https://vimeo.com/666431163}.)
-
 \begin{CodeChunk}
@@
 
-\caption[(ref:casefifa-cap)]{(ref:casefifa-cap)}\label{fig:casefifa}
+\caption[Exploring the wages relative to skill measurements in the FIFA 2020 data]{Exploring the wages relative to skill measurements in the FIFA 2020 data.
+Star offensive player (L.
+Messi) is the PI, and he is compared with a top defensive player (V.
+van Dijk).
+The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
+Defense (`def`) is the chosen variable to vary.
+It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
+The increased contribution in defense comes at the expense of offensive and reaction skills.
+The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
+(A video of the animated radial tour can be found at https://vimeo.com/666431163.)}\label{fig:casefifa}
 \end{figure}
@@
 
-(ref:caseames-cap) Exploring an observation with an extreme residual as
-the PI in relation to an observation with an accurate prediction for a
-similarly priced house in a random forest fit to the Ames housing data.
-The local explanation indicates a sizable attribution to Lot Area
-(\texttt{LtA}), while the CI has minimal attribution to this variable.
-The PI has a higher predicted value than the CI in the attribution
-projection. Reducing the contribution of Lot Area brings these two
-prices in line. This suggests that if the model did not value Lot Area
-so highly for this observation, then the observed sales price would be
-quite similar. That is, the large residual is due to a lack of factoring
-in the Lot Area for the prediction of PI's sales price. (A video showing
-the animation is at \url{https://vimeo.com/666431134}.)
-
 \begin{CodeChunk}
@@
 
-\caption[(ref:caseames-cap)]{(ref:caseames-cap)}\label{fig:caseames}
+\caption[Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data]{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
+The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
+The PI has a higher predicted value than the CI in the attribution projection.
+Reducing the contribution of Lot Area brings these two prices in line.
+This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar.
+That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
+(A video showing the animation is at https://vimeo.com/666431134.)}\label{fig:caseames}
 \end{figure}

### 2023-05-20 motivating_letter.Rmd
modified motivating letter

@@
 
+Editorial Team, \
+Journal of Statistical Software,
+
 Dear Editors,
+\bigskip
 
-Please consider our paper titled "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections" for publication in the Journal of Computational and Graphical Statistics.
+Please consider our paper titled "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections" for publication in the Journal of Statistical Software.
+
+We feel that JSS is a good fit for this work because it describes new software for visualising local explanations diagnostics for understanding model fits.
+This work can also be used to assess the reliability of the information reported by these diagnostics.
+
 We'd like to acknowledge that it was previously submitted to the Journal of Machine Learning Research.
@@
 However, some useful feedback was provided, and below we have described how we have improved the paper based on their reviews.
-
-We feel that JCGS is a good fit for this work because it describes useful new visual approaches for understanding model fits based on the relatively new local explanations diagnostics.
-This work can also be used to assess the reliability of the information reported by these diagnostics.
-
-Here is an explanation of how the paper was improved based on the review from JMLR:
-
-*(a) The writing of the paper is informal, lacks clarity and depth in many parts that are critical to understanding the method proposed, and requires significant improvement.
-The method of radial tour, which (apparently?) seems to be the most important contribution of the paper, is practically not explained at all.
-Up to Section 4, when the description of the interactive visualization tool starts, all we get for the explanation of the radial tour is one sentence, which is drastically not enough.
-This is observable throughout the text as a whole, and it is very hard to point exactly where and when it happens.
-In general, the paper's writing is informal, lacking the depth necessary to make it reproducible and auditable.
-Without clear, detailed explanations of the methods used, it becomes hard to accept it in a deeply technical journal such as JMLR.*
-
-*(b) The paper does not seem to fit the scope of the journal.
-Even though the idea behind the radial tour seems interesting (although it is hard to know because the authors have skipped explaining it in details), it seems that one of the most important contributions of the paper is actually an interactive visualization tool.
-As such, I believe that JMLR is not the right venue for this paper.
-The authors should've chosen a visualization venue for this, where they'll be able to get a much more interested audience and much more relevant feedback too.
-I'd recommend IEEE VIS, IEEE TVCG, EuroVis, Computer Graphics Forum, Computers & Graphics, Information Visualization journal, or the PacificVis conference.*
-
-* *Abstract: While it is not wrong, not all XAI techniques use local explanations; some definitely do. The wording here could be a bit more accurate.*
-
-* *The introduction seems to end a bit abruptly. Maybe a couple of sentences about the results and outcomes of the research?*
-
-* *Page 3: "Consider a highly nonlinear model." I feel like this (and the reasoning that follows) is a bit of an informal statement.
-Although intuitively I agree with the authors, is there any objective (or at least semi-formal) way to explain this problem?
-What is highly non-linear (vs.
-just linear)?
-Why, exactly, is it that it is so hard to interpret nonlinear models?
-That would improve the argumentation here quite a bit.*
-
-* *"The attribution of feature importance depends on the sequence of the included features." This sentence feels out of place.*
-
-* *"modified breakdown plots" -- Since this is taking an important place in the paper, with a figure attached to it, I think it needs to be better explained.
-What are exactly modified breakdown plots?*
-
-* *"The magnitude of the contributions depends on the sequence in which they appear." -- This is unclear.
-The figure has no label for the horizontal axes, which makes it very hard to interpret.
-And, in case, the sequence alters the magnitude of the contributions, then it should be explained how the sequence is determined.
-Right now, this explanation does not help to make the visualization or the method presented in Figure 1 easier to understand.*
-
-* *"The attribution space corresponds to the local explanations for each instance; feature importance in the
-vicinity of the instance." -- I think this needs to be better explained.
-The concept of the attribution space remains obscure to me.
-It seems to be a very interesting concept and very important to the paper.*
-
-* *"The projection attribution of the primary instance (PI) is examined and typically viewed with an optional
-comparison instance (CI)." -- The concept of PI and CI remains confusing to me. Needs a clearer explanation.*
-
-* *Section 4.2 as a whole -- This explanation is overly confusing and informal/subjective. It needs to be significantly revised to become more fluid and clear.*
-
-* *I don't believe that sections 4.7 and 4.8 are needed.*
+Here is an explanation of how the paper was improved based on the review from JMLR is available at https://github.com/nspyrison/cheem_paper/issues/2.
+We have also modified it to fit the style of JSS.
+\bigskip
+\bigskip
 
@@
 
-Nick Spyrison
+Nick Spyrison, PhD
+
+\url{https://github.com/nspyrison/}\
+spyrison@gmail.com

### 2023-05-20 motivating_letter.Rmd
US spelling

@@
 
-We feel that JSS is a good fit for this work because it describes new software for visualising local explanations diagnostics for understanding model fits.
+We feel that JSS is a good fit for this work because it describes new software for visualizing local explanations diagnostics for understanding model fits.
 This work can also be used to assess the reliability of the information reported by these diagnostics.
@@
 We'd like to acknowledge that it was previously submitted to the Journal of Machine Learning Research.
-The primary reason was that the editor felt it had too much visualisation to be a good fit for the journal.
+The primary reason was that the editor felt it had too much visualization to be a good fit for the journal.
 However, some useful feedback was provided, and below we have described how we have improved the paper based on their reviews.

### 2023-05-21 jss/cheem_paper.Rmd
abstract substantially changed and improved

@@
 abstract: >
-    The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
-This trade-off has led to the emergence of eXplainable AI (XAI).
-XAI attempts to shed light on how models use predictors to arrive at a prediction.
-"Local variable attributions", which provide a point estimate of the linear variable importance in the vicinity of one observation, are one local XAI method.
-These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface as well as understand the structure of local explanations for observations, whether outlier are present, whether segments can be identified.
-Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
+    The increased predictive power of machine learning models comes at the cost of increased complexity and loss of interpretability, particularly in comparison to parametric statistical models.
+This trade-off has led to the emergence of eXplainable AI (XAI) which provides methods, such as local explanations (LEs) and local variable attributions (LVAs), to shed light on how a model use predictors to arrive at a prediction.
+These provide a point estimate of the linear variable importance in the vicinity of a single observation.
+However, LVAs tend not to effectively handle association between predictors.
+To understand how the interaction between predictors affects the variable importance estimate, we can convert LVAs into linear projections and use the radial tour.
+This is also useful for learning how a model has made a mistake, or the effect of outliers, or clustering of observations.
+The approach is llustrated with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
 The methods are implemented in the R package cheem, available on CRAN.

### 2023-05-21 jss/cheem_paper.tex
abstract substantially changed and improved

@@
 \Abstract{
-The increased predictive power of nonlinear models comes at the cost of
-interpretability of its terms. This trade-off has led to the emergence
-of eXplainable AI (XAI). XAI attempts to shed light on how models use
-predictors to arrive at a prediction. ``Local variable attributions'',
-which provide a point estimate of the linear variable importance in the
-vicinity of one observation, are one local XAI method. These can be
-considered to be linear projections and can be further explored to
-understand the interactions between variables used to make predictions
-across the predictive model surface as well as understand the structure
-of local explanations for observations, whether outlier are present,
-whether segments can be identified. Here we describe interactive linear
-interpolation used for exploration at any observation and illustrate
+The increased predictive power of machine learning models comes at the
+cost of increased complexity and loss of interpretability, particularly
+in comparison to parametric statistical models. This trade-off has led
+to the emergence of eXplainable AI (XAI) which provides methods, such as
+local explanations (LEs) and local variable attributions (LVAs), to shed
+light on how a model use predictors to arrive at a prediction. These
+provide a point estimate of the linear variable importance in the
+vicinity of a single observation. However, LVAs tend not to effectively
+handle association between predictors. To understand how the interaction
+between predictors affects the variable importance estimate, we can
+convert LVAs into linear projections and use the radial tour. This is
+also useful for learning how a model has made a mistake, or the effect
+of outliers, or clustering of observations. The approach is llustrated
 with examples from categorical (penguin species, chocolate types) and

### 2023-05-21 jss/cheem_paper.Rmd
added paragraph on available R packages

@@
 Explainable Artificial Intelligence (XAI) is an emerging field of research focused on methods for the interpreting of models [@adadi_peeking_2018; @arrieta_explainable_2020].
-A class of techniques called _local explanations_, provide methods to approximate linear variable importance at the location of each observation or the predictions at a specific point in the data domain.
+A class of techniques, called _local explanations_ (LEs), provide methods to approximate linear variable importance, called local variable attributions (LVAs), at the location of each observation or the predictions at a specific point in the data domain.
 Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
-There are common approaches for visualizing high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
+There are common approaches for visualizing high-dimensional data as a whole, but what is needed are new approaches for viewing these individual LVAs relative to the whole.
 
 <!-- Data visualization tours -->
-For multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] of linear data projections onto a lower-dimensional space, could be an element of XAI, complementing local explanations.
+For multivariate data visualization, a _tour_ [@asimov_grand_1985; @buja_grand_1986; @lee_state_2021] of linear data projections onto a lower-dimensional space, could be an element of XAI, complementing LVAs.
 <!--Tours are viewed as an animation over minor changes to the projection basis.
@@
 In a _manual_ or radial tour [@cook_manual_1997; @spyrison_spinifex_2020], the path of linear projections is defined by changing the contribution of a selected variable.
-We propose to use this to scrutinize a local explanation.
+We propose to use this to scrutinize the LVAs.
 <!-- Additional interactive elements in a graphical user interface should allow the user to identify an observation of interest and then explore its local explanation by changing variable contribution with the radial tour.
@@
 The remainder of this paper is organized as follows.
-Section \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
+Section \ref{sec:explanations} covers the background of the LEs and the traditional visuals produced.
 Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
@@
 Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative response variables.
-These show how the local explanations can be used to get an overview of the model's use of predictors, and to investigate errors in the model predictions.
+These show how the LVAs can be used to get an overview of the model's use of predictors, and to investigate errors in the model predictions.
 Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
@@
 <!-- Reminder of local explanation -->
-Local explanations shed light on nonlinear model fits by estimating linear variable importance in the vicinity of a single observation.  There are many approaches for calculating local explanations.
+LVAs shed light on machine learning model fits by estimating linear variable importance in the vicinity of a single observation.  There are many approaches for calculating LVAs.
 <!-- Taxonomy of local explanations -->
@@
 It includes a large number of model-specific explanations such as deepLIFT [@shrikumar_not_2016; @shrikumar_learning_2017], a popular recursive method for estimating importance in neural networks.
-There are fewer model-agnostic methods, of which LIME, [@ribeiro_why_2016] SHaply Additive exPlanations (SHAP), [@lundberg_unified_2017], are popular.
+There are fewer model-agnostic methods, of which LIME [@ribeiro_why_2016] and SHaply Additive exPlanations (SHAP) [@lundberg_unified_2017], are popular.
 
@@
 <!-- SHAP -->
-We will be focusing on SHAP values in this this paper, but the approach is applicable for any method used to calculate the local explanations.
+We will be focusing on SHAP values in this this paper, but the approach is applicable for any method used to calculate the LVAs.
 SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
 The term "SHAP" refers to @shapley_value_1953's method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
-@strumbelj_efficient_2010 introduced SHAP for local explanations in machine learning models.
+@strumbelj_efficient_2010 introduced SHAP for LEs in machine learning models.
 Variable importance can depend on the sequence in which variables are entered into the model fitting process, thus for any sequence we get a set of variable contribution values for a single observation.
@@
 
-The approach is related to partial dependence plots (see for example @molnar_interpretable_2020), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
+The approach is related to partial dependence plots (see for example, chapter 8 of @molnar2022), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
 Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
@@
 
+There are numerous R packages currently available which provide functions for computing SHAP values, including `fastshap` [@fastshap], `kernelshap` [@kernelshap], `shapr` [@shapr], `shapviz` [@shapviz], `PPtreeregViz` [@PPtreeregViz], and the package `DALEX` has many resources [@biecek_dalex_2018].
+There are many more packages only available through Github, like `treeshap` [kominsarczyk_treeshap_2021] that is used for this work.
+@molnar2022 provides good explanations of the different methods and how to apply them to different models.
+
 # Tours and the Radial Tour {#sec:tour}
@@
 
-To explore the local explanations, coordinated views [@roberts_state_2007] [also known as ensemble graphics, @unwin_ensemble_2018] are provided in the _cheem viewer_ application.
-There are two primary plots: the __global view__ to give the context of all of the SHAP values and the __radial tour view__ to explore the local explanations with user-controlled rotation.
+To explore the LVAs, coordinated views [@roberts_state_2007] [also known as ensemble graphics, @unwin_ensemble_2018] are provided in the _cheem viewer_ application.
+There are two primary plots: the __global view__ to give the context of all of the SHAP values and the __radial tour view__ to explore the LVAs with user-controlled rotation.
 There are numerous user inputs, including variable selection for the radial tour and observation selection for making comparisons.
@@
 <!-- Local explanations as par coords -->
-The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
+The LVAs for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
 These are depicted as a vertical parallel coordinate plot [@ocagne_coordonnees_1885].
@@
 <!-- Note on time of execution -->
-The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
+The time to preprocess the data will vary significantly with the complexity of the model and the LE.
 For reference, the FIFA data contained 5000 observations of nine explanatory variables took 2.5 seconds to fit a random forest model of modest hyperparameters.
@@
 PCA and statistics of the variables and attributions took 2.8 seconds.
-These run times were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
+These run times were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the LVA.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
@@
 
-```{r casechocolates, echo=F, out.width="100%", fig.cap = "Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
+```{r casechocolates, echo=F, out.width="100%", fig.cap = "Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the LVA does not help understand the prediction for this observation.
 So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
@@
 
-```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+```{r casechocolatesinverse, echo=F, out.width="100%", fig.cap = "Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
@@
 The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
+The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the LVA.
 (A video of the animated radial tour can be found at https://vimeo.com/666431163.)"}
@@
 
-Figure \ref{fig:casefifa} tests the support of the local explanation.
+Figure \ref{fig:casefifa} tests the support of the LVA.
 Offensive and reaction skills (`off` and `rct`) are both crucial to explaining a star offensive player.
@@
 ```{r caseames, echo=F, out.width="90%", fig.cap = "Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
-The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
+The LVA indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
 The PI has a higher predicted value than the CI in the attribution projection.
@@
 There is a clear need to extend the interpretability of black box models.
-With techniques such as SHAP, LIME, Break-down, one can calculate local explanations, i.e.
+With techniques such as SHAP, LIME, Break-down, one can calculate LEs, i.e.
 for every observation in the data.
 These techniques quantify for each observation how strongly particular variables affect the model's predictions.
-Surprisingly few techniques allow us to understand the global distribution of these local explanations.
-Unsupervised data exploration techniques applied to data show how useful they are for identifying outliers, identifying segments of observations or discovering correlations between variables.
+Surprisingly few techniques allow us to understand the global distribution of these LEs.
+Unsupervised data exploration techniques applied to data show how useful they are for identifying outliers, identifying clusters of observations or discovering correlations between variables.
 All of these tasks can be performed for a set of explanations.
 
-To address this challenge in this paper provides a technique that builds on local explanations to explore the variable importance local to an observation.
-The local explanations of an attribution projection from which variable contributions are varied using a radial tour.
+To address this challenge in this paper provides a technique that builds on LEs to explore the variable importance local to an observation.
+The LVA is converted into an attribution projection from which variable contributions are varied using a radial tour.
 Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular variables.
@@
 
-This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
-Local explanations focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
+This approach has been illustrated using four data examples of random forest models with the tree SHAP LVA.
+LEs focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
 They can also form the basis of learning how the model has got it wrong, with when the observation is misclassified or has a large residual.
@@
 An implementation is provided in the open-source __R__ package __cheem__, available on CRAN at https://CRAN.R-project.org/package=cheem.
-Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
-The local explanations need to be pre-computed and uploaded.
+Example data sets are provided, and you can upload your data after model fitting and computing the LVAs.
+The LVAs need to be pre-computed and uploaded.
 Examples show how to do this for tree SHAP values, using  __treeshap__ (tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__  @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively).
@@
 Both views are created first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive `html` widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and the free ebook, _Explanatory Model Analysis_ [@biecek_explanatory_2021] were a boon to understanding local explanations and how to apply them.
+__DALEX__ [@biecek_dalex_2018] and  _Explanatory Model Analysis_ [@biecek_explanatory_2021] are helpfule for understanding LEs and how to apply them.
 
@@
 
-Follow the examples provided with the package to compute the local explanation (using `?cheem_ls`).
+Follow the examples provided with the package to compute the LVAs (using `?cheem_ls`).
 The application expects the output returned by `cheem_ls()`, saved to an `rds` file with `saveRDS()` to be uploaded.

### 2023-05-21 jss/cheem_paper.tex
added paragraph on available R packages

@@
 \citep{adadi_peeking_2018, arrieta_explainable_2020}. A class of
-techniques called \emph{local explanations}, provide methods to
-approximate linear variable importance at the location of each
-observation or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to comprehensively
-visualize them to understand a model. There are common approaches for
-visualizing high-dimensional data as a whole, but what is needed are new
-approaches for viewing these individual local explanations, in relation
-to the whole.
+techniques, called \emph{local explanations} (LEs), provide methods to
+approximate linear variable importance, called local variable
+attributions (LVAs), at the location of each observation or the
+predictions at a specific point in the data domain. Because these are
+point-specific, it is challenging to comprehensively visualize them to
+understand a model. There are common approaches for visualizing
+high-dimensional data as a whole, but what is needed are new approaches
+for viewing these individual LVAs relative to the whole.
 
@@
 data projections onto a lower-dimensional space, could be an element of
-XAI, complementing local explanations. Applying tours to model
-interpretation is recommended in \citet{wickham_visualizing_2015}
-primarily to examine the fitted model in the space of the data.
-\citet{cook_interactive_2007} describe the use of tours for exploring
-classification boundaries and model diagnostics
+XAI, complementing LVAs. Applying tours to model interpretation is
+recommended in \citet{wickham_visualizing_2015} primarily to examine the
+fitted model in the space of the data. \citet{cook_interactive_2007}
+describe the use of tours for exploring classification boundaries and
+model diagnostics
 \citep{Caragea2008, lee_pptree_2013, da_silva_projection_2021}. There
@@
 projections is defined by changing the contribution of a selected
-variable. We propose to use this to scrutinize a local explanation. This
-approach could be considered to be a counter-factual, what-if analysis,
-such as \emph{ceteris paribus} (``other things held constant'') profiles
+variable. We propose to use this to scrutinize the LVAs. This approach
+could be considered to be a counter-factual, what-if analysis, such as
+\emph{ceteris paribus} (``other things held constant'') profiles
 \citep{biecek_ceterisparibus_2020}.
@@
 The remainder of this paper is organized as follows. Section
-\ref{sec:explanations} covers the background of the local explanation
-and the traditional visuals produced. Section \ref{sec:tour} explains
-the tours and particularly the radial manual tour. Section
-\ref{sec:cheemviewer} discusses the visual layout in the graphical user
-interface and how it facilitates analysis, data pre-processing, and
-package infrastructure. Illustrations are provided in Section
-\ref{sec:casestudies} for a range of supervised learning tasks with
-categorical and quantitative response variables. These show how the
-local explanations can be used to get an overview of the model's use of
-predictors, and to investigate errors in the model predictions. Section
-\ref{sec:cheemdiscussion} concludes with a summary of the insights
-gained. The methods are implemented in the \textbf{R} package
-\textbf{cheem}.
+\ref{sec:explanations} covers the background of the LEs and the
+traditional visuals produced. Section \ref{sec:tour} explains the tours
+and particularly the radial manual tour. Section \ref{sec:cheemviewer}
+discusses the visual layout in the graphical user interface and how it
+facilitates analysis, data pre-processing, and package infrastructure.
+Illustrations are provided in Section \ref{sec:casestudies} for a range
+of supervised learning tasks with categorical and quantitative response
+variables. These show how the LVAs can be used to get an overview of the
+model's use of predictors, and to investigate errors in the model
+predictions. Section \ref{sec:cheemdiscussion} concludes with a summary
+of the insights gained. The methods are implemented in the \textbf{R}
+package \textbf{cheem}.
 
@@
 
-Local explanations shed light on nonlinear model fits by estimating
-linear variable importance in the vicinity of a single observation.
-There are many approaches for calculating local explanations. A
-comprehensive summary of the taxonomy of currently available methods is
-provided in Figure 6 by \citet{arrieta_explainable_2020}. It includes a
-large number of model-specific explanations such as deepLIFT
+LVAs shed light on machine learning model fits by estimating linear
+variable importance in the vicinity of a single observation. There are
+many approaches for calculating LVAs. A comprehensive summary of the
+taxonomy of currently available methods is provided in Figure 6 by
+\citet{arrieta_explainable_2020}. It includes a large number of
+model-specific explanations such as deepLIFT
 \citep{shrikumar_not_2016, shrikumar_learning_2017}, a popular recursive
 method for estimating importance in neural networks. There are fewer
-model-agnostic methods, of which LIME, \citep{ribeiro_why_2016} SHaply
-Additive exPlanations (SHAP), \citep{lundberg_unified_2017}, are
+model-agnostic methods, of which LIME \citep{ribeiro_why_2016} and
+SHaply Additive exPlanations (SHAP) \citep{lundberg_unified_2017}, are
 popular.
@@
 We will be focusing on SHAP values in this this paper, but the approach
-is applicable for any method used to calculate the local explanations.
-SHAP calculates the variable contributions of one observation by
-examining the effect of other variables on the predictions. The term
-``SHAP'' refers to \citet{shapley_value_1953}'s method to evaluate an
-individual's contribution in cooperative games by assessing this
-player's performance in the presence or absence of other players.
-\citet{strumbelj_efficient_2010} introduced SHAP for local explanations
-in machine learning models. Variable importance can depend on the
-sequence in which variables are entered into the model fitting process,
-thus for any sequence we get a set of variable contribution values for a
-single observation. These values will add up to the difference between
-the fitted value for the observation, and the average fitted value for
-all observations. Using all possible sequence, or permutation, gives
+is applicable for any method used to calculate the LVAs. SHAP calculates
+the variable contributions of one observation by examining the effect of
+other variables on the predictions. The term ``SHAP'' refers to
+\citet{shapley_value_1953}'s method to evaluate an individual's
+contribution in cooperative games by assessing this player's performance
+in the presence or absence of other players.
+\citet{strumbelj_efficient_2010} introduced SHAP for LEs in machine
+learning models. Variable importance can depend on the sequence in which
+variables are entered into the model fitting process, thus for any
+sequence we get a set of variable contribution values for a single
+observation. These values will add up to the difference between the
+fitted value for the observation, and the average fitted value for all
+observations. Using all possible sequence, or permutation, gives
 multiple values for each variable, which are averaged to get the SHAP
@@
 
-The approach is related to partial dependence plots (see for example
-\citet{molnar_interpretable_2020}), used to explain the effect of a
+The approach is related to partial dependence plots (see for example,
+chapter 8 of \citet{molnar2022}), used to explain the effect of a
 variable by predicting the response for a range of values on this
@@
 
+There are numerous R packages currently available which provide
+functions for computing SHAP values, including \texttt{fastshap}
+\citep{fastshap}, \texttt{kernelshap} \citep{kernelshap}, \texttt{shapr}
+\citep{shapr}, \texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
+\citep{PPtreeregViz}, and the package \texttt{DALEX} has many resources
+\citep{biecek_dalex_2018}. There are many more packages only available
+through Github, like \texttt{treeshap}
+{[}kominsarczyk\_treeshap\_2021{]} that is used for this work.
+\citet{molnar2022} provides good explanations of the different methods
+and how to apply them to different models.
+
 \hypertarget{sec:tour}{%
@@
 
-To explore the local explanations, coordinated views
-\citep{roberts_state_2007} \citep[also known as ensemble
-graphics,][]{unwin_ensemble_2018} are provided in the \emph{cheem
-viewer} application. There are two primary plots: the \textbf{global
-view} to give the context of all of the SHAP values and the
-\textbf{radial tour view} to explore the local explanations with
+To explore the LVAs, coordinated views \citep{roberts_state_2007}
+\citep[also known as ensemble graphics,][]{unwin_ensemble_2018} are
+provided in the \emph{cheem viewer} application. There are two primary
+plots: the \textbf{global view} to give the context of all of the SHAP
+values and the \textbf{radial tour view} to explore the LVAs with
 user-controlled rotation. There are numerous user inputs, including
@@
 
-The local explanations for all observations are normalized (sum of
-squares equals 1), and thus, the relative importance of variables can be
-compared across all observations. These are depicted as a vertical
-parallel coordinate plot \citep{ocagne_coordonnees_1885}. (The SHAP
-values of the PI and CI are shown as dashed and dotted lines,
-respectively.) One should obtain a sense of the overall importance of
-variables from this plot. The more important variables will have larger
-values, and in the case of classification tasks variables which have
-different magnitude for different classes are more globally important.
-For example, Figure \ref{fig:classificationcase}e suggests that
-\texttt{bl} is important for distinguishing the green class from the
-other two. For regression, one might generally observe which variables
-have low values for all observations (not important). For example,
-\texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have
-a range of high and low values (e.g., \texttt{off}, \texttt{def}),
-suggesting they are important for some observations and not important
-for others.
+The LVAs for all observations are normalized (sum of squares equals 1),
+and thus, the relative importance of variables can be compared across
+all observations. These are depicted as a vertical parallel coordinate
+plot \citep{ocagne_coordonnees_1885}. (The SHAP values of the PI and CI
+are shown as dashed and dotted lines, respectively.) One should obtain a
+sense of the overall importance of variables from this plot. The more
+important variables will have larger values, and in the case of
+classification tasks variables which have different magnitude for
+different classes are more globally important. For example, Figure
+\ref{fig:classificationcase}e suggests that \texttt{bl} is important for
+distinguishing the green class from the other two. For regression, one
+might generally observe which variables have low values for all
+observations (not important). For example, \texttt{BMI} and \texttt{pwr}
+in Figure \ref{fig:regressioncase}e, have a range of high and low values
+(e.g., \texttt{off}, \texttt{def}), suggesting they are important for
+some observations and not important for others.
 
@@
 The time to preprocess the data will vary significantly with the
-complexity of the model and local explanation. For reference, the FIFA
-data contained 5000 observations of nine explanatory variables took 2.5
+complexity of the model and the LE. For reference, the FIFA data
+contained 5000 observations of nine explanatory variables took 2.5
 seconds to fit a random forest model of modest hyperparameters.
@@
 modern laptop, but suffice to say that most of the time will be spent on
-the local attribution. An increase in model complexity or data
-dimensionality will quickly become an obstacle. Its reduced
-computational complexity makes tree SHAP an excellent candidate to
-start. Alternatively, the package \textbf{fastshap} claims extremely low
-run times, attributed to fewer calls to the prediction function, partial
-implementation in C++, and efficient use of logical subsetting,
-\citet{greenwell_fastshap_2020}.
+the LVA. An increase in model complexity or data dimensionality will
+quickly become an obstacle. Its reduced computational complexity makes
+tree SHAP an excellent candidate to start. Alternatively, the package
+\textbf{fastshap} claims extremely low run times, attributed to fewer
+calls to the prediction function, partial implementation in C++, and
+efficient use of logical subsetting, \citet{greenwell_fastshap_2020}.
 
@@
 
-\caption[Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
+\caption[Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the LVA does not help understand the prediction for this observation.
 So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
@@
 
-\caption[Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+\caption[Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
@@
 The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
+The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the LVA.
 (A video of the animated radial tour can be found at https://vimeo.com/666431163.)}\label{fig:casefifa}
@@
 
-Figure \ref{fig:casefifa} tests the support of the local explanation.
-Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both
-crucial to explaining a star offensive player. If either of them were
-rotated out, the other would be rotated into the frame, maintaining a
-far-right position. However, increasing the contribution of a variable
-with low importance would rotate both variables out of the frame.
+Figure \ref{fig:casefifa} tests the support of the LVA. Offensive and
+reaction skills (\texttt{off} and \texttt{rct}) are both crucial to
+explaining a star offensive player. If either of them were rotated out,
+the other would be rotated into the frame, maintaining a far-right
+position. However, increasing the contribution of a variable with low
+importance would rotate both variables out of the frame.
 
@@
 \caption[Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data]{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
-The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
+The LVA indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
 The PI has a higher predicted value than the CI in the attribution projection.
@@
 models. With techniques such as SHAP, LIME, Break-down, one can
-calculate local explanations, i.e.~for every observation in the data.
-These techniques quantify for each observation how strongly particular
-variables affect the model's predictions. Surprisingly few techniques
-allow us to understand the global distribution of these local
-explanations. Unsupervised data exploration techniques applied to data
-show how useful they are for identifying outliers, identifying segments
-of observations or discovering correlations between variables. All of
-these tasks can be performed for a set of explanations.
+calculate LEs, i.e.~for every observation in the data. These techniques
+quantify for each observation how strongly particular variables affect
+the model's predictions. Surprisingly few techniques allow us to
+understand the global distribution of these LEs. Unsupervised data
+exploration techniques applied to data show how useful they are for
+identifying outliers, identifying clusters of observations or
+discovering correlations between variables. All of these tasks can be
+performed for a set of explanations.
 
 To address this challenge in this paper provides a technique that builds
-on local explanations to explore the variable importance local to an
-observation. The local explanations of an attribution projection from
-which variable contributions are varied using a radial tour. Several
-diagnostic plots are provided to assist with understanding the
-sensitivity of the prediction to particular variables. A global view
-shows the data space, explanation space, and residual plot. The user can
-interactively select observations to compare, contrast, and study
-further. Then the radial tour is used to explore the variable
-sensitivity identified by the attribution projection.
+on LEs to explore the variable importance local to an observation. The
+LVA is converted into an attribution projection from which variable
+contributions are varied using a radial tour. Several diagnostic plots
+are provided to assist with understanding the sensitivity of the
+prediction to particular variables. A global view shows the data space,
+explanation space, and residual plot. The user can interactively select
+observations to compare, contrast, and study further. Then the radial
+tour is used to explore the variable sensitivity identified by the
+attribution projection.
 
 This approach has been illustrated using four data examples of random
-forest models with the tree SHAP local explanation. Local explanations
-focus on the model fit, and help to dissect which variables are most
-responsible for the fitted value. They can also form the basis of
-learning how the model has got it wrong, with when the observation is
-misclassified or has a large residual.
+forest models with the tree SHAP LVA. LEs focus on the model fit, and
+help to dissect which variables are most responsible for the fitted
+value. They can also form the basis of learning how the model has got it
+wrong, with when the observation is misclassified or has a large
+residual.
 
@@
 provided, and you can upload your data after model fitting and computing
-the local explanations. The local explanations need to be pre-computed
-and uploaded. Examples show how to do this for tree SHAP values, using
-\textbf{treeshap} (tree based models from \textbf{gbm},
-\textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or
-\textbf{xgboost} \citet{greenwell_gbm_2020}; \citet{shi_lightgbm_2022};
-\citet{liaw_classification_2002}; \citet{wright_ranger_2017};
-\citet{chen_xgboost_2021}, respectively). The SHAP and oscillation
-explanations could be easily added using \texttt{DALEX::explain()}
+the LVAs. The LVAs need to be pre-computed and uploaded. Examples show
+how to do this for tree SHAP values, using \textbf{treeshap} (tree based
+models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest},
+\textbf{ranger}, or \textbf{xgboost} \citet{greenwell_gbm_2020};
+\citet{shi_lightgbm_2022}; \citet{liaw_classification_2002};
+\citet{wright_ranger_2017}; \citet{chen_xgboost_2021}, respectively).
+The SHAP and oscillation explanations could be easily added using
+\texttt{DALEX::explain()}
 \citep{biecek_dalex_2018, biecek_explanatory_2021}.
@@
 \citep{sievert_interactive_2020}. \textbf{DALEX}
-\citep{biecek_dalex_2018} and the free ebook, \emph{Explanatory Model
-Analysis} \citep{biecek_explanatory_2021} were a boon to understanding
-local explanations and how to apply them.
+\citep{biecek_dalex_2018} and \emph{Explanatory Model Analysis}
+\citep{biecek_explanatory_2021} are helpfule for understanding LEs and
+how to apply them.
 
@@
 
-Follow the examples provided with the package to compute the local
-explanation (using \texttt{?cheem\_ls}). The application expects the
-output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file
-with \texttt{saveRDS()} to be uploaded.
+Follow the examples provided with the package to compute the LVAs (using
+\texttt{?cheem\_ls}). The application expects the output returned by
+\texttt{cheem\_ls()}, saved to an \texttt{rds} file with
+\texttt{saveRDS()} to be uploaded.
 

### 2023-05-21 jss/cheem_paper.Rmd
added ExplainPrediction and flashlight references

@@
     affiliation: |
-      | Warsaw University of 
-      | Technology and University
+      | University of Warsaw
+      | Warsaw University of Technology
 title:
@@
 
-There are numerous R packages currently available which provide functions for computing SHAP values, including `fastshap` [@fastshap], `kernelshap` [@kernelshap], `shapr` [@shapr], `shapviz` [@shapviz], `PPtreeregViz` [@PPtreeregViz], and the package `DALEX` has many resources [@biecek_dalex_2018].
-There are many more packages only available through Github, like `treeshap` [kominsarczyk_treeshap_2021] that is used for this work.
+There are numerous R packages currently available which provide functions for computing SHAP values, including `fastshap` [@fastshap], `kernelshap` [@kernelshap], `shapr` [@shapr], `shapviz` [@shapviz], `PPtreeregViz` [@PPtreeregViz], `ExplainPrediction` [@ExplainPrediction], `flashlight` [@flashlight] and the package `DALEX` has many resources [@biecek_dalex_2018].
+There are many more packages only available through Github, like `treeshap` [@kominsarczyk_treeshap_2021] that is used for this work.
 @molnar2022 provides good explanations of the different methods and how to apply them to different models.

### 2023-05-21 jss/cheem_paper.tex
added ExplainPrediction and flashlight references

@@
 University \And Przemyslaw
-Biecek~\orcidlink{0000-0001-8423-1823}\\Warsaw University of\\
-Technology and University
+Biecek~\orcidlink{0000-0001-8423-1823}\\University of Warsaw\\
+Warsaw University of Technology
 }
@@
 \citep{shapr}, \texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
-\citep{PPtreeregViz}, and the package \texttt{DALEX} has many resources
-\citep{biecek_dalex_2018}. There are many more packages only available
-through Github, like \texttt{treeshap}
-{[}kominsarczyk\_treeshap\_2021{]} that is used for this work.
-\citet{molnar2022} provides good explanations of the different methods
-and how to apply them to different models.
+\citep{PPtreeregViz}, \texttt{ExplainPrediction}
+\citep{ExplainPrediction}, \texttt{flashlight} \citep{flashlight} and
+the package \texttt{DALEX} has many resources \citep{biecek_dalex_2018}.
+There are many more packages only available through Github, like
+\texttt{treeshap} \citep{kominsarczyk_treeshap_2021} that is used for
+this work. \citet{molnar2022} provides good explanations of the
+different methods and how to apply them to different models.
 

### 2023-05-22 jss/cheem_paper.Rmd
-fix app link -minor lint&wordsmithing -motiviating letter dates

@@
   \usepackage{amsmath}
+  \usepackage{datetime}
 output: rticles::jss_article
@@
 - A version of the cheem viewer shiny app can be directly accessed at
-https://ebsmonash.shinyapps.io/cheem_initial/.
+https://ebsmonash.shinyapps.io/cheem/.
 - The development version of the package is available at https://github.com/nspyrison/cheem, and

### 2023-05-22 paper.Rmd
-fix app link -minor lint&wordsmithing -motiviating letter dates

@@
 ---
-title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
-type: ARTICLE TEMPLATE
+documentclass: jss
+title: 
+  formatted: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
+  plain: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
+  short: "Exploring Local Explanations"
 author:
   - name: Nicholas Spyrison
-    affiliation: a
-    orcid: 0000-0002-8417-0212
-    email: spyrison@gmail.com
+    affiliation: Monash University
+    address: >
+      Department of Econometrics and Business Statistics
+      Monash University
+    email: \email{spyrison@gmail.com}
   - name: Dianne Cook
-    affiliation: a
-    orcid: 0000-0002-3813-7155
-    email: dicook@monash.edu
+    affiliation: Monash University
   - name: Przemyslaw Biecek
-    affiliation: b
-    orcid: 0000-0001-8423-1823
-    email: przemyslaw.biecek@gmail.com
-affiliation:
-  - num: a
-    address: |
-      Monash University, Melbourne, Australia
-  - num: b
-    address: |
-      Warsaw University of Technology and University, Warsaw, Poland
-abstract: |
+    affiliation: Warsaw University of Technology and University
+abstract: >
   The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
@@
 The methods are implemented in the __R__ package __cheem__, available on CRAN.
-keywords: | 
-  explainable artificial intelligence; nonlinear model interpretability; visual analytics; local explanations; grand tour; radial tour
+keywords:  
+  formatted: [explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour]
+  plain: [explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour]
 bibliography: paper.bib
-header-includes: |
-  \usepackage{hyperref}
-  \usepackage[utf8]{inputenc}
-  \def\tightlist{}
-output: rticles::tf_article
+preamble: >
+  \usepackage{amsmath}
+  \usepackage[english]{babel}
+  \usepackage{float}
+  \usepackage{booktabs}
+  \usepackage{datetime}
+output:
+  bookdown::pdf_book:
+    base_format: rticles::jss_article
+    keep_tex: yes
+    citation_package: 'natbib'
+cls: jss.cls
 ---
@@
 )
+options(tinytex.clean = FALSE)
+options(tinytex.verbose = TRUE)
 ```

### 2023-05-22 paper.tex
-fix app link -minor lint&wordsmithing -motiviating letter dates

--- 2023-05-21 jss/cheem_paper.tex
@@
 \author{
-Nicholas Spyrison~\orcidlink{0000-0002-8417-0212}\\Monash
-University \And Dianne Cook~\orcidlink{0000-0002-3813-7155}\\Monash
-University \And Przemyslaw
-Biecek~\orcidlink{0000-0001-8423-1823}\\Warsaw University of\\
-Technology and University
+Nicholas Spyrison\\Monash University \And Dianne Cook\\Monash University \And Przemyslaw Biecek\\Warsaw University of Technology and University
 }
-\title{Exploring Local Explanations of Nonlinear Models Using Animated
-Linear Projections}
+\title{Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections}
 
 \Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}
-\Plaintitle{Exploring Local Explanations of Nonlinear Models Using
-Animated Linear Projections}
+\Plaintitle{Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections}
 \Shorttitle{Exploring Local Explanations}
@@
 \Abstract{
-The increased predictive power of machine learning models comes at the
-cost of increased complexity and loss of interpretability, particularly
-in comparison to parametric statistical models. This trade-off has led
-to the emergence of eXplainable AI (XAI) which provides methods, such as
-local explanations (LEs) and local variable attributions (LVAs), to shed
-light on how a model use predictors to arrive at a prediction. These
-provide a point estimate of the linear variable importance in the
-vicinity of a single observation. However, LVAs tend not to effectively
-handle association between predictors. To understand how the interaction
-between predictors affects the variable importance estimate, we can
-convert LVAs into linear projections and use the radial tour. This is
-also useful for learning how a model has made a mistake, or the effect
-of outliers, or clustering of observations. The approach is llustrated
-with examples from categorical (penguin species, chocolate types) and
-quantitative (soccer/football salaries, house prices) response models.
-The methods are implemented in the R package cheem, available on CRAN.
+The increased predictive power of nonlinear models comes at the cost of interpretability of its terms.
+This trade-off has led to the emergence of eXplainable AI (XAI).
+XAI attempts to shed light on how models use predictors to arrive at a prediction.
+``Local explanations'', which provide a point estimate of the linear variable importance in the vicinity of one observation, are one XAI method.
+These can be considered to be linear projections and can be further explored to understand the interactions between variables used to make predictions across the predictive model surface.
+Here we describe interactive linear interpolation used for exploration at any observation and illustrate with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
+The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
 }
 
-\Keywords{explainable artificial intelligence, nonlinear model
-interpretability, visual analytics, local explanations, grand
-tour, radial tour}
-\Plainkeywords{explainable artificial intelligence, nonlinear model
-interpretability, visual analytics, local explanations, grand
-tour, radial tour}
+\Keywords{explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour}
+\Plainkeywords{explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour}
 
@@
     Monash University\\
-    Dept of Human Centred Computing\\
-Faculty of Information Technology\\
+    Department of Econometrics and Business Statistics Monash University\\
   E-mail: \email{spyrison@gmail.com}\\
-  URL: \url{https://nspyrison.netlify.app}\\~\\
-      Dianne Cook\\
-    Monash University\\
-    Dept of Econometrics and Business Statistics\\
 
@@
 
-\usepackage{amsmath}
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\usepackage{amsmath} \usepackage[english]{babel} \usepackage{float} \usepackage{booktabs} \usepackage{datetime}
 
@@
 
-There are different reasons and purposes for fitting a model. According
-to the taxonomies of \citet{breiman_statistical_2001} and
-\citet{shmueli_explain_2010}, it can be useful to group models into two
-types: explanatory and predictive. Explanatory modeling is used for
-inferential purposes, while predictive modeling focuses solely on the
-performance of an objective function. The intended use of the model has
-important implications for its selection and development.
-Interpretability is critical in explanatory modeling to draw meaningful
-inferential conclusions, such as which variables most contribute to a
-prediction or whether some observations are less well fit.
+There are different reasons and purposes for fitting a model.
+According to the taxonomies of \citet{breiman_statistical_2001} and \citet{shmueli_explain_2010}, it can be useful to group models into two types: explanatory and predictive.
+Explanatory modeling is used for inferential purposes, while predictive modeling focuses solely on the performance of an objective function.
+The intended use of the model has important implications for its selection and development.
+Interpretability is critical in explanatory modeling to draw meaningful inferential conclusions, such as which variables most contribute to a prediction or whether some observations are less well fit.
 Interpretability becomes more difficult when the model is nonlinear.
-Nonlinear models occur in statistical models with polynomial or
-interaction terms between quantitative predictors, and almost all
-computational models such as random forests, support-vector machines, or
-neural networks
-\citep[e.g.][]{breiman_random_2001, boser_training_1992, anderson_introduction_1995}.
+Nonlinear models occur in statistical models with polynomial or interaction terms between quantitative predictors, and almost all computational models such as random forests, support-vector machines, or neural networks \citep[e.g.][]{breiman_random_2001, boser_training_1992, anderson_introduction_1995}.
 
-In linear models interpretation of the importance of variables is
-relatively straight forward, one adjusts for the covariance of multiple
-variables when examining the relationship with the response. The
-interpretation is valid for the full domain of the predictors. In
-nonlinear models one needs to consider the model in small neighborhoods
-of the domain to make any assessment of variable importance. Even though
-this is difficult, it is especially important to interpret model fits as
-we become more dependent on nonlinear models for routine aspects of life
-to avoid issues described in \citet{stahl-ethics}. Understanding how
-nonlinear models behave when usage extrapolates outside the domain of
-predictors, either in sub-spaces where few samples were provided in the
-training set, or extending outside the domain. It is especially
-important because nonlinear models can vary wildly and predictions can
-be dramatically wrong in these areas.
+In linear models interpretation of the importance of variables is relatively straight forward, one adjusts for the covariance of multiple variables when examining the relationship with the response.
+The interpretation is valid for the full domain of the predictors.
+In nonlinear models one needs to consider the model in small neighborhoods of the domain to make any assessment of variable importance.
+Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues described in \citet{stahl-ethics}.
+Understanding how nonlinear models behave when usage extrapolates outside the domain of predictors, either in sub-spaces where few samples were provided in the training set, or extending outside the domain.
+It is especially important because nonlinear models can vary wildly and predictions can be dramatically wrong in these areas.
 
-Explainable Artificial Intelligence (XAI) is an emerging field of
-research focused on methods for the interpreting of models
-\citep{adadi_peeking_2018, arrieta_explainable_2020}. A class of
-techniques called \emph{local explanations}, provide methods to
-approximate linear variable importance at the location of each
-observation or the predictions at a specific point in the data domain.
-Because these are point-specific, it is challenging to comprehensively
-visualize them to understand a model. There are common approaches for
-visualizing high-dimensional data as a whole, but what is needed are new
-approaches for viewing these individual local explanations, in relation
-to the whole.
+Explainable Artificial Intelligence (XAI) is an emerging field of research focused on methods for the interpreting of models \citep{adadi_peeking_2018, arrieta_explainable_2020}.
+A class of techniques called \emph{local explanations}, provide methods to approximate linear variable importance at the location of each observation or the predictions at a specific point in the data domain.
+Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
+There are common approaches for visualizing high-dimensional data as a whole, but what is needed are new approaches for viewing these individual local explanations, in relation to the whole.
 
-For multivariate data visualization, a \emph{tour}
-\citep{asimov_grand_1985, buja_grand_1986, lee_state_2021} of linear
-data projections onto a lower-dimensional space, could be an element of
-XAI, complementing local explanations. Applying tours to model
-interpretation is recommended in \citet{wickham_visualizing_2015}
-primarily to examine the fitted model in the space of the data.
-\citet{cook_interactive_2007} describe the use of tours for exploring
-classification boundaries and model diagnostics
-\citep{Caragea2008, lee_pptree_2013, da_silva_projection_2021}. There
-are various types of tours. In a \emph{manual} or radial tour
-\citep{cook_manual_1997, spyrison_spinifex_2020}, the path of linear
-projections is defined by changing the contribution of a selected
-variable. We propose to use this to scrutinize a local explanation. This
-approach could be considered to be a counter-factual, what-if analysis,
-such as \emph{ceteris paribus} (``other things held constant'') profiles
-\citep{biecek_ceterisparibus_2020}.
+For multivariate data visualization, a \emph{tour} \citep{asimov_grand_1985, buja_grand_1986, lee_state_2021} of linear data projections onto a lower-dimensional space, could be an element of XAI, complementing local explanations.
+Applying tours to model interpretation is recommended in \citet{wickham_visualizing_2015} primarily to examine the fitted model in the space of the data.
+\citet{cook_interactive_2007} describe the use of tours for exploring classification boundaries and model diagnostics \citep{Caragea2008, lee_pptree_2013, da_silva_projection_2021}.
+There are various types of tours.
+In a \emph{manual} or radial tour \citep{cook_manual_1997, spyrison_spinifex_2020}, the path of linear projections is defined by changing the contribution of a selected variable.
+We propose to use this to scrutinize a local explanation.
+This approach could be considered to be a counter-factual, what-if analysis, such as \emph{ceteris paribus} (``other things held constant'') profiles \citep{biecek_ceterisparibus_2020}.
 
-The remainder of this paper is organized as follows. Section
-\ref{sec:explanations} covers the background of the local explanation
-and the traditional visuals produced. Section \ref{sec:tour} explains
-the tours and particularly the radial manual tour. Section
-\ref{sec:cheemviewer} discusses the visual layout in the graphical user
-interface and how it facilitates analysis, data pre-processing, and
-package infrastructure. Illustrations are provided in Section
-\ref{sec:casestudies} for a range of supervised learning tasks with
-categorical and quantitative response variables. These show how the
-local explanations can be used to get an overview of the model's use of
-predictors, and to investigate errors in the model predictions. Section
-\ref{sec:cheemdiscussion} concludes with a summary of the insights
-gained. The methods are implemented in the \textbf{R} package
-\textbf{cheem}.
+The remainder of this paper is organized as follows.
+Section \ref{sec:explanations} covers the background of the local explanation and the traditional visuals produced.
+Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data pre-processing, and package infrastructure.
+Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative response variables.
+These show how the local explanations can be used to get an overview of the model's use of predictors, and to investigate errors in the model predictions.
+Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
+The methods are implemented in the \textbf{R} package \textbf{cheem}.
 
@@
 
-Local explanations shed light on nonlinear model fits by estimating
-linear variable importance in the vicinity of a single observation.
-There are many approaches for calculating local explanations. A
-comprehensive summary of the taxonomy of currently available methods is
-provided in Figure 6 by \citet{arrieta_explainable_2020}. It includes a
-large number of model-specific explanations such as deepLIFT
-\citep{shrikumar_not_2016, shrikumar_learning_2017}, a popular recursive
-method for estimating importance in neural networks. There are fewer
-model-agnostic methods, of which LIME, \citep{ribeiro_why_2016} SHaply
-Additive exPlanations (SHAP), \citep{lundberg_unified_2017}, are
-popular.
+Local explanations shed light on nonlinear model fits by estimating linear variable importance in the vicinity of a single observation. There are many approaches for calculating local explanations.
+A comprehensive summary of the taxonomy of currently available methods is provided in Figure 6 by \citet{arrieta_explainable_2020}.
+It includes a large number of model-specific explanations such as deepLIFT \citep{shrikumar_not_2016, shrikumar_learning_2017}, a popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic methods, of which LIME, \citep{ribeiro_why_2016} SHaply Additive exPlanations (SHAP), \citep{lundberg_unified_2017}, are popular.
 
-These observation-level explanations are used in various ways depending
-on the data. In image classification, where pixels correspond to
-predictors, saliency maps overlay or offset a heatmap to indicate
-important pixels \citep{simonyan_deep_2014}. For example, pixels
-corresponding to snow may be highlighted as important contributors when
-distinguishing if a picture contains a coyote or husky. In text
-analysis, word-level contextual sentiment analysis highlights the
-sentiment and magnitude of influential words \citep{vanni_textual_2018}.
-In the case of numeric regression, they are used to explain additive
-contributions of variables from the model intercept to the observation's
-prediction \citep{ribeiro_why_2016}.
+These observation-level explanations are used in various ways depending on the data.
+In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap to indicate important pixels \citep{simonyan_deep_2014}.
+For example, pixels corresponding to snow may be highlighted as important contributors when distinguishing if a picture contains a coyote or husky.
+In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words \citep{vanni_textual_2018}.
+In the case of numeric regression, they are used to explain additive contributions of variables from the model intercept to the observation's prediction \citep{ribeiro_why_2016}.
 
-We will be focusing on SHAP values in this this paper, but the approach
-is applicable for any method used to calculate the local explanations.
-SHAP calculates the variable contributions of one observation by
-examining the effect of other variables on the predictions. The term
-``SHAP'' refers to \citet{shapley_value_1953}'s method to evaluate an
-individual's contribution in cooperative games by assessing this
-player's performance in the presence or absence of other players.
-\citet{strumbelj_efficient_2010} introduced SHAP for local explanations
-in machine learning models. Variable importance can depend on the
-sequence in which variables are entered into the model fitting process,
-thus for any sequence we get a set of variable contribution values for a
-single observation. These values will add up to the difference between
-the fitted value for the observation, and the average fitted value for
-all observations. Using all possible sequence, or permutation, gives
-multiple values for each variable, which are averaged to get the SHAP
-value for an observation. It can be helpful to standardize variables
-prior to computing SHAP values, if they have been measured on different
-scales.
+We will be focusing on SHAP values in this this paper, but the approach is applicable for any method used to calculate the local explanations.
+SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
+The term ``SHAP'' refers to \citet{shapley_value_1953}'s method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
+\citet{strumbelj_efficient_2010} introduced SHAP for local explanations in machine learning models.
+Variable importance can depend on the sequence in which variables are entered into the model fitting process, thus for any sequence we get a set of variable contribution values for a single observation.
+These values will add up to the difference between the fitted value for the observation, and the average fitted value for all observations.
+Using all possible sequence, or permutation, gives multiple values for each variable, which are averaged to get the SHAP value for an observation.
+It can be helpful to standardize variables prior to computing SHAP values, if they have been measured on different scales.
 
-The approach is related to partial dependence plots (see for example
-\citet{molnar_interpretable_2020}), used to explain the effect of a
-variable by predicting the response for a range of values on this
-variable after fixing the value of all other variables to their mean.
-Though partial dependence plots are a global approximation of the
-variable importance, while SHAP is specific to one observation.
+The approach is related to partial dependence plots (see for example \citet{molnar_interpretable_2020}), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
+Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
 
@@
 Reaction skills are important for both players.
-Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
+Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}(\#fig:shapdistrbd)
 \end{figure}
@@
 
-We use 2020 season FIFA data \citep{leone_fifa_2020} to illustrate SHAP
-following the procedures described in \citet{biecek_explanatory_2021}.
-There are 5000 observations of nine predictor variables measuring
-players' skills and one response variable, wages (in euros). A random
-forest model is fit regressing players' wages on the skill variables. In
-this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are
-compared for a star offensive player (L. Messi) and a prominent
-defensive player (V. van Dijk). We are interested in knowing how the
-skill variables locally contribute to the wage prediction of each
-player. A difference in the attribution of the variable importance
-across the two positions of the players can be expected. This would be
-interpreted as how a player's salary depends on which combination of
-skills. Panel (a) is a version of a breakdown plot
-\citep{gosiewska_ibreakdown_2019} where just three sequences of
-variables are shown, for two observations. A breakdown plot shows the
-absolute values of the variable attribution for an observation, usually
-sorted from highest value to the lowest. There is no scale on the
-horizontal axis here because values are considered relative to each
-other. Here we can see how the variable contribution can change
-depending on sequence, relative to both players. (Note that the order of
-the variables is different in each plot, because they have been sorted
-by biggest average contribution across both players.) For all sequences,
-and for both players \texttt{reaction} has the strongest contribution,
-with perhaps more importance for the defensive player. Then it differs
-by player: for Messi \texttt{offense} and \texttt{movement} have the
-strongest contributions, and for van Dijk it is \texttt{defense} and
-\texttt{power}, regardless of the variable sequence.
+We use 2020 season FIFA data \citep{leone_fifa_2020} to illustrate SHAP following the procedures described in \citet{biecek_explanatory_2021}.
+There are 5000 observations of nine predictor variables measuring players' skills and one response variable, wages (in euros).
+A random forest model is fit regressing players' wages on the skill variables.
+In this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are compared for a star offensive player (L.
+Messi) and a prominent defensive player (V.
+van Dijk).
+We are interested in knowing how the skill variables locally contribute to the wage prediction of each player.
+A difference in the attribution of the variable importance across the two positions of the players can be expected.
+This would be interpreted as how a player's salary depends on which combination of skills.
+Panel (a) is a version of a breakdown plot \citep{gosiewska_ibreakdown_2019} where just three sequences of variables are shown, for two observations.
+A breakdown plot shows the absolute values of the variable attribution for an observation, usually sorted from highest value to the lowest.
+There is no scale on the horizontal axis here because values are considered relative to each other.
+Here we can see how the variable contribution can change depending on sequence, relative to both players.
+(Note that the order of the variables is different in each plot, because they have been sorted by biggest average contribution across both players.) For all sequences, and for both players \texttt{reaction} has the strongest contribution, with perhaps more importance for the defensive player.
+Then it differs by player: for Messi \texttt{offense} and \texttt{movement} have the strongest contributions, and for van Dijk it is \texttt{defense} and \texttt{power}, regardless of the variable sequence.
 
-Panel (b) shows the differences in the player's median values (large
-dots) for 25 such sequences (tick marks). We can see that the wage
-predictions for the two players come from different combinations of
-skills sets, as might be expected for players who's value on the team
-depends on their offensive or defensive prowess. It is also interesting
-to see from the distribution of values across the different sequence of
-variables, that there is some multimodality. For example, look at the
-SHAP values for \texttt{reaction} for Messi, and in some sequences
-reaction has a much lower contribution than others. This suggests that
-other variables (\texttt{offense}, \texttt{movement} probably) can
-substitute for \texttt{reaction} in the wage prediction.
+Panel (b) shows the differences in the player's median values (large dots) for 25 such sequences (tick marks).
+We can see that the wage predictions for the two players come from different combinations of skills sets, as might be expected for players who's value on the team depends on their offensive or defensive prowess.
+It is also interesting to see from the distribution of values across the different sequence of variables, that there is some multimodality.
+For example, look at the SHAP values for \texttt{reaction} for Messi, and in some sequences reaction has a much lower contribution than others.
+This suggests that other variables (\texttt{offense}, \texttt{movement} probably) can substitute for \texttt{reaction} in the wage prediction.
 
-This can also be considered similar to examining the coefficients from
-all subsets regression, as described in
-\citet{wickham_visualizing_2015}. Various models that are similarly good
-might use different combinations of the variables. Examining the
-coefficients from multiple models helps to understand the relative
-importance of each variable in the context of all other variables. This
-is similar to the approach here with SHAP values, that by examining the
-variation in values across different permutations of variables, we can
-gain more understanding of the relationship between the response and
-predictors.
+This can also be considered similar to examining the coefficients from all subsets regression, as described in \citet{wickham_visualizing_2015}.
+Various models that are similarly good might use different combinations of the variables.
+Examining the coefficients from multiple models helps to understand the relative importance of each variable in the context of all other variables.
+This is similar to the approach here with SHAP values, that by examining the variation in values across different permutations of variables, we can gain more understanding of the relationship between the response and predictors.
 
-For the application, we use \emph{tree SHAP}, a variant of SHAP that
-enjoys a lower computational complexity
-\citep{lundberg_consistent_2018}. Instead of aggregating over sequences
-of the variables, tree SHAP calculates observation-level variable
-importance by exploring the structure of the decision trees. Tree SHAP
-is only compatible with tree-based models. so random forests are used
-for illustration.
+For the application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity \citep{lundberg_consistent_2018}.
+Instead of aggregating over sequences of the variables, tree SHAP calculates observation-level variable importance by exploring the structure of the decision trees.
+Tree SHAP is only compatible with tree-based models.
+so random forests are used for illustration.
 
@@
 
-A \emph{tour} enables the viewing of high-dimensional data by animating
-many linear projections with small incremental changes. It is achieved
-by following a path of linear projections (bases) of high-dimensional
-space. One key variable of the tour is the object permanence of the data
-points; one can track the relative change of observations in time and
-gain information about the relationships between points across multiple
-variables. There are various types of tours that are distinguished by
-how the paths are generated \citep{lee_state_2021, cook_grand_2008}.
+A \emph{tour} enables the viewing of high-dimensional data by animating many linear projections with small incremental changes.
+It is achieved by following a path of linear projections (bases) of high-dimensional space.
+One key variable of the tour is the object permanence of the data points; one can track the relative change of observations in time and gain information about the relationships between points across multiple variables.
+There are various types of tours that are distinguished by how the paths are generated \citep{lee_state_2021, cook_grand_2008}.
 
-The manual tour \citep{cook_manual_1997} defines its path by changing a
-selected variable's contribution to a basis to allow the variable to
-contribute more or less to the projection. The requirement constrains
-the contribution of all other variables that a basis needs to be
-orthonormal (column correspond to vectors, with unit length, and
-orthogonal to each other). The manual tour is primarily used to assess
-the importance of a variable to structure visible in a projection. It
-also lends itself to pre-computation queued in advance or computed
-on-the-fly for human-in-the-loop analysis
-\citep{karwowski_international_2006}.
+The manual tour \citep{cook_manual_1997} defines its path by changing a selected variable's contribution to a basis to allow the variable to contribute more or less to the projection.
+The requirement constrains the contribution of all other variables that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
+It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis \citep{karwowski_international_2006}.
 
-A version of the manual tour called a \emph{radial tour} is implemented
-in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
-In a radial tour, the selected variable can change its magnitude of
-contribution but not its angle; it must move along the direction of its
-original contribution. The implementation allows for pre-computation and
-interactive re-calculation to focus on a different variable.
+A version of the manual tour called a \emph{radial tour} is implemented in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
+In a radial tour, the selected variable can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and interactive re-calculation to focus on a different variable.
 
@@
 As this variable contribution is reduced in the plot (middle, right) we can see that the bimodality decreases.
-Thus \textsf{bd} is an important variable contributing to the bimodal structure.}\label{fig:radialtour}
+Thus \textsf{bd} is an important variable contributing to the bimodal structure.}(\#fig:radialtour)
 \end{figure}
@@
 
-To explore the local explanations, coordinated views
-\citep{roberts_state_2007} \citep[also known as ensemble
-graphics,][]{unwin_ensemble_2018} are provided in the \emph{cheem
-viewer} application. There are two primary plots: the \textbf{global
-view} to give the context of all of the SHAP values and the
-\textbf{radial tour view} to explore the local explanations with
-user-controlled rotation. There are numerous user inputs, including
-variable selection for the radial tour and observation selection for
-making comparisons. There are different plots used for the categorical
-and quantitative responses. Figures \ref{fig:classificationcase} and
-\ref{fig:regressioncase} are screenshots showing the cheem viewer for
-the two primary tasks: classification (categorical response) and
-regression (quantitative response).
+To explore the local explanations, coordinated views \citep{roberts_state_2007} \citep[also known as ensemble graphics,][]{unwin_ensemble_2018} are provided in the \emph{cheem viewer} application.
+There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values and the \textbf{radial tour view} to explore the local explanations with user-controlled rotation.
+There are numerous user inputs, including variable selection for the radial tour and observation selection for making comparisons.
+There are different plots used for the categorical and quantitative responses.
+Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 
-The global view provides context for all observations and facilitates
-the exploration of the separability of the data and attribution spaces.
+The global view provides context for all observations and facilitates the exploration of the separability of the data and attribution spaces.
 The attribution space refers to the SHAP values for each observation.
-These spaces both have dimensionality \(n \times p\), where \(n\) is the
-number of observations and \(p\) is the number of variables.
+These spaces both have dimensionality \(n \times p\), where \(n\) is the number of observations and \(p\) is the number of variables.
 
-The visualization is composed of the first two principal components of
-the data (left) and the attribution (middle) spaces. These single 2D
-projections will not reveal all of the structure of higher-dimensional
-space, but they are helpful visual summaries. In addition, a plot of the
-observed against predicted response values is also provided (Figures
-\ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help
-identify observations poorly predicted by the model. For classification
-tasks, color indicates the predicted class, and misclassified
-observations are circled in red. Linked brushing between the plots is
-provided, and a tabular display of selected points helps to facilitate
-exploration of the spaces and the model (shown in Figures
-\ref{fig:regressioncase}d).
+The visualization is composed of the first two principal components of the data (left) and the attribution (middle) spaces.
+These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
+In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify observations poorly predicted by the model.
+For classification tasks, color indicates the predicted class, and misclassified observations are circled in red.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
 
-While the comparison of these spaces is interesting, the primary purpose
-of the global view is to enable the selection of particular observations
-to explore in detail. We have designed it to enable a comparison between
-an observation that is interesting is some way, perhaps misclassified,
-or poorly predicted, relative to an observation with similar predictor
-values but more expected prediction. For brevity, we call the
-interesting observation the primary investigation (PI) and the other is
-the comparison investigation (CI). These observations are highlighted as
-asterisk and \(\times\), respectively.
+While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular observations to explore in detail.
+We have designed it to enable a comparison between an observation that is interesting is some way, perhaps misclassified, or poorly predicted, relative to an observation with similar predictor values but more expected prediction.
+For brevity, we call the interesting observation the primary investigation (PI) and the other is the comparison investigation (CI).
+These observations are highlighted as asterisk and \(\times\), respectively.
 
@@
 
-There are two plots in this part of the interface. The first (Figures
-\ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a
-display of the SHAP values for all observations. This will generally
-give the global view of variables important for the fit as a whole, but
-it will also highlight observations which have different patterns. The
-second plot is the radial tour, which for classification is a density
-plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for
-regression are scatterplots of the observed response values, and
-residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
+There are two plots in this part of the interface.
+The first (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a display of the SHAP values for all observations.
+This will generally give the global view of variables important for the fit as a whole, but it will also highlight observations which have different patterns.
+The second plot is the radial tour, which for classification is a density plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for regression are scatterplots of the observed response values, and residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
 
-The local explanations for all observations are normalized (sum of
-squares equals 1), and thus, the relative importance of variables can be
-compared across all observations. These are depicted as a vertical
-parallel coordinate plot \citep{ocagne_coordonnees_1885}. (The SHAP
-values of the PI and CI are shown as dashed and dotted lines,
-respectively.) One should obtain a sense of the overall importance of
-variables from this plot. The more important variables will have larger
-values, and in the case of classification tasks variables which have
-different magnitude for different classes are more globally important.
-For example, Figure \ref{fig:classificationcase}e suggests that
-\texttt{bl} is important for distinguishing the green class from the
-other two. For regression, one might generally observe which variables
-have low values for all observations (not important). For example,
-\texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have
-a range of high and low values (e.g., \texttt{off}, \texttt{def}),
-suggesting they are important for some observations and not important
-for others.
+The local explanations for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
+These are depicted as a vertical parallel coordinate plot \citep{ocagne_coordonnees_1885}.
+(The SHAP values of the PI and CI are shown as dashed and dotted lines, respectively.) One should obtain a sense of the overall importance of variables from this plot.
+The more important variables will have larger values, and in the case of classification tasks variables which have different magnitude for different classes are more globally important.
+For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
+For regression, one might generally observe which variables have low values for all observations (not important).
+For example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some observations and not important for others.
 
-A bar chart is overlaid to represent the projection shown in the radial
-tour at right. It starts from the SHAP values of the PI, but if the user
-changes the projection the length of these bars will reflect this
-change. (The PI is interactively selected by clicking on a point in the
-global view). By scaling the SHAP value it becomes an (attribution)
-projection.
+A bar chart is overlaid to represent the projection shown in the radial tour at right.
+It starts from the SHAP values of the PI, but if the user changes the projection the length of these bars will reflect this change.
+(The PI is interactively selected by clicking on a point in the global view).
+By scaling the SHAP value it becomes an (attribution) projection.
 
-The attribution projection of the PI is the initial 1D basis in a radial
-tour, displayed as a density plot for a categorical response (Figure
-\ref{fig:classificationcase}f) and as scatterplots for a quantitative
-response (Figure \ref{fig:regressioncase}f). The PI and CI are indicated
-by vertical dashed and dotted lines, respectively. The radial tour
-varies the contribution of the selected variable. This is viewed as an
-animation of the projections from many intermediate bases. Doing so
-tests the sensitivity of structure (class separation or strength of
-relationship) to the variable's contribution. For classification, if the
-separation between classes diminishes when the variable contribution is
-reduced, this suggests that the variable is important for class
-separation. For regression, if the relationship scatterplot weakens when
-the variable contribution is reduced, indicating that the variable is
-important for accurately predicting the response.
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
+The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The radial tour varies the contribution of the selected variable.
+This is viewed as an animation of the projections from many intermediate bases.
+Doing so tests the sensitivity of structure (class separation or strength of relationship) to the variable's contribution.
+For classification, if the separation between classes diminishes when the variable contribution is reduced, this suggests that the variable is important for class separation.
+For regression, if the relationship scatterplot weakens when the variable contribution is reduced, indicating that the variable is important for accurately predicting the response.
 
@@
 
-Selecting a misclassified observation as PI and a correctly classified
-point nearby in data space as CI makes it easier to examine the
-variables most responsible for the error. The global view (Figure
-\ref{fig:classificationcase}c) displays the model confusion matrix. The
-radial tour is 1D and displays as density where color indicates class.
-An animation slider enables users to vary the contribution of variables
-to explore the sensitivity of the separation to that variable.
+Selecting a misclassified observation as PI and a correctly classified point nearby in data space as CI makes it easier to examine the variables most responsible for the error.
+The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
+The radial tour is 1D and displays as density where color indicates class.
+An animation slider enables users to vary the contribution of variables to explore the sensitivity of the separation to that variable.
 
@@
 The black bar is the variable being changed in the radial tour.
-Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
+Panel (f) is the resulting data projection indicated as density in the classification case.}(\#fig:classificationcase)
 \end{figure}
@@
 
-Selecting an inaccurately predicted observation as PI and an accurately
-predicted observation with similar variable values as CI is a helpful
-way to understand how the model is failing or not. The global view
-(Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed
-vs predicted values, which should exhibit a strong relationship if the
-model is a good fit. The points can be colored by a statistic, residual,
-a measure of outlyingness (log Mahalanobis distance), or correlation to
-aid in understanding the structure identified in these spaces.
+Selecting an inaccurately predicted observation as PI and an accurately predicted observation with similar variable values as CI is a helpful way to understand how the model is failing or not.
+The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation to aid in understanding the structure identified in these spaces.
 
-In the radial tour view, the observed response and the residuals
-(vertical) are plotted against the attribution projection of the PI
-(horizontal). The attribution projection can be interpreted similarly to
-the predicted value from the global view plot. It represents a linear
-combination of the variables, and a good fit would be indicated when
-there is a strong relationship with the observed values. This can be
-viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a variable is varied, if the value of the PI does
-not change much, it would indicate that the prediction for this
-observation is NOT sensitive to that variable. Conversely, if the
-predicted value varies substantially, the prediction is very sensitive
-to that variable, suggesting that the variable is very important for the
-PI's prediction.
+In the radial tour view, the observed response and the residuals (vertical) are plotted against the attribution projection of the PI (horizontal).
+The attribution projection can be interpreted similarly to the predicted value from the global view plot.
+It represents a linear combination of the variables, and a good fit would be indicated when there is a strong relationship with the observed values.
+This can be viewed as a local linear approximation if the fitted model is nonlinear.
+As the contribution of a variable is varied, if the value of the PI does not change much, it would indicate that the prediction for this observation is NOT sensitive to that variable.
+Conversely, if the predicted value varies substantially, the prediction is very sensitive to that variable, suggesting that the variable is very important for the PI's prediction.
 
@@
 Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the variable attribution from all observations, and horizontal bars show the contribution to the current basis.
-Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}\label{fig:regressioncase}
+Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}(\#fig:regressioncase)
 \end{figure}
@@
 
-The application has several reactive inputs that affect the data used,
-aesthetic display, and tour manipulation. These reactive inputs make the
-software flexible and extensible (Figure \ref{fig:classificationcase}a
-\& d). The application also has more exploratory interactions to help
-link points across displays, reveal structures found in different
-spaces, and access the original data.
+The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
+These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a \& d).
+The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
 
-A tooltip displays observation number/name and classification
-information while the cursor hovers over a point. Linked brushing allows
-the selection of points (left click and drag) where those points will be
-highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
-The information corresponding to the selected points is populated on a
-dynamic table (Figure \ref{fig:classificationcase}d). These interactions
-aid exploration of the spaces and, finally, the identification of
-primary and comparison observations.
+A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
+The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
+These interactions aid exploration of the spaces and, finally, the identification of primary and comparison observations.
 
@@
 
-It is vital to mitigate the render time of visuals, especially when
-users may want to iterate many explorations. All computational
-operations should be prepared before run time. The work remaining when
-an application is run solely reacts to inputs and rendering visuals and
-tables. Below discusses the steps and details of the reprocessing.
+It is vital to mitigate the render time of visuals, especially when users may want to iterate many explorations.
+All computational operations should be prepared before run time.
+The work remaining when an application is run solely reacts to inputs and rendering visuals and tables.
+Below discusses the steps and details of the reprocessing.
 
@@
 \item
-  \textbf{Data:} predictors and response are unscaled complete numerical
-  matrix. Most models and local explanations are scale-invariant. Keep
-  normality assumptions of the model in mind.
+  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep normality assumptions of the model in mind.
 \item
-  \textbf{Model:} any model and compatible explanation could be explored
-  with this method. Currently, random forest models are applied via the
-  package \textbf{randomForest} \citep{liaw_classification_2002},
-  compatibility tree SHAP. Modest hyperparameters are used, namely: 125
-  trees, the number of variables at each split, mtry = \(\sqrt{p}\) or
-  \(p/3\) for classification and regression, and minimum size of
-  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
-  classification and regression.
+  \textbf{Model:} any model and compatible explanation could be explored with this method.
+Currently, random forest models are applied via the package \textbf{randomForest} \citep{liaw_classification_2002}, compatibility tree SHAP.
+Modest hyperparameters are used, namely: 125 trees, the number of variables at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
 \item
-  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
-  observation using the package \textbf{treeshap}
-  \citep{kominsarczyk_treeshap_2021}. We opt to find the attribution of
-  each observation in the training data and not fit to fit variable
-  interactions.
+  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} \citep{kominsarczyk_treeshap_2021}.
+We opt to find the attribution of each observation in the training data and not fit to fit variable interactions.
 \item
-  \textbf{Cheem viewer:} after the model and full explanation space are
-  calculated, each variable is scaled by standard deviations away from
-  the mean to achieve common support for visuals. Statistics for mapping
-  to color are computed on the scaled spaces.
+  \textbf{Cheem viewer:} after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
+Statistics for mapping to color are computed on the scaled spaces.
 \end{itemize}
 
-The time to preprocess the data will vary significantly with the
-complexity of the model and local explanation. For reference, the FIFA
-data contained 5000 observations of nine explanatory variables took 2.5
-seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 270 seconds
-total. PCA and statistics of the variables and attributions took 2.8
-seconds. These run times were from a non-parallelized session on a
-modern laptop, but suffice to say that most of the time will be spent on
-the local attribution. An increase in model complexity or data
-dimensionality will quickly become an obstacle. Its reduced
-computational complexity makes tree SHAP an excellent candidate to
-start. Alternatively, the package \textbf{fastshap} claims extremely low
-run times, attributed to fewer calls to the prediction function, partial
-implementation in C++, and efficient use of logical subsetting,
-\citet{greenwell_fastshap_2020}.
+The time to preprocess the data will vary significantly with the complexity of the model and local explanation.
+For reference, the FIFA data contained 5000 observations of nine explanatory variables took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 270 seconds total.
+PCA and statistics of the variables and attributions took 2.8 seconds.
+These run times were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the local attribution.
+An increase in model complexity or data dimensionality will quickly become an obstacle.
+Its reduced computational complexity makes tree SHAP an excellent candidate to start.
+Alternatively, the package \textbf{fastshap} claims extremely low run times, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, \citet{greenwell_fastshap_2020}.
 
@@
 
-To illustrate the cheem method it is applied to modern data sets, two
-classification examples and then two of regression.
+To illustrate the cheem method it is applied to modern data sets, two classification examples and then two of regression.
 
 \hypertarget{palmer-penguin-species-classification}{%
-\subsection{Palmer Penguin, Species
-Classification}\label{palmer-penguin-species-classification}}
+\subsection{Palmer Penguin, Species Classification}\label{palmer-penguin-species-classification}}
 
-The Palmer penguins data
-\citep{gorman_ecological_2014, horst_palmerpenguins_2020} was collected
-on three species of penguins foraging near Palmer Station, Antarctica.
-The data is publicly available to substitute for the overly-used iris
-data and is quite similar in form. After removing incomplete
-observations, there are 333 observations of four physical measurements,
-bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length
-(\texttt{fl}), and body mass (\texttt{bm}) for this illustration. A
-random forest model was fit with species as the response variable.
+The Palmer penguins data \citep{gorman_ecological_2014, horst_palmerpenguins_2020} was collected on three species of penguins foraging near Palmer Station, Antarctica.
+The data is publicly available to substitute for the overly-used iris data and is quite similar in form.
+After removing incomplete observations, there are 333 observations of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), and body mass (\texttt{bm}) for this illustration.
+A random forest model was fit with species as the response variable.
+
+(ref:casepenguins-cap) Examining the SHAP values for a random forest model classifying Palmer penguin species.
+The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
+The radial view shows varying the contribution of \texttt{fl} from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
+(The animation of the radial tour is at \url{https://vimeo.com/666431172}.)
 
@@
 
-\caption[Examining the SHAP values for a random forest model classifying Palmer penguin species]{Examining the SHAP values for a random forest model classifying Palmer penguin species.
-The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
-The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right).
-(The animation of the radial tour is at https://vimeo.com/666431172.)}\label{fig:casepenguins}
+\caption[(ref:casepenguins-cap)]{(ref:casepenguins-cap)}(\#fig:casepenguins)
 \end{figure}
@@
 
-Figure \ref{fig:casepenguins} shows plots from the cheem viewer for
-exploring the random forest model on the penguins data. Panel (a) shows
-the global view, and panel (b) shows several 1D projections generated
-with the radial tour. Penguin 243, a Gentoo (purple), is the PI because
-it has been misclassified as a Chinstrap (orange).
+Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
+Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
+Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
+
+(ref:casepenguinsblfl-cap) Checking what is learned from the cheem viewer.
+This is a plot of flipper length (\texttt{fl}) and bill length (\texttt{bl}), where an asterisk highlights the PI.
+A Gentoo (purple) misclassified as a Chinstrap (orange).
+The PI has an unusually small \texttt{fl} length which is why it is confused with a Chinstrap.
 
@@
 
-\caption[Checking what is learned from the cheem viewer]{Checking what is learned from the cheem viewer.
-This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI.
-A Gentoo (purple) misclassified as a Chinstrap (orange).
-The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
+\caption[(ref:casepenguinsblfl-cap)]{(ref:casepenguinsblfl-cap)}(\#fig:casepenguinsblfl)
 \end{figure}
@@
 
-There is more separation visible in the attribution space than in the
-data space, as would be expected. The predicted vs observed plot reveals
-a handful of misclassified observations. A Gentoo that has been wrongly
-labeled as a Chinstrap is selected for illustration. The PI is a
-misclassified point (represented by the asterisk in the global view and
-a dashed vertical line in the tour view). The CI is a correctly
-classified point (represented by an \(\times\) and a vertical dotted
-line).
+There is more separation visible in the attribution space than in the data space, as would be expected.
+The predicted vs observed plot reveals a handful of misclassified observations.
+A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
+The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
+The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
 
-The radial tour starts from the attribution projection of the
-misclassified observation (b, left). The important variables identified
-by SHAP in the (wrong) prediction for this observation are mostly
-\texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and
-\texttt{bm}. This projection is a view where the Gentoo (purple) looks
-much more likely for this observation than Chinstrap. That is, this
-combination of variables is not particularly useful because the PI looks
-very much like other Gentoo penguins. The radial tour is used to vary
-the contribution of flipper length (\texttt{fl}) to explore this. (In
-our exploration, this was the third variable explored. It is typically
-helpful to explore the variables with more significant contributions,
-here \texttt{bl} and \texttt{bd}. Still, when doing this, nothing was
-revealed about how the PI differed from other Gentoos). On varying
-\texttt{fl} as it contributes increasingly to the projection (b, right),
-more and more, this penguin looks like a Chinstrap. This suggests that
-\texttt{fl} should be considered an important variable for explaining
-the (wrong) prediction.
+The radial tour starts from the attribution projection of the misclassified observation (b, left).
+The important variables identified by SHAP in the (wrong) prediction for this observation are mostly \texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
+This projection is a view where the Gentoo (purple) looks much more likely for this observation than Chinstrap.
+That is, this combination of variables is not particularly useful because the PI looks very much like other Gentoo penguins.
+The radial tour is used to vary the contribution of flipper length (\texttt{fl}) to explore this.
+(In our exploration, this was the third variable explored.
+It is typically helpful to explore the variables with more significant contributions, here \texttt{bl} and \texttt{bd}.
+Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
+On varying \texttt{fl} as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
+This suggests that \texttt{fl} should be considered an important variable for explaining the (wrong) prediction.
 
-Figure \ref{fig:casepenguinsblfl} confirms that flipper length
-(\texttt{fl}) is vital for the confusion of the PI as a Chinstrap. Here,
-flipper length and body length are plotted, and the PI can be seen to be
-closer to the Chinstrap group in these two variables, mainly because it
-has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard observation to account
-for, as decision trees can only partition only vertical and horizontal
-lines.
+Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{fl}) is vital for the confusion of the PI as a Chinstrap.
+Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two variables, mainly because it has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard observation to account for, as decision trees can only partition only vertical and horizontal lines.
 
 \hypertarget{chocolates-milkdark-classification}{%
-\subsection{Chocolates, Milk/Dark
-Classification}\label{chocolates-milkdark-classification}}
+\subsection{Chocolates, Milk/Dark Classification}\label{chocolates-milkdark-classification}}
 
-The chocolates data set consists of 88 observations of ten nutritional
-measurements determined from their labels and labeled as either milk or
-dark. Dark chocolate is considered healthier than milk. Students
-collected the data during the Iowa State University class STAT503 from
-nutritional information on the manufacturer's websites and were
-normalized to 100g equivalents. The data is available in the
-\textbf{cheem} package. A random forest model is used for the
-classification of chocolate types.
+The chocolates data set consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
+Dark chocolate is considered healthier than milk.
+Students collected the data during the Iowa State University class STAT503 from nutritional information on the manufacturer's websites and were normalized to 100g equivalents.
+The data is available in the \textbf{cheem} package.
+A random forest model is used for the classification of chocolate types.
 
-It could be interesting to examine the nutritional properties of any
-dark chocolates that have been misclassified as milk. A reason to do
-this is that a dark chocolate, nutritionally more like milk should not
-be considered a healthy alternative. It is interesting to explore which
-nutritional variables contribute most to misclassification.
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
+A reason to do this is that a dark chocolate, nutritionally more like milk should not be considered a healthy alternative.
+It is interesting to explore which nutritional variables contribute most to misclassification.
+
+(ref:casechocolates-cap) Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
+So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
+When Sugar is zero, Fiber contributes strongly towards the left.
+In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
+This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
+(A video of the tour animation can be found at \url{https://vimeo.com/666431143}.)
 
@@
 
-\caption[Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the local explanation for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
-From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the local explanation does not help understand the prediction for this observation.
-So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
-When Sugar is zero, Fiber contributes strongly towards the left.
-In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
-This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate.
-(A video of the tour animation can be found at https://vimeo.com/666431143.)}\label{fig:casechocolates}
+\caption[(ref:casechocolates-cap)]{(ref:casechocolates-cap)}(\#fig:casechocolates)
 \end{figure}
@@
 
-This type of exploration is shown in Figure \ref{fig:casechocolates},
-where a chocolate labeled dark but predicted to be milk is chosen as the
-PI (observation 22). It is compared with a CI that is a correctly
-classified dark chocolate (observation 7). The PCA plot and the tree
-SHAP PCA plots (a) show a big difference between the two chocolate types
-but with confusion for a handful of observations. The misclassifications
-are more apparent in the observed vs predicted plot and can be seen to
-be mistaken in both ways: milk to dark and dark to milk.
+This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+It is compared with a CI that is a correctly classified dark chocolate (observation 7).
+The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
 
-The attribution projection for chocolate 22 suggests that Fiber, Sugars,
-and Calories are most responsible for its incorrect prediction. The way
-to read this plot is to see that Fiber has a large negative value while
-Sugars and Calories have reasonably large positive values. In the
-density plot, observations on the very left of the display would have
-high values of Fiber (matching the negative projection coefficient) and
-low values of Sugars and Calories. The opposite would be interpreting a
-point with high values in this plot. The dark chocolates (orange) are
-primarily on the left, and this is a reason why they are considered to
-be healthier: high fiber and low sugar. The density of milk chocolates
-is further to the right, indicating that they generally have low fiber
-and high sugar.
+The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+The opposite would be interpreting a point with high values in this plot.
+The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
+The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
 
-The PI (dashed line) can be viewed against the CI (dotted line). Now,
-one needs to pay attention to the parallel plot of the SHAP values,
-which are local to a particular observation, and the density plot, which
-is the same projection of all observations as specified by the SHAP
-values of the PI. The variable contribution of the two different
-predictions can be quickly compared in the parallel coordinate plot. The
-PI differs from the comparison primarily on the Fiber variable, which
-suggests that this is the reason for the incorrect prediction.
+The PI (dashed line) can be viewed against the CI (dotted line).
+Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the PI.
+The variable contribution of the two different predictions can be quickly compared in the parallel coordinate plot.
+The PI differs from the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
 
-From the density plot, which is the attribution projection corresponding
-to the PI, both observations are more like dark chocolates. Varying the
-contribution of Sugars and altogether removing it from the projection is
-where the difference becomes apparent. When a frame with contribution
-primarily from Fiber is examined observation 22 looks more like a milk
-chocolate.
+From the density plot, which is the attribution projection corresponding to the PI, both observations are more like dark chocolates.
+Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
+When a frame with contribution primarily from Fiber is examined observation 22 looks more like a milk chocolate.
 
-It would also be interesting to explore an inverse misclassification. In
-this case, a milk chocolate is selected while it was misclassified as a
-dark chocolate. Chocolate 84 is just this case and is compared with a
-correctly predicted milk chocolate (observation 71). The corresponding
-global view and radial tour frames are shown in Figure
-\ref{fig:casechocolatesinverse}.
+It would also be interesting to explore an inverse misclassification.
+In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
+Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (observation 71).
+The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
+
+(ref:casechocolatesinverse-cap) Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+In the attribution projection, the PI could be either milk or dark.
+Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates.
+The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
+If Sodium had a larger contribution to the prediction (like in this view).
+the PI would look more like other milk chocolates.
+(A video of the tour animation can be found at \url{https://vimeo.com/666431148}.)
 
@@
 
-\caption[Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the local explanation for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
-In the attribution projection, the PI could be either milk or dark.
-Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates.
-The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied.
-If Sodium had a larger contribution to the prediction (like in this view).
-the PI would look more like other milk chocolates.
-(A video of the tour animation can be found at https://vimeo.com/666431148.)}\label{fig:casechocolatesinverse}
+\caption[(ref:casechocolatesinverse-cap)]{(ref:casechocolatesinverse-cap)}(\#fig:casechocolatesinverse)
 \end{figure}
@@
 
-The difference of position in the tree SHAP PCA with the previous case
-is quite significant; this gives a higher-level sense that the
-attributions should be quite different. Looking at the attribution
-projection, this is found to be the case. Previously, Fiber was
-essential while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat have high attributions here,
-while they were unimportant in the preceding case.
+The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
+Looking at the attribution projection, this is found to be the case.
+Previously, Fiber was essential while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
 
-Comparing the attribution with the CI (dotted line), large discrepancies
-in Sodium and Fiber are identified. The contribution of Sodium is
-selected to be varied. Even in the initial projection, the observation
-looks slightly more like its observed milk than predicted dark
-chocolate. The misclassification appears least supported when the basis
-reaches sodium attribution of typical dark chocolate.
+Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
+The contribution of Sodium is selected to be varied.
+Even in the initial projection, the observation looks slightly more like its observed milk than predicted dark chocolate.
+The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
 
@@
 
-The 2020 season FIFA data \citep{leone_fifa_2020, biecek_dalex_2018}
-contains many skill measurements of soccer/football players and wage
-information. Nine higher-level skill groupings were identified and
-aggregated from highly correlated variables. A random forest model is
-fit from these predictors, regressing player wages {[}2020 euros{]}. The
-model was fit from 5000 observations before being thinned to 500 players
-to mitigate occlusion and render time. Continuing from the exploration
-in Section \textbackslash ref\{sec:explanations), we are interested to
-see the difference in attribution based on the exogenous player
-position. That is, the model should be able to use multiple linear
-profiles to better predict the wages from different field positions of
-players despite not having this information. A leading offensive fielder
-(L. Messi) is compared with a top defensive fielder (V. van Dijk). The
-same observations were used in Figure \ref{fig:shapdistrbd}.
+The 2020 season FIFA data \citep{leone_fifa_2020, biecek_dalex_2018} contains many skill measurements of soccer/football players and wage information.
+Nine higher-level skill groupings were identified and aggregated from highly correlated variables.
+A random forest model is fit from these predictors, regressing player wages {[}2020 euros{]}.
+The model was fit from 5000 observations before being thinned to 500 players to mitigate occlusion and render time.
+Continuing from the exploration in Section \textbackslash ref\{sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
+A leading offensive fielder (L.
+Messi) is compared with a top defensive fielder (V.
+van Dijk).
+The same observations were used in Figure \ref{fig:shapdistrbd}.
+
+(ref:casefifa-cap) Exploring the wages relative to skill measurements in the FIFA 2020 data.
+Star offensive player (L.
+Messi) is the PI, and he is compared with a top defensive player (V.
+van Dijk).
+The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
+Defense (\texttt{def}) is the chosen variable to vary.
+It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
+The increased contribution in defense comes at the expense of offensive and reaction skills.
+The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
+(A video of the animated radial tour can be found at \url{https://vimeo.com/666431163}.)
 
@@
 
-\caption[Exploring the wages relative to skill measurements in the FIFA 2020 data]{Exploring the wages relative to skill measurements in the FIFA 2020 data.
-Star offensive player (L.
-Messi) is the PI, and he is compared with a top defensive player (V.
-van Dijk).
-The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages.
-Defense (`def`) is the chosen variable to vary.
-It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot).
-The increased contribution in defense comes at the expense of offensive and reaction skills.
-The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the local explanation.
-(A video of the animated radial tour can be found at https://vimeo.com/666431163.)}\label{fig:casefifa}
+\caption[(ref:casefifa-cap)]{(ref:casefifa-cap)}(\#fig:casefifa)
 \end{figure}
@@
 Figure \ref{fig:casefifa} tests the support of the local explanation.
-Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both
-crucial to explaining a star offensive player. If either of them were
-rotated out, the other would be rotated into the frame, maintaining a
-far-right position. However, increasing the contribution of a variable
-with low importance would rotate both variables out of the frame.
+Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
+If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
+However, increasing the contribution of a variable with low importance would rotate both variables out of the frame.
 
-The contribution from \texttt{def} will be varied to contrast with
-offensive skills. As the contribution of defensive skills increases,
-Messi's is no longer separated from the group. Players with high values
-in defensive skills are now the rightmost points. In terms of what-if
-analysis, the difference between the data mean and his predicted wages
-would be halved if Messi's tree SHAP attributions were at these levels.
+The contribution from \texttt{def} will be varied to contrast with offensive skills.
+As the contribution of defensive skills increases, Messi's is no longer separated from the group.
+Players with high values in defensive skills are now the rightmost points.
+In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions were at these levels.
+
+(ref:caseames-cap) Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
+The local explanation indicates a sizable attribution to Lot Area (\texttt{LtA}), while the CI has minimal attribution to this variable.
+The PI has a higher predicted value than the CI in the attribution projection.
+Reducing the contribution of Lot Area brings these two prices in line.
+This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar.
+That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
+(A video showing the animation is at \url{https://vimeo.com/666431134}.)
 
@@
 
-\caption[Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data]{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
-The local explanation indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
-The PI has a higher predicted value than the CI in the attribution projection.
-Reducing the contribution of Lot Area brings these two prices in line.
-This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar.
-That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price.
-(A video showing the animation is at https://vimeo.com/666431134.)}\label{fig:caseames}
+\caption[(ref:caseames-cap)]{(ref:caseames-cap)}(\#fig:caseames)
 \end{figure}
@@
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{Ames Housing 2018, Sales Price
-Regression}\label{ames-housing-2018-sales-price-regression}}
+\subsection{Ames Housing 2018, Sales Price Regression}\label{ames-housing-2018-sales-price-regression}}
 
-Ames housing data 2018 \citep{de_cock_ames_2011, prevek18_ames_2018} was
-subset to North Ames (the neighborhood with the most house sales). The
-remaining are 338 house sales. A random forest model was fit, predicting
-the sale price {[}USD{]} from the property variables: Lot Area
-(\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built
-(\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms
-(\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of
-Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and
-Garage Area (\texttt{GrA}). Using interactions with the global view, a
-house with an extreme negative residual and an accurate observation with
-a similar prediction is selected.
+Ames housing data 2018 \citep{de_cock_ames_2011, prevek18_ames_2018} was subset to North Ames (the neighborhood with the most house sales).
+The remaining are 338 house sales.
+A random forest model was fit, predicting the sale price {[}USD{]} from the property variables: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+Using interactions with the global view, a house with an extreme negative residual and an accurate observation with a similar prediction is selected.
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under
-prediction with an enormous Lot Area contribution. The CI has a similar
-predicted price though the prediction was accurate and gives almost no
-attribution to lot size. The attribution projection places observations
-with high Living Areas to the right. The contribution of Living Area
-contrasts the contribution of this variable. As the contribution of Lot
-Area decreases, the predictive power decreases for the PI, while the CI
-remains stationary. This large importance in the Living Area is
-relatively uncommon. Boosting tree models may be more resilient to such
-an under-prediction as they would up-weighting this residual and force
-its inclusion in the final model.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
+The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
+The attribution projection places observations with high Living Areas to the right.
+The contribution of Living Area contrasts the contribution of this variable.
+As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
+This large importance in the Living Area is relatively uncommon.
+Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
 
@@
 
-There is a clear need to extend the interpretability of black box
-models. With techniques such as SHAP, LIME, Break-down, one can
-calculate local explanations, i.e.~for every observation in the data.
-These techniques quantify for each observation how strongly particular
-variables affect the model's predictions. Surprisingly few techniques
-allow us to understand the global distribution of these local
-explanations. Unsupervised data exploration techniques applied to data
-show how useful they are for identifying outliers, identifying segments
-of observations or discovering correlations between variables. All of
-these tasks can be performed for a set of explanations.
+There is a clear need to extend the interpretability of black box models.
+This paper provides a technique that builds on local explanations to explore the variable importance local to an observation.
+The local explanations of an attribution projection from which variable contributions are varied using a radial tour.
+Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular variables.
+A global view shows the data space, explanation space, and residual plot.
+The user can interactively select observations to compare, contrast, and study further.
+Then the radial tour is used to explore the variable sensitivity identified by the attribution projection.
 
-To address this challenge in this paper provides a technique that builds
-on local explanations to explore the variable importance local to an
-observation. The local explanations of an attribution projection from
-which variable contributions are varied using a radial tour. Several
-diagnostic plots are provided to assist with understanding the
-sensitivity of the prediction to particular variables. A global view
-shows the data space, explanation space, and residual plot. The user can
-interactively select observations to compare, contrast, and study
-further. Then the radial tour is used to explore the variable
-sensitivity identified by the attribution projection.
+This approach has been illustrated using four data examples of random forest models with the tree SHAP local explanation.
+Local explanations focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
+They can also form the basis of learning how the model has got it wrong, with when the observation is misclassified or has a large residual.
 
-This approach has been illustrated using four data examples of random
-forest models with the tree SHAP local explanation. Local explanations
-focus on the model fit, and help to dissect which variables are most
-responsible for the fitted value. They can also form the basis of
-learning how the model has got it wrong, with when the observation is
-misclassified or has a large residual.
+In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
+This was verified by making a follow-up plot of the data.
+The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
+In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area is not being effectively used by the random forest model.
 
-In the penguins example, we showed how the misclassification of a
-penguin arose due to it having an unusually small flipper size compared
-to others of its species. This was verified by making a follow-up plot
-of the data. The chocolates example shows how a dark chocolate was
-misclassified primarily due to its attribution to Fiber, and a milk
-chocolate was misclassified as dark due to its lowish Sodium value. In
-the FIFA example, we show how low Messi's salary would be if it depended
-on defensive skills. In the Ames housing data, an inaccurate prediction
-for a house was likely due to the Lot Area is not being effectively used
-by the random forest model.
-
-This analysis is manually intensive and thus only feasible for
-investigating a few observations. The recommended approach is to
-investigate an observation where the model has not predicted accurately
-and compare it with an observation with similar predictor values where
-model fitted well. The radial tour launches from the attribution
-projection to enable exploration of the sensitivity of the prediction to
-any variable. It can be helpful to make additional plots of the
-variables and responses to cross-check interpretations made from the
-cheem viewer. This methodology provides an additional tool in the box
-for studying model fitting.
+This analysis is manually intensive and thus only feasible for investigating a few observations.
+The recommended approach is to investigate an observation where the model has not predicted accurately and compare it with an observation with similar predictor values where model fitted well.
+The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any variable.
+It can be helpful to make additional plots of the variables and responses to cross-check interpretations made from the cheem viewer.
+This methodology provides an additional tool in the box for studying model fitting.
 
@@
 
-An implementation is provided in the open-source \textbf{R} package
-\textbf{cheem}, available on CRAN at
-\url{https://CRAN.R-project.org/package=cheem}. Example data sets are
-provided, and you can upload your data after model fitting and computing
-the local explanations. The local explanations need to be pre-computed
-and uploaded. Examples show how to do this for tree SHAP values, using
-\textbf{treeshap} (tree based models from \textbf{gbm},
-\textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or
-\textbf{xgboost} \citet{greenwell_gbm_2020}; \citet{shi_lightgbm_2022};
-\citet{liaw_classification_2002}; \citet{wright_ranger_2017};
-\citet{chen_xgboost_2021}, respectively). The SHAP and oscillation
-explanations could be easily added using \texttt{DALEX::explain()}
-\citep{biecek_dalex_2018, biecek_explanatory_2021}.
+An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
+Example data sets are provided, and you can upload your data after model fitting and computing the local explanations.
+The local explanations need to be pre-computed and uploaded.
+Examples show how to do this for tree SHAP values, using \textbf{treeshap} (tree based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} \citet{greenwell_gbm_2020}; \citet{shi_lightgbm_2022}; \citet{liaw_classification_2002}; \citet{wright_ranger_2017}; \citet{chen_xgboost_2021}, respectively).
+The SHAP and oscillation explanations could be easily added using \texttt{DALEX::explain()} \citep{biecek_dalex_2018, biecek_explanatory_2021}.
 
 The application was made with \textbf{shiny} \citep{chang_shiny_2021}.
-The tour visual is built with \textbf{spinifex}
-\citep{spyrison_spinifex_2020}. Both views are created first with
-\textbf{ggplot2} \citep{wickham_ggplot2_2016} and then rendered as
-interactive \texttt{html} widgets with \textbf{plotly}
-\citep{sievert_interactive_2020}. \textbf{DALEX}
-\citep{biecek_dalex_2018} and the free ebook, \emph{Explanatory Model
-Analysis} \citep{biecek_explanatory_2021} were a boon to understanding
-local explanations and how to apply them.
+The tour visual is built with \textbf{spinifex} \citep{spyrison_spinifex_2020}.
+Both views are created first with \textbf{ggplot2} \citep{wickham_ggplot2_2016} and then rendered as interactive \texttt{html} widgets with \textbf{plotly} \citep{sievert_interactive_2020}.
+\textbf{DALEX} \citep{biecek_dalex_2018} and the free ebook, \emph{Explanatory Model Analysis} \citep{biecek_explanatory_2021} were a boon to understanding local explanations and how to apply them.
 
-The package can be installed from CRAN, and the application can be run
-using the following \textbf{R} code:
+The package can be installed from CRAN, and the application can be run using the following \textbf{R} code:
 
@@
 \item
-  The development version of the package is available at
-  \url{https://github.com/nspyrison/cheem}, and
+  The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
 \item
-  Documentation of the package can be found at
-  \url{https://nspyrison.github.io/cheem/}.
+  Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
 \end{itemize}
 
-Follow the examples provided with the package to compute the local
-explanation (using \texttt{?cheem\_ls}). The application expects the
-output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file
-with \texttt{saveRDS()} to be uploaded.
+Follow the examples provided with the package to compute the local explanation (using \texttt{?cheem\_ls}).
+The application expects the output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file with \texttt{saveRDS()} to be uploaded.
 
@@
 
-Kim Marriott provided advice on many aspects of this work, especially on
-the explanations in the applications section. This research was
-supported by the Australian Government Research Training Program (RTP)
-scholarships. Thanks to Jieyang Chong for helping proofread this
-article. The namesake, Cheem, refers to a fictional race of humanoid
-trees from Doctor Who lore. \textbf{DALEX} pulls on from that universe,
-and we initially apply tree SHAP explanations specific to tree-based
-models.
+Kim Marriott provided advice on many aspects of this work, especially on the explanations in the applications section.
+This research was supported by the Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
 

### 2023-05-22 jss/cheem_paper.Rmd
edits through-out

@@
 To understand how the interaction between predictors affects the variable importance estimate, we can convert LVAs into linear projections and use the radial tour.
-This is also useful for learning how a model has made a mistake, or the effect of outliers, or clustering of observations.
-The approach is llustrated with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
+This is also useful for learning how a model has made a mistake, or the effect of outliers, or the clustering of observations.
+The approach is illustrated with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
 The methods are implemented in the R package cheem, available on CRAN.
@@
 Interpretability becomes more difficult when the model is nonlinear.
-Nonlinear models occur in statistical models with polynomial or interaction terms between quantitative predictors, and almost all computational models such as random forests, support-vector machines, or neural networks [e.g.
+Nonlinear models occur in statistical models with polynomial or interaction terms between quantitative predictors, and almost all computational models such as random forests, support vector machines, or neural networks [e.g.
 @breiman_random_2001; @boser_training_1992; @anderson_introduction_1995].
@@
 <!-- nonlinear leading to interpretability issues -->
-In linear models interpretation of the importance of variables is relatively straight forward, one adjusts for the covariance of multiple variables when examining the relationship with the response.
+In linear models interpretation of the importance of variables is relatively straightforward, one adjusts for the covariance of multiple variables when examining the relationship with the response.
 The interpretation is valid for the full domain of the predictors.
-In nonlinear models one needs to consider the model in small neighborhoods of the domain to make any assessment of variable importance.
+In nonlinear models, one needs to consider the model in small neighborhoods of the domain to make any assessment of variable importance.
 Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues described in @stahl-ethics.
@@
 <!-- tours and models -->
-Applying tours to model interpretation is recommended in @wickham_visualizing_2015 primarily to examine the fitted model in the space of the data.
+Applying tours to model interpretation is recommended by @wickham_visualizing_2015 primarily to examine the fitted model in the space of the data.
 @cook_interactive_2007 describe the use of tours for exploring classification boundaries and model diagnostics [@Caragea2008; @lee_pptree_2013; @da_silva_projection_2021].
@@
 Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative response variables.
-These show how the LVAs can be used to get an overview of the model's use of predictors, and to investigate errors in the model predictions.
+These show how the LVAs can be used to get an overview of the model's use of predictors and to investigate errors in the model predictions.
 Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
@@
 <!-- SHAP -->
-We will be focusing on SHAP values in this this paper, but the approach is applicable for any method used to calculate the LVAs.
+We will be focusing on SHAP values in this paper, but the approach is applicable to any method used to calculate the LVAs.
 SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
@@
 These values will add up to the difference between the fitted value for the observation, and the average fitted value for all observations.
-Using all possible sequence, or permutation, gives multiple values for each variable, which are averaged to get the SHAP value for an observation.
-It can be helpful to standardize variables prior to computing SHAP values, if they have been measured on different scales.
+Using all possible sequences, or permutations, gives multiple values for each variable, which are averaged to get the SHAP value for an observation.
+It can be helpful to standardize variables prior to computing SHAP values if they have been measured on different scales.
 
-The approach is related to partial dependence plots (see for example, chapter 8 of @molnar2022), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
+The approach is related to partial dependence plots (for example see chapter 8 of @molnar2022), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
 Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
@@
 Panel (a) is a version of a breakdown plot [@gosiewska_ibreakdown_2019] where just three sequences of variables are shown, for two observations.
-A breakdown plot shows the absolute values of the variable attribution for an observation, usually sorted from highest value to the lowest.
+A breakdown plot shows the absolute values of the variable attribution for an observation, usually sorted from the highest value to the lowest.
 There is no scale on the horizontal axis here because values are considered relative to each other.
 Here we can see how the variable contribution can change depending on sequence, relative to both players.
-(Note that the order of the variables is different in each plot, because they have been sorted by biggest average contribution across both players.) For all sequences, and for both players `reaction` has the strongest contribution, with perhaps more importance for the defensive player.
+(Note that the order of the variables is different in each plot because they have been sorted by the biggest average contribution across both players.) For all sequences, and for both players `reaction` has the strongest contribution, with perhaps more importance for the defensive player.
 Then it differs by player: for Messi `offense` and `movement` have the strongest contributions, and for van Dijk it is `defense` and `power`, regardless of the variable sequence.
@@
 Panel (b) shows the differences in the player's median values (large dots) for 25 such sequences (tick marks).
-We can see that the wage predictions for the two players come from different combinations of skills sets, as might be expected for players who's value on the team depends on their offensive or defensive prowess.
-It is also interesting to see from the distribution of values across the different sequence of variables, that there is some multimodality.
-For example, look at the SHAP values for `reaction` for Messi, and in some sequences reaction has a much lower contribution than others.
+We can see that the wage predictions for the two players come from different combinations of skill sets, as might be expected for players whose value on the team depends on their offensive or defensive prowess.
+It is also interesting to see from the distribution of values across the different sequences of variables, that there is some multimodality.
+For example, look at the SHAP values for `reaction` for Messi, and in some sequences, reaction has a much lower contribution than others.
 This suggests that other variables (`offense`, `movement` probably) can substitute for `reaction` in the wage prediction.
@@
 
-There are numerous R packages currently available which provide functions for computing SHAP values, including `fastshap` [@fastshap], `kernelshap` [@kernelshap], `shapr` [@shapr], `shapviz` [@shapviz], `PPtreeregViz` [@PPtreeregViz], `ExplainPrediction` [@ExplainPrediction], `flashlight` [@flashlight] and the package `DALEX` has many resources [@biecek_dalex_2018].
+There are numerous R packages currently available that provide functions for computing SHAP values, including `fastshap` [@fastshap], `kernelshap` [@kernelshap], `shapr` [@shapr], `shapviz` [@shapviz], `PPtreeregViz` [@PPtreeregViz], `ExplainPrediction` [@ExplainPrediction], `flashlight` [@flashlight], and the package `DALEX` has many resources [@biecek_dalex_2018].
 There are many more packages only available through Github, like `treeshap` [@kominsarczyk_treeshap_2021] that is used for this work.
@@
 The manual tour [@cook_manual_1997] defines its path by changing a selected variable's contribution to a basis to allow the variable to contribute more or less to the projection.
-The requirement constrains the contribution of all other variables that a basis needs to be orthonormal (column correspond to vectors, with unit length, and orthogonal to each other).
-The manual tour is primarily used to assess the importance of a variable to structure visible in a projection.
-It also lends itself to pre-computation queued in advance or computed on-the-fly for human-in-the-loop analysis [@karwowski_international_2006].
+The requirement constrains the contribution of all other variables that a basis needs to be orthonormal (columns correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a variable to the structure visible in a projection.
+It also lends itself to pre-computation queued in advance or computed on the fly for human-in-the-loop analysis [@karwowski_international_2006].
 
@@
 
-```{r radialtour, echo=F, fig.align='center', out.width="99%", fig.cap = "The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to structure in the plot.
+```{r radialtour, echo=F, fig.align='center', out.width="99%", fig.cap = "The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to the structure in the plot.
 Here we have a 1D projection of the penguins data displayed as a density plot.
@@
 In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify observations poorly predicted by the model.
-<!-- Interactions --> For classification tasks, color indicates the predicted class, and misclassified observations are circled in red.
-Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
+<!-- Interactions --> For classification tasks, color indicates the predicted class and misclassified observations are circled in red.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate the exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
 
 While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular observations to explore in detail.
-We have designed it to enable a comparison between an observation that is interesting is some way, perhaps misclassified, or poorly predicted, relative to an observation with similar predictor values but more expected prediction.
-For brevity, we call the interesting observation the primary investigation (PI) and the other is the comparison investigation (CI).
-These observations are highlighted as asterisk and $\times$, respectively.
+We have designed it to enable a comparison between an observation that is interesting in some way, perhaps misclassified, or poorly predicted, relative to an observation with similar predictor values but a more expected prediction.
+For brevity, we call the interesting observation the primary investigation (PI), and the other is the comparison investigation (CI).
+These observations are highlighted as an asterisk and $\times$, respectively.
 
@@
 The first (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a display of the SHAP values for all observations.
-This will generally give the global view of variables important for the fit as a whole, but it will also highlight observations which have different patterns.
+This will generally give the global view of variables important for the fit as a whole, but it will also highlight observations that have different patterns.
 The second plot is the radial tour, which for classification is a density plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for regression are scatterplots of the observed response values, and residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
@@
 (The SHAP values of the PI and CI are shown as dashed and dotted lines, respectively.) One should obtain a sense of the overall importance of variables from this plot.
-The more important variables will have larger values, and in the case of classification tasks variables which have different magnitude for different classes are more globally important.
+The more important variables will have larger values, and in the case of classification tasks variables that have different magnitudes for different classes are more globally important.
 For example, Figure \ref{fig:classificationcase}e suggests that `bl` is important for distinguishing the green class from the other two.
@@
 <!-- Segue of PI to attribution projection -->
-A bar chart is overlaid to represent the projection shown in the radial tour at right.
+A bar chart is overlaid to represent the projection shown in the radial tour on the right.
 It starts from the SHAP values of the PI, but if the user changes the projection the length of these bars will reflect this change.
@@
 Four selected points are highlighted in the PC spaces and tabularly displayed.
-Coloring on a statistic (c) highlights structure organized in the attribution space.
+Coloring on a statistic (c) highlights the structure organized in the attribution space.
 Interactive tabular display (d) populates when observations are selected.
@@
 <!-- Exploratory interactions -->
-A tooltip displays observation number/name and classification information while the cursor hovers over a point.
+A tooltip displays the observation number/name and classification information while the cursor hovers over a point.
 Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a & b).
 The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
-These interactions aid exploration of the spaces and, finally, the identification of primary and comparison observations.
+These interactions aid the exploration of the spaces and, finally, the identification of primary and comparison observations.
 
@@
 <!-- Switch to rmd bullets, consistent with thesis -->
-- __Data:__ predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep normality assumptions of the model in mind. 
+- __Data:__ predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep the normality assumptions of the model in mind.
 - __Model:__ any model and compatible explanation could be explored with this method.
@@
 The time to preprocess the data will vary significantly with the complexity of the model and the LE.
-For reference, the FIFA data contained 5000 observations of nine explanatory variables took 2.5 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 270 seconds total.
+For reference, the FIFA data contained 5000 observations of nine explanatory variables that took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 270 seconds in total.
 PCA and statistics of the variables and attributions took 2.8 seconds.
-These run times were from a non-parallelized session on a modern laptop, but suffice to say that most of the time will be spent on the LVA.
+These run times were from a non-parallelized session on a modern laptop, but suffice it to say that most of the time will be spent on the LVA.
 An increase in model complexity or data dimensionality will quickly become an obstacle.
 Its reduced computational complexity makes tree SHAP an excellent candidate to start.
-Alternatively, the package __fastshap__ claims extremely low run times, attributed to fewer calls to the prediction function, partial implementation in C++, and efficient use of logical subsetting, @greenwell_fastshap_2020.
+Alternatively, some package and methods use approximate calculations of LEs, such as __fastshap__ @greenwell_fastshap_2020.
 
@@
 The predicted vs observed plot reveals a handful of misclassified observations.
-A Gentoo that has been wrongly labeled as a Chinstrap is selected for illustration.
+A Gentoo which has been wrongly labeled as a Chinstrap is selected for illustration.
 The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
@@
 Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
-On varying `fl` as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
+On varying `fl`, as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
 This suggests that `fl` should be considered an important variable for explaining the (wrong) prediction.
@@
 So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber.
-When Sugar is zero, Fiber contributes strongly towards the left.
+When Sugar is zero, Fiber contributes strongly toward the left.
 In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber.
@@
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under prediction with an enormous Lot Area contribution.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under-prediction with an enormous Lot Area contribution.
 The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
@@
 
-To address this challenge in this paper provides a technique that builds on LEs to explore the variable importance local to an observation.
+To address this challenge this paper provides a technique that builds on LEs to explore the variable importance local to an observation.
 The LVA is converted into an attribution projection from which variable contributions are varied using a radial tour.
@@
 This approach has been illustrated using four data examples of random forest models with the tree SHAP LVA.
-LEs focus on the model fit, and help to dissect which variables are most responsible for the fitted value.
-They can also form the basis of learning how the model has got it wrong, with when the observation is misclassified or has a large residual.
+LEs focus on the model fit and help to dissect which variables are most responsible for the fitted value.
+They can also form the basis of learning how the model has got it wrong, when the observation is misclassified or has a large residual.
 
@@
 The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
-In the FIFA example, we show how low Messi's salary would be if it depended on defensive skills.
-In the Ames housing data, an inaccurate prediction for a house was likely due to the Lot Area is not being effectively used by the random forest model.
+In the FIFA example, we show how low Messi's salary would be if it depended on their defensive skill.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the lot area not being effectively used by the random forest model.
 
 This analysis is manually intensive and thus only feasible for investigating a few observations.
-The recommended approach is to investigate an observation where the model has not predicted accurately and compare it with an observation with similar predictor values where model fitted well.
+The recommended approach is to investigate an observation where the model has not predicted accurately and compare it with an observation with similar predictor values where the model fitted well.
 The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any variable.
@@
 The LVAs need to be pre-computed and uploaded.
-Examples show how to do this for tree SHAP values, using __treeshap__ (tree based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__ @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively).
+Examples show how to do this for tree SHAP values, using __treeshap__ (tree-based models from __gbm__, __lightgbm__, __randomForest__, __ranger__, or __xgboost__ @greenwell_gbm_2020; @shi_lightgbm_2022; @liaw_classification_2002; @wright_ranger_2017; @chen_xgboost_2021, respectively).
 The SHAP and oscillation explanations could be easily added using `DALEX::explain()` [@biecek_dalex_2018; @biecek_explanatory_2021].
@@
 Both views are created first with __ggplot2__ [@wickham_ggplot2_2016] and then rendered as interactive `html` widgets with __plotly__ [@sievert_interactive_2020].
-__DALEX__ [@biecek_dalex_2018] and _Explanatory Model Analysis_ [@biecek_explanatory_2021] are helpfule for understanding LEs and how to apply them.
+__DALEX__ [@biecek_dalex_2018] and _Explanatory Model Analysis_ [@biecek_explanatory_2021] are helpful for understanding LEs and how to apply them.
 

### 2023-05-22 jss/cheem_paper.tex
-fix app link -minor lint&wordsmithing -motiviating letter dates
edits through-out

@@
 also useful for learning how a model has made a mistake, or the effect
-of outliers, or clustering of observations. The approach is llustrated
-with examples from categorical (penguin species, chocolate types) and
-quantitative (soccer/football salaries, house prices) response models.
-The methods are implemented in the R package cheem, available on CRAN.
+of outliers, or the clustering of observations. The approach is
+illustrated with examples from categorical (penguin species, chocolate
+types) and quantitative (soccer/football salaries, house prices)
+response models. The methods are implemented in the R package cheem,
+available on CRAN.
 }
@@
 
-\usepackage{amsmath}
+\usepackage{amsmath} \usepackage{datetime}
 
@@
 interaction terms between quantitative predictors, and almost all
-computational models such as random forests, support-vector machines, or
+computational models such as random forests, support vector machines, or
 neural networks
@@
 In linear models interpretation of the importance of variables is
-relatively straight forward, one adjusts for the covariance of multiple
+relatively straightforward, one adjusts for the covariance of multiple
 variables when examining the relationship with the response. The
 interpretation is valid for the full domain of the predictors. In
-nonlinear models one needs to consider the model in small neighborhoods
+nonlinear models, one needs to consider the model in small neighborhoods
 of the domain to make any assessment of variable importance. Even though
@@
 XAI, complementing LVAs. Applying tours to model interpretation is
-recommended in \citet{wickham_visualizing_2015} primarily to examine the
+recommended by \citet{wickham_visualizing_2015} primarily to examine the
 fitted model in the space of the data. \citet{cook_interactive_2007}
@@
 variables. These show how the LVAs can be used to get an overview of the
-model's use of predictors, and to investigate errors in the model
+model's use of predictors and to investigate errors in the model
 predictions. Section \ref{sec:cheemdiscussion} concludes with a summary
@@
 
-We will be focusing on SHAP values in this this paper, but the approach
-is applicable for any method used to calculate the LVAs. SHAP calculates
-the variable contributions of one observation by examining the effect of
+We will be focusing on SHAP values in this paper, but the approach is
+applicable to any method used to calculate the LVAs. SHAP calculates the
+variable contributions of one observation by examining the effect of
 other variables on the predictions. The term ``SHAP'' refers to
@@
 fitted value for the observation, and the average fitted value for all
-observations. Using all possible sequence, or permutation, gives
+observations. Using all possible sequences, or permutations, gives
 multiple values for each variable, which are averaged to get the SHAP
 value for an observation. It can be helpful to standardize variables
-prior to computing SHAP values, if they have been measured on different
+prior to computing SHAP values if they have been measured on different
 scales.
 
-The approach is related to partial dependence plots (see for example,
+The approach is related to partial dependence plots (for example see
 chapter 8 of \citet{molnar2022}), used to explain the effect of a
@@
 absolute values of the variable attribution for an observation, usually
-sorted from highest value to the lowest. There is no scale on the
+sorted from the highest value to the lowest. There is no scale on the
 horizontal axis here because values are considered relative to each
@@
 depending on sequence, relative to both players. (Note that the order of
-the variables is different in each plot, because they have been sorted
-by biggest average contribution across both players.) For all sequences,
-and for both players \texttt{reaction} has the strongest contribution,
-with perhaps more importance for the defensive player. Then it differs
-by player: for Messi \texttt{offense} and \texttt{movement} have the
-strongest contributions, and for van Dijk it is \texttt{defense} and
-\texttt{power}, regardless of the variable sequence.
+the variables is different in each plot because they have been sorted by
+the biggest average contribution across both players.) For all
+sequences, and for both players \texttt{reaction} has the strongest
+contribution, with perhaps more importance for the defensive player.
+Then it differs by player: for Messi \texttt{offense} and
+\texttt{movement} have the strongest contributions, and for van Dijk it
+is \texttt{defense} and \texttt{power}, regardless of the variable
+sequence.
 
@@
 predictions for the two players come from different combinations of
-skills sets, as might be expected for players who's value on the team
+skill sets, as might be expected for players whose value on the team
 depends on their offensive or defensive prowess. It is also interesting
-to see from the distribution of values across the different sequence of
+to see from the distribution of values across the different sequences of
 variables, that there is some multimodality. For example, look at the
-SHAP values for \texttt{reaction} for Messi, and in some sequences
+SHAP values for \texttt{reaction} for Messi, and in some sequences,
 reaction has a much lower contribution than others. This suggests that
@@
 
-There are numerous R packages currently available which provide
-functions for computing SHAP values, including \texttt{fastshap}
-\citep{fastshap}, \texttt{kernelshap} \citep{kernelshap}, \texttt{shapr}
-\citep{shapr}, \texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
+There are numerous R packages currently available that provide functions
+for computing SHAP values, including \texttt{fastshap} \citep{fastshap},
+\texttt{kernelshap} \citep{kernelshap}, \texttt{shapr} \citep{shapr},
+\texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
 \citep{PPtreeregViz}, \texttt{ExplainPrediction}
-\citep{ExplainPrediction}, \texttt{flashlight} \citep{flashlight} and
+\citep{ExplainPrediction}, \texttt{flashlight} \citep{flashlight}, and
 the package \texttt{DALEX} has many resources \citep{biecek_dalex_2018}.
@@
 the contribution of all other variables that a basis needs to be
-orthonormal (column correspond to vectors, with unit length, and
+orthonormal (columns correspond to vectors, with unit length, and
 orthogonal to each other). The manual tour is primarily used to assess
-the importance of a variable to structure visible in a projection. It
-also lends itself to pre-computation queued in advance or computed
-on-the-fly for human-in-the-loop analysis
+the importance of a variable to the structure visible in a projection.
+It also lends itself to pre-computation queued in advance or computed on
+the fly for human-in-the-loop analysis
 \citep{karwowski_international_2006}.
@@
 
-\caption{The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to structure in the plot.
+\caption{The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to the structure in the plot.
 Here we have a 1D projection of the penguins data displayed as a density plot.
@@
 identify observations poorly predicted by the model. For classification
-tasks, color indicates the predicted class, and misclassified
+tasks, color indicates the predicted class and misclassified
 observations are circled in red. Linked brushing between the plots is
 provided, and a tabular display of selected points helps to facilitate
-exploration of the spaces and the model (shown in Figures
+the exploration of the spaces and the model (shown in Figures
 \ref{fig:regressioncase}d).
@@
 to explore in detail. We have designed it to enable a comparison between
-an observation that is interesting is some way, perhaps misclassified,
+an observation that is interesting in some way, perhaps misclassified,
 or poorly predicted, relative to an observation with similar predictor
-values but more expected prediction. For brevity, we call the
-interesting observation the primary investigation (PI) and the other is
+values but a more expected prediction. For brevity, we call the
+interesting observation the primary investigation (PI), and the other is
 the comparison investigation (CI). These observations are highlighted as
-asterisk and \(\times\), respectively.
+an asterisk and \(\times\), respectively.
 
@@
 give the global view of variables important for the fit as a whole, but
-it will also highlight observations which have different patterns. The
+it will also highlight observations that have different patterns. The
 second plot is the radial tour, which for classification is a density
@@
 important variables will have larger values, and in the case of
-classification tasks variables which have different magnitude for
+classification tasks variables that have different magnitudes for
 different classes are more globally important. For example, Figure
@@
 A bar chart is overlaid to represent the projection shown in the radial
-tour at right. It starts from the SHAP values of the PI, but if the user
-changes the projection the length of these bars will reflect this
+tour on the right. It starts from the SHAP values of the PI, but if the
+user changes the projection the length of these bars will reflect this
 change. (The PI is interactively selected by clicking on a point in the
@@
 
-A tooltip displays observation number/name and classification
+A tooltip displays the observation number/name and classification
 information while the cursor hovers over a point. Linked brushing allows
@@
 dynamic table (Figure \ref{fig:classificationcase}d). These interactions
-aid exploration of the spaces and, finally, the identification of
+aid the exploration of the spaces and, finally, the identification of
 primary and comparison observations.
@@
   matrix. Most models and local explanations are scale-invariant. Keep
-  normality assumptions of the model in mind.
+  the normality assumptions of the model in mind.
 \item
@@
 complexity of the model and the LE. For reference, the FIFA data
-contained 5000 observations of nine explanatory variables took 2.5
+contained 5000 observations of nine explanatory variables that took 2.5
 seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 270 seconds
+Extracting the tree SHAP values of each observation took 270 seconds in
 total. PCA and statistics of the variables and attributions took 2.8
 seconds. These run times were from a non-parallelized session on a
-modern laptop, but suffice to say that most of the time will be spent on
-the LVA. An increase in model complexity or data dimensionality will
+modern laptop, but suffice it to say that most of the time will be spent
+on the LVA. An increase in model complexity or data dimensionality will
 quickly become an obstacle. Its reduced computational complexity makes
-tree SHAP an excellent candidate to start. Alternatively, the package
-\textbf{fastshap} claims extremely low run times, attributed to fewer
-calls to the prediction function, partial implementation in C++, and
-efficient use of logical subsetting, \citet{greenwell_fastshap_2020}.
+tree SHAP an excellent candidate to start. Alternatively, some package
+and methods use approximate calculations of LEs, such as
+\textbf{fastshap} \citet{greenwell_fastshap_2020}.
 
@@
 data space, as would be expected. The predicted vs observed plot reveals
-a handful of misclassified observations. A Gentoo that has been wrongly
+a handful of misclassified observations. A Gentoo which has been wrongly
 labeled as a Chinstrap is selected for illustration. The PI is a
@@
 revealed about how the PI differed from other Gentoos). On varying
-\texttt{fl} as it contributes increasingly to the projection (b, right),
-more and more, this penguin looks like a Chinstrap. This suggests that
-\texttt{fl} should be considered an important variable for explaining
-the (wrong) prediction.
+\texttt{fl}, as it contributes increasingly to the projection (b,
+right), more and more, this penguin looks like a Chinstrap. This
+suggests that \texttt{fl} should be considered an important variable for
+explaining the (wrong) prediction.
 
@@
 
-Figure \ref{fig:caseames} selects the house sale 74, a sizable under
-prediction with an enormous Lot Area contribution. The CI has a similar
-predicted price though the prediction was accurate and gives almost no
-attribution to lot size. The attribution projection places observations
-with high Living Areas to the right. The contribution of Living Area
-contrasts the contribution of this variable. As the contribution of Lot
-Area decreases, the predictive power decreases for the PI, while the CI
-remains stationary. This large importance in the Living Area is
-relatively uncommon. Boosting tree models may be more resilient to such
-an under-prediction as they would up-weighting this residual and force
-its inclusion in the final model.
+Figure \ref{fig:caseames} selects the house sale 74, a sizable
+under-prediction with an enormous Lot Area contribution. The CI has a
+similar predicted price though the prediction was accurate and gives
+almost no attribution to lot size. The attribution projection places
+observations with high Living Areas to the right. The contribution of
+Living Area contrasts the contribution of this variable. As the
+contribution of Lot Area decreases, the predictive power decreases for
+the PI, while the CI remains stationary. This large importance in the
+Living Area is relatively uncommon. Boosting tree models may be more
+resilient to such an under-prediction as they would up-weighting this
+residual and force its inclusion in the final model.
 
@@
 
-To address this challenge in this paper provides a technique that builds
-on LEs to explore the variable importance local to an observation. The
-LVA is converted into an attribution projection from which variable
+To address this challenge this paper provides a technique that builds on
+LEs to explore the variable importance local to an observation. The LVA
+is converted into an attribution projection from which variable
 contributions are varied using a radial tour. Several diagnostic plots
@@
 This approach has been illustrated using four data examples of random
-forest models with the tree SHAP LVA. LEs focus on the model fit, and
+forest models with the tree SHAP LVA. LEs focus on the model fit and
 help to dissect which variables are most responsible for the fitted
 value. They can also form the basis of learning how the model has got it
-wrong, with when the observation is misclassified or has a large
-residual.
+wrong, when the observation is misclassified or has a large residual.
 
@@
 the FIFA example, we show how low Messi's salary would be if it depended
-on defensive skills. In the Ames housing data, an inaccurate prediction
-for a house was likely due to the Lot Area is not being effectively used
-by the random forest model.
+on their defensive skill. In the Ames housing data, an inaccurate
+prediction for a house was likely due to the lot area not being
+effectively used by the random forest model.
 
@@
 and compare it with an observation with similar predictor values where
-model fitted well. The radial tour launches from the attribution
+the model fitted well. The radial tour launches from the attribution
 projection to enable exploration of the sensitivity of the prediction to
@@
 the LVAs. The LVAs need to be pre-computed and uploaded. Examples show
-how to do this for tree SHAP values, using \textbf{treeshap} (tree based
+how to do this for tree SHAP values, using \textbf{treeshap} (tree-based
 models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest},
@@
 \citep{biecek_dalex_2018} and \emph{Explanatory Model Analysis}
-\citep{biecek_explanatory_2021} are helpfule for understanding LEs and
+\citep{biecek_explanatory_2021} are helpful for understanding LEs and
 how to apply them.
@@
   A version of the cheem viewer shiny app can be directly accessed at
-  \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
+  \url{https://ebsmonash.shinyapps.io/cheem/}.
 \item

### 2023-07-10 jss/cheem_paper.Rmd
failed attempt to go to springer nature template for Computational Statistics.

@@
 ---
-documentclass: jss
+output:
+  #bookdown::html_document2: default
+  bookdown::pdf_document2:
+    keep_tex: true
+#documentclass: sn-jnl
 author:
@@
   \usepackage{datetime}
-output: rticles::jss_article
+  \usepackage{subfig}
+#output: rticles::arxiv_article
 bibliography: paper.bib

### 2023-07-10 jss/cheem_paper.tex
failed attempt to go to springer nature template for Computational Statistics.

@@
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
 \documentclass[
-]{jss}
-
-%% recommended packages
-\usepackage{orcidlink,thumbpdf,lmodern}
-
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
   \usepackage[utf8]{inputenc}
-
-\author{
-Nicholas Spyrison~\orcidlink{0000-0002-8417-0212}\\Monash
-University \And Dianne Cook~\orcidlink{0000-0002-3813-7155}\\Monash
-University \And Przemyslaw
-Biecek~\orcidlink{0000-0001-8423-1823}\\University of Warsaw\\
-Warsaw University of Technology
-}
-\title{Exploring Local Explanations of Nonlinear Models Using Animated
-Linear Projections}
-
-\Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}
-\Plaintitle{Exploring Local Explanations of Nonlinear Models Using
-Animated Linear Projections}
-\Shorttitle{Exploring Local Explanations}
-
-\Abstract{
-The increased predictive power of machine learning models comes at the
-cost of increased complexity and loss of interpretability, particularly
-in comparison to parametric statistical models. This trade-off has led
-to the emergence of eXplainable AI (XAI) which provides methods, such as
-local explanations (LEs) and local variable attributions (LVAs), to shed
-light on how a model use predictors to arrive at a prediction. These
-provide a point estimate of the linear variable importance in the
-vicinity of a single observation. However, LVAs tend not to effectively
-handle association between predictors. To understand how the interaction
-between predictors affects the variable importance estimate, we can
-convert LVAs into linear projections and use the radial tour. This is
-also useful for learning how a model has made a mistake, or the effect
-of outliers, or the clustering of observations. The approach is
-illustrated with examples from categorical (penguin species, chocolate
-types) and quantitative (soccer/football salaries, house prices)
-response models. The methods are implemented in the R package cheem,
-available on CRAN.
-}
-
-\Keywords{explainable artificial intelligence, nonlinear model
-interpretability, visual analytics, local explanations, grand
-tour, radial tour}
-\Plainkeywords{explainable artificial intelligence, nonlinear model
-interpretability, visual analytics, local explanations, grand
-tour, radial tour}
-
-%% publication information
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{}
-%% \Acceptdate{2012-06-04}
-
-\Address{
-    Nicholas Spyrison\\
-    Monash University\\
-    Dept of Human Centred Computing\\
-Faculty of Information Technology\\
-  E-mail: \email{spyrison@gmail.com}\\
-  URL: \url{https://nspyrison.netlify.app}\\~\\
-      Dianne Cook\\
-    Monash University\\
-    Dept of Econometrics and Business Statistics\\
-
-    }
-
-% tightlist command for lists without linebreak
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\usepackage{amsmath} \usepackage{datetime}
+\setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdfkeywords={true},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\title{true}
+\author{true \and true \and true}
+\date{}
 
 \begin{document}
-
+\maketitle
+\begin{abstract}
+The increased predictive power of machine learning models comes at the cost of increased complexity and loss of interpretability, particularly in comparison to parametric statistical models.
+This trade-off has led to the emergence of eXplainable AI (XAI) which provides methods, such as local explanations (LEs) and local variable attributions (LVAs), to shed light on how a model use predictors to arrive at a prediction.
+These provide a point estimate of the linear variable importance in the vicinity of a single observation.
+However, LVAs tend not to effectively handle association between predictors.
+To understand how the interaction between predictors affects the variable importance estimate, we can convert LVAs into linear projections and use the radial tour.
+This is also useful for learning how a model has made a mistake, or the effect of outliers, or the clustering of observations.
+The approach is illustrated with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
+The methods are implemented in the R package cheem, available on CRAN.
+\end{abstract}
+
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
 \hypertarget{sec:intro}{%
@@
 
-There are different reasons and purposes for fitting a model. According
-to the taxonomies of \citet{breiman_statistical_2001} and
-\citet{shmueli_explain_2010}, it can be useful to group models into two
-types: explanatory and predictive. Explanatory modeling is used for
-inferential purposes, while predictive modeling focuses solely on the
-performance of an objective function. The intended use of the model has
-important implications for its selection and development.
-Interpretability is critical in explanatory modeling to draw meaningful
-inferential conclusions, such as which variables most contribute to a
-prediction or whether some observations are less well fit.
+There are different reasons and purposes for fitting a model.
+According to the taxonomies of Breiman (2001b) and Shmueli (2010), it can be useful to group models into two types: explanatory and predictive.
+Explanatory modeling is used for inferential purposes, while predictive modeling focuses solely on the performance of an objective function.
+The intended use of the model has important implications for its selection and development.
+Interpretability is critical in explanatory modeling to draw meaningful inferential conclusions, such as which variables most contribute to a prediction or whether some observations are less well fit.
 Interpretability becomes more difficult when the model is nonlinear.
-Nonlinear models occur in statistical models with polynomial or
-interaction terms between quantitative predictors, and almost all
-computational models such as random forests, support vector machines, or
-neural networks
-\citep[e.g.][]{breiman_random_2001, boser_training_1992, anderson_introduction_1995}.
-
-In linear models interpretation of the importance of variables is
-relatively straightforward, one adjusts for the covariance of multiple
-variables when examining the relationship with the response. The
-interpretation is valid for the full domain of the predictors. In
-nonlinear models, one needs to consider the model in small neighborhoods
-of the domain to make any assessment of variable importance. Even though
-this is difficult, it is especially important to interpret model fits as
-we become more dependent on nonlinear models for routine aspects of life
-to avoid issues described in \citet{stahl-ethics}. Understanding how
-nonlinear models behave when usage extrapolates outside the domain of
-predictors, either in sub-spaces where few samples were provided in the
-training set, or extending outside the domain. It is especially
-important because nonlinear models can vary wildly and predictions can
-be dramatically wrong in these areas.
-
-Explainable Artificial Intelligence (XAI) is an emerging field of
-research focused on methods for the interpreting of models
-\citep{adadi_peeking_2018, arrieta_explainable_2020}. A class of
-techniques, called \emph{local explanations} (LEs), provide methods to
-approximate linear variable importance, called local variable
-attributions (LVAs), at the location of each observation or the
-predictions at a specific point in the data domain. Because these are
-point-specific, it is challenging to comprehensively visualize them to
-understand a model. There are common approaches for visualizing
-high-dimensional data as a whole, but what is needed are new approaches
-for viewing these individual LVAs relative to the whole.
-
-For multivariate data visualization, a \emph{tour}
-\citep{asimov_grand_1985, buja_grand_1986, lee_state_2021} of linear
-data projections onto a lower-dimensional space, could be an element of
-XAI, complementing LVAs. Applying tours to model interpretation is
-recommended by \citet{wickham_visualizing_2015} primarily to examine the
-fitted model in the space of the data. \citet{cook_interactive_2007}
-describe the use of tours for exploring classification boundaries and
-model diagnostics
-\citep{Caragea2008, lee_pptree_2013, da_silva_projection_2021}. There
-are various types of tours. In a \emph{manual} or radial tour
-\citep{cook_manual_1997, spyrison_spinifex_2020}, the path of linear
-projections is defined by changing the contribution of a selected
-variable. We propose to use this to scrutinize the LVAs. This approach
-could be considered to be a counter-factual, what-if analysis, such as
-\emph{ceteris paribus} (``other things held constant'') profiles
-\citep{biecek_ceterisparibus_2020}.
-
-The remainder of this paper is organized as follows. Section
-\ref{sec:explanations} covers the background of the LEs and the
-traditional visuals produced. Section \ref{sec:tour} explains the tours
-and particularly the radial manual tour. Section \ref{sec:cheemviewer}
-discusses the visual layout in the graphical user interface and how it
-facilitates analysis, data pre-processing, and package infrastructure.
-Illustrations are provided in Section \ref{sec:casestudies} for a range
-of supervised learning tasks with categorical and quantitative response
-variables. These show how the LVAs can be used to get an overview of the
-model's use of predictors and to investigate errors in the model
-predictions. Section \ref{sec:cheemdiscussion} concludes with a summary
-of the insights gained. The methods are implemented in the \textbf{R}
-package \textbf{cheem}.
+Nonlinear models occur in statistical models with polynomial or interaction terms between quantitative predictors, and almost all computational models such as random forests, support vector machines, or neural networks (e.g.
+Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995).
+
+In linear models interpretation of the importance of variables is relatively straightforward, one adjusts for the covariance of multiple variables when examining the relationship with the response.
+The interpretation is valid for the full domain of the predictors.
+In nonlinear models, one needs to consider the model in small neighborhoods of the domain to make any assessment of variable importance.
+Even though this is difficult, it is especially important to interpret model fits as we become more dependent on nonlinear models for routine aspects of life to avoid issues described in Stahl (2021).
+Understanding how nonlinear models behave when usage extrapolates outside the domain of predictors, either in sub-spaces where few samples were provided in the training set, or extending outside the domain.
+It is especially important because nonlinear models can vary wildly and predictions can be dramatically wrong in these areas.
+
+Explainable Artificial Intelligence (XAI) is an emerging field of research focused on methods for the interpreting of models (Adadi and Berrada 2018; Barredo Arrieta et al.
+2020).
+A class of techniques, called \emph{local explanations} (LEs), provide methods to approximate linear variable importance, called local variable attributions (LVAs), at the location of each observation or the predictions at a specific point in the data domain.
+Because these are point-specific, it is challenging to comprehensively visualize them to understand a model.
+There are common approaches for visualizing high-dimensional data as a whole, but what is needed are new approaches for viewing these individual LVAs relative to the whole.
+
+For multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and Asimov 1986; S.
+Lee et al.
+2021) of linear data projections onto a lower-dimensional space, could be an element of XAI, complementing LVAs.
+Applying tours to model interpretation is recommended by Wickham, Cook, and Hofmann (2015) primarily to examine the fitted model in the space of the data.
+Cook, Swayne, and Buja (2007) describe the use of tours for exploring classification boundaries and model diagnostics (Caragea et al.
+2008; Y.
+D.
+Lee et al.
+2013; da Silva, Cook, and Lee 2021).
+There are various types of tours.
+In a \emph{manual} or radial tour (Cook and Buja 1997; Spyrison and Cook 2020), the path of linear projections is defined by changing the contribution of a selected variable.
+We propose to use this to scrutinize the LVAs.
+This approach could be considered to be a counter-factual, what-if analysis, such as \emph{ceteris paribus} (``other things held constant'') profiles (Biecek 2020).
+
+The remainder of this paper is organized as follows.
+Section \ref{sec:explanations} covers the background of the LEs and the traditional visuals produced.
+Section \ref{sec:tour} explains the tours and particularly the radial manual tour.
+Section \ref{sec:cheemviewer} discusses the visual layout in the graphical user interface and how it facilitates analysis, data pre-processing, and package infrastructure.
+Illustrations are provided in Section \ref{sec:casestudies} for a range of supervised learning tasks with categorical and quantitative response variables.
+These show how the LVAs can be used to get an overview of the model's use of predictors and to investigate errors in the model predictions.
+Section \ref{sec:cheemdiscussion} concludes with a summary of the insights gained.
+The methods are implemented in the \textbf{R} package \textbf{cheem}.
 
@@
 
-LVAs shed light on machine learning model fits by estimating linear
-variable importance in the vicinity of a single observation. There are
-many approaches for calculating LVAs. A comprehensive summary of the
-taxonomy of currently available methods is provided in Figure 6 by
-\citet{arrieta_explainable_2020}. It includes a large number of
-model-specific explanations such as deepLIFT
-\citep{shrikumar_not_2016, shrikumar_learning_2017}, a popular recursive
-method for estimating importance in neural networks. There are fewer
-model-agnostic methods, of which LIME \citep{ribeiro_why_2016} and
-SHaply Additive exPlanations (SHAP) \citep{lundberg_unified_2017}, are
-popular.
-
-These observation-level explanations are used in various ways depending
-on the data. In image classification, where pixels correspond to
-predictors, saliency maps overlay or offset a heatmap to indicate
-important pixels \citep{simonyan_deep_2014}. For example, pixels
-corresponding to snow may be highlighted as important contributors when
-distinguishing if a picture contains a coyote or husky. In text
-analysis, word-level contextual sentiment analysis highlights the
-sentiment and magnitude of influential words \citep{vanni_textual_2018}.
-In the case of numeric regression, they are used to explain additive
-contributions of variables from the model intercept to the observation's
-prediction \citep{ribeiro_why_2016}.
-
-We will be focusing on SHAP values in this paper, but the approach is
-applicable to any method used to calculate the LVAs. SHAP calculates the
-variable contributions of one observation by examining the effect of
-other variables on the predictions. The term ``SHAP'' refers to
-\citet{shapley_value_1953}'s method to evaluate an individual's
-contribution in cooperative games by assessing this player's performance
-in the presence or absence of other players.
-\citet{strumbelj_efficient_2010} introduced SHAP for LEs in machine
-learning models. Variable importance can depend on the sequence in which
-variables are entered into the model fitting process, thus for any
-sequence we get a set of variable contribution values for a single
-observation. These values will add up to the difference between the
-fitted value for the observation, and the average fitted value for all
-observations. Using all possible sequences, or permutations, gives
-multiple values for each variable, which are averaged to get the SHAP
-value for an observation. It can be helpful to standardize variables
-prior to computing SHAP values if they have been measured on different
-scales.
-
-The approach is related to partial dependence plots (for example see
-chapter 8 of \citet{molnar2022}), used to explain the effect of a
-variable by predicting the response for a range of values on this
-variable after fixing the value of all other variables to their mean.
-Though partial dependence plots are a global approximation of the
-variable importance, while SHAP is specific to one observation.
-
-\begin{CodeChunk}
+LVAs shed light on machine learning model fits by estimating linear variable importance in the vicinity of a single observation. There are many approaches for calculating LVAs.
+A comprehensive summary of the taxonomy of currently available methods is provided in Figure 6 by Barredo Arrieta et al.
+(2020).
+It includes a large number of model-specific explanations such as deepLIFT (Shrikumar et al.
+2016; Shrikumar, Greenside, and Kundaje 2017), a popular recursive method for estimating importance in neural networks.
+There are fewer model-agnostic methods, of which LIME (Ribeiro, Singh, and Guestrin 2016) and SHaply Additive exPlanations (SHAP) (Lundberg and Lee 2017), are popular.
+
+These observation-level explanations are used in various ways depending on the data.
+In image classification, where pixels correspond to predictors, saliency maps overlay or offset a heatmap to indicate important pixels (Simonyan, Vedaldi, and Zisserman 2014).
+For example, pixels corresponding to snow may be highlighted as important contributors when distinguishing if a picture contains a coyote or husky.
+In text analysis, word-level contextual sentiment analysis highlights the sentiment and magnitude of influential words (Vanni et al.
+2018).
+In the case of numeric regression, they are used to explain additive contributions of variables from the model intercept to the observation's prediction (Ribeiro, Singh, and Guestrin 2016).
+
+We will be focusing on SHAP values in this paper, but the approach is applicable to any method used to calculate the LVAs.
+SHAP calculates the variable contributions of one observation by examining the effect of other variables on the predictions.
+The term ``SHAP'' refers to Shapley (1953)'s method to evaluate an individual's contribution in cooperative games by assessing this player's performance in the presence or absence of other players.
+Strumbelj and Kononenko (2010) introduced SHAP for LEs in machine learning models.
+Variable importance can depend on the sequence in which variables are entered into the model fitting process, thus for any sequence we get a set of variable contribution values for a single observation.
+These values will add up to the difference between the fitted value for the observation, and the average fitted value for all observations.
+Using all possible sequences, or permutations, gives multiple values for each variable, which are averaged to get the SHAP value for an observation.
+It can be helpful to standardize variables prior to computing SHAP values if they have been measured on different scales.
+
+The approach is related to partial dependence plots (for example see chapter 8 of Molnar (2022)), used to explain the effect of a variable by predicting the response for a range of values on this variable after fixing the value of all other variables to their mean.
+Though partial dependence plots are a global approximation of the variable importance, while SHAP is specific to one observation.
+
 \begin{figure}
@@
 
-\caption[Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors]{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
+\caption{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
 A star offensive and defensive player are compared, L.
@@
 \end{figure}
-\end{CodeChunk}
-
-We use 2020 season FIFA data \citep{leone_fifa_2020} to illustrate SHAP
-following the procedures described in \citet{biecek_explanatory_2021}.
-There are 5000 observations of nine predictor variables measuring
-players' skills and one response variable, wages (in euros). A random
-forest model is fit regressing players' wages on the skill variables. In
-this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are
-compared for a star offensive player (L. Messi) and a prominent
-defensive player (V. van Dijk). We are interested in knowing how the
-skill variables locally contribute to the wage prediction of each
-player. A difference in the attribution of the variable importance
-across the two positions of the players can be expected. This would be
-interpreted as how a player's salary depends on which combination of
-skills. Panel (a) is a version of a breakdown plot
-\citep{gosiewska_ibreakdown_2019} where just three sequences of
-variables are shown, for two observations. A breakdown plot shows the
-absolute values of the variable attribution for an observation, usually
-sorted from the highest value to the lowest. There is no scale on the
-horizontal axis here because values are considered relative to each
-other. Here we can see how the variable contribution can change
-depending on sequence, relative to both players. (Note that the order of
-the variables is different in each plot because they have been sorted by
-the biggest average contribution across both players.) For all
-sequences, and for both players \texttt{reaction} has the strongest
-contribution, with perhaps more importance for the defensive player.
-Then it differs by player: for Messi \texttt{offense} and
-\texttt{movement} have the strongest contributions, and for van Dijk it
-is \texttt{defense} and \texttt{power}, regardless of the variable
-sequence.
-
-Panel (b) shows the differences in the player's median values (large
-dots) for 25 such sequences (tick marks). We can see that the wage
-predictions for the two players come from different combinations of
-skill sets, as might be expected for players whose value on the team
-depends on their offensive or defensive prowess. It is also interesting
-to see from the distribution of values across the different sequences of
-variables, that there is some multimodality. For example, look at the
-SHAP values for \texttt{reaction} for Messi, and in some sequences,
-reaction has a much lower contribution than others. This suggests that
-other variables (\texttt{offense}, \texttt{movement} probably) can
-substitute for \texttt{reaction} in the wage prediction.
-
-This can also be considered similar to examining the coefficients from
-all subsets regression, as described in
-\citet{wickham_visualizing_2015}. Various models that are similarly good
-might use different combinations of the variables. Examining the
-coefficients from multiple models helps to understand the relative
-importance of each variable in the context of all other variables. This
-is similar to the approach here with SHAP values, that by examining the
-variation in values across different permutations of variables, we can
-gain more understanding of the relationship between the response and
-predictors.
-
-For the application, we use \emph{tree SHAP}, a variant of SHAP that
-enjoys a lower computational complexity
-\citep{lundberg_consistent_2018}. Instead of aggregating over sequences
-of the variables, tree SHAP calculates observation-level variable
-importance by exploring the structure of the decision trees. Tree SHAP
-is only compatible with tree-based models. so random forests are used
-for illustration.
-
-There are numerous R packages currently available that provide functions
-for computing SHAP values, including \texttt{fastshap} \citep{fastshap},
-\texttt{kernelshap} \citep{kernelshap}, \texttt{shapr} \citep{shapr},
-\texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
-\citep{PPtreeregViz}, \texttt{ExplainPrediction}
-\citep{ExplainPrediction}, \texttt{flashlight} \citep{flashlight}, and
-the package \texttt{DALEX} has many resources \citep{biecek_dalex_2018}.
-There are many more packages only available through Github, like
-\texttt{treeshap} \citep{kominsarczyk_treeshap_2021} that is used for
-this work. \citet{molnar2022} provides good explanations of the
-different methods and how to apply them to different models.
+
+We use 2020 season FIFA data (Leone 2020) to illustrate SHAP following the procedures described in Biecek and Burzykowski (2021).
+There are 5000 observations of nine predictor variables measuring players' skills and one response variable, wages (in euros).
+A random forest model is fit regressing players' wages on the skill variables.
+In this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are compared for a star offensive player (L.
+Messi) and a prominent defensive player (V.
+van Dijk).
+We are interested in knowing how the skill variables locally contribute to the wage prediction of each player.
+A difference in the attribution of the variable importance across the two positions of the players can be expected.
+This would be interpreted as how a player's salary depends on which combination of skills.
+Panel (a) is a version of a breakdown plot (Gosiewska and Biecek 2019) where just three sequences of variables are shown, for two observations.
+A breakdown plot shows the absolute values of the variable attribution for an observation, usually sorted from the highest value to the lowest.
+There is no scale on the horizontal axis here because values are considered relative to each other.
+Here we can see how the variable contribution can change depending on sequence, relative to both players.
+(Note that the order of the variables is different in each plot because they have been sorted by the biggest average contribution across both players.) For all sequences, and for both players \texttt{reaction} has the strongest contribution, with perhaps more importance for the defensive player.
+Then it differs by player: for Messi \texttt{offense} and \texttt{movement} have the strongest contributions, and for van Dijk it is \texttt{defense} and \texttt{power}, regardless of the variable sequence.
+
+Panel (b) shows the differences in the player's median values (large dots) for 25 such sequences (tick marks).
+We can see that the wage predictions for the two players come from different combinations of skill sets, as might be expected for players whose value on the team depends on their offensive or defensive prowess.
+It is also interesting to see from the distribution of values across the different sequences of variables, that there is some multimodality.
+For example, look at the SHAP values for \texttt{reaction} for Messi, and in some sequences, reaction has a much lower contribution than others.
+This suggests that other variables (\texttt{offense}, \texttt{movement} probably) can substitute for \texttt{reaction} in the wage prediction.
+
+This can also be considered similar to examining the coefficients from all subsets regression, as described in Wickham, Cook, and Hofmann (2015).
+Various models that are similarly good might use different combinations of the variables.
+Examining the coefficients from multiple models helps to understand the relative importance of each variable in the context of all other variables.
+This is similar to the approach here with SHAP values, that by examining the variation in values across different permutations of variables, we can gain more understanding of the relationship between the response and predictors.
+
+For the application, we use \emph{tree SHAP}, a variant of SHAP that enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
+Instead of aggregating over sequences of the variables, tree SHAP calculates observation-level variable importance by exploring the structure of the decision trees.
+Tree SHAP is only compatible with tree-based models.
+so random forests are used for illustration.
+
+There are numerous R packages currently available that provide functions for computing SHAP values, including \texttt{fastshap} (Greenwell 2021), \texttt{kernelshap} (Mayer and Watson 2023), \texttt{shapr} (Sellereite, Jullum, and Redelmeier 2023), \texttt{shapviz} (Mayer 2023b), \texttt{PPtreeregViz} (E.-K.
+Lee and Cho 2022), \texttt{ExplainPrediction} (Robnik-Sikonja 2018), \texttt{flashlight} (Mayer 2023a), and the package \texttt{DALEX} has many resources (Biecek 2018).
+There are many more packages only available through Github, like \texttt{treeshap} (Kominsarczyk et al.
+2021) that is used for this work.
+Molnar (2022) provides good explanations of the different methods and how to apply them to different models.
 
@@
 
-A \emph{tour} enables the viewing of high-dimensional data by animating
-many linear projections with small incremental changes. It is achieved
-by following a path of linear projections (bases) of high-dimensional
-space. One key variable of the tour is the object permanence of the data
-points; one can track the relative change of observations in time and
-gain information about the relationships between points across multiple
-variables. There are various types of tours that are distinguished by
-how the paths are generated \citep{lee_state_2021, cook_grand_2008}.
-
-The manual tour \citep{cook_manual_1997} defines its path by changing a
-selected variable's contribution to a basis to allow the variable to
-contribute more or less to the projection. The requirement constrains
-the contribution of all other variables that a basis needs to be
-orthonormal (columns correspond to vectors, with unit length, and
-orthogonal to each other). The manual tour is primarily used to assess
-the importance of a variable to the structure visible in a projection.
-It also lends itself to pre-computation queued in advance or computed on
-the fly for human-in-the-loop analysis
-\citep{karwowski_international_2006}.
-
-A version of the manual tour called a \emph{radial tour} is implemented
-in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
-In a radial tour, the selected variable can change its magnitude of
-contribution but not its angle; it must move along the direction of its
-original contribution. The implementation allows for pre-computation and
-interactive re-calculation to focus on a different variable.
-
-\begin{CodeChunk}
+A \emph{tour} enables the viewing of high-dimensional data by animating many linear projections with small incremental changes.
+It is achieved by following a path of linear projections (bases) of high-dimensional space.
+One key variable of the tour is the object permanence of the data points; one can track the relative change of observations in time and gain information about the relationships between points across multiple variables.
+There are various types of tours that are distinguished by how the paths are generated (S.
+Lee et al.
+2021; Cook et al.
+2008).
+
+The manual tour (Cook and Buja 1997) defines its path by changing a selected variable's contribution to a basis to allow the variable to contribute more or less to the projection.
+The requirement constrains the contribution of all other variables that a basis needs to be orthonormal (columns correspond to vectors, with unit length, and orthogonal to each other).
+The manual tour is primarily used to assess the importance of a variable to the structure visible in a projection.
+It also lends itself to pre-computation queued in advance or computed on the fly for human-in-the-loop analysis (Karwowski 2006).
+
+A version of the manual tour called a \emph{radial tour} is implemented in Spyrison and Cook (2020) and forms the basis of this new work.
+In a radial tour, the selected variable can change its magnitude of contribution but not its angle; it must move along the direction of its original contribution.
+The implementation allows for pre-computation and interactive re-calculation to focus on a different variable.
+
 \begin{figure}
@@
 \end{figure}
-\end{CodeChunk}
 
@@
 
-To explore the LVAs, coordinated views \citep{roberts_state_2007}
-\citep[also known as ensemble graphics,][]{unwin_ensemble_2018} are
-provided in the \emph{cheem viewer} application. There are two primary
-plots: the \textbf{global view} to give the context of all of the SHAP
-values and the \textbf{radial tour view} to explore the LVAs with
-user-controlled rotation. There are numerous user inputs, including
-variable selection for the radial tour and observation selection for
-making comparisons. There are different plots used for the categorical
-and quantitative responses. Figures \ref{fig:classificationcase} and
-\ref{fig:regressioncase} are screenshots showing the cheem viewer for
-the two primary tasks: classification (categorical response) and
-regression (quantitative response).
+To explore the LVAs, coordinated views (Roberts 2007) (also known as ensemble graphics, Unwin and Valero-Mora 2018) are provided in the \emph{cheem viewer} application.
+There are two primary plots: the \textbf{global view} to give the context of all of the SHAP values and the \textbf{radial tour view} to explore the LVAs with user-controlled rotation.
+There are numerous user inputs, including variable selection for the radial tour and observation selection for making comparisons.
+There are different plots used for the categorical and quantitative responses.
+Figures \ref{fig:classificationcase} and \ref{fig:regressioncase} are screenshots showing the cheem viewer for the two primary tasks: classification (categorical response) and regression (quantitative response).
 
@@
 
-The global view provides context for all observations and facilitates
-the exploration of the separability of the data and attribution spaces.
+The global view provides context for all observations and facilitates the exploration of the separability of the data and attribution spaces.
 The attribution space refers to the SHAP values for each observation.
-These spaces both have dimensionality \(n \times p\), where \(n\) is the
-number of observations and \(p\) is the number of variables.
-
-The visualization is composed of the first two principal components of
-the data (left) and the attribution (middle) spaces. These single 2D
-projections will not reveal all of the structure of higher-dimensional
-space, but they are helpful visual summaries. In addition, a plot of the
-observed against predicted response values is also provided (Figures
-\ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help
-identify observations poorly predicted by the model. For classification
-tasks, color indicates the predicted class and misclassified
-observations are circled in red. Linked brushing between the plots is
-provided, and a tabular display of selected points helps to facilitate
-the exploration of the spaces and the model (shown in Figures
-\ref{fig:regressioncase}d).
-
-While the comparison of these spaces is interesting, the primary purpose
-of the global view is to enable the selection of particular observations
-to explore in detail. We have designed it to enable a comparison between
-an observation that is interesting in some way, perhaps misclassified,
-or poorly predicted, relative to an observation with similar predictor
-values but a more expected prediction. For brevity, we call the
-interesting observation the primary investigation (PI), and the other is
-the comparison investigation (CI). These observations are highlighted as
-an asterisk and \(\times\), respectively.
+These spaces both have dimensionality \(n \times p\), where \(n\) is the number of observations and \(p\) is the number of variables.
+
+The visualization is composed of the first two principal components of the data (left) and the attribution (middle) spaces.
+These single 2D projections will not reveal all of the structure of higher-dimensional space, but they are helpful visual summaries.
+In addition, a plot of the observed against predicted response values is also provided (Figures \ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help identify observations poorly predicted by the model.
+For classification tasks, color indicates the predicted class and misclassified observations are circled in red.
+Linked brushing between the plots is provided, and a tabular display of selected points helps to facilitate the exploration of the spaces and the model (shown in Figures \ref{fig:regressioncase}d).
+
+While the comparison of these spaces is interesting, the primary purpose of the global view is to enable the selection of particular observations to explore in detail.
+We have designed it to enable a comparison between an observation that is interesting in some way, perhaps misclassified, or poorly predicted, relative to an observation with similar predictor values but a more expected prediction.
+For brevity, we call the interesting observation the primary investigation (PI), and the other is the comparison investigation (CI).
+These observations are highlighted as an asterisk and \(\times\), respectively.
 
@@
 
-There are two plots in this part of the interface. The first (Figures
-\ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a
-display of the SHAP values for all observations. This will generally
-give the global view of variables important for the fit as a whole, but
-it will also highlight observations that have different patterns. The
-second plot is the radial tour, which for classification is a density
-plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for
-regression are scatterplots of the observed response values, and
-residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
-
-The LVAs for all observations are normalized (sum of squares equals 1),
-and thus, the relative importance of variables can be compared across
-all observations. These are depicted as a vertical parallel coordinate
-plot \citep{ocagne_coordonnees_1885}. (The SHAP values of the PI and CI
-are shown as dashed and dotted lines, respectively.) One should obtain a
-sense of the overall importance of variables from this plot. The more
-important variables will have larger values, and in the case of
-classification tasks variables that have different magnitudes for
-different classes are more globally important. For example, Figure
-\ref{fig:classificationcase}e suggests that \texttt{bl} is important for
-distinguishing the green class from the other two. For regression, one
-might generally observe which variables have low values for all
-observations (not important). For example, \texttt{BMI} and \texttt{pwr}
-in Figure \ref{fig:regressioncase}e, have a range of high and low values
-(e.g., \texttt{off}, \texttt{def}), suggesting they are important for
-some observations and not important for others.
-
-A bar chart is overlaid to represent the projection shown in the radial
-tour on the right. It starts from the SHAP values of the PI, but if the
-user changes the projection the length of these bars will reflect this
-change. (The PI is interactively selected by clicking on a point in the
-global view). By scaling the SHAP value it becomes an (attribution)
-projection.
-
-The attribution projection of the PI is the initial 1D basis in a radial
-tour, displayed as a density plot for a categorical response (Figure
-\ref{fig:classificationcase}f) and as scatterplots for a quantitative
-response (Figure \ref{fig:regressioncase}f). The PI and CI are indicated
-by vertical dashed and dotted lines, respectively. The radial tour
-varies the contribution of the selected variable. This is viewed as an
-animation of the projections from many intermediate bases. Doing so
-tests the sensitivity of structure (class separation or strength of
-relationship) to the variable's contribution. For classification, if the
-separation between classes diminishes when the variable contribution is
-reduced, this suggests that the variable is important for class
-separation. For regression, if the relationship scatterplot weakens when
-the variable contribution is reduced, indicating that the variable is
-important for accurately predicting the response.
+There are two plots in this part of the interface.
+The first (Figures \ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a display of the SHAP values for all observations.
+This will generally give the global view of variables important for the fit as a whole, but it will also highlight observations that have different patterns.
+The second plot is the radial tour, which for classification is a density plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for regression are scatterplots of the observed response values, and residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).
+
+The LVAs for all observations are normalized (sum of squares equals 1), and thus, the relative importance of variables can be compared across all observations.
+These are depicted as a vertical parallel coordinate plot (Ocagne 1885).
+(The SHAP values of the PI and CI are shown as dashed and dotted lines, respectively.) One should obtain a sense of the overall importance of variables from this plot.
+The more important variables will have larger values, and in the case of classification tasks variables that have different magnitudes for different classes are more globally important.
+For example, Figure \ref{fig:classificationcase}e suggests that \texttt{bl} is important for distinguishing the green class from the other two.
+For regression, one might generally observe which variables have low values for all observations (not important).
+For example, \texttt{BMI} and \texttt{pwr} in Figure \ref{fig:regressioncase}e, have a range of high and low values (e.g., \texttt{off}, \texttt{def}), suggesting they are important for some observations and not important for others.
+
+A bar chart is overlaid to represent the projection shown in the radial tour on the right.
+It starts from the SHAP values of the PI, but if the user changes the projection the length of these bars will reflect this change.
+(The PI is interactively selected by clicking on a point in the global view).
+By scaling the SHAP value it becomes an (attribution) projection.
+
+The attribution projection of the PI is the initial 1D basis in a radial tour, displayed as a density plot for a categorical response (Figure \ref{fig:classificationcase}f) and as scatterplots for a quantitative response (Figure \ref{fig:regressioncase}f).
+The PI and CI are indicated by vertical dashed and dotted lines, respectively.
+The radial tour varies the contribution of the selected variable.
+This is viewed as an animation of the projections from many intermediate bases.
+Doing so tests the sensitivity of structure (class separation or strength of relationship) to the variable's contribution.
+For classification, if the separation between classes diminishes when the variable contribution is reduced, this suggests that the variable is important for class separation.
+For regression, if the relationship scatterplot weakens when the variable contribution is reduced, indicating that the variable is important for accurately predicting the response.
 
@@
 
-Selecting a misclassified observation as PI and a correctly classified
-point nearby in data space as CI makes it easier to examine the
-variables most responsible for the error. The global view (Figure
-\ref{fig:classificationcase}c) displays the model confusion matrix. The
-radial tour is 1D and displays as density where color indicates class.
-An animation slider enables users to vary the contribution of variables
-to explore the sensitivity of the separation to that variable.
-
-\begin{CodeChunk}
+Selecting a misclassified observation as PI and a correctly classified point nearby in data space as CI makes it easier to examine the variables most responsible for the error.
+The global view (Figure \ref{fig:classificationcase}c) displays the model confusion matrix.
+The radial tour is 1D and displays as density where color indicates class.
+An animation slider enables users to vary the contribution of variables to explore the sensitivity of the separation to that variable.
+
 \begin{figure}
@@
 
-\caption[Overview of the cheem viewer for classification tasks (categorical response)]{Overview of the cheem viewer for classification tasks (categorical response).
+\caption{Overview of the cheem viewer for classification tasks (categorical response).
 Global view inputs, (a), set the PI, CI, and color statistic.
@@
 
-Selecting an inaccurately predicted observation as PI and an accurately
-predicted observation with similar variable values as CI is a helpful
-way to understand how the model is failing or not. The global view
-(Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed
-vs predicted values, which should exhibit a strong relationship if the
-model is a good fit. The points can be colored by a statistic, residual,
-a measure of outlyingness (log Mahalanobis distance), or correlation to
-aid in understanding the structure identified in these spaces.
-
-In the radial tour view, the observed response and the residuals
-(vertical) are plotted against the attribution projection of the PI
-(horizontal). The attribution projection can be interpreted similarly to
-the predicted value from the global view plot. It represents a linear
-combination of the variables, and a good fit would be indicated when
-there is a strong relationship with the observed values. This can be
-viewed as a local linear approximation if the fitted model is nonlinear.
-As the contribution of a variable is varied, if the value of the PI does
-not change much, it would indicate that the prediction for this
-observation is NOT sensitive to that variable. Conversely, if the
-predicted value varies substantially, the prediction is very sensitive
-to that variable, suggesting that the variable is very important for the
-PI's prediction.
-
-\begin{CodeChunk}
+Selecting an inaccurately predicted observation as PI and an accurately predicted observation with similar variable values as CI is a helpful way to understand how the model is failing or not.
+The global view (Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed vs predicted values, which should exhibit a strong relationship if the model is a good fit.
+The points can be colored by a statistic, residual, a measure of outlyingness (log Mahalanobis distance), or correlation to aid in understanding the structure identified in these spaces.
+
+In the radial tour view, the observed response and the residuals (vertical) are plotted against the attribution projection of the PI (horizontal).
+The attribution projection can be interpreted similarly to the predicted value from the global view plot.
+It represents a linear combination of the variables, and a good fit would be indicated when there is a strong relationship with the observed values.
+This can be viewed as a local linear approximation if the fitted model is nonlinear.
+As the contribution of a variable is varied, if the value of the PI does not change much, it would indicate that the prediction for this observation is NOT sensitive to that variable.
+Conversely, if the predicted value varies substantially, the prediction is very sensitive to that variable, suggesting that the variable is very important for the PI's prediction.
+
 \begin{figure}
@@
 
-\caption[Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables]{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
+\caption{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
@@
 
-The application has several reactive inputs that affect the data used,
-aesthetic display, and tour manipulation. These reactive inputs make the
-software flexible and extensible (Figure \ref{fig:classificationcase}a
-\& d). The application also has more exploratory interactions to help
-link points across displays, reveal structures found in different
-spaces, and access the original data.
-
-A tooltip displays the observation number/name and classification
-information while the cursor hovers over a point. Linked brushing allows
-the selection of points (left click and drag) where those points will be
-highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
-The information corresponding to the selected points is populated on a
-dynamic table (Figure \ref{fig:classificationcase}d). These interactions
-aid the exploration of the spaces and, finally, the identification of
-primary and comparison observations.
+The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation.
+These reactive inputs make the software flexible and extensible (Figure \ref{fig:classificationcase}a \& d).
+The application also has more exploratory interactions to help link points across displays, reveal structures found in different spaces, and access the original data.
+
+A tooltip displays the observation number/name and classification information while the cursor hovers over a point.
+Linked brushing allows the selection of points (left click and drag) where those points will be highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
+The information corresponding to the selected points is populated on a dynamic table (Figure \ref{fig:classificationcase}d).
+These interactions aid the exploration of the spaces and, finally, the identification of primary and comparison observations.
 
@@
 \item
-  \textbf{Data:} predictors and response are unscaled complete numerical
-  matrix. Most models and local explanations are scale-invariant. Keep
-  the normality assumptions of the model in mind.
+  \textbf{Data:} predictors and response are unscaled complete numerical matrix. Most models and local explanations are scale-invariant. Keep the normality assumptions of the model in mind.
 \item
-  \textbf{Model:} any model and compatible explanation could be explored
-  with this method. Currently, random forest models are applied via the
-  package \textbf{randomForest} \citep{liaw_classification_2002},
-  compatibility tree SHAP. Modest hyperparameters are used, namely: 125
-  trees, the number of variables at each split, mtry = \(\sqrt{p}\) or
-  \(p/3\) for classification and regression, and minimum size of
-  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
-  classification and regression.
+  \textbf{Model:} any model and compatible explanation could be explored with this method.
+Currently, random forest models are applied via the package \textbf{randomForest} (Liaw and Wiener 2002), compatibility tree SHAP.
+Modest hyperparameters are used, namely: 125 trees, the number of variables at each split, mtry = \(\sqrt{p}\) or \(p/3\) for classification and regression, and minimum size of terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for classification and regression.
 \item
-  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
-  observation using the package \textbf{treeshap}
-  \citep{kominsarczyk_treeshap_2021}. We opt to find the attribution of
-  each observation in the training data and not fit to fit variable
-  interactions.
+  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each} observation using the package \textbf{treeshap} (Kominsarczyk et al.
+2021).
+We opt to find the attribution of each observation in the training data and not fit to fit variable interactions.
 \item
-  \textbf{Cheem viewer:} after the model and full explanation space are
-  calculated, each variable is scaled by standard deviations away from
-  the mean to achieve common support for visuals. Statistics for mapping
-  to color are computed on the scaled spaces.
+  \textbf{Cheem viewer:} after the model and full explanation space are calculated, each variable is scaled by standard deviations away from the mean to achieve common support for visuals.
+Statistics for mapping to color are computed on the scaled spaces.
 \end{itemize}
 
-The time to preprocess the data will vary significantly with the
-complexity of the model and the LE. For reference, the FIFA data
-contained 5000 observations of nine explanatory variables that took 2.5
-seconds to fit a random forest model of modest hyperparameters.
-Extracting the tree SHAP values of each observation took 270 seconds in
-total. PCA and statistics of the variables and attributions took 2.8
-seconds. These run times were from a non-parallelized session on a
-modern laptop, but suffice it to say that most of the time will be spent
-on the LVA. An increase in model complexity or data dimensionality will
-quickly become an obstacle. Its reduced computational complexity makes
-tree SHAP an excellent candidate to start. Alternatively, some package
-and methods use approximate calculations of LEs, such as
-\textbf{fastshap} \citet{greenwell_fastshap_2020}.
+The time to preprocess the data will vary significantly with the complexity of the model and the LE.
+For reference, the FIFA data contained 5000 observations of nine explanatory variables that took 2.5 seconds to fit a random forest model of modest hyperparameters.
+Extracting the tree SHAP values of each observation took 270 seconds in total.
+PCA and statistics of the variables and attributions took 2.8 seconds.
+These run times were from a non-parallelized session on a modern laptop, but suffice it to say that most of the time will be spent on the LVA.
+An increase in model complexity or data dimensionality will quickly become an obstacle.
+Its reduced computational complexity makes tree SHAP an excellent candidate to start.
+Alternatively, some package and methods use approximate calculations of LEs, such as \textbf{fastshap} Greenwell (2020).
 
@@
 
-To illustrate the cheem method it is applied to modern data sets, two
-classification examples and then two of regression.
+To illustrate the cheem method it is applied to modern data sets, two classification examples and then two of regression.
 
 \hypertarget{palmer-penguin-species-classification}{%
-\subsection{Palmer Penguin, Species
-Classification}\label{palmer-penguin-species-classification}}
-
-The Palmer penguins data
-\citep{gorman_ecological_2014, horst_palmerpenguins_2020} was collected
-on three species of penguins foraging near Palmer Station, Antarctica.
-The data is publicly available to substitute for the overly-used iris
-data and is quite similar in form. After removing incomplete
-observations, there are 333 observations of four physical measurements,
-bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length
-(\texttt{fl}), and body mass (\texttt{bm}) for this illustration. A
-random forest model was fit with species as the response variable.
-
-\begin{CodeChunk}
+\subsection{Palmer Penguin, Species Classification}\label{palmer-penguin-species-classification}}
+
+The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst, Hill, and Gorman 2020) was collected on three species of penguins foraging near Palmer Station, Antarctica.
+The data is publicly available to substitute for the overly-used iris data and is quite similar in form.
+After removing incomplete observations, there are 333 observations of four physical measurements, bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length (\texttt{fl}), and body mass (\texttt{bm}) for this illustration.
+A random forest model was fit with species as the response variable.
+
 \begin{figure}
@@
 
-\caption[Examining the SHAP values for a random forest model classifying Palmer penguin species]{Examining the SHAP values for a random forest model classifying Palmer penguin species.
+\caption{Examining the SHAP values for a random forest model classifying Palmer penguin species.
 The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b).
@@
 \end{figure}
-\end{CodeChunk}
-
-Figure \ref{fig:casepenguins} shows plots from the cheem viewer for
-exploring the random forest model on the penguins data. Panel (a) shows
-the global view, and panel (b) shows several 1D projections generated
-with the radial tour. Penguin 243, a Gentoo (purple), is the PI because
-it has been misclassified as a Chinstrap (orange).
-
-\begin{CodeChunk}
+
+Figure \ref{fig:casepenguins} shows plots from the cheem viewer for exploring the random forest model on the penguins data.
+Panel (a) shows the global view, and panel (b) shows several 1D projections generated with the radial tour.
+Penguin 243, a Gentoo (purple), is the PI because it has been misclassified as a Chinstrap (orange).
+
 \begin{figure}
@@
 
-\caption[Checking what is learned from the cheem viewer]{Checking what is learned from the cheem viewer.
+\caption{Checking what is learned from the cheem viewer.
 This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI.
@@
 \end{figure}
-\end{CodeChunk}
-
-There is more separation visible in the attribution space than in the
-data space, as would be expected. The predicted vs observed plot reveals
-a handful of misclassified observations. A Gentoo which has been wrongly
-labeled as a Chinstrap is selected for illustration. The PI is a
-misclassified point (represented by the asterisk in the global view and
-a dashed vertical line in the tour view). The CI is a correctly
-classified point (represented by an \(\times\) and a vertical dotted
-line).
-
-The radial tour starts from the attribution projection of the
-misclassified observation (b, left). The important variables identified
-by SHAP in the (wrong) prediction for this observation are mostly
-\texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and
-\texttt{bm}. This projection is a view where the Gentoo (purple) looks
-much more likely for this observation than Chinstrap. That is, this
-combination of variables is not particularly useful because the PI looks
-very much like other Gentoo penguins. The radial tour is used to vary
-the contribution of flipper length (\texttt{fl}) to explore this. (In
-our exploration, this was the third variable explored. It is typically
-helpful to explore the variables with more significant contributions,
-here \texttt{bl} and \texttt{bd}. Still, when doing this, nothing was
-revealed about how the PI differed from other Gentoos). On varying
-\texttt{fl}, as it contributes increasingly to the projection (b,
-right), more and more, this penguin looks like a Chinstrap. This
-suggests that \texttt{fl} should be considered an important variable for
-explaining the (wrong) prediction.
-
-Figure \ref{fig:casepenguinsblfl} confirms that flipper length
-(\texttt{fl}) is vital for the confusion of the PI as a Chinstrap. Here,
-flipper length and body length are plotted, and the PI can be seen to be
-closer to the Chinstrap group in these two variables, mainly because it
-has an unusually low value of flipper length relative to other Gentoos.
-From this view, it makes sense that it is a hard observation to account
-for, as decision trees can only partition only vertical and horizontal
-lines.
+
+There is more separation visible in the attribution space than in the data space, as would be expected.
+The predicted vs observed plot reveals a handful of misclassified observations.
+A Gentoo which has been wrongly labeled as a Chinstrap is selected for illustration.
+The PI is a misclassified point (represented by the asterisk in the global view and a dashed vertical line in the tour view).
+The CI is a correctly classified point (represented by an \(\times\) and a vertical dotted line).
+
+The radial tour starts from the attribution projection of the misclassified observation (b, left).
+The important variables identified by SHAP in the (wrong) prediction for this observation are mostly \texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
+This projection is a view where the Gentoo (purple) looks much more likely for this observation than Chinstrap.
+That is, this combination of variables is not particularly useful because the PI looks very much like other Gentoo penguins.
+The radial tour is used to vary the contribution of flipper length (\texttt{fl}) to explore this.
+(In our exploration, this was the third variable explored.
+It is typically helpful to explore the variables with more significant contributions, here \texttt{bl} and \texttt{bd}.
+Still, when doing this, nothing was revealed about how the PI differed from other Gentoos).
+On varying \texttt{fl}, as it contributes increasingly to the projection (b, right), more and more, this penguin looks like a Chinstrap.
+This suggests that \texttt{fl} should be considered an important variable for explaining the (wrong) prediction.
+
+Figure \ref{fig:casepenguinsblfl} confirms that flipper length (\texttt{fl}) is vital for the confusion of the PI as a Chinstrap.
+Here, flipper length and body length are plotted, and the PI can be seen to be closer to the Chinstrap group in these two variables, mainly because it has an unusually low value of flipper length relative to other Gentoos.
+From this view, it makes sense that it is a hard observation to account for, as decision trees can only partition only vertical and horizontal lines.
 
 \hypertarget{chocolates-milkdark-classification}{%
-\subsection{Chocolates, Milk/Dark
-Classification}\label{chocolates-milkdark-classification}}
-
-The chocolates data set consists of 88 observations of ten nutritional
-measurements determined from their labels and labeled as either milk or
-dark. Dark chocolate is considered healthier than milk. Students
-collected the data during the Iowa State University class STAT503 from
-nutritional information on the manufacturer's websites and were
-normalized to 100g equivalents. The data is available in the
-\textbf{cheem} package. A random forest model is used for the
-classification of chocolate types.
-
-It could be interesting to examine the nutritional properties of any
-dark chocolates that have been misclassified as milk. A reason to do
-this is that a dark chocolate, nutritionally more like milk should not
-be considered a healthy alternative. It is interesting to explore which
-nutritional variables contribute most to misclassification.
-
-\begin{CodeChunk}
+\subsection{Chocolates, Milk/Dark Classification}\label{chocolates-milkdark-classification}}
+
+The chocolates data set consists of 88 observations of ten nutritional measurements determined from their labels and labeled as either milk or dark.
+Dark chocolate is considered healthier than milk.
+Students collected the data during the Iowa State University class STAT503 from nutritional information on the manufacturer's websites and were normalized to 100g equivalents.
+The data is available in the \textbf{cheem} package.
+A random forest model is used for the classification of chocolate types.
+
+It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk.
+A reason to do this is that a dark chocolate, nutritionally more like milk should not be considered a healthy alternative.
+It is interesting to explore which nutritional variables contribute most to misclassification.
+
 \begin{figure}
@@
 
-\caption[Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
+\caption{Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green).
 From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the LVA does not help understand the prediction for this observation.
@@
 \end{figure}
-\end{CodeChunk}
-
-This type of exploration is shown in Figure \ref{fig:casechocolates},
-where a chocolate labeled dark but predicted to be milk is chosen as the
-PI (observation 22). It is compared with a CI that is a correctly
-classified dark chocolate (observation 7). The PCA plot and the tree
-SHAP PCA plots (a) show a big difference between the two chocolate types
-but with confusion for a handful of observations. The misclassifications
-are more apparent in the observed vs predicted plot and can be seen to
-be mistaken in both ways: milk to dark and dark to milk.
-
-The attribution projection for chocolate 22 suggests that Fiber, Sugars,
-and Calories are most responsible for its incorrect prediction. The way
-to read this plot is to see that Fiber has a large negative value while
-Sugars and Calories have reasonably large positive values. In the
-density plot, observations on the very left of the display would have
-high values of Fiber (matching the negative projection coefficient) and
-low values of Sugars and Calories. The opposite would be interpreting a
-point with high values in this plot. The dark chocolates (orange) are
-primarily on the left, and this is a reason why they are considered to
-be healthier: high fiber and low sugar. The density of milk chocolates
-is further to the right, indicating that they generally have low fiber
-and high sugar.
-
-The PI (dashed line) can be viewed against the CI (dotted line). Now,
-one needs to pay attention to the parallel plot of the SHAP values,
-which are local to a particular observation, and the density plot, which
-is the same projection of all observations as specified by the SHAP
-values of the PI. The variable contribution of the two different
-predictions can be quickly compared in the parallel coordinate plot. The
-PI differs from the comparison primarily on the Fiber variable, which
-suggests that this is the reason for the incorrect prediction.
-
-From the density plot, which is the attribution projection corresponding
-to the PI, both observations are more like dark chocolates. Varying the
-contribution of Sugars and altogether removing it from the projection is
-where the difference becomes apparent. When a frame with contribution
-primarily from Fiber is examined observation 22 looks more like a milk
-chocolate.
-
-It would also be interesting to explore an inverse misclassification. In
-this case, a milk chocolate is selected while it was misclassified as a
-dark chocolate. Chocolate 84 is just this case and is compared with a
-correctly predicted milk chocolate (observation 71). The corresponding
-global view and radial tour frames are shown in Figure
-\ref{fig:casechocolatesinverse}.
-
-\begin{CodeChunk}
+
+This type of exploration is shown in Figure \ref{fig:casechocolates}, where a chocolate labeled dark but predicted to be milk is chosen as the PI (observation 22).
+It is compared with a CI that is a correctly classified dark chocolate (observation 7).
+The PCA plot and the tree SHAP PCA plots (a) show a big difference between the two chocolate types but with confusion for a handful of observations.
+The misclassifications are more apparent in the observed vs predicted plot and can be seen to be mistaken in both ways: milk to dark and dark to milk.
+
+The attribution projection for chocolate 22 suggests that Fiber, Sugars, and Calories are most responsible for its incorrect prediction.
+The way to read this plot is to see that Fiber has a large negative value while Sugars and Calories have reasonably large positive values.
+In the density plot, observations on the very left of the display would have high values of Fiber (matching the negative projection coefficient) and low values of Sugars and Calories.
+The opposite would be interpreting a point with high values in this plot.
+The dark chocolates (orange) are primarily on the left, and this is a reason why they are considered to be healthier: high fiber and low sugar.
+The density of milk chocolates is further to the right, indicating that they generally have low fiber and high sugar.
+
+The PI (dashed line) can be viewed against the CI (dotted line).
+Now, one needs to pay attention to the parallel plot of the SHAP values, which are local to a particular observation, and the density plot, which is the same projection of all observations as specified by the SHAP values of the PI.
+The variable contribution of the two different predictions can be quickly compared in the parallel coordinate plot.
+The PI differs from the comparison primarily on the Fiber variable, which suggests that this is the reason for the incorrect prediction.
+
+From the density plot, which is the attribution projection corresponding to the PI, both observations are more like dark chocolates.
+Varying the contribution of Sugars and altogether removing it from the projection is where the difference becomes apparent.
+When a frame with contribution primarily from Fiber is examined observation 22 looks more like a milk chocolate.
+
+It would also be interesting to explore an inverse misclassification.
+In this case, a milk chocolate is selected while it was misclassified as a dark chocolate.
+Chocolate 84 is just this case and is compared with a correctly predicted milk chocolate (observation 71).
+The corresponding global view and radial tour frames are shown in Figure \ref{fig:casechocolatesinverse}.
+
 \begin{figure}
@@
 
-\caption[Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
+\caption{Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange).
 In the attribution projection, the PI could be either milk or dark.
@@
 \end{figure}
-\end{CodeChunk}
-
-The difference of position in the tree SHAP PCA with the previous case
-is quite significant; this gives a higher-level sense that the
-attributions should be quite different. Looking at the attribution
-projection, this is found to be the case. Previously, Fiber was
-essential while it is absent from the attribution in this case.
-Conversely, Calories from Fat and Total Fat have high attributions here,
-while they were unimportant in the preceding case.
-
-Comparing the attribution with the CI (dotted line), large discrepancies
-in Sodium and Fiber are identified. The contribution of Sodium is
-selected to be varied. Even in the initial projection, the observation
-looks slightly more like its observed milk than predicted dark
-chocolate. The misclassification appears least supported when the basis
-reaches sodium attribution of typical dark chocolate.
+
+The difference of position in the tree SHAP PCA with the previous case is quite significant; this gives a higher-level sense that the attributions should be quite different.
+Looking at the attribution projection, this is found to be the case.
+Previously, Fiber was essential while it is absent from the attribution in this case.
+Conversely, Calories from Fat and Total Fat have high attributions here, while they were unimportant in the preceding case.
+
+Comparing the attribution with the CI (dotted line), large discrepancies in Sodium and Fiber are identified.
+The contribution of Sodium is selected to be varied.
+Even in the initial projection, the observation looks slightly more like its observed milk than predicted dark chocolate.
+The misclassification appears least supported when the basis reaches sodium attribution of typical dark chocolate.
 
@@
 
-The 2020 season FIFA data \citep{leone_fifa_2020, biecek_dalex_2018}
-contains many skill measurements of soccer/football players and wage
-information. Nine higher-level skill groupings were identified and
-aggregated from highly correlated variables. A random forest model is
-fit from these predictors, regressing player wages {[}2020 euros{]}. The
-model was fit from 5000 observations before being thinned to 500 players
-to mitigate occlusion and render time. Continuing from the exploration
-in Section \textbackslash ref\{sec:explanations), we are interested to
-see the difference in attribution based on the exogenous player
-position. That is, the model should be able to use multiple linear
-profiles to better predict the wages from different field positions of
-players despite not having this information. A leading offensive fielder
-(L. Messi) is compared with a top defensive fielder (V. van Dijk). The
-same observations were used in Figure \ref{fig:shapdistrbd}.
-
-\begin{CodeChunk}
+The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill measurements of soccer/football players and wage information.
+Nine higher-level skill groupings were identified and aggregated from highly correlated variables.
+A random forest model is fit from these predictors, regressing player wages {[}2020 euros{]}.
+The model was fit from 5000 observations before being thinned to 500 players to mitigate occlusion and render time.
+Continuing from the exploration in Section \textbackslash ref\{sec:explanations), we are interested to see the difference in attribution based on the exogenous player position.
+That is, the model should be able to use multiple linear profiles to better predict the wages from different field positions of players despite not having this information.
+A leading offensive fielder (L.
+Messi) is compared with a top defensive fielder (V.
+van Dijk).
+The same observations were used in Figure \ref{fig:shapdistrbd}.
+
 \begin{figure}
@@
 
-\caption[Exploring the wages relative to skill measurements in the FIFA 2020 data]{Exploring the wages relative to skill measurements in the FIFA 2020 data.
+\caption{Exploring the wages relative to skill measurements in the FIFA 2020 data.
 Star offensive player (L.
@@
 \end{figure}
-\end{CodeChunk}
-
-Figure \ref{fig:casefifa} tests the support of the LVA. Offensive and
-reaction skills (\texttt{off} and \texttt{rct}) are both crucial to
-explaining a star offensive player. If either of them were rotated out,
-the other would be rotated into the frame, maintaining a far-right
-position. However, increasing the contribution of a variable with low
-importance would rotate both variables out of the frame.
-
-The contribution from \texttt{def} will be varied to contrast with
-offensive skills. As the contribution of defensive skills increases,
-Messi's is no longer separated from the group. Players with high values
-in defensive skills are now the rightmost points. In terms of what-if
-analysis, the difference between the data mean and his predicted wages
-would be halved if Messi's tree SHAP attributions were at these levels.
-
-\begin{CodeChunk}
+
+Figure \ref{fig:casefifa} tests the support of the LVA.
+Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both crucial to explaining a star offensive player.
+If either of them were rotated out, the other would be rotated into the frame, maintaining a far-right position.
+However, increasing the contribution of a variable with low importance would rotate both variables out of the frame.
+
+The contribution from \texttt{def} will be varied to contrast with offensive skills.
+As the contribution of defensive skills increases, Messi's is no longer separated from the group.
+Players with high values in defensive skills are now the rightmost points.
+In terms of what-if analysis, the difference between the data mean and his predicted wages would be halved if Messi's tree SHAP attributions were at these levels.
+
 \begin{figure}
@@
 
-\caption[Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data]{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
+\caption{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data.
 The LVA indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable.
@@
 \end{figure}
-\end{CodeChunk}
 
 \hypertarget{ames-housing-2018-sales-price-regression}{%
-\subsection{Ames Housing 2018, Sales Price
-Regression}\label{ames-housing-2018-sales-price-regression}}
-
-Ames housing data 2018 \citep{de_cock_ames_2011, prevek18_ames_2018} was
-subset to North Ames (the neighborhood with the most house sales). The
-remaining are 338 house sales. A random forest model was fit, predicting
-the sale price {[}USD{]} from the property variables: Lot Area
-(\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built
-(\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms
-(\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of
-Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and
-Garage Area (\texttt{GrA}). Using interactions with the global view, a
-house with an extreme negative residual and an accurate observation with
-a similar prediction is selected.
-
-Figure \ref{fig:caseames} selects the house sale 74, a sizable
-under-prediction with an enormous Lot Area contribution. The CI has a
-similar predicted price though the prediction was accurate and gives
-almost no attribution to lot size. The attribution projection places
-observations with high Living Areas to the right. The contribution of
-Living Area contrasts the contribution of this variable. As the
-contribution of Lot Area decreases, the predictive power decreases for
-the PI, while the CI remains stationary. This large importance in the
-Living Area is relatively uncommon. Boosting tree models may be more
-resilient to such an under-prediction as they would up-weighting this
-residual and force its inclusion in the final model.
+\subsection{Ames Housing 2018, Sales Price Regression}\label{ames-housing-2018-sales-price-regression}}
+
+Ames housing data 2018 (De Cock 2011; prevek18 2018) was subset to North Ames (the neighborhood with the most house sales).
+The remaining are 338 house sales.
+A random forest model was fit, predicting the sale price {[}USD{]} from the property variables: Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and Garage Area (\texttt{GrA}).
+Using interactions with the global view, a house with an extreme negative residual and an accurate observation with a similar prediction is selected.
+
+Figure \ref{fig:caseames} selects the house sale 74, a sizable under-prediction with an enormous Lot Area contribution.
+The CI has a similar predicted price though the prediction was accurate and gives almost no attribution to lot size.
+The attribution projection places observations with high Living Areas to the right.
+The contribution of Living Area contrasts the contribution of this variable.
+As the contribution of Lot Area decreases, the predictive power decreases for the PI, while the CI remains stationary.
+This large importance in the Living Area is relatively uncommon.
+Boosting tree models may be more resilient to such an under-prediction as they would up-weighting this residual and force its inclusion in the final model.
 
@@
 
-There is a clear need to extend the interpretability of black box
-models. With techniques such as SHAP, LIME, Break-down, one can
-calculate LEs, i.e.~for every observation in the data. These techniques
-quantify for each observation how strongly particular variables affect
-the model's predictions. Surprisingly few techniques allow us to
-understand the global distribution of these LEs. Unsupervised data
-exploration techniques applied to data show how useful they are for
-identifying outliers, identifying clusters of observations or
-discovering correlations between variables. All of these tasks can be
-performed for a set of explanations.
-
-To address this challenge this paper provides a technique that builds on
-LEs to explore the variable importance local to an observation. The LVA
-is converted into an attribution projection from which variable
-contributions are varied using a radial tour. Several diagnostic plots
-are provided to assist with understanding the sensitivity of the
-prediction to particular variables. A global view shows the data space,
-explanation space, and residual plot. The user can interactively select
-observations to compare, contrast, and study further. Then the radial
-tour is used to explore the variable sensitivity identified by the
-attribution projection.
-
-This approach has been illustrated using four data examples of random
-forest models with the tree SHAP LVA. LEs focus on the model fit and
-help to dissect which variables are most responsible for the fitted
-value. They can also form the basis of learning how the model has got it
-wrong, when the observation is misclassified or has a large residual.
-
-In the penguins example, we showed how the misclassification of a
-penguin arose due to it having an unusually small flipper size compared
-to others of its species. This was verified by making a follow-up plot
-of the data. The chocolates example shows how a dark chocolate was
-misclassified primarily due to its attribution to Fiber, and a milk
-chocolate was misclassified as dark due to its lowish Sodium value. In
-the FIFA example, we show how low Messi's salary would be if it depended
-on their defensive skill. In the Ames housing data, an inaccurate
-prediction for a house was likely due to the lot area not being
-effectively used by the random forest model.
-
-This analysis is manually intensive and thus only feasible for
-investigating a few observations. The recommended approach is to
-investigate an observation where the model has not predicted accurately
-and compare it with an observation with similar predictor values where
-the model fitted well. The radial tour launches from the attribution
-projection to enable exploration of the sensitivity of the prediction to
-any variable. It can be helpful to make additional plots of the
-variables and responses to cross-check interpretations made from the
-cheem viewer. This methodology provides an additional tool in the box
-for studying model fitting.
+There is a clear need to extend the interpretability of black box models.
+With techniques such as SHAP, LIME, Break-down, one can calculate LEs, i.e.~for every observation in the data.
+These techniques quantify for each observation how strongly particular variables affect the model's predictions.
+Surprisingly few techniques allow us to understand the global distribution of these LEs.
+Unsupervised data exploration techniques applied to data show how useful they are for identifying outliers, identifying clusters of observations or discovering correlations between variables.
+All of these tasks can be performed for a set of explanations.
+
+To address this challenge this paper provides a technique that builds on LEs to explore the variable importance local to an observation.
+The LVA is converted into an attribution projection from which variable contributions are varied using a radial tour.
+Several diagnostic plots are provided to assist with understanding the sensitivity of the prediction to particular variables.
+A global view shows the data space, explanation space, and residual plot.
+The user can interactively select observations to compare, contrast, and study further.
+Then the radial tour is used to explore the variable sensitivity identified by the attribution projection.
+
+This approach has been illustrated using four data examples of random forest models with the tree SHAP LVA.
+LEs focus on the model fit and help to dissect which variables are most responsible for the fitted value.
+They can also form the basis of learning how the model has got it wrong, when the observation is misclassified or has a large residual.
+
+In the penguins example, we showed how the misclassification of a penguin arose due to it having an unusually small flipper size compared to others of its species.
+This was verified by making a follow-up plot of the data.
+The chocolates example shows how a dark chocolate was misclassified primarily due to its attribution to Fiber, and a milk chocolate was misclassified as dark due to its lowish Sodium value.
+In the FIFA example, we show how low Messi's salary would be if it depended on their defensive skill.
+In the Ames housing data, an inaccurate prediction for a house was likely due to the lot area not being effectively used by the random forest model.
+
+This analysis is manually intensive and thus only feasible for investigating a few observations.
+The recommended approach is to investigate an observation where the model has not predicted accurately and compare it with an observation with similar predictor values where the model fitted well.
+The radial tour launches from the attribution projection to enable exploration of the sensitivity of the prediction to any variable.
+It can be helpful to make additional plots of the variables and responses to cross-check interpretations made from the cheem viewer.
+This methodology provides an additional tool in the box for studying model fitting.
 
@@
 
-An implementation is provided in the open-source \textbf{R} package
-\textbf{cheem}, available on CRAN at
-\url{https://CRAN.R-project.org/package=cheem}. Example data sets are
-provided, and you can upload your data after model fitting and computing
-the LVAs. The LVAs need to be pre-computed and uploaded. Examples show
-how to do this for tree SHAP values, using \textbf{treeshap} (tree-based
-models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest},
-\textbf{ranger}, or \textbf{xgboost} \citet{greenwell_gbm_2020};
-\citet{shi_lightgbm_2022}; \citet{liaw_classification_2002};
-\citet{wright_ranger_2017}; \citet{chen_xgboost_2021}, respectively).
-The SHAP and oscillation explanations could be easily added using
-\texttt{DALEX::explain()}
-\citep{biecek_dalex_2018, biecek_explanatory_2021}.
-
-The application was made with \textbf{shiny} \citep{chang_shiny_2021}.
-The tour visual is built with \textbf{spinifex}
-\citep{spyrison_spinifex_2020}. Both views are created first with
-\textbf{ggplot2} \citep{wickham_ggplot2_2016} and then rendered as
-interactive \texttt{html} widgets with \textbf{plotly}
-\citep{sievert_interactive_2020}. \textbf{DALEX}
-\citep{biecek_dalex_2018} and \emph{Explanatory Model Analysis}
-\citep{biecek_explanatory_2021} are helpful for understanding LEs and
-how to apply them.
-
-The package can be installed from CRAN, and the application can be run
-using the following \textbf{R} code:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> install.packages("cheem", dependencies = TRUE)
-R> library("cheem")
-R> run_app()
-\end{CodeInput}
-\end{CodeChunk}
+An implementation is provided in the open-source \textbf{R} package \textbf{cheem}, available on CRAN at \url{https://CRAN.R-project.org/package=cheem}.
+Example data sets are provided, and you can upload your data after model fitting and computing the LVAs.
+The LVAs need to be pre-computed and uploaded.
+Examples show how to do this for tree SHAP values, using \textbf{treeshap} (tree-based models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or \textbf{xgboost} Greenwell et al.
+(2020); Shi et al.
+(2022); Liaw and Wiener (2002); Wright and Ziegler (2017); Chen et al.
+(2021), respectively).
+The SHAP and oscillation explanations could be easily added using \texttt{DALEX::explain()} (Biecek 2018; Biecek and Burzykowski 2021).
+
+The application was made with \textbf{shiny} (Chang et al.
+2021).
+The tour visual is built with \textbf{spinifex} (Spyrison and Cook 2020).
+Both views are created first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive \texttt{html} widgets with \textbf{plotly} (Sievert 2020).
+\textbf{DALEX} (Biecek 2018) and \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) are helpful for understanding LEs and how to apply them.
+
+The package can be installed from CRAN, and the application can be run using the following \textbf{R} code:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
+\FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
+\FunctionTok{run\_app}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \item
-  The development version of the package is available at
-  \url{https://github.com/nspyrison/cheem}, and
+  The development version of the package is available at \url{https://github.com/nspyrison/cheem}, and
 \item
-  Documentation of the package can be found at
-  \url{https://nspyrison.github.io/cheem/}.
+  Documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
 \end{itemize}
 
-Follow the examples provided with the package to compute the LVAs (using
-\texttt{?cheem\_ls}). The application expects the output returned by
-\texttt{cheem\_ls()}, saved to an \texttt{rds} file with
-\texttt{saveRDS()} to be uploaded.
+Follow the examples provided with the package to compute the LVAs (using \texttt{?cheem\_ls}).
+The application expects the output returned by \texttt{cheem\_ls()}, saved to an \texttt{rds} file with \texttt{saveRDS()} to be uploaded.
 
@@
 
-Kim Marriott provided advice on many aspects of this work, especially on
-the explanations in the applications section. This research was
-supported by the Australian Government Research Training Program (RTP)
-scholarships. Thanks to Jieyang Chong for helping proofread this
-article. The namesake, Cheem, refers to a fictional race of humanoid
-trees from Doctor Who lore. \textbf{DALEX} pulls on from that universe,
-and we initially apply tree SHAP explanations specific to tree-based
-models.
-
-\renewcommand\refname{References}
-\bibliography{paper.bib}
+Kim Marriott provided advice on many aspects of this work, especially on the explanations in the applications section.
+This research was supported by the Australian Government Research Training Program (RTP) scholarships.
+Thanks to Jieyang Chong for helping proofread this article.
+The namesake, Cheem, refers to a fictional race of humanoid trees from Doctor Who lore.
+\textbf{DALEX} pulls on from that universe, and we initially apply tree SHAP explanations specific to tree-based models.
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-adadi_peeking_2018}{}}%
+Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''} \emph{IEEE Access} 6: 52138--60.
+
+\leavevmode\vadjust pre{\hypertarget{ref-anderson_introduction_1995}{}}%
+Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-asimov_grand_1985}{}}%
+Asimov, Daniel.
+1985.
+{``The {Grand} {Tour}: A {Tool} for {Viewing} {Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
+\url{https://doi.org/10.1137/0906011}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-arrieta_explainable_2020}{}}%
+Barredo Arrieta, Alejandro, Natalia Diaz-Rodriguez, Javier Del Ser, Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador Garcia, et al.
+2020.
+{``Explainable Artificial Intelligence (XAI): Concepts, Taxonomies, Opportunities and Challenges Toward Responsible AI.''} \emph{Information Fusion} 58: 82--115.
+https://doi.org/\url{https://doi.org/10.1016/j.inffus.2019.12.012}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_dalex_2018}{}}%
+Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive Models in {R}.''} \emph{The Journal of Machine Learning Research} 19 (1): 3245--49.
+
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_ceterisparibus_2020}{}}%
+---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-biecek_explanatory_2021}{}}%
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-boser_training_1992}{}}%
+Boser, Bernhard E., Isabelle M.
+Guyon, and Vladimir N.
+Vapnik.
+1992.
+{``A Training Algorithm for Optimal Margin Classifiers.''} In \emph{Proceedings of the Fifth Annual Workshop on {Computational} Learning Theory}, 144--52.
+
+\leavevmode\vadjust pre{\hypertarget{ref-breiman_random_2001}{}}%
+Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45 (1): 5--32.
+
+\leavevmode\vadjust pre{\hypertarget{ref-breiman_statistical_2001}{}}%
+---------. 2001b. {``Statistical Modeling: {The} Two Cultures (with Comments and a Rejoinder by the Author).''} \emph{Statistical Science} 16 (3): 199--231.
+
+\leavevmode\vadjust pre{\hypertarget{ref-buja_grand_1986}{}}%
+Buja, Andreas, and Daniel Asimov.
+1986.
+{``Grand {Tour} {Methods}: {An} {Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer} {Science} and {Statistics}}, 63--67.
+New York, NY, USA: Elsevier North-Holland, Inc.
+\url{http://dl.acm.org/citation.cfm?id=26036.26046}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Caragea2008}{}}%
+Caragea, Doina, Dianne Cook, Hadley Wickham, and Vasant Honavar.
+2008.
+{``Visual Methods for Examining SVM Classifiers.''} In \emph{Visual Data Mining: Theory, Techniques and Tools for Visual Analytics}, edited by Simeon J.
+Simoff, Michael H.
+Böhlen, and Arturas Mazeika, 136--53.
+Berlin, Heidelberg: Springer Berlin Heidelberg.
+\url{https://doi.org/10.1007/978-3-540-71080-6_10}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chang_shiny_2021}{}}%
+Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
+2021.
+\emph{Shiny: Web Application Framework for r}.
+\url{https://CRAN.R-project.org/package=shiny}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chen_xgboost_2021}{}}%
+Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al.
+2021.
+{``Xgboost: {Extreme} {Gradient} {Boosting}.''} \url{https://CRAN.R-project.org/package=xgboost}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
+Cook, Dianne, and Andreas Buja.
+1997.
+{``Manual {Controls} for {High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
+\url{https://doi.org/10.2307/1390747}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cook_grand_2008}{}}%
+Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham.
+2008.
+{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual} {Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
+Berlin, Heidelberg: Springer Berlin Heidelberg.
+\url{https://doi.org/10.1007/978-3-540-33037-0_13}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cook_interactive_2007}{}}%
+Cook, Dianne, Deborah F. Swayne, and A. Buja. 2007. \emph{Interactive and {Dynamic} {Graphics} for {Data} {Analysis}: {With} {R} and {GGobi}}. Springer Science \& Business Media.
+
+\leavevmode\vadjust pre{\hypertarget{ref-da_silva_projection_2021}{}}%
+da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee.
+2021.
+{``A {Projection} {Pursuit} {Forest} {Algorithm} for {Supervised} {Classification}.''} \emph{Journal of Computational and Graphical Statistics}, 1--21.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de_cock_ames_2011}{}}%
+De Cock, Dean. 2011. {``Ames, {Iowa}: {Alternative} to the {Boston} Housing Data as an End of Semester Regression Project.''} \emph{Journal of Statistics Education} 19 (3).
+
+\leavevmode\vadjust pre{\hypertarget{ref-gorman_ecological_2014}{}}%
+Gorman, Kristen B., Tony D.
+Williams, and William R.
+Fraser.
+2014.
+{``Ecological Sexual Dimorphism and Environmental Variability Within a Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS One} 9 (3): e90081.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gosiewska_ibreakdown_2019}{}}%
+Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}: {Uncertainty} of Model Explanations for Non-Additive Predictive Models.''} \emph{arXiv Preprint arXiv:1903.11420}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-greenwell_fastshap_2020}{}}%
+Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley} {Values}}. \url{https://CRAN.R-project.org/package=fastshap}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-fastshap}{}}%
+---------. 2021. \emph{Fastshap: Fast Approximate Shapley Values}. \url{https://CRAN.R-project.org/package=fastshap}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-greenwell_gbm_2020}{}}%
+Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M. Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression} {Models}.''} \url{https://CRAN.R-project.org/package=gbm}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-horst_palmerpenguins_2020}{}}%
+Horst, Allison Marie, Alison Presmanes Hill, and Kristen B.
+Gorman.
+2020.
+{``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-karwowski_international_2006}{}}%
+Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of {Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kominsarczyk_treeshap_2021}{}}%
+Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw Biecek. 2021. {``Treeshap.''} Model Oriented. \url{https://github.com/ModelOriented/treeshap}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-PPtreeregViz}{}}%
+Lee, Eun-Kyung, and HyunSun Cho. 2022. \emph{PPtreeregViz: Projection Pursuit Regression Tree Visualization}. \url{https://CRAN.R-project.org/package=PPtreeregViz}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lee_state_2021}{}}%
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas Spyrison, Earo Wang, and H.
+Sherry Zhang.
+2021.
+{``The State-of-the-Art on Tours for Dynamic Visualization of High-Dimensional Data.''} \emph{WIREs Computational Statistics} n/a (n/a): e1573.
+\url{https://doi.org/10.1002/wics.1573}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lee_pptree_2013}{}}%
+Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013. {``{PPtree}: {Projection} Pursuit Classification Tree.''} \emph{Electronic Journal of Statistics} 7: 1369--86.
+
+\leavevmode\vadjust pre{\hypertarget{ref-leone_fifa_2020}{}}%
+Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''} \url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-liaw_classification_2002}{}}%
+Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression by {randomForest}.''} \emph{R News} 2 (3): 18--22.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lundberg_consistent_2018}{}}%
+Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv Preprint arXiv:1802.03888}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lundberg_unified_2017}{}}%
+Lundberg, Scott M., and Su-In Lee.
+2017.
+{``A Unified Approach to Interpreting Model Predictions.''} In \emph{Proceedings of the 31st International Conference on Neural Information Processing Systems}, 4768--77.
+
+\leavevmode\vadjust pre{\hypertarget{ref-flashlight}{}}%
+Mayer, Michael. 2023a. \emph{Flashlight: Shed Light on Black Box Machine Learning Models}. \url{https://CRAN.R-project.org/package=flashlight}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shapviz}{}}%
+---------. 2023b. \emph{Shapviz: SHAP Visualizations}. \url{https://CRAN.R-project.org/package=shapviz}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kernelshap}{}}%
+Mayer, Michael, and David Watson. 2023. \emph{Kernelshap: Kernel SHAP}. \url{https://CRAN.R-project.org/package=kernelshap}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-molnar2022}{}}%
+Molnar, Christoph. 2022. \emph{Interpretable Machine Learning: A Guide for Making Black Box Models Explainable}. 2nd ed. \url{https://christophm.github.io/interpretable-ml-book}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ocagne_coordonnees_1885}{}}%
+Ocagne, Maurice d'.
+1885.
+\emph{Coordonnées Parallèles Et Axiales.
+{Méthode} de Transformation Géométrique Et Procédé Nouveau de Calcul Graphique Déduits de La Considération Des Coordonnées Parallèles, Par {Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.
+
+\leavevmode\vadjust pre{\hypertarget{ref-prevek18_ames_2018}{}}%
+prevek18. 2018. {``Ames {Housing} {Dataset}.''} \emph{Kaggle.com}. \url{https://kaggle.com/prevek18/ames-housing-dataset}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ribeiro_why_2016}{}}%
+Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
+2016.
+{``"{Why} {Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any} {Classifier}.''} In \emph{Proceedings of the 22nd {ACM} {SIGKDD} {International} {Conference} on {Knowledge} {Discovery} and {Data} {Mining}}, 1135--44.
+{KDD} '16.
+New York, NY, USA: Association for Computing Machinery.
+\url{https://doi.org/10.1145/2939672.2939778}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-roberts_state_2007}{}}%
+Roberts, Jonathan C.
+2007.
+{``State of the Art: {Coordinated} \& Multiple Views in Exploratory Visualization.''} In \emph{Fifth International Conference on Coordinated and Multiple Views in Exploratory Visualization ({CMV} 2007)}, 61--71.
+IEEE.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ExplainPrediction}{}}%
+Robnik-Sikonja, Marko. 2018. \emph{ExplainPrediction: Explanation of Predictions for Classification and Regression Models}. \url{https://CRAN.R-project.org/package=ExplainPrediction}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shapr}{}}%
+Sellereite, Nikolai, Martin Jullum, and Annabelle Redelmeier. 2023. \emph{Shapr: Prediction Explanation with Dependence-Aware Shapley Values}. \url{https://CRAN.R-project.org/package=shapr}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shapley_value_1953}{}}%
+Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shi_lightgbm_2022}{}}%
+Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas Finley, Taifeng Wang, et al.
+2022.
+{``Lightgbm: {Light} {Gradient} {Boosting} {Machine}.''} \url{https://CRAN.R-project.org/package=lightgbm}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shmueli_explain_2010}{}}%
+Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical Science} 25 (3): 289--310.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_learning_2017}{}}%
+Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje.
+2017.
+{``Learning Important Features Through Propagating Activation Differences.''} In \emph{International {Conference} on {Machine} {Learning}}, 3145--53.
+PMLR.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shrikumar_not_2016}{}}%
+Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul Kundaje.
+2016.
+{``Not Just a Black Box: {Learning} Important Features Through Propagating Activation Differences.''} \emph{arXiv Preprint arXiv:1605.01713}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sievert_interactive_2020}{}}%
+Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simonyan_deep_2014}{}}%
+Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman.
+2014.
+{``Deep Inside Convolutional Networks: {Visualising} Image Classification Models and Saliency Maps.''} In \emph{In {Workshop} at {International} {Conference} on {Learning} {Representations}}.
+Citeseer.
+
+\leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
+Spyrison, Nicholas, and Dianne Cook.
+2020.
+{``Spinifex: An {R} {Package} for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of {Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
+\url{https://doi.org/10.32614/RJ-2020-027}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stahl-ethics}{}}%
+Stahl, Bernd Carsten. 2021. {``Ethical Issues of AI.''} \emph{Artificial Intelligence for a Better Future}, 35--53. \url{https://doi.org/10.1007/978-3-030-69978-9_4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-strumbelj_efficient_2010}{}}%
+Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation of Individual Classifications Using Game Theory.''} \emph{The Journal of Machine Learning Research} 11: 1--18.
+
+\leavevmode\vadjust pre{\hypertarget{ref-unwin_ensemble_2018}{}}%
+Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''} \emph{Journal of Computational and Graphical Statistics} 27 (1): 157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanni_textual_2018}{}}%
+Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and Damon Mayaffre.
+2018.
+{``Textual {Deconvolution} {Saliency} ({TDS}): A Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the 56th {Annual} {Meeting} of the {Association} for {Computational} {Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham_ggplot2_2016}{}}%
+Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data} {Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham_visualizing_2015}{}}%
+Wickham, Hadley, Dianne Cook, and Heike Hofmann.
+2015.
+{``Visualizing Statistical Models: {Removing} the Blindfold.''} \emph{Statistical Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
+\url{https://doi.org/10.1002/sam.11271}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wright_ranger_2017}{}}%
+Wright, Marvin N., and Andreas Ziegler.
+2017.
+{``Ranger: {A} {Fast} {Implementation} of {Random} {Forests} for {High} {Dimensional} {Data} in {C}++ and {R}.''} \emph{Journal of Statistical Software} 77 (1): 1--17.
+\url{https://doi.org/10.18637/jss.v077.i01}.
+
+\end{CSLReferences}
 

### 2023-07-10 paper.tex
failed attempt to go to springer nature template for Computational Statistics.

--- 2023-05-21 jss/cheem_paper.tex
@@
 \documentclass[
-]{jss}
+]{sn-jnl}
 
@@
 }
-\title{Exploring Local Explanations of Nonlinear Models Using Animated
-Linear Projections}
+\title{}
 
 \Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}
-\Plaintitle{Exploring Local Explanations of Nonlinear Models Using
-Animated Linear Projections}
-\Shorttitle{Exploring Local Explanations}
 

### 2023-10-25 response_to_review.Rmd
response file in Rmd made earlier by DC added

---
title: "Response to review"
---

**Overview**

Thank you to the editor and reviews for your helpful feedback on our paper. We have made the changes requested, and explain them below, in *italicised* text.

**Reviewer #1**

This paper proposes an interactive visualization tool and its implementation as an R package to intuitively understand LVA in non-linear models.

General Comments

You have not clearly stated the usefulness of the radial tour; please explain the advantage of using the SHAP values of PI and CI as the contribution of 1D projection compared to drawing a static density plot or scatterplot. Is it possible to correctly evaluate the sensitivity of variables by 1d projection for nonlinear models? Also, in the Case Studies in Section 5, please provide an explanation in relation to the Radial tour operation.

*TODO (DC):*

The results are very different between the examples provided in the shiny application in the R cheem package and the examples presented in the paper. Please make sure to use random number seeds to achieve similar results. In particular, in the Case Studies of classification, there is no misclassification in the examples in the cheem package, and the misclassification is not indicated by the red circle as shown in the paper.

*TODO (NS):*

The panel that is treeshap in the global view in the paper is attribution in the cheem package. Please unify either one.

*TODO (NS):*

In the global view in the paper, the model panel has jitter processing, but the cheem package does not.

*TODO (NS):*

Discussion is just a Conclusion. For example, please supplement any discussion on the following.
- Abstract and Introduction are described in the context of XAI, but the tools proposed in this paper are not likely to be able to handle large data sets.
- Concerning the convenience of the package: the application by the cheem package outputs little information (e.g., axis information), which makes it difficult to use intuitively.

*TODO (DC):*

Individual Comments

(p.7 l.18) Figures 3b, 4a -> 3c, 4b?

*TODO (DC):*

(p.8 l.7) Is the SHAP value of the CI taken into account in the radial tour?

*TODO (DC):*

(p.8 l.8) "The PI is ... global view" : The shiny app in the cheem package does not support interactive selection by clicking in the global view. (cheem version 0.3.0 on Windows 11 / Google Chrome)

*TODO (PB):*

(p.10 Fig4) Residual plot is not displayed in cheem version 0.3.0 on Windows 11 / Google Chrome.

*TODO (PB):*

(p.11 l.5) Figure 3a & b -> 4a & b?

*TODO (PB):*

(p.11 l.7) Figure 3d -> 4d?

*TODO (PB):*

(p.21 l.10) https://nspyrison.github.io/cheem/ is not available.

*TODO (NS): And Nick, if you need help we will need to do a zoom call. I think that I mentioned how to fix this some time ago, but not sure you understood.*

(references) There are some characters like "???".

*TODO (PB):*

**Reviewer #2:**

I was able to run your software using publicly available packages.
Unfortunately, I could not play videos shown at the following URLs.

*We cannot address this unless the error messages are provided. The videos at each of the addresses are playable by us and several colleagues who also checked.*

I think the visualization you propose is very clear.
However, I have the impression that it is difficult to understand, so I expect detailed explanations.
For example, detailed explanations of Figure 3 e and Figure 4 e would be helpful.
I also think that explanations of the data in Figure 3 and Figure 4 would make it even easier to understand.

*TODO: expand explanations*

For small details, I think it is necessary to check page 5, line 38 and page 11, line 29.

*TODO: not sure what the problem is here, maybe some typo*

### 2023-10-31 review/cheem_ Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections/figures/manual.r
Uploaded a copy of the reviews in progress from Overleaf.

@@ -1,0 +1,24 @@
+# Figure to explain manual tour
+library(tourr)
+library(palmerpenguins)
+library(dplyr)
+library(ggplot2)
+
+penguins <- penguins %>%
+  na.omit() 
+penguins_sub <- penguins[,c(1, 3:6)] %>% 
+  mutate(across(where(is.numeric),  ~ scale(.)[,1])) %>%
+  rename(bl = bill_length_mm,
+         bd = bill_depth_mm,
+         fl = flipper_length_mm,
+         bm = body_mass_g)
+
+s <- basis_random(4, 1)
+render_gif(penguins_sub[,-1],
+           radial_tour(start=s, mvar=2),
+           display_dist(),
+           "penguins1d.gif",
+           width=300,
+           height=300)
+
+ggplot(penguins_sub, aes(x=bd, y=bm)) + geom_point()

### 2023-11-06 review/cheem_ Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections/sn-article.tex

--- 2023-07-10 paper.tex
@@
-\documentclass[
-]{sn-jnl}
-
-%% recommended packages
-\usepackage{orcidlink,thumbpdf,lmodern}
-
-\usepackage[utf8]{inputenc}
-
-\author{
-Nicholas Spyrison~\orcidlink{0000-0002-8417-0212}\\Monash
-University \And Dianne Cook~\orcidlink{0000-0002-3813-7155}\\Monash
-University \And Przemyslaw
-Biecek~\orcidlink{0000-0001-8423-1823}\\University of Warsaw\\
-Warsaw University of Technology
-}
-\title{}
-
-\Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}
-
-\Abstract{
-The increased predictive power of machine learning models comes at the
-cost of increased complexity and loss of interpretability, particularly
-in comparison to parametric statistical models. This trade-off has led
-to the emergence of eXplainable AI (XAI) which provides methods, such as
-local explanations (LEs) and local variable attributions (LVAs), to shed
-light on how a model use predictors to arrive at a prediction. These
-provide a point estimate of the linear variable importance in the
-vicinity of a single observation. However, LVAs tend not to effectively
-handle association between predictors. To understand how the interaction
-between predictors affects the variable importance estimate, we can
-convert LVAs into linear projections and use the radial tour. This is
-also useful for learning how a model has made a mistake, or the effect
-of outliers, or the clustering of observations. The approach is
-illustrated with examples from categorical (penguin species, chocolate
-types) and quantitative (soccer/football salaries, house prices)
-response models. The methods are implemented in the R package cheem,
-available on CRAN.
-}
-
-\Keywords{explainable artificial intelligence, nonlinear model
-interpretability, visual analytics, local explanations, grand
-tour, radial tour}
-\Plainkeywords{explainable artificial intelligence, nonlinear model
-interpretability, visual analytics, local explanations, grand
-tour, radial tour}
-
-%% publication information
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{}
-%% \Acceptdate{2012-06-04}
-
-\Address{
-    Nicholas Spyrison\\
-    Monash University\\
-    Dept of Human Centred Computing\\
-Faculty of Information Technology\\
-  E-mail: \email{spyrison@gmail.com}\\
-  URL: \url{https://nspyrison.netlify.app}\\~\\
-      Dianne Cook\\
-    Monash University\\
-    Dept of Econometrics and Business Statistics\\
-
-    }
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\usepackage{amsmath} \usepackage{datetime}
+%Version 2.1 April 2023
+% See section 11 of the User Manual for version history
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
+
+%%=======================================================%%
+%% to print line numbers in the margin use lineno option %%
+%%=======================================================%%
+
+%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%======================================================%%
+%% to compile with pdflatex/xelatex use pdflatex option %%
+%%======================================================%%
+
+%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%Note: the following reference styles support Namedate and Numbered referencing.
+By default the style follows the most common style.
+To switch between the options you can add or remove Numbered in the optional parenthesis.
+%%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst, sn-mathphys.bst. %
+
+%%\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals
+%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+%%\documentclass[sn-mathphys,Numbered]{sn-jnl}% Math and Physical Sciences Reference Style
+%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
+%%\documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
+%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style 
+\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
+%%\documentclass[default]{sn-jnl}% Default
+%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published 
+%%%%  by Springer Nature. The guidance has been prepared in partnership with 
+%%%%  production teams to conform to Springer Nature technical requirements. 
+%%%%  Editorial and presentation requirements differ among journal portfolios and 
+%%%%  research disciplines. You may find sections in this template are irrelevant 
+%%%%  to your work and are empowered to omit any such section if allowed by the 
+%%%%  journal you intend to submit to. The submission guidelines and policies 
+%%%%  of the journal take precedence. A detailed User Manual is available in the 
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+%\jyear{2021}%
+
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+%%\unnumbered% uncomment this for unnumbered level heads
 
 \begin{document}
+
+\title[Article Title]{Article Title}
+
+%%=============================================================%%
+%% Prefix	-> \pfx{Dr}
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% NatureName	-> \tanm{Poet Laureate} -> Title after name
+%% Degrees	-> \dgr{MSc, PhD}
+%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
+%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
+%%=============================================================%%
+
+\title{Exploring Local Explanations of Nonlinear Models Using Animated Linear
+Projections}
+
+\author*[1]{\fnm{Nicholas} \sur{Spyrison}}\email{spyrison@gmail.com}
+
+\author[2]{\fnm{Dianne} \sur{Cook}}\email{dicook@monash.edu}
+
+\author[3]{\fnm{Przemyslaw} \sur{Biecek}}\email{przemyslaw.biecek@pw.edu.pl}
+
+\affil*[1]{\orgdiv{Faculty of Information Technology}, \orgname{Monash
+University}, \orgaddress{\street{Wellington Rd}, \city{Clayton}, \postcode{VIC 3800}, \country{Australia}}}
+
+\affil[2]{\orgdiv{Dept of Econometrics and Business Statistics}, \orgname{Monash University}, \orgaddress{\street{Wellington Rd}, \city{Clayton}, \postcode{VIC 3800}, \country{Australia}}}
+
+\affil[3]{\orgdiv{Faculty of Mathematics and Information Science}, \orgname{Warsaw University of Technology}, \orgaddress{\street{Koszykowa 75}, \city{Warsaw}, \postcode{00-662}, \country{Poland}}}
+
+%%==================================%%
+%% sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{The increased predictive power of machine learning models comes at the cost of increased complexity and loss of interpretability, particularly in comparison to parametric statistical models.
+This trade-off has led to the emergence of eXplainable AI (XAI) which provides methods, such as local explanations (LEs) and local variable attributions (LVAs), to shed light on how a model use predictors to arrive at a prediction.
+These provide a point estimate of the linear variable importance in the vicinity of a single observation.
+However, LVAs tend not to effectively handle association between predictors.
+To understand how the interaction between predictors affects the variable importance estimate, we can convert LVAs into linear projections and use the radial tour.
+This is also useful for learning how a model has made a mistake, or the effect of outliers, or the clustering of observations.
+The approach is illustrated with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models.
+The methods are implemented in the R package cheem, available on CRAN.}
+
+%%================================%%
+%% Sample for structured abstract %%
+%%================================%%
+
+\keywords{explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour}
+
+%%\pacs[JEL Classification]{D8, H51}
+
+%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}
+
+\maketitle
 
@@
 
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=0.85\linewidth]{./figures/shap_distr_bd}
-
+\begin{figure}[h!]
+{\centering \includegraphics[width=0.85\linewidth]{figures/shap_distr_bd} 
 }
 
-\caption[Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors]{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
+\caption{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors.
 A star offensive and defensive player are compared, L.
@@
 
-There are numerous R packages currently available that provide functions
-for computing SHAP values, including \texttt{fastshap} \citep{fastshap},
+There are numerous R packages currently available on CRAN that provide functions
+for computing SHAP and other LVA values, including \texttt{treeshap} \citep{kominsarczyk_treeshap_2023}, \texttt{fastshap} \citep{fastshap},
 \texttt{kernelshap} \citep{kernelshap}, \texttt{shapr} \citep{shapr},
@@
 the package \texttt{DALEX} has many resources \citep{biecek_dalex_2018}.
-There are many more packages only available through Github, like
-\texttt{treeshap} \citep{kominsarczyk_treeshap_2021} that is used for
-this work. \citet{molnar2022} provides good explanations of the
+\citet{molnar2022} provides good explanations of the
 different methods and how to apply them to different models.
@@
 original contribution. The implementation allows for pre-computation and
-interactive re-calculation to focus on a different variable.
-
-\begin{CodeChunk}
-\begin{figure}
+interactive re-calculation to focus on a different variable. In this work, the radial tour allows us to explore the sensitivity of LVA to the prediction of a model.
+
+\begin{figure}[h!]
 
@@
 observed against predicted response values is also provided (Figures
-\ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help
+\ref{fig:classificationcase}c, \ref{fig:regressioncase}b) to help
 identify observations poorly predicted by the model. For classification
@@
 observations are circled in red. Linked brushing between the plots is
-provided, and a tabular display of selected points helps to facilitate
-the exploration of the spaces and the model (shown in Figures
+provided (click and drag), and a tabular display of selected points helps to facilitate
+the exploration of the spaces and the model (shown in Figure
 \ref{fig:regressioncase}d).
@@
 \subsection{Radial Tour}\label{radial-tour}}
+
+HERE
 
@@
 user changes the projection the length of these bars will reflect this
-change. (The PI is interactively selected by clicking on a point in the
-global view). By scaling the SHAP value it becomes an (attribution)
+change. By scaling the SHAP value it becomes an (attribution)
 projection.
@@
 tests the sensitivity of structure (class separation or strength of
-relationship) to the variable's contribution. For classification, if the
-separation between classes diminishes when the variable contribution is
+relationship) to the variable's contribution. The CI attribution of 
+the CI does not impact the bases but it highlighted from context. 
+For classification, if the separation between classes diminishes when the variable contribution is
 reduced, this suggests that the variable is important for class
@@
 
-\begin{CodeChunk}
-\begin{figure}
+\begin{figure}[h!]
 
@@
 
-\begin{CodeChunk}
-\begin{figure}
+\begin{figure}[h!]
 
 {\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions} 
-
 }
 
-\caption[Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables]{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
+\caption{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables.
 Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values.
@@
 the selection of points (left click and drag) where those points will be
-highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
+highlighted across plots (Figure \ref{fig:regressioncase}a \& b).
 The information corresponding to the selected points is populated on a
-dynamic table (Figure \ref{fig:classificationcase}d). These interactions
+dynamic table (Figure \ref{fig:regressioncase}d). These interactions
 aid the exploration of the spaces and, finally, the identification of
@@
   package \textbf{randomForest} \citep{liaw_classification_2002},
-  compatibility tree SHAP. Modest hyperparameters are used, namely: 125
-  trees, the number of variables at each split, mtry = \(\sqrt{p}\) or
-  \(p/3\) for classification and regression, and minimum size of
-  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
-  classification and regression.
+  compatibility tree SHAP. Modest hyperparameters are used. Namely, 
+  classification models use 125 trees, number of variables at each 
+  split (mtry) of \(\sqrt{p}\), and minimum terminal node size of 
+  \(max(1, n/500)\). While regression models use 125 tree, \(p/3\) variables at split, and 
+  \(max(5, n/500)\) minimum terminal node size.
 \item
@@
   observation using the package \textbf{treeshap}
-  \citep{kominsarczyk_treeshap_2021}. We opt to find the attribution of
+  \citep{kominsarczyk_treeshap_2023}. We opt to find the attribution of
   each observation in the training data and not fit to fit variable
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
+\begin{lstlisting}
 R> install.packages("cheem", dependencies = TRUE)
@@
 R> run_app()
-\end{CodeInput}
-\end{CodeChunk}
-
-Alternatively,
-
-\begin{itemize}
-\tightlist
-\item
-  A version of the cheem viewer shiny app can be directly accessed at
-  \url{https://ebsmonash.shinyapps.io/cheem/}.
-\item
-  The development version of the package is available at
-  \url{https://github.com/nspyrison/cheem}, and
-\item
-  Documentation of the package can be found at
-  \url{https://nspyrison.github.io/cheem/}.
-\end{itemize}
+\end{lstlisting}
+
+Alternatively, a version of the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem/}, the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.
 
@@
 
-\hypertarget{acknowledgments}{%
-\subsection*{Acknowledgments}\label{acknowledgments}}
-\addcontentsline{toc}{subsection}{Acknowledgments}
+\backmatter
+
+\bmhead{Acknowledgments}
 
@@
 
-\renewcommand\refname{References}
-\bibliography{paper.bib}
+%%=============================================%%
+%% For submissions to Nature Portfolio Journals %%
+%% please use the heading ``Extended Data''.   %%
+%%=============================================%%
+
+%%=============================================================%%
+%% Sample for another appendix section			       %%
+%%=============================================================%%
+
+%% \section{Example of another appendix section}\label{secA2}%
+%% Appendices may be used for helpful, supporting or essential material that would otherwise 
+%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
+%% tables and equations etc.
+
+%%===========================================================================================%%
+%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
+%% system, please include the references within the manuscript file itself. You may do this  %%
+%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
+%% file, and delete the associated \verb+\bibliography+ commands.                            %%
+%%===========================================================================================%%
+
+\bibliography{sn-bibliography}
 
